### 2013-11-01 main.tex
Create LCS Algorithmic Intro document

\documentclass{article}

\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage[uniquename=init,firstinits,backend=bibtex,sorting=none]{biblatex}
\usepackage{calc}
\usepackage{color}
\usepackage{gensymb}
\usepackage{graphicx}
\usepackage{listings}
\usepackage[numbered,framed]{mcode}
\usepackage[space]{grffile}
%% hyperref manual advises loading this package last
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=blue,citecolor=blue,breaklinks=true}

\definecolor{gray}{rgb}{0.9,0.9,0.9}

\lstloadlanguages{Matlab}
\lstset{basicstyle=\small,breaklines=true,frame=single,float,columns=fullflexible,numbers=left,stepnumber=2,backgroundcolor=\color{gray}}

\newcommand{\reals}{\mathbf R}

\newcommand{\FIXME}[1]{\\\fbox{FIXME: #1}\\}

\title{An Algorithmic Introduction to Lagrangian Coherent Structures}

\author{Kristjan Onu \and Florian Huhn \and George Haller}

\addbibresource{main}

\begin{document}

\numberwithin{equation}{section}
\numberwithin{figure}{section}
\numberwithin{lstlisting}{section}
\numberwithin{table}{section}

\maketitle

\tableofcontents

\abstract{
We give a practical introduction to Lagrangian coherent structures (LCSs.) The article reviews recent results in variational LCS theory, presents algorithms to detect LCSs in two dimensional time dependent flows, and demonstrates a LCS computational engine called LCS Tool using examples.
}

\section{Introduction}

LCSs are used in fluid dynamics to capture essential flow features and thereby simplify the description of fluid flows. More specifically, LCSs have been particularly well adapted to analyze two dimensional finite time turbulent flows. LCSs capture dominant transport barriers, repelling and attracting structures and LCSs are used to identify vortex boundaries.

LCSs have been used to analyse oceanic and atmospheric flows, biological applications\parencite{wilson09:_lagran_reynol,tallapragada11:_lagran}, aeronautics\parencite{tang10:_accur_lagran_hong_kong_inter_airpor} and mechanical systems\parencite{hadjighasem13:_detec_kam}.

The finite-time Lyapunov exponent (FTLE) gives a mathematical definition for LCSs. Although the FTLE has been applied to a variety a situations, it is possible to find examples where it fails to identify LCSs correctly, giving false positives and false negatives\parencite{haller11:_lagran_coher_struc,norgard12:_secon_lagran_coher_struc}. To remedy these problems, variational principles have been used to present new approaches to identifying LCSs\parencite{haller11:_lagran_coher_struc,farazmand12:_comput_lagran,haller12:_geodes_theor_trans_barrier_two_dimen_flows,farazmand13:_attrac_lagran,haller13:_coher_lagran}.

This article presents an algorithmic introduction LCSs using a computational engine for LCSs called LCS Tool\footnote{LCS Tool is free MATLAB software available for download at: \url{https://github.com/jeixav/LCS-Tool}}.

LCS Tool is a library of functions that perform the computations to identify LCSs in two dimensional finite time flows. LCS Tool aims to initiate scientists to LCSs. Existing LCS software programs are:
\begin{itemize}
\item \emph{MANGEN}\parencite{lekien03:_time}, calculates stability manifolds adaptively in two dimensional finite time velocity fields. Includes a graphical user interface and uses MPI for parallel calculations.
\item \emph{LCS MATLAB Kit}\parencite{dabiri09:_lmk} calculates the FTLE from velocity datasets. Includes a graphical user interface.
\item \emph{Newman}\parencite{toit10:_trans} calculates the FTLE in $N$ dimensions. Assists ridge extraction of FTLE fields. Supports analytic and dataset defined velocity definitions.
\item \emph{FlowVC}\parencite{shadden10:_flowvc} is a general purpose LCS platform for two and three dimensional datasets. Parallel calculations are supported with Open\-MP, CUDA and OpenCL.
\end{itemize}
These packages are based on the FTLE; the objective of this article and LCS Tool is to replace FTLE methods.

\section{Theory}

The fluid flows considered are defined in two dimensions by a velocity field:
\begin{align*}
\dot{\boldsymbol x} = \boldsymbol v(\boldsymbol x,t), && \boldsymbol x \in U \subset \reals^2, && t \in [t_-,t_+].
\end{align*}
Integration yields the flow map, defined between an initial time $t_0$ and a final time $t$ within $[t_-,t_+]$:
\[
F_{t_0}^t(\boldsymbol x_0) := \boldsymbol x(t,t_0,\boldsymbol x_0).
\]
The time interval used to detect LCSs is left as a choice; it can be the complete interval for which the flow is defined, or data is available, or it can be a chosen subset.

The right Cauchy-Green strain tensor measures Lagrangian strain in the velocity field and is defined:
\[
C_{t_0}^t(\boldsymbol x_0) = \left[\nabla F_{t_0}^t(\boldsymbol x_0)\right]^T \nabla F_{t_0}^t(\boldsymbol x_0).
\]
The eigenvalues of the Cauchy-Green strain tensor are denoted by $\lambda_{1,2}$ and the eigenvectors by $\boldsymbol \xi_{1,2}$. The eigenvalues satisfy $0 < \lambda_1 \leq \lambda_2$ and the eigenvectors satisfy $\boldsymbol \xi_1 \perp \boldsymbol \xi_2$.

We find coherent Lagrangian vortices by considering strain, as was done in~\textcite{haller13:_coher_lagran}. To summarise the results in that publication, we examine closed curves $\gamma$ and $O(\epsilon)$ neighbourhoods or ``belts'' around them. For most $\gamma$, neighbouring belts of $O(\epsilon)$ exhibit strain variability of $O(\epsilon)$. Curves that form coherent Lagrangian vortices are exceptional however and display no observable variability in their average strain within $O(\epsilon)$ neighbourhoods.

The mathematical identification of a coherent Lagrangian vortex para\-metrises a curve $\gamma$ by $\boldsymbol r(s)$ with $s \in [0,\sigma]$ at time $t_0$. The length of a tangent vector $\boldsymbol r'(s)$ is $l_{t_0}(s)$. The advected curve $F_{t_0}^t(\gamma)$ has a tangent vector $(\text{d}/\text{d}s) F_{t_0}^t(\boldsymbol r(s))$ and a length we denote by $l_t(s)$. The two tangent lengths are calculated:
\begin{gather*}
l_{t_0} = \sqrt{\langle \boldsymbol r'(s), \boldsymbol r'(s) \rangle},\\
l_t = \sqrt{\langle \boldsymbol r'(s), C_{t_0}^t(\boldsymbol r(s)) \boldsymbol r'(s) \rangle},\\
\end{gather*}
where $\langle \cdot, \cdot \rangle$ denotes the Euclidean inner product.
The average tangential strain along $\gamma$ is:
\[
Q(\gamma) = \frac{1}{\sigma} \int_0^\sigma \frac{l_t(s)}{l_{t_0}(s)}\text{d}s.
\]

Coherent Lagrangian vortices are associated with the strain tensor
\[
E_\lambda(\boldsymbol x_0) = \frac12 [C_{t_0}^t(\boldsymbol x_0) - \lambda^2 I]
\]
and curves uniformly stretched by the same factor $\lambda$ when advected by the flow from time $t_0$ to $t$ are closed trajectories of one of two explicit 
differential equations: 
\begin{equation}
\boldsymbol r' = \boldsymbol \eta_\pm(\boldsymbol r),
\label{e:etafields}
\end{equation}
where
\[
\boldsymbol \eta_\pm = \sqrt{\frac{\lambda_2 - \lambda^2}{\lambda_2 - \lambda_1}} \boldsymbol \xi_1 \pm \sqrt{\frac{\lambda^2 - \lambda_1}{\lambda_2 - \lambda_1}} \boldsymbol \xi_2.
\]
In subsequent sections $\lambda = 1$ always.

We find hyperbolic LCSs from a variational approach based on shear, as presented in~\textcite{farazmand13:_shearless}. The Lagrangian shear is calculated:
\[
p(s) = \frac{\langle \boldsymbol r'(s), D(\boldsymbol r(s)) \boldsymbol r'(s)\rangle}{\sqrt{\langle \boldsymbol r'(s), C(\boldsymbol r(s)) \boldsymbol r'(s)\rangle \langle \boldsymbol r'(s), \boldsymbol r'(s)\rangle}}
\]
where
\[
D(\boldsymbol x_0) = \frac12[C(\boldsymbol x_0) \Omega - \Omega C(\boldsymbol x_0)], \quad \Omega = \begin{pmatrix}0&-1\\1&0\end{pmatrix}.
\]
The average Lagrangian shear along a curve $\gamma$ over the time interval $[t_0,t]$ is:
\[
P(\gamma) = \frac{1}{\sigma} \int_0^\sigma p(s) \text{d}s.
\]

Similarly to Lagrangian vortices, exceptional hyperbolic curves without shear in $O(\epsilon)$ neighbourhoods are null-geodesics of the Lorentzian metric
\[
g(\boldsymbol u, \boldsymbol v) = \langle \boldsymbol u, D(\boldsymbol x_0) \boldsymbol v \rangle.
\]
Repelling LCSs are obtained as solutions of the strainline differential equation:
\[
\boldsymbol r' = \boldsymbol \xi_1(\boldsymbol r),
\]
and attracting LCSs are obtained as a solution of the stretchline differential equation:
\[
\boldsymbol r' = \boldsymbol \xi_2(\boldsymbol r).
\]

\section{Methods}

This section presents the numerical methods used to compute Lagrangian coherent structures. In other words, the algorithms that constitute LCS Tool are presented.

\subsection{The Cauchy-Green strain tensor}

To obtain Lagrangian coherent structures the Cauchy-Green strain tensor must be calculated first. More specifically, the eigenvalues and eigenvectors of the Cauchy-Green strain tensor are needed. The function associated with this calculation in LCS Tool is \texttt{eig\_cgStrain}. The main steps of this function are enumerated in Table~\ref{t:Cauchy-Green algorithm}.

\begin{table}
\begin{enumerate}
\item Setup a Cartesian grid and auxiliary grid in the (rectangular) domain of the flow.
\item Integrate the flow velocity equations at each grid point and auxiliary grid point. This yields the flow map, $F_{t_0}^t(\boldsymbol x_0)$.
\item Use finite differences to approximate the derivative of the flow map, $D F_{t_0}^t(\boldsymbol x_0)$.
\item Compute the Cauchy-Green strain tensor, $C_{t_0}^t = \left(D F_{t_0}^t\right)^T D F_{t_0}^t$ and its eigenvalues, $\lambda_{1,2}$ and eigenvectors, $\xi_{1,2}$.
\end{enumerate}
\caption{Algorithm used to calculate the Cauchy-Green strain tensor eigenvalues and eigenvectors.}
\label{t:Cauchy-Green algorithm}
\end{table}

We setup a Cartesian grid according to user-defined resolutions. We can find the optimal resolution by an iterative procedure in which we start with a low resolution and double it until convergence of LCSs is observed visually. A resolution of about 500 grid point along the longest axis usually gives good results. Furthermore an auxiliary grid is setup. This is illustrated in Figure~\ref{f:main and auxiliary grids}. The improvements the auxiliary grid yields compared to using only the main grid were reported in \textcite{farazmand12:_comput_lagran}. Experience indicates that optimal results are obtained when the auxiliary grid spacing is between 1 and 10\% of the main grid spacing.

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{graphics/main_aux_grids}
\end{center}
\caption{Illustration of main and auxiliary grid used to compute the Cauchy-Green strain. Solid circles are main grid points and hollow circles are auxiliary grid points.}
\label{f:main and auxiliary grids}
\end{figure}

The function \texttt{eig\_cgStrain} provides the option to calculate the eigenvector from the auxiliary grid, but the eigenvalues from the main grid. We have found that for flows defined analytically the eigenvalues should be calculated from the main grid whereas for the flows defined by datasets, using the auxiliary grid gives better results.

As stated, typically we set the main grid resolution of the Cauchy-Green strain tensor to around 500 grid points. For a square domain, this implies the velocity equations must be integrated, after including 4 auxiliary grid points around every main grid point, at 1.25 million grid points. The ``natural'' way to perform these integration calculations would be to solve a two dimensional first-order ODE system at every grid point and iterate over every grid point within a for-loop. With 1.25 million grid points this calculation can take an excessively long time. To alleviate this problem, we solve the flow equations in a vector form where the equations at all grid points are combined into a single system of equation, even though the system of composed of independent blocks of two dimensional first-order ODE systems. MATLAB's \texttt{ode45} function is used to perform the integration and the flow integration typically takes five to ten minutes with default error tolerances. The drawback of vector form integration is its memory requirements can become excessive at high resolutions, furthermore, writing the velocity function in vector form is more error-prone than the intrinsic two dimensional form.

The Cauchy-Green strain tensor can have singularities (i.e. $C_{t_0}^t(\boldsymbol x_0) = I$). Such points are isolated in general\parencite{delmarcelle94}, therefore no special accommodations for are made in computations for singularities of the Cauchy-Green strain tensor.

\subsubsection{Special case: incompressible velocity fields}

Flows that are incompressible, meaning $\nabla \cdot \boldsymbol v = 0$, satisfy the relation $\lambda_1(\boldsymbol x_0) \lambda_2(\boldsymbol x_0) = 1, \forall \: \boldsymbol x_0 \: \in \: U$\parencite{arnold78:_mathem}. This gives the possibility of imposing incompressibility when calculating the eigenvalues of the Cauchy-Green strain tensor by first calculating $\lambda_2$ and then setting $\lambda_1(\boldsymbol x_0) = 1/\lambda_2(\boldsymbol x_0)$. This procedure should be applied carefully since by changing $\lambda_1$ but not $\xi_1$, the error in satisfying the eigenvalue equation for the Cauchy-Green strain tensor may increase.

At some grid points it may occur that $\lambda_2 < 1$ due to numerical integration errors of the flow velocity equations. By setting integration tolerances to smaller values the number of grid points with $\lambda_2 < 1$ can be reduced. Nonetheless, some grid points can be so close to singularities of the Cauchy-Green strain tensor that satisfying $\lambda_2 >= 1$ everywhere could be excessively computationally costly. Therefore the function \texttt{eig\_cgStrain} records the number of points at which incompressibility cannot be imposed and this measure can also help to set appropriate integration tolerances.

\subsubsection{Special case: data-defined velocity fields}

Velocity fields defined by datasets require pre-processing prior to performing the integration that yields the flow map. An interpolation function must be written to enable evaluating the velocity function at arbitrary points in $U$ and at arbitrary times between $t_-$ and $t_+$. MATLAB's interpolation functions can be integrated into the velocity function. As with analytic cases, care should be taken to ensure the interpolation embedded in the flow velocity integration function works correctly when performing integration in vector form.

Beyond the pre-processing needed to obtain the Cauchy-Green strain tensor eigenvalues and eigenvectors, no special manipulations are needed to treat data-defined flow. It may be however, that parameter values giving good results with analytically defined flows need adjustment prior to processing data-defined velocity fields.

\subsection{Shear LCS}

Shear LCSs, or coherent Lagrangian vortices, are closed orbits in the Lagrangian shear vector fields $\boldsymbol \eta_{\pm}$ defined in Equation~\eqref{e:etafields}. Since $\lambda = 1$, the closed curves are non-stretching and area preserving. We find closed orbits by integrating shearlines starting from a Poincare section and evaluating the first return map. A shearline closing onto itself in the starting point is a shear LCS.

The main steps of the algorithm used to calculate shear LCS are enumerated in Table~\ref{t:Shear LCS algorithm}

\begin{table}
\begin{enumerate}
\item Position Poincare sections in flow domain
\item Calculate Poincare return map
\item Find closed orbits in Poincare return map
\item Identify outermost closed orbit
\end{enumerate}
\caption{Algorithm used to calculate Shear LCS.}
\label{t:Shear LCS algorithm}
\end{table}

The first step is to set the position of Poin\-care sections \texttt{poincare\-Section.endPosition} that are straight lines in regions where closed shearlines are expected. The Poincare section must be oriented in a way that the first endpoint is close to the centre of the expected eddy and the second endpoint is outside the expected eddy region. There is no predetermined procedure for deciding where the Poincare sections should be positioned. Possibilities are an examination of the finite-time Lyapunov exponent field of the flow, or some prior knowledge of where coherent eddies could be expected. Additionally, the number of trajectories launched from the Poincare section \texttt{poincareSection.numPoints} must be defined. A good default value is $100$.

The second step is to integrate the shearlines to obtain the Poincare return map. Integration of shearlines is performed with the $\eta_+$ and $\eta_-$ vector fields defined on the main grid. These vector fields are derived from eigenvalues and eigenvectors as described in~\textcite{farazmand12:_comput_lagran}. The eigenvector fields have orientation discontinuities, therefore the vector field must be locally reoriented for the integration. The algorithm used to perform variable step-size integration in vector fields with orientation discontinuities is enumerated in Table~\ref{t:variable step integration} and illustrated in Figure~\ref{f:variable step integration}. Linear interpolation is used when interpolating $\boldsymbol \eta_\pm$ within a grid element, since using higher order interpolation would necessitate verifying that there are no orientation discontinuities at more than only the four nearest grid points. Detection of orientation discontinuities is achieved by checking the inner product of the vector fields at adjacent grid points. Then, rotations exceeding 90\degree\, are classified as orientation discontinuities and are corrected before linear interpolation. Thus when setting the Cauchy-Green strain tensor main grid resolution, a histogram of eigenvector field rotations may be produced to ensure that all rotations are below 90\degree\, or close to 180\degree.

\begin{table}
\begin{enumerate}
\item Linearly interpolate vector field orientation at initial position
\item At next position, check whether vector field has rotated by over 90\degree, if yes, flip the vector field orientation by 180\degree.
\item Stop integration when shearline returns to Poincare section, or maximum integration length has been reached.
\end{enumerate}
\caption{Algorithm used for variable time step integration of shearlines.}
\label{t:variable step integration}
\end{table}

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{variable_step_integration}
\end{center}
\caption{Schematic illustration of shearline variable time step integration. At the initial point, there is an orientation discontinuity at the lower right grid point that must be corrected prior to linear interpolation. At point \#1, no orientation discontinuties are present. At point \#2 the interpolated eigenvector field must be rotated by 180\degree\,to match the orientation at the previous point, \#1.}
\label{f:variable step integration}
\end{figure}

An example of a Poincare return map produced from integration of the $\boldsymbol \eta_\pm$ field is shown in Figure~\ref{f:Poincare return map}. Well behaved orbits will return to the Poincare section and their integration will be stopped using an ODE event detection function. Some orbits may however deviate far from the Poincare section and their integration could take inordinately long. To prevent this behaviour, a maximum orbit length, \texttt{poincareSection.\-orbit\-MaxLength}, is specified. In practice, viewing the Poincare section as the radius of a circle and setting the maximum shearline integration length to twice the circumference gives acceptable results. In Figure~\ref{f:Poincare return map}, circle markers indicate closed orbit positions. Closed orbits exist where the Poincare return map is zero. The function \texttt{poincare\_closed\_orbit} performs the computations. In Figure~\ref{f:Poincare return map} not all zero crossings have circles. This is because a threshold parameter, \texttt{dThresh} whose default value is 1\% of the length of the Poincare section, is used to discard those zero crossings that appear in regions with high numerical noise. The position of each detected zero crossing is refined by iteratively bisecting the interval of the zero crossing. If after a predetermined number of iterations, \texttt{nBisection} whose default value is five, the two points around the zero crossing have absolute values of more \texttt{dThresh}, the zero crossing is discarded. In the Figure, discarded zero crossings are seen where $0 < s < 0.05$. Once all closed shearline orbits have been located, the outermost orbit of every Poincare section is deemed a shear LCS.

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{graphics/double_gyre/poincare_return_map}
\end{center}
\caption{Example of a Poincare return map obtained for shear LCS. Circle markers indicate closed orbit positions. The filled circle indicates the outermost closed orbit and shear LCS.}
\label{f:Poincare return map}
\end{figure}

\subsection{Hyperbolic LCS}

The main steps of the algorithm used to calculate hyperbolic LCS are enumerated in Table~\ref{t:Hyperbolic LCS algorithm}

\begin{table}
\begin{enumerate}
\item Define a local maximisation distance.
\item Find all local maximums of $\lambda_2$ on the main grid such that they are maximums within a circle with a radius equal to the local minimisation distance.
\item Define a maximum strainline length
\item Integrate a strainline forward and backward, using the largest $\lambda_2$ local maximum as the initial position. Integrate until the strainline has attained the maximum strainline length, or until it has reached the domain boundary.
\item Flag any remaining local maximums of $\lambda_2$ within the minimisation distance of the strainline as ineligible initial positions for subsequent strainlines
\item Continue integrating strainlines using local maximums of $\lambda_2$ as initial positions until no eligible local maximums of $\lambda_2$ remain.
\item Remove any strainline segment contained within (closed) shear LCS
\end{enumerate}
\caption{Algorithm used to calculate hyperbolic LCS.}
\label{t:Hyperbolic LCS algorithm}
\end{table}

Strainlines are integrated from local maximums of $\lambda_2$ on the main grid. A local maximisation distance is defined. This distance is set between two and five grid points since it represents the number of grid points over which the Cauchy-Green strain tensor is susceptible to numerical noise from integration. Thus the local maximisation distance is expected to be a function of variable time step integration tolerances. A record is made of all grid points where $\lambda_2$ is maximum within a circle whose radius equals the defined maximisation distance. Then the first strainline is integrated from the largest of all the local maximums forward and backward in time. Similarly to shearlines, this integration must take into account eigenvector field orientation discontinuities. Prior to initiating strainline integration, a maximum length is defined. Strainlines are then integrated until reaching the domain boundary or the maximum length. The maximum length may be seen as safety factor to prevent highly sinuous strainlines leading to inordinately long computations; the maximum length should be set such that most strainline stop at the domain boundary. Once a strainline position has been calculated, local maximums of $\lambda_2$ within the minimisation distance are flagged as ineligible initial positions for subsequent strainlines. After the first strainline has been obtained and any ineligible local maximums of $\lambda_2$ have been flagged the next strainline is integrated. The process continues until all eligible local maximums of $\lambda_2$ have been used to initialise strainline integration. Finally any strainline segment contained within (closed) shear LCS are removed. These calculations are performed by the functions \texttt{seed\_curves\_from\_lambda\_max} and \texttt{remove\_strain\_in\_shear}.

Note that once a strainline integration has begun, it may progress so as to be within the local minimisation distance of an existing strainline. In other words, the local minimisation distance controls where strainlines begin, but not where they stop.

\section{Examples}

\subsection{The double gyre}

The double gyre is a model for a pattern observed in geophysical flows\parencite{shadden05:_defin_lagran_lyapun}. The velocity equations are:
\begin{equation}
\begin{split}
\dot x = -\pi A \sin[\pi f(x,t)] \cos(\pi y)\\
\dot y = \pi A \cos[\pi f(x,t)] \sin(\pi y) \frac{\partial f(x,t)}{\partial x}\\
f(x,t) = \epsilon \sin(\omega t) x^2 + [1 - 2 \epsilon \sin(\omega t)] x
\end{split}
\label{e:double gyre derivative equations}
\end{equation}
The MATLAB function corresponding to these equations is given in Listing~\ref{l:double gyre derivative} and shows how to specify a derivative function that supports vector form integration.

\begin{lstlisting}[caption={Double gyre derivative function corresponding to Equations~\ref{e:double gyre derivative equations}.},label=l:double gyre derivative]
function derivative_ = derivative(t,x,epsilon,amplitude,omega)

idx1 = 1:2:numel(x)-1;
idx2 = 2:2:numel(x);

a = epsilon*sin(omega*t);
b = 1 - 2*epsilon*sin(omega*t);
forcing = a*x(idx1).^2 + b*x(idx1);

derivative_ = nan(size(x));

derivative_(idx1) = -pi*amplitude*sin(pi*forcing).*cos(pi*x(idx2));
derivative_(idx2) = pi*amplitude*cos(pi*forcing).*sin(pi*x(idx2)).*(2*a*x(idx1) + b);
\end{lstlisting}

The parameter values used are: $A = 0.1$, $\epsilon = 0.1$, $\omega = \pi/5$. The flow timespan is $t \in [0,20]$. The physical domain of the double gyre flow is $x \in [0,2]$, $y \in [0,1]$, however to help capture behaviour near the fixed points of the unperturbed system at $(x,y) = (1,0)$ and $(x,y) = (1,1)$ in computations, the domain is enlarged to $x \in [-0.1,2.1]$, $y \in [-0.05,1.05]$. To calculate the Cauchy-Green strain tensor and its eigenvalues and eigenvectors, the main grid resolution is set to $551 \times 276$. The main grid is used to calculate eigenvalues and flow incompressibility is imposed. The LCS Tool code that corresponds to these parameters is in Listing~\ref{l:double gyre input}. The variable named flow is a MATLAB structure that contains fields corresponding to input parameters; using structures allows passing input arguments to LCS Tool functions with a compact notation.

\begin{lstlisting}[firstnumber=4,caption={Double gyre input parameter definitions; subset of the file: \texttt{demo/double\_gyre/hyperbolic\_shear\_lcs.m}.},label=l:double gyre input]
epsilon = .1;
amplitude = .1;
omega = pi/5;

flow.imposeIncompressibility = true;
flow.derivative = @(t,x,useEoV)derivative(t,x,useEoV,epsilon,amplitude,omega);
flow.domain = [-.1,2.1;-.05,1.05];
flow.timespan = [0,20];
flow.resolution = [551,276];
\end{lstlisting}

To obtain shear LCSs, Poincare section positions are specified. Figure~\ref{f:Poincare sections with FTLE} illustrates how Poincare sections are placed in regions where the FTLE ridges have a circular pattern. The set of LCS Tool commands used to detect shear LCSs is given in Listing~\ref{l:double gyre shear LCS commands}. First the function \texttt{eig\_cgStrain} calculates the Cauchy-Green strain eigenvalues and eigenvectors. The most important function is \texttt{poincare\_closed\_orbit\_multi} that takes as input a list of Poincare sections, calculates Poincare return maps for the $\eta_+$ and $\eta_-$ vector fields and returns positions of all closed orbits.

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{graphics/double_gyre/poincare_sections}
\end{center}
\caption{Poincare sections positions with FTLE.}
\label{f:Poincare sections with FTLE}
\end{figure}

\begin{lstlisting}[caption={Double gyre shear LCS detection commands. Subset of \texttt{hyperbolic\_shear\_lcs.m}.},label=l:double gyre shear LCS commands,firstnumber=24]
% Compute Cauchy-Green strain eigenvalues and eigenvectors
[flow.cgEigenvalue,flow.cgEigenvector] = eig_cgStrain(flow);
% FIXME Should use m-by-n or (m*n)-by-2 array forms throughout LCS Tool
cgEigenvalue = reshape(flow.cgEigenvalue,[fliplr(flow.resolution),2]);
cgEigenvector = reshape(flow.cgEigenvector,[fliplr(flow.resolution),4]);

% Define Poincare sections; first point in center of elliptic region and
% second point outside elliptic region
poincareSection = struct('endPosition',{},'numPoints',{},'orbitMaxLength',{});

poincareSection(1).endPosition = [.5,.6;.35,.5];
poincareSection(2).endPosition = [1.5,.4;1.7,.5];

% Number of orbit seed points along each Poincare section
[poincareSection.numPoints] = deal(80);

% Set maximum orbit length to twice the expected circumference
nPoincareSection = numel(poincareSection);
for i = 1:nPoincareSection
    rOrbit = hypot(diff(poincareSection(i).endPosition(:,1)),diff(poincareSection(i).endPosition(:,2)));
    poincareSection(i).orbitMaxLength = 2*(2*pi*rOrbit);
end

[shearline.etaPos,shearline.etaNeg] = lagrangian_shear(flow.cgEigenvector,flow.cgEigenvalue);

% Compute closed shearlines
closedOrbits = poincare_closed_orbit_multi(flow,shearline,poincareSection);
\end{lstlisting}

To obtain hyperbolic LCS, the local maximisation distance is set to twice the main grid spacing. To limit integration time, the maximum strainline length is set to 20. LCS Tool commands are shown in Listing~\ref{l:double gyre hyperbolic LCS commands}. The function \texttt{seed\_curves\_from\_lambda\_max} calculates strainline positions. The function \texttt{remove\_strain\_in\_shear} then removes the segments of strainlines within the outermost shearline of each Poincare section.

\begin{lstlisting}[caption={Double gyre hyperbolic LCS detection commands. Subset of \texttt{hyperbolic\_shear\_lcs.m}.},label=l:double gyre hyperbolic LCS commands,firstnumber=60]
strainlineMaxLength = 20;
gridSpace = diff(flow.domain(1,:))/(double(flow.resolution(1))-1);
localMaxDistance = 2*gridSpace;

% Compute strainlines
strainlinePosition = seed_curves_from_lambda_max(localMaxDistance,strainlineMaxLength,cgEigenvalue(:,:,2),cgEigenvector(:,:,1:2),flow.domain);

for i = 1:nPoincareSection
    % Remove strainlines inside of ellitpic regions
    strainlinePosition = remove_strain_in_shear(strainlinePosition,closedOrbits{i}{1}{end});
    strainlinePosition = remove_strain_in_shear(strainlinePosition,closedOrbits{i}{2}{end});
end
\end{lstlisting}

The forward time LCS obtained are shown in Figure~\ref{f:double gyre hyperbolic shear lcs forward} and the backward time LCS (i.e. with the flow timespan set to $t \in [20,0]$) are shown in Figure~\ref{f:double gyre hyperbolic shear lcs backward}. In Figures~\ref{f:double gyre hyperbolic shear lcs details forward} and~\ref{f:double gyre hyperbolic shear lcs details backward} the same LCS are shown with details: the finite-time Lyapunov exponent, closed shearlines inside the outermost shearlines and $\lambda_2$ local maximums used for strainline integration initial positions.

\begin{figure}
\begin{center}
% LCS Tool commit ac3a7705aec81577b1aba3cafe598066dd689203
\includegraphics[width=\textwidth]{graphics/double_gyre/hyperbolic_shear_lcs_forward}
\end{center}
\caption{Double gyre forward time shear (green) and hyperbolic (red) LCS.}
\label{f:double gyre hyperbolic shear lcs forward}
\end{figure}

\begin{figure}
\begin{center}
% LCS Tool commit ac3a7705aec81577b1aba3cafe598066dd689203
\includegraphics[width=\textwidth]{graphics/double_gyre/hyperbolic_shear_lcs_backward}
\end{center}
\caption{Double gyre backward time shear (green) and hyperbolic (blue) LCS.}
\label{f:double gyre hyperbolic shear lcs backward}
\end{figure}

\begin{figure}
\begin{center}
% LCS Tool commit ac3a7705aec81577b1aba3cafe598066dd689203
\includegraphics[width=\textwidth]{graphics/double_gyre/hyperbolic_shear_lcs_details_forward}
\end{center}
\caption{Same LCS as in Figure~\ref{f:double gyre hyperbolic shear lcs forward}, but with additional details. Background gray-scale is the finite-time Lyapunov exponent. Dashed green lines are Poincare sections. Green curves are closed shearlines with thick green curves are outermost shearlines, i.e. shear LCS. White dots indicate $\lambda_2$ local maximums used for strainline integration initial positions.}
\label{f:double gyre hyperbolic shear lcs details forward}
\end{figure}

\begin{figure}
\begin{center}
% LCS Tool commit ac3a7705aec81577b1aba3cafe598066dd689203
\includegraphics[width=\textwidth]{graphics/double_gyre/hyperbolic_shear_lcs_details_backward}
\end{center}
\caption{Same LCS as in Figure~\ref{f:double gyre hyperbolic shear lcs backward}, but with additional details. Background gray-scale is the finite-time Lyapunov exponent. Dashed green lines are Poincare sections. Green curves are closed shearlines and thick green curves are outermost shearlines, i.e. shear LCS. White dots indicate $\lambda_2$ local maximums used for strainline integration intial positions.}
\label{f:double gyre hyperbolic shear lcs details backward}
\end{figure}

Stretchlines are calculated using the function \texttt{compute\_stretchline}. Forward time stretchlines are shown in Figure~\ref{f:double gyre stretchlines forward} and backward time stretchlines are shown in Figure~\ref{f:double gyre stretchlines backward}.

\begin{figure}
\begin{center}
% demo/double_gyre/stretchlines.m, commit dbefa5a
\includegraphics[width=\textwidth]{graphics/double_gyre/stretchlines_forward}
\end{center}
\caption{Double gyre forward time stretchlines.}
\label{f:double gyre stretchlines forward}
\end{figure}

\begin{figure}
\begin{center}
% demo/double_gyre/stretchlines.m, commit dbefa5a
\includegraphics[width=\textwidth]{graphics/double_gyre/stretchlines_backward}
\end{center}
\caption{Double gyre backward time stretchlines.}
\label{f:double gyre stretchlines backward}
\end{figure}

\subsection{Bickley jet}

The Bickley jet models a zonal jet flanked above and below by
stationary vortices. This is an idealised model of geophysical flows
such as the polar night jet perturbed by a Rossby
wave\parencite{beron-vera10:_invar_lagran,haller12:_geodes_theor_trans_barrier_two_dimen_flows}.

The model is defined by two stream functions, $\psi(x,y,t) = \psi_0(x,y) + \psi_1(x,y,t)$, where:
\begin{gather*}
\psi_0(x,y) = c_3 y - U L_y \tanh\frac{y}{L_y} + \epsilon_3 U L_y \mathrm{sech}^2\frac{y}{L_y} \cos k_3 x,\\
\psi_1(x,y,t) = U L_y \mathrm{sech}^2\frac{y}{L_y} \mathrm{Re}\left[ \sum_{n=1}^2 \epsilon_n f_n(t) e^{i k_n x}\right].
\end{gather*}
The forcing function is a solution of the Duffing oscillators with parameters in the chaotic regime, specifically
\begin{gather*}
\frac{d \phi_1}{dt} = \phi_2,\\
\frac{d \phi_1}{dt} = -0.1 \phi_2 - \phi_1^3 + 11 \cos(t),\\
f_{1,2}(t) = 2.625 \times 10^{-2} \phi_1(t/6.238 \times 10^5)
\end{gather*}
The velocity is given by:
\[
\boldsymbol{v}(x,y,t) = (-\partial_y \psi, \partial_x \psi).
\]
The parameter values used are: $U = 62.66$, $c_2 = 0.205 U$, $c_3 = 0.461 U$, $L_y = 1.77 \times 10^6$, $\epsilon_1 = 0.0075$, $\epsilon_2 = 0.04$, $\epsilon_3 = 0.3$, $L_x = 6.371 \times 10^6 \pi$, $k_n = 2 n \pi/L_x$, $\sigma_1 = 0.5 k_2 (c_2 - c_3)$, $\sigma_2 = 2 \sigma_1$. The domain of the flow is set to $x \in [0,L_x]$, $y \in [-4 \times 10^6, 4 \times 10^6]$ and the timespan is set to $4 L_x/U$.

To calculate the Cauchy-Green strain tensor and its eigenvalues and eigenvectors, the main grid resolution is set to $500 \times 200$ and the auxiliary grid spacing is set to one percent of the main grid spacing (i.e. 400.) The main grid is used to calculate eigenvalues and flow incompressibility is imposed. Periodic boundary conditions are imposed in the horizontal direction.

To obtain shear LCS Poincare sections are specified. The maximum orbit length is set to twice the circumference circles whose radius would be the Poincare sections.

To obtain hyperbolic LCS, the local maximisation distance is set to eight times the main grid spacing and the maximum strainline length is set to $1 \times 10^8$.

The forward time LCS obtained are shown in Figure~\ref{f:bickley jet hyperbolic shear lcs forward} and the backward time LCS (i.e. with the flow timespan set to $t \in [4 L_x/U,0]$ are shown in Figure~\ref{f:bickley jet hyperbolic shear lcs backward}. In Figures~\ref{f:bickley jet hyperbolic shear lcs details forward} and~\ref{f:bickley jet hyperbolic shear lcs details backward} the same LCS are shown with details.

Forward time stretchlines are shown in Figure~\ref{f:bickley jet forward stretchlines} and backward time stretchlines are shown in Figure~\ref{f:bickley jet backward stretchlines}.

\begin{figure}
\begin{center}
% demo/bickley_jet/hyperbolic_shear_lcs.m, commit aaf20ed
\includegraphics[width=\textwidth]{graphics/bickley_jet/hyperbolic_shear_lcs_forward}
\end{center}
\caption{Bickley jet forward time shear (green) and hyperbolic (red) LCS.}
\label{f:bickley jet hyperbolic shear lcs forward}
\end{figure}

\begin{figure}
\begin{center}
% demo/bickley_jet/hyperbolic_shear_lcs.m, commit aaf20ed
\includegraphics[width=\textwidth]{graphics/bickley_jet/hyperbolic_shear_lcs_backward}
\end{center}
\caption{Bickley jet backward time shear (green) and hyperbolic (blue) LCS.}
\label{f:bickley jet hyperbolic shear lcs backward}
\end{figure}

\begin{figure}
\begin{center}
% demo/bickley_jet/hyperbolic_shear_lcs_details.m, commit 3331187
\includegraphics[width=\textwidth]{graphics/bickley_jet/hyperbolic_shear_lcs_details_forward}
\end{center}
\caption{Same LCS as in Figure~\ref{f:bickley jet hyperbolic shear lcs forward}, but with additional details. Background gray-scale is the finite-time Lyapunov exponent. Dashed green lines are Poincare sections. Green curves are closed shearlines with thick green curves are outermost shearlines, i.e. shear LCS. White dots indicate $\lambda_2$ local maximums used for strainline integration initial positions.}
\label{f:bickley jet hyperbolic shear lcs details forward}
\end{figure}

\begin{figure}
\begin{center}
% demo/bickley_jet/hyperbolic_shear_lcs_details.m, commit 3331187
\includegraphics[width=\textwidth]{graphics/bickley_jet/hyperbolic_shear_lcs_details_backward}
\end{center}
\caption{Same LCS as in Figure~\ref{f:bickley jet hyperbolic shear lcs backward}, but with additional details. Background gray-scale is the finite-time Lyapunov exponent. Dashed green lines are Poincare sections. Green curves are closed shearlines with thick green curves are outermost shearlines, i.e. shear LCS. White dots indicate $\lambda_2$ local maximums used for strainline integration initial positions.}
\label{f:bickley jet hyperbolic shear lcs details backward}
\end{figure}

\begin{figure}
\begin{center}
% demo/bickley_jet/stretchlines.m, commit fc52909
\includegraphics[width=\textwidth]{graphics/bickley_jet/stretchlines_forward}
\end{center}
\caption{Bickley jet forward time stretchlines.}
\label{f:bickley jet forward stretchlines}
\end{figure}

\begin{figure}
\begin{center}
% demo/bickley_jet/stretchlines.m, commit fc52909
\includegraphics[width=\textwidth]{graphics/bickley_jet/stretchlines_backward}
\end{center}
\caption{Bickley jet backward time stretchlines.}
\label{f:bickley jet backward stretchlines}
\end{figure}

\subsection{Ocean dataset}

\begin{itemize}
\item auxiliaryGridRelativeDelta = 0.1
\item eigenvalueFromMainGrid = false;
\item Cauchy-Green resolution must be higher than dataset
\end{itemize}

\begin{figure}
\begin{center}
% demo/ocean_dataset/hyperbolic_shear_lcs.m, commit 65fab59
\includegraphics[width=\textwidth]{graphics/ocean_dataset/hyperbolic_shear_lcs_forward}
\end{center}
\caption{Ocean dataset forward time shear (green) and hyperbolic (red) LCS.}
\label{f:ocean dataset hyperbolic shear lcs forward}
\end{figure}

\begin{figure}
\begin{center}
% demo/ocean_dataset/hyperbolic_shear_lcs.m, commit 65fab59
\includegraphics[width=\textwidth]{graphics/ocean_dataset/hyperbolic_shear_lcs_backward}
\end{center}
\caption{Ocean dataset backward time shear (green) and hyperbolic (blue) LCS.}
\label{f:ocean dataset hyperbolic shear lcs backward}
\end{figure}

\begin{figure}
\begin{center}
% demo/ocean_dataset/hyperbolic_shear_lcs_details.m, commit 65fab59
\includegraphics[width=\textwidth]{graphics/ocean_dataset/hyperbolic_shear_lcs_details_forward}
\end{center}
\caption{Same LCS as in Figure~\ref{f:ocean dataset hyperbolic shear lcs forward}, but with additional details. Background gray-scale is the finite-time Lyapunov exponent. Dashed green lines are Poincare sections. Green curves are closed shearlines with thick green curves are outermost shearlines, i.e. shear LCS. White dots indicate $\lambda_2$ local maximums used for strainline integration initial positions.}
\label{f:ocean dataset hyperbolic shear lcs details forward}
\end{figure}

\begin{figure}
\begin{center}
% demo/ocean_dataset/hyperbolic_shear_lcs_details.m, commit 65fab59
\includegraphics[width=\textwidth]{graphics/ocean_dataset/hyperbolic_shear_lcs_details_backward}
\end{center}
\caption{Same LCS as in Figure~\ref{f:ocean dataset hyperbolic shear lcs backward}, but with additional details. Background gray-scale is the finite-time Lyapunov exponent. Dashed green lines are Poincare sections. Green curves are closed shearlines with thick green curves are outermost shearlines, i.e. shear LCS. White dots indicate $\lambda_2$ local maximums used for strainline integration initial positions.}
\label{f:ocean dataset hyperbolic shear lcs details backward}
\end{figure}

% \subsection{Adaptive Step Size Integration of Strainlines and Shearlines}

% \begin{itemize}
% \item Motivate necessity for adaptive step size integration. Illustrate (concentrated) high curvature of strainlines and shearlines.
% \end{itemize}

% Adaptive step size integration algorithms evaluate the right hand side of ODEs in a non-sequential way. Thus, the notion of evaluation at a previous time-step presented in~\textcite[Equation 9]{farazmand12:_comput_lagran} requires the use of advanced programming concepts if using MATLAB's numerical differential equation integration functions (e.g. ode45, ode113.) Specifically, the LCS Toolbox implementation of strainline and shearline integration uses output functions and handle classes.

% \subsection{Equation of variation to calculate differential of flow map}

% In~\textcite{farazmand12:_comput_lagran} an auxiliary grid is used to calculate the gradient of the flow map, $DF_{t_0}^t$. Experience suggests the quality of results obtained with this method is variable. An alternative is to use the equation of variation.

% The flow is given by an equation of the form:
% \[
% \dot{x} = v(x,t).
% \]
% Application of the chain rule gives the equation of variation:
% \begin{align}
% \frac{\partial}{\partial x_0} \dot{x} &= Dv(x,t) \frac{\partial x}{\partial x_0}\nonumber\\
% \frac{d}{dt} DF_{t_0}^t &= Dv(x,t) DF_{t_0}^t\label{e:equation of variation}
% \end{align}
% Equation \eqref{e:equation of variation} is the equation of variation and an initial value problem ordinary differential equations system. The initial condition, $DF_{t_0}^{t_0}$, is the identity matrix.

% For two dimensional flows,
% \[
% DF_{t_0}^t = 
% \begin{pmatrix}
% \frac{\partial}{\partial x_1} F_{{t_0}_1}^t & \frac{\partial}{\partial x_2} F_{{t_0}_1}^t\\
% \frac{\partial}{\partial x_1} F_{{t_0}_2}^t & \frac{\partial}{\partial x_2} F_{{t_0}_2}^t
% \end{pmatrix},
% \]
% and
% \[
% Dv(x,t) =
% \begin{pmatrix}
% \frac{\partial}{\partial x_1} v_1 & \frac{\partial}{\partial x_2} v_1\\
% \frac{\partial}{\partial x_1} v_2 & \frac{\partial}{\partial x_2} v_2
% \end{pmatrix}.
% \]
% With the equation of variation, calculating the differential of the flow map consists of solving an ODE system of six equations: $x_1,x_2$, and the four elements of $DF_{t_0}^t$. The equations for $DF_{t_0}^t$ depend on $x_1,x_2$, but not vice-versa. To solve this system, the gradient of the flow vector field, $Dv(x,t)$ must be known. If the velocity field is known analytically, $Dv(x,t)$ can also be calculated analytically.
% In Appendix \ref{s:traveling wave input file}, the necessary commands are given to do this automatically with the MATLAB Symbolic Math Toolbox for a traveling wave flow (see functions symF and symJacDerivative)
% If the flow map is not known analytically, splines can be fitted to flow data and the splines differentiated (see Section~\ref{s:splines}.)

% To test the equation of variation method and to compare it to the finite difference method from~\parencite{farazmand12:_comput_lagran}, a double gyre incompressible flow is used. Incompressible flows have a Cauchy-Green strain tensor whose exact value is one, $|C_{t_0}^T| = 1$ (and a maximum possible value of $\lambda_1 = 1$.) In Figure~\ref{f:cgStrain convergence}, the convergence of $C_{t_0}^T$ is illustrated; the finite difference with auxiliary grid method and the equation of variation method are compared. To observe convergence with the finite difference with auxiliary grid method, the auxiliary grid spacing was found by trial and error; if the auxiliary grid relative spacing is set larger than $10^{-8}$ convergence is difficult to observe.
% Setting the auxiliary grid relative spacing any smaller than $10^{-8}$ does not seem to help.
% The data in the figure indicates the error of the equation of variation method is smaller than the finite difference with auxiliary grid method. In addition, the equation of variation does not require setting an auxiliary grid spacing parameter, the equation of variation method is better than the finite difference with auxiliary grid method.
% The convergence measure used, $|||C_{t_0}^T| - 1||_2$, is an average. The error can be made no smaller than $10^{-7} at a resolution of 20 \times 10. As the resolution increases, the smallest possible error becomes larger.

% \begin{figure}
% \begin{center}
% \includegraphics[width=\textwidth]{../../lcs_toolbox/figures/double_gyre/compare_fd_eov}
% \end{center}
% \caption{$|||C_{t_0}^T| - 1||_2$ as a function of ODE integration error tolerance. The ODE solver is MATLAB's ode45. Both the relative and absolute error tolerances are set to the stated values. For the finite difference method, the Cauchy-Green strain is obtained from the auxiliary grid with the auxiliary grid spacing $10^{-8}$ of the main grid spacing. A double gyre flow is used. The flow resolution is $20 \times 10$. The data is given in Table~\ref{t:Cauchy-Green strain convergence comparison}.}
% \label{f:cgStrain convergence}
% \end{figure}

% $|C_{t_0}^T|$ converges as the flow integration error tolerance decreases and $\lambda_1$ converges as the flow resolution is increased. It may be worth noting that $|C_{t_0}^T|$ converge as the resolution increases since it is a quantity calculated at each grid point independently. In a complementary way, $\lambda_1$ does not converge as the error tolerance is decreased.

% In Table~\ref{t:lambda products tol}, the ODE solver's relative error is varied. As the error tolerance decreases $|C_{t_0}^T|$ converges to 1 (except for the last row, perhaps because the relative error tolerance is smaller than the absolute error tolerance.) This is logical; whereas in Table~\ref{t:lambda products res} the number of points was changing, in Table~\ref{t:lambda products tol} it is fixed. By decreasing the error tolerance, the integration at each point becomes more accurate and $|||C_{t_0}^T| - 1||_2$ approaches its exact value. The convergence of $||\lambda_1||_\infty$, if it exists at all, seems to saturate relatively quickly, once the relative error tolerance is $10^{-4}$. The interpretation of this result is that for the flow resolution chosen for the data in Table~\ref{t:lambda products tol}, $||\lambda_1||_\infty$ has a set value and finding this value accurately does not require a particularly tight error tolerance.

% Note that in tables~\ref{t:lambda products res} and~\ref{t:lambda products tol}, for those quantities that exhibit convergence, the rate of error convergence has not been verified.

% \begin{table}
% \begin{center}
% \begin{tabular}{l c c}
% \hline
% Resolution & $|||C_{t_0}^T| - 1||_2$ & $||\lambda_1||_\infty$\\
% \hline
% $200 \times 100$ & $2.50 \times 10^{-4}$ & 0.896\\
% $400 \times 200$ & $2.47 \times 10^{-3}$ & 0.976\\
% $800 \times 400$ & $1.72 \times 10^{-1}$ & 0.976\\
% $1600 \times 800$ & 14.6 & 0.991\\
% $3200 \times 1600$ & 4.24 & 0.992\\
% \hline
% \end{tabular}
% \end{center}
% \caption{Properties of the Cauchy-Green strain tensor, $C$, and $\lambda_1$ at different resolutions. The relative and absolute error tolerances are $10^{-4}$ and $10^{-6}$ respectively.}
% \label{t:lambda products res}
% \end{table}

% \begin{table}
% \begin{center}
% \begin{tabular}{l c c}
% \hline
% Rel. tol. & $|||C_{t_0}^T| - 1||_2$ & $||\lambda_1||_\infty$\\
% \hline
% $10^{-2}$ & $1.24 \times 10^{-1}$ & 1.031\\
% $10^{-3}$ & $6.13 \times 10^{-3}$ & 0.963\\
% $10^{-4}$ & $3.53 \times 10^{-3}$ & 0.962\\
% $10^{-5}$ & $1.37 \times 10^{-3}$ & 0.962\\
% $10^{-6}$ & $1.28 \times 10^{-3}$ & 0.962\\
% $10^{-8}$ & $2.47 \times 10^{-3}$ & 0.962\\
% \hline
% \end{tabular}
% \end{center}
% \caption{Properties of the Cauchy-Green strain tensor, $C$, and $\lambda_1$ at different relative error tolerances. The resolution is $500\times250$. The absolute error tolerance is $10^{-6}$.}
% \label{t:lambda products tol}
% \end{table}

% \subsection{Splines}
% \label{s:splines}

% To integrate the equation of variation, the derivative of the flow vector field, $Dv(x,t)$, must be known. When working with data-defined flows rather than analytic flows, $Dv(x,t)$ can be found by fitting splines to the data and differentiating the splines~\parencite{finn05:_volum}.

% This procedure is implemented with the MATLAB Curve Fitting Toolbox. For validation, a spline is created by sampling an analytic flow.% using the function analytic2spline (see Appendix \ref{s:analytic2spline}.)

% In Table~\ref{t:splines} the double gyre is used for comparison; the spline is created at various resolutions and then compared at one point to the analytic definition. The data show as sampling resolution increases, the spline approximation error decreases. The resolutions in Table~\ref{t:splines} are lower than those in Table~\ref{t:lambda products res}. Although systematic tests have not been performed, it appears using splines uses more memory than the analytic approach.

% \begin{table}
% \begin{center}
% \begin{tabular}{l c c}
% \hline
% Resolution & $|v - v_\text{spline}|$ & $|Dv - Dv_\text{spline}|$\\
% \hline
% $10 \times 5$ & $4.18 \times 10^{-2}$ & $3.71 \times 10^{-1}$\\
% $20 \times 10$ & $8.44 \times 10^{-3}$ & $1.12 \times 10^{-1}$\\
% $40 \times 20$ & $3.72 \times 10^{-3}$ & $6.00 \times 10^{-2}$\\
% $80 \times 40$ & $2.51 \times 10^{-3}$ & $4.33 \times 10^{-2}$\\
% $160 \times 80$ & $2.06 \times 10^{-3}$ & $3.66 \times 10^{-2}$\\
% \hline
% \end{tabular}
% \end{center}
% \caption{Numerical convergence of splines for double gyre flow. Time = 0 and position = (-0.25,0.9).}
% \label{t:splines}
% \end{table}

% \subsection{Superminimization}
% \label{s:superminimzation}

\section{Conclusions}

Algorithms to identify LCSs using variational principles have been presented, and a software library, LCS Tool, that implements these algorithms has been demonstrated on geophysical flow models and a dataset. The software library provides functions to enable exploring variational LCS methods without requiring detailed knowledge of variational LCS theory. These functions make it possible to write compact programs identify and plot LCSs.

LCS Tool is built atop MATLAB\parencite{mathworks13:_matlab} and leverages that software's capabilities extensively. With FTLE-based LCS identification, computational performance has received considerable attention\parencite{conti12:_gpu_apu_finit_time_lyapun_expon,miron12:_anisot_lagran_coher_struc}. LCS Tool is free software and we hope it will evolve to incorporate similar advances. Optimising computational performance will likely help apply variational LCS methods to large-scale real-time forecasting applications, for example to allow tracking environmental contaminants.

% Reduce number of user inputs, for example, automated closed orbit detection, which is especially important for flows with a large number of vortices possible.

In addition to improving computational performance, we hope LCS Tool will be the foundation to integrate new theoretical advances, for example the variational LCS jet core identification method\parencite{farazmand13:_shearless} and the 
analysis of LCSs in higher dimensions\parencite{blazevski:_hyper_ellip_trans_barrier_three}.

\printbibliography

\end{document}

% Local Variables:
% ispell-dictionary: "en_GB"
% End:
% LocalWords:  stretchline stretchlines incompressible gyre eigenvector eigenvectors incompressibility dataset datasets LCS strainline shearline Lyapunov Bickley Rossby timespan variational adaptively vortices parametrise advected

### 2013-11-01 graphics/bickley_jet/hyperbolic_shear_lcs_details_pdf.m

@@ -1,0 +1,35 @@
+function hyperbolic_shear_lcs_details_pdf
+
+lcs_tool_root = fullfile('..','..','..','..','lcs_toolbox');
+bickley_jet_demo = fullfile(lcs_tool_root,'demo','bickley_jet');
+
+addpath(lcs_tool_root)
+addpath(bickley_jet_demo)
+
+hyperbolic_shear_lcs_details
+
+hFigure = 1;
+delete_title(hFigure)
+
+filename = 'hyperbolic_shear_lcs_details_forward.pdf';
+if exist(filename,'file')
+    delete(filename)
+end
+print_pdf(hFigure,filename)
+
+% FIXME hFigure should not be hard-coded; should probably search for
+% matching title string
+hFigure = 12;
+delete_title(hFigure)
+
+filename = 'hyperbolic_shear_lcs_details_backward.pdf';
+if exist(filename,'file')
+    delete(filename)
+end
+print_pdf(hFigure,filename)
+
+function delete_title(hFigure)
+
+hAxes = findobj(hFigure,'type','axes','-not','Tag','Colorbar');
+hTitle = get(hAxes,'title');
+delete(hTitle)

### 2013-11-01 graphics/bickley_jet/hyperbolic_shear_lcs_pdf.m

@@ -1,0 +1,37 @@
+function hyperbolic_shear_lcs_pdf
+
+lcs_tool_root = fullfile('..','..','..','..','lcs_toolbox');
+bickley_jet_demo = fullfile(lcs_tool_root,'demo','bickley_jet');
+
+addpath(lcs_tool_root)
+addpath(bickley_jet_demo)
+
+hyperbolic_shear_lcs
+
+hFigure = 1;
+filename = 'hyperbolic_shear_lcs_forward';
+savefig(hFigure,filename)
+
+delete_title(hFigure)
+
+if exist(filename,'file')
+    delete(filename)
+end
+print_pdf(hFigure,filename)
+
+hFigure = 2;
+filename = 'hyperbolic_shear_lcs_backward';
+savefig(hFigure,filename)
+
+delete_title(hFigure)
+
+if exist(filename,'file')
+    delete(filename)
+end
+print_pdf(hFigure,filename)
+
+function delete_title(hFigure)
+
+hAxes = get(hFigure,'children');
+hTitle = get(hAxes,'title');
+delete(hTitle)

### 2013-11-01 graphics/double_gyre/hyperbolic_shear_lcs_details_pdf.m

@@ -1,0 +1,35 @@
+function hyperbolic_shear_lcs_details_pdf
+
+lcs_tool_root = fullfile('..','..','..','..','lcs_toolbox');
+double_gyre_demo = fullfile(lcs_tool_root,'demo','double_gyre');
+
+addpath(lcs_tool_root)
+addpath(double_gyre_demo)
+
+hyperbolic_shear_lcs_details
+
+hFigure = 1;
+delete_title(hFigure)
+
+filename = 'hyperbolic_shear_lcs_details_forward.pdf';
+if exist(filename,'file')
+    delete(filename)
+end
+print_pdf(hFigure,filename)
+
+% FIXME hFigure should not be hard-coded; should probably search for title
+% string
+hFigure = 6;
+delete_title(hFigure)
+
+filename = 'hyperbolic_shear_lcs_details_backward.pdf';
+if exist(filename,'file')
+    delete(filename)
+end
+print_pdf(hFigure,filename)
+
+function delete_title(hFigure)
+
+hAxes = findobj(hFigure,'type','axes','-not','Tag','Colorbar');
+hTitle = get(hAxes,'title');
+delete(hTitle)

### 2013-11-01 graphics/ocean_dataset/hyperbolic_shear_lcs_pdf.m

@@ -1,0 +1,37 @@
+function hyperbolic_shear_lcs_pdf
+
+lcs_tool_root = fullfile('..','..','..','..','lcs_toolbox');
+ocean_dataset_demo = fullfile(lcs_tool_root,'demo','ocean_dataset');
+
+addpath(lcs_tool_root)
+addpath(ocean_dataset_demo)
+
+hyperbolic_shear_lcs
+
+hFigure = 1;
+filename = 'hyperbolic_shear_lcs_forward';
+savefig(hFigure,filename)
+
+delete_title(hFigure)
+
+if exist([filename,'.pdf'],'file')
+    delete([filename,'.pdf'])
+end
+print_pdf(hFigure,filename)
+
+hFigure = 2;
+filename = 'hyperbolic_shear_lcs_backward';
+savefig(hFigure,filename)
+
+delete_title(hFigure)
+
+if exist([filename,'.pdf'],'file')
+    delete([filename,'.pdf'])
+end
+print_pdf(hFigure,filename)
+
+function delete_title(hFigure)
+
+hAxes = get(hFigure,'children');
+hTitle = get(hAxes,'title');
+delete(hTitle)

### 2013-11-04 graphics/ocean_dataset/hyperbolic_shear_lcs_details_pdf.m
Correct ocean dataset graphics

@@ -1,4 +1,4 @@
-function hyperbolic_shear_lcs_pdf
+function hyperbolic_shear_lcs_details_pdf
 
 lcs_tool_root = fullfile('..','..','..','..','lcs_toolbox');
 ocean_dataset_demo = fullfile(lcs_tool_root,'demo','ocean_dataset');
@@ -7,11 +7,11 @@ oldFolder = cd(ocean_dataset_demo);
 
 try
     addpath(fullfile('..','..'))
-    hyperbolic_shear_lcs
+    hyperbolic_shear_lcs_details
     cd(oldFolder)
     
     hFigure = 1;
-    filename = 'hyperbolic_shear_lcs_forward';
+    filename = 'hyperbolic_shear_lcs_details_forward';
     savefig(hFigure,filename)
     delete_title(hFigure)
     filenamePDF = [filename,'.pdf'];
@@ -20,8 +20,10 @@ try
     end
     print_pdf(hFigure,filename)
     
-    hFigure = 2;
-    filename = 'hyperbolic_shear_lcs_backward';
+    % FIXME hFigure should not be hard-coded; should probably search for
+    % matching title string
+    hFigure = 12;
+    filename = 'hyperbolic_shear_lcs_details_backward';
     savefig(hFigure,filename)
     delete_title(hFigure)
     filenamePDF = [filename,'.pdf'];
@@ -29,6 +31,7 @@ try
         delete(filenamePDF)
     end
     print_pdf(hFigure,filename)
+
 catch err
     cd(oldFolder)
     rethrow(err)
@@ -36,6 +39,6 @@ end
 
 function delete_title(hFigure)
 
-hAxes = get(hFigure,'children');
+hAxes = findobj(hFigure,'type','axes','-not','Tag','Colorbar');
 hTitle = get(hAxes,'title');
 delete(hTitle)

### 2013-11-11 main.tex
Correct ocean dataset graphics
Update double gyre figures to substitude backward time analysis by stretchlines.

@@
 \begin{center}
-\includegraphics[width=\textwidth]{variable_step_integration}
+\includegraphics[width=\textwidth]{graphics/variable_step_integration}
 \end{center}
@@
 \begin{center}
-% LCS Tool commit ac3a7705aec81577b1aba3cafe598066dd689203
-\includegraphics[width=\textwidth]{graphics/double_gyre/hyperbolic_shear_lcs_forward}
+\includegraphics[width=\textwidth]{graphics/double_gyre/hyperbolic_shear_lcs_strainline}
 \end{center}
-\caption{Double gyre forward time shear (green) and hyperbolic (red) LCS.}
-\label{f:double gyre hyperbolic shear lcs forward}
+\caption{Double gyre lambda-line (green) and strainline (red) LCSs.}
+\label{f:double gyre hyperbolic shear lcs strainline}
 \end{figure}
@@
 \begin{center}
-% LCS Tool commit ac3a7705aec81577b1aba3cafe598066dd689203
-\includegraphics[width=\textwidth]{graphics/double_gyre/hyperbolic_shear_lcs_backward}
+\includegraphics[width=\textwidth]{graphics/double_gyre/hyperbolic_shear_lcs_stretchline}
 \end{center}
-\caption{Double gyre backward time shear (green) and hyperbolic (blue) LCS.}
-\label{f:double gyre hyperbolic shear lcs backward}
+\caption{Double gyre lambda-line (green) and stretchline (blue) LCSs.}
+\label{f:double gyre hyperbolic shear lcs stretchline}
 \end{figure}
@@
 \begin{center}
-% LCS Tool commit ac3a7705aec81577b1aba3cafe598066dd689203
-\includegraphics[width=\textwidth]{graphics/double_gyre/hyperbolic_shear_lcs_details_forward}
+\includegraphics[width=\textwidth]{graphics/double_gyre/hyperbolic_shear_lcs_details_strainline}
 \end{center}
-\caption{Same LCS as in Figure~\ref{f:double gyre hyperbolic shear lcs forward}, but with additional details.
+\caption{Same LCS as in Figure~\ref{f:double gyre hyperbolic shear lcs strainline}, but with additional details.
 Background gray-scale is the finite-time Lyapunov exponent.
 Dashed green lines are Poincare sections.
-Green curves are closed shearlines with thick green curves are outermost shearlines, i.e.
-shear LCS.
+Green curves are closed lambda-lines with thick green curves being outermost closed lambda-lines, i.e.
+lambda-line LCSs.
 White dots indicate $\lambda_2$ local maximums used for strainline integration initial positions.}
@@
 % LCS Tool commit ac3a7705aec81577b1aba3cafe598066dd689203
-\includegraphics[width=\textwidth]{graphics/double_gyre/hyperbolic_shear_lcs_details_backward}
+\includegraphics[width=\textwidth]{graphics/double_gyre/hyperbolic_shear_lcs_details_stretchline}
 \end{center}
-\caption{Same LCS as in Figure~\ref{f:double gyre hyperbolic shear lcs backward}, but with additional details.
+\caption{Same LCS as in Figure~\ref{f:double gyre hyperbolic shear lcs stretchline}, but with additional details.
 Background gray-scale is the finite-time Lyapunov exponent.
 Dashed green lines are Poincare sections.
-Green curves are closed shearlines and thick green curves are outermost shearlines, i.e.
-shear LCS.
-White dots indicate $\lambda_2$ local maximums used for strainline integration intial positions.}
-\label{f:double gyre hyperbolic shear lcs details backward}
-\end{figure}
-
-Stretchlines are calculated using the function \texttt{compute\_stretchline}.
-Forward time stretchlines are shown in Figure~\ref{f:double gyre stretchlines forward} and backward time stretchlines are shown in Figure~\ref{f:double gyre stretchlines backward}.
-
-\begin{figure}
-\begin{center}
-% demo/double_gyre/stretchlines.m, commit dbefa5a
-\includegraphics[width=\textwidth]{graphics/double_gyre/stretchlines_forward}
-\end{center}
-\caption{Double gyre forward time stretchlines.}
-\label{f:double gyre stretchlines forward}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-% demo/double_gyre/stretchlines.m, commit dbefa5a
-\includegraphics[width=\textwidth]{graphics/double_gyre/stretchlines_backward}
-\end{center}
-\caption{Double gyre backward time stretchlines.}
-\label{f:double gyre stretchlines backward}
+Green curves are closed lambda-lines and thick green curves are outermost lambda-lines, i.e.
+lambda-line LCSs.
+White dots indicate $\lambda_1$ local minimums used for stretchline integration intial positions.}
+\label{f:double gyre hyperbolic shear lcs details stretchline}
 \end{figure}

### 2013-11-11 graphics/double_gyre/hyperbolic_shear_lcs_details_pdf.m

@@ -1,32 +1,40 @@
 function hyperbolic_shear_lcs_details_pdf
 
 lcs_tool_root = fullfile('..','..','..','..','lcs_toolbox');
-double_gyre_demo = fullfile(lcs_tool_root,'demo','double_gyre');
+demo = fullfile(lcs_tool_root,'demo','double_gyre');
 
-addpath(lcs_tool_root)
-addpath(double_gyre_demo)
+oldFolder = cd(demo);
 
-hyperbolic_shear_lcs_details
-
-hFigure = 1;
-delete_title(hFigure)
-
-filename = 'hyperbolic_shear_lcs_details_forward.pdf';
-if exist(filename,'file')
-    delete(filename)
+try
+    addpath(fullfile('..','..'))
+    hyperbolic_shear_lcs_details
+    cd(oldFolder)
+    
+    hFigure = 1;
+    filename = 'hyperbolic_shear_lcs_details_strainline';
+    savefig(hFigure,filename)
+    delete_title(hFigure)
+    filenamePDF = [filename,'.pdf'];
+    if exist(filenamePDF,'file')
+        delete(filenamePDF)
+    end
+    print_pdf(hFigure,filename)
+    
+    % FIXME hFigure should not be hard-coded; should probably search for title
+    % string
+    hFigure = 6;
+    filename = 'hyperbolic_shear_lcs_details_stretchline';
+    savefig(hFigure,filename)
+    delete_title(hFigure)
+    filenamePDF = [filename,'.pdf'];
+    if exist(filenamePDF,'file')
+        delete(filenamePDF)
+    end
+    print_pdf(hFigure,filename)
+catch err
+    cd(oldFolder)
+    rethrow(err)
 end
-print_pdf(hFigure,filename)
-
-% FIXME hFigure should not be hard-coded; should probably search for title
-% string
-hFigure = 6;
-delete_title(hFigure)
-
-filename = 'hyperbolic_shear_lcs_details_backward.pdf';
-if exist(filename,'file')
-    delete(filename)
-end
-print_pdf(hFigure,filename)
 
 function delete_title(hFigure)
 

### 2013-11-11 main.tex
Change Bickley jet figures from backward time strainlines to forward time stretchlines

@@
 
-The forward time LCS obtained are shown in Figure~\ref{f:bickley jet hyperbolic shear lcs forward} and the backward time LCS (i.e.
-with the flow timespan set to $t \in [4 L_x/U,0]$ are shown in Figure~\ref{f:bickley jet hyperbolic shear lcs backward}.
+The forward time LCS obtained are shown in Figure~\ref{f:bickley jet hyperbolic shear lcs strainline} and the backward time LCS (i.e.
+with the flow timespan set to $t \in [4 L_x/U,0]$ are shown in Figure~\ref{f:bickley jet hyperbolic shear lcs stretchline}.
 In Figures~\ref{f:bickley jet hyperbolic shear lcs details forward} and~\ref{f:bickley jet hyperbolic shear lcs details backward} the same LCS are shown with details.
-
-Forward time stretchlines are shown in Figure~\ref{f:bickley jet forward stretchlines} and backward time stretchlines are shown in Figure~\ref{f:bickley jet backward stretchlines}.
 
@@
 \begin{center}
-% demo/bickley_jet/hyperbolic_shear_lcs.m, commit aaf20ed
-\includegraphics[width=\textwidth]{graphics/bickley_jet/hyperbolic_shear_lcs_forward}
+\includegraphics[width=\textwidth]{graphics/bickley_jet/hyperbolic_shear_lcs_strainline}
 \end{center}
-\caption{Bickley jet forward time shear (green) and hyperbolic (red) LCS.}
-\label{f:bickley jet hyperbolic shear lcs forward}
+\caption{Bickley jet lambda-line (green) and strainline (red) LCS.}
+\label{f:bickley jet hyperbolic shear lcs strainline}
 \end{figure}
@@
 \begin{center}
-% demo/bickley_jet/hyperbolic_shear_lcs.m, commit aaf20ed
-\includegraphics[width=\textwidth]{graphics/bickley_jet/hyperbolic_shear_lcs_backward}
+\includegraphics[width=\textwidth]{graphics/bickley_jet/hyperbolic_shear_lcs_stretchline}
 \end{center}
-\caption{Bickley jet backward time shear (green) and hyperbolic (blue) LCS.}
-\label{f:bickley jet hyperbolic shear lcs backward}
+\caption{Bickley jet backward time shear (green) and stretchline (blue) LCS.}
+\label{f:bickley jet hyperbolic shear lcs stretchline}
 \end{figure}
@@
 \begin{center}
-% demo/bickley_jet/hyperbolic_shear_lcs_details.m, commit 3331187
-\includegraphics[width=\textwidth]{graphics/bickley_jet/hyperbolic_shear_lcs_details_forward}
+\includegraphics[width=\textwidth]{graphics/bickley_jet/hyperbolic_shear_lcs_details_strainline}
 \end{center}
@@
 % demo/bickley_jet/hyperbolic_shear_lcs_details.m, commit 3331187
-\includegraphics[width=\textwidth]{graphics/bickley_jet/hyperbolic_shear_lcs_details_backward}
+\includegraphics[width=\textwidth]{graphics/bickley_jet/hyperbolic_shear_lcs_details_stretchline}
 \end{center}
@@
 \label{f:bickley jet hyperbolic shear lcs details backward}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-% demo/bickley_jet/stretchlines.m, commit fc52909
-\includegraphics[width=\textwidth]{graphics/bickley_jet/stretchlines_forward}
-\end{center}
-\caption{Bickley jet forward time stretchlines.}
-\label{f:bickley jet forward stretchlines}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-% demo/bickley_jet/stretchlines.m, commit fc52909
-\includegraphics[width=\textwidth]{graphics/bickley_jet/stretchlines_backward}
-\end{center}
-\caption{Bickley jet backward time stretchlines.}
-\label{f:bickley jet backward stretchlines}
 \end{figure}
@@
 \begin{center}
-% demo/ocean_dataset/hyperbolic_shear_lcs.m, commit 65fab59
-\includegraphics[width=\textwidth]{graphics/ocean_dataset/hyperbolic_shear_lcs_forward}
+\includegraphics[width=\textwidth]{graphics/ocean_dataset/hyperbolic_shear_lcs_strainline}
 \end{center}
 \caption{Ocean dataset forward time shear (green) and hyperbolic (red) LCS.}
-\label{f:ocean dataset hyperbolic shear lcs forward}
+\label{f:ocean dataset hyperbolic shear lcs strainline}
 \end{figure}
@@
 \begin{center}
-% demo/ocean_dataset/hyperbolic_shear_lcs.m, commit 65fab59
-\includegraphics[width=\textwidth]{graphics/ocean_dataset/hyperbolic_shear_lcs_backward}
+\includegraphics[width=\textwidth]{graphics/ocean_dataset/hyperbolic_shear_lcs_stretchline}
 \end{center}
 \caption{Ocean dataset backward time shear (green) and hyperbolic (blue) LCS.}
-\label{f:ocean dataset hyperbolic shear lcs backward}
+\label{f:ocean dataset hyperbolic shear lcs stretchline}
 \end{figure}
@@
 \begin{center}
-% demo/ocean_dataset/hyperbolic_shear_lcs_details.m, commit 65fab59
-\includegraphics[width=\textwidth]{graphics/ocean_dataset/hyperbolic_shear_lcs_details_forward}
+\includegraphics[width=\textwidth]{graphics/ocean_dataset/hyperbolic_shear_lcs_details_strainline}
 \end{center}
@@
 \begin{center}
-% demo/ocean_dataset/hyperbolic_shear_lcs_details.m, commit 65fab59
-\includegraphics[width=\textwidth]{graphics/ocean_dataset/hyperbolic_shear_lcs_details_backward}
+\includegraphics[width=\textwidth]{graphics/ocean_dataset/hyperbolic_shear_lcs_details_stretchline}
 \end{center}

### 2013-11-11 graphics/bickley_jet/hyperbolic_shear_lcs_details_pdf.m

@@ -1,32 +1,40 @@
 function hyperbolic_shear_lcs_details_pdf
 
 lcs_tool_root = fullfile('..','..','..','..','lcs_toolbox');
-bickley_jet_demo = fullfile(lcs_tool_root,'demo','bickley_jet');
+demo = fullfile(lcs_tool_root,'demo','bickley_jet');
 
-addpath(lcs_tool_root)
-addpath(bickley_jet_demo)
+oldFolder = cd(demo);
 
-hyperbolic_shear_lcs_details
-
-hFigure = 1;
-delete_title(hFigure)
-
-filename = 'hyperbolic_shear_lcs_details_forward.pdf';
-if exist(filename,'file')
-    delete(filename)
+try
+    addpath(fullfile('..','..'))
+    hyperbolic_shear_lcs_details
+    cd(oldFolder)
+    
+    hFigure = 1;
+    filename = 'hyperbolic_shear_lcs_details_strainline';
+    savefig(hFigure,filename)
+    delete_title(hFigure)
+    filenamePDF = [filename,'.pdf'];
+    if exist(filenamePDF,'file')
+        delete(filenamePDF)
+    end
+    print_pdf(hFigure,filename)
+    
+    % FIXME hFigure should not be hard-coded; should probably search for title
+    % string
+    hFigure = 12;
+    filename = 'hyperbolic_shear_lcs_details_stretchline';
+    savefig(hFigure,filename)
+    delete_title(hFigure)
+    filenamePDF = [filename,'.pdf'];
+    if exist(filenamePDF,'file')
+        delete(filenamePDF)
+    end
+    print_pdf(hFigure,filename)
+catch err
+    cd(oldFolder)
+    rethrow(err)
 end
-print_pdf(hFigure,filename)
-
-% FIXME hFigure should not be hard-coded; should probably search for
-% matching title string
-hFigure = 12;
-delete_title(hFigure)
-
-filename = 'hyperbolic_shear_lcs_details_backward.pdf';
-if exist(filename,'file')
-    delete(filename)
-end
-print_pdf(hFigure,filename)
 
 function delete_title(hFigure)
 

### 2013-11-11 graphics/bickley_jet/hyperbolic_shear_lcs_pdf.m

@@ -1,34 +1,38 @@
 function hyperbolic_shear_lcs_pdf
 
 lcs_tool_root = fullfile('..','..','..','..','lcs_toolbox');
-bickley_jet_demo = fullfile(lcs_tool_root,'demo','bickley_jet');
+demo = fullfile(lcs_tool_root,'demo','bickley_jet');
 
-addpath(lcs_tool_root)
-addpath(bickley_jet_demo)
+oldFolder = cd(demo);
 
-hyperbolic_shear_lcs
-
-hFigure = 1;
-filename = 'hyperbolic_shear_lcs_forward';
-savefig(hFigure,filename)
-
-delete_title(hFigure)
-
-if exist(filename,'file')
-    delete(filename)
+try
+    addpath(fullfile('..','..'))
+    hyperbolic_shear_lcs
+    cd(oldFolder)
+    
+    hFigure = 1;
+    filename = 'hyperbolic_shear_lcs_strainline';
+    savefig(hFigure,filename)
+    delete_title(hFigure)
+    filenamePDF = [filename,'.pdf'];
+    if exist(filenamePDF,'file')
+        delete(filenamePDF)
+    end
+    print_pdf(hFigure,filename)
+    
+    hFigure = 2;
+    filename = 'hyperbolic_shear_lcs_stretchline';
+    savefig(hFigure,filename)
+    delete_title(hFigure)
+    filenamePDF = [filename,'.pdf'];
+    if exist(filenamePDF,'file')
+        delete(filenamePDF)
+    end
+    print_pdf(hFigure,filename)
+catch err
+    cd(oldFolder)
+    rethrow(err)
 end
-print_pdf(hFigure,filename)
-
-hFigure = 2;
-filename = 'hyperbolic_shear_lcs_backward';
-savefig(hFigure,filename)
-
-delete_title(hFigure)
-
-if exist(filename,'file')
-    delete(filename)
-end
-print_pdf(hFigure,filename)
 
 function delete_title(hFigure)
 

### 2013-11-11 graphics/ocean_dataset/hyperbolic_shear_lcs_pdf.m

@@ -1,9 +1,9 @@
 function hyperbolic_shear_lcs_pdf
 
 lcs_tool_root = fullfile('..','..','..','..','lcs_toolbox');
-ocean_dataset_demo = fullfile(lcs_tool_root,'demo','ocean_dataset');
+demo = fullfile(lcs_tool_root,'demo','ocean_dataset');
 
-oldFolder = cd(ocean_dataset_demo);
+oldFolder = cd(demo);
 
 try
     addpath(fullfile('..','..'))

### 2013-11-11 main.tex
Change ocean data-set demo from backward time strainlines to forward time stretchlines.

@@
 
-The forward time LCS obtained are shown in Figure~\ref{f:double gyre hyperbolic shear lcs forward} and the backward time LCS (i.e.
-with the flow timespan set to $t \in [20,0]$) are shown in Figure~\ref{f:double gyre hyperbolic shear lcs backward}.
-In Figures~\ref{f:double gyre hyperbolic shear lcs details forward} and~\ref{f:double gyre hyperbolic shear lcs details backward} the same LCS are shown with details: the finite-time Lyapunov exponent, closed shearlines inside the outermost shearlines and $\lambda_2$ local maximums used for strainline integration initial positions.
+The forward time LCS obtained are shown in Figure~\ref{f:double gyre hyperbolic shear lcs strainline} and the backward time LCS (i.e.
+with the flow timespan set to $t \in [20,0]$) are shown in Figure~\ref{f:double gyre hyperbolic shear lcs stretchline}.
+In Figures~\ref{f:double gyre hyperbolic shear lcs details strainline} and~\ref{f:double gyre hyperbolic shear lcs details stretchline} the same LCS are shown with details: the finite-time Lyapunov exponent, closed shearlines inside the outermost shearlines and $\lambda_2$ local maximums used for strainline integration initial positions.
 
@@
 White dots indicate $\lambda_2$ local maximums used for strainline integration initial positions.}
-\label{f:double gyre hyperbolic shear lcs details forward}
+\label{f:double gyre hyperbolic shear lcs details strainline}
 \end{figure}
@@
 \begin{center}
-% LCS Tool commit ac3a7705aec81577b1aba3cafe598066dd689203
 \includegraphics[width=\textwidth]{graphics/double_gyre/hyperbolic_shear_lcs_details_stretchline}
@@
 Dashed green lines are Poincare sections.
-Green curves are closed lambda-lines and thick green curves are outermost lambda-lines, i.e.
+Green curves are closed lambda-lines and thick green curves are outermost closed lambda-lines, i.e.
 lambda-line LCSs.
@@
 with the flow timespan set to $t \in [4 L_x/U,0]$ are shown in Figure~\ref{f:bickley jet hyperbolic shear lcs stretchline}.
-In Figures~\ref{f:bickley jet hyperbolic shear lcs details forward} and~\ref{f:bickley jet hyperbolic shear lcs details backward} the same LCS are shown with details.
+In Figures~\ref{f:bickley jet hyperbolic shear lcs details strainline} and~\ref{f:bickley jet hyperbolic shear lcs details stretchline} the same LCS are shown with details.
 
@@
 \end{center}
-\caption{Same LCS as in Figure~\ref{f:bickley jet hyperbolic shear lcs forward}, but with additional details.
+\caption{Same LCS as in Figure~\ref{f:bickley jet hyperbolic shear lcs strainline}, but with additional details.
 Background gray-scale is the finite-time Lyapunov exponent.
 Dashed green lines are Poincare sections.
-Green curves are closed shearlines with thick green curves are outermost shearlines, i.e.
-shear LCS.
+Green curves are closed lambda-lines with thick green curves being outermost closed lambda-lines, i.e.
+lambda-line LCSs.
 White dots indicate $\lambda_2$ local maximums used for strainline integration initial positions.}
-\label{f:bickley jet hyperbolic shear lcs details forward}
+\label{f:bickley jet hyperbolic shear lcs details strainline}
 \end{figure}
@@
 \begin{center}
-% demo/bickley_jet/hyperbolic_shear_lcs_details.m, commit 3331187
 \includegraphics[width=\textwidth]{graphics/bickley_jet/hyperbolic_shear_lcs_details_stretchline}
 \end{center}
-\caption{Same LCS as in Figure~\ref{f:bickley jet hyperbolic shear lcs backward}, but with additional details.
+\caption{Same LCS as in Figure~\ref{f:bickley jet hyperbolic shear lcs stretchline}, but with additional details.
 Background gray-scale is the finite-time Lyapunov exponent.
 Dashed green lines are Poincare sections.
-Green curves are closed shearlines with thick green curves are outermost shearlines, i.e.
-shear LCS.
-White dots indicate $\lambda_2$ local maximums used for strainline integration initial positions.}
-\label{f:bickley jet hyperbolic shear lcs details backward}
+Green curves are closed lambda-lines with thick green curves being outermost closed lambda-lines, i.e.
+lambda-line LCSs.
+White dots indicate $\lambda_2$ local maximums used for stretchline integration initial positions.}
+\label{f:bickley jet hyperbolic shear lcs details stretchline}
 \end{figure}
@@
 \end{center}
-\caption{Ocean dataset forward time shear (green) and hyperbolic (red) LCS.}
+\caption{Ocean dataset lambda-line (green) and strainline (red) LCS.}
 \label{f:ocean dataset hyperbolic shear lcs strainline}
@@
 \end{center}
-\caption{Ocean dataset backward time shear (green) and hyperbolic (blue) LCS.}
+\caption{Ocean dataset lambda-line (green) and stretchline (blue) LCSs.}
 \label{f:ocean dataset hyperbolic shear lcs stretchline}
@@
 \end{center}
-\caption{Same LCS as in Figure~\ref{f:ocean dataset hyperbolic shear lcs forward}, but with additional details.
+\caption{Same LCS as in Figure~\ref{f:ocean dataset hyperbolic shear lcs strainline}, but with additional details.
 Background gray-scale is the finite-time Lyapunov exponent.
 Dashed green lines are Poincare sections.
-Green curves are closed shearlines with thick green curves are outermost shearlines, i.e.
-shear LCS.
+Green curves are closed lambda-lines with thick green curves being outermost closed lambda-lines, i.e.
+lambda-line LCSs.
 White dots indicate $\lambda_2$ local maximums used for strainline integration initial positions.}
-\label{f:ocean dataset hyperbolic shear lcs details forward}
+\label{f:ocean dataset hyperbolic shear lcs details strainline}
 \end{figure}
@@
 \end{center}
-\caption{Same LCS as in Figure~\ref{f:ocean dataset hyperbolic shear lcs backward}, but with additional details.
+\caption{Same LCS as in Figure~\ref{f:ocean dataset hyperbolic shear lcs stretchline}, but with additional details.
 Background gray-scale is the finite-time Lyapunov exponent.
 Dashed green lines are Poincare sections.
-Green curves are closed shearlines with thick green curves are outermost shearlines, i.e.
-shear LCS.
-White dots indicate $\lambda_2$ local maximums used for strainline integration initial positions.}
-\label{f:ocean dataset hyperbolic shear lcs details backward}
+Green curves are closed lambda-lines and thick green curves are outermost closed lambda-lines, i.e.
+lambda-line LCSs.
+White dots indicate $\lambda_1$ local minimums used for stretchline integration intial positions.}
+\label{f:ocean dataset hyperbolic shear lcs details stretchline}
 \end{figure}

### 2013-11-11 graphics/ocean_dataset/hyperbolic_shear_lcs_details_pdf.m

@@ -11,7 +11,7 @@ try
     cd(oldFolder)
     
     hFigure = 1;
-    filename = 'hyperbolic_shear_lcs_details_forward';
+    filename = 'hyperbolic_shear_lcs_details_strainline';
     savefig(hFigure,filename)
     delete_title(hFigure)
     filenamePDF = [filename,'.pdf'];
@@ -23,7 +23,7 @@ try
     % FIXME hFigure should not be hard-coded; should probably search for
     % matching title string
     hFigure = 12;
-    filename = 'hyperbolic_shear_lcs_details_backward';
+    filename = 'hyperbolic_shear_lcs_details_stretchline';
     savefig(hFigure,filename)
     delete_title(hFigure)
     filenamePDF = [filename,'.pdf'];

### 2013-11-11 graphics/ocean_dataset/hyperbolic_shear_lcs_pdf.m

@@ -11,7 +11,7 @@ try
     cd(oldFolder)
     
     hFigure = 1;
-    filename = 'hyperbolic_shear_lcs_forward';
+    filename = 'hyperbolic_shear_lcs_strainline';
     savefig(hFigure,filename)
     delete_title(hFigure)
     filenamePDF = [filename,'.pdf'];
@@ -21,7 +21,7 @@ try
     print_pdf(hFigure,filename)
     
     hFigure = 2;
-    filename = 'hyperbolic_shear_lcs_backward';
+    filename = 'hyperbolic_shear_lcs_stretchline';
     savefig(hFigure,filename)
     delete_title(hFigure)
     filenamePDF = [filename,'.pdf'];

### 2013-11-12 main.tex
Add graphics of Bickley jet FTLE at various resolutions

@@
 
+Figure~\ref{f:Bickley jet FTLE resolutions} shows the FTLE at various resolutions.
+Visual inspection of these figures indicates that a resolution of $500 \times 199$ is acceptable since increasing the resolution to $1000 \times 398$ does not change the appearance of the FTLE noticeably.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=.85\textwidth]{graphics/bickley_jet/ftle_100_39}
+\includegraphics[width=.85\textwidth]{graphics/bickley_jet/ftle_250_99}
+\includegraphics[width=.85\textwidth]{graphics/bickley_jet/ftle_500_199}
+\includegraphics[width=.85\textwidth]{graphics/bickley_jet/ftle_1000_398}
+\end{center}
+\caption{Bickley jet FTLE at various resolutions.}
+\label{f:Bickley jet FTLE resolutions}
+\end{figure}
+
 To calculate the Cauchy-Green strain tensor and its eigenvalues and eigenvectors, the main grid resolution is set to $500 \times 200$ and the auxiliary grid spacing is set to one percent of the main grid spacing (i.e.

### 2013-11-12 graphics/bickley_jet/ftle_resolutions.m

@@ -1,0 +1,43 @@
+% Plot FTLE at various resolutions
+
+function ftle_resolutions
+
+lcs_tool_root = fullfile('..','..','..','..','lcs_toolbox');
+demo = fullfile(lcs_tool_root,'demo','bickley_jet');
+
+oldFolder = cd(demo);
+addpath(fullfile('..','..'))
+
+resolutionX = [100,250,500,1000];
+
+%% Input parameters
+u = 62.66;
+lengthX = pi*earthRadius;
+lengthY = 1.77e6;
+epsilon = [.075,.4,.3];
+domain = [0,lengthX;[-1,1]*2.25*lengthY];
+timespan = [0,4*lengthX/u];
+
+perturbationCase = 3;
+lDerivative = @(t,x,~)derivative(t,x,false,u,lengthX,lengthY,epsilon,perturbationCase);
+incompressible = true;
+
+for m = 1:numel(resolutionX)
+   gridSpace = diff(domain(1,:))/(double(resolutionX(m))-1);
+   resolutionY = round(diff(domain(2,:))/gridSpace);
+   resolution = [resolutionX(m),resolutionY];
+   
+   cgEigenvalue = eig_cgStrain(lDerivative,domain,resolution,timespan,'incompressible',incompressible);
+   cgEigenvalue2 = reshape(cgEigenvalue(:,2),fliplr(resolution));
+   ftle_ = ftle(cgEigenvalue2,diff(timespan));
+
+   hAxes = setup_figure(domain);
+   title(hAxes,['Resolution: (',num2str(resolution(1)),',',num2str(resolution(2)),')'])
+   plot_ftle(hAxes,domain,resolution,ftle_);
+   colormap(hAxes,flipud(gray))
+   filename = fullfile(oldFolder,['ftle_',num2str(resolution(1)),'_',num2str(resolution(2))]);
+   print_pdf(gcf,filename)
+   close(gcf)
+end
+
+cd(oldFolder)

### 2013-11-13 main.tex
Update Bickley jet section

@@
 such as the polar night jet perturbed by a Rossby
-wave\parencite{beron-vera10:_invar_lagran,haller12:_geodes_theor_trans_barrier_two_dimen_flows}.
+wave\parencite{haller12:_geodes_theor_trans_barrier_two_dimen_flows,beron-vera10:_invar_lagran}.
 
-The model is defined by two stream functions, $\psi(x,y,t) = \psi_0(x,y) + \psi_1(x,y,t)$, where:
+The velocity is given by $\boldsymbol{v}(x,y,t) = (-\partial_y \psi, \partial_x \psi)$ and
 \begin{gather*}
+\psi(x,y,t) = \psi_0(x,y) + \psi_1(x,y,t),\\
 \psi_0(x,y) = c_3 y - U L_y \tanh\frac{y}{L_y} + \epsilon_3 U L_y \mathrm{sech}^2\frac{y}{L_y} \cos k_3 x,\\
@@
 \end{gather*}
-The forcing function is a solution of the Duffing oscillators with parameters in the chaotic regime, specifically
+Periodic boundary conditions are imposed in the $x$-direction.The forcing function we choose is a chaotic solution of the Duffing oscillators with parameters in the chaotic regime, specifically
 \begin{gather*}
@@
 \end{gather*}
-The velocity is given by:
-\[
-\boldsymbol{v}(x,y,t) = (-\partial_y \psi, \partial_x \psi).
-\]
 The parameter values used are: $U = 62.66$, $c_2 = 0.205 U$, $c_3 = 0.461 U$, $L_y = 1.77 \times 10^6$, $\epsilon_1 = 0.0075$, $\epsilon_2 = 0.04$, $\epsilon_3 = 0.3$, $L_x = 6.371 \times 10^6 \pi$, $k_n = 2 n \pi/L_x$, $\sigma_1 = 0.5 k_2 (c_2 - c_3)$, $\sigma_2 = 2 \sigma_1$.
-The domain of the flow is set to $x \in [0,L_x]$, $y \in [-4 \times 10^6, 4 \times 10^6]$ and the timespan is set to $4 L_x/U$.
+The domain of the flow is set to $x \in [0,L_x]$, $y \in [-2.25 L_y, 2.25 L_y]$ and the timespan is $t \in [0,4 L_x/U]$.
+We skip listing the file that defines the velocity equations; in regards to input and output variables, it is similar to the double gyre function (Listing~\ref{l:double gyre derivative}.) The LCS Tool script that specifies the parameters we are using is given in Listing~\ref{l:Bickley jet input}.
+Note the use of the \texttt{periodicBc} variable to specify periodic boundary conditions.
+
+\begin{lstlisting}[caption={Bickley jet input file: \texttt{demo/bickley\_jet/hyperbolic\_shear\_lcs.m}.},label=l:Bickley jet input]
+%% Input parameters
+u = 62.66;
+lengthX = pi*earthRadius;
+lengthY = 1.77e6;
+epsilon = [.075,.4,.3];
+domain = [0,lengthX;[-1,1]*2.25*lengthY];
+
+% Make x and y grid spacing as equal as possible
+resolutionX = 500;
+gridSpace = diff(domain(1,:))/(double(resolutionX)-1);
+resolutionY = round(diff(domain(2,:))/gridSpace);
+resolution = [resolutionX,resolutionY];
+
+timespan = [0,4*lengthX/u];
+
+%% Velocity definition
+perturbationCase = 3;
+derivative = @(t,x,~)derivative(t,x,false,u,lengthX,lengthY,epsilon,perturbationCase);
+incompressible = true;
+periodicBc = [true,false];
+\end{lstlisting}
 
@@
 
-To calculate the Cauchy-Green strain tensor and its eigenvalues and eigenvectors, the main grid resolution is set to $500 \times 200$ and the auxiliary grid spacing is set to one percent of the main grid spacing (i.e.
-400.) The main grid is used to calculate eigenvalues and flow incompressibility is imposed.
-Periodic boundary conditions are imposed in the horizontal direction.
+To calculate the Cauchy-Green eigenvalues and eigenvectors we therefore set the grid resolution to $500 \times 199$.
+We also specify the flow is incompressible.
+Commands to obtain lambda-lines are given in listing~\ref{l:Bickley jet lambda-line commands}.
+The primary input needed are the positions of the Poincare sections which we determine
 
-To obtain shear LCS Poincare sections are specified. The maximum orbit length is set to twice the circumference circles whose radius would be the Poincare sections.
+\begin{lstlisting}[caption={LCS Tool commands Bickley jet lambda-lines, subset of \texttt{demo/bickley\_jet/hyperbolic\_shear\_lcs.m}.},label=l:Bickley jet lambda-line commands]
+lambda = 1;
+lambdaLineOdeSolverOptions = odeset('relTol',1e-4);
 
-To obtain hyperbolic LCS, the local maximisation distance is set to eight times the main grid spacing and the maximum strainline length is set to $1 \times 10^8$.
+poincareSection = struct('endPosition',{},'numPoints',{},'orbitMaxLength',{});
+poincareSection(1).endPosition = [6.5,-1.4;5.2,-1.4]*1e6;
+poincareSection(2).endPosition = [1.35e7,-1.4e6;1.5e7,-.5e6];
+poincareSection(3).endPosition = [3.25,1.5;1.4,2.6]*1e6;
+poincareSection(4).endPosition = [1e7,1.5e6;8e6,2.6e6];
+poincareSection(5).endPosition = [1.65e7,1.5e6;1.5e7,2.6e6];
+
+% Number of orbit seed points along each Poincare section
+[poincareSection.numPoints] = deal(80);
+
+% Set maximum orbit length to twice the expected circumference
+nPoincareSection = numel(poincareSection);
+for i = 1:nPoincareSection
+    rOrbit = hypot(diff(poincareSection(i).endPosition(:,1)),diff(poincareSection(i).endPosition(:,2)));
+    poincareSection(i).orbitMaxLength = 2*(2*pi*rOrbit);
+end
+
+[shearline.etaPos,shearline.etaNeg] = lambda_line(cgEigenvector,cgEigenvalue,lambda);
+closedLambdaLine = poincare_closed_orbit_multi(domain,resolution,shearline,poincareSection,'odeSolverOptions',lambdaLineOdeSolverOptions,'periodicBc',periodicBc);
+\end{lstlisting}
+
+To obtain strainline hyperbolic LCS, we must set a local maximisation distance and a maximum LCS length. To demonstrate the
 

### 2013-11-13 graphics/bickley_jet/ftle_resolutions.m

@@ -5,8 +5,7 @@ function ftle_resolutions
 lcs_tool_root = fullfile('..','..','..','..','lcs_toolbox');
 demo = fullfile(lcs_tool_root,'demo','bickley_jet');
 
-oldFolder = cd(demo);
-addpath(fullfile('..','..'))
+addpath(lcs_tool_root,demo)
 
 resolutionX = [100,250,500,1000];
 
@@ -32,12 +31,10 @@ for m = 1:numel(resolutionX)
    ftle_ = ftle(cgEigenvalue2,diff(timespan));
 
    hAxes = setup_figure(domain);
-   title(hAxes,['Resolution: (',num2str(resolution(1)),',',num2str(resolution(2)),')'])
+   title(hAxes,['Resolution: ',num2str(resolution(1)),'\times',num2str(resolution(2))])
    plot_ftle(hAxes,domain,resolution,ftle_);
    colormap(hAxes,flipud(gray))
-   filename = fullfile(oldFolder,['ftle_',num2str(resolution(1)),'_',num2str(resolution(2))]);
+   filename = ['ftle_',num2str(resolution(1)),'_',num2str(resolution(2))];
    print_pdf(gcf,filename)
    close(gcf)
 end
-
-cd(oldFolder)

### 2013-11-13 graphics/bickley_jet/hyperbolic_shear_lcs_details_pdf.m

@@ -3,38 +3,21 @@ function hyperbolic_shear_lcs_details_pdf
 lcs_tool_root = fullfile('..','..','..','..','lcs_toolbox');
 demo = fullfile(lcs_tool_root,'demo','bickley_jet');
 
-oldFolder = cd(demo);
+addpath(lcs_tool_root,demo)
 
-try
-    addpath(fullfile('..','..'))
-    hyperbolic_shear_lcs_details
-    cd(oldFolder)
+hyperbolic_shear_lcs_details
     
-    hFigure = 1;
-    filename = 'hyperbolic_shear_lcs_details_strainline';
-    savefig(hFigure,filename)
-    delete_title(hFigure)
-    filenamePDF = [filename,'.pdf'];
-    if exist(filenamePDF,'file')
-        delete(filenamePDF)
-    end
-    print_pdf(hFigure,filename)
+hFigure = 1;
+delete_title(hFigure)
+filename = 'hyperbolic_shear_lcs_details_strainline';
+print_pdf(hFigure,filename)
     
-    % FIXME hFigure should not be hard-coded; should probably search for title
-    % string
-    hFigure = 12;
-    filename = 'hyperbolic_shear_lcs_details_stretchline';
-    savefig(hFigure,filename)
-    delete_title(hFigure)
-    filenamePDF = [filename,'.pdf'];
-    if exist(filenamePDF,'file')
-        delete(filenamePDF)
-    end
-    print_pdf(hFigure,filename)
-catch err
-    cd(oldFolder)
-    rethrow(err)
-end
+% FIXME hFigure number should not be hard-coded; should probably search for
+% title string
+hFigure = 12;
+delete_title(hFigure)
+filename = 'hyperbolic_shear_lcs_details_stretchline';
+print_pdf(hFigure,filename)
 
 function delete_title(hFigure)
 

### 2013-12-09 main.tex
Update double gyre example

@@
 \begin{center}
-\includegraphics[width=\textwidth]{graphics/double_gyre/poincare_return_map}
+%\includegraphics[width=\textwidth]{graphics/double_gyre/poincare_return_map}
 \end{center}
@@
 
-The parameter values used are: $A = 0.1$, $\epsilon = 0.1$, $\omega = \pi/5$.
-The flow timespan is $t \in [0,20]$.
-The physical domain of the double gyre flow is $x \in [0,2]$, $y \in [0,1]$, however to help capture behaviour near the fixed points of the unperturbed system at $(x,y) = (1,0)$ and $(x,y) = (1,1)$ in computations, the domain is enlarged to $x \in [-0.1,2.1]$, $y \in [-0.05,1.05]$.
-To calculate the Cauchy-Green strain tensor and its eigenvalues and eigenvectors, the main grid resolution is set to $551 \times 276$.
-The main grid is used to calculate eigenvalues and flow incompressibility is imposed.
-The LCS Tool code that corresponds to these parameters is in Listing~\ref{l:double gyre input}.
-The variable named flow is a MATLAB structure that contains fields corresponding to input parameters; using structures allows passing input arguments to LCS Tool functions with a compact notation.
-
-\begin{lstlisting}[firstnumber=4,caption={Double gyre input parameter definitions; subset of the file: \texttt{demo/double\_gyre/hyperbolic\_shear\_lcs.m}.},label=l:double gyre input]
-epsilon = .1;
-amplitude = .1;
-omega = pi/5;
-
-flow.imposeIncompressibility = true;
-flow.derivative = @(t,x,useEoV)derivative(t,x,useEoV,epsilon,amplitude,omega);
-flow.domain = [-.1,2.1;-.05,1.05];
-flow.timespan = [0,20];
-flow.resolution = [551,276];
-\end{lstlisting}
-
-To obtain shear LCSs, Poincare section positions are specified.
-Figure~\ref{f:Poincare sections with FTLE} illustrates how Poincare sections are placed in regions where the FTLE ridges have a circular pattern.
-The set of LCS Tool commands used to detect shear LCSs is given in Listing~\ref{l:double gyre shear LCS commands}.
-First the function \texttt{eig\_cgStrain} calculates the Cauchy-Green strain eigenvalues and eigenvectors.
-The most important function is \texttt{poincare\_closed\_orbit\_multi} that takes as input a list of Poincare sections, calculates Poincare return maps for the $\eta_+$ and $\eta_-$ vector fields and returns positions of all closed orbits.
+In what follows, the parameter values are: $A = 0.1$, $\epsilon = 0.1$, $\omega = \pi/5$.
+The flow timespan is $t \in [0,5]$ and the domain is $x \in [0,2]$, $y \in [0,1]$.
+In Figure~\ref{f:double gyre FTLE}, we show the FTLE at a resolution of $500\times250$.
 
@@
 \begin{center}
-\includegraphics[width=\textwidth]{graphics/double_gyre/poincare_sections}
+\includegraphics[width=\textwidth]{graphics/double_gyre/ftle}
 \end{center}
-\caption{Poincare sections positions with FTLE.}
-\label{f:Poincare sections with FTLE}
+\caption{Double gyre FTLE.}
+\label{f:double gyre FTLE}
 \end{figure}
 
-\begin{lstlisting}[caption={Double gyre shear LCS detection commands. Subset of \texttt{hyperbolic\_shear\_lcs.m}.},label=l:double gyre shear LCS commands,firstnumber=24]
-% Compute Cauchy-Green strain eigenvalues and eigenvectors
-[flow.cgEigenvalue,flow.cgEigenvector] = eig_cgStrain(flow);
-% FIXME Should use m-by-n or (m*n)-by-2 array forms throughout LCS Tool
-cgEigenvalue = reshape(flow.cgEigenvalue,[fliplr(flow.resolution),2]);
-cgEigenvector = reshape(flow.cgEigenvector,[fliplr(flow.resolution),4]);
+By examining the FTLE, we can position Poincare sections to detect lambda-line LCSs.
+An LCS Tool script to perform this operation is given in Listing~\ref{l:double gyre lambda-line LCS} where two Poincare sections are defined.
+In the script we have set error tolerances for integration of the Cauchy-Green strain tensor and lambda lines.
+These numerical parameters help achieve convergence.
 
-% Define Poincare sections; first point in center of elliptic region and
-% second point outside elliptic region
-poincareSection = struct('endPosition',{},'numPoints',{},'orbitMaxLength',{});
+\lstinputlisting[caption={Double gyre script for lambda-line LCSs},label=l:double gyre lambda-line LCS]{listings/double_gyre/lambda_lcs.m}
 
-poincareSection(1).endPosition = [.5,.6;.35,.5];
-poincareSection(2).endPosition = [1.5,.4;1.7,.5];
+In Figure~\ref{fig:double_gyre_lambda_lcs_convergence} the resolution of the  Cauchy-Green strain tensor is varied from $500 \times 250$ to $1000 \times 500$.
+The location of the outermost closed lambda-line changes little, demonstrating convergence.
 
-% Number of orbit seed points along each Poincare section
-[poincareSection.numPoints] = deal(80);
+\begin{figure}[hbt]
+  \centering
+  \includegraphics[width=.8\textwidth]{graphics/double_gyre/lambda_lcs_convergence_500}
+\includegraphics[width=.8\textwidth]{graphics/double_gyre/lambda_lcs_convergence_750}
+\includegraphics[width=.8\textwidth]{graphics/double_gyre/lambda_lcs_convergence_1000}
+  \caption{Double gyre closed lambda-line at varying resolutions}
+  \label{fig:double_gyre_lambda_lcs_convergence}
+\end{figure}
 
-% Set maximum orbit length to twice the expected circumference
-nPoincareSection = numel(poincareSection);
-for i = 1:nPoincareSection
-    rOrbit = hypot(diff(poincareSection(i).endPosition(:,1)),diff(poincareSection(i).endPosition(:,2)));
-    poincareSection(i).orbitMaxLength = 2*(2*pi*rOrbit);
-end
-
-[shearline.etaPos,shearline.etaNeg] = lagrangian_shear(flow.cgEigenvector,flow.cgEigenvalue);
-
-% Compute closed shearlines
-closedOrbits = poincare_closed_orbit_multi(flow,shearline,poincareSection);
-\end{lstlisting}
-
-To obtain hyperbolic LCS, the local maximisation distance is set to twice the main grid spacing.
-To limit integration time, the maximum strainline length is set to 20.
-LCS Tool commands are shown in Listing~\ref{l:double gyre hyperbolic LCS commands}.
-The function \texttt{seed\_curves\_from\_lambda\_max} calculates strainline positions.
-The function \texttt{remove\_strain\_in\_shear} then removes the segments of strainlines within the outermost shearline of each Poincare section.
-
-\begin{lstlisting}[caption={Double gyre hyperbolic LCS detection commands. Subset of \texttt{hyperbolic\_shear\_lcs.m}.},label=l:double gyre hyperbolic LCS commands,firstnumber=60]
-strainlineMaxLength = 20;
-gridSpace = diff(flow.domain(1,:))/(double(flow.resolution(1))-1);
-localMaxDistance = 2*gridSpace;
-
-% Compute strainlines
-strainlinePosition = seed_curves_from_lambda_max(localMaxDistance,strainlineMaxLength,cgEigenvalue(:,:,2),cgEigenvector(:,:,1:2),flow.domain);
-
-for i = 1:nPoincareSection
-    % Remove strainlines inside of ellitpic regions
-    strainlinePosition = remove_strain_in_shear(strainlinePosition,closedOrbits{i}{1}{end});
-    strainlinePosition = remove_strain_in_shear(strainlinePosition,closedOrbits{i}{2}{end});
-end
-\end{lstlisting}
-
-The forward time LCS obtained are shown in Figure~\ref{f:double gyre hyperbolic shear lcs strainline} and the backward time LCS (i.e.
-with the flow timespan set to $t \in [20,0]$) are shown in Figure~\ref{f:double gyre hyperbolic shear lcs stretchline}.
-In Figures~\ref{f:double gyre hyperbolic shear lcs details strainline} and~\ref{f:double gyre hyperbolic shear lcs details stretchline} the same LCS are shown with details: the finite-time Lyapunov exponent, closed shearlines inside the outermost shearlines and $\lambda_2$ local maximums used for strainline integration initial positions.
+We continue the example by adding the calculation of hyperbolic strainline and stretchline LCSs.
+Figures~\ref{fig:double_gyre_lambda_strain_lcs} and~\ref{fig:double_gyre_lambda_stretch_lcs} show the results and listing~\ref{l:double_gyre_lambda_strain_stretch_lcs} gives the LCS Tool commands.
+Note that the local maximisation distance is usually set larger for stretchlines than for strainlines.
+This is due to the fact stretchlines do not follow $\lambda_1$ minimum ridges.
 
 \begin{figure}
-\begin{center}
-\includegraphics[width=\textwidth]{graphics/double_gyre/hyperbolic_shear_lcs_strainline}
-\end{center}
-\caption{Double gyre lambda-line (green) and strainline (red) LCSs.}
-\label{f:double gyre hyperbolic shear lcs strainline}
+  \centering
+  \includegraphics[width=\textwidth]{graphics/double_gyre/lambda_strain_lcs}
+  \caption{Double gyre lambda-line and strainline LCSs}
+  \label{fig:double_gyre_lambda_strain_lcs}
 \end{figure}
@@
 \begin{figure}
-\begin{center}
-\includegraphics[width=\textwidth]{graphics/double_gyre/hyperbolic_shear_lcs_stretchline}
-\end{center}
-\caption{Double gyre lambda-line (green) and stretchline (blue) LCSs.}
-\label{f:double gyre hyperbolic shear lcs stretchline}
+  \centering
+  \includegraphics[width=\textwidth]{graphics/double_gyre/lambda_stretch_lcs}
+  \caption{Double gyre lambda-line and stretchline LCSs}
+  \label{fig:double_gyre_lambda_stretch_lcs}
 \end{figure}
 
-\begin{figure}
-\begin{center}
-\includegraphics[width=\textwidth]{graphics/double_gyre/hyperbolic_shear_lcs_details_strainline}
-\end{center}
-\caption{Same LCS as in Figure~\ref{f:double gyre hyperbolic shear lcs strainline}, but with additional details.
-Background gray-scale is the finite-time Lyapunov exponent.
-Dashed green lines are Poincare sections.
-Green curves are closed lambda-lines with thick green curves being outermost closed lambda-lines, i.e.
-lambda-line LCSs.
-White dots indicate $\lambda_2$ local maximums used for strainline integration initial positions.}
-\label{f:double gyre hyperbolic shear lcs details strainline}
-\end{figure}
+\lstinputlisting[caption={LCS Tool commands for strainline and stretchline LCSs},label=l:double_gyre_lambda_strain_stretch_lcs]{listings/double_gyre/lambda_strain_stretch_lcs.m}
 
-\begin{figure}
-\begin{center}
-\includegraphics[width=\textwidth]{graphics/double_gyre/hyperbolic_shear_lcs_details_stretchline}
-\end{center}
-\caption{Same LCS as in Figure~\ref{f:double gyre hyperbolic shear lcs stretchline}, but with additional details.
-Background gray-scale is the finite-time Lyapunov exponent.
-Dashed green lines are Poincare sections.
-Green curves are closed lambda-lines and thick green curves are outermost closed lambda-lines, i.e.
-lambda-line LCSs.
-White dots indicate $\lambda_1$ local minimums used for stretchline integration intial positions.}
-\label{f:double gyre hyperbolic shear lcs details stretchline}
-\end{figure}
-
+\clearpage
 \subsection{Bickley jet}
@@
 \end{gather*}
-Periodic boundary conditions are imposed in the $x$-direction.The forcing function we choose is a chaotic solution of the Duffing oscillators with parameters in the chaotic regime, specifically
+The forcing function we choose is a chaotic solution of the Duffing oscillator, specifically
 \begin{gather*}
@@
 \end{gather*}
-The parameter values used are: $U = 62.66$, $c_2 = 0.205 U$, $c_3 = 0.461 U$, $L_y = 1.77 \times 10^6$, $\epsilon_1 = 0.0075$, $\epsilon_2 = 0.04$, $\epsilon_3 = 0.3$, $L_x = 6.371 \times 10^6 \pi$, $k_n = 2 n \pi/L_x$, $\sigma_1 = 0.5 k_2 (c_2 - c_3)$, $\sigma_2 = 2 \sigma_1$.
-The domain of the flow is set to $x \in [0,L_x]$, $y \in [-2.25 L_y, 2.25 L_y]$ and the timespan is $t \in [0,4 L_x/U]$.
-We skip listing the file that defines the velocity equations; in regards to input and output variables, it is similar to the double gyre function (Listing~\ref{l:double gyre derivative}.) The LCS Tool script that specifies the parameters we are using is given in Listing~\ref{l:Bickley jet input}.
-Note the use of the \texttt{periodicBc} variable to specify periodic boundary conditions.
+The parameters used are: $U = 62.66$, $c_2 = 0.205 U$, $c_3 = 0.461 U$, $L_y = 1.77 \times 10^6$, $\epsilon_1 = 0.0075$, $\epsilon_2 = 0.04$, $\epsilon_3 = 0.3$, $L_x = 6.371 \times 10^6 \pi$, $k_n = 2 n \pi/L_x$, $\sigma_1 = 0.5 k_2 (c_2 - c_3)$, $\sigma_2 = 2 \sigma_1$.
+
+Figure~\ref{f:Bickley jet FTLE} shows the overall appearance of the Bickley jet.
+To get a sense of the numerical parameters required to detect LCSs in this flow, we analyse lambda-line closed orbits of one vortex for a short timespan.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=.85\textwidth]{graphics/bickley_jet/ftle_overview}
+\end{center}
+\caption{Bickley jet FTLE}
+\label{f:Bickley jet FTLE}
+\end{figure}
+
+% $[0,2L_x/U]$.
+% Loose closed orbits as increase lambda-line integration accuracy
+
+First, we set a Poincare section so that the first point is near the centre of the region where we seek a vortex and the second point is outside this region, or at least a reasonable distance from the outermost closed lambda-line we find.
+
+% We analyse lambda-line LCSs in a subdomain of this flow.
+% 1. FTLE field
+% 2. Poincare sections
+% 3. Convergence by varying resolution and other parameters.
+
+Start with numerical parameters that give answers quickly and increase accuracy until observe convergence visually.
 
@@
 \end{lstlisting}
-
-Figure~\ref{f:Bickley jet FTLE resolutions} shows the FTLE at various resolutions.
-Visual inspection of these figures indicates that a resolution of $500 \times 199$ is acceptable since increasing the resolution to $1000 \times 398$ does not change the appearance of the FTLE noticeably.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=.85\textwidth]{graphics/bickley_jet/ftle_100_39}
-\includegraphics[width=.85\textwidth]{graphics/bickley_jet/ftle_250_99}
-\includegraphics[width=.85\textwidth]{graphics/bickley_jet/ftle_500_199}
-\includegraphics[width=.85\textwidth]{graphics/bickley_jet/ftle_1000_398}
-\end{center}
-\caption{Bickley jet FTLE at various resolutions.}
-\label{f:Bickley jet FTLE resolutions}
-\end{figure}
 
@@
 
+For the last example, we choose a measured geophysical flow where, in contrast to the previous examples, the two dimensional incompressible velocity field is presented as a dataset with discrete temporal and spatial resolution.
+Our region of interest is a small domain in the South Atlantic Ocean, where exceptionally coherent eddies, Agulhas rings, were recently found using methods provided by LCS Tool\parencite{haller13:_coher_lagran}.
+The oceanic velocity field is obtained from altimetric satelite measurements of the sea surface height under the assumption that the flow is geostrophic.
+Under this assumption the sea surface height $\eta$ serves as a streamfunction and in a longitude-latitude $(\varphi,\theta)$ coordinate system the evolution of a fluid particle is given by
+\begin{eqnarray}
+\dot{\varphi}(\varphi,\theta,t) = -\frac{g}{R^2 f(\theta) \cos\theta}\partial_{\theta}\eta(\varphi,\theta,t)\label{e:dtlon}\\
+\dot{\theta}(\varphi,\theta,t) = +\frac{g}{R^2 f(\theta) \cos\theta}\partial_{\varphi}\eta(\varphi,\theta,t)
+\label{e:dtlat}
+\end{eqnarray}
+where $g$ is the gravitational acceleration, $R$ is the mean radius of Earth, and $f(\theta):=2\Omega\sin\theta$ is the Coriolis parameter with $\Omega$ being the Earth's mean angular velocity.
+The data is given at a spatial resolution of $1/4\degree$ and temporal resolution of $7\,\mathrm{days}$.
+Due to the discrete data, defining the right hand side of Eq.\,(\ref{e:dtlon}) and (\ref{e:dtlat}) involves interpolation in space and time with a spline interpolation scheme.
+An interpolant is generated first, then the function \texttt{flowdata\_derivative} evaluates the interpolant at the needed coordinates.
+The interpolation time was chosen to be $T=90\,\mathrm{days}$, a multiple of the turn over time of the mesoscale eddies in the domain.
+The resolution of the grid of initial conditions is set to $400\times400$ which corresponds to a resolution of roughly $0.015^\circ$ and gives good results.
+With this choice the resolution of the tracer grid is of the order of $15$ times higher than the resolution of the velocity field.
+Figures~\ref{f:ocean dataset hyperbolic shear lcs strainline} and \ref{f:ocean dataset hyperbolic shear lcs stretchline} show elliptic and hyperbolic LCS on 22 November 2006, i.e., the same time instant shown in~\textcite{haller13:_coher_lagran,beron-vera13:_objec_agulh}.
+The analysis with LCS Tool reveals a large coherent eddy bounded by a closed lambda-line at $(3,-32)$ and four further smaller coherent eddy cores.
+
 \begin{itemize}
-\item auxiliaryGridRelativeDelta = 0.1
-\item eigenvalueFromMainGrid = false;
-\item Cauchy-Green resolution must be higher than dataset
+\item Cauchy-Green resolution higher than dataset
 \end{itemize}
@@
 
+\subsection*{Acknowledgements}
+
+The altimeter products used in this work are produced by SSALTO/DUACS and distributed by AVISO, with support from CNES \url{http://www.aviso.oceanobs.com/duacs}.
+
 \printbibliography
@@
 % LocalWords:  stretchline stretchlines incompressible gyre eigenvector eigenvectors incompressibility dataset datasets LCS strainline shearline Lyapunov Bickley Rossby timespan variational adaptively vortices parametrise advected
+%  LocalWords:  strainlines

### 2013-12-09 graphics/double_gyre/ftle_.m

@@ -1,0 +1,38 @@
+function ftle_
+
+%% Input parameters
+epsilon = .1;
+amplitude = .1;
+omega = pi/5;
+domain = [0,2;0,1];
+resolutionX = 500;
+timespan = [0,5];
+
+%% Velocity definition
+lDerivative = @(t,x,~)derivative(t,x,false,epsilon,amplitude,omega);
+incompressible = true;
+
+%% LCS parameters
+cgStrainOdeSolverOptions = odeset('relTol',1e-5);
+
+% Make x and y grid spacing as equal as possible
+gridSpace = diff(domain(1,:))/(double(resolutionX)-1);
+resolutionY = round(diff(domain(2,:))/gridSpace);
+resolution = [resolutionX,resolutionY];
+    
+hAxes = setup_figure(domain);
+title(hAxes,['Resolution: ',num2str(resolution(1)),'\times',num2str(resolution(2))])
+    
+%% Cauchy-Green strain eigenvalues and eigenvectors
+s = warning('off','eig_cgStrain:unequalDelta');
+cgEigenvalue = eig_cgStrain(lDerivative,domain,resolution,timespan,'incompressible',incompressible,'odeSolverOptions',cgStrainOdeSolverOptions);
+warning(s)
+
+% Plot finite-time Lyapunov exponent
+cgEigenvalue2 = reshape(cgEigenvalue(:,2),fliplr(resolution));
+ftle_ = ftle(cgEigenvalue2,diff(timespan));
+plot_ftle(hAxes,domain,resolution,ftle_);
+colormap(hAxes,flipud(gray))
+delete(get(hAxes,'title'))
+hFigure = get(hAxes,'parent');
+print_pdf(hFigure,'ftle.pdf')

### 2013-12-09 graphics/double_gyre/hyperbolic_shear_lcs_details_pdf.m

@@ -1,43 +1,0 @@
-function hyperbolic_shear_lcs_details_pdf
-
-lcs_tool_root = fullfile('..','..','..','..','lcs_toolbox');
-demo = fullfile(lcs_tool_root,'demo','double_gyre');
-
-oldFolder = cd(demo);
-
-try
-    addpath(fullfile('..','..'))
-    hyperbolic_shear_lcs_details
-    cd(oldFolder)
-    
-    hFigure = 1;
-    filename = 'hyperbolic_shear_lcs_details_strainline';
-    savefig(hFigure,filename)
-    delete_title(hFigure)
-    filenamePDF = [filename,'.pdf'];
-    if exist(filenamePDF,'file')
-        delete(filenamePDF)
-    end
-    print_pdf(hFigure,filename)
-    
-    % FIXME hFigure should not be hard-coded; should probably search for title
-    % string
-    hFigure = 6;
-    filename = 'hyperbolic_shear_lcs_details_stretchline';
-    savefig(hFigure,filename)
-    delete_title(hFigure)
-    filenamePDF = [filename,'.pdf'];
-    if exist(filenamePDF,'file')
-        delete(filenamePDF)
-    end
-    print_pdf(hFigure,filename)
-catch err
-    cd(oldFolder)
-    rethrow(err)
-end
-
-function delete_title(hFigure)
-
-hAxes = findobj(hFigure,'type','axes','-not','Tag','Colorbar');
-hTitle = get(hAxes,'title');
-delete(hTitle)

### 2013-12-09 graphics/double_gyre/hyperbolic_shear_lcs_pdf.m

@@ -1,41 +1,0 @@
-function hyperbolic_shear_lcs_pdf
-
-lcs_tool_root = fullfile('..','..','..','..','lcs_toolbox');
-demo = fullfile(lcs_tool_root,'demo','double_gyre');
-
-oldFolder = cd(demo);
-
-try
-    addpath(fullfile('..','..'))
-    hyperbolic_shear_lcs
-    cd(oldFolder)
-    
-    hFigure = 1;
-    filename = 'hyperbolic_shear_lcs_strainline';
-    savefig(hFigure,filename)
-    delete_title(hFigure)
-    filenamePDF = [filename,'.pdf'];
-    if exist(filenamePDF,'file')
-        delete(filenamePDF)
-    end
-    print_pdf(hFigure,filename)
-    
-    hFigure = 2;
-    filename = 'hyperbolic_shear_lcs_stretchline';
-    savefig(hFigure,filename)
-    delete_title(hFigure)
-    filenamePDF = [filename,'.pdf'];
-    if exist(filenamePDF,'file')
-        delete(filenamePDF)
-    end
-    print_pdf(hFigure,filename)
-catch err
-    cd(oldFolder)
-    rethrow(err)
-end
-
-function delete_title(hFigure)
-
-hAxes = get(hFigure,'children');
-hTitle = get(hAxes,'title');
-delete(hTitle)

### 2013-12-09 graphics/double_gyre/lambda_lcs_convergence.m

@@ -1,0 +1,79 @@
+function lambda_lcs_convergence
+
+%% Input parameters
+epsilon = .1;
+amplitude = .1;
+omega = pi/5;
+domain = [0,2;0,1];
+resolutionX = 500:250:1000;
+timespan = [0,5];
+
+%% Velocity definition
+lDerivative = @(t,x,~)derivative(t,x,false,epsilon,amplitude,omega);
+incompressible = true;
+
+%% LCS parameters
+cgStrainOdeSolverOptions = odeset('relTol',1e-5);
+
+% Lambda-lines
+lambda = 1;
+lambdaLineOdeSolverOptions = odeset('relTol',1e-6);
+poincareSection = struct('endPosition',{},'numPoints',{},'orbitMaxLength',{});
+poincareSection(1).endPosition = [.55,.55;.2,.5];
+poincareSection(2).endPosition = [1.53,.45;1.9,.5];
+
+% Number of orbit seed points along each Poincare section
+[poincareSection.numPoints] = deal(100);
+
+% Set maximum orbit length to twice the expected circumference
+nPoincareSection = numel(poincareSection);
+for i = 1:nPoincareSection
+    rOrbit = hypot(diff(poincareSection(i).endPosition(:,1)),diff(poincareSection(i).endPosition(:,2)));
+    poincareSection(i).orbitMaxLength = 2*(2*pi*rOrbit);
+end
+
+% Graphics properties
+lambdaLineColor = [0,.6,0];
+
+for m = 1:numel(resolutionX)
+    % Make x and y grid spacing as equal as possible
+    lResolutionX = resolutionX(m);
+    gridSpace = diff(domain(1,:))/(double(lResolutionX)-1);
+    resolutionY = round(diff(domain(2,:))/gridSpace);
+    resolution = [lResolutionX,resolutionY];
+    disp(['Resolution: ',num2str(resolution)])
+    
+    hAxes = setup_figure(domain);
+    title(hAxes,['Resolution: ',num2str(resolution(1)),'\times',num2str(resolution(2))])
+    
+    %% Cauchy-Green strain eigenvalues and eigenvectors
+    s = warning('off','eig_cgStrain:unequalDelta');
[…]

### 2013-12-09 graphics/double_gyre/lambda_strain_lcs.m

@@ -1,0 +1,72 @@
+function lambda_strain_lcs
+
+%% Input parameters
+epsilon = .1;
+amplitude = .1;
+omega = pi/5;
+domain = [0,2;0,1];
+resolution = [750,375];
+timespan = [0,5];
+
+%% Velocity definition
+lDerivative = @(t,x,~)derivative(t,x,false,epsilon,amplitude,omega);
+incompressible = true;
+
+%% LCS parameters
+cgStrainOdeSolverOptions = odeset('relTol',1e-5);
+
+% Lambda-lines
+poincareSection = struct('endPosition',{},'numPoints',{},'orbitMaxLength',{});
+poincareSection(1).endPosition = [.55,.55;.2,.5];
+poincareSection(2).endPosition = [1.53,.45;1.9,.5];
+[poincareSection.numPoints] = deal(100);
+nPoincareSection = numel(poincareSection);
+for i = 1:nPoincareSection
+    rOrbit = hypot(diff(poincareSection(i).endPosition(:,1)),diff(poincareSection(i).endPosition(:,2)));
+    poincareSection(i).orbitMaxLength = 2*(2*pi*rOrbit);
+end
+lambda = 1;
+lambdaLineOdeSolverOptions = odeset('relTol',1e-6);
+
+% Strainlines
+strainlineMaxLength = 20;
+gridSpace = diff(domain(1,:))/(double(resolution(1))-1);
+strainlineLocalMaxDistance = 2*gridSpace;
+
+% Graphics properties
+strainlineColor = 'r';
+lambdaLineColor = [0,.6,0];
+
+hAxes = setup_figure(domain);
+hFigure = get(hAxes,'parent');
+
+%% Cauchy-Green strain eigenvalues and eigenvectors
+s = warning('off','eig_cgStrain:unequalDelta');
+[cgEigenvector,cgEigenvalue] = eig_cgStrain(lDerivative,domain,resolution,timespan,'incompressible',incompressible,'odeSolverOptions',cgStrainOdeSolverOptions);
+warning(s)
+
+%% Lambda-line LCSs
+[shearline.etaPos,shearline.etaNeg] = lambda_line(cgEigenvector,cgEigenvalue,lambda);
+s = warning('off','integrate_line:isDiscontinuousLargeAngle');
[…]

### 2013-12-09 graphics/double_gyre/lambda_stretch_lcs.m

@@ -1,0 +1,109 @@
+function lambda_stretch_lcs
+
+%% Input parameters
+epsilon = .1;
+amplitude = .1;
+omega = pi/5;
+domain = [0,2;0,1];
+resolution = [750,375];
+timespan = [0,5];
+
+%% Velocity definition
+lDerivative = @(t,x,~)derivative(t,x,false,epsilon,amplitude,omega);
+incompressible = true;
+
+%% LCS parameters
+cgStrainOdeSolverOptions = odeset('relTol',1e-5);
+
+% Lambda-lines
+poincareSection = struct('endPosition',{},'numPoints',{},'orbitMaxLength',{});
+poincareSection(1).endPosition = [.55,.55;.2,.5];
+poincareSection(2).endPosition = [1.53,.45;1.9,.5];
+[poincareSection.numPoints] = deal(100);
+nPoincareSection = numel(poincareSection);
+for i = 1:nPoincareSection
+    rOrbit = hypot(diff(poincareSection(i).endPosition(:,1)),diff(poincareSection(i).endPosition(:,2)));
+    poincareSection(i).orbitMaxLength = 2*(2*pi*rOrbit);
+end
+lambda = 1;
+lambdaLineOdeSolverOptions = odeset('relTol',1e-6);
+
+% Stretchlines
+stretchlineMaxLength = 20;
+gridSpace = diff(domain(1,:))/(double(resolution(1))-1);
+stretchlineLocalMaxDistance = 10*gridSpace;
+
+% Graphics properties
+stretchlineColor = 'b';
+lambdaLineColor = [0,.6,0];
+lcsInitialPositionMarkerSize = 2;
+
+hAxes = setup_figure(domain);
+hFigure = get(hAxes,'parent');
+
+%% Cauchy-Green strain eigenvalues and eigenvectors
+s = warning('off','eig_cgStrain:unequalDelta');
+[cgEigenvector,cgEigenvalue] = eig_cgStrain(lDerivative,domain,resolution,timespan,'incompressible',incompressible,'odeSolverOptions',cgStrainOdeSolverOptions);
+warning(s)
+
+% Plot finite-time Lyapunov exponent
+cgEigenvalue2 = reshape(cgEigenvalue(:,2),fliplr(resolution));
+ftle_ = ftle(cgEigenvalue2,diff(timespan));
+plot_ftle(hAxes,domain,resolution,ftle_);
[…]

### 2013-12-09 graphics/double_gyre/startup.m

@@ -1,0 +1,4 @@
+lcs_tool_root = fullfile('..','..','lcs_tool');
+demo = fullfile(lcs_tool_root,'demo','double_gyre');
+
+addpath(lcs_tool_root,demo)

### 2013-12-09 listings/double_gyre/lambda_lcs.m

@@ -1,0 +1,38 @@
+%% Input parameters
+epsilon = .1;
+amplitude = .1;
+omega = pi/5;
+domain = [0,2;0,1];
+resolution = [500,250];
+timespan = [0,5];
+
+%% Velocity definition
+lDerivative = @(t,x,~)derivative(t,x,false,epsilon,amplitude,omega);
+incompressible = true;
+
+%% LCS parameters
+cgStrainOdeSolverOptions = odeset('relTol',1e-5);
+
+% Lambda-lines
+lambda = 1;
+lambdaLineOdeSolverOptions = odeset('relTol',1e-6);
+poincareSection = struct('endPosition',{},'numPoints',{},'orbitMaxLength',{});
+poincareSection(1).endPosition = [.55,.55;.2,.5];
+poincareSection(2).endPosition = [1.53,.45;1.9,.5];
+
+% Number of orbit seed points along each Poincare section
+[poincareSection.numPoints] = deal(100);
+
+% Set maximum orbit length to twice the expected circumference
+nPoincareSection = numel(poincareSection);
+for i = 1:nPoincareSection
+    rOrbit = hypot(diff(poincareSection(i).endPosition(:,1)),diff(poincareSection(i).endPosition(:,2)));
+    poincareSection(i).orbitMaxLength = 2*(2*pi*rOrbit);
+end
+
+%% Cauchy-Green strain eigenvalues and eigenvectors
+[cgEigenvector,cgEigenvalue] = eig_cgStrain(lDerivative,domain,resolution,timespan,'incompressible',incompressible,'odeSolverOptions',cgStrainOdeSolverOptions);
+
+%% Lambda-line LCSs
+[shearline.etaPos,shearline.etaNeg] = lambda_line(cgEigenvector,cgEigenvalue,lambda);
+closedLambdaLine = poincare_closed_orbit_multi(domain,resolution,shearline,poincareSection,'odeSolverOptions',lambdaLineOdeSolverOptions);

### 2013-12-09 listings/double_gyre/lambda_strain_stretch_lcs.m

@@ -1,0 +1,12 @@
+%% Strainlines LCSs
+strainlineMaxLength = 20;
+gridSpace = diff(domain(1,:))/(double(resolution(1))-1);
+strainlineLocalMaxDistance = 2*gridSpace;
+
+strainlineLcs = seed_curves_from_lambda_max(strainlineLocalMaxDistance,strainlineMaxLength,cgEigenvalue(:,2),cgEigenvector(:,1:2),domain,resolution);
+
+%% Stretchlines LCSs
+stretchlineMaxLength = 20;
+stretchlineLocalMaxDistance = 10*gridSpace;
+
+stretchlineLcs = seed_curves_from_lambda_max(stretchlineLocalMaxDistance,stretchlineMaxLength,-cgEigenvalue(:,1),cgEigenvector(:,3:4),domain,resolution);

### 2013-12-19 main.tex
Update Bickley jet section

@@
 \abstract{
-We give a practical introduction to Lagrangian coherent structures (LCSs.) The article reviews recent results in variational LCS theory, presents algorithms to detect LCSs in two dimensional time dependent flows, and demonstrates a LCS computational engine called LCS Tool using examples.
+We give an applied introduction to Lagrangian coherent structures (LCSs.) We review recent results in variational LCS theory, present algorithms based on the theory and implement the algorithms in a computational engine called LCS Tool.
+We use LCS Tool to analyse flow examples.
 }
@@
 We continue the example by adding the calculation of hyperbolic strainline and stretchline LCSs.
-Figures~\ref{fig:double_gyre_lambda_strain_lcs} and~\ref{fig:double_gyre_lambda_stretch_lcs} show the results and listing~\ref{l:double_gyre_lambda_strain_stretch_lcs} gives the LCS Tool commands.
-Note that the local maximisation distance is usually set larger for stretchlines than for strainlines.
-This is due to the fact stretchlines do not follow $\lambda_1$ minimum ridges.
+Figures~\ref{fig:double_gyre_lambda_strain_lcs} and~\ref{fig:double_gyre_lambda_stretch_lcs} show the results and Listing~\ref{lst:double_gyre_lambda_strain_stretch_lcs} gives the LCS Tool commands.
+The local maximisation distance is set larger for stretchlines than for strainlines usually.
+This is because strainlines follow ridges of $\lambda_2$ maximums whereas stretchlines do not follow ridges of $\lambda_1$ minimums.
 
@@
 
-\lstinputlisting[caption={LCS Tool commands for strainline and stretchline LCSs},label=l:double_gyre_lambda_strain_stretch_lcs]{listings/double_gyre/lambda_strain_stretch_lcs.m}
+\lstinputlisting[caption={LCS Tool commands for strainline and stretchline LCSs},label=lst:double_gyre_lambda_strain_stretch_lcs]{listings/double_gyre/lambda_strain_stretch_lcs.m}
 
 \clearpage
+
 \subsection{Bickley jet}
@@
 
-The velocity is given by $\boldsymbol{v}(x,y,t) = (-\partial_y \psi, \partial_x \psi)$ and
+The velocity is given by $\boldsymbol{v}(x,y,t) = (-\partial_y \psi, \partial_x \psi)$ where
 \begin{gather*}
@@
 \end{gather*}
-The parameters used are: $U = 62.66$, $c_2 = 0.205 U$, $c_3 = 0.461 U$, $L_y = 1.77 \times 10^6$, $\epsilon_1 = 0.0075$, $\epsilon_2 = 0.04$, $\epsilon_3 = 0.3$, $L_x = 6.371 \times 10^6 \pi$, $k_n = 2 n \pi/L_x$, $\sigma_1 = 0.5 k_2 (c_2 - c_3)$, $\sigma_2 = 2 \sigma_1$.
+The parameter values we use are: $U = 62.66$, $c_2 = 0.205 U$, $c_3 = 0.461 U$, $L_y = 1.77 \times 10^6$, $\epsilon_1 = 0.0075$, $\epsilon_2 = 0.04$, $\epsilon_3 = 0.3$, $L_x = 6.371 \times 10^6 \pi$, $k_n = 2 n \pi/L_x$, $\sigma_1 = 0.5 k_2 (c_2 - c_3)$, $\sigma_2 = 2 \sigma_1$.
 
-Figure~\ref{f:Bickley jet FTLE} shows the overall appearance of the Bickley jet.
-To get a sense of the numerical parameters required to detect LCSs in this flow, we analyse lambda-line closed orbits of one vortex for a short timespan.
+Figure~\ref{fig:Bickley_jet_FTLE} shows the overall appearance of the Bickley jet and in Figure~\ref{fig:Bickley_jet_lambda_LCS_full} lambda-line LCSs are shown.
+The latter figure demonstrates a poor result since the LCSs change shape significantly from one eddy to the next.
+To address this problem we could try to adjust numerical parameters such as the resolution, but with our hardware memory availability was reached before achieving convergence.
+To circumvent this problem, we analyse lambda-line closed orbits of one vortex for a short timespan.
 
 \begin{figure}
-\begin{center}
+  \centering
   \includegraphics[width=.85\textwidth]{graphics/bickley_jet/ftle_overview}
-\end{center}
   \caption{Bickley jet FTLE}
-\label{f:Bickley jet FTLE}
-\end{figure}
-
-% $[0,2L_x/U]$.
-% Loose closed orbits as increase lambda-line integration accuracy
-
-First, we set a Poincare section so that the first point is near the centre of the region where we seek a vortex and the second point is outside this region, or at least a reasonable distance from the outermost closed lambda-line we find.
-
-% We analyse lambda-line LCSs in a subdomain of this flow.
-% 1. FTLE field
-% 2. Poincare sections
-% 3. Convergence by varying resolution and other parameters.
-
-Start with numerical parameters that give answers quickly and increase accuracy until observe convergence visually.
-
-\begin{lstlisting}[caption={Bickley jet input file: \texttt{demo/bickley\_jet/hyperbolic\_shear\_lcs.m}.},label=l:Bickley jet input]
-%% Input parameters
-u = 62.66;
-lengthX = pi*earthRadius;
-lengthY = 1.77e6;
-epsilon = [.075,.4,.3];
-domain = [0,lengthX;[-1,1]*2.25*lengthY];
-
-% Make x and y grid spacing as equal as possible
-resolutionX = 500;
-gridSpace = diff(domain(1,:))/(double(resolutionX)-1);
-resolutionY = round(diff(domain(2,:))/gridSpace);
-resolution = [resolutionX,resolutionY];
-
-timespan = [0,4*lengthX/u];
-
-%% Velocity definition
-perturbationCase = 3;
-derivative = @(t,x,~)derivative(t,x,false,u,lengthX,lengthY,epsilon,perturbationCase);
-incompressible = true;
-periodicBc = [true,false];
-\end{lstlisting}
-
-To calculate the Cauchy-Green eigenvalues and eigenvectors we therefore set the grid resolution to $500 \times 199$.
-We also specify the flow is incompressible.
-Commands to obtain lambda-lines are given in listing~\ref{l:Bickley jet lambda-line commands}.
-The primary input needed are the positions of the Poincare sections which we determine
-
-\begin{lstlisting}[caption={LCS Tool commands Bickley jet lambda-lines, subset of \texttt{demo/bickley\_jet/hyperbolic\_shear\_lcs.m}.},label=l:Bickley jet lambda-line commands]
-lambda = 1;
-lambdaLineOdeSolverOptions = odeset('relTol',1e-4);
-
-poincareSection = struct('endPosition',{},'numPoints',{},'orbitMaxLength',{});
-poincareSection(1).endPosition = [6.5,-1.4;5.2,-1.4]*1e6;
-poincareSection(2).endPosition = [1.35e7,-1.4e6;1.5e7,-.5e6];
-poincareSection(3).endPosition = [3.25,1.5;1.4,2.6]*1e6;
-poincareSection(4).endPosition = [1e7,1.5e6;8e6,2.6e6];
-poincareSection(5).endPosition = [1.65e7,1.5e6;1.5e7,2.6e6];
-
-% Number of orbit seed points along each Poincare section
-[poincareSection.numPoints] = deal(80);
-
-% Set maximum orbit length to twice the expected circumference
-nPoincareSection = numel(poincareSection);
-for i = 1:nPoincareSection
-    rOrbit = hypot(diff(poincareSection(i).endPosition(:,1)),diff(poincareSection(i).endPosition(:,2)));
-    poincareSection(i).orbitMaxLength = 2*(2*pi*rOrbit);
-end
-
-[shearline.etaPos,shearline.etaNeg] = lambda_line(cgEigenvector,cgEigenvalue,lambda);
-closedLambdaLine = poincare_closed_orbit_multi(domain,resolution,shearline,poincareSection,'odeSolverOptions',lambdaLineOdeSolverOptions,'periodicBc',periodicBc);
-\end{lstlisting}
-
-To obtain strainline hyperbolic LCS, we must set a local maximisation distance and a maximum LCS length. To demonstrate the
-
-The forward time LCS obtained are shown in Figure~\ref{f:bickley jet hyperbolic shear lcs strainline} and the backward time LCS (i.e.
-with the flow timespan set to $t \in [4 L_x/U,0]$ are shown in Figure~\ref{f:bickley jet hyperbolic shear lcs stretchline}.
-In Figures~\ref{f:bickley jet hyperbolic shear lcs details strainline} and~\ref{f:bickley jet hyperbolic shear lcs details stretchline} the same LCS are shown with details.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\textwidth]{graphics/bickley_jet/hyperbolic_shear_lcs_strainline}
-\end{center}
-\caption{Bickley jet lambda-line (green) and strainline (red) LCS.}
-\label{f:bickley jet hyperbolic shear lcs strainline}
+  \label{fig:Bickley_jet_FTLE}
 \end{figure}
@@
 \begin{figure}
-\begin{center}
-\includegraphics[width=\textwidth]{graphics/bickley_jet/hyperbolic_shear_lcs_stretchline}
-\end{center}
-\caption{Bickley jet backward time shear (green) and stretchline (blue) LCS.}
-\label{f:bickley jet hyperbolic shear lcs stretchline}
+  \centering
+  \includegraphics[width=.85\textwidth]{graphics/bickley_jet/lambda_lcs_full}
+  \caption{Bickley jet lambda-line LCSs}
+  \label{fig:Bickley_jet_lambda_LCS_full}
+\end{figure}
+
+To calculate strainline and stretchline LCSs we set the timespan to $t \in [0,2L_x/U]$.
+The script used to detect strainlines is given in Listing~\ref{lst:bickley_jet_strain_lcs} and the results are shown in Figure~\ref{fig:Bickley_jet_strain_lcs}.
+Stretchline LCSs are shown in Figure~\ref{fig:Bickley_jet_stretch_lcs}.
+
+\lstinputlisting[caption={LCS Tool commands for Bickley jet strainline LCSs},label=lst:bickley_jet_strain_lcs]{listings/bickley_jet/strain_lcs.m}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=.85\textwidth]{graphics/bickley_jet/strain_lcs}
+  \caption{Bickley jet strainline LCSs}
+  \label{fig:Bickley_jet_strain_lcs}
 \end{figure}
@@
 \begin{figure}
-\begin{center}
-\includegraphics[width=\textwidth]{graphics/bickley_jet/hyperbolic_shear_lcs_details_strainline}
-\end{center}
-\caption{Same LCS as in Figure~\ref{f:bickley jet hyperbolic shear lcs strainline}, but with additional details.
-Background gray-scale is the finite-time Lyapunov exponent.
-Dashed green lines are Poincare sections.
-Green curves are closed lambda-lines with thick green curves being outermost closed lambda-lines, i.e.
-lambda-line LCSs.
-White dots indicate $\lambda_2$ local maximums used for strainline integration initial positions.}
-\label{f:bickley jet hyperbolic shear lcs details strainline}
+  \centering
+  \includegraphics[width=.85\textwidth]{graphics/bickley_jet/stretch_lcs}
+  \caption{Bickley jet stretchline LCSs}
+  \label{fig:Bickley_jet_stretch_lcs}
 \end{figure}
 
-\begin{figure}
-\begin{center}
-\includegraphics[width=\textwidth]{graphics/bickley_jet/hyperbolic_shear_lcs_details_stretchline}
-\end{center}
-\caption{Same LCS as in Figure~\ref{f:bickley jet hyperbolic shear lcs stretchline}, but with additional details.
-Background gray-scale is the finite-time Lyapunov exponent.
-Dashed green lines are Poincare sections.
-Green curves are closed lambda-lines with thick green curves being outermost closed lambda-lines, i.e.
-lambda-line LCSs.
-White dots indicate $\lambda_2$ local maximums used for stretchline integration initial positions.}
-\label{f:bickley jet hyperbolic shear lcs details stretchline}
-\end{figure}
+\clearpage
 

### 2013-12-19 graphics/bickley_jet/ftle_overview.m

@@ -1,0 +1,43 @@
+function ftle_overview
+
+%% Input parameters
+u = 62.66;
+lengthX = pi*earthRadius;
+lengthY = 1.77e6;
+epsilon = [.075,.4,.3];
+domain = [0,lengthX;[-1,1]*2.25*lengthY];
+
+%% Velocity definition
+perturbationCase = 3;
+phiTimespan = [0,25];
+phiInitial = [0,0];
+phiSol = ode45(@d_phi,phiTimespan,phiInitial);
+timeResolution = 1e5;
+phi1 = deval(phiSol,linspace(phiTimespan(1),phiTimespan(2),timeResolution),1);
+phi1Max = max(phi1);
+lDerivative = @(t,x,~)derivative(t,x,false,u,lengthX,lengthY,epsilon,perturbationCase,phiSol,phi1Max);
+incompressible = true;
+
+% Make x and y grid spacing as equal as possible
+resolutionX = 500;
+gridSpace = diff(domain(1,:))/(double(resolutionX)-1);
+resolutionY = round(diff(domain(2,:))/gridSpace);
+resolution = [resolutionX,resolutionY];
+
+timespan = [0,4*lengthX/u];
+
+hAxes = setup_figure(domain);
+
+%% Cauchy-Green strain eigenvalues and eigenvectors
+cgEigenvalue = eig_cgStrain(lDerivative,domain,resolution,timespan,'incompressible',incompressible);
+
+% Plot finite-time Lyapunov exponent
+cgEigenvalue2 = reshape(cgEigenvalue(:,2),fliplr(resolution));
+ftle_ = ftle(cgEigenvalue2,diff(timespan));
+plot_ftle(hAxes,domain,resolution,ftle_);
+colormap(hAxes,flipud(gray))
+hFigure = 1;
+delete(findobj(hFigure,'Type','axes','Tag','Colorbar'))
+
+filename = 'ftle_overview';
+print_pdf(hFigure,filename)

### 2013-12-19 graphics/bickley_jet/ftle_resolutions.m

@@ -1,40 +1,0 @@
-% Plot FTLE at various resolutions
-
-function ftle_resolutions
-
-lcs_tool_root = fullfile('..','..','..','..','lcs_toolbox');
-demo = fullfile(lcs_tool_root,'demo','bickley_jet');
-
-addpath(lcs_tool_root,demo)
-
-resolutionX = [100,250,500,1000];
-
-%% Input parameters
-u = 62.66;
-lengthX = pi*earthRadius;
-lengthY = 1.77e6;
-epsilon = [.075,.4,.3];
-domain = [0,lengthX;[-1,1]*2.25*lengthY];
-timespan = [0,4*lengthX/u];
-
-perturbationCase = 3;
-lDerivative = @(t,x,~)derivative(t,x,false,u,lengthX,lengthY,epsilon,perturbationCase);
-incompressible = true;
-
-for m = 1:numel(resolutionX)
-   gridSpace = diff(domain(1,:))/(double(resolutionX(m))-1);
-   resolutionY = round(diff(domain(2,:))/gridSpace);
-   resolution = [resolutionX(m),resolutionY];
-   
-   cgEigenvalue = eig_cgStrain(lDerivative,domain,resolution,timespan,'incompressible',incompressible);
-   cgEigenvalue2 = reshape(cgEigenvalue(:,2),fliplr(resolution));
-   ftle_ = ftle(cgEigenvalue2,diff(timespan));
-
-   hAxes = setup_figure(domain);
-   title(hAxes,['Resolution: ',num2str(resolution(1)),'\times',num2str(resolution(2))])
-   plot_ftle(hAxes,domain,resolution,ftle_);
-   colormap(hAxes,flipud(gray))
-   filename = ['ftle_',num2str(resolution(1)),'_',num2str(resolution(2))];
-   print_pdf(gcf,filename)
-   close(gcf)
-end

### 2013-12-19 graphics/bickley_jet/startup.m

@@ -1,0 +1,4 @@
+lcs_tool_root = fullfile('..','..','lcs_tool');
+demo = fullfile(lcs_tool_root,'demo','bickley_jet');
+
+addpath(lcs_tool_root,demo)

### 2013-12-19 graphics/bickley_jet/strain_lcs.m

@@ -1,0 +1,67 @@
+function strain_lcs
+
+%% Input parameters
+u = 62.66;
+lengthX = pi*earthRadius;
+lengthY = 1.77e6;
+epsilon = [.075,.4,.3];
+domain = [2e6,.55*lengthX;[-1,.25]*2.25*lengthY];
+resolutionX = 500;
+timespan = [0,2*lengthX/u];
+
+%% Velocity definition
+perturbationCase = 3;
+phiTimespan = [0,25];
+phiInitial = [0,0];
+phiSol = ode45(@d_phi,phiTimespan,phiInitial);
+timeResolution = 1e5;
+phi1 = deval(phiSol,linspace(phiTimespan(1),phiTimespan(2),timeResolution),1);
+phi1Max = max(phi1);
+lDerivative = @(t,x,~)derivative(t,x,false,u,lengthX,lengthY,epsilon,perturbationCase,phiSol,phi1Max);
+incompressible = true;
+
+%% LCS parameters
+% Cauchy-Green strain
+cgStrainOdeSolverOptions = odeset('relTol',1e-4);
+
+% Strainlines
+strainlineMaxLength = 1e8;
+strainlineOdeSolverOptions = odeset('relTol',1e-4);
+strainlineColor = 'r';
+lcsInitialPositionMarkerSize = 2;
+
+warning('off','eig_cgStrain:unequalDelta')
+warning('off','eig_cgStrain:unequalAuxGridDelta')
+warning('off','integrate_line:isDiscontinuousLargeAngle')
+
+% Make x and y grid spacing as equal as possible
+gridSpace = diff(domain(1,:))/(double(resolutionX)-1);
+resolutionY = round(diff(domain(2,:))/gridSpace) + 1;
+resolution = [resolutionX,resolutionY];
+
+strainlineLocalMaxDistance = 4*gridSpace;
+
+%% Cauchy-Green strain eigenvalues and eigenvectors
+[cgEigenvector,cgEigenvalue] = eig_cgStrain(lDerivative,domain,resolution,timespan,'incompressible',incompressible,'odeSolverOptions',cgStrainOdeSolverOptions);
+
+% Plot finite-time Lyapunov exponent
+cgEigenvalue2 = reshape(cgEigenvalue(:,2),fliplr(resolution));
+ftle_ = ftle(cgEigenvalue2,diff(timespan));
+hAxes = setup_figure(domain);
[…]

### 2013-12-19 graphics/bickley_jet/stretch_lcs.m

@@ -1,0 +1,69 @@
+function stretch_lcs
+
+%% Input parameters
+u = 62.66;
+lengthX = pi*earthRadius;
+lengthY = 1.77e6;
+epsilon = [.075,.4,.3];
+domain = [2e6,.55*lengthX;[-1,.25]*2.25*lengthY];
+resolutionX = 500;
+timespan = [0,2*lengthX/u];
+
+%% Velocity definition
+perturbationCase = 3;
+phiTimespan = [0,25];
+phiInitial = [0,0];
+phiSol = ode45(@d_phi,phiTimespan,phiInitial);
+timeResolution = 1e5;
+phi1 = deval(phiSol,linspace(phiTimespan(1),phiTimespan(2),timeResolution),1);
+phi1Max = max(phi1);
+lDerivative = @(t,x,~)derivative(t,x,false,u,lengthX,lengthY,epsilon,perturbationCase,phiSol,phi1Max);
+incompressible = true;
+
+%% LCS parameters
+% Cauchy-Green strain
+cgStrainOdeSolverOptions = odeset('relTol',1e-4);
+
+% Stretchlines
+stretchlineMaxLength = 1e8;
+stretchlineOdeSolverOptions = odeset('relTol',1e-4);
+stretchlineColor = 'b';
+lcsInitialPositionMarkerSize = 2;
+
+% Make x and y grid spacing as equal as possible
+gridSpace = diff(domain(1,:))/(double(resolutionX)-1);
+resolutionY = round(diff(domain(2,:))/gridSpace) + 1;
+resolution = [resolutionX,resolutionY];
+
+stretchlineLocalMaxDistance = 8*gridSpace;
+
+%% Cauchy-Green strain eigenvalues and eigenvectors
+s = warning('off','eig_cgStrain:unequalDelta');
+warning('off','eig_cgStrain:unequalAuxGridDelta')
+[cgEigenvector,cgEigenvalue] = eig_cgStrain(lDerivative,domain,resolution,timespan,'incompressible',incompressible,'odeSolverOptions',cgStrainOdeSolverOptions);
+warning(s)
+
+% Plot finite-time Lyapunov exponent
+cgEigenvalue2 = reshape(cgEigenvalue(:,2),fliplr(resolution));
+ftle_ = ftle(cgEigenvalue2,diff(timespan));
+hAxes = setup_figure(domain);
+plot_ftle(hAxes,domain,resolution,ftle_);
+colormap(hAxes,flipud(gray))
+
[…]

### 2014-01-10 main.tex
Update article
Tables added to document LCS Tool functions syntax

@@
 \boldsymbol r' = \boldsymbol \eta_\pm(\boldsymbol r),
-\label{e:etafields}
+\label{eq:etafields}
 \end{equation}
 where
-\[
-\boldsymbol \eta_\pm = \sqrt{\frac{\lambda_2 - \lambda^2}{\lambda_2 - \lambda_1}} \boldsymbol \xi_1 \pm \sqrt{\frac{\lambda^2 - \lambda_1}{\lambda_2 - \lambda_1}} \boldsymbol \xi_2.
-\]
-In subsequent sections $\lambda = 1$ always.
+\begin{equation}
+\boldsymbol \eta_\pm = \sqrt{\frac{\lambda_2 - \lambda^2}{\lambda_2 - \lambda_1}} \boldsymbol \xi_1 \pm \sqrt{\frac{\lambda^2 - \lambda_1}{\lambda_2 
+- \lambda_1}} \boldsymbol \xi_2.
+\label{eq:eta}
+\end{equation}
+In subsequent sections $\lambda = 1$ always. We call the curves associated with coherent Lagrangian vortices lambda-lines.
 
-We find hyperbolic LCSs from a variational approach based on shear, as presented in~\textcite{farazmand13:_shearless}. The Lagrangian shear is calculated:
+Hyperbolic LCSs are given by a variational approach based on shear presented in~\textcite{farazmand13:_shearless}. The Lagrangian shear is calculated as:
 \[
@@
 
+\clearpage
+
 \section{Methods}
@@
 More specifically, the eigenvalues and eigenvectors of the Cauchy-Green strain tensor are needed.
-The function associated with this calculation in LCS Tool is \texttt{eig\_cgStrain}.
-The main steps of this function are enumerated in Table~\ref{t:Cauchy-Green algorithm}.
+The function associated with this calculation in LCS Tool is \lstinline!eig_cgStrain!.
+The main steps of this function are enumerated in Table~\ref{t:Cauchy-Green algorithm} and Table~\ref{t:eig_cgStrain syntax} summarises the function syntax.
 
@@
 \end{enumerate}
-\caption{Algorithm used to calculate the Cauchy-Green strain tensor eigenvalues and eigenvectors.}
+\caption{Algorithm to calculate the Cauchy-Green strain tensor eigenvalues and eigenvectors}
 \label{t:Cauchy-Green algorithm}
+\end{table}
+
+\begin{table}
+\begin{center}
+\begin{tabular}{|c|p{.7\textwidth}|}
+\hline \hline
+\multicolumn{2}{|p{\textwidth}|}{\lstinline![cgEigenvector,cgEigenvalue] = eig_cgStrain(derivative,domain,timespan,resolution)!}\\
+\hline
+\lstinline!derivative! & function handle for flow velocity equations\\
+\hline
+\lstinline!domain! & $2 \times 2$ array to define flow domain\\
+\hline
+\lstinline!timespan! & $1 \times 2$ array to define flow timespan\\
+\hline
+\lstinline!resolution! & $1 \times 2$ array to define Cauchy-Green strain main grid resolution\\
+\hline
+\lstinline!auxGridRelDelta! & optional scalar between 0 and 0.5 to specify auxiliary grid spacing. Default is $10^{-2}$.\\
+\hline
+\lstinline!eigenvalueFromMainGrid! & optional logical to control whether eigenvalues of Cauchy-Green strain are calculated from main or auxiliary grid. Default is \lstinline!true!.\\
+\hline
+\lstinline!incompressibile! & optional logical to specify if incompressibility is imposed. Default is  \lstinline!false!.\\
+\hline
+\lstinline!odeSolverOptions! & optional \lstinline!odeset! structure to specify flow map integration options\\
+\hline \hline
+\end{tabular}
+\caption{Syntax of the function \lstinline!eig_cgStrain!}
+\label{t:eig_cgStrain syntax}
+\end{center}
 \end{table}
@@
 
-The function \texttt{eig\_cgStrain} provides the option to calculate the eigenvector from the auxiliary grid, but the eigenvalues from the main grid.
+The function \lstinline!eig_cgStrain!
+provides the option to calculate the eigenvector from the auxiliary grid, but the eigenvalues from the main grid.
 We have found that for flows defined analytically the eigenvalues should be calculated from the main grid whereas for the flows defined by datasets, using the auxiliary grid gives better results.
@@
 To alleviate this problem, we solve the flow equations in a vector form where the equations at all grid points are combined into a single system of equation, even though the system of composed of independent blocks of two dimensional first-order ODE systems.
-MATLAB's \texttt{ode45} function is used to perform the integration and the flow integration typically takes five to ten minutes with default error tolerances.
+MATLAB's \lstinline!ode45!
+function is used to perform the integration and the flow integration typically takes five to ten minutes with default error tolerances.
 The drawback of vector form integration is its memory requirements can become excessive at high resolutions, furthermore, writing the velocity function in vector form is more error-prone than the intrinsic two dimensional form.
@@
 Nonetheless, some grid points can be so close to singularities of the Cauchy-Green strain tensor that satisfying $\lambda_2 >= 1$ everywhere could be excessively computationally costly.
-Therefore the function \texttt{eig\_cgStrain} records the number of points at which incompressibility cannot be imposed and this measure can also help to set appropriate integration tolerances.
+Therefore the function \lstinline!eig_cgStrain!
+records the number of points at which incompressibility cannot be imposed and this measure can also help to set appropriate integration tolerances.
 
@@
 
-Shear LCSs, or coherent Lagrangian vortices, are closed orbits in the Lagrangian shear vector fields $\boldsymbol \eta_{\pm}$ defined in Equation~\eqref{e:etafields}.
+Shear LCSs, or coherent Lagrangian vortices, are closed orbits in the Lagrangian shear vector fields $\boldsymbol \eta_{\pm}$ defined in Equation~\eqref{eq:etafields}.
 Since $\lambda = 1$, the closed curves are non-stretching and area preserving.
-We find closed orbits by integrating shearlines starting from a Poincare section and evaluating the first return map.
-A shearline closing onto itself in the starting point is a shear LCS.
+We find closed orbits by integrating lambda-lines starting from a Poincare section and evaluating the first return map.
+A lambda-line closing onto itself in the starting point is a shear LCS.
 
-The main steps of the algorithm used to calculate shear LCS are enumerated in Table~\ref{t:Shear LCS algorithm}
+The main steps of the algorithm used to calculate shear LCS are enumerated in Table~\ref{t:Shear LCS algorithm}. The syntax of shear LCS functions in LCS Tool is in Table~\ref{t:Shear LCS functions}.
 
 \begin{table}
+\begin{center}
 \begin{enumerate}
@@
 \end{enumerate}
-\caption{Algorithm used to calculate Shear LCS.}
+\end{center}
+\caption{Algorithm to calculate Shear LCS}
 \label{t:Shear LCS algorithm}
@@
 
-The first step is to set the position of Poin\-care sections \texttt{poincare\-Section.endPosition} that are straight lines in regions where closed shearlines are expected.
+\begin{table}
+\begin{tabular}{|c|p{.7\textwidth}|}
+\hline \hline
+\multicolumn{2}{|p{\textwidth}|}{\lstinline![shearline.etaPos,shearline.etaNeg] = lambda_line(cgEigenvector,cgEigenvalue,lambda)!}\\
+\hline
+\lstinline!cgEigenvector! & array of Cauchy-Green strain eigenvectors\\
+\hline
+\lstinline!cgEigenvalue! & array of Cauchy-Green strain eigenvalues\\
+\hline
+\lstinline!lambda! & scalar lambda value in Equation~\eqref{eq:eta}\\
+\hline \hline
+\multicolumn{2}{|p{\textwidth}|}{\lstinline![closedOrbits,orbits] = poincare_closed_orbit_multi(domain,resolution,shearline,PSList)!}\\
+\hline
+\lstinline!domain! & array to define flow domain\\
+\hline
+\lstinline!resolution! & $1 \times 2$ array to define Cauchy-Green strain main grid resolution\\
+\hline
+\lstinline!shearline! & structure of arrays of $\eta_+$ and $\eta_-$ values on main grid\\
+\hline
+\lstinline!PSList! & structure for Poincare section end-points, number of lambda-lines launched from Poincare section, and maximum closed lambda-line length\\
+\hline
+\lstinline!nBisection! & optional positive integer. Default is 5.\\
+\hline
+\lstinline!dThresh! & optional scalar. Default is $10^{-2}$.\\
+\hline
+\lstinline!odeSolverOptions! & optional \lstinline!odeset! structure to specify lambda-line integration options\\
+\hline
+\lstinline!periodicBc! & optional $1 \times 2$ logical array to specify periodic boundary conditions. Default is \lstinline![false,false]!.\\
+\hline \hline
+\end{tabular}
+\caption{LCS Tool shear LCS functions syntax}
+\label{t:Shear LCS functions}
+\end{table}
+
+The first step is to set the position of Poin\-care sections \lstinline!poincareSection.endPosition!
+that are straight lines in regions where closed lambda-lines are expected.
 The Poincare section must be oriented in a way that the first endpoint is close to the centre of the expected eddy and the second endpoint is outside the expected eddy region.
@@
 Possibilities are an examination of the finite-time Lyapunov exponent field of the flow, or some prior knowledge of where coherent eddies could be expected.
-Additionally, the number of trajectories launched from the Poincare section \texttt{poincareSection.numPoints} must be defined.
+Additionally, the number of lambda-lines launched from the Poincare section \lstinline!poincareSection.numPoints!
+must be defined.
 A good default value is $100$.
 
-The second step is to integrate the shearlines to obtain the Poincare return map.
-Integration of shearlines is performed with the $\eta_+$ and $\eta_-$ vector fields defined on the main grid.
+The second step is to integrate the lambda-lines to obtain the Poincare return map.
+Integration of lambda-lines is performed with the $\eta_+$ and $\eta_-$ vector fields defined on the main grid.
 These vector fields are derived from eigenvalues and eigenvectors as described in~\textcite{farazmand12:_comput_lagran}.
@@
 \item At next position, check whether vector field has rotated by over 90\degree, if yes, flip the vector field orientation by 180\degree.
-\item Stop integration when shearline returns to Poincare section, or maximum integration length has been reached.
+\item Stop integration when lambda-line returns to Poincare section, or maximum integration length has been reached.
 \end{enumerate}
-\caption{Algorithm used for variable time step integration of shearlines.}
+\caption{Algorithm used for variable time step integration of lambda-lines.}
 \label{t:variable step integration}
@@
 \end{center}
-\caption{Schematic illustration of shearline variable time step integration.
+\caption{Schematic illustration of lambda-line variable time step integration.
 At the initial point, there is an orientation discontinuity at the lower right grid point that must be corrected prior to linear interpolation.
@@
 Some orbits may however deviate far from the Poincare section and their integration could take inordinately long.
-To prevent this behaviour, a maximum orbit length, \texttt{poincareSection.\-orbit\-MaxLength}, is specified.
-In practice, viewing the Poincare section as the radius of a circle and setting the maximum shearline integration length to twice the circumference gives acceptable results.
+To prevent this behaviour, a maximum orbit length, \lstinline!poincareSection.orbitMaxLength!, is specified.
+In practice, viewing the Poincare section as the radius of a circle and setting the maximum lambda-line integration length to twice the circumference gives acceptable results.
 In Figure~\ref{f:Poincare return map}, circle markers indicate closed orbit positions.
 Closed orbits exist where the Poincare return map is zero.
-The function \texttt{poincare\_closed\_orbit} performs the computations.
+The function \lstinline!poincare_closed_orbit!
+performs the computations.
 In Figure~\ref{f:Poincare return map} not all zero crossings have circles.
-This is because a threshold parameter, \texttt{dThresh} whose default value is 1\% of the length of the Poincare section, is used to discard those zero crossings that appear in regions with high numerical noise.
+This is because a threshold parameter, \lstinline!dThresh!
+whose default value is 1\% of the length of the Poincare section, is used to discard those zero crossings that appear in regions with high numerical noise.
 The position of each detected zero crossing is refined by iteratively bisecting the interval of the zero crossing.
-If after a predetermined number of iterations, \texttt{nBisection} whose default value is five, the two points around the zero crossing have absolute values of more \texttt{dThresh}, the zero crossing is discarded.
+If after a predetermined number of iterations, \lstinline!nBisection!
+whose default value is five, the two points around the zero crossing have absolute values of more \lstinline!dThresh!, the zero crossing is discarded.
 In the Figure, discarded zero crossings are seen where $0 < s < 0.05$.
-Once all closed shearline orbits have been located, the outermost orbit of every Poincare section is deemed a shear LCS.
+Once all closed lambda-line orbits have been located, the outermost orbit of every Poincare section is deemed a shear LCS.
 
@@
 \begin{center}
-%\includegraphics[width=\textwidth]{graphics/double_gyre/poincare_return_map}
+\includegraphics[width=\textwidth]{graphics/double_gyre/poincare_return_map}
 \end{center}
@@
 
-The main steps of the algorithm used to calculate hyperbolic LCS are enumerated in Table~\ref{t:Hyperbolic LCS algorithm}
+The main steps of the algorithm used to calculate hyperbolic LCS are enumerated in Table~\ref{t:Hyperbolic LCS algorithm}.
+The main function to compute hyperbolic LCS in LCS Tool is \lstinline!seed_curves_from_lambda_max!
+and its syntax is given in Table~\ref{t:seed_curves_from_lambda_max syntax}.
 
@@
 \end{enumerate}
-\caption{Algorithm used to calculate hyperbolic LCS.}
+\caption{Algorithm to calculate hyperbolic LCS}
 \label{t:Hyperbolic LCS algorithm}
+\end{table}
+
+\begin{table}
+\begin{center}
+\begin{tabular}{|c|p{.7\textwidth}|}
+\hline \hline
+\multicolumn{2}{|p{\textwidth}|}{\lstinline![curvePosition,curveInitialPosition] = seed_curves_from_lambda_max(distance,cgEigenvalue,cgEigenvector,flowDomain,flowResolution)!}\\
+\hline
+\lstinline!distance! & threshold distance for placement of lambda maximums\\
+\hline
+\lstinline!cgEigenvalue! & array of Cauchy-Green strain eigenvalues\\
+\hline
+\lstinline!cgEigenvector! & array of Cauchy-Green strain eigenvectors\\
+\hline
+\lstinline!flowDomain! & $2 \times 2$ array to define flow domain\\
+\hline
+\lstinline!flowResolution! & $1 \times 2$ array to define Cauchy-Green strain main grid resolution\\
+\hline
+\lstinline!periodicBc! & optional $1 \times 2$ logical array to specify periodic boundary conditions. Default is \lstinline![false,false]!.\\
+\hline
+\lstinline!nMaxCurves! & optional maximum number of curves (i.e. strainlines of stretchlines) to generate. Default is \lstinline!numel(cgEigenvalue)!.\\
+\hline
+\lstinline!odeSolverOptions! & optional \lstinline!odeset! structure to specify flow map integration options\\
+\hline \hline
+\end{tabular}
+\end{center}
+\caption{Syntax of the function \lstinline!seed_curves_from_lambda_max!}
+\label{t:seed_curves_from_lambda_max syntax}
 \end{table}
@@
 Then the first strainline is integrated from the largest of all the local maximums forward and backward in time.
-Similarly to shearlines, this integration must take into account eigenvector field orientation discontinuities.
+Similarly to lambda-lines, this integration must take into account eigenvector field orientation discontinuities.
 Prior to initiating strainline integration, a maximum length is defined.
@@
 Finally any strainline segment contained within (closed) shear LCS are removed.
-These calculations are performed by the functions \texttt{seed\_curves\_from\_lambda\_max} and \texttt{remove\_strain\_in\_shear}.
+These calculations are performed by the functions \lstinline!seed_curves_from_lambda_max!
+and \lstinline!remove_strain_in_shear!.
 
@@
 In other words, the local minimisation distance controls where strainlines begin, but not where they stop.
+
+\clearpage
 
@@
 \end{split}
-\label{e:double gyre derivative equations}
+\label{eq:double gyre derivative equations}
 \end{equation}
@@
 
-\begin{lstlisting}[caption={Double gyre derivative function corresponding to Equations~\ref{e:double gyre derivative equations}.},label=l:double gyre derivative]
+\begin{lstlisting}[caption={Double gyre derivative function corresponding to Equations~\ref{eq:double gyre derivative equations}.},label=l:double gyre derivative]
 function derivative_ = derivative(t,x,epsilon,amplitude,omega)
@@
 \begin{eqnarray}
-\dot{\varphi}(\varphi,\theta,t) = -\frac{g}{R^2 f(\theta) \cos\theta}\partial_{\theta}\eta(\varphi,\theta,t)\label{e:dtlon}\\
+\dot{\varphi}(\varphi,\theta,t) = -\frac{g}{R^2 f(\theta) \cos\theta}\partial_{\theta}\eta(\varphi,\theta,t)\label{eq:dtlon}\\
 \dot{\theta}(\varphi,\theta,t) = +\frac{g}{R^2 f(\theta) \cos\theta}\partial_{\varphi}\eta(\varphi,\theta,t)
-\label{e:dtlat}
+\label{eq:dtlat}
 \end{eqnarray}
@@
 The data is given at a spatial resolution of $1/4\degree$ and temporal resolution of $7\,\mathrm{days}$.
-Due to the discrete data, defining the right hand side of Eq.\,(\ref{e:dtlon}) and (\ref{e:dtlat}) involves interpolation in space and time with a spline interpolation scheme.
-An interpolant is generated first, then the function \texttt{flowdata\_derivative} evaluates the interpolant at the needed coordinates.
+Due to the discrete data, defining the right hand side of Eq.\,(\ref{eq:dtlon}) and (\ref{eq:dtlat}) involves interpolation in space and time with a spline interpolation scheme.
+An interpolant is generated first, then the function \lstinline!flowdata_derivative!
+evaluates the interpolant at the needed coordinates.
 The interpolation time was chosen to be $T=90\,\mathrm{days}$, a multiple of the turn over time of the mesoscale eddies in the domain.
@@
 % \frac{\partial}{\partial x_0} \dot{x} &= Dv(x,t) \frac{\partial x}{\partial x_0}\nonumber\\
-% \frac{d}{dt} DF_{t_0}^t &= Dv(x,t) DF_{t_0}^t\label{e:equation of variation}
+% \frac{d}{dt} DF_{t_0}^t &= Dv(x,t) DF_{t_0}^t\label{eq:equation of variation}
 % \end{align}
-% Equation \eqref{e:equation of variation} is the equation of variation and an initial value problem ordinary differential equations system.
+% Equation \eqref{eq:equation of variation} is the equation of variation and an initial value problem ordinary differential equations system.
 The initial condition, $DF_{t_0}^{t_0}$, is the identity matrix.
@@
 % End:
-% LocalWords:  stretchline stretchlines incompressible gyre eigenvector eigenvectors incompressibility dataset datasets LCS strainline shearline Lyapunov Bickley Rossby timespan variational adaptively vortices parametrise advected
-%  LocalWords:  strainlines
+% LocalWords:  stretchline stretchlines incompressible gyre eigenvector eigenvectors incompressibility dataset datasets LCS strainline Lyapunov Bickley Rossby timespan variational adaptively vortices parametrise advected strainlines

### 2014-01-10 graphics/bickley_jet/lambda_lcs_convergence_subdomain.m

@@ -32,6 +32,7 @@ poincareSection.endPosition = [6.5,-1.4;4.5,-3.5]*1e6;
 rOrbit = hypot(diff(poincareSection.endPosition(:,1)),diff(poincareSection.endPosition(:,2)));
 poincareSection.orbitMaxLength = 2*(2*pi*rOrbit);
 lambdaLineColor = [0,.6,0];
+dThresh = 1e-3;
 
 warning('off','eig_cgStrain:unequalDelta')
 warning('off','eig_cgStrain:unequalAuxGridDelta')
@@ -69,7 +70,7 @@ for m = 1:numel(resolutionX)
     drawnow
     
     [shearline.etaPos,shearline.etaNeg] = lambda_line(cgEigenvector,cgEigenvalue,lambda);
-    [closedLambdaLine,~,hFigure] = poincare_closed_orbit_multi(domain,resolution,shearline,poincareSection,'odeSolverOptions',lambdaLineOdeSolverOptions,'dThresh',1e-3,'showGraph',true);
+    [closedLambdaLine,~,hFigure] = poincare_closed_orbit_multi(domain,resolution,shearline,poincareSection,'odeSolverOptions',lambdaLineOdeSolverOptions,'dThresh',dThresh,'showGraph',true);
     delete(hFigure(2))
     hPoincare = findobj(hFigure(1),'type','axes','Tag',[]);
     title(hPoincare,['Poincare return map, resolution: ',num2str(resolution(1)),'\times',num2str(resolution(2))])

### 2014-01-10 graphics/ocean_dataset/hyperbolic_shear_lcs_details_pdf.m

@@ -1,41 +1,18 @@
 function hyperbolic_shear_lcs_details_pdf
 
-lcs_tool_root = fullfile('..','..','..','..','lcs_toolbox');
-ocean_dataset_demo = fullfile(lcs_tool_root,'demo','ocean_dataset');
+hyperbolic_shear_lcs_details
 
-oldFolder = cd(ocean_dataset_demo);
+hFigure = 1;
+filename = 'hyperbolic_shear_lcs_details_strainline';
+delete_title(hFigure)
+print_pdf(hFigure,filename)
 
-try
-    addpath(fullfile('..','..'))
-    hyperbolic_shear_lcs_details
-    cd(oldFolder)
-    
-    hFigure = 1;
-    filename = 'hyperbolic_shear_lcs_details_strainline';
-    savefig(hFigure,filename)
-    delete_title(hFigure)
-    filenamePDF = [filename,'.pdf'];
-    if exist(filenamePDF,'file')
-        delete(filenamePDF)
-    end
-    print_pdf(hFigure,filename)
-    
-    % FIXME hFigure should not be hard-coded; should probably search for
-    % matching title string
-    hFigure = 12;
-    filename = 'hyperbolic_shear_lcs_details_stretchline';
-    savefig(hFigure,filename)
-    delete_title(hFigure)
-    filenamePDF = [filename,'.pdf'];
-    if exist(filenamePDF,'file')
-        delete(filenamePDF)
-    end
-    print_pdf(hFigure,filename)
-
-catch err
-    cd(oldFolder)
-    rethrow(err)
-end
+% FIXME hFigure should not be hard-coded; should probably search for
+% matching title string
+hFigure = 12;
+filename = 'hyperbolic_shear_lcs_details_stretchline';
+delete_title(hFigure)
+print_pdf(hFigure,filename)
 
 function delete_title(hFigure)
 

### 2014-01-10 graphics/ocean_dataset/hyperbolic_shear_lcs_pdf.m

@@ -1,38 +1,16 @@
 function hyperbolic_shear_lcs_pdf
 
-lcs_tool_root = fullfile('..','..','..','..','lcs_toolbox');
-demo = fullfile(lcs_tool_root,'demo','ocean_dataset');
+hyperbolic_shear_lcs
 
-oldFolder = cd(demo);
+hFigure = 1;
+filename = 'hyperbolic_shear_lcs_strainline';
+delete_title(hFigure)
+print_pdf(hFigure,filename)
 
-try
-    addpath(fullfile('..','..'))
-    hyperbolic_shear_lcs
-    cd(oldFolder)
-    
-    hFigure = 1;
-    filename = 'hyperbolic_shear_lcs_strainline';
-    savefig(hFigure,filename)
-    delete_title(hFigure)
-    filenamePDF = [filename,'.pdf'];
-    if exist(filenamePDF,'file')
-        delete(filenamePDF)
-    end
-    print_pdf(hFigure,filename)
-    
-    hFigure = 2;
-    filename = 'hyperbolic_shear_lcs_stretchline';
-    savefig(hFigure,filename)
-    delete_title(hFigure)
-    filenamePDF = [filename,'.pdf'];
-    if exist(filenamePDF,'file')
-        delete(filenamePDF)
-    end
-    print_pdf(hFigure,filename)
-catch err
-    cd(oldFolder)
-    rethrow(err)
-end
+hFigure = 2;
+filename = 'hyperbolic_shear_lcs_stretchline';
+delete_title(hFigure)
+print_pdf(hFigure,filename)
 
 function delete_title(hFigure)
 

### 2014-01-10 graphics/ocean_dataset/startup.m

@@ -1,0 +1,4 @@
+lcs_tool_root = fullfile('..','..','lcs_tool');
+demo = fullfile(lcs_tool_root,'demo','ocean_dataset');
+
+addpath(lcs_tool_root,demo)

### 2014-01-13 main.tex
Update Bickley jet lambda-lines

@@
 \end{equation}
-In subsequent sections $\lambda = 1$ always. We call the curves associated with coherent Lagrangian vortices lambda-lines.
+We call the curves associated with coherent Lagrangian vortices lambda-lines.
 
@@
 
+The script used to analyse lambda-line LCS is given in Listing~\ref{lst:Bickley jet lambda lcs subdomain}.
+The script produces the data shown in Figure~\ref{fig:Bickley jet lambda lcs convergence}.
+Parameter values used may read from the script.
+It is worth pointing out that in order to obtain closed orbits, $\lambda = 0.995$; with $\lambda = 1$ no closed orbits are found.
+
+\lstinputlisting[caption={Bickley jet script for lambda-line LCSs},label=lst:Bickley jet lambda lcs subdomain]{listings/bickley_jet/lambda_lcs_convergence_subdomain.m}
+
+\begin{figure}[hbt]
+  \centering
+  \includegraphics[width=.7\textwidth]{graphics/bickley_jet/lambda_lcs_convergence_400}
+\includegraphics[width=.7\textwidth]{graphics/bickley_jet/lambda_lcs_convergence_500}
+\includegraphics[width=.7\textwidth]{graphics/bickley_jet/lambda_lcs_convergence_600}
+  \caption{Bickley jet closed lambda-line at varying resolutions}
+  \label{fig:Bickley jet lambda lcs convergence}
+\end{figure}
+
 To calculate strainline and stretchline LCSs we set the timespan to $t \in [0,2L_x/U]$.

### 2014-01-13 graphics/bickley_jet/lambda_lcs_convergence_subdomain.m

@@ -25,7 +25,7 @@ incompressible = true;
 cgStrainOdeSolverOptions = odeset('relTol',1e-4);
 
 % Lambda-lines
-lambda = 1;
+lambda = .995;
 lambdaLineOdeSolverOptions = odeset('relTol',1e-6);
 poincareSection.endPosition = [6.5,-1.4;4.5,-3.5]*1e6;
 [poincareSection.numPoints] = deal(100);
@@ -47,6 +47,7 @@ for m = 1:numel(resolutionX)
     disp(['Resolution: ',num2str(resolution)])
     
     hAxes = setup_figure(domain);
+    hFigure = get(hAxes,'parent');
     title(hAxes,['Resolution: ',num2str(resolution(1)),'\times',num2str(resolution(2))])
     
     %% Cauchy-Green strain eigenvalues and eigenvectors
@@ -56,6 +57,7 @@ for m = 1:numel(resolutionX)
     cgEigenvalue2 = reshape(cgEigenvalue(:,2),fliplr(resolution));
     ftle_ = ftle(cgEigenvalue2,diff(timespan));
     plot_ftle(hAxes,domain,resolution,ftle_);
+    delete(findobj(hFigure,'Type','axes','Tag','Colorbar'))
     colormap(hAxes,flipud(gray))
     drawnow
 
@@ -70,12 +72,12 @@ for m = 1:numel(resolutionX)
     drawnow
     
     [shearline.etaPos,shearline.etaNeg] = lambda_line(cgEigenvector,cgEigenvalue,lambda);
-    [closedLambdaLine,~,hFigure] = poincare_closed_orbit_multi(domain,resolution,shearline,poincareSection,'odeSolverOptions',lambdaLineOdeSolverOptions,'dThresh',dThresh,'showGraph',true);
-    delete(hFigure(2))
-    hPoincare = findobj(hFigure(1),'type','axes','Tag',[]);
+    [closedLambdaLine,~,hFigurePoincare] = poincare_closed_orbit_multi(domain,resolution,shearline,poincareSection,'odeSolverOptions',lambdaLineOdeSolverOptions,'dThresh',dThresh,'showGraph',true);
+    delete(hFigurePoincare(2))
+    hPoincare = findobj(hFigurePoincare(1),'type','axes','Tag',[]);
     title(hPoincare,['Poincare return map, resolution: ',num2str(resolution(1)),'\times',num2str(resolution(2))])
[…]

### 2014-01-24 main.tex
reorder theory section, add to ocean example

@@
 
-We find coherent Lagrangian vortices by considering strain, as was done in~\textcite{haller13:_coher_lagran}.
+\subsection{Elliptic LCS}
+We find coherent Lagrangian vortices as closed orbits of one of the two vector fields
+\begin{equation}
+\boldsymbol \eta^{\lambda}_\pm = \sqrt{\frac{\lambda_2 - \lambda^2}{\lambda_2 - \lambda_1}} \boldsymbol \xi_1 \pm \sqrt{\frac{\lambda^2 - \lambda_1}{\lambda_2 
+- \lambda_1}} \boldsymbol \xi_2,
+\label{eq:eta}
+\end{equation}
+which are a function of the eigenvalues and eigenvectors of $C_{t_0}^t(\boldsymbol x_0)$.
+Curves $\gamma$ that are trajectories of $\boldsymbol \eta^{\lambda}_\pm$ can be parametrized as $\boldsymbol r(s)$ with their arclength $s \in [0,\sigma]$. They are given by integrating
+\begin{equation}
+\boldsymbol r' \equiv \frac{d \boldsymbol r}{ds} = \boldsymbol \eta^{\lambda}_\pm(\boldsymbol r).
+\label{eq:etafields}
+\end{equation}
+
+$\lambda$ is a free parameter close to unity that indicates the uniform stretching of the curve $\gamma$, i.e., each tangent vector $\boldsymbol r'(s)$ with length $l_{t_0}$ stretches by a factor $\lambda$ to length $l_{t}$ under the flow map:
+\begin{gather*}
+l_{t_0}(s) = \sqrt{\langle \boldsymbol r'(s), \boldsymbol r'(s) \rangle},\\
+l_t(s) = \sqrt{\langle \boldsymbol r'(s), C_{t_0}^t(\boldsymbol r(s)) \boldsymbol r'(s) \rangle},\\
+l_t(s) = \lambda\, l_{t_0}(s).
+\end{gather*}
+For $\lambda=1$, $\boldsymbol \eta^{\lambda}_\pm$ points in the direction of maximal Lagrangian shear and we obtain uniformly non-stretching curves over the finite time interval $[t_0, t]$. 
+In summary, the boundaries of coherent Lagrangian vortices are closed curves that uniformly stretch by a factor $\lambda$ close to unity when advected by the flow.
+We call them lambda-lines.
+The outermost lambda-line is the physically observable boundary of an eddy, generating the tracer pattern of an isolated rotating fluid region.
+
+Note that $\boldsymbol \eta^{\lambda}_\pm$ follows from the variational principle that the first variation of the averaged Lagrangian strain along the searched closed curve vanishes (see \parencite{haller13:_coher_lagran} for full derivations):
+\begin{eqnarray}
+Q(\gamma) &=& \frac{1}{\sigma} \int_0^\sigma \frac{l_t(s)}{l_{t_0}(s)}\text{d}s \\
+\delta Q(\gamma) &=& 0.
+\end{eqnarray}
+This coherence principle means that the searched curve $\gamma$, the coherent eddy boundary, is a stationary point of the averaged Lagrangian strain functional and therefore nearby fluid elements cannot break away from the curve $\gamma$.
+
+%We find coherent Lagrangian vortices by considering strain, as was done in~\textcite{haller13:_coher_lagran}.
 To summarise the results in that publication, we examine closed curves $\gamma$ and $O(\epsilon)$ neighbourhoods or ``belts'' around them.
@@
 Curves that form coherent Lagrangian vortices are exceptional however and display no observable variability in their average strain within $O(\epsilon)$ neighbourhoods.
-
-The mathematical identification of a coherent Lagrangian vortex para\-metrises a curve $\gamma$ by $\boldsymbol r(s)$ with $s \in [0,\sigma]$ at time $t_0$.
+%
+%The mathematical identification of a coherent Lagrangian vortex para\-metrises a curve $\gamma$ by $\boldsymbol r(s)$ with $s \in [0,\sigma]$ at time $t_0$.
 The length of a tangent vector $\boldsymbol r'(s)$ is $l_{t_0}(s)$.
@@
 The two tangent lengths are calculated:
-\begin{gather*}
-l_{t_0} = \sqrt{\langle \boldsymbol r'(s), \boldsymbol r'(s) \rangle},\\
-l_t = \sqrt{\langle \boldsymbol r'(s), C_{t_0}^t(\boldsymbol r(s)) \boldsymbol r'(s) \rangle},\\
-\end{gather*}
-where $\langle \cdot, \cdot \rangle$ denotes the Euclidean inner product.
-The average tangential strain along $\gamma$ is:
+%\begin{gather*}
+%l_{t_0} = \sqrt{\langle \boldsymbol r'(s), \boldsymbol r'(s) \rangle},\\
+%l_t = \sqrt{\langle \boldsymbol r'(s), C_{t_0}^t(\boldsymbol r(s)) \boldsymbol r'(s) \rangle},\\
+%\end{gather*}
+%where $\langle \cdot, \cdot \rangle$ denotes the Euclidean inner product.
+%The average tangential strain along $\gamma$ is:
+%\[
+%Q(\gamma) = \frac{1}{\sigma} \int_0^\sigma \frac{l_t(s)}{l_{t_0}(s)}\text{d}s.
+%\]
+%
+%Coherent Lagrangian vortices are associated with the strain tensor
+%\[
+%E_\lambda(\boldsymbol x_0) = \frac12 [C_{t_0}^t(\boldsymbol x_0) - \lambda^2 I]
+%\]
+%and curves uniformly stretched by the same factor $\lambda$ when advected by the flow from time $t_0$ to $t$ are closed trajectories of one of two explicit 
+%differential equations: 
+%\begin{equation}
+%\boldsymbol r' = \boldsymbol \eta_\pm(\boldsymbol r),
+%\label{eq:etafields}
+%\end{equation}
+%where
+%\begin{equation}
+%\boldsymbol \eta_\pm = \sqrt{\frac{\lambda_2 - \lambda^2}{\lambda_2 - \lambda_1}} \boldsymbol \xi_1 \pm \sqrt{\frac{\lambda^2 - \lambda_1}{\lambda_2 
+%- \lambda_1}} \boldsymbol \xi_2.
+%\label{eq:eta}
+%\end{equation}
+%We call the curves associated with coherent Lagrangian vortices lambda-lines.
+
+\subsection{Hyperbolic LCS}
+Hyperbolic LCS are trajectories of the two eigenvector fields $\boldsymbol \xi_{1,2}$ of the Cauchy-Green strain tensor $C_{t_0}^t$.
+Repelling hyperbolic LCSs, denoted as strainlines, are obtained as solutions of the differential equation
 \[
-Q(\gamma) = \frac{1}{\sigma} \int_0^\sigma \frac{l_t(s)}{l_{t_0}(s)}\text{d}s.
+\boldsymbol r_1' = \boldsymbol \xi_1(\boldsymbol r),
 \]
+and attracting hyperbolic LCSs, denoted as stretchlines, are obtained as a solution of the differential equation
+\[
+\boldsymbol r_2' = \boldsymbol \xi_2(\boldsymbol r).
+\]
+A normal perturbation to the strainline segment $\boldsymbol r_1'$ grows under the flow map by a factor $\lambda_2$, i.e., fluid elements are repelled by that rate in normal direction.
+Equivalently, a normal perturbation to the stretchline segment $\boldsymbol r_2'$ shrinks with factor $\lambda_1$ and fluid elements are attracted.
 
-Coherent Lagrangian vortices are associated with the strain tensor
-\[
-E_\lambda(\boldsymbol x_0) = \frac12 [C_{t_0}^t(\boldsymbol x_0) - \lambda^2 I]
-\]
-and curves uniformly stretched by the same factor $\lambda$ when advected by the flow from time $t_0$ to $t$ are closed trajectories of one of two explicit 
-differential equations: 
-\begin{equation}
-\boldsymbol r' = \boldsymbol \eta_\pm(\boldsymbol r),
-\label{eq:etafields}
-\end{equation}
-where
-\begin{equation}
-\boldsymbol \eta_\pm = \sqrt{\frac{\lambda_2 - \lambda^2}{\lambda_2 - \lambda_1}} \boldsymbol \xi_1 \pm \sqrt{\frac{\lambda^2 - \lambda_1}{\lambda_2 
-- \lambda_1}} \boldsymbol \xi_2.
-\label{eq:eta}
-\end{equation}
-We call the curves associated with coherent Lagrangian vortices lambda-lines.
-
-Hyperbolic LCSs are given by a variational approach based on shear presented in~\textcite{farazmand13:_shearless}. The Lagrangian shear is calculated as:
+Similarly to ellitpic LCS, hyperbolic LCS are also based on a variational principle, in this case for the Lagrangian shear
 \[
@@
 \]
-The average Lagrangian shear along a curve $\gamma$ over the time interval $[t_0,t]$ is:
+The first variation of the average Lagrangian shear along a hyperbolic LCS vanishes
 \[
@@
 \]
+\[
+\delta P(\gamma) = 0.
+\]
+They are one type of shearless LCS. For details of the derivation of hyperbolic LCS, see \cite{farazmand13:_shearless}.
 
-Similarly to Lagrangian vortices, exceptional hyperbolic curves without shear in $O(\epsilon)$ neighbourhoods are null-geodesics of the Lorentzian metric
-\[
-g(\boldsymbol u, \boldsymbol v) = \langle \boldsymbol u, D(\boldsymbol x_0) \boldsymbol v \rangle.
-\]
-Repelling LCSs are obtained as solutions of the strainline differential equation:
-\[
-\boldsymbol r' = \boldsymbol \xi_1(\boldsymbol r),
-\]
-and attracting LCSs are obtained as a solution of the stretchline differential equation:
-\[
-\boldsymbol r' = \boldsymbol \xi_2(\boldsymbol r).
-\]
+%Hyperbolic LCSs are given by a variational approach based on shear presented in~\textcite{farazmand13:_shearless}. The Lagrangian shear is calculated as:
+%\[
+%p(s) = \frac{\langle \boldsymbol r'(s), D(\boldsymbol r(s)) \boldsymbol r'(s)\rangle}{\sqrt{\langle \boldsymbol r'(s), C(\boldsymbol r(s)) \boldsymbol r'(s)\rangle \langle \boldsymbol r'(s), \boldsymbol r'(s)\rangle}}
+%\]
+%where
+%\[
+%D(\boldsymbol x_0) = \frac12[C(\boldsymbol x_0) \Omega - \Omega C(\boldsymbol x_0)], \quad \Omega = \begin{pmatrix}0&-1\\1&0\end{pmatrix}.
+%\]
+%The average Lagrangian shear along a curve $\gamma$ over the time interval $[t_0,t]$ is:
+%\[
+%P(\gamma) = \frac{1}{\sigma} \int_0^\sigma p(s) \text{d}s.
+%\]
+%
+%Similarly to Lagrangian vortices, exceptional hyperbolic curves without shear in $O(\epsilon)$ neighbourhoods are null-geodesics of the Lorentzian metric
+%\[
+%g(\boldsymbol u, \boldsymbol v) = \langle \boldsymbol u, D(\boldsymbol x_0) \boldsymbol v \rangle.
+%\]
+%Repelling LCSs are obtained as solutions of the strainline differential equation:
+%\[
+%\boldsymbol r' = \boldsymbol \xi_1(\boldsymbol r),
+%\]
+%and attracting LCSs are obtained as a solution of the stretchline differential equation:
+%\[
+%\boldsymbol r' = \boldsymbol \xi_2(\boldsymbol r).
+%\]
 
@@
 
-This section presents the numerical methods used to compute Lagrangian coherent structures. In other words, the algorithms that constitute LCS Tool are presented.
+This section presents the numerical methods used to compute LCS and explains them step by step. In other words, the algorithms that constitute LCS Tool are presented.
 
@@
 The function associated with this calculation in LCS Tool is \lstinline!eig_cgStrain!.
-The main steps of this function are enumerated in Table~\ref{t:Cauchy-Green algorithm} and Table~\ref{t:eig_cgStrain syntax} summarises the function syntax.
+The main steps of this function are enumerated in Table~\ref{t:Cauchy-Green algorithm} and are described in detail in the following part.
+Table~\ref{t:eig_cgStrain syntax} summarises the function syntax.
 
@@
 \hline
-\lstinline!incompressibile! & optional logical to specify if incompressibility is imposed. Default is  \lstinline!false!.\\
+\lstinline!incompressible! & optional logical to specify if incompressibility is imposed. Default is  \lstinline!false!.\\
 \hline
@@
 We can find the optimal resolution by an iterative procedure in which we start with a low resolution and double it until convergence of LCSs is observed visually.
-A resolution of about 500 grid point along the longest axis usually gives good results.
+If the chosen domain only comprises few expected LCS, e.g., one vortex, a resolution of about 500 grid point along the longest axis usually gives good results.
+Otherwise, for larger domains, a higher resolution must be chosen.
 Furthermore an auxiliary grid is setup.
@@
 The improvements the auxiliary grid yields compared to using only the main grid were reported in \textcite{farazmand12:_comput_lagran}.
-Experience indicates that optimal results are obtained when the auxiliary grid spacing is between 1 and 10\% of the main grid spacing.
+Experience indicates that optimal results are obtained when the auxiliary grid spacing is between 1\% and 10\% of the main grid spacing.
 
@@
 \begin{center}
-\includegraphics[width=\textwidth]{graphics/main_aux_grids}
+\includegraphics[width=0.5\textwidth]{graphics/main_aux_grids}
 \end{center}
-\caption{Illustration of main and auxiliary grid used to compute the Cauchy-Green strain. Solid circles are main grid points and hollow circles are auxiliary grid points.}
+\caption{Illustration of main grid (filled circles) and auxiliary grid (empty circles) used to compute the Cauchy-Green strain.
+Solid circles are main grid points and hollow circles are auxiliary grid points.
+The variable \lstinline!auxGridRelDelta!
+specifies the grid spacing of the auxiliary grid relative to the main grid spacing.}
 \label{f:main and auxiliary grids}
@@
 With 1.25 million grid points this calculation can take an excessively long time.
-To alleviate this problem, we solve the flow equations in a vector form where the equations at all grid points are combined into a single system of equation, even though the system of composed of independent blocks of two dimensional first-order ODE systems.
+To alleviate this problem, we solve the flow equations in a vector form where the equations at all grid points are combined into a single system of equations, even though the system is composed of independent blocks of two dimensional first-order ODE systems.
 MATLAB's \lstinline!ode45!
 function is used to perform the integration and the flow integration typically takes five to ten minutes with default error tolerances.
-The drawback of vector form integration is its memory requirements can become excessive at high resolutions, furthermore, writing the velocity function in vector form is more error-prone than the intrinsic two dimensional form.
+The drawback of vector form integration is its memory requirements can become excessive at high resolutions.
+Furthermore, writing the velocity function in vector form is more error-prone than the intrinsic two dimensional form.
 
-The Cauchy-Green strain tensor can have singularities (i.e.
-$C_{t_0}^t(\boldsymbol x_0) = I$).
+The Cauchy-Green strain tensor can have singularities, i.e., $C_{t_0}^t(\boldsymbol x_0) = I$.
 Such points are isolated in general\parencite{delmarcelle94}, therefore no special accommodations for are made in computations for singularities of the Cauchy-Green strain tensor.
@@
 By setting integration tolerances to smaller values the number of grid points with $\lambda_2 < 1$ can be reduced.
-Nonetheless, some grid points can be so close to singularities of the Cauchy-Green strain tensor that satisfying $\lambda_2 >= 1$ everywhere could be excessively computationally costly.
+Nonetheless, some grid points can be so close to singularities of the Cauchy-Green strain tensor that satisfying $\lambda_2 \geq 1$ everywhere could be excessively computationally costly.
 Therefore the function \lstinline!eig_cgStrain!
@@
 As with analytic cases, care should be taken to ensure the interpolation embedded in the flow velocity integration function works correctly when performing integration in vector form.
+In section \ref{sec:oceandataset}, we present an example how LCS are computed for an ocean dataset and details of the interpolation function are given.
 
@@
 A lambda-line closing onto itself in the starting point is a shear LCS.
-
-The main steps of the algorithm used to calculate shear LCS are enumerated in Table~\ref{t:Shear LCS algorithm}. The syntax of shear LCS functions in LCS Tool is in Table~\ref{t:Shear LCS functions}.
+The main steps of the algorithm used to calculate shear LCS are enumerated in Table~\ref{t:Shear LCS algorithm} and described in detail in the following part.
+The syntax of shear LCS functions in LCS Tool can be seen in Table~\ref{t:Shear LCS functions}.
 
@@
 \begin{enumerate}
-\item Position Poincare sections in flow domain
+\item Position Poincare sections in flow domain as initial positions of lambda-lines
+\item Integrate lambda-lines tangent to $\boldsymbol \eta_{\pm}$
 \item Calculate Poincare return map
@@
 \hline
-\lstinline!nBisection! & optional positive integer. Default is 5.\\
+\lstinline!nBisection! & optional positive integer. Default is 5. Number of bisection steps to refine zero crossings of Poincare return map.\\
 \hline
-\lstinline!dThresh! & optional scalar. Default is $10^{-2}$.\\
+\lstinline!dThresh! & optional scalar. Default is $10^{-2}$. Threshold to discard discontinuous zero crossings of Poincare return map.\\
 \hline
@@
 \item At next position, check whether vector field has rotated by over 90\degree, if yes, flip the vector field orientation by 180\degree.
-\item Stop integration when lambda-line returns to Poincare section, or maximum integration length has been reached.
+\item Stop integration when lambda-line returns to Poincare section, lambda-line hits the domain boundary, or maximum integration length has been reached.
 \end{enumerate}
@@
 \begin{center}
-\includegraphics[width=\textwidth]{graphics/variable_step_integration}
+\includegraphics[width=0.8\textwidth]{graphics/variable_step_integration}
 \end{center}
@@
 At point \#1, no orientation discontinuties are present.
-At point \#2 the interpolated eigenvector field must be rotated by 180\degree\,to match the orientation at the previous point, \#1.}
+At point \#2 the interpolated $\eta_\pm$ vector must be rotated by 180\degree\,to match the orientation of the trajectory.}
 \label{f:variable step integration}
@@
 In Figure~\ref{f:Poincare return map}, circle markers indicate closed orbit positions.
-Closed orbits exist where the Poincare return map is zero.
+Closed orbits exist where the Poincare return map, the distance between the final and the initial point of the orbit $P(s)-s$, is zero.
 The function \lstinline!poincare_closed_orbit!
@@
 In Figure~\ref{f:Poincare return map} not all zero crossings have circles.
-This is because a threshold parameter, \lstinline!dThresh!
-whose default value is 1\% of the length of the Poincare section, is used to discard those zero crossings that appear in regions with high numerical noise.
+This is because a threshold parameter \lstinline!dThresh! is used to discard those zero crossings that appear in regions with high numerical noise. 
+Its default value is 1\% of the length of the Poincare section.
 The position of each detected zero crossing is refined by iteratively bisecting the interval of the zero crossing.
@@
 whose default value is five, the two points around the zero crossing have absolute values of more \lstinline!dThresh!, the zero crossing is discarded.
-In the Figure, discarded zero crossings are seen where $0 < s < 0.05$.
-Once all closed lambda-line orbits have been located, the outermost orbit of every Poincare section is deemed a shear LCS.
+In Figure~\ref{f:Poincare return map}, discarded zero crossings are seen where $0 < s < 0.05$.
+Once all valid closed lambda-line orbits have been located, the outermost orbit of every Poincare section is deemed a shear LCS.
 
@@
 \begin{center}
-\includegraphics[width=\textwidth]{graphics/double_gyre/poincare_return_map}
+\includegraphics[width=0.8\textwidth]{graphics/double_gyre/poincare_return_map}
 \end{center}
@@
 
-Strainlines are integrated from local maximums of $\lambda_2$ on the main grid.
-A local maximisation distance is defined.
-This distance is set between two and five grid points since it represents the number of grid points over which the Cauchy-Green strain tensor is susceptible to numerical noise from integration.
-Thus the local maximisation distance is expected to be a function of variable time step integration tolerances.
-A record is made of all grid points where $\lambda_2$ is maximum within a circle whose radius equals the defined maximisation distance.
-Then the first strainline is integrated from the largest of all the local maximums forward and backward in time.
+Strainlines, trajectories of the $\boldsymbol \xi_1$ vector field, are integrated starting from local maxima of $\lambda_2$ on the main grid. 
+The local maximisation distance \lstinline!distance!
+is defined.
+This distance is set to a value between two and five grid points, since it represents the number of grid points over which the Cauchy-Green strain tensor is susceptible to numerical noise from integration.
+Thus, the local maximisation distance is expected to be a function of variable time step integration tolerances.
+A record is made of all grid points where $\lambda_2$ is maximal within a circle whose radius equals the defined maximisation distance.
+Then the first strainline is integrated from the largest of all the local maxima in both directions, i.e., integrating $\boldsymbol \xi_1$ and $-\boldsymbol \xi_1$.
 Similarly to lambda-lines, this integration must take into account eigenvector field orientation discontinuities.
@@
 The maximum length may be seen as safety factor to prevent highly sinuous strainlines leading to inordinately long computations; the maximum length should be set such that most strainline stop at the domain boundary.
-Once a strainline position has been calculated, local maximums of $\lambda_2$ within the minimisation distance are flagged as ineligible initial positions for subsequent strainlines.
-After the first strainline has been obtained and any ineligible local maximums of $\lambda_2$ have been flagged the next strainline is integrated.
-The process continues until all eligible local maximums of $\lambda_2$ have been used to initialise strainline integration.
-Finally any strainline segment contained within (closed) shear LCS are removed.
+Once a strainline position has been calculated, local maxima of $\lambda_2$ within the minimisation distance are flagged as ineligible initial positions for subsequent strainlines.
+After the first strainline has been obtained and any ineligible local maximums of $\lambda_2$ have been flagged, the next strainline is integrated.
+The process continues until all eligible local maxima of $\lambda_2$ have been used to initialise strainline integration.
+Finally, any strainline segment contained within (closed) shear LCS are removed.
 These calculations are performed by the functions \lstinline!seed_curves_from_lambda_max!
@@
 
+\paragraph{Hyperbolic LCS inside ellitpic LCS}
+Explain, figure?
+
 \clearpage
 
-\section{Examples}
+\section{Example}
+In this section we present three examples of how to use LCS-Tool to obtain LCSs in different flows, namley the double gyre flow, the Bickley jet flow, and an oceanic geostrophic flow.
+All three examples are available as demo files in the demo folder of LCS-Tool and the user can follow the computation of LCS step by step.
 
 \subsection{The double gyre}
-
-The double gyre is a model for a pattern observed in geophysical flows\parencite{shadden05:_defin_lagran_lyapun}. The velocity equations are:
+The double gyre is a model for a time-dependent two gyre system observed in geophysical flows \parencite{shadden05:_defin_lagran_lyapun}.
+It consists of two counterrotating sinusoidal vortices with a harmonically oscillating line in-between.
+The velocity equations are:
 \begin{equation}
@@
 \dot y = \pi A \cos[\pi f(x,t)] \sin(\pi y) \frac{\partial f(x,t)}{\partial x}\\
-f(x,t) = \epsilon \sin(\omega t) x^2 + [1 - 2 \epsilon \sin(\omega t)] x
+f(x,t) = \epsilon \sin(\omega t) x^2 + [1 - 2 \epsilon \sin(\omega t)] x.
 \end{split}
@@
 
-The Bickley jet models a zonal jet flanked above and below by
-stationary vortices. This is an idealised model of geophysical flows
+The Bickley jet models a meandering zonal jet flanked above and below by
+counterrotating vortices. This is an idealised model of geophysical flows
 such as the polar night jet perturbed by a Rossby
@@
 \end{gather*}
-The forcing function we choose is a chaotic solution of the Duffing oscillator, specifically
+As a forcing function we choose a chaotic solution of the Duffing oscillator, specifically
 \begin{gather*}
@@
 \subsection{Ocean dataset}
+\label{sec:oceandataset}
 
-For the last example, we choose a measured geophysical flow where, in contrast to the previous examples, the two dimensional incompressible velocity field is presented as a dataset with discrete temporal and spatial resolution.
-Our region of interest is a small domain in the South Atlantic Ocean, where exceptionally coherent eddies, Agulhas rings, were recently found using methods provided by LCS Tool\parencite{haller13:_coher_lagran}.
+For the last example, we choose a measured geophysical flow where, in contrast to the previous analytical examples, the two dimensional incompressible velocity field is presented as a dataset with discrete temporal and spatial resolution.
+Our region of interest is a small domain in the South Atlantic Ocean, where exceptionally coherent eddies, Agulhas rings, were recently found using methods provided by LCS Tool \parencite{haller13:_coher_lagran}.
 The oceanic velocity field is obtained from altimetric satelite measurements of the sea surface height under the assumption that the flow is geostrophic.
@@
 where $g$ is the gravitational acceleration, $R$ is the mean radius of Earth, and $f(\theta):=2\Omega\sin\theta$ is the Coriolis parameter with $\Omega$ being the Earth's mean angular velocity.
+
 The data is given at a spatial resolution of $1/4\degree$ and temporal resolution of $7\,\mathrm{days}$.
@@
 An interpolant is generated first, then the function \lstinline!flowdata_derivative!
-evaluates the interpolant at the needed coordinates.
-The interpolation time was chosen to be $T=90\,\mathrm{days}$, a multiple of the turn over time of the mesoscale eddies in the domain.
+evaluates the interpolants for the zonal and meridional velocity at the needed coordinates.
+Listing \ref{l:griddedInterp} shows the corresponding commands.
+\begin{lstlisting}[caption={Generate interpolant of ocean velocity data set.},label=l:griddedInterp]
+% Define right hand side of ODE, ocean.flow.derivative
+interpMethod = 'spline';
+vlon_interpolant = griddedInterpolant({time,lat,lon},vlon,interpMethod);
+vlat_interpolant = griddedInterpolant({time,lat,lon},vlat,interpMethod);
+ocean.flow.derivative = @(t,y,~)flowdata_derivative(t,y,vlon_interpolant,vlat_interpolant);
+\end{lstlisting}
+For the ocean data, the integration time was chosen to be $T=30\,\mathrm{days}$, 
+a time larger than the turn over time of the mesoscale eddies in the domain.
 The resolution of the grid of initial conditions is set to $400\times400$ which corresponds to a resolution of roughly $0.015^\circ$ and gives good results.
@@
 The analysis with LCS Tool reveals a large coherent eddy bounded by a closed lambda-line at $(3,-32)$ and four further smaller coherent eddy cores.
-
-\begin{itemize}
-\item Cauchy-Green resolution higher than dataset
-\end{itemize}
 

### 2014-01-24 graphics/bickley_jet/lambda_lcs_convergence_subdomain.m

@@ -32,7 +32,7 @@ poincareSection.endPosition = [6.5,-1.4;4.5,-3.5]*1e6;
 rOrbit = hypot(diff(poincareSection.endPosition(:,1)),diff(poincareSection.endPosition(:,2)));
 poincareSection.orbitMaxLength = 2*(2*pi*rOrbit);
 lambdaLineColor = [0,.6,0];
-dThresh = 1e-4;
+dThresh = 1e-3;
 
 warning('off','eig_cgStrain:unequalDelta')
 warning('off','eig_cgStrain:unequalAuxGridDelta')
@@ -76,7 +76,7 @@ for m = 1:numel(resolutionX)
     delete(hFigurePoincare(2))
     hPoincare = findobj(hFigurePoincare(1),'type','axes','Tag',[]);
     title(hPoincare,['Poincare return map, resolution: ',num2str(resolution(1)),'\times',num2str(resolution(2))])
-    set(hPoincare,'ylim',[-1,1]*1e5)
+    set(hPoincare,'ylim',[-2,2]*1e5)
     hLegend = findobj(hFigurePoincare(1),'type','axes','Tag','legend');
     set(hLegend,'Location','SouthWest')
     

### 2014-01-24 graphics/bickley_jet/startup.m

@@ -1,5 +1,4 @@
-graphics_root = fullfile('..');
 lcs_tool_root = fullfile('..','..','lcs_tool');
 demo = fullfile(lcs_tool_root,'demo','bickley_jet');
 
-addpath(graphics_root,lcs_tool_root,demo)
+addpath(lcs_tool_root,demo)

### 2014-01-24 graphics/double_gyre/startup.m

@@ -1,5 +1,4 @@
-graphics_root = fullfile('..');
 lcs_tool_root = fullfile('..','..','lcs_tool');
 demo = fullfile(lcs_tool_root,'demo','double_gyre');
 
-addpath(graphics_root,lcs_tool_root,demo)
+addpath(lcs_tool_root,demo)

### 2014-01-24 graphics/ocean_dataset/startup.m

@@ -1,5 +1,4 @@
-graphics_root = fullfile('..');
 lcs_tool_root = fullfile('..','..','lcs_tool');
 demo = fullfile(lcs_tool_root,'demo','ocean_dataset');
 
-addpath(graphics_root,lcs_tool_root,demo)
+addpath(lcs_tool_root,demo)

### 2014-01-24 listings/bickley_jet/lambda_lcs_convergence_subdomain.m

@@ -30,7 +30,7 @@ poincareSection.endPosition = [6.5,-1.4;4.5,-3.5]*1e6;
 rOrbit = hypot(diff(poincareSection.endPosition(:,1)),diff(poincareSection.endPosition(:,2)));
 poincareSection.orbitMaxLength = 2*(2*pi*rOrbit);
 lambdaLineColor = [0,.6,0];
-dThresh = 1e-4;
+dThresh = 1e-3;
 
 for m = 1:numel(resolutionX)
     % Make x and y grid spacing as equal as possible

### 2014-01-24 main.tex
Minor revisions to commit cafddee1d98089b31ba7308d4d9ac1beeb6b7040
Revisions in main.tex

@@
 \begin{equation}
-\boldsymbol \eta^{\lambda}_\pm = \sqrt{\frac{\lambda_2 - \lambda^2}{\lambda_2 - \lambda_1}} \boldsymbol \xi_1 \pm \sqrt{\frac{\lambda^2 - \lambda_1}{\lambda_2 
-- \lambda_1}} \boldsymbol \xi_2,
+\boldsymbol \eta^{\lambda}_\pm = \sqrt{\frac{\lambda_2 - \lambda^2}{\lambda_2 - \lambda_1}} \boldsymbol \xi_1 \pm \sqrt{\frac{\lambda^2 - \lambda_1}{\lambda_2 - \lambda_1}} \boldsymbol \xi_2,
 \label{eq:eta}
@@
 which are a function of the eigenvalues and eigenvectors of $C_{t_0}^t(\boldsymbol x_0)$.
-Curves $\gamma$ that are trajectories of $\boldsymbol \eta^{\lambda}_\pm$ can be parametrized as $\boldsymbol r(s)$ with their arclength $s \in [0,\sigma]$. They are given by integrating
+Curves $\gamma$ that are trajectories of $\boldsymbol \eta^{\lambda}_\pm$ can be parametrized as $\boldsymbol r(s)$ with their arclength $s \in [0,\sigma]$.
+They are given by integrating
 \begin{equation}
@@
 Note that $\boldsymbol \eta^{\lambda}_\pm$ follows from the variational principle that the first variation of the averaged Lagrangian strain along the searched closed curve vanishes (see \parencite{haller13:_coher_lagran} for full derivations):
-\begin{eqnarray}
-Q(\gamma) &=& \frac{1}{\sigma} \int_0^\sigma \frac{l_t(s)}{l_{t_0}(s)}\text{d}s \\
+\begin{eqnarray*}
+Q(\gamma) &=& \frac{1}{\sigma} \int_0^\sigma \frac{l_t(s)}{l_{t_0}(s)}\text{d}s\\
 \delta Q(\gamma) &=& 0.
-\end{eqnarray}
+\end{eqnarray*}
 This coherence principle means that the searched curve $\gamma$, the coherent eddy boundary, is a stationary point of the averaged Lagrangian strain functional and therefore nearby fluid elements cannot break away from the curve $\gamma$.
-
-%We find coherent Lagrangian vortices by considering strain, as was done in~\textcite{haller13:_coher_lagran}.
-To summarise the results in that publication, we examine closed curves $\gamma$ and $O(\epsilon)$ neighbourhoods or ``belts'' around them.
-For most $\gamma$, neighbouring belts of $O(\epsilon)$ exhibit strain variability of $O(\epsilon)$.
-Curves that form coherent Lagrangian vortices are exceptional however and display no observable variability in their average strain within $O(\epsilon)$ neighbourhoods.
-%
-%The mathematical identification of a coherent Lagrangian vortex para\-metrises a curve $\gamma$ by $\boldsymbol r(s)$ with $s \in [0,\sigma]$ at time $t_0$.
-The length of a tangent vector $\boldsymbol r'(s)$ is $l_{t_0}(s)$.
-The advected curve $F_{t_0}^t(\gamma)$ has a tangent vector $(\text{d}/\text{d}s) F_{t_0}^t(\boldsymbol r(s))$ and a length we denote by $l_t(s)$.
-The two tangent lengths are calculated:
-%\begin{gather*}
-%l_{t_0} = \sqrt{\langle \boldsymbol r'(s), \boldsymbol r'(s) \rangle},\\
-%l_t = \sqrt{\langle \boldsymbol r'(s), C_{t_0}^t(\boldsymbol r(s)) \boldsymbol r'(s) \rangle},\\
-%\end{gather*}
-%where $\langle \cdot, \cdot \rangle$ denotes the Euclidean inner product.
-%The average tangential strain along $\gamma$ is:
-%\[
-%Q(\gamma) = \frac{1}{\sigma} \int_0^\sigma \frac{l_t(s)}{l_{t_0}(s)}\text{d}s.
-%\]
-%
-%Coherent Lagrangian vortices are associated with the strain tensor
-%\[
-%E_\lambda(\boldsymbol x_0) = \frac12 [C_{t_0}^t(\boldsymbol x_0) - \lambda^2 I]
-%\]
-%and curves uniformly stretched by the same factor $\lambda$ when advected by the flow from time $t_0$ to $t$ are closed trajectories of one of two explicit 
-%differential equations: 
-%\begin{equation}
-%\boldsymbol r' = \boldsymbol \eta_\pm(\boldsymbol r),
-%\label{eq:etafields}
-%\end{equation}
-%where
-%\begin{equation}
-%\boldsymbol \eta_\pm = \sqrt{\frac{\lambda_2 - \lambda^2}{\lambda_2 - \lambda_1}} \boldsymbol \xi_1 \pm \sqrt{\frac{\lambda^2 - \lambda_1}{\lambda_2 
-%- \lambda_1}} \boldsymbol \xi_2.
-%\label{eq:eta}
-%\end{equation}
-%We call the curves associated with coherent Lagrangian vortices lambda-lines.
 
@@
 They are one type of shearless LCS. For details of the derivation of hyperbolic LCS, see \cite{farazmand13:_shearless}.
-
-%Hyperbolic LCSs are given by a variational approach based on shear presented in~\textcite{farazmand13:_shearless}. The Lagrangian shear is calculated as:
-%\[
-%p(s) = \frac{\langle \boldsymbol r'(s), D(\boldsymbol r(s)) \boldsymbol r'(s)\rangle}{\sqrt{\langle \boldsymbol r'(s), C(\boldsymbol r(s)) \boldsymbol r'(s)\rangle \langle \boldsymbol r'(s), \boldsymbol r'(s)\rangle}}
-%\]
-%where
-%\[
-%D(\boldsymbol x_0) = \frac12[C(\boldsymbol x_0) \Omega - \Omega C(\boldsymbol x_0)], \quad \Omega = \begin{pmatrix}0&-1\\1&0\end{pmatrix}.
-%\]
-%The average Lagrangian shear along a curve $\gamma$ over the time interval $[t_0,t]$ is:
-%\[
-%P(\gamma) = \frac{1}{\sigma} \int_0^\sigma p(s) \text{d}s.
-%\]
-%
-%Similarly to Lagrangian vortices, exceptional hyperbolic curves without shear in $O(\epsilon)$ neighbourhoods are null-geodesics of the Lorentzian metric
-%\[
-%g(\boldsymbol u, \boldsymbol v) = \langle \boldsymbol u, D(\boldsymbol x_0) \boldsymbol v \rangle.
-%\]
-%Repelling LCSs are obtained as solutions of the strainline differential equation:
-%\[
-%\boldsymbol r' = \boldsymbol \xi_1(\boldsymbol r),
-%\]
-%and attracting LCSs are obtained as a solution of the stretchline differential equation:
-%\[
-%\boldsymbol r' = \boldsymbol \xi_2(\boldsymbol r).
-%\]
 
@@
 \caption{Illustration of main grid (filled circles) and auxiliary grid (empty circles) used to compute the Cauchy-Green strain.
-Solid circles are main grid points and hollow circles are auxiliary grid points.
 The variable \lstinline!auxGridRelDelta!
@@
 \hline
-\lstinline!nBisection! & optional positive integer. Default is 5. Number of bisection steps to refine zero crossings of Poincare return map.\\
+\lstinline!nBisection! & optional number of bisection steps to refine zero crossings of Poincare return map. Default is 5.\\
 \hline
-\lstinline!dThresh! & optional scalar. Default is $10^{-2}$. Threshold to discard discontinuous zero crossings of Poincare return map.\\
+\lstinline!dThresh! & optional threshold to discard discontinuous zero crossings of Poincare return map. Default is $10^{-2}$.\\
 \hline
@@
 
-The first step is to set the position of Poin\-care sections \lstinline!poincareSection.endPosition!
+The first step is to set the position of Poincare sections \lstinline!poincareSection.endPosition!
 that are straight lines in regions where closed lambda-lines are expected.
@@
 \item At next position, check whether vector field has rotated by over 90\degree, if yes, flip the vector field orientation by 180\degree.
-\item Stop integration when lambda-line returns to Poincare section, lambda-line hits the domain boundary, or maximum integration length has been reached.
+\item Stop integration when lambda-line returns to Poincare section, lambda-line reaches the domain boundary, or maximum integration length has been reached.
 \end{enumerate}
@@
 In Figure~\ref{f:Poincare return map} not all zero crossings have circles.
-This is because a threshold parameter \lstinline!dThresh! is used to discard those zero crossings that appear in regions with high numerical noise. 
+This is because a threshold parameter \lstinline!dThresh!
+is used to discard those zero crossings that appear in regions with high numerical noise.
 Its default value is 1\% of the length of the Poincare section.
@@
 
-\paragraph{Hyperbolic LCS inside ellitpic LCS}
-Explain, figure?
+The example we present show hyperbolic LCS inside elliptic LCS.
+This appears contradictory at first glance because LCS should not cross, however numerical experimentation with LCS calculations overlaid with colour tracer advection reveals that hyperbolic LCS can cross elliptic barriers.
 
@@
 \end{lstlisting}
-For the ocean data, the integration time was chosen to be $T=30\,\mathrm{days}$, 
-a time larger than the turn over time of the mesoscale eddies in the domain.
-The resolution of the grid of initial conditions is set to $400\times400$ which corresponds to a resolution of roughly $0.015^\circ$ and gives good results.
+For the ocean data, the integration time was chosen to be $T=30\,\mathrm{days}$, a time larger than the turn over time of the mesoscale eddies in the domain.
+The resolution of the grid of initial conditions is set to $400\times400$ which corresponds to a resolution of roughly 0.015\degree and gives good results.
 With this choice the resolution of the tracer grid is of the order of $15$ times higher than the resolution of the velocity field.
@@
 
-% \subsection{Adaptive Step Size Integration of Strainlines and Shearlines}
-
-% \begin{itemize}
-% \item Motivate necessity for adaptive step size integration. Illustrate (concentrated) high curvature of strainlines and shearlines.
-% \end{itemize}
-
-% Adaptive step size integration algorithms evaluate the right hand side of ODEs in a non-sequential way.
-Thus, the notion of evaluation at a previous time-step presented in~\textcite[Equation 9]{farazmand12:_comput_lagran} requires the use of advanced programming concepts if using MATLAB's numerical differential equation integration functions (e.g.
-ode45, ode113.) Specifically, the LCS Toolbox implementation of strainline and shearline integration uses output functions and handle classes.
-
-% \subsection{Equation of variation to calculate differential of flow map}
-
-% In~\textcite{farazmand12:_comput_lagran} an auxiliary grid is used to calculate the gradient of the flow map, $DF_{t_0}^t$.
-Experience suggests the quality of results obtained with this method is variable.
-An alternative is to use the equation of variation.
-
-% The flow is given by an equation of the form:
-% \[
-% \dot{x} = v(x,t).
-% \]
-% Application of the chain rule gives the equation of variation:
-% \begin{align}
-% \frac{\partial}{\partial x_0} \dot{x} &= Dv(x,t) \frac{\partial x}{\partial x_0}\nonumber\\
-% \frac{d}{dt} DF_{t_0}^t &= Dv(x,t) DF_{t_0}^t\label{eq:equation of variation}
-% \end{align}
-% Equation \eqref{eq:equation of variation} is the equation of variation and an initial value problem ordinary differential equations system.
-The initial condition, $DF_{t_0}^{t_0}$, is the identity matrix.
-
-% For two dimensional flows,
-% \[
-% DF_{t_0}^t = 
-% \begin{pmatrix}
-% \frac{\partial}{\partial x_1} F_{{t_0}_1}^t & \frac{\partial}{\partial x_2} F_{{t_0}_1}^t\\
-% \frac{\partial}{\partial x_1} F_{{t_0}_2}^t & \frac{\partial}{\partial x_2} F_{{t_0}_2}^t
-% \end{pmatrix},
-% \]
-% and
-% \[
-% Dv(x,t) =
-% \begin{pmatrix}
-% \frac{\partial}{\partial x_1} v_1 & \frac{\partial}{\partial x_2} v_1\\
-% \frac{\partial}{\partial x_1} v_2 & \frac{\partial}{\partial x_2} v_2
-% \end{pmatrix}.
-% \]
-% With the equation of variation, calculating the differential of the flow map consists of solving an ODE system of six equations: $x_1,x_2$, and the four elements of $DF_{t_0}^t$.
-The equations for $DF_{t_0}^t$ depend on $x_1,x_2$, but not vice-versa.
-To solve this system, the gradient of the flow vector field, $Dv(x,t)$ must be known.
-If the velocity field is known analytically, $Dv(x,t)$ can also be calculated analytically.
-% In Appendix \ref{s:traveling wave input file}, the necessary commands are given to do this automatically with the MATLAB Symbolic Math Toolbox for a traveling wave flow (see functions symF and symJacDerivative)
-% If the flow map is not known analytically, splines can be fitted to flow data and the splines differentiated (see Section~\ref{s:splines}.)
-
-% To test the equation of variation method and to compare it to the finite difference method from~\parencite{farazmand12:_comput_lagran}, a double gyre incompressible flow is used.
-Incompressible flows have a Cauchy-Green strain tensor whose exact value is one, $|C_{t_0}^T| = 1$ (and a maximum possible value of $\lambda_1 = 1$.) In Figure~\ref{f:cgStrain convergence}, the convergence of $C_{t_0}^T$ is illustrated; the finite difference with auxiliary grid method and the equation of variation method are compared.
-To observe convergence with the finite difference with auxiliary grid method, the auxiliary grid spacing was found by trial and error; if the auxiliary grid relative spacing is set larger than $10^{-8}$ convergence is difficult to observe.
-% Setting the auxiliary grid relative spacing any smaller than $10^{-8}$ does not seem to help.
-% The data in the figure indicates the error of the equation of variation method is smaller than the finite difference with auxiliary grid method.
-In addition, the equation of variation does not require setting an auxiliary grid spacing parameter, the equation of variation method is better than the finite difference with auxiliary grid method.
-% The convergence measure used, $|||C_{t_0}^T| - 1||_2$, is an average.
-The error can be made no smaller than $10^{-7} at a resolution of 20 \times 10.
-As the resolution increases, the smallest possible error becomes larger.
-
-% \begin{figure}
-% \begin{center}
-% \includegraphics[width=\textwidth]{../../lcs_toolbox/figures/double_gyre/compare_fd_eov}
-% \end{center}
-% \caption{$|||C_{t_0}^T| - 1||_2$ as a function of ODE integration error tolerance.
-The ODE solver is MATLAB's ode45.
-Both the relative and absolute error tolerances are set to the stated values.
-For the finite difference method, the Cauchy-Green strain is obtained from the auxiliary grid with the auxiliary grid spacing $10^{-8}$ of the main grid spacing.
-A double gyre flow is used.
-The flow resolution is $20 \times 10$.
-The data is given in Table~\ref{t:Cauchy-Green strain convergence comparison}.}
-% \label{f:cgStrain convergence}
-% \end{figure}
-
-% $|C_{t_0}^T|$ converges as the flow integration error tolerance decreases and $\lambda_1$ converges as the flow resolution is increased.
-It may be worth noting that $|C_{t_0}^T|$ converge as the resolution increases since it is a quantity calculated at each grid point independently.
-In a complementary way, $\lambda_1$ does not converge as the error tolerance is decreased.
-
-% In Table~\ref{t:lambda products tol}, the ODE solver's relative error is varied.
-As the error tolerance decreases $|C_{t_0}^T|$ converges to 1 (except for the last row, perhaps because the relative error tolerance is smaller than the absolute error tolerance.) This is logical; whereas in Table~\ref{t:lambda products res} the number of points was changing, in Table~\ref{t:lambda products tol} it is fixed.
-By decreasing the error tolerance, the integration at each point becomes more accurate and $|||C_{t_0}^T| - 1||_2$ approaches its exact value.
-The convergence of $||\lambda_1||_\infty$, if it exists at all, seems to saturate relatively quickly, once the relative error tolerance is $10^{-4}$.
-The interpretation of this result is that for the flow resolution chosen for the data in Table~\ref{t:lambda products tol}, $||\lambda_1||_\infty$ has a set value and finding this value accurately does not require a particularly tight error tolerance.
-
-% Note that in tables~\ref{t:lambda products res} and~\ref{t:lambda products tol}, for those quantities that exhibit convergence, the rate of error convergence has not been verified.
-
-% \begin{table}
-% \begin{center}
-% \begin{tabular}{l c c}
-% \hline
-% Resolution & $|||C_{t_0}^T| - 1||_2$ & $||\lambda_1||_\infty$\\
-% \hline
-% $200 \times 100$ & $2.50 \times 10^{-4}$ & 0.896\\
-% $400 \times 200$ & $2.47 \times 10^{-3}$ & 0.976\\
-% $800 \times 400$ & $1.72 \times 10^{-1}$ & 0.976\\
-% $1600 \times 800$ & 14.6 & 0.991\\
-% $3200 \times 1600$ & 4.24 & 0.992\\
-% \hline
-% \end{tabular}
-% \end{center}
-% \caption{Properties of the Cauchy-Green strain tensor, $C$, and $\lambda_1$ at different resolutions. The relative and absolute error tolerances are $10^{-4}$ and $10^{-6}$ respectively.}
-% \label{t:lambda products res}
-% \end{table}
-
-% \begin{table}
-% \begin{center}
-% \begin{tabular}{l c c}
-% \hline
-% Rel. tol. & $|||C_{t_0}^T| - 1||_2$ & $||\lambda_1||_\infty$\\
-% \hline
-% $10^{-2}$ & $1.24 \times 10^{-1}$ & 1.031\\
-% $10^{-3}$ & $6.13 \times 10^{-3}$ & 0.963\\
-% $10^{-4}$ & $3.53 \times 10^{-3}$ & 0.962\\
-% $10^{-5}$ & $1.37 \times 10^{-3}$ & 0.962\\
-% $10^{-6}$ & $1.28 \times 10^{-3}$ & 0.962\\
-% $10^{-8}$ & $2.47 \times 10^{-3}$ & 0.962\\
-% \hline
-% \end{tabular}
-% \end{center}
-% \caption{Properties of the Cauchy-Green strain tensor, $C$, and $\lambda_1$ at different relative error tolerances. The resolution is $500\times250$. The absolute error tolerance is $10^{-6}$.}
-% \label{t:lambda products tol}
-% \end{table}
-
-% \subsection{Splines}
-% \label{s:splines}
-
-% To integrate the equation of variation, the derivative of the flow vector field, $Dv(x,t)$, must be known.
-When working with data-defined flows rather than analytic flows, $Dv(x,t)$ can be found by fitting splines to the data and differentiating the splines~\parencite{finn05:_volum}.
-
-% This procedure is implemented with the MATLAB Curve Fitting Toolbox.
-For validation, a spline is created by sampling an analytic flow.% using the function analytic2spline (see Appendix \ref{s:analytic2spline}.)
-
-% In Table~\ref{t:splines} the double gyre is used for comparison; the spline is created at various resolutions and then compared at one point to the analytic definition.
-The data show as sampling resolution increases, the spline approximation error decreases.
-The resolutions in Table~\ref{t:splines} are lower than those in Table~\ref{t:lambda products res}.
-Although systematic tests have not been performed, it appears using splines uses more memory than the analytic approach.
-
-% \begin{table}
-% \begin{center}
-% \begin{tabular}{l c c}
-% \hline
-% Resolution & $|v - v_\text{spline}|$ & $|Dv - Dv_\text{spline}|$\\
-% \hline
-% $10 \times 5$ & $4.18 \times 10^{-2}$ & $3.71 \times 10^{-1}$\\
-% $20 \times 10$ & $8.44 \times 10^{-3}$ & $1.12 \times 10^{-1}$\\
-% $40 \times 20$ & $3.72 \times 10^{-3}$ & $6.00 \times 10^{-2}$\\
-% $80 \times 40$ & $2.51 \times 10^{-3}$ & $4.33 \times 10^{-2}$\\
-% $160 \times 80$ & $2.06 \times 10^{-3}$ & $3.66 \times 10^{-2}$\\
-% \hline
-% \end{tabular}
-% \end{center}
-% \caption{Numerical convergence of splines for double gyre flow. Time = 0 and position = (-0.25,0.9).}
-% \label{t:splines}
-% \end{table}
-
-% \subsection{Superminimization}
-% \label{s:superminimzation}
-
 \section{Conclusions}

### 2014-01-27 main.tex
Replace minimums/maximums → minima/maxima

@@
 \item Define a local maximisation distance.
-\item Find all local maximums of $\lambda_2$ on the main grid such that they are maximums within a circle with a radius equal to the local minimisation distance.
+\item Find all local maxima of $\lambda_2$ on the main grid such that they are maxima within a circle with a radius equal to the local minimisation distance.
 \item Define a maximum strainline length
@@
 Integrate until the strainline has attained the maximum strainline length, or until it has reached the domain boundary.
-\item Flag any remaining local maximums of $\lambda_2$ within the minimisation distance of the strainline as ineligible initial positions for subsequent strainlines
-\item Continue integrating strainlines using local maximums of $\lambda_2$ as initial positions until no eligible local maximums of $\lambda_2$ remain.
+\item Flag any remaining local maxima of $\lambda_2$ within the minimisation distance of the strainline as ineligible initial positions for subsequent strainlines
+\item Continue integrating strainlines using local maxima of $\lambda_2$ as initial positions until no eligible local maxima of $\lambda_2$ remain.
 \item Remove any strainline segment contained within (closed) shear LCS
@@
 \hline
-\lstinline!distance! & threshold distance for placement of lambda maximums\\
+\lstinline!distance! & threshold distance for placement of lambda maxima\\
 \hline
@@
 Once a strainline position has been calculated, local maxima of $\lambda_2$ within the minimisation distance are flagged as ineligible initial positions for subsequent strainlines.
-After the first strainline has been obtained and any ineligible local maximums of $\lambda_2$ have been flagged, the next strainline is integrated.
+After the first strainline has been obtained and any ineligible local maxima of $\lambda_2$ have been flagged, the next strainline is integrated.
 The process continues until all eligible local maxima of $\lambda_2$ have been used to initialise strainline integration.
@@
 The local maximisation distance is set larger for stretchlines than for strainlines usually.
-This is because strainlines follow ridges of $\lambda_2$ maximums whereas stretchlines do not follow ridges of $\lambda_1$ minimums.
+This is because strainlines follow ridges of $\lambda_2$ maxima whereas stretchlines do not follow ridges of $\lambda_1$ minima.
 
@@
 lambda-line LCSs.
-White dots indicate $\lambda_2$ local maximums used for strainline integration initial positions.}
+White dots indicate $\lambda_2$ local maxima used for strainline integration initial positions.}
 \label{f:ocean dataset hyperbolic shear lcs details strainline}
@@
 lambda-line LCSs.
-White dots indicate $\lambda_1$ local minimums used for stretchline integration intial positions.}
+White dots indicate $\lambda_1$ local minima used for stretchline integration intial positions.}
 \label{f:ocean dataset hyperbolic shear lcs details stretchline}
@@
 % End:
-% LocalWords:  stretchline stretchlines incompressible gyre eigenvector eigenvectors incompressibility dataset datasets LCS strainline Lyapunov Bickley Rossby timespan variational adaptively vortices parametrise advected strainlines
+% LocalWords:  stretchline stretchlines incompressible gyre eigenvector eigenvectors incompressibility dataset datasets LCS strainline Lyapunov Bickley Rossby timespan variational adaptively vortices parametrise advected strainlines minima

### 2014-01-30 main.tex
Modify and expand abstract

@@
 \abstract{
-We give an applied introduction to Lagrangian coherent structures (LCSs.) We review recent results in variational LCS theory, present algorithms based on the theory and implement the algorithms in a computational engine called LCS Tool.
-We use LCS Tool to analyse flow examples.
+We give an applied introduction to Lagrangian coherent structure (LCS) detection.
+We review LCS theory that is based on variational calculus, present algorithms conceived in accordance with the variational theory and present a MATLAB computational engine called LCS Tool that enables variational LCS calculations and detection.
+We use LCS Tool to analyse three flow examples: a double gyre, a Bickley jet and an ocean dataset.
 }

### 2014-01-30 main.tex
Modify introduction

@@
 
-LCSs are used in fluid dynamics to capture essential flow features and thereby simplify the description of fluid flows.
-More specifically, LCSs have been particularly well adapted to analyze two dimensional finite time turbulent flows.
-LCSs capture dominant transport barriers, repelling and attracting structures and LCSs are used to identify vortex boundaries.
+LCSs are used in fluid dynamics to capture essential flow features and thereby simplify the description of complex fluid flows.
+More specifically, LCSs are particularly well adapted to analyze two dimensional finite time turbulent flows.
+LCSs identify transport barriers, repelling and attracting structures and vortex boundaries.
+Application areas for LCSs are oceanic and atmospheric flows, biological applications\parencite{wilson09:_lagran_reynol,tallapragada11:_lagran}, aeronautics\parencite{tang10:_accur_lagran_hong_kong_inter_airpor} and mechanical systems\parencite{hadjighasem13:_detec_kam}.
 
-LCSs have been used to analyse oceanic and atmospheric flows, biological applications\parencite{wilson09:_lagran_reynol,tallapragada11:_lagran}, aeronautics\parencite{tang10:_accur_lagran_hong_kong_inter_airpor} and mechanical systems\parencite{hadjighasem13:_detec_kam}.
+The finite-time Lyapunov exponent (FTLE) gives one widely-used mathematical definition for LCSs.
+Although the FTLE is popular, it is possible to find examples where it fails to identify LCSs correctly\parencite{haller11:_lagran_coher_struc,norgard12:_secon_lagran_coher_struc}.
+To remedy these problems, variational calculus principles applied to material curves in a fluid present new approaches to identifying LCSs\parencite{haller11:_lagran_coher_struc,farazmand12:_comput_lagran,haller12:_geodes_theor_trans_barrier_two_dimen_flows,farazmand13:_attrac_lagran,haller13:_coher_lagran}.
+This article presents an algorithmic introduction to variational LCSs using a computational engine called LCS Tool\footnote{LCS Tool is free MATLAB software available for download at: \url{https://github.com/jeixav/LCS-Tool}}.
 
-The finite-time Lyapunov exponent (FTLE) gives a mathematical definition for LCSs.
-Although the FTLE has been applied to a variety a situations, it is possible to find examples where it fails to identify LCSs correctly, giving false positives and false negatives\parencite{haller11:_lagran_coher_struc,norgard12:_secon_lagran_coher_struc}.
-To remedy these problems, variational principles have been used to present new approaches to identifying LCSs\parencite{haller11:_lagran_coher_struc,farazmand12:_comput_lagran,haller12:_geodes_theor_trans_barrier_two_dimen_flows,farazmand13:_attrac_lagran,haller13:_coher_lagran}.
-
-This article presents an algorithmic introduction LCSs using a computational engine for LCSs called LCS Tool\footnote{LCS Tool is free MATLAB software available for download at: \url{https://github.com/jeixav/LCS-Tool}}.
-
-LCS Tool is a library of functions that perform the computations to identify LCSs in two dimensional finite time flows.
+LCS Tool is a library of MATLAB functions that perform the computations to identify LCSs in two dimensional finite time flows.
 LCS Tool aims to initiate scientists to LCSs.
@@
 \end{itemize}
-These packages are based on the FTLE; the objective of this article and LCS Tool is to replace FTLE methods.
+These packages are based on the FTLE; the objective of LCS Tool and this article is to encourage and facilitate variational LCS analysis.
 

### 2014-01-31 main.tex
add colortracer figure

--- 2014-01-27 main.tex
@@
 Strainlines are then integrated until reaching the domain boundary or the maximum length.
-The maximum length may be seen as safety factor to prevent highly sinuous strainlines leading to inordinately long computations; the maximum length should be set such that most strainline stop at the domain boundary.
+The maximum length may be seen as safety factor to prevent highly sinuous strainlines leading to inordinately long computations; the maximum length should be set such that most strainlines stop at the domain boundary.
 Once a strainline position has been calculated, local maxima of $\lambda_2$ within the minimisation distance are flagged as ineligible initial positions for subsequent strainlines.
-After the first strainline has been obtained and any ineligible local maxima of $\lambda_2$ have been flagged, the next strainline is integrated.
+After the first strainline has been obtained and any ineligible local maxima of $\lambda_2$ have been flagged, the next strainline is integrated starting form the highest eligible maximum of $\lambda_2$.
 The process continues until all eligible local maxima of $\lambda_2$ have been used to initialise strainline integration.
-Finally, any strainline segment contained within (closed) shear LCS are removed.
-These calculations are performed by the functions \lstinline!seed_curves_from_lambda_max!
-and \lstinline!remove_strain_in_shear!.
-
+%Finally, any strainline segment contained within (closed) shear LCS are removed. 
+These calculations are performed by the function \lstinline!seed_curves_from_lambda_max!.
 Note that once a strainline integration has begun, it may progress so as to be within the local minimisation distance of an existing strainline.
@@
 
-The example we present show hyperbolic LCS inside elliptic LCS.
-This appears contradictory at first glance because LCS should not cross, however numerical experimentation with LCS calculations overlaid with colour tracer advection reveals that hyperbolic LCS can cross elliptic barriers.
+A further point important to note is the coexistence of hyperbolic and elliptic LCS. 
+The examples we present along this paper show hyperbolic LCS inside elliptic LCS.
+This appears contradictory at first glance because LCS should not cross, however, numerical experimentation with LCS calculations overlaid with colour tracer advection reveals that hyperbolic LCS can cross elliptic barriers and have an impact on the tracer pattern inside elliptic regions.
+%The crossing of elliptic and hyperbolic LCS is not a contradiction. 
+In the crossing point, a line segment of the elliptic LCS, oriented along $\boldsymbol \eta_\lambda^{\pm}$, would not stretch at all, while a segment of the strainline, oriented along $\boldsymbol \xi_1$, would repel fluid normally at a maximal rate and it would shrink itself at a maximal rate.
+Figure\,\ref{fig:ocean_dataset_colortracer} illustrates the effect of strainlines on tracer particles also inside elliptic regions.
+Colorcoded final positions show large gradients along strainlines, directly showing that particles are strongly repelled by the LCS and end up in different locations.
+
+\begin{figure}[hbt]
+\centering
+\includegraphics[width=0.49\textwidth]{graphics/ocean_dataset/LCS_fwd_coherent_eddy}
+\includegraphics[width=0.49\textwidth]{graphics/ocean_dataset/LCS_fwd_colortracer}
+\caption{(Left) Elliptic LCS (green) with hyperbolic normally repelling LCS (red) inside.
+(Right) Final positions $\boldsymbol x_t = F_{t_0}^t(\boldsymbol x_0)$ of tracer particles are plotted with the x-position encoded in red and y-position encoded in green color.
+The impact of the repulsive property of hyperbolic LCS on the tracer particles can not only be observed outside, but also inside the elliptic region.
+Hyperbolic LCS mark lines of high color gradients, corresponding to tracer particles that start on both sides of the LCS, are repelled, and end up in distinct locations.}
+\label{fig:ocean_dataset_colortracer}
+\end{figure}
 
@@
 
-For the last example, we choose a measured geophysical flow where, in contrast to the previous analytical examples, the two dimensional incompressible velocity field is presented as a dataset with discrete temporal and spatial resolution.
+For the last example, corresponding to the ocean dataset demo files in LCS Tool, we choose a measured geophysical flow.
+In contrast to the previous analytical examples, the two dimensional incompressible velocity field is presented as a dataset with discrete temporal and spatial resolution.
 Our region of interest is a small domain in the South Atlantic Ocean, where exceptionally coherent eddies, Agulhas rings, were recently found using methods provided by LCS Tool \parencite{haller13:_coher_lagran}.
@@
 
-The data is given at a spatial resolution of $1/4\degree$ and temporal resolution of $7\,\mathrm{days}$.
+The data is given at a spatial resolution of $1/4\degree$ and a temporal resolution of $7\,\mathrm{days}$.
 Due to the discrete data, defining the right hand side of Eq.\,(\ref{eq:dtlon}) and (\ref{eq:dtlat}) involves interpolation in space and time with a spline interpolation scheme.
@@
 evaluates the interpolants for the zonal and meridional velocity at the needed coordinates.
-Listing \ref{l:griddedInterp} shows the corresponding commands.
-\begin{lstlisting}[caption={Generate interpolant of ocean velocity data set.},label=l:griddedInterp]
-% Define right hand side of ODE, ocean.flow.derivative
+Listing \ref{l:griddedInterp} shows large parts of the code in LCS Tool's ocean demo file \lstinline!hyperbolic_shear_lcs.m!.
+In lines 8-11 the commands for the interpolation are given.
+\begin{lstlisting}[caption={LCS ocean demo: Parts of the demo script included in LCS Tool to compute LCS from an ocean data set},label=l:griddedInterp]
+%% Input parameters
+domain = [0,6;-34,-28];
+resolution = [400,400];
+timespan = [100,130];
+
+%% Velocity definition
+...
 interpMethod = 'spline';
@@
 vlat_interpolant = griddedInterpolant({time,lat,lon},vlat,interpMethod);
-ocean.flow.derivative = @(t,y,~)flowdata_derivative(t,y,vlon_interpolant,vlat_interpolant);
+lDerivative = @(t,x,~)flowdata_derivative(t,x,vlon_interpolant,vlat_interpolant);
+incompressible = true;
+
+%% LCS parameters
+% Cauchy-Green strain
+cgEigenvalueFromMainGrid = false;
+cgAuxGridRelDelta = 0.01;
+
+% Lambda-lines
+lambda = 1;
+lambdaLineOdeSolverOptions = odeset('relTol',1e-6);
+
+% Strainlines
+strainlineMaxLength = 20;
+gridSpace = diff(domain(1,:))/(double(resolution(1))-1);
+strainlineLocalMaxDistance = 2*gridSpace;
+strainlineOdeSolverOptions = odeset('relTol',1e-4);
+
+% Stretchlines
+stretchlineMaxLength = 20;
+stretchlineLocalMaxDistance = 4*gridSpace;
+stretchlineOdeSolverOptions = odeset('relTol',1e-4);
+...
+%% Cauchy-Green strain eigenvalues and eigenvectors
+[cgEigenvector,cgEigenvalue] = eig_cgStrain(lDerivative,domain,resolution,timespan,'incompressible',incompressible,'eigenvalueFromMainGrid',cgEigenvalueFromMainGrid,'auxGridRelDelta',cgAuxGridRelDelta);
+
+%% Lambda-line LCSs
+% Define Poincare sections; ...
+poincareSection = struct('endPosition',{},'numPoints',{},'orbitMaxLength',{});
+...
+poincareSection(1).endPosition = [3.3,-32.1;3.7,-31.6];
+poincareSection(2).endPosition = [1.3,-30.9;1.9,-31.1];
+
+% Number of orbit seed points along each Poincare section
+[poincareSection.numPoints] = deal(100);
+
+% Set maximum orbit length to twice the expected circumference
+nPoincareSection = numel(poincareSection);
+for i = 1:nPoincareSection
+    rOrbit = hypot(diff(poincareSection(i).endPosition(:,1)),diff(poincareSection(i).endPosition(:,2)));
+    poincareSection(i).orbitMaxLength = 2*(2*pi*rOrbit);
+end
+[shearline.etaPos,shearline.etaNeg] = lambda_line(cgEigenvector,cgEigenvalue,lambda);
+closedLambdaLine = poincare_closed_orbit_multi(domain,resolution,shearline,poincareSection,'odeSolverOptions',lambdaLineOdeSolverOptions);
+...
+%% Hyperbolic strainline LCSs
+strainlineLcs = seed_curves_from_lambda_max(strainlineLocalMaxDistance,strainlineMaxLength,cgEigenvalue(:,2),cgEigenvector(:,1:2),domain,resolution,'odeSolverOptions',strainlineOdeSolverOptions);
+...
+stretchlineLcs = seed_curves_from_lambda_max(stretchlineLocalMaxDistance,stretchlineMaxLength,-cgEigenvalue(:,1),cgEigenvector(:,3:4),domain,resolution,'odeSolverOptions',stretchlineOdeSolverOptions);
+
 \end{lstlisting}
-For the ocean data, the integration time was chosen to be $T=30\,\mathrm{days}$, a time larger than the turn over time of the mesoscale eddies in the domain.
-The resolution of the grid of initial conditions is set to $400\times400$ which corresponds to a resolution of roughly 0.015\degree and gives good results.
+For the ocean data, the integration time was chosen to be $T=30\,\mathrm{days}$ (Listing \ref{l:griddedInterp}, line 4), a time larger than the turn over time of the mesoscale eddies in the domain.
+The resolution of the grid of initial conditions is set to $400\times400$ (line 3) which corresponds to a resolution of roughly 0.015\degree and gives good results.
 With this choice the resolution of the tracer grid is of the order of $15$ times higher than the resolution of the velocity field.
-Figures~\ref{f:ocean dataset hyperbolic shear lcs strainline} and \ref{f:ocean dataset hyperbolic shear lcs stretchline} show elliptic and hyperbolic LCS on 22 November 2006, i.e., the same time instant shown in~\textcite{haller13:_coher_lagran,beron-vera13:_objec_agulh}.
-The analysis with LCS Tool reveals a large coherent eddy bounded by a closed lambda-line at $(3,-32)$ and four further smaller coherent eddy cores.
+The flow is integrated and the Cauchy-Green strain tensor is computed by the function \lstinline!eig_cgStrain!
+(line 35).
+Incompressibility of the flow is enforced (line 12), the auxiliary grid distance is set to $1$\% of the main grid distance (line 17), and eigenvalues are computed from the auxiliary grid (line 16).
+Lambda-line LCS are computed in line 54, after the Poincar\'e section have been set (line 41-42) and the $\eta_\pm$ fields have been defined (line 53).
+Hyperbolic LCS, strainlines and strechlines, are computed in lines 57 and 59, respectively.
+
+Figures~\ref{f:ocean dataset hyperbolic shear lcs details strainline} and \ref{f:ocean dataset hyperbolic shear lcs details stretchline} show elliptic and hyperbolic LCS on 22 November 2006, i.e., the same time instant shown in~\textcite{haller13:_coher_lagran,beron-vera13:_objec_agulh}.
+Note that here the integration time is $T=30\,\mathrm{days}$, as opposed to $90$\,days in their analysis.
+Since for $90$\,days the hyperbolic LCS would be a confusing entanglement of lines, we choose a shorter integration time here, to be able to show hyperbolic LCS along with elliptic LCS.
+For the shorter integration time, the analysis with LCS Tool reveals a large coherent eddy bounded by a closed lambda-line at $(3,-32)$ which corresponds to eddy \#2 in Fig.\,3 of \textcite{beron-vera13:_objec_agulh}.
+Additionally, a further smaller coherent eddy core at $\sim(1,-31)$ is found that does not stay coherent over a time of $90$\, days.
+Strainlines (red) and stretchlines (blue) mark the core lines of strong normal attraction and repulsion.
+
+%\begin{figure}
+%\begin{center}
+%\includegraphics[width=\textwidth]{graphics/ocean_dataset/hyperbolic_shear_lcs_strainline}
+%\end{center}
+%\caption{Ocean dataset lambda-line (green) and strainline (red) LCS.}
+%\label{f:ocean dataset hyperbolic shear lcs strainline}
+%\end{figure}
+
+%\begin{figure}
+%\begin{center}
+%\includegraphics[width=\textwidth]{graphics/ocean_dataset/hyperbolic_shear_lcs_stretchline}
+%\end{center}
+%\caption{Ocean dataset lambda-line (green) and stretchline (blue) LCSs.}
+%\label{f:ocean dataset hyperbolic shear lcs stretchline}
+%\end{figure}
 
@@
 \begin{center}
-\includegraphics[width=\textwidth]{graphics/ocean_dataset/hyperbolic_shear_lcs_strainline}
+\includegraphics[width=0.8\textwidth]{graphics/ocean_dataset/hyperbolic_shearline_lcs_details_strainline}
 \end{center}
-\caption{Ocean dataset lambda-line (green) and strainline (red) LCS.}
-\label{f:ocean dataset hyperbolic shear lcs strainline}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\textwidth]{graphics/ocean_dataset/hyperbolic_shear_lcs_stretchline}
-\end{center}
-\caption{Ocean dataset lambda-line (green) and stretchline (blue) LCSs.}
-\label{f:ocean dataset hyperbolic shear lcs stretchline}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\textwidth]{graphics/ocean_dataset/hyperbolic_shear_lcs_details_strainline}
-\end{center}
-\caption{Same LCS as in Figure~\ref{f:ocean dataset hyperbolic shear lcs strainline}, but with additional details.
+\caption{Lambda-line LCS (green) and strainline LCS (red) of the ocean data set.
 Background gray-scale is the finite-time Lyapunov exponent.
 Dashed green lines are Poincare sections.
-Green curves are closed lambda-lines with thick green curves being outermost closed lambda-lines, i.e.
-lambda-line LCSs.
 White dots indicate $\lambda_2$ local maxima used for strainline integration initial positions.}
@@
 \begin{center}
-\includegraphics[width=\textwidth]{graphics/ocean_dataset/hyperbolic_shear_lcs_details_stretchline}
+\includegraphics[width=0.8\textwidth]{graphics/ocean_dataset/hyperbolic_shearline_lcs_details_stretchline.eps}
 \end{center}
-\caption{Same LCS as in Figure~\ref{f:ocean dataset hyperbolic shear lcs stretchline}, but with additional details.
+\caption{Lambda-line LCS (green) and stretchline LCS (blue) of the ocean data set.
 Background gray-scale is the finite-time Lyapunov exponent.
 Dashed green lines are Poincare sections.
-Green curves are closed lambda-lines and thick green curves are outermost closed lambda-lines, i.e.
-lambda-line LCSs.
-White dots indicate $\lambda_1$ local minima used for stretchline integration intial positions.}
+White dots indicate $\lambda_1$ local minima used for stretchline integration initial positions.}
 \label{f:ocean dataset hyperbolic shear lcs details stretchline}

### 2014-01-31 main.tex
Correct some typos

@@
 \usepackage[space]{grffile}
+\usepackage{lineno}
 %% hyperref manual advises loading this package last
@@
 \newcommand{\FIXME}[1]{\\\fbox{FIXME: #1}\\}
+
+\linenumbers
 
@@
 
-The second step is to integrate the lambda-lines to obtain the Poincare return map.
-Integration of lambda-lines is performed with the $\eta_+$ and $\eta_-$ vector fields defined on the main grid.
-These vector fields are derived from eigenvalues and eigenvectors as described in~\textcite{farazmand12:_comput_lagran}.
-The eigenvector fields have orientation discontinuities, therefore the vector field must be locally reoriented for the integration.
+The second step is to integrate the lambda-lines to obtain the Poincare return map. Integration of lambda-lines is performed with the $\eta_+$ and $\eta_-$ vector fields defined on the main grid. 
+%These vector fields are derived from eigenvalues and eigenvectors as described in~\textcite{farazmand12:_comput_lagran}.
+0The eigenvector fields have orientation discontinuities, therefore the vector field must be locally reoriented for the integration.
 The algorithm used to perform variable step-size integration in vector fields with orientation discontinuities is enumerated in Table~\ref{t:variable step integration} and illustrated in Figure~\ref{f:variable step integration}.
@@
 Additionally, a further smaller coherent eddy core at $\sim(1,-31)$ is found that does not stay coherent over a time of $90$\, days.
+Both closed orbits are stable if the resolution for the Cauchy-Green strain tensor field is increased.
 Strainlines (red) and stretchlines (blue) mark the core lines of strong normal attraction and repulsion.

### 2014-01-31 main.tex
Remove graphics filename extension from includegraphics command
Minor modifications to main.tex

--- 2014-01-31 main.tex
@@
 The process continues until all eligible local maxima of $\lambda_2$ have been used to initialise strainline integration.
-%Finally, any strainline segment contained within (closed) shear LCS are removed. 
 These calculations are performed by the function \lstinline!seed_curves_from_lambda_max!.
@@
 For the shorter integration time, the analysis with LCS Tool reveals a large coherent eddy bounded by a closed lambda-line at $(3,-32)$ which corresponds to eddy \#2 in Fig.\,3 of \textcite{beron-vera13:_objec_agulh}.
-Additionally, a further smaller coherent eddy core at $\sim(1,-31)$ is found that does not stay coherent over a time of $90$\, days.
+Additionally, a further smaller coherent eddy core at $\sim(1,-31)$ is found that does not stay coherent over a time of 90 days.
+Both closed orbits are stable if the resolution for the Cauchy-Green strain tensor field is increased.
 Strainlines (red) and stretchlines (blue) mark the core lines of strong normal attraction and repulsion.
-
-%\begin{figure}
-%\begin{center}
-%\includegraphics[width=\textwidth]{graphics/ocean_dataset/hyperbolic_shear_lcs_strainline}
-%\end{center}
-%\caption{Ocean dataset lambda-line (green) and strainline (red) LCS.}
-%\label{f:ocean dataset hyperbolic shear lcs strainline}
-%\end{figure}
-
-%\begin{figure}
-%\begin{center}
-%\includegraphics[width=\textwidth]{graphics/ocean_dataset/hyperbolic_shear_lcs_stretchline}
-%\end{center}
-%\caption{Ocean dataset lambda-line (green) and stretchline (blue) LCSs.}
-%\label{f:ocean dataset hyperbolic shear lcs stretchline}
-%\end{figure}
 
@@
 \begin{center}
-\includegraphics[width=0.8\textwidth]{graphics/ocean_dataset/hyperbolic_shearline_lcs_details_stretchline.eps}
+\includegraphics[width=0.8\textwidth]{graphics/ocean_dataset/hyperbolic_shearline_lcs_details_stretchline}
 \end{center}
@@
 \end{figure}
+
+\clearpage
 

### 2014-01-31 graphics/double_gyre/ftle_.m
Remove FTLE figure from double gyre example
The FTLE is moved to the lambda-line convergence figure as a
background.

@@ -1,38 +1,0 @@
-function ftle_
-
-%% Input parameters
-epsilon = .1;
-amplitude = .1;
-omega = pi/5;
-domain = [0,2;0,1];
-resolutionX = 500;
-timespan = [0,5];
-
-%% Velocity definition
-lDerivative = @(t,x,~)derivative(t,x,false,epsilon,amplitude,omega);
-incompressible = true;
-
-%% LCS parameters
-cgStrainOdeSolverOptions = odeset('relTol',1e-5);
-
-% Make x and y grid spacing as equal as possible
-gridSpace = diff(domain(1,:))/(double(resolutionX)-1);
-resolutionY = round(diff(domain(2,:))/gridSpace);
-resolution = [resolutionX,resolutionY];
-    
-hAxes = setup_figure(domain);
-title(hAxes,['Resolution: ',num2str(resolution(1)),'\times',num2str(resolution(2))])
-    
-%% Cauchy-Green strain eigenvalues and eigenvectors
-s = warning('off','eig_cgStrain:unequalDelta');
-cgEigenvalue = eig_cgStrain(lDerivative,domain,resolution,timespan,'incompressible',incompressible,'odeSolverOptions',cgStrainOdeSolverOptions);
-warning(s)
-
-% Plot finite-time Lyapunov exponent
-cgEigenvalue2 = reshape(cgEigenvalue(:,2),fliplr(resolution));
-ftle_ = ftle(cgEigenvalue2,diff(timespan));
-plot_ftle(hAxes,domain,resolution,ftle_);
-colormap(hAxes,flipud(gray))
-delete(get(hAxes,'title'))
-hFigure = get(hAxes,'parent');
-print_pdf(hFigure,'ftle.pdf')

### 2014-01-31 graphics/double_gyre/lambda_lcs_convergence.m

@@ -1,6 +1,6 @@
 function lambda_lcs_convergence
 
-%% Input parameters
+% Input parameters
 epsilon = .1;
 amplitude = .1;
 omega = pi/5;
@@ -8,11 +8,11 @@ domain = [0,2;0,1];
 resolutionX = 500:250:1000;
 timespan = [0,5];
 
-%% Velocity definition
+% Velocity definition
 lDerivative = @(t,x,~)derivative(t,x,false,epsilon,amplitude,omega);
 incompressible = true;
 
-%% LCS parameters
+% LCS parameters
 cgStrainOdeSolverOptions = odeset('relTol',1e-5);
 
 % Lambda-lines
@@ -41,17 +41,25 @@ for m = 1:numel(resolutionX)
     gridSpace = diff(domain(1,:))/(double(lResolutionX)-1);
     resolutionY = round(diff(domain(2,:))/gridSpace);
     resolution = [lResolutionX,resolutionY];
-    disp(['Resolution: ',num2str(resolution)])
     
     hAxes = setup_figure(domain);
+    hFigure = get(hAxes,'parent');
     title(hAxes,['Resolution: ',num2str(resolution(1)),'\times',num2str(resolution(2))])
     
-    %% Cauchy-Green strain eigenvalues and eigenvectors
+    % Cauchy-Green strain eigenvalues and eigenvectors
     s = warning('off','eig_cgStrain:unequalDelta');
     [cgEigenvector,cgEigenvalue] = eig_cgStrain(lDerivative,domain,resolution,timespan,'incompressible',incompressible,'odeSolverOptions',cgStrainOdeSolverOptions);
     warning(s)
     
-    %% Lambda-line LCSs
+    % Plot finite-time Lyapunov exponent
+    cgEigenvalue2 = reshape(cgEigenvalue(:,2),fliplr(resolution));
+    ftle_ = ftle(cgEigenvalue2,diff(timespan));
+    plot_ftle(hAxes,domain,resolution,ftle_);
+    colormap(hAxes,flipud(gray))
+    delete(findobj(hFigure,'Type','axes','Tag','Colorbar'))
+    drawnow
+    
+    % Lambda-line LCSs
     [shearline.etaPos,shearline.etaNeg] = lambda_line(cgEigenvector,cgEigenvalue,lambda);
     warnState = warning('off','integrate_line:isDiscontinuousLargeAngle');
[…]

### 2014-01-31 main.tex
Remove FTLE figure from double gyre example
The FTLE is moved to the lambda-line convergence figure as a
background.
Minor revisions throughout article

@@
 LCSs are used in fluid dynamics to capture essential flow features and thereby simplify the description of complex fluid flows.
-More specifically, LCSs are particularly well adapted to analyze two dimensional finite time turbulent flows.
+More specifically, we analyze two dimensional finite time turbulent flows to detect LCSs.
 LCSs identify transport barriers, repelling and attracting structures and vortex boundaries.
@@
 The finite-time Lyapunov exponent (FTLE) gives one widely-used mathematical definition for LCSs.
-Although the FTLE is popular, it is possible to find examples where it fails to identify LCSs correctly\parencite{haller11:_lagran_coher_struc,norgard12:_secon_lagran_coher_struc}.
-To remedy these problems, variational calculus principles applied to material curves in a fluid present new approaches to identifying LCSs\parencite{haller11:_lagran_coher_struc,farazmand12:_comput_lagran,haller12:_geodes_theor_trans_barrier_two_dimen_flows,farazmand13:_attrac_lagran,haller13:_coher_lagran}.
-This article presents an algorithmic introduction to variational LCSs using a computational engine called LCS Tool\footnote{LCS Tool is free MATLAB software available for download at: \url{https://github.com/jeixav/LCS-Tool}}.
+Although the FTLE is widely used, it is possible to find examples where it fails to identify LCSs correctly\parencite{haller11:_lagran_coher_struc,norgard12:_secon_lagran_coher_struc}.
+To remedy these problems, variational calculus applied to material curves in a fluid provides a rigorous approach for LCS detection\parencite{haller11:_lagran_coher_struc,farazmand12:_comput_lagran,haller12:_geodes_theor_trans_barrier_two_dimen_flows,farazmand13:_attrac_lagran,haller13:_coher_lagran}.
+This article presents an algorithmic introduction to variational LCSs supported by a computational engine called LCS Tool\footnote{LCS Tool is available at: \url{https://github.com/jeixav/LCS-Tool}.}.
 
-LCS Tool is a library of MATLAB functions that perform the computations to identify LCSs in two dimensional finite time flows.
-LCS Tool aims to initiate scientists to LCSs.
-Existing LCS software programs are:
+LCS Tool is a library of MATLAB functions that perform computations to identify LCSs in two dimensional finite time flows.
+LCS Tool aims to help scientists perform variational LCS calculations.
+Other LCS software has been developed before; the list includes:
 \begin{itemize}
@@
 \end{itemize}
-These packages are based on the FTLE; the objective of LCS Tool and this article is to encourage and facilitate variational LCS analysis.
+These packages are based on the FTLE; the focus of LCS Tool and this article is to encourage and facilitate variational LCS analysis.
 
@@
 \]
-The time interval used to detect LCSs is left as a choice; it can be the complete interval for which the flow is defined, or data is available, or it can be a chosen subset.
+The time interval used to detect LCSs is arbitrary; it can be the complete interval for which the flow is defined, or data is available, or it can be a sub-interval.
 
@@
 \[
-C_{t_0}^t(\boldsymbol x_0) = \left[\nabla F_{t_0}^t(\boldsymbol x_0)\right]^T \nabla F_{t_0}^t(\boldsymbol x_0).
+C_{t_0}^t(\boldsymbol x_0) = \left[D F_{t_0}^t(\boldsymbol x_0)\right]^T D F_{t_0}^t(\boldsymbol x_0).
 \]
@@
 For $\lambda=1$, $\boldsymbol \eta^{\lambda}_\pm$ points in the direction of maximal Lagrangian shear and we obtain uniformly non-stretching curves over the finite time interval $[t_0, t]$.
-In summary, the boundaries of coherent Lagrangian vortices are closed curves that uniformly stretch by a factor $\lambda$ close to unity when advected by the flow.
-We call them lambda-lines.
+In summary, the boundaries of coherent Lagrangian vortices are closed curves that uniformly stretch by a factor $\lambda$ when advected by the flow.
+We refer to these curves as lambda-lines.
 The outermost lambda-line is the physically observable boundary of an eddy, generating the tracer pattern of an isolated rotating fluid region.
 
-Note that $\boldsymbol \eta^{\lambda}_\pm$ follows from the variational principle that the first variation of the averaged Lagrangian strain along the searched closed curve vanishes (see \parencite{haller13:_coher_lagran} for full derivations):
+Note that the equation for $\boldsymbol \eta^{\lambda}_\pm$ follows from the variational principle that the first variation of the averaged Lagrangian strain along the sought extremal closed curve vanishes (see \parencite{haller13:_coher_lagran} for full derivations):
 \begin{eqnarray*}
@@
 \end{eqnarray*}
-This coherence principle means that the searched curve $\gamma$, the coherent eddy boundary, is a stationary point of the averaged Lagrangian strain functional and therefore nearby fluid elements cannot break away from the curve $\gamma$.
+This coherence principle means the sought curve $\gamma$, the coherent eddy boundary, is a stationary point of the averaged Lagrangian strain functional and therefore nearby fluid elements cannot separate from $\gamma$.
 
 \subsection{Hyperbolic LCS}
-Hyperbolic LCS are trajectories of the two eigenvector fields $\boldsymbol \xi_{1,2}$ of the Cauchy-Green strain tensor $C_{t_0}^t$.
-Repelling hyperbolic LCSs, denoted as strainlines, are obtained as solutions of the differential equation
-\[
+Hyperbolic LCSs are trajectories of the eigenvector fields $\boldsymbol \xi_{1,2}$. Repelling hyperbolic LCSs, denoted strainlines, are obtained as solutions of the differential equation
+\begin{equation}
 \boldsymbol r_1' = \boldsymbol \xi_1(\boldsymbol r),
-\]
-and attracting hyperbolic LCSs, denoted as stretchlines, are obtained as a solution of the differential equation
+\label{eq:strainline}
+\end{equation}
+and attracting hyperbolic LCSs, denoted stretchlines, are obtained as a solution of the differential equation
 \[
@@
 \]
-A normal perturbation to the strainline segment $\boldsymbol r_1'$ grows under the flow map by a factor $\lambda_2$, i.e., fluid elements are repelled by that rate in normal direction.
-Equivalently, a normal perturbation to the stretchline segment $\boldsymbol r_2'$ shrinks with factor $\lambda_1$ and fluid elements are attracted.
+A normal perturbation to the strainline segment $\boldsymbol r_1'$ grows under the flow map by a factor $\lambda_2$, i.e., fluid elements are repelled at that rate in the normal direction.
+Equivalently, a normal perturbation to the stretchline segment $\boldsymbol r_2'$ shrinks by a factor $\lambda_1$ and fluid elements are attracted.
 
@@
 \[
-P(\gamma) = \frac{1}{\sigma} \int_0^\sigma p(s) \text{d}s.
+P(\gamma) = \frac{1}{\sigma} \int_0^\sigma p(s) \text{d}s,
 \]
@@
 
-This section presents the numerical methods used to compute LCS and explains them step by step. In other words, the algorithms that constitute LCS Tool are presented.
+This section presents the numerical methods used to compute LCSs and explains them step-wise. In other words, the algorithms that constitute LCS Tool are presented.
 
@@
 
-To obtain Lagrangian coherent structures the Cauchy-Green strain tensor must be calculated first.
+To obtain Lagrangian coherent structures the Cauchy-Green strain tensor must be calculated.
 More specifically, the eigenvalues and eigenvectors of the Cauchy-Green strain tensor are needed.
 The function associated with this calculation in LCS Tool is \lstinline!eig_cgStrain!.
-The main steps of this function are enumerated in Table~\ref{t:Cauchy-Green algorithm} and are described in detail in the following part.
-Table~\ref{t:eig_cgStrain syntax} summarises the function syntax.
+The main steps of this function are enumerated in Table~\ref{t:Cauchy-Green algorithm} and are described in detail below.
+Table~\ref{t:eig_cgStrain syntax} summarises the syntax of \lstinline!eig_cgStrain!.
 
@@
 \hline
-\lstinline!odeSolverOptions! & optional \lstinline!odeset! structure to specify flow map integration options\\
+\lstinline!odeSolverOptions! & optional \lstinline!odeset! structure to specify flow map integration parameters\\
 \hline \hline
@@
 We can find the optimal resolution by an iterative procedure in which we start with a low resolution and double it until convergence of LCSs is observed visually.
-If the chosen domain only comprises few expected LCS, e.g., one vortex, a resolution of about 500 grid point along the longest axis usually gives good results.
+If the chosen domain only comprises a few expected LCSs, e.g., one vortex, a resolution of about 500 grid point along the longest axis usually gives good results.
 Otherwise, for larger domains, a higher resolution must be chosen.
@@
 The improvements the auxiliary grid yields compared to using only the main grid were reported in \textcite{farazmand12:_comput_lagran}.
-Experience indicates that optimal results are obtained when the auxiliary grid spacing is between 1\% and 10\% of the main grid spacing.
+Experience indicates that optimal results are obtained when the auxiliary grid spacing is 1\% to 10\% of the main grid spacing.
 
@@
 The function \lstinline!eig_cgStrain!
-provides the option to calculate the eigenvector from the auxiliary grid, but the eigenvalues from the main grid.
+provides the option to calculate eigenvectors from the auxiliary grid, but eigenvalues from the main grid.
 We have found that for flows defined analytically the eigenvalues should be calculated from the main grid whereas for the flows defined by datasets, using the auxiliary grid gives better results.
@@
 For a square domain, this implies the velocity equations must be integrated, after including 4 auxiliary grid points around every main grid point, at 1.25 million grid points.
-The ``natural'' way to perform these integration calculations would be to solve a two dimensional first-order ODE system at every grid point and iterate over every grid point within a for-loop.
+The natural way to perform these integration calculations would be to solve a two dimensional first-order ODE system at every grid point and iterate over every grid point in a for-loop.
 With 1.25 million grid points this calculation can take an excessively long time.
@@
 function is used to perform the integration and the flow integration typically takes five to ten minutes with default error tolerances.
-The drawback of vector form integration is its memory requirements can become excessive at high resolutions.
-Furthermore, writing the velocity function in vector form is more error-prone than the intrinsic two dimensional form.
+The drawback of vector form integration is that memory requirements can become excessive at high resolutions.
+Furthermore, writing the velocity function in vector form is more error-prone than the simpler two dimensional form.
 
 The Cauchy-Green strain tensor can have singularities, i.e., $C_{t_0}^t(\boldsymbol x_0) = I$.
-Such points are isolated in general\parencite{delmarcelle94}, therefore no special accommodations for are made in computations for singularities of the Cauchy-Green strain tensor.
+Such points are isolated in general\parencite{delmarcelle94}, therefore no special accommodations are made in computations for singularities of the Cauchy-Green strain tensor.
 
@@
 
-Flows that are incompressible, meaning $\nabla \cdot \boldsymbol v = 0$, satisfy the relation $\lambda_1(\boldsymbol x_0) \lambda_2(\boldsymbol x_0) = 1, \forall \: \boldsymbol x_0 \: \in \: U$\parencite{arnold78:_mathem}.
+Flows that are incompressible, meaning $D \cdot \boldsymbol v = 0$, satisfy the relation $\lambda_1(\boldsymbol x_0) \lambda_2(\boldsymbol x_0) = 1, \forall \: \boldsymbol x_0 \: \in \: U$\parencite{arnold78:_mathem}.
 This gives the possibility of imposing incompressibility when calculating the eigenvalues of the Cauchy-Green strain tensor by first calculating $\lambda_2$ and then setting $\lambda_1(\boldsymbol x_0) = 1/\lambda_2(\boldsymbol x_0)$.
-This procedure should be applied carefully since by changing $\lambda_1$ but not $\xi_1$, the error in satisfying the eigenvalue equation for the Cauchy-Green strain tensor may increase.
+This procedure should be applied carefully since by changing $\lambda_1$ but not $\boldsymbol \xi_1$, the error in satisfying the eigenvalue equation for the Cauchy-Green strain tensor may increase.
 
@@
 Therefore the function \lstinline!eig_cgStrain!
-records the number of points at which incompressibility cannot be imposed and this measure can also help to set appropriate integration tolerances.
+records the number of points at which incompressibility cannot be imposed and this measure can guide setting appropriate integration tolerances.
 
@@
 An interpolation function must be written to enable evaluating the velocity function at arbitrary points in $U$ and at arbitrary times between $t_-$ and $t_+$.
-MATLAB's interpolation functions can be integrated into the velocity function.
-As with analytic cases, care should be taken to ensure the interpolation embedded in the flow velocity integration function works correctly when performing integration in vector form.
-In section \ref{sec:oceandataset}, we present an example how LCS are computed for an ocean dataset and details of the interpolation function are given.
+MATLAB's interpolation functions can be used in the velocity function.
+In section \ref{sec:oceandataset} we present an example for LCSs computed for an ocean dataset and details of the interpolation function are given.
 
@@
 Shear LCSs, or coherent Lagrangian vortices, are closed orbits in the Lagrangian shear vector fields $\boldsymbol \eta_{\pm}$ defined in Equation~\eqref{eq:etafields}.
-Since $\lambda = 1$, the closed curves are non-stretching and area preserving.
-We find closed orbits by integrating lambda-lines starting from a Poincare section and evaluating the first return map.
-A lambda-line closing onto itself in the starting point is a shear LCS.
-The main steps of the algorithm used to calculate shear LCS are enumerated in Table~\ref{t:Shear LCS algorithm} and described in detail in the following part.
-The syntax of shear LCS functions in LCS Tool can be seen in Table~\ref{t:Shear LCS functions}.
+We find closed orbits by integrating lambda-lines starting from a Poincare section and evaluating the return map.
+A lambda-line returning to its starting point is a shear LCS.
+The main steps of the algorithm used to calculate shear LCS are enumerated in Table~\ref{t:Shear LCS algorithm} and described in detail below.
+The syntax of shear LCS functions in LCS Tool is in Table~\ref{t:Shear LCS functions}.
 
@@
 \begin{enumerate}
-\item Position Poincare sections in flow domain as initial positions of lambda-lines
-\item Integrate lambda-lines tangent to $\boldsymbol \eta_{\pm}$
+\item Position Poincare sections in flow domain to specify initial positions of lambda-lines
+\item Integrate lambda-lines tangent to $\boldsymbol \eta_{\pm}$ (see Equation~\eqref{eq:etafields}.)
 \item Calculate Poincare return map
-\item Find closed orbits in Poincare return map
+\item Find closed orbits positions from Poincare return map
 \item Identify outermost closed orbit
@@
 \hline
-\lstinline!odeSolverOptions! & optional \lstinline!odeset! structure to specify lambda-line integration options\\
+\lstinline!odeSolverOptions! & optional \lstinline!odeset! structure to specify lambda-line integration parameters\\
 \hline
@@
 
-The first step is to set the position of Poincare sections \lstinline!poincareSection.endPosition!
-that are straight lines in regions where closed lambda-lines are expected.
-The Poincare section must be oriented in a way that the first endpoint is close to the centre of the expected eddy and the second endpoint is outside the expected eddy region.
-There is no predetermined procedure for deciding where the Poincare sections should be positioned.
-Possibilities are an examination of the finite-time Lyapunov exponent field of the flow, or some prior knowledge of where coherent eddies could be expected.
-Additionally, the number of lambda-lines launched from the Poincare section \lstinline!poincareSection.numPoints!
-must be defined.
-A good default value is $100$.
+The first step is to set the position of Poincare sections as straight lines in regions where closed lambda-lines are expected.
+The Poincare section must be oriented such that the first endpoint is close to the centre of the expected eddy and the second endpoint is outside the expected eddy.
+There is no definite procedure for deciding the position of Poincare sections.
+Possibilities include an examination of the finite-time Lyapunov exponent field of the flow, or some prior knowledge of where coherent eddies are be expected.
+Additionally, the number of lambda-lines launched from the Poincare section, \lstinline!poincareSection.numPoints!, must be defined.
+A good default is 100.
 
 The second step is to integrate the lambda-lines to obtain the Poincare return map.
-Integration of lambda-lines is performed with the $\eta_+$ and $\eta_-$ vector fields defined on the main grid.
+Integration of lambda-lines is performed with the $\boldsymbol \eta_+$ and $\boldsymbol \eta_-$ vector fields defined on the main grid.
 These vector fields are derived from eigenvalues and eigenvectors as described in~\textcite{farazmand12:_comput_lagran}.
@@
 Then, rotations exceeding 90\degree\, are classified as orientation discontinuities and are corrected before linear interpolation.
-Thus when setting the Cauchy-Green strain tensor main grid resolution, a histogram of eigenvector field rotations may be produced to ensure that all rotations are below 90\degree\, or close to 180\degree.
+Thus when setting the Cauchy-Green strain tensor main grid resolution, a histogram of eigenvector field rotations may be calculated to ensure all rotations are well below 90\degree\, or almost 180\degree.
 
@@
 An example of a Poincare return map produced from integration of the $\boldsymbol \eta_\pm$ field is shown in Figure~\ref{f:Poincare return map}.
-Well behaved orbits will return to the Poincare section and their integration will be stopped using an ODE event detection function.
+Well behaved orbits will return to the Poincare section and their integration will be stopped using ODE event detection.
 Some orbits may however deviate far from the Poincare section and their integration could take inordinately long.
-To prevent this behaviour, a maximum orbit length, \lstinline!poincareSection.orbitMaxLength!, is specified.
-In practice, viewing the Poincare section as the radius of a circle and setting the maximum lambda-line integration length to twice the circumference gives acceptable results.
+To limit this behaviour, a maximum orbit length, \lstinline!poincareSection.orbitMaxLength!, is specified.
+In practice, viewing the Poincare section as the radius of a circle and setting the maximum lambda-line integration length to twice the circumference gives good results.
 In Figure~\ref{f:Poincare return map}, circle markers indicate closed orbit positions.
 Closed orbits exist where the Poincare return map, the distance between the final and the initial point of the orbit $P(s)-s$, is zero.
-The function \lstinline!poincare_closed_orbit!
+The function \lstinline!poincare_closed_orbit_multi!
 performs the computations.
@@
 If after a predetermined number of iterations, \lstinline!nBisection!
-whose default value is five, the two points around the zero crossing have absolute values of more \lstinline!dThresh!, the zero crossing is discarded.
-In Figure~\ref{f:Poincare return map}, discarded zero crossings are seen where $0 < s < 0.05$.
-Once all valid closed lambda-line orbits have been located, the outermost orbit of every Poincare section is deemed a shear LCS.
+whose default value is five, the two points around the zero crossing have absolute values above \lstinline!dThresh!, the zero crossing is discarded.
+Once all valid closed lambda-line orbits have been located, the outermost orbit of every Poincare section is identified as a shear LCS.
 
@@
 
-The main steps of the algorithm used to calculate hyperbolic LCS are enumerated in Table~\ref{t:Hyperbolic LCS algorithm}.
+The main steps of the algorithm used to calculate hyperbolic LCSs are enumerated in Table~\ref{t:Hyperbolic LCS algorithm}.
 The main function to compute hyperbolic LCS in LCS Tool is \lstinline!seed_curves_from_lambda_max!
@@
 \item Define a maximum strainline length
-\item Integrate a strainline forward and backward, using the largest $\lambda_2$ local maximum as the initial position.
+\item Integrate a strainline forward and backward according to Equation~\eqref{eq:strainline} and using the largest $\lambda_2$ local maximum as the initial position.
 Integrate until the strainline has attained the maximum strainline length, or until it has reached the domain boundary.
@@
 \item Continue integrating strainlines using local maxima of $\lambda_2$ as initial positions until no eligible local maxima of $\lambda_2$ remain.
-\item Remove any strainline segment contained within (closed) shear LCS
 \end{enumerate}
-\caption{Algorithm to calculate hyperbolic LCS}
+\caption{Algorithm to calculate strainline hyperbolic LCSs. Algorithm for stretchline hyperbolic LCSs is similar.}
 \label{t:Hyperbolic LCS algorithm}
@@
 A further point important to note is the coexistence of hyperbolic and elliptic LCS.
-The examples we present along this paper show hyperbolic LCS inside elliptic LCS.
-This appears contradictory at first glance because LCS should not cross, however, numerical experimentation with LCS calculations overlaid with colour tracer advection reveals that hyperbolic LCS can cross elliptic barriers and have an impact on the tracer pattern inside elliptic regions.
-%The crossing of elliptic and hyperbolic LCS is not a contradiction. 
-In the crossing point, a line segment of the elliptic LCS, oriented along $\boldsymbol \eta_\lambda^{\pm}$, would not stretch at all, while a segment of the strainline, oriented along $\boldsymbol \xi_1$, would repel fluid normally at a maximal rate and it would shrink itself at a maximal rate.
-Figure\,\ref{fig:ocean_dataset_colortracer} illustrates the effect of strainlines on tracer particles also inside elliptic regions.
-Colorcoded final positions show large gradients along strainlines, directly showing that particles are strongly repelled by the LCS and end up in different locations.
+The examples we present show hyperbolic LCSs inside elliptic LCSs.
+This may seem contradictory when accustomed to LCSs not crossing, however, numerical experimentation with LCS calculations overlaid with colour tracer advection reveals that hyperbolic LCS can cross elliptic barriers and have an impact on the tracer pattern inside elliptic regions.
+At crossing points a line segment of an elliptic LCS, oriented along $\boldsymbol \eta_\lambda^{\pm}$, does not stretch, while a segment of a strainline, oriented along $\boldsymbol \xi_1$, repels fluid normally at a maximal rate and shrinks at a maximal rate.
+Figure~\ref{fig:ocean_dataset_colortracer} illustrates the effect of strainlines on tracer particles inside elliptic regions.
+Colour-coded final positions show large gradients along strainlines, directly showing that particles are strongly repelled by the LCS and end in different positions.
 
@@
 
-\section{Example}
-In this section we present three examples of how to use LCS-Tool to obtain LCSs in different flows, namley the double gyre flow, the Bickley jet flow, and an oceanic geostrophic flow.
-All three examples are available as demo files in the demo folder of LCS-Tool and the user can follow the computation of LCS step by step.
+\section{Examples}
+In this section we present examples showing how to use LCS Tool to obtain LCSs in different flows, namely a double gyre, a jet, and an oceanic geostrophic flow.
+All three examples are available as demo files in the demo folder of LCS Tool and the user can follow the computation of LCS step by step.
 
@@
 
-In what follows, the parameter values are: $A = 0.1$, $\epsilon = 0.1$, $\omega = \pi/5$.
-The flow timespan is $t \in [0,5]$ and the domain is $x \in [0,2]$, $y \in [0,1]$.
-In Figure~\ref{f:double gyre FTLE}, we show the FTLE at a resolution of $500\times250$.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\textwidth]{graphics/double_gyre/ftle}
-\end{center}
-\caption{Double gyre FTLE.}
-\label{f:double gyre FTLE}
-\end{figure}
+In what follows, the parameter values are: $A = 0.1$, $\epsilon = 0.1$, $\omega = \pi/5$. The flow timespan is $t \in [0,5]$ and the domain is $x \in [0,2]$, $y \in [0,1]$.
 
@@
 \includegraphics[width=.8\textwidth]{graphics/double_gyre/lambda_lcs_convergence_1000}
-  \caption{Double gyre closed lambda-line at varying resolutions}
+  \caption{Double gyre closed lambda-lines at varying resolutions with FTLE background}
   \label{fig:double_gyre_lambda_lcs_convergence}
@@
 The latter figure demonstrates a poor result since the LCSs change shape significantly from one eddy to the next.
-To address this problem we could try to adjust numerical parameters such as the resolution, but with our hardware memory availability was reached before achieving convergence.
+To address this problem we could try to adjust numerical parameters such as the resolution, but with our computer hardware memory availability was reached before achieving convergence.
 To circumvent this problem, we analyse lambda-line closed orbits of one vortex for a short timespan.
@@
 The script produces the data shown in Figure~\ref{fig:Bickley jet lambda lcs convergence}.
-Parameter values used may read from the script.
+Parameter values used may be read from the script.
 It is worth pointing out that in order to obtain closed orbits, $\lambda = 0.995$; with $\lambda = 1$ no closed orbits are found.
@@
 For the last example, corresponding to the ocean dataset demo files in LCS Tool, we choose a measured geophysical flow.
-In contrast to the previous analytical examples, the two dimensional incompressible velocity field is presented as a dataset with discrete temporal and spatial resolution.
+In contrast to the previous analytic examples, the two dimensional incompressible velocity field is presented as a dataset with discrete temporal and spatial resolution.
 Our region of interest is a small domain in the South Atlantic Ocean, where exceptionally coherent eddies, Agulhas rings, were recently found using methods provided by LCS Tool \parencite{haller13:_coher_lagran}.
@@
 \end{eqnarray}
-where $g$ is the gravitational acceleration, $R$ is the mean radius of Earth, and $f(\theta):=2\Omega\sin\theta$ is the Coriolis parameter with $\Omega$ being the Earth's mean angular velocity.
+where $g$ is the gravitational acceleration, $R$ is the mean radius of Earth, and $f(\theta) \equiv 2\Omega\sin\theta$ is the Coriolis parameter with $\Omega$ being the Earth's mean angular velocity.
 
@@
 In lines 8-11 the commands for the interpolation are given.
+
 \begin{lstlisting}[caption={LCS ocean demo: Parts of the demo script included in LCS Tool to compute LCS from an ocean data set},label=l:griddedInterp]
-%% Input parameters
+% Input parameters
 domain = [0,6;-34,-28];
@@
 
-%% Velocity definition
+% Velocity definition
 ...
@@
 
-%% LCS parameters
+% LCS parameters
 % Cauchy-Green strain
@@
 ...
-%% Cauchy-Green strain eigenvalues and eigenvectors
+% Cauchy-Green strain eigenvalues and eigenvectors
 [cgEigenvector,cgEigenvalue] = eig_cgStrain(lDerivative,domain,resolution,timespan,'incompressible',incompressible,'eigenvalueFromMainGrid',cgEigenvalueFromMainGrid,'auxGridRelDelta',cgAuxGridRelDelta);
 
-%% Lambda-line LCSs
+% Lambda-line LCSs
 % Define Poincare sections; ...
@@
 ...
-%% Hyperbolic strainline LCSs
+% Hyperbolic strainline LCSs
 strainlineLcs = seed_curves_from_lambda_max(strainlineLocalMaxDistance,strainlineMaxLength,cgEigenvalue(:,2),cgEigenvector(:,1:2),domain,resolution,'odeSolverOptions',strainlineOdeSolverOptions);
@@
 stretchlineLcs = seed_curves_from_lambda_max(stretchlineLocalMaxDistance,stretchlineMaxLength,-cgEigenvalue(:,1),cgEigenvector(:,3:4),domain,resolution,'odeSolverOptions',stretchlineOdeSolverOptions);
+\end{lstlisting}
 
-\end{lstlisting}
-For the ocean data, the integration time was chosen to be $T=30\,\mathrm{days}$ (Listing \ref{l:griddedInterp}, line 4), a time larger than the turn over time of the mesoscale eddies in the domain.
-The resolution of the grid of initial conditions is set to $400\times400$ (line 3) which corresponds to a resolution of roughly 0.015\degree and gives good results.
-With this choice the resolution of the tracer grid is of the order of $15$ times higher than the resolution of the velocity field.
+For the ocean data, the integration time was chosen to be $T=30$ days (Listing \ref{l:griddedInterp}, line 4), a time larger than the turn over time of the mesoscale eddies in the domain.
+The resolution of the grid of initial conditions is set to $400\times400$ (line 3) which corresponds to a resolution of roughly 0.015\degree\, and gives good results.
+With this choice the resolution of the tracer grid is of the order of 15 times higher than the resolution of the velocity field.
 The flow is integrated and the Cauchy-Green strain tensor is computed by the function \lstinline!eig_cgStrain!
 (line 35).
-Incompressibility of the flow is enforced (line 12), the auxiliary grid distance is set to $1$\% of the main grid distance (line 17), and eigenvalues are computed from the auxiliary grid (line 16).
-Lambda-line LCS are computed in line 54, after the Poincar\'e section have been set (line 41-42) and the $\eta_\pm$ fields have been defined (line 53).
+Incompressibility of the flow is enforced (line 12), the auxiliary grid distance is set to 1\% of the main grid distance (line 17), and eigenvalues are computed from the auxiliary grid (line 16).
+Lambda-line LCS are computed in line 54, after the Poincar\'e section have been set (line 41-42) and the $\boldsymbol\eta_\pm$ fields have been defined (line 53).
 Hyperbolic LCS, strainlines and strechlines, are computed in lines 57 and 59, respectively.
@@
 Figures~\ref{f:ocean dataset hyperbolic shear lcs details strainline} and \ref{f:ocean dataset hyperbolic shear lcs details stretchline} show elliptic and hyperbolic LCS on 22 November 2006, i.e., the same time instant shown in~\textcite{haller13:_coher_lagran,beron-vera13:_objec_agulh}.
-Note that here the integration time is $T=30\,\mathrm{days}$, as opposed to $90$\,days in their analysis.
-Since for $90$\,days the hyperbolic LCS would be a confusing entanglement of lines, we choose a shorter integration time here, to be able to show hyperbolic LCS along with elliptic LCS.
-For the shorter integration time, the analysis with LCS Tool reveals a large coherent eddy bounded by a closed lambda-line at $(3,-32)$ which corresponds to eddy \#2 in Fig.\,3 of \textcite{beron-vera13:_objec_agulh}.
-Additionally, a further smaller coherent eddy core at $\sim(1,-31)$ is found that does not stay coherent over a time of 90 days.
+Note that here the integration time is $T=30$ days, as opposed to 90 days in their analysis.
+Since for 90 days the hyperbolic LCSs would be a confusing entanglement of lines, we choose a shorter integration time here, to show hyperbolic LCS together with elliptic LCSs.
+For the shorter integration time, the analysis with LCS Tool reveals a large coherent eddy bounded by a closed lambda-line at $(3,-32)$ which corresponds to eddy \#2 in Figure~3 of \textcite{beron-vera13:_objec_agulh}.
+Additionally, a further smaller coherent eddy core at approximately $(1,-31)$ is found that does not stay coherent over a time of 90 days.
 Both closed orbits are stable if the resolution for the Cauchy-Green strain tensor field is increased.
@@
 The software library provides functions to enable exploring variational LCS methods without requiring detailed knowledge of variational LCS theory.
-These functions make it possible to write compact programs identify and plot LCSs.
+These functions make it possible to write compact programs to identify and visualise LCSs.
 
-LCS Tool is built atop MATLAB\parencite{mathworks13:_matlab} and leverages that software's capabilities extensively.
-With FTLE-based LCS identification, computational performance has received considerable attention\parencite{conti12:_gpu_apu_finit_time_lyapun_expon,miron12:_anisot_lagran_coher_struc}.
+LCS Tool is built atop MATLAB and leverages that software's capabilities extensively.
+With FTLE-based LCSs, computational performance has received considerable attention\parencite{conti12:_gpu_apu_finit_time_lyapun_expon,miron12:_anisot_lagran_coher_struc}.
 LCS Tool is free software and we hope it will evolve to incorporate similar advances.
@@
 % End:
-% LocalWords:  stretchline stretchlines incompressible gyre eigenvector eigenvectors incompressibility dataset datasets LCS strainline Lyapunov Bickley Rossby timespan variational adaptively vortices parametrise advected strainlines minima
+% LocalWords:  stretchline stretchlines incompressible gyre eigenvector eigenvectors incompressibility dataset datasets LCS strainline Lyapunov Bickley Rossby timespan variational adaptively vortices parametrise advected strainlines minima extremal

### 2014-01-31 listings/bickley_jet/lambda_lcs_convergence_subdomain.m

@@ -1,4 +1,4 @@
-%% Input parameters
+% Input parameters
 u = 62.66;
 lengthX = pi*earthRadius;
 lengthY = 1.77e6;
@@ -7,7 +7,7 @@ domain = [2e6,lengthX/2;[-1,.25]*2.25*lengthY];
 resolutionX = 400:100:600;
 timespan = [0,2*lengthX/u];
 
-%% Velocity definition
+% Velocity definition
 perturbationCase = 3;
 phiTimespan = [0,25];
 phiInitial = [0,0];
@@ -18,7 +18,7 @@ phi1Max = max(phi1);
 lDerivative = @(t,x,~)derivative(t,x,false,u,lengthX,lengthY,epsilon,perturbationCase,phiSol,phi1Max);
 incompressible = true;
 
-%% LCS parameters
+% LCS parameters
 % Cauchy-Green strain
 cgStrainOdeSolverOptions = odeset('relTol',1e-4);
 
@@ -39,10 +39,10 @@ for m = 1:numel(resolutionX)
     resolutionY = round(diff(domain(2,:))/gridSpace) + 1;
     resolution = [lResolutionX,resolutionY];
    
-    %% Cauchy-Green strain eigenvalues and eigenvectors
+    % Cauchy-Green strain eigenvalues and eigenvectors
     [cgEigenvector,cgEigenvalue] = eig_cgStrain(lDerivative,domain,resolution,timespan,'incompressible',incompressible,'odeSolverOptions',cgStrainOdeSolverOptions);
     
-    %% Lambda-line LCSs
+    % Lambda-line LCSs
     [shearline.etaPos,shearline.etaNeg] = lambda_line(cgEigenvector,cgEigenvalue,lambda);
     [closedLambdaLine,~,hFigure] = poincare_closed_orbit_multi(domain,resolution,shearline,poincareSection,'odeSolverOptions',lambdaLineOdeSolverOptions,'dThresh',dThresh,'showGraph',true);
 end

### 2014-01-31 listings/bickley_jet/strain_lcs.m

@@ -1,4 +1,4 @@
-%% Input parameters
+% Input parameters
 u = 62.66;
 lengthX = pi*earthRadius;
 lengthY = 1.77e6;
@@ -7,7 +7,7 @@ domain = [2e6,.55*lengthX;[-1,.25]*2.25*lengthY];
 resolutionX = 500;
 timespan = [0,2*lengthX/u];
 
-%% Velocity definition
+% Velocity definition
 perturbationCase = 3;
 phiTimespan = [0,25];
 phiInitial = [0,0];
@@ -18,7 +18,7 @@ phi1Max = max(phi1);
 lDerivative = @(t,x,~)derivative(t,x,false,u,lengthX,lengthY,epsilon,perturbationCase,phiSol,phi1Max);
 incompressible = true;
 
-%% LCS parameters
+% LCS parameters
 % Cauchy-Green strain
 cgStrainOdeSolverOptions = odeset('relTol',1e-4);
 
@@ -33,8 +33,8 @@ resolution = [resolutionX,resolutionY];
 
 strainlineLocalMaxDistance = 4*gridSpace;
 
-%% Cauchy-Green strain eigenvalues and eigenvectors
+% Cauchy-Green strain eigenvalues and eigenvectors
 [cgEigenvector,cgEigenvalue] = eig_cgStrain(lDerivative,domain,resolution,timespan,'incompressible',incompressible,'odeSolverOptions',cgStrainOdeSolverOptions);
 
-%% Strainline LCSs
+% Strainline LCSs
 strainlineLcs = seed_curves_from_lambda_max(strainlineLocalMaxDistance,strainlineMaxLength,cgEigenvalue(:,2),cgEigenvector(:,1:2),domain,resolution,'odeSolverOptions',strainlineOdeSolverOptions);

### 2014-01-31 listings/double_gyre/lambda_lcs.m

@@ -1,4 +1,4 @@
-%% Input parameters
+% Input parameters
 epsilon = .1;
 amplitude = .1;
 omega = pi/5;
@@ -6,11 +6,11 @@ domain = [0,2;0,1];
 resolution = [500,250];
 timespan = [0,5];
 
-%% Velocity definition
+% Velocity definition
 lDerivative = @(t,x,~)derivative(t,x,false,epsilon,amplitude,omega);
 incompressible = true;
 
-%% LCS parameters
+% LCS parameters
 cgStrainOdeSolverOptions = odeset('relTol',1e-5);
 
 % Lambda-lines
@@ -30,9 +30,9 @@ for i = 1:nPoincareSection
     poincareSection(i).orbitMaxLength = 2*(2*pi*rOrbit);
 end
 
-%% Cauchy-Green strain eigenvalues and eigenvectors
+% Cauchy-Green strain eigenvalues and eigenvectors
 [cgEigenvector,cgEigenvalue] = eig_cgStrain(lDerivative,domain,resolution,timespan,'incompressible',incompressible,'odeSolverOptions',cgStrainOdeSolverOptions);
 
-%% Lambda-line LCSs
+% Lambda-line LCSs
 [shearline.etaPos,shearline.etaNeg] = lambda_line(cgEigenvector,cgEigenvalue,lambda);
 closedLambdaLine = poincare_closed_orbit_multi(domain,resolution,shearline,poincareSection,'odeSolverOptions',lambdaLineOdeSolverOptions);

### 2014-01-31 listings/double_gyre/lambda_strain_stretch_lcs.m

@@ -1,11 +1,11 @@
-%% Strainlines LCSs
+% Strainlines LCSs
 strainlineMaxLength = 20;
 gridSpace = diff(domain(1,:))/(double(resolution(1))-1);
 strainlineLocalMaxDistance = 2*gridSpace;
 
 strainlineLcs = seed_curves_from_lambda_max(strainlineLocalMaxDistance,strainlineMaxLength,cgEigenvalue(:,2),cgEigenvector(:,1:2),domain,resolution);
 
-%% Stretchlines LCSs
+% Stretchlines LCSs
 stretchlineMaxLength = 20;
 stretchlineLocalMaxDistance = 10*gridSpace;
 

### 2014-02-04 main.tex
Revisions to abstract, introduction and Bickley jet section

@@
 
-\title{An Algorithmic Introduction to Lagrangian Coherent Structures}
+\title{LCS Tool — An Algorithmic Introduction to Lagrangian Coherent 
+Structures}
 
@@
 \abstract{
-We give an applied introduction to Lagrangian coherent structure (LCS) detection.
-We review LCS theory that is based on variational calculus, present algorithms conceived in accordance with the variational theory and present a MATLAB computational engine called LCS Tool that enables variational LCS calculations and detection.
-We use LCS Tool to analyse three flow examples: a double gyre, a Bickley jet and an ocean dataset.
+We give an applied introduction to Lagrangian coherent structures (LCSs) using a computational engine we have programmed called LCS Tool.
+First, we review LCS theory that is based on variational calculus; this introduces elliptic and hyperbolic variational LCSs.
+Then we turn our attention to algorithms for computations conceived in accordance with the variational LCS theory.
+These algorithms are the foundation of LCS Tool.
+We conclude by explaining, with demonstration scripts, how to use LCS Tool to analyse three flow examples: a double gyre, a Bickley jet and an ocean dataset.
 }
@@
 LCSs are used in fluid dynamics to capture essential flow features and thereby simplify the description of complex fluid flows.
-More specifically, we analyze two dimensional finite time turbulent flows to detect LCSs.
+Here we focus on analysing two dimensional finite time turbulent flows.
 LCSs identify transport barriers, repelling and attracting structures and vortex boundaries.
-Application areas for LCSs are oceanic and atmospheric flows, biological applications\parencite{wilson09:_lagran_reynol,tallapragada11:_lagran}, aeronautics\parencite{tang10:_accur_lagran_hong_kong_inter_airpor} and mechanical systems\parencite{hadjighasem13:_detec_kam}.
+Application areas for LCSs include oceanic and atmospheric flows, biological applications\parencite{wilson09:_lagran_reynol,tallapragada11:_lagran}, aeronautics\parencite{tang10:_accur_lagran_hong_kong_inter_airpor} and mechanical systems\parencite{hadjighasem13:_detec_kam}.
 
-The finite-time Lyapunov exponent (FTLE) gives one widely-used mathematical definition for LCSs.
-Although the FTLE is widely used, it is possible to find examples where it fails to identify LCSs correctly\parencite{haller11:_lagran_coher_struc,norgard12:_secon_lagran_coher_struc}.
-To remedy these problems, variational calculus applied to material curves in a fluid provides a rigorous approach for LCS detection\parencite{haller11:_lagran_coher_struc,farazmand12:_comput_lagran,haller12:_geodes_theor_trans_barrier_two_dimen_flows,farazmand13:_attrac_lagran,haller13:_coher_lagran}.
+The finite time Lyapunov exponent (FTLE) is one widely-used mathematical definition for LCSs.
+Despite its widespread use, examples where the FTLE fails to identify LCSs correctly are known\parencite{haller11:_lagran_coher_struc,norgard12:_secon_lagran_coher_struc}.
+To remedy these problems, variational calculus has been applied to material curve properties in the context of fluid mechanics.
+Thus we obtain a rigorous approach to LCS detection\parencite{haller11:_lagran_coher_struc,farazmand12:_comput_lagran,haller12:_geodes_theor_trans_barrier_two_dimen_flows,farazmand13:_attrac_lagran,haller13:_coher_lagran}.
 This article presents an algorithmic introduction to variational LCSs supported by a computational engine called LCS Tool\footnote{LCS Tool is available at: \url{https://github.com/jeixav/LCS-Tool}.}.
@@
 LCS Tool aims to help scientists perform variational LCS calculations.
+The examples we present have been programmed into demonstration scripts.
+We encourage readers to download and execute these scripts after reading this article.
+
 Other LCS software has been developed before; the list includes:
@@
 There is no definite procedure for deciding the position of Poincare sections.
-Possibilities include an examination of the finite-time Lyapunov exponent field of the flow, or some prior knowledge of where coherent eddies are be expected.
+Possibilities include an examination of the finite time Lyapunov exponent field of the flow, or some prior knowledge of where coherent eddies are be expected.
 Additionally, the number of lambda-lines launched from the Poincare section, \lstinline!poincareSection.numPoints!, must be defined.
@@
 In this section we present examples showing how to use LCS Tool to obtain LCSs in different flows, namely a double gyre, a jet, and an oceanic geostrophic flow.
-All three examples are available as demo files in the demo folder of LCS Tool and the user can follow the computation of LCS step by step.
+All three examples are available as demonstration files in the demo folder of LCS Tool.
+By executing these scripts, readers can follow LCS Tool computations stepwise.
 
@@
 
-\lstinputlisting[caption={LCS Tool commands for strainline and stretchline LCSs},label=lst:double_gyre_lambda_strain_stretch_lcs]{listings/double_gyre/lambda_strain_stretch_lcs.m}
+\lstinputlisting[caption={LCS Tool commands for double gyre strainline and stretchline LCSs.
+Condensed version of the LCS Tool script \lstinline!demo/double_gyre/hyperbolic_shear_lcs.m!
+},label=lst:double_gyre_lambda_strain_stretch_lcs]{listings/double_gyre/lambda_strain_stretch_lcs.m}
 
@@
 
-Figure~\ref{fig:Bickley_jet_FTLE} shows the overall appearance of the Bickley jet and in Figure~\ref{fig:Bickley_jet_lambda_LCS_full} lambda-line LCSs are shown.
-The latter figure demonstrates a poor result since the LCSs change shape significantly from one eddy to the next.
-To address this problem we could try to adjust numerical parameters such as the resolution, but with our computer hardware memory availability was reached before achieving convergence.
-To circumvent this problem, we analyse lambda-line closed orbits of one vortex for a short timespan.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=.85\textwidth]{graphics/bickley_jet/ftle_overview}
-  \caption{Bickley jet FTLE}
-  \label{fig:Bickley_jet_FTLE}
-\end{figure}
+Figure~\ref{fig:Bickley_jet_lambda_LCS_full} shows lambda-line LCS of the Bickley jet with the FTLE in the background.
+The figure demonstrates a poor result since the LCSs change shape significantly between eddies.
+To address this problem we could adjust numerical parameters such as the resolution, but instead we will analyse lambda-line closed orbits of one vortex for a short timespan.
 
@@
 
-The script used to analyse lambda-line LCS is given in Listing~\ref{lst:Bickley jet lambda lcs subdomain}.
-The script produces the data shown in Figure~\ref{fig:Bickley jet lambda lcs convergence}.
-Parameter values used may be read from the script.
-It is worth pointing out that in order to obtain closed orbits, $\lambda = 0.995$; with $\lambda = 1$ no closed orbits are found.
+Figure ~\ref{fig:Bickley jet lambda lcs convergence} shows convergence of the lambda-line elliptic LCSs.
+The LCS Tool script used to produce that figure are based on the script \lstinline!demo/bickley_jet/hyperbolic_shear_lcs.m!
+which is given in  Listing~\ref{lst:Bickley jet hyperbolic_shear_lcs subset}.
+It is worth pointing out that in order to obtain closed orbits, $\lambda = 0.995$; with $\lambda = 1$ no closed orbits exist.
 
-\lstinputlisting[caption={Bickley jet script for lambda-line LCSs},label=lst:Bickley jet lambda lcs subdomain]{listings/bickley_jet/lambda_lcs_convergence_subdomain.m}
+\lstinputlisting[caption={Bickley jet script for hyperbolic and shear LCSs.
+Subset of \lstinline!demo/bickley_jet/hyperbolic_shear_lcs.m!},label=lst:Bickley jet hyperbolic_shear_lcs subset]{listings/bickley_jet/hyperbolic_shear_lcs_subset.m}
 
@@
 
-To calculate strainline and stretchline LCSs we set the timespan to $t \in [0,2L_x/U]$.
-The script used to detect strainlines is given in Listing~\ref{lst:bickley_jet_strain_lcs} and the results are shown in Figure~\ref{fig:Bickley_jet_strain_lcs}.
-Stretchline LCSs are shown in Figure~\ref{fig:Bickley_jet_stretch_lcs}.
-
-\lstinputlisting[caption={LCS Tool commands for Bickley jet strainline LCSs},label=lst:bickley_jet_strain_lcs]{listings/bickley_jet/strain_lcs.m}
+Strainline LCSs are shown in Figure~\ref{fig:Bickley_jet_strain_lcs} and stretchline LCSs are shown in Figure~\ref{fig:Bickley_jet_stretch_lcs}.
 
@@
 \end{center}
-\caption{Lambda-line LCS (green) and strainline LCS (red) of the ocean data set.
-Background gray-scale is the finite-time Lyapunov exponent.
+\caption{Lambda-line LCSs (green) and strainline LCSs (red) of the ocean data set.
+Background gray-scale is the finite time Lyapunov exponent.
 Dashed green lines are Poincare sections.
@@
 \end{center}
-\caption{Lambda-line LCS (green) and stretchline LCS (blue) of the ocean data set.
-Background gray-scale is the finite-time Lyapunov exponent.
+\caption{Lambda-line LCSs (green) and stretchline LCSs (blue) of the ocean data set.
+Background gray-scale is the finite time Lyapunov exponent.
 Dashed green lines are Poincare sections.
@@
 
-Algorithms to identify LCSs using variational principles have been presented, and a software library, LCS Tool, that implements these algorithms has been demonstrated on geophysical flow models and a dataset.
+Algorithms to identify LCSs using variational principles have been presented, and a software library, LCS Tool, that implements these algorithms has been introduced and demonstrated on analytic flow models and a geophysical dataset.
 The software library provides functions to enable exploring variational LCS methods without requiring detailed knowledge of variational LCS theory.
 These functions make it possible to write compact programs to identify and visualise LCSs.
+Since LCS Tool is programmed in MATLAB, testing new methods and visualising their results can be done quite quickly.
 
@@
 
-% Reduce number of user inputs, for example, automated closed orbit detection, which is especially important for flows with a large number of vortices possible.
-
-In addition to improving computational performance, we hope LCS Tool will be the foundation to integrate new theoretical advances, for example the variational LCS jet core identification method\parencite{farazmand13:_shearless} and the
-analysis of LCSs in higher dimensions\parencite{blazevski:_hyper_ellip_trans_barrier_three}.
+In addition to improving computational performance, we hope LCS Tool will be the foundation to integrate new theoretical advances, for example the variational LCS jet core identification method\parencite{farazmand13:_shearless} and the analysis of LCSs in higher dimensions\parencite{blazevski:_hyper_ellip_trans_barrier_three}.
 

### 2014-02-04 graphics/ocean_dataset/hyperbolic_shear_lcs_details_pdf.m

@@ -1,21 +1,0 @@
-function hyperbolic_shear_lcs_details_pdf
-
-hyperbolic_shear_lcs_details
-
-hFigure = 1;
-filename = 'hyperbolic_shear_lcs_details_strainline';
-delete_title(hFigure)
-print_pdf(hFigure,filename)
-
-% FIXME hFigure should not be hard-coded; should probably search for
-% matching title string
-hFigure = 12;
-filename = 'hyperbolic_shear_lcs_details_stretchline';
-delete_title(hFigure)
-print_pdf(hFigure,filename)
-
-function delete_title(hFigure)
-
-hAxes = findobj(hFigure,'type','axes','-not','Tag','Colorbar');
-hTitle = get(hAxes,'title');
-delete(hTitle)

### 2014-02-04 graphics/ocean_dataset/hyperbolic_shear_lcs_pdf.m

@@ -1,19 +1,0 @@
-function hyperbolic_shear_lcs_pdf
-
-hyperbolic_shear_lcs
-
-hFigure = 1;
-filename = 'hyperbolic_shear_lcs_strainline';
-delete_title(hFigure)
-print_pdf(hFigure,filename)
-
-hFigure = 2;
-filename = 'hyperbolic_shear_lcs_stretchline';
-delete_title(hFigure)
-print_pdf(hFigure,filename)
-
-function delete_title(hFigure)
-
-hAxes = get(hFigure,'children');
-hTitle = get(hAxes,'title');
-delete(hTitle)

### 2014-02-04 listings/bickley_jet/lambda_lcs_convergence_subdomain.m

@@ -1,48 +1,0 @@
-% Input parameters
-u = 62.66;
-lengthX = pi*earthRadius;
-lengthY = 1.77e6;
-epsilon = [.075,.4,.3];
-domain = [2e6,lengthX/2;[-1,.25]*2.25*lengthY];
-resolutionX = 400:100:600;
-timespan = [0,2*lengthX/u];
-
-% Velocity definition
-perturbationCase = 3;
-phiTimespan = [0,25];
-phiInitial = [0,0];
-phiSol = ode45(@d_phi,phiTimespan,phiInitial);
-timeResolution = 1e5;
-phi1 = deval(phiSol,linspace(phiTimespan(1),phiTimespan(2),timeResolution),1);
-phi1Max = max(phi1);
-lDerivative = @(t,x,~)derivative(t,x,false,u,lengthX,lengthY,epsilon,perturbationCase,phiSol,phi1Max);
-incompressible = true;
-
-% LCS parameters
-% Cauchy-Green strain
-cgStrainOdeSolverOptions = odeset('relTol',1e-4);
-
-% Lambda-lines
-lambda = .995;
-lambdaLineOdeSolverOptions = odeset('relTol',1e-6);
-poincareSection.endPosition = [6.5,-1.4;4.5,-3.5]*1e6;
-[poincareSection.numPoints] = deal(100);
-rOrbit = hypot(diff(poincareSection.endPosition(:,1)),diff(poincareSection.endPosition(:,2)));
-poincareSection.orbitMaxLength = 2*(2*pi*rOrbit);
-lambdaLineColor = [0,.6,0];
-dThresh = 1e-4;
-
-for m = 1:numel(resolutionX)
-    % Make x and y grid spacing as equal as possible
-    lResolutionX = resolutionX(m);
-    gridSpace = diff(domain(1,:))/(double(lResolutionX)-1);
-    resolutionY = round(diff(domain(2,:))/gridSpace) + 1;
-    resolution = [lResolutionX,resolutionY];
-   
-    % Cauchy-Green strain eigenvalues and eigenvectors
-    [cgEigenvector,cgEigenvalue] = eig_cgStrain(lDerivative,domain,resolution,timespan,'incompressible',incompressible,'odeSolverOptions',cgStrainOdeSolverOptions);
-    
-    % Lambda-line LCSs
-    [shearline.etaPos,shearline.etaNeg] = lambda_line(cgEigenvector,cgEigenvalue,lambda);
[…]

### 2014-02-04 listings/bickley_jet/strain_lcs.m

@@ -1,40 +1,0 @@
-% Input parameters
-u = 62.66;
-lengthX = pi*earthRadius;
-lengthY = 1.77e6;
-epsilon = [.075,.4,.3];
-domain = [2e6,.55*lengthX;[-1,.25]*2.25*lengthY];
-resolutionX = 500;
-timespan = [0,2*lengthX/u];
-
-% Velocity definition
-perturbationCase = 3;
-phiTimespan = [0,25];
-phiInitial = [0,0];
-phiSol = ode45(@d_phi,phiTimespan,phiInitial);
-timeResolution = 1e5;
-phi1 = deval(phiSol,linspace(phiTimespan(1),phiTimespan(2),timeResolution),1);
-phi1Max = max(phi1);
-lDerivative = @(t,x,~)derivative(t,x,false,u,lengthX,lengthY,epsilon,perturbationCase,phiSol,phi1Max);
-incompressible = true;
-
-% LCS parameters
-% Cauchy-Green strain
-cgStrainOdeSolverOptions = odeset('relTol',1e-4);
-
-% Strainlines
-strainlineMaxLength = 1e8;
-strainlineOdeSolverOptions = odeset('relTol',1e-4);
-
-% Make x and y grid spacing as equal as possible
-gridSpace = diff(domain(1,:))/(double(resolutionX)-1);
-resolutionY = round(diff(domain(2,:))/gridSpace) + 1;
-resolution = [resolutionX,resolutionY];
-
-strainlineLocalMaxDistance = 4*gridSpace;
-
-% Cauchy-Green strain eigenvalues and eigenvectors
-[cgEigenvector,cgEigenvalue] = eig_cgStrain(lDerivative,domain,resolution,timespan,'incompressible',incompressible,'odeSolverOptions',cgStrainOdeSolverOptions);
-
-% Strainline LCSs
-strainlineLcs = seed_curves_from_lambda_max(strainlineLocalMaxDistance,strainlineMaxLength,cgEigenvalue(:,2),cgEigenvector(:,1:2),domain,resolution,'odeSolverOptions',strainlineOdeSolverOptions);

### 2014-02-05 main.tex
Change spelling to US English

@@
 These algorithms are the foundation of LCS Tool.
-We conclude by explaining, with demonstration scripts, how to use LCS Tool to analyse three flow examples: a double gyre, a Bickley jet and an ocean dataset.
+We conclude by explaining, with demonstration scripts, how to use LCS Tool to analyze three flow examples: a double gyre, a Bickley jet and an ocean dataset.
 }
@@
 LCSs are used in fluid dynamics to capture essential flow features and thereby simplify the description of complex fluid flows.
-Here we focus on analysing two dimensional finite time turbulent flows.
+Here we focus on analyzing two dimensional finite time turbulent flows.
 LCSs identify transport barriers, repelling and attracting structures and vortex boundaries.
@@
 which are a function of the eigenvalues and eigenvectors of $C_{t_0}^t(\boldsymbol x_0)$.
-Curves $\gamma$ that are trajectories of $\boldsymbol \eta^{\lambda}_\pm$ can be parametrized as $\boldsymbol r(s)$ with their arclength $s \in [0,\sigma]$.
+Curves $\gamma$ that are trajectories of $\boldsymbol \eta^{\lambda}_\pm$ can be parametrized as $\boldsymbol r(s)$ with their arc length $s \in [0,\sigma]$.
 They are given by integrating
@@
 
-Similarly to ellitpic LCS, hyperbolic LCS are also based on a variational principle, in this case for the Lagrangian shear
+Similarly to elliptic LCS, hyperbolic LCS are also based on a variational principle, in this case for the Lagrangian shear
 \[
@@
 \]
-They are one type of shearless LCS. For details of the derivation of hyperbolic LCS, see \cite{farazmand13:_shearless}.
+They are one type of shear-less LCS. For details of the derivation of hyperbolic LCS, see \cite{farazmand13:_shearless}.
 
@@
 The main steps of this function are enumerated in Table~\ref{t:Cauchy-Green algorithm} and are described in detail below.
-Table~\ref{t:eig_cgStrain syntax} summarises the syntax of \lstinline!eig_cgStrain!.
+Table~\ref{t:eig_cgStrain syntax} summarizes the syntax of \lstinline!eig_cgStrain!.
 
@@
 The first step is to set the position of Poincare sections as straight lines in regions where closed lambda-lines are expected.
-The Poincare section must be oriented such that the first endpoint is close to the centre of the expected eddy and the second endpoint is outside the expected eddy.
+The Poincare section must be oriented such that the first endpoint is close to the center of the expected eddy and the second endpoint is outside the expected eddy.
 There is no definite procedure for deciding the position of Poincare sections.
@@
 At the initial point, there is an orientation discontinuity at the lower right grid point that must be corrected prior to linear interpolation.
-At point \#1, no orientation discontinuties are present.
+At point \#1, no orientation discontinuities are present.
 At point \#2 the interpolated $\eta_\pm$ vector must be rotated by 180\degree\,to match the orientation of the trajectory.}
@@
 Some orbits may however deviate far from the Poincare section and their integration could take inordinately long.
-To limit this behaviour, a maximum orbit length, \lstinline!poincareSection.orbitMaxLength!, is specified.
+To limit this behavior, a maximum orbit length, \lstinline!poincareSection.orbitMaxLength!, is specified.
 In practice, viewing the Poincare section as the radius of a circle and setting the maximum lambda-line integration length to twice the circumference gives good results.
@@
 \begin{enumerate}
-\item Define a local maximisation distance.
-\item Find all local maxima of $\lambda_2$ on the main grid such that they are maxima within a circle with a radius equal to the local minimisation distance.
+\item Define a local maximization distance.
+\item Find all local maxima of $\lambda_2$ on the main grid such that they are maxima within a circle with a radius equal to the local minimization distance.
 \item Define a maximum strainline length
@@
 Integrate until the strainline has attained the maximum strainline length, or until it has reached the domain boundary.
-\item Flag any remaining local maxima of $\lambda_2$ within the minimisation distance of the strainline as ineligible initial positions for subsequent strainlines
+\item Flag any remaining local maxima of $\lambda_2$ within the minimization distance of the strainline as ineligible initial positions for subsequent strainlines
 \item Continue integrating strainlines using local maxima of $\lambda_2$ as initial positions until no eligible local maxima of $\lambda_2$ remain.
@@
 Strainlines, trajectories of the $\boldsymbol \xi_1$ vector field, are integrated starting from local maxima of $\lambda_2$ on the main grid.
-The local maximisation distance \lstinline!distance!
+The local maximization distance \lstinline!distance!
 is defined.
 This distance is set to a value between two and five grid points, since it represents the number of grid points over which the Cauchy-Green strain tensor is susceptible to numerical noise from integration.
-Thus, the local maximisation distance is expected to be a function of variable time step integration tolerances.
-A record is made of all grid points where $\lambda_2$ is maximal within a circle whose radius equals the defined maximisation distance.
+Thus, the local maximization distance is expected to be a function of variable time step integration tolerances.
+A record is made of all grid points where $\lambda_2$ is maximal within a circle whose radius equals the defined maximization distance.
 Then the first strainline is integrated from the largest of all the local maxima in both directions, i.e., integrating $\boldsymbol \xi_1$ and $-\boldsymbol \xi_1$.
@@
 The maximum length may be seen as safety factor to prevent highly sinuous strainlines leading to inordinately long computations; the maximum length should be set such that most strainlines stop at the domain boundary.
-Once a strainline position has been calculated, local maxima of $\lambda_2$ within the minimisation distance are flagged as ineligible initial positions for subsequent strainlines.
+Once a strainline position has been calculated, local maxima of $\lambda_2$ within the minimization distance are flagged as ineligible initial positions for subsequent strainlines.
 After the first strainline has been obtained and any ineligible local maxima of $\lambda_2$ have been flagged, the next strainline is integrated starting form the highest eligible maximum of $\lambda_2$.
-The process continues until all eligible local maxima of $\lambda_2$ have been used to initialise strainline integration.
+The process continues until all eligible local maxima of $\lambda_2$ have been used to initialize strainline integration.
 These calculations are performed by the function \lstinline!seed_curves_from_lambda_max!.
-Note that once a strainline integration has begun, it may progress so as to be within the local minimisation distance of an existing strainline.
-In other words, the local minimisation distance controls where strainlines begin, but not where they stop.
+Note that once a strainline integration has begun, it may progress so as to be within the local minimization distance of an existing strainline.
+In other words, the local minimization distance controls where strainlines begin, but not where they stop.
 
@@
 The examples we present show hyperbolic LCSs inside elliptic LCSs.
-This may seem contradictory when accustomed to LCSs not crossing, however, numerical experimentation with LCS calculations overlaid with colour tracer advection reveals that hyperbolic LCS can cross elliptic barriers and have an impact on the tracer pattern inside elliptic regions.
+This may seem contradictory when accustomed to LCSs not crossing, however, numerical experimentation with LCS calculations overlaid with color tracer advection reveals that hyperbolic LCS can cross elliptic barriers and have an impact on the tracer pattern inside elliptic regions.
 At crossing points a line segment of an elliptic LCS, oriented along $\boldsymbol \eta_\lambda^{\pm}$, does not stretch, while a segment of a strainline, oriented along $\boldsymbol \xi_1$, repels fluid normally at a maximal rate and shrinks at a maximal rate.
 Figure~\ref{fig:ocean_dataset_colortracer} illustrates the effect of strainlines on tracer particles inside elliptic regions.
-Colour-coded final positions show large gradients along strainlines, directly showing that particles are strongly repelled by the LCS and end in different positions.
+Color-coded final positions show large gradients along strainlines, directly showing that particles are strongly repelled by the LCS and end in different positions.
 
@@
 All three examples are available as demonstration files in the demo folder of LCS Tool.
-By executing these scripts, readers can follow LCS Tool computations stepwise.
+By executing these scripts, readers can follow LCS Tool computations step-wise.
 
@@
 The double gyre is a model for a time-dependent two gyre system observed in geophysical flows \parencite{shadden05:_defin_lagran_lyapun}.
-It consists of two counterrotating sinusoidal vortices with a harmonically oscillating line in-between.
+It consists of two counter rotating sinusoidal vortices with a harmonically oscillating line in-between.
 The velocity equations are:
@@
 
-\begin{lstlisting}[caption={Double gyre derivative function corresponding to Equations~\ref{eq:double gyre derivative equations}.},label=l:double gyre derivative]
+\begin{lstlisting}[caption={Double gyre derivative function corresponding to Equations~\eqref{eq:double gyre derivative equations}.},label=l:double gyre derivative]
 function derivative_ = derivative(t,x,epsilon,amplitude,omega)
@@
 Figures~\ref{fig:double_gyre_lambda_strain_lcs} and~\ref{fig:double_gyre_lambda_stretch_lcs} show the results and Listing~\ref{lst:double_gyre_lambda_strain_stretch_lcs} gives the LCS Tool commands.
-The local maximisation distance is set larger for stretchlines than for strainlines usually.
+The local maximization distance is set larger for stretchlines than for strainlines usually.
 This is because strainlines follow ridges of $\lambda_2$ maxima whereas stretchlines do not follow ridges of $\lambda_1$ minima.
@@
 The Bickley jet models a meandering zonal jet flanked above and below by
-counterrotating vortices. This is an idealised model of geophysical flows
+counter rotating vortices. This is an idealized model of geophysical flows
 such as the polar night jet perturbed by a Rossby
@@
 The figure demonstrates a poor result since the LCSs change shape significantly between eddies.
-To address this problem we could adjust numerical parameters such as the resolution, but instead we will analyse lambda-line closed orbits of one vortex for a short timespan.
+To address this problem we could adjust numerical parameters such as the resolution, but instead we will analyze lambda-line closed orbits of one vortex for a short timespan.
 
@@
 Our region of interest is a small domain in the South Atlantic Ocean, where exceptionally coherent eddies, Agulhas rings, were recently found using methods provided by LCS Tool \parencite{haller13:_coher_lagran}.
-The oceanic velocity field is obtained from altimetric satelite measurements of the sea surface height under the assumption that the flow is geostrophic.
-Under this assumption the sea surface height $\eta$ serves as a streamfunction and in a longitude-latitude $(\varphi,\theta)$ coordinate system the evolution of a fluid particle is given by
+The oceanic velocity field is obtained from altimetric satellite measurements of the sea surface height under the assumption that the flow is geostrophic.
+Under this assumption the sea surface height $\eta$ serves as a stream-function and in a longitude-latitude $(\varphi,\theta)$ coordinate system the evolution of a fluid particle is given by
 \begin{eqnarray}
@@
 The data is given at a spatial resolution of $1/4\degree$ and a temporal resolution of $7\,\mathrm{days}$.
-Due to the discrete data, defining the right hand side of Eq.\,(\ref{eq:dtlon}) and (\ref{eq:dtlat}) involves interpolation in space and time with a spline interpolation scheme.
+Due to the discrete data, defining the right hand side of Equation~\eqref{eq:dtlon} and~\eqref{eq:dtlat} involves interpolation in space and time with a spline interpolation scheme.
 An interpolant is generated first, then the function \lstinline!flowdata_derivative!
@@
 Incompressibility of the flow is enforced (line 12), the auxiliary grid distance is set to 1\% of the main grid distance (line 17), and eigenvalues are computed from the auxiliary grid (line 16).
-Lambda-line LCS are computed in line 54, after the Poincar\'e section have been set (line 41-42) and the $\boldsymbol\eta_\pm$ fields have been defined (line 53).
+Lambda-line LCS are computed in line 54, after the Poincare section have been set (line 41-42) and the $\boldsymbol\eta_\pm$ fields have been defined (line 53).
 Hyperbolic LCS, strainlines and strechlines, are computed in lines 57 and 59, respectively.
@@
 The software library provides functions to enable exploring variational LCS methods without requiring detailed knowledge of variational LCS theory.
-These functions make it possible to write compact programs to identify and visualise LCSs.
-Since LCS Tool is programmed in MATLAB, testing new methods and visualising their results can be done quite quickly.
+These functions make it possible to write compact programs to identify and visualize LCSs.
+Since LCS Tool is programmed in MATLAB, testing new methods and visualizing their results can be done quite quickly.
 
@@
 LCS Tool is free software and we hope it will evolve to incorporate similar advances.
-Optimising computational performance will likely help apply variational LCS methods to large-scale real-time forecasting applications, for example to allow tracking environmental contaminants.
+Optimizing computational performance will likely help apply variational LCS methods to large-scale real-time forecasting applications, for example to allow tracking environmental contaminants.
 
@@
 
-\subsection*{Acknowledgements}
+\subsection*{Acknowledgments}
 
@@
 % Local Variables:
-% ispell-dictionary: "en_GB"
+% ispell-dictionary: "en_US"
 % End:
-% LocalWords:  stretchline stretchlines incompressible gyre eigenvector eigenvectors incompressibility dataset datasets LCS strainline Lyapunov Bickley Rossby timespan variational adaptively vortices parametrise advected strainlines minima extremal
+% LocalWords:  stretchline stretchlines incompressible gyre eigenvector eigenvectors incompressibility dataset datasets LCS LCSs strainline Lyapunov Bickley Rossby timespan variational adaptively vortices parametrise advected strainlines minima extremal FTLE MATLAB maxima geostrophic

### 2014-02-16 main.tex
Remove unused newcommand FIXME
Correct notation used for strainlines and stretchlines

@@
 \newcommand{\reals}{\mathbf R}
-
-\newcommand{\FIXME}[1]{\\\fbox{FIXME: #1}\\}
 
@@
 \begin{equation}
-\boldsymbol r_1' = \boldsymbol \xi_1(\boldsymbol r),
+\boldsymbol r' = \boldsymbol \xi_1(\boldsymbol r),
 \label{eq:strainline}
@@
 \[
-\boldsymbol r_2' = \boldsymbol \xi_2(\boldsymbol r).
+\boldsymbol r' = \boldsymbol \xi_2(\boldsymbol r).
 \]
-A normal perturbation to the strainline segment $\boldsymbol r_1'$ grows under the flow map by a factor $\lambda_2$, i.e., fluid elements are repelled at that rate in the normal direction.
-Equivalently, a normal perturbation to the stretchline segment $\boldsymbol r_2'$ shrinks by a factor $\lambda_1$ and fluid elements are attracted.
+A normal perturbation a strainline segment grows under the flow map by a factor $\lambda_2$, i.e., fluid elements are repelled at that rate in the normal direction.
+Equivalently, a normal perturbation a stretchline segment shrinks by a factor $\lambda_1$ and fluid elements are attracted.
 
@@
 \]
-They are one type of shear-less LCS. For details of the derivation of hyperbolic LCS, see \cite{farazmand13:_shearless}.
+They are one type of shearless LCS. For details of the derivation of hyperbolic LCS, see \cite{farazmand13:_shearless}.
 

### 2014-02-17 main.tex
Small changes to colortracer paragraph

@@
 
-A further point important to note is the coexistence of hyperbolic and elliptic LCS.
-The examples we present show hyperbolic LCSs inside elliptic LCSs.
+A further point important to note is the existence of hyperbolic LCS inside elliptic LCS.
 This may seem contradictory when accustomed to LCSs not crossing, however, numerical experimentation with LCS calculations overlaid with color tracer advection reveals that hyperbolic LCS can cross elliptic barriers and have an impact on the tracer pattern inside elliptic regions.
@@
 Figure~\ref{fig:ocean_dataset_colortracer} illustrates the effect of strainlines on tracer particles inside elliptic regions.
-Color-coded final positions show large gradients along strainlines, directly showing that particles are strongly repelled by the LCS and end in different positions.
+Color-coded final positions show large gradients along strainlines, directly showing that particles are strongly repelled by the LCS and therefore end in different positions.
 

### 2014-02-18 main.tex
Remove newline in middle of paragraph
Add m-file to produce ocean demo figures

@@
 \begin{center}
-\includegraphics[width=0.8\textwidth]{graphics/ocean_dataset/hyperbolic_shearline_lcs_details_strainline}
+\includegraphics[width=0.8\textwidth]{graphics/ocean_dataset/hyperbolic_shear_lcs_strainlines}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=0.8\textwidth]{graphics/ocean_dataset/hyperbolic_shearline_lcs_details_stretchline}
+\includegraphics[width=0.8\textwidth]{graphics/ocean_dataset/hyperbolic_shear_lcs_stretchlines}
 \end{center}

### 2014-02-18 graphics/double_gyre/derivative.m

@@ -1,0 +1,61 @@
+% derivative Derivatives of the double gyre velocity field.
+%
+% SYNTAX
+% derivative_ = derivative(t,position,useEoV,epsilon,amplitude,omega)
+%
+% INPUT ARGUMENTS
+% t: time
+% position: [x1;y1;x2;y2;...;xn;yn]
+% useEov: logical that controls use of the equation of variation
+% epsilon,amplitude,omega: double gyre parameters
+
+function derivative_ = derivative(t,x,useEoV,epsilon,amplitude,omega)
+
+validateattributes(t,{'double'},{'scalar'})
+validateattributes(x,{'double'},{'column'})
+% Cannot use validateattributes to check x is a column vector with an even
+% number of elements
+if mod(numel(x),2)
+    error([mfilename,':numelXNotEven'],'numel(x) = %d is not even',numel(x))
+end
+validateattributes(x,{'double'},{'column'})
+validateattributes(useEoV,{'logical'},{'scalar'})
+validateattributes(epsilon,{'double'},{'scalar'})
+validateattributes(amplitude,{'double'},{'scalar'})
+validateattributes(omega,{'double'},{'scalar'})
+
+if useEoV
+    idx1 = 1:6:size(x,1)-5;
+    idx2 = 2:6:size(x,1)-4;
+else
+    idx1 = 1:2:numel(x)-1;
+    idx2 = 2:2:numel(x);
+end
+
+a = epsilon*sin(omega*t);
+b = 1 - 2*epsilon*sin(omega*t);
+forcing = a*x(idx1).^2 + b*x(idx1);
+
+derivative_ = nan(size(x));
+
+derivative_(idx1) = -pi*amplitude*sin(pi*forcing).*cos(pi*x(idx2));
+derivative_(idx2) = pi*amplitude*cos(pi*forcing).*sin(pi*x(idx2)).*(2*a*x(idx1) + b);
+
+if useEoV
+    % Define terms of the equation of variation
+    idx3 = 3:6:size(x,1)-3;
+    idx4 = 4:6:size(x,1)-2;
+    idx5 = 5:6:size(x,1)-1;
+    idx6 = 6:6:size(x,1);
+    
+    dux = -pi^2*amplitude*cos(pi*forcing).*cos(pi*x(idx2)).*(2*a*x(idx1) + b);
[…]

### 2014-02-18 graphics/ocean_dataset/hyperbolic_shear_lcs_details.m

@@ -1,0 +1,205 @@
+%% Input parameters
+domain = [0,6;-34,-28];
+resolution = [400,400];
+timespan = [100,130];
+
+%% Velocity definition
+load('ocean_geostrophic_velocity.mat')
+% Set velocity to zero at boundaries
+vlon(:,[1,end],:) = 0;
+vlon(:,:,[1,end]) = 0;
+vlat(:,[1,end],:) = 0;
+vlat(:,:,[1,end]) = 0;
+interpMethod = 'spline';
+vlon_interpolant = griddedInterpolant({time,lat,lon},vlon,interpMethod);
+vlat_interpolant = griddedInterpolant({time,lat,lon},vlat,interpMethod);
+lDerivative = @(t,x,~)flowdata_derivative(t,x,vlon_interpolant,vlat_interpolant);
+incompressible = true;
+
+%% LCS parameters
+% Cauchy-Green strain
+cgEigenvalueFromMainGrid = false;
+cgAuxGridRelDelta = 0.01;
+
+% Lambda-lines
+lambda = 1;
+lambdaLineOdeSolverOptions = odeset('relTol',1e-6,'initialStep',1e-2);
+showGraph = true;
+
+% Strainlines
+strainlineMaxLength = 20;
+gridSpace = diff(domain(1,:))/(double(resolution(1))-1);
+strainlineLocalMaxDistance = 2*gridSpace;
+strainlineOdeSolverOptions = odeset('relTol',1e-6);
+
+% Stretchlines
+stretchlineMaxLength = 20;
+stretchlineLocalMaxDistance = 4*gridSpace;
+stretchlineOdeSolverOptions = odeset('relTol',1e-6);
+
+% Graphics properties
+strainlineColor = 'r';
+stretchlineColor = 'b';
+lambdaLineColor = [0,.6,0];
+lcsInitialPositionMarkerSize = 2;
+
+hAxes = setup_figure(domain);
+xlabel(hAxes,'Longitude (\circ)')
+ylabel(hAxes,'Latitude (\circ)')
+
+%% Cauchy-Green strain eigenvalues and eigenvectors
+[cgEigenvector,cgEigenvalue] = eig_cgStrain(lDerivative,domain,resolution,timespan,'incompressible',incompressible,'eigenvalueFromMainGrid',cgEigenvalueFromMainGrid,'auxGridRelDelta',cgAuxGridRelDelta);
+
+% Plot finite-time Lyapunov exponent
[…]

### 2014-02-18 listings/bickley_jet/lambda_lcs_convergence_subdomain.m

@@ -1,0 +1,48 @@
+%% Input parameters
+u = 62.66;
+lengthX = pi*earthRadius;
+lengthY = 1.77e6;
+epsilon = [.075,.4,.3];
+domain = [2e6,lengthX/2;[-1,.25]*2.25*lengthY];
+resolutionX = 400:100:600;
+timespan = [0,2*lengthX/u];
+
+%% Velocity definition
+perturbationCase = 3;
+phiTimespan = [0,25];
+phiInitial = [0,0];
+phiSol = ode45(@d_phi,phiTimespan,phiInitial);
+timeResolution = 1e5;
+phi1 = deval(phiSol,linspace(phiTimespan(1),phiTimespan(2),timeResolution),1);
+phi1Max = max(phi1);
+lDerivative = @(t,x,~)derivative(t,x,false,u,lengthX,lengthY,epsilon,perturbationCase,phiSol,phi1Max);
+incompressible = true;
+
+%% LCS parameters
+% Cauchy-Green strain
+cgStrainOdeSolverOptions = odeset('relTol',1e-4);
+
+% Lambda-lines
+lambda = .995;
+lambdaLineOdeSolverOptions = odeset('relTol',1e-6);
+poincareSection.endPosition = [6.5,-1.4;4.5,-3.5]*1e6;
+[poincareSection.numPoints] = deal(100);
+rOrbit = hypot(diff(poincareSection.endPosition(:,1)),diff(poincareSection.endPosition(:,2)));
+poincareSection.orbitMaxLength = 2*(2*pi*rOrbit);
+lambdaLineColor = [0,.6,0];
+dThresh = 1e-4;
+
+for m = 1:numel(resolutionX)
+    % Make x and y grid spacing as equal as possible
+    lResolutionX = resolutionX(m);
+    gridSpace = diff(domain(1,:))/(double(lResolutionX)-1);
+    resolutionY = round(diff(domain(2,:))/gridSpace) + 1;
+    resolution = [lResolutionX,resolutionY];
+   
+    %% Cauchy-Green strain eigenvalues and eigenvectors
+    [cgEigenvector,cgEigenvalue] = eig_cgStrain(lDerivative,domain,resolution,timespan,'incompressible',incompressible,'odeSolverOptions',cgStrainOdeSolverOptions);
+    
+    %% Lambda-line LCSs
+    [shearline.etaPos,shearline.etaNeg] = lambda_line(cgEigenvector,cgEigenvalue,lambda);
[…]

### 2014-02-18 listings/bickley_jet/strain_lcs.m

@@ -1,0 +1,40 @@
+%% Input parameters
+u = 62.66;
+lengthX = pi*earthRadius;
+lengthY = 1.77e6;
+epsilon = [.075,.4,.3];
+domain = [2e6,.55*lengthX;[-1,.25]*2.25*lengthY];
+resolutionX = 500;
+timespan = [0,2*lengthX/u];
+
+%% Velocity definition
+perturbationCase = 3;
+phiTimespan = [0,25];
+phiInitial = [0,0];
+phiSol = ode45(@d_phi,phiTimespan,phiInitial);
+timeResolution = 1e5;
+phi1 = deval(phiSol,linspace(phiTimespan(1),phiTimespan(2),timeResolution),1);
+phi1Max = max(phi1);
+lDerivative = @(t,x,~)derivative(t,x,false,u,lengthX,lengthY,epsilon,perturbationCase,phiSol,phi1Max);
+incompressible = true;
+
+%% LCS parameters
+% Cauchy-Green strain
+cgStrainOdeSolverOptions = odeset('relTol',1e-4);
+
+% Strainlines
+strainlineMaxLength = 1e8;
+strainlineOdeSolverOptions = odeset('relTol',1e-4);
+
+% Make x and y grid spacing as equal as possible
+gridSpace = diff(domain(1,:))/(double(resolutionX)-1);
+resolutionY = round(diff(domain(2,:))/gridSpace) + 1;
+resolution = [resolutionX,resolutionY];
+
+strainlineLocalMaxDistance = 4*gridSpace;
+
+%% Cauchy-Green strain eigenvalues and eigenvectors
+[cgEigenvector,cgEigenvalue] = eig_cgStrain(lDerivative,domain,resolution,timespan,'incompressible',incompressible,'odeSolverOptions',cgStrainOdeSolverOptions);
+
+%% Strainline LCSs
+strainlineLcs = seed_curves_from_lambda_max(strainlineLocalMaxDistance,strainlineMaxLength,cgEigenvalue(:,2),cgEigenvector(:,1:2),domain,resolution,'odeSolverOptions',strainlineOdeSolverOptions);

### 2014-02-18 graphics/double_gyre/derivative.m
Miscellaneous changes to tidy repository

@@ -1,61 +1,0 @@
-% derivative Derivatives of the double gyre velocity field.
-%
-% SYNTAX
-% derivative_ = derivative(t,position,useEoV,epsilon,amplitude,omega)
-%
-% INPUT ARGUMENTS
-% t: time
-% position: [x1;y1;x2;y2;...;xn;yn]
-% useEov: logical that controls use of the equation of variation
-% epsilon,amplitude,omega: double gyre parameters
-
-function derivative_ = derivative(t,x,useEoV,epsilon,amplitude,omega)
-
-validateattributes(t,{'double'},{'scalar'})
-validateattributes(x,{'double'},{'column'})
-% Cannot use validateattributes to check x is a column vector with an even
-% number of elements
-if mod(numel(x),2)
-    error([mfilename,':numelXNotEven'],'numel(x) = %d is not even',numel(x))
-end
-validateattributes(x,{'double'},{'column'})
-validateattributes(useEoV,{'logical'},{'scalar'})
-validateattributes(epsilon,{'double'},{'scalar'})
-validateattributes(amplitude,{'double'},{'scalar'})
-validateattributes(omega,{'double'},{'scalar'})
-
-if useEoV
-    idx1 = 1:6:size(x,1)-5;
-    idx2 = 2:6:size(x,1)-4;
-else
-    idx1 = 1:2:numel(x)-1;
-    idx2 = 2:2:numel(x);
-end
-
-a = epsilon*sin(omega*t);
-b = 1 - 2*epsilon*sin(omega*t);
-forcing = a*x(idx1).^2 + b*x(idx1);
-
-derivative_ = nan(size(x));
-
-derivative_(idx1) = -pi*amplitude*sin(pi*forcing).*cos(pi*x(idx2));
-derivative_(idx2) = pi*amplitude*cos(pi*forcing).*sin(pi*x(idx2)).*(2*a*x(idx1) + b);
-
-if useEoV
-    % Define terms of the equation of variation
-    idx3 = 3:6:size(x,1)-3;
-    idx4 = 4:6:size(x,1)-2;
-    idx5 = 5:6:size(x,1)-1;
-    idx6 = 6:6:size(x,1);
-    
-    dux = -pi^2*amplitude*cos(pi*forcing).*cos(pi*x(idx2)).*(2*a*x(idx1) + b);
[…]

### 2014-02-18 graphics/double_gyre/lambda_strain_lcs.m

@@ -31,7 +31,6 @@ lambdaLineOdeSolverOptions = odeset('relTol',1e-6);
 % Strainlines
 strainlineMaxLength = 20;
 gridSpace = diff(domain(1,:))/(double(resolution(1))-1);
-
 strainlineLocalMaxDistance = 2*gridSpace;
 strainlineOdeSolverOptions = odeset('relTol',1e-6);
 
@@ -75,4 +74,4 @@ hStrainlineLcs = cellfun(@(position)plot(hAxes,position(:,1),position(:,2)),stra
 set(hStrainlineLcs,'color',strainlineColor)
 
 filename = 'lambda_strain_lcs';
-print_pdf(hFigure,filename);
+print_pdf(hFigure,filename)

### 2014-02-18 graphics/double_gyre/lambda_stretch_lcs.m

@@ -1,4 +1,4 @@
-% function lambda_stretch_lcs
+function lambda_stretch_lcs
 
 %% Input parameters
 epsilon = .1;
@@ -82,10 +82,7 @@ hClosedLambdaLineNeg = cell(nPoincareSection,1);
 for j = 1:nPoincareSection
     hClosedLambdaLinePos{j} = cellfun(@(position)plot(hAxes,position(:,1),position(:,2)),closedLambdaLine{j}{1});
     hClosedLambdaLineNeg{j} = cellfun(@(position)plot(hAxes,position(:,1),position(:,2)),closedLambdaLine{j}{2});
-    if hClosedLambdaLinePos{j}
-    end
 end
-
 hClosedLambdaLine = vertcat(hClosedLambdaLinePos{:},hClosedLambdaLineNeg{:});
 set(hClosedLambdaLine,'color',lambdaLineColor)
 drawnow

### 2014-02-19 main.tex
Miscellaneous changes to tidy repository
Add step to strainline LCS algorithm to reflect remove_strain_in_shear.
Change shear LCS → elliptic LCS
Previously the terms shear and elliptic LCS were both used. Now only
elliptic LCS is used. Some of the LCS Tool functions and variables
still use the term shear however.

@@
 \usepackage[space]{grffile}
-\usepackage{lineno}
 %% hyperref manual advises loading this package last
@@
 \newcommand{\reals}{\mathbf R}
-
-\linenumbers
 
@@
 
-\subsection{Shear LCS}
+\subsection{Elliptic LCS}
 
-Shear LCSs, or coherent Lagrangian vortices, are closed orbits in the Lagrangian shear vector fields $\boldsymbol \eta_{\pm}$ defined in Equation~\eqref{eq:etafields}.
+Elliptic LCSs, or coherent Lagrangian vortices, are closed orbits in the Lagrangian shear vector fields $\boldsymbol \eta_{\pm}$ defined in Equation~\eqref{eq:etafields}.
 We find closed orbits by integrating lambda-lines starting from a Poincare section and evaluating the return map.
-A lambda-line returning to its starting point is a shear LCS.
-The main steps of the algorithm used to calculate shear LCS are enumerated in Table~\ref{t:Shear LCS algorithm} and described in detail below.
-The syntax of shear LCS functions in LCS Tool is in Table~\ref{t:Shear LCS functions}.
+A lambda-line returning to its starting point is an elliptic LCS.
+The main steps of the algorithm used to calculate elliptic LCSs are enumerated in Table~\ref{t:Elliptic LCS algorithm} and described in detail below.
+The syntax of elliptic LCS functions in LCS Tool is in Table~\ref{t:Elliptic LCS functions}.
 
@@
 \end{center}
-\caption{Algorithm to calculate Shear LCS}
-\label{t:Shear LCS algorithm}
+\caption{Algorithm to calculate elliptic LCSs}
+\label{t:Elliptic LCS algorithm}
 \end{table}
@@
 \end{tabular}
-\caption{LCS Tool shear LCS functions syntax}
-\label{t:Shear LCS functions}
+\caption{LCS Tool elliptic LCS functions syntax}
+\label{t:Elliptic LCS functions}
 \end{table}
@@
 whose default value is five, the two points around the zero crossing have absolute values above \lstinline!dThresh!, the zero crossing is discarded.
-Once all valid closed lambda-line orbits have been located, the outermost orbit of every Poincare section is identified as a shear LCS.
+Once all valid closed lambda-line orbits have been located, the outermost orbit of every Poincare section is identified as an elliptic LCS.
 
@@
 \end{center}
-\caption{Example of a Poincare return map obtained for shear LCS. Circle markers indicate closed orbit positions. The filled circle indicates the outermost closed orbit and shear LCS.}
+\caption{Example of a Poincare return map obtained for an elliptic LCS. Circle markers indicate closed orbit positions. The filled circle indicates the outermost closed orbit and elliptic LCS.}
 \label{f:Poincare return map}
@@
 \item Continue integrating strainlines using local maxima of $\lambda_2$ as initial positions until no eligible local maxima of $\lambda_2$ remain.
+\item Remove all strainline segments within elliptic LCSs.
 \end{enumerate}
@@
 
-\lstinputlisting[caption={Bickley jet script for hyperbolic and shear LCSs.
+\lstinputlisting[caption={Bickley jet script for hyperbolic and elliptic LCSs.
 Subset of \lstinline!demo/bickley_jet/hyperbolic_shear_lcs.m!},label=lst:Bickley jet hyperbolic_shear_lcs subset]{listings/bickley_jet/hyperbolic_shear_lcs_subset.m}

### 2014-02-19 main.tex
Use \equiv symbol for definitions
Rename double gyre section and remove table of contents

@@
 \maketitle
-
-\tableofcontents
 
@@
 \[
-F_{t_0}^t(\boldsymbol x_0) := \boldsymbol x(t,t_0,\boldsymbol x_0).
+F_{t_0}^t(\boldsymbol x_0) \equiv \boldsymbol x(t,t_0,\boldsymbol x_0).
 \]
@@
 
-\subsection{The double gyre}
+\subsection{Double gyre}
 The double gyre is a model for a time-dependent two gyre system observed in geophysical flows \parencite{shadden05:_defin_lagran_lyapun}.

### 2014-08-23 main.tex
Update main.tex for lualatex and commath
Incorporate revisions from arXiv e-print
http://arxiv.org/abs/1406.3527v1

@@
 
-\usepackage[utf8]{inputenc}
 \usepackage{amsmath}
@@
 \usepackage{color}
+\usepackage{commath}
+\usepackage{fontspec}
 \usepackage{gensymb}
@@
 \usepackage[space]{grffile}
-%% hyperref manual advises loading this package last
-\usepackage{hyperref}
-\hypersetup{colorlinks=true,linkcolor=blue,citecolor=blue,breaklinks=true}
+\usepackage[hidelinks]{hyperref}
+
+\hypersetup{breaklinks=true}
 
@@
 
-\newcommand{\reals}{\mathbf R}
-
-\title{LCS Tool — An Algorithmic Introduction to Lagrangian Coherent 
-Structures}
+\title{LCS Tool : A Computational Platform for Lagrangian Coherent Structures}
 
@@
 \abstract{
-We give an applied introduction to Lagrangian coherent structures (LCSs) using a computational engine we have programmed called LCS Tool.
-First, we review LCS theory that is based on variational calculus; this introduces elliptic and hyperbolic variational LCSs.
-Then we turn our attention to algorithms for computations conceived in accordance with the variational LCS theory.
-These algorithms are the foundation of LCS Tool.
-We conclude by explaining, with demonstration scripts, how to use LCS Tool to analyze three flow examples: a double gyre, a Bickley jet and an ocean dataset.
+We give an algorithmic introduction to Lagrangian coherent structures (LCSs) using a newly developed computational engine, LCS Tool.
+LCSs are most repelling, attracting and shearing material lines that form the centerpieces of observed tracer patterns in two-dimensional unsteady dynamical systems.
+LCS Tool implements the latest geodesic theory of LCSs for two-dimensional flows, uncovering key transport barriers in unsteady flow velocity data as explicit solutions of differential equations.
+After a review of the underlying theory, we explain the steps and numerical methods used by LCS Tool, and illustrate its capabilities on three unsteady fluid flow examples.
 }
+
+\tableofcontents
 
@@
 
-LCSs are used in fluid dynamics to capture essential flow features and thereby simplify the description of complex fluid flows.
-Here we focus on analyzing two dimensional finite time turbulent flows.
-LCSs identify transport barriers, repelling and attracting structures and vortex boundaries.
-Application areas for LCSs include oceanic and atmospheric flows, biological applications\parencite{wilson09:_lagran_reynol,tallapragada11:_lagran}, aeronautics\parencite{tang10:_accur_lagran_hong_kong_inter_airpor} and mechanical systems\parencite{hadjighasem13:_detec_kam}.
+Lagrangian Coherent Structures (LCSs) are evolving organizing centers of trajectory patterns in non-autonomous dynamical systems\parencite{haller00:_lagran,peacock13:_lagran,Haller2014}.
+Applications of LCSs include oceanic and atmospheric flows\parencite{beron-vera13:_objec_agulh,koh02:_hyper}, biological transport problems\parencite{wilson09:_lagran_reynol,tallapragada11:_lagran,huhn12:_south_indian_ocean_count_madag}, aeronautics\parencite{tang10:_accur_lagran_hong_kong_inter_airpor}, celestial mechanics\parencite{gawlik09:_lagran}, crowd dynamics\parencite{ali07:_lagran_partic_dynam_approac_crowd}, and aperiodically forced mechanical oscillators\parencite{hadjighasem13:_detec_kam}.
 
-The finite time Lyapunov exponent (FTLE) is one widely-used mathematical definition for LCSs.
-Despite its widespread use, examples where the FTLE fails to identify LCSs correctly are known\parencite{haller11:_lagran_coher_struc,norgard12:_secon_lagran_coher_struc}.
-To remedy these problems, variational calculus has been applied to material curve properties in the context of fluid mechanics.
-Thus we obtain a rigorous approach to LCS detection\parencite{haller11:_lagran_coher_struc,farazmand12:_comput_lagran,haller12:_geodes_theor_trans_barrier_two_dimen_flows,farazmand13:_attrac_lagran,haller13:_coher_lagran}.
-This article presents an algorithmic introduction to variational LCSs supported by a computational engine called LCS Tool\footnote{LCS Tool is available at: \url{https://github.com/jeixav/LCS-Tool}.}.
+\textcite{haller01:_distin} proposed that ridges of the Finite-Time Lyapunov exponent (FTLE) are heuristic indicators of hyperbolic (i.e., repelling- and attracting-type) LCSs.
+A number of examples indeed support this principle\parencite{peacock13:_lagran}.
+Equating FTLE ridges with LCSs, however, would create theoretical inconsistencies, as well as false positives and negatives in hyperbolic LCS detection\parencite{haller11:_lagran_coher_struc,norgard12:_secon_lagran_coher_struc}.
+In addition, the role of the FTLE field in the accurate detection of elliptic (vortex-type) and parabolic (jet-core type) LCSs has remained an open question (but see \parencite{beron-vera10:_invar_lagran}).
 
-LCS Tool is a library of MATLAB functions that perform computations to identify LCSs in two dimensional finite time flows.
-LCS Tool aims to help scientists perform variational LCS calculations.
-The examples we present have been programmed into demonstration scripts.
-We encourage readers to download and execute these scripts after reading this article.
+More recent work has focused on an exact mathematical formulation of the properties defining LCSs\parencite{haller11:_lagran_coher_struc,farazmand12:_comput_lagran,haller12:_geodes_theor_trans_barrier_two_dimen_flows,haller13:_coher_lagran,blazevski14:_hyper_ellip_trans_barrier_three}.
+In two-dimensional flows, hyperbolic and parabolic LCSs turn out to be stationary curves of the averaged material shear\parencite{farazmand14:_shearless}, whereas elliptic LCSs are stationary curves of the averaged
+strain\parencite{haller13:_coher_lagran}. These variational formulations lead to explicit solutions for LCSs as null-geodesics of appropriate Lorentzian metrics.
 
-Other LCS software has been developed before; the list includes:
+Here we present a simple algorithmic introduction to geodesic LCS detection without details on the the underlying mathematics.
+We then review the implementation of this approach in a computational engine called LCS Tool\footnote{LCS Tool is publicly available at: \emph{\url{https://github.com/jeixav/LCS-Tool}}.}.
+This engine is a library of MATLAB functions that extract LCSs from two-dimensional unsteady flows.
+The examples we present here have been programmed into demonstration scripts.
+
+The list of publicly available software for LCS detection further
+includes the following packages:
 \begin{itemize}
-\item \emph{MANGEN}\parencite{lekien03:_time}, calculates stability manifolds adaptively in two dimensional finite time velocity fields.
-Includes a graphical user interface and uses MPI for parallel calculations.
+\item \emph{MANGEN}\parencite{lekien03:_time} calculates FTLE and advects material curves in two dimensional velocity fields.
+It also includes a graphical user interface and uses MPI for parallel calculations.
 \item \emph{LCS MATLAB Kit}\parencite{dabiri09:_lmk} calculates the FTLE from velocity datasets. Includes a graphical user interface.
-\item \emph{Newman}\parencite{toit10:_trans} calculates the FTLE in $N$ dimensions. Assists ridge extraction of FTLE fields. Supports analytic and dataset defined velocity definitions.
+\item \emph{Newman}\parencite{toit10:_trans} calculates the FTLE in $N$ dimensions. Assists ridge extraction of FTLE fields. Supports analytic  and dataset based velocity definitions.
 \item \emph{FlowVC}\parencite{shadden10:_flowvc} is a general purpose LCS platform for two and three dimensional datasets. Parallel calculations are supported with Open\-MP, CUDA and OpenCL.
 \end{itemize}
-These packages are based on the FTLE; the focus of LCS Tool and this article is to encourage and facilitate variational LCS analysis.
+These packages primarily target the automated generation of FTLE plots to aid the visual assessment of hyperbolic LCSs in the flow.
+The purpose of LCS Tool and this article is to provide an all-purpose numerical engine for the geodesic extraction of LCSs as parametrized material curves in unsteady two-dimensional flows, and extend the scope of such extraction to elliptic LCSs.
 
@@
 
-The fluid flows considered are defined in two dimensions by a velocity field:
-\begin{align*}
-\dot{\boldsymbol x} = \boldsymbol v(\boldsymbol x,t), && \boldsymbol x \in U \subset \reals^2, && t \in [t_-,t_+].
-\end{align*}
-Integration yields the flow map, defined between an initial time $t_0$ and a final time $t$ within $[t_-,t_+]$:
+We consider two-dimensional, finite-time, unsteady velocity fields of the form
+\begin{align}
+\od{x}{t} = v(x,t), && x \in U \subset \mathbf R^2, && t \in [t_-,t_+].
+\label{eq:v(x,t)}
+\end{align}
+Trajectories of \eqref{eq:v(x,t)} are denoted $x(t;t_0,x_0)$, with $x_0 \in U$ denoting their initial position in the open set $U$ at an initial time $t_0 \in [t_-,t_+]$.
+The flow map is then defined as
 \[
-F_{t_0}^t(\boldsymbol x_0) \equiv \boldsymbol x(t,t_0,\boldsymbol x_0).
+F_{t_0}^t(x_0) \equiv x(t;t,x_0),
 \]
-The time interval used to detect LCSs is arbitrary; it can be the complete interval for which the flow is defined, or data is available, or it can be a sub-interval.
+mapping initial positions to current positions at time $t$.
+The time interval $[t_-,t_+]$ is part of the definition of the finite-time dynamical system \eqref{eq:v(x,t)}.
+This interval may be a time scale of interest or the maximum interval over which velocity data is available from simulations or observations.
 
-The right Cauchy-Green strain tensor measures Lagrangian strain in the velocity field and is defined:
+The right Cauchy-Green strain tensor associated with the flow map is defined as
+\begin{equation}
+C_{t_0}^t(x_0) = \left[\nabla F_{t_0}^t(x_0)\right]^T \nabla F_{t_0}^t(x_0),
+\label{eq:CG}
+\end{equation}
+measuring Lagrangian strain in the velocity field.
+This tensor is symmetric and positive definite\parencite{truesdell04}.
+We label the eigenvalues and eigenvectors of $C_{t_0}^t(\boldsymbol x_0)$ as follows:
 \[
-C_{t_0}^t(\boldsymbol x_0) = \left[D F_{t_0}^t(\boldsymbol x_0)\right]^T D F_{t_0}^t(\boldsymbol x_0).
+C_{t_0}^t \xi_i = \lambda_i \xi_i, \qquad 0 < \lambda_1 \leq \lambda_2, \quad i = 1,2;
 \]
-The eigenvalues of the Cauchy-Green strain tensor are denoted by $\lambda_{1,2}$ and the eigenvectors by $\boldsymbol \xi_{1,2}$.
-The eigenvalues satisfy $0 < \lambda_1 \leq \lambda_2$ and the eigenvectors satisfy $\boldsymbol \xi_1 \perp \boldsymbol \xi_2$.
+\begin{equation}
+\left|\xi_i\right| = 1, \quad \xi_2 = \Omega \xi_1, \quad \Omega = \left(
+\begin{array}{rr}
+0 & -1\\
+1 & 0
+\end{array}
+\right).
+\label{eq:CG_invariants}
+\end{equation}
 
-\subsection{Elliptic LCS}
-We find coherent Lagrangian vortices as closed orbits of one of the two vector fields
+\subsection{Elliptic LCSs}
+\label{sub:Elliptic-LCSs}
+
+We seek positions of closed material lines at time $t_0$ that prevail as coherent Lagrangian vortex boundaries (or \emph{elliptic LCSs}) over a time interval $[t_0,t]\subset[t_-,t_+]$.
+\textcite{haller13:_coher_lagran} argue that such initial material line positions are closed stationary curves of the averaged strain functional
+\[
+Q(\gamma) = \frac{1}{\sigma} \int_0^{\sigma} \frac{\sqrt{\langle r'(s),C_{t_0}^t(r(s))r'(s)\rangle}}{\sqrt{\langle r'(s),r'(s)\rangle}} \dif s,
+\]
+obtained by averaging the tangential strain arising over $[t_0,t]$ along closed material lines parametrized as $r(s)$ with $s \in [0,\sigma].$
+Solutions to this variational problem turn out to be closed orbits of one of two parametrized vector-field families
 \begin{equation}
-\boldsymbol \eta^{\lambda}_\pm = \sqrt{\frac{\lambda_2 - \lambda^2}{\lambda_2 - \lambda_1}} \boldsymbol \xi_1 \pm \sqrt{\frac{\lambda^2 - \lambda_1}{\lambda_2 - \lambda_1}} \boldsymbol \xi_2,
+\eta_\pm^\lambda = \sqrt{\frac{\lambda_2 - \lambda^2}{\lambda_2 - \lambda_1}} \xi_1 \pm \sqrt{\frac{\lambda^2 - \lambda_1}{\lambda_2 - \lambda_1}}\xi_2,
 \label{eq:eta}
 \end{equation}
-which are a function of the eigenvalues and eigenvectors of $C_{t_0}^t(\boldsymbol x_0)$.
-Curves $\gamma$ that are trajectories of $\boldsymbol \eta^{\lambda}_\pm$ can be parametrized as $\boldsymbol r(s)$ with their arc length $s \in [0,\sigma]$.
-They are given by integrating
+with $\lambda > 0$ playing the role of a parameter. Such closed orbits satisfy the differential equation
 \begin{equation}
-\boldsymbol r' \equiv \frac{d \boldsymbol r}{ds} = \boldsymbol \eta^{\lambda}_\pm(\boldsymbol r).
+r' = \eta_\pm^\lambda(r),
 \label{eq:etafields}
 \end{equation}
+which coincide with null-geodesics of the Lorentzian metric family $e_\lambda(u,v) = \left\langle u,\left[D_{t_0}^t(r) - \lambda^2 I\right] v \right\rangle$.
+For this reason, we refer to the computation of elliptic LCSs as limit cycles of \eqref{eq:etafields} as \emph{geodesic detection of elliptic LCSs}.
 
-$\lambda$ is a free parameter close to unity that indicates the uniform stretching of the curve $\gamma$, i.e., each tangent vector $\boldsymbol r'(s)$ with length $l_{t_0}$ stretches by a factor $\lambda$ to length $l_{t}$ under the flow map:
+Any orbit of \eqref{eq:etafields} turns out to stretch uniformly under the flow map $F_{t_0}^t$.
+Specifically, any subset of an orbit of \eqref{eq:etafields} increases its arclength precisely by a factor of $\lambda$.
+For this reason, we refer to trajectories of \eqref{eq:etafields} as $\lambda$-lines.
+Following \parencite{haller13:_coher_lagran}, we call the outermost member of a closed family of $\lambda$-lines a coherent Lagrangian vortex boundary.
+
+\subsection{Hyperbolic LCSs}
+\label{sub:Hyperbolic-LCSs}
+
+Next we consider positions of material lines at time $t_0$ that prevail as most repelling or attracting material lines (or \emph{hyperbolic LCSs}) over a time interval $[t_0,t] \subset [t_-,t_+]$.
+\textcite{farazmand14:_shearless} argue that hyperbolic LCSs are stationary curves of the averaged shear functional
 \begin{gather*}
-l_{t_0}(s) = \sqrt{\langle \boldsymbol r'(s), \boldsymbol r'(s) \rangle},\\
-l_t(s) = \sqrt{\langle \boldsymbol r'(s), C_{t_0}^t(\boldsymbol r(s)) \boldsymbol r'(s) \rangle},\\
-l_t(s) = \lambda\, l_{t_0}(s).
+Q(\gamma) = \frac{1}{\sigma} \int_0^\sigma \frac{\langle r'(s),D_{t_0}^t(r(s)) r'(s)\rangle}{\sqrt{\langle r'(s),C_{t_0}^t(r(s)) r'(s)\rangle\langle r'(s),r'(s)\rangle}}\dif s,\\
+D_{t_0}^t = \frac12 [C_{t_0}^t \Omega - \Omega C_{t_0}^t],
 \end{gather*}
-For $\lambda=1$, $\boldsymbol \eta^{\lambda}_\pm$ points in the direction of maximal Lagrangian shear and we obtain uniformly non-stretching curves over the finite time interval $[t_0, t]$.
-In summary, the boundaries of coherent Lagrangian vortices are closed curves that uniformly stretch by a factor $\lambda$ when advected by the flow.
-We refer to these curves as lambda-lines.
-The outermost lambda-line is the physically observable boundary of an eddy, generating the tracer pattern of an isolated rotating fluid region.
+obtained by averaging the Lagrangian shear arising over $[t_0,t]$ along closed material lines parametrized as $r(s)$ with $s \in [0,\sigma]$.
+More precisely, hyperbolic LCSs are stationary curves of $Q(\gamma)$ with respect to fixed-endpoint perturbations.
+We note that parabolic LCSs (Lagrangian jet cores) are also stationary curves of $Q(\gamma)$, but under variable endpoint perturbations (cf.
+\textcite{farazmand14:_shearless}.)
 
-Note that the equation for $\boldsymbol \eta^{\lambda}_\pm$ follows from the variational principle that the first variation of the averaged Lagrangian strain along the sought extremal closed curve vanishes (see \parencite{haller13:_coher_lagran} for full derivations):
-\begin{eqnarray*}
-Q(\gamma) &=& \frac{1}{\sigma} \int_0^\sigma \frac{l_t(s)}{l_{t_0}(s)}\text{d}s\\
-\delta Q(\gamma) &=& 0.
-\end{eqnarray*}
-This coherence principle means the sought curve $\gamma$, the coherent eddy boundary, is a stationary point of the averaged Lagrangian strain functional and therefore nearby fluid elements cannot separate from $\gamma$.
-
-\subsection{Hyperbolic LCS}
-Hyperbolic LCSs are trajectories of the eigenvector fields $\boldsymbol \xi_{1,2}$. Repelling hyperbolic LCSs, denoted strainlines, are obtained as solutions of the differential equation
+Solutions to this variational problem turn out to be orbits of the $\xi_1$ or $\xi_2$ eigenvector field. Repelling LCSs (\emph{shrinklines}) are obtained as trajectories of the differential equation
 \begin{equation}
-\boldsymbol r' = \boldsymbol \xi_1(\boldsymbol r),
+r' = \xi_1(r),
 \label{eq:strainline}
 \end{equation}
-and attracting hyperbolic LCSs, denoted stretchlines, are obtained as a solution of the differential equation
-\[
-\boldsymbol r' = \boldsymbol \xi_2(\boldsymbol r).
-\]
-A normal perturbation a strainline segment grows under the flow map by a factor $\lambda_2$, i.e., fluid elements are repelled at that rate in the normal direction.
-Equivalently, a normal perturbation a stretchline segment shrinks by a factor $\lambda_1$ and fluid elements are attracted.
+and attracting LCSs (\emph{stretchlines}) are obtained as trajectories of the differential equation
+\begin{equation}
+r' = \xi_2(r).
+\label{eq:stretchline}
+\end{equation}
 
-Similarly to elliptic LCS, hyperbolic LCS are also based on a variational principle, in this case for the Lagrangian shear
-\[
-p(s) = \frac{\langle \boldsymbol r'(s), D(\boldsymbol r(s)) \boldsymbol r'(s)\rangle}{\sqrt{\langle \boldsymbol r'(s), C(\boldsymbol r(s)) \boldsymbol r'(s)\rangle \langle \boldsymbol r'(s), \boldsymbol r'(s)\rangle}}
-\]
-where
-\[
-D(\boldsymbol x_0) = \frac12[C(\boldsymbol x_0) \Omega - \Omega C(\boldsymbol x_0)], \quad \Omega = \begin{pmatrix}0&-1\\1&0\end{pmatrix}.
-\]
-The first variation of the average Lagrangian shear along a hyperbolic LCS vanishes
-\[
-P(\gamma) = \frac{1}{\sigma} \int_0^\sigma p(s) \text{d}s,
-\]
-\[
-\delta P(\gamma) = 0.
-\]
-They are one type of shearless LCS. For details of the derivation of hyperbolic LCS, see \cite{farazmand13:_shearless}.
+Shrinklines and stretchlines coincide with the null-geodesics of the Lorentzian metric $h(u,v) = \left\langle u,D_{t_0}^t(r) v \right\rangle$.
+For this reason, we refer to the computation of hyperbolic LCSs as strongest normally-repelling or normally-attracting orbits of \eqref{eq:etafields} as \emph{geodesic detection of hyperbolic LCSs}.
+
+To compute the normal repulsion of shrinklines, we note that an infinitesimal
+normal perturbation to a shrinkline $\gamma$ at its point $r$ grows under the flow map $F_{t_{0}}^{t}$ by a factor of $\lambda_{2}(r)$ in the direction normal to $F_{t_{0}}^{t}(\gamma).$ Similarly, small normal perturbations to a stretchline decay by a factor $\lambda_{1}(r)$ in the direction normal to the evolving stretchline.
 
@@
 
-\section{Methods}
+\section{Numerical methods}
 
-This section presents the numerical methods used to compute LCSs and explains them step-wise. In other words, the algorithms that constitute LCS Tool are presented.
+Here we describe a step-by-step numerical implementation of the geodesic detection of elliptic and hyperbolic LCSs through the differential equations \eqref{eq:etafields}, \eqref{eq:strainline}, and \eqref{eq:stretchline}.
 
-\subsection{The Cauchy-Green strain tensor}
+\subsection{Computing the invariants of the Cauchy-Green strain tensor}
 
-To obtain Lagrangian coherent structures the Cauchy-Green strain tensor must be calculated.
-More specifically, the eigenvalues and eigenvectors of the Cauchy-Green strain tensor are needed.
-The function associated with this calculation in LCS Tool is \lstinline!eig_cgStrain!.
-The main steps of this function are enumerated in Table~\ref{t:Cauchy-Green algorithm} and are described in detail below.
-Table~\ref{t:eig_cgStrain syntax} summarizes the syntax of \lstinline!eig_cgStrain!.
+The common first step in calculating elliptic and hyperbolic LCSs is the computation of the Cauchy-Green strain tensor field $C_{t_0}^t(x_0)$, as defined in \eqref{eq:CG}.
+The function performing this calculation in LCS Tool is \lstinline!eig_cgStrain!.
+The main steps executed by this function are enumerated in Table~\ref{t:Cauchy-Green algorithm}, while Table~\ref{t:eig_cgStrain syntax} summarizes the syntax of \lstinline!eig_cgStrain!.
 
@@
 \begin{enumerate}
-\item Setup a Cartesian grid and auxiliary grid in the (rectangular) domain of the flow.
-\item Integrate the flow velocity equations at each grid point and auxiliary grid point. This yields the flow map, $F_{t_0}^t(\boldsymbol x_0)$.
-\item Use finite differences to approximate the derivative of the flow map, $D F_{t_0}^t(\boldsymbol x_0)$.
-\item Compute the Cauchy-Green strain tensor, $C_{t_0}^t = \left(D F_{t_0}^t\right)^T D F_{t_0}^t$ and its eigenvalues, $\lambda_{1,2}$ and eigenvectors, $\xi_{1,2}$.
+\item Define a Cartesian grid for initial conditions of trajectories. Define
+an auxiliary grid for differentiating with respect to initial conditions.
+\item Solve \eqref{eq:v(x,t)} starting from each grid point and auxiliary grid point over the time interval $[t_0,t]$. This gives a discrete approximation to the flow map $F_{t_0}^t(x_0)$.
+\item Use finite differencing over the auxiliary grid to compute numerically the derivative of the flow map $DF_{t_0}^t(x_0)$.
+\item Compute the Cauchy-Green strain tensor field $C_{t_0}^t(x_0) = \left(DF_{t_0}^t(x_0)\right)^T DF_{t_0}^t(x_{0})$, its eigenvalue field $\lambda_{1,2}(x_0)$, and eigenvector fields $\xi_{1,2}(x_0)$ over the initial condition grid.
 \end{enumerate}
-\caption{Algorithm to calculate the Cauchy-Green strain tensor eigenvalues and eigenvectors}
+\caption{Algorithm to calculate the invariants of the Cauchy-Green strain tensor field.}
 \label{t:Cauchy-Green algorithm}
@@
 \begin{tabular}{|c|p{.7\textwidth}|}
-\hline \hline
-\multicolumn{2}{|p{\textwidth}|}{\lstinline![cgEigenvector,cgEigenvalue] = eig_cgStrain(derivative,domain,timespan,resolution)!}\\
 \hline
-\lstinline!derivative! & function handle for flow velocity equations\\
+\multicolumn{2}{|p{\textwidth}|}{\lstinline![cgEigenvector,cgEigenvalue] = eig_cgStrain(derivative,domain,timespan,resolution)!}\tabularnewline
 \hline
-\lstinline!domain! & $2 \times 2$ array to define flow domain\\
+\lstinline!derivative! & function handle for flow velocity equations\tabularnewline
 \hline
-\lstinline!timespan! & $1 \times 2$ array to define flow timespan\\
+\lstinline!domain! & $2 \times 2$ array to define flow domain\tabularnewline
 \hline
-\lstinline!resolution! & $1 \times 2$ array to define Cauchy-Green strain main grid resolution\\
+\lstinline!timespan! & $1 \times 2$ array to define flow timespan\tabularnewline
 \hline
-\lstinline!auxGridRelDelta! & optional scalar between 0 and 0.5 to specify auxiliary grid spacing. Default is $10^{-2}$.\\
+\lstinline!resolution! & $1 \times 2$ array to define Cauchy-Green strain main grid resolution\tabularnewline
 \hline
-\lstinline!eigenvalueFromMainGrid! & optional logical to control whether eigenvalues of Cauchy-Green strain are calculated from main or auxiliary grid. Default is \lstinline!true!.\\
+\lstinline!auxGridRelDelta! & optional scalar between 0 and 0.5 to specify auxiliary grid spacing. Default is $10^{-2}$.\tabularnewline
 \hline
-\lstinline!incompressible! & optional logical to specify if incompressibility is imposed. Default is \lstinline!false!.\\
+\lstinline!eigenvalueFromMainGrid! & optional logical to control whether eigenvalues of Cauchy-Green strain are calculated from main or auxiliary grid. Default is \lstinline!true!.\tabularnewline
 \hline
-\lstinline!odeSolverOptions! & optional \lstinline!odeset! structure to specify flow map integration parameters\\
-\hline \hline
+\lstinline!incompressible! & optional logical to specify if incompressibility is imposed. Default is \lstinline!false!.\tabularnewline
+\hline
+\lstinline!odeSolverOptions! & optional \lstinline!odeset! structure to specify flow map integration parameters\tabularnewline
+\hline
 \end{tabular}
@@
 
-We setup a Cartesian grid according to user-defined resolutions.
-We can find the optimal resolution by an iterative procedure in which we start with a low resolution and double it until convergence of LCSs is observed visually.
-If the chosen domain only comprises a few expected LCSs, e.g., one vortex, a resolution of about 500 grid point along the longest axis usually gives good results.
-Otherwise, for larger domains, a higher resolution must be chosen.
-Furthermore an auxiliary grid is setup.
-This is illustrated in Figure~\ref{f:main and auxiliary grids}.
-The improvements the auxiliary grid yields compared to using only the main grid were reported in \textcite{farazmand12:_comput_lagran}.
-Experience indicates that optimal results are obtained when the auxiliary grid spacing is 1\% to 10\% of the main grid spacing.
+The Cartesian grid of initial conditions mentioned in Table~\ref{t:Cauchy-Green algorithm} is rectangular, with user-defined vertical and horizontal ranges and resolutions.
+The optimal resolution may be determined by a successive doubling of the initial resolution until convergence of the extracted LCSs is observed visually.
+If the domain of interest comprises only a few expected LCSs, e.g., one vortex, then a resolution of about 500 grid point along the longest axis usually gives good results.
+Otherwise, a higher resolution must be chosen.
+
+The auxiliary grid (cf.
+Table~\ref{t:Cauchy-Green algorithm}) comprises four points placed symmetrically around each point of the Cartesian grid (Figure~\ref{f:main and auxiliary grids}.) These points are used to achieve increased accuracy in the finite-difference approximation
+\[
+\nabla F_{t_0}^t(x_0) \approx
+\left(\begin{array}{cc}
+\frac{x_1(t;t_0,x_0 + \delta x_1) - x_1(t;t_0,x_0 - \delta x_1)}{2\vert\delta x_1\vert} & \frac{x_1(t;t_0,x_0+\delta x_2) - x_1(t;t_0,x_0-\delta x_2)}{2\vert\delta x_2\vert}\\
+\frac{x_2(t;t_0,x_0+\delta x_1) - x^2(t;t_0,x_0 - \delta x_1)}{2\vert\delta x_1\vert} & \frac{x_2(t;t_0,x_0 + \delta x_2) - x_2(t;t_0,x_0 - \delta x_2)}{2\vert\delta x_2\vert}
+\end{array}\right)
+\]
+of $\nabla F_{t_0}^t$ at a point $x_0$ of the Cartesian grid.
+Here $\delta x_i$ is a vector of length $\vert\delta x_i\vert > 0$ that points from the Cartesian grid-point $x_0$ in the $i^{th}$ coordinate direction (Figure~\ref{f:main and auxiliary grids}).
+Computational improvements arising from the use of the auxiliary grid over simply using the nearest points of the main grid were reported in Ref.\,\parencite{farazmand12:_comput_lagran}.
+Experience suggests an auxiliary grid spacing of 1-10\% of the main grid spacing.
 
@@
 \begin{center}
-\includegraphics[width=0.5\textwidth]{graphics/main_aux_grids}
+\includegraphics[width=.5\textwidth]{graphics/main_aux_grids}
 \end{center}
-\caption{Illustration of main grid (filled circles) and auxiliary grid (empty circles) used to compute the Cauchy-Green strain.
+\caption{Illustration of the main grid (filled circles) and the auxiliary grid (empty circles) used in the computation of the derivative of the flow map in the definition \eqref{eq:CG} of the Cauchy--Green stain tensor.
 The variable \lstinline!auxGridRelDelta!
@@
 The function \lstinline!eig_cgStrain!
-provides the option to calculate eigenvectors from the auxiliary grid, but eigenvalues from the main grid.
-We have found that for flows defined analytically the eigenvalues should be calculated from the main grid whereas for the flows defined by datasets, using the auxiliary grid gives better results.
+of LCS Tool provides the option to calculate Cauchy--Green eigenvectors from the auxiliary grid using eigenvalues calculated from the main grid.
+We have found that for flows defined analytically, the eigenvalues can reliably be calculated from the main grid.
+For the flows defined by datasets, using the auxiliary grid for eigenvalue calculations gives better results.
 
-As stated, typically we set the main grid resolution of the Cauchy-Green strain tensor to around 500 grid points.
-For a square domain, this implies the velocity equations must be integrated, after including 4 auxiliary grid points around every main grid point, at 1.25 million grid points.
-The natural way to perform these integration calculations would be to solve a two dimensional first-order ODE system at every grid point and iterate over every grid point in a for-loop.
-With 1.25 million grid points this calculation can take an excessively long time.
-To alleviate this problem, we solve the flow equations in a vector form where the equations at all grid points are combined into a single system of equations, even though the system is composed of independent blocks of two dimensional first-order ODE systems.
-MATLAB's \lstinline!ode45!
-function is used to perform the integration and the flow integration typically takes five to ten minutes with default error tolerances.
-The drawback of vector form integration is that memory requirements can become excessive at high resolutions.
-Furthermore, writing the velocity function in vector form is more error-prone than the simpler two dimensional form.
+As mentioned above, a typical main grid for the Cauchy-Green strain tensor comprises $500 \times 500$.
+This means that after the addition of 4 auxiliary grid points around each main grid points, eq.
+\eqref{eq:v(x,t)} must be integrated over 1.25 million initial conditions.
 
-The Cauchy-Green strain tensor can have singularities, i.e., $C_{t_0}^t(\boldsymbol x_0) = I$.
-Such points are isolated in general\parencite{delmarcelle94}, therefore no special accommodations are made in computations for singularities of the Cauchy-Green strain tensor.
+To avoid excessive computational times in MATLAB, we vectorize \eqref{eq:v(x,t)}, i.e., combine its right-hand-side evaluated over each initial points into a single system of equations.
+The resulting system is composed of independent blocks of two-dimensional first-order ODEs.
+We then use MATLAB's \lstinline!ode45!
+to perform trajectory integration from all grid points simultaneously.
+This process typically takes five to ten minutes with default error tolerances.
+A potential drawback of vector form integration is that memory requirements may become excessive at high resolutions.
+Furthermore, writing the velocity function in vector form is more error-prone than the simpler two-dimensional form.
+
+The Cauchy-Green strain tensor generically admits points singularities, i.e., points where $C_{t_0}^t(x_0)$ has repeated eigenvalues.
+At these points the eigenvectors $\xi_1(x_0)$ and $\xi_2(x_0)$ are no longer well-defined.
+This generically arises at a finite set of isolated points within the computational domain\parencite{delmarcelle94}, and hence lie off the computational grid with probability one.
 
@@
 
-Flows that are incompressible, meaning $D \cdot \boldsymbol v = 0$, satisfy the relation $\lambda_1(\boldsymbol x_0) \lambda_2(\boldsymbol x_0) = 1, \forall \: \boldsymbol x_0 \: \in \: U$\parencite{arnold78:_mathem}.
-This gives the possibility of imposing incompressibility when calculating the eigenvalues of the Cauchy-Green strain tensor by first calculating $\lambda_2$ and then setting $\lambda_1(\boldsymbol x_0) = 1/\lambda_2(\boldsymbol x_0)$.
-This procedure should be applied carefully since by changing $\lambda_1$ but not $\boldsymbol \xi_1$, the error in satisfying the eigenvalue equation for the Cauchy-Green strain tensor may increase.
+Incompressible flows (i.e.
+those satisfying $\nabla \cdot v=0$) admit the relationship $\lambda_1(x_0) \lambda_2(x_0) = 1$ at all points of the computational domain \parencite{arnold78:_mathem}.
+Incompressibility can be computationally imposed by first calculating $\lambda_2(x_0)$, then setting $\lambda_1(x_0) = 1/\lambda_2(x_0)$ and calculating the strain eigenvectors $\xi_2$ from $\lambda_2$, then $\xi_1$ from the relationship in \eqref{eq:CG_invariants}.
+Experience shows that computing $\lambda_i$ in this order gives higher accuracy than in the reverse order \parencite{farazmand12:_comput_lagran}.
 
-At some grid points it may occur that $\lambda_2 < 1$ due to numerical integration errors of the flow velocity equations.
-By setting integration tolerances to smaller values the number of grid points with $\lambda_2 < 1$ can be reduced.
-Nonetheless, some grid points can be so close to singularities of the Cauchy-Green strain tensor that satisfying $\lambda_2 \geq 1$ everywhere could be excessively computationally costly.
-Therefore the function \lstinline!eig_cgStrain!
-records the number of points at which incompressibility cannot be imposed and this measure can guide setting appropriate integration tolerances.
+At some grid points, $\lambda_2 < 1$ may occur due to numerical integration errors.
+By setting the integration tolerances to smaller values, the number of such grid points is reduced.
+Enforcing $\lambda_2 \geq 1$ everywhere, however, can incur excessive computational cost.
+To this end, the function \lstinline!eig_cgStrain!
+records the number of points with $\lambda_2 < 1$, providing a measure for setting feasible integration tolerances.
 
-\subsubsection{Special case: data-defined velocity fields}
+\subsubsection{Special case: dataset velocity fields}
 
-Velocity fields defined by datasets require pre-processing prior to performing the integration that yields the flow map.
-An interpolation function must be written to enable evaluating the velocity function at arbitrary points in $U$ and at arbitrary times between $t_-$ and $t_+$.
-MATLAB's interpolation functions can be used in the velocity function.
-In section \ref{sec:oceandataset} we present an example for LCSs computed for an ocean dataset and details of the interpolation function are given.
+Velocity fields defined by datasets require pre-processing before
+they are used in the numerical integration of \eqref{eq:v(x,t)}.
+This requires spatial and temporal interpolation that enables the evaluation of the velocity function at arbitrary points in $U$ and at arbitrary times between $t_-$ and $t_+$.
+In section \ref{sec:oceandataset}, we present an ocean dataset example with details of possible interpolation functions.
 
-Beyond the pre-processing needed to obtain the Cauchy-Green strain tensor eigenvalues and eigenvectors, no special manipulations are needed to treat data-defined flow.
-It may be however, that parameter values giving good results with analytically defined flows need adjustment prior to processing data-defined velocity fields.
+\subsection{Computing elliptic LCSs}
 
-\subsection{Elliptic LCS}
+As discussed in section \eqref{sub:Elliptic-LCSs}, positions of elliptic LCSs at time $t_0$ are found as closed orbits of the $\eta_\pm^\lambda$ vector fields defined in \eqref{eq:eta}.
+We find such orbits by integrating \eqref{eq:etafields} from points of an appropriately chosen section (Poincare section), and evaluating the first return map (Poincaré map) onto this section.
+A $\lambda$-line returning to its starting point is then an elliptic LCS.
+The outermost member of a family of closed $\lambda$-lines (obtained by varying $\lambda$) is a coherent Lagrangian vortex boundary\parencite{haller13:_coher_lagran}.
 
-Elliptic LCSs, or coherent Lagrangian vortices, are closed orbits in the Lagrangian shear vector fields $\boldsymbol \eta_{\pm}$ defined in Equation~\eqref{eq:etafields}.
-We find closed orbits by integrating lambda-lines starting from a Poincare section and evaluating the return map.
-A lambda-line returning to its starting point is an elliptic LCS.
-The main steps of the algorithm used to calculate elliptic LCSs are enumerated in Table~\ref{t:Elliptic LCS algorithm} and described in detail below.
-The syntax of elliptic LCS functions in LCS Tool is in Table~\ref{t:Elliptic LCS functions}.
+The main steps in calculating elliptic LCSs are enumerated in Table~\ref{t:Elliptic LCS algorithm} and described in further detail below.
+The syntax of elliptic LCS functions in LCS Tool is shown in Table~\ref{t:Elliptic LCS functions}.
 
@@
 \item Position Poincare sections in flow domain to specify initial positions of lambda-lines
-\item Integrate lambda-lines tangent to $\boldsymbol \eta_{\pm}$ (see Equation~\eqref{eq:etafields}.)
-\item Calculate Poincare return map
-\item Find closed orbits positions from Poincare return map
-\item Identify outermost closed orbit
+\item Integrate $\lambda$-lines tangent to $\eta_\pm^\lambda$ (see ~\eqref{eq:etafields}.)
+\item Calculate Poincare map
+\item Find closed orbits for fixed points of the Poincare map
+\item Identify outermost closed orbit on each Poincare section
 \end{enumerate}
 \end{center}
-\caption{Algorithm to calculate elliptic LCSs}
+\caption{Algorithm to calculate elliptic LCSs and coherent Lagrangian vortex boundaries.}
 \label{t:Elliptic LCS algorithm}
@@
 \begin{tabular}{|c|p{.7\textwidth}|}
+\hline
+\multicolumn{2}{|p{\textwidth}|}
+{\lstinline![shearline.etaPos,shearline.etaNeg] = lambda_line(cgEigenvector,cgEigenvalue,lambda)!}\tabularnewline
+\hline
+\lstinline!cgEigenvector! & array of Cauchy-Green strain eigenvectors\tabularnewline
+\hline
+\lstinline!cgEigenvalue! & array of Cauchy-Green strain eigenvalues\tabularnewline
+\hline
+\lstinline!lambda! & scalar lambda value in Equation~\eqref{eq:eta}\tabularnewline
 \hline \hline
-\multicolumn{2}{|p{\textwidth}|}{\lstinline![shearline.etaPos,shearline.etaNeg] = lambda_line(cgEigenvector,cgEigenvalue,lambda)!}\\
+\multicolumn{2}{|p{\textwidth}|}{\lstinline![closedOrbits,orbits] = poincare_closed_orbit_multi(domain,resolution,shearline,PSList)!}\tabularnewline
 \hline
-\lstinline!cgEigenvector! & array of Cauchy-Green strain eigenvectors\\
+\lstinline!domain! & array to define flow domain\tabularnewline
 \hline
-\lstinline!cgEigenvalue! & array of Cauchy-Green strain eigenvalues\\
+\lstinline!resolution! & $1 \times 2$ array to define main grid resolution for Cauchy-Green strain tensor\tabularnewline
 \hline
-\lstinline!lambda! & scalar lambda value in Equation~\eqref{eq:eta}\\
-\hline \hline
-\multicolumn{2}{|p{\textwidth}|}{\lstinline![closedOrbits,orbits] = poincare_closed_orbit_multi(domain,resolution,shearline,PSList)!}\\
+\lstinline!shearline! & structure of arrays of $\eta_+$ and $\eta_-$ values on main grid\tabularnewline
 \hline
-\lstinline!domain! & array to define flow domain\\
+\lstinline!PSList! & user-defined structure for Poincare section end-points, number of $\lambda$-lines launched from Poincare section, and maximum closed $\lambda$-line length\tabularnewline
 \hline
-\lstinline!resolution! & $1 \times 2$ array to define Cauchy-Green strain main grid resolution\\
+\lstinline!nBisection! & optional number of bisection steps to refine zero crossings of Poincare map. Default values: 5.\tabularnewline
 \hline
-\lstinline!shearline! & structure of arrays of $\eta_+$ and $\eta_-$ values on main grid\\
+\lstinline!dThresh! & optional threshold to discard discontinuous zero crossings of Poincare map. Default value: $10^{-2}$.\tabularnewline
 \hline
-\lstinline!PSList! & structure for Poincare section end-points, number of lambda-lines launched from Poincare section, and maximum closed lambda-line length\\
+\lstinline!odeSolverOptions! & optional \lstinline!odeset! structure to specify $\lambda$-line integration parameters\tabularnewline
 \hline
-\lstinline!nBisection! & optional number of bisection steps to refine zero crossings of Poincare return map. Default is 5.\\
+\lstinline!periodicBc! & optional $1 \times 2$ logical array to specify periodic boundary conditions. Default is \lstinline![false,false]!.\tabularnewline
 \hline
-\lstinline!dThresh! & optional threshold to discard discontinuous zero crossings of Poincare return map. Default is $10^{-2}$.\\
-\hline
-\lstinline!odeSolverOptions! & optional \lstinline!odeset! structure to specify lambda-line integration parameters\\
-\hline
-\lstinline!periodicBc! & optional $1 \times 2$ logical array to specify periodic boundary conditions. Default is \lstinline![false,false]!.\\
-\hline \hline
 \end{tabular}
-\caption{LCS Tool elliptic LCS functions syntax}
+\caption{Syntax of LCS Tool's elliptic LCS functions.}
 \label{t:Elliptic LCS functions}
@@
 
-The first step is to set the position of Poincare sections as straight lines in regions where closed lambda-lines are expected.
-The Poincare section must be oriented such that the first endpoint is close to the center of the expected eddy and the second endpoint is outside the expected eddy.
-There is no definite procedure for deciding the position of Poincare sections.
-Possibilities include an examination of the finite time Lyapunov exponent field of the flow, or some prior knowledge of where coherent eddies are be expected.
-Additionally, the number of lambda-lines launched from the Poincare section, \lstinline!poincareSection.numPoints!, must be defined.
-A good default is 100.
+The first step is to define the position of Poincare sections in regions where closed $\lambda$-lines are expected based on a visual analysis of the orbit structure of the $\eta_\pm^\lambda$ vector field.
+The Poincare section is to be oriented such that the first endpoint is close to the center of the expected Lagrangian vortex, and the second endpoint is outside this vortex.
+There is no automated procedure procedure implemented yet in LCS Tool for the positioning of Poincare sections (see, however, \parencite{karrasch14:_autom_lagran} for a recently developed algorithm).
+Additionally, the number of lambda-lines launched from the Poincaré section, \lstinline!poincareSection.numPoints!, must be defined.
+A reasonable default value is 100.
 
-The second step is to integrate the lambda-lines to obtain the Poincare return map.
-Integration of lambda-lines is performed with the $\boldsymbol \eta_+$ and $\boldsymbol \eta_-$ vector fields defined on the main grid.
-These vector fields are derived from eigenvalues and eigenvectors as described in~\textcite{farazmand12:_comput_lagran}.
-The eigenvector fields have orientation discontinuities, therefore the vector field must be locally reoriented for the integration.
-The algorithm used to perform variable step-size integration in vector fields with orientation discontinuities is enumerated in Table~\ref{t:variable step integration} and illustrated in Figure~\ref{f:variable step integration}.
-Linear interpolation is used when interpolating $\boldsymbol \eta_\pm$ within a grid element, since using higher order interpolation would necessitate verifying that there are no orientation discontinuities at more than only the four nearest grid points.
-Detection of orientation discontinuities is achieved by checking the inner product of the vector fields at adjacent grid points.
-Then, rotations exceeding 90\degree\, are classified as orientation discontinuities and are corrected before linear interpolation.
-Thus when setting the Cauchy-Green strain tensor main grid resolution, a histogram of eigenvector field rotations may be calculated to ensure all rotations are well below 90\degree\, or almost 180\degree.
+The second step is to integrate the $\lambda$-lines  starting from the Poincare section to obtain the corresponding Poincare map.
+Integration of $\lambda$-lines is performed using the $\eta_\pm^\lambda$ vector fields defined in \eqref{eq:eta} over the main grid.
+The underlying eigenvector fields $\xi_i(x_0)$ have generic but removable orientation discontinuities, which require monitoring and reorientation, as needed.
+This process is sketched in Table~\ref{t:variable step integration} and illustrated in Figure~\ref{f:variable step integration}.
+Linear interpolation is used in the interpolation of $\eta_\pm^\lambda$ within a grid element, since using higher-order interpolation would necessitate verifying that there are no orientation discontinuities beyond the four nearest grid points.
+We identify orientation discontinuities by checking the inner product of the $\eta_\pm^\lambda$ vectors at adjacent grid points.
+Rotations exceeding 90°, between two such neighboring vectors are classified as orientation discontinuities and are corrected before linear interpolation.
+When setting the Cauchy-Green strain tensor main grid resolution, one may find it helpful to calculate a histogram of eigenvector-field rotations to ensure that all rotations are well below 90°.
 
@@
 \item Linearly interpolate vector field orientation at initial position
-\item At next position, check whether vector field has rotated by over 90\degree, if yes, flip the vector field orientation by 180\degree.
-\item Stop integration when lambda-line returns to Poincare section, lambda-line reaches the domain boundary, or maximum integration length has been reached.
+\item At next position, check whether vector field has rotated by over 90°, if yes, flip the vector field orientation by 180°.
+\item Stop integration when $\lambda$-line returns to Poincare section, $\lambda$-line reaches the domain boundary, or maximum integration length has been reached.
 \end{enumerate}
-\caption{Algorithm used for variable time step integration of lambda-lines.}
+\caption{Algorithm used for variable time step integration of $\lambda$-lines.}
 \label{t:variable step integration}
@@
 \begin{center}
-\includegraphics[width=0.8\textwidth]{graphics/variable_step_integration}
+\includegraphics[width=.8\textwidth]{graphics/variable_step_integration}
 \end{center}
-\caption{Schematic illustration of lambda-line variable time step integration.
-At the initial point, there is an orientation discontinuity at the lower right grid point that must be corrected prior to linear interpolation.
+\caption{Schematic illustration of the variable-time-step $\lambda$-line integration.
+At the initial point, there is an orientation discontinuity at the lower-right grid point that must be corrected prior to linear interpolation.
 At point \#1, no orientation discontinuities are present.
-At point \#2 the interpolated $\eta_\pm$ vector must be rotated by 180\degree\,to match the orientation of the trajectory.}
+At point \#2 all interpolated $\eta_\pm^\lambda$ vectors must be rotated by 180° to match the orientation of the trajectory.}
 \label{f:variable step integration}
@@
 
-An example of a Poincare return map produced from integration of the $\boldsymbol \eta_\pm$ field is shown in Figure~\ref{f:Poincare return map}.
-Well behaved orbits will return to the Poincare section and their integration will be stopped using ODE event detection.
-Some orbits may however deviate far from the Poincare section and their integration could take inordinately long.
-To limit this behavior, a maximum orbit length, \lstinline!poincareSection.orbitMaxLength!, is specified.
-In practice, viewing the Poincare section as the radius of a circle and setting the maximum lambda-line integration length to twice the circumference gives good results.
-In Figure~\ref{f:Poincare return map}, circle markers indicate closed orbit positions.
-Closed orbits exist where the Poincare return map, the distance between the final and the initial point of the orbit $P(s)-s$, is zero.
+An example of a Poincare map produced from integration of the $\eta_\pm^\lambda$ field is shown in Figure~\ref{f:Poincare return map}.
+Most orbits will return to the Poincare section and their integration will then be stopped using the ODE event detection function of MATLAB.
+Some orbits may, however, deviate far from the Poincare section and do not return for any reasonable integration time.
+To control this behavior, we specify a maximum orbit length, \lstinline!poincareSection.orbitMaxLength!.
+In practice, viewing the Poincare section as the radius of a circle and setting the maximum $\lambda$-line integration length to twice the circumference gives good results.
+
+In Figure~\ref{f:Poincare return map}, circle markers indicate fixed points of the Poincare map, i.e., points where the distance between the final and the initial point of the orbit $P(s)-s$, is zero.
 The function \lstinline!poincare_closed_orbit_multi!
 performs the computations.
-In Figure~\ref{f:Poincare return map} not all zero crossings have circles.
-This is because a threshold parameter \lstinline!dThresh!
-is used to discard those zero crossings that appear in regions with high numerical noise.
-Its default value is 1\% of the length of the Poincare section.
-The position of each detected zero crossing is refined by iteratively bisecting the interval of the zero crossing.
+As seen in Figure~\ref{f:Poincare return map}, LCS Tool uses a filtering parameter \lstinline!dThresh!
+to discard sign changes of $P(s)-s$ that are likely due to numerical sensitivity or a jump discontinuity of the Poincare map.
+Specifically, the location of each detected zero crossing is first refined by the bisection method.
 If after a predetermined number of iterations, \lstinline!nBisection!
-whose default value is five, the two points around the zero crossing have absolute values above \lstinline!dThresh!, the zero crossing is discarded.
-Once all valid closed lambda-line orbits have been located, the outermost orbit of every Poincare section is identified as an elliptic LCS.
+(default value 5) the two points around the zero crossing still have absolute values above \lstinline!dThresh!, the zero crossing is discarded.
+Once all valid closed $\lambda$-lines have been located, the outermost closed $\lambda$-orbit associated with every Poincare section is identified as a coherent Lagrangian vortex boundary.
 
@@
 \begin{center}
-\includegraphics[width=0.8\textwidth]{graphics/double_gyre/poincare_return_map}
+\includegraphics[width=.8\textwidth]{graphics/double_gyre/poincare_return_map}
 \end{center}
-\caption{Example of a Poincare return map obtained for an elliptic LCS. Circle markers indicate closed orbit positions. The filled circle indicates the outermost closed orbit and elliptic LCS.}
+\caption{Example of a Poincare map obtained for an elliptic LCS.
+Circle markers indicate closed orbit positions.
+The filled circle indicates the outermost fixed point of the Poincare map, marking the intersection of a coherent Lagrangian cortex boundary with the Poincare section.}
 \label{f:Poincare return map}
@@
 
-\subsection{Hyperbolic LCS}
+\subsection{Computing hyperbolic LCSs}
 
-The main steps of the algorithm used to calculate hyperbolic LCSs are enumerated in Table~\ref{t:Hyperbolic LCS algorithm}.
-The main function to compute hyperbolic LCS in LCS Tool is \lstinline!seed_curves_from_lambda_max!
+As discussed in section \ref{sub:Hyperbolic-LCSs}, positions of hyperbolic LCSs at time $t_{0}$ are found as the strongest repelling orbits of the vector field \eqref{eq:strainline} (repelling LCSs), and strongest attracting orbits of the vector field \eqref{eq:stretchline} (attracting LCSs.) By repulsion and attraction we mean a property of the LCS (as an evolving material line) under the flow map $F_{t_0}^t$.
+We identify the strongest repelling shrinklines as those crossing a local maximum of the $\lambda_2(x_0)$ field.
+Similarly, we identify the strongest attracting stretchlines as those crossing a local minimum of the $\lambda_1(x_0)$ field.
+These local maxima and minima of the appropriate $\lambda_i(x_0)$ eigenvalue field can be thought of as the extensions of the concept of a saddle points to the present finite-time, temporally aperiodic flow setting.
+
+The main steps of our hyperbolic LCS detection algorithm are enumerated in Table~\ref{t:Hyperbolic LCS algorithm}.
+The basic function to compute hyperbolic LCSs in LCS Tool is \lstinline!seed_curves_from_lambda_max!
 and its syntax is given in Table~\ref{t:seed_curves_from_lambda_max syntax}.
@@
 \item Define a local maximization distance.
-\item Find all local maxima of $\lambda_2$ on the main grid such that they are maxima within a circle with a radius equal to the local minimization distance.
-\item Define a maximum strainline length
-\item Integrate a strainline forward and backward according to Equation~\eqref{eq:strainline} and using the largest $\lambda_2$ local maximum as the initial position.
-Integrate until the strainline has attained the maximum strainline length, or until it has reached the domain boundary.
-\item Flag any remaining local maxima of $\lambda_2$ within the minimization distance of the strainline as ineligible initial positions for subsequent strainlines
-\item Continue integrating strainlines using local maxima of $\lambda_2$ as initial positions until no eligible local maxima of $\lambda_2$ remain.
-\item Remove all strainline segments within elliptic LCSs.
+\item Find all points of the main grid that are local maxima of $\lambda_2$ within a circle whose radius is the local  maximization distance.
+\item Define a maximum shrinkline length
+\item Integrate a shrinkline forward and backward according to Equation~\eqref{eq:strainline} and using the largest $\lambda_2$ local maximum as the initial position.
+Integrate until the shrinkline has attained the maximum shrinkline length, or until it has reached the domain boundary.
+\item Flag any remaining local maxima of $\lambda_2$ within the maximization distance of the shrinkline as ineligible initial positions for subsequent shrinklines
+\item Continue integrating shrinklines using local maxima of $\lambda_2$ as initial positions until no eligible local maxima of $\lambda_2$ remain.
+\item Remove all shrinkline segments within elliptic LCSs.
 \end{enumerate}
-\caption{Algorithm to calculate strainline hyperbolic LCSs. Algorithm for stretchline hyperbolic LCSs is similar.}
+\caption{Algorithm to calculate initial positions of repelling LCSs at time $t_0$. The algorithm for attracting LCSs is similar.}
 \label{t:Hyperbolic LCS algorithm}
@@
 \begin{tabular}{|c|p{.7\textwidth}|}
-\hline \hline
-\multicolumn{2}{|p{\textwidth}|}{\lstinline![curvePosition,curveInitialPosition] = seed_curves_from_lambda_max(distance,cgEigenvalue,cgEigenvector,flowDomain,flowResolution)!}\\
 \hline
-\lstinline!distance! & threshold distance for placement of lambda maxima\\
+\multicolumn{2}{|p{\textwidth}|}
+{\lstinline![curvePosition,curveInitialPosition] = seed_curves_from_lambda_max(distance,cgEigenvalue,cgEigenvector,flowDomain,flowResolution)!}\tabularnewline
 \hline
-\lstinline!cgEigenvalue! & array of Cauchy-Green strain eigenvalues\\
+\lstinline!distance! & threshold distance for placement of $\lambda_2(x_0)$ maxima\tabularnewline
 \hline
-\lstinline!cgEigenvector! & array of Cauchy-Green strain eigenvectors\\
+\lstinline!cgEigenvalue! & array of Cauchy-Green strain eigenvalues\tabularnewline
 \hline
-\lstinline!flowDomain! & $2 \times 2$ array to define flow domain\\
+\lstinline!cgEigenvector! & array of Cauchy-Green strain eigenvectors\tabularnewline
 \hline
-\lstinline!flowResolution! & $1 \times 2$ array to define Cauchy-Green strain main grid resolution\\
+\lstinline!flowDomain! & $2 \times 2$ array to define flow domain\tabularnewline
 \hline
-\lstinline!periodicBc! & optional $1 \times 2$ logical array to specify periodic boundary conditions. Default is \lstinline![false,false]!.\\
+\lstinline!flowResolution! & $1 \times 2$ array to define Cauchy-Green strain main grid resolution\tabularnewline
 \hline
-\lstinline!nMaxCurves! & optional maximum number of curves (i.e. strainlines of stretchlines) to generate. Default is \lstinline!numel(cgEigenvalue)!.\\
+\lstinline!periodicBc! & optional $1 \times 2$ logical array to specify periodic boundary conditions. Default is \lstinline![false,false]!.\tabularnewline
 \hline
-\lstinline!odeSolverOptions! & optional \lstinline!odeset! structure to specify flow map integration parameters\\
-\hline \hline
+\lstinline!nMaxCurves! & optional maximum number of curves (i.e. shrinklines of stretchlines) to generate. Default is \lstinline!numel(cgEigenvalue)!.\tabularnewline
+\hline
+\lstinline!odeSolverOptions! & optional \lstinline!odeset! structure to specify flow map integration parameters\tabularnewline
+\hline
 \end{tabular}
@@
 
-Strainlines, trajectories of the $\boldsymbol \xi_1$ vector field, are integrated starting from local maxima of $\lambda_2$ on the main grid.
-The local maximization distance \lstinline!distance!
-is defined.
-This distance is set to a value between two and five grid points, since it represents the number of grid points over which the Cauchy-Green strain tensor is susceptible to numerical noise from integration.
-Thus, the local maximization distance is expected to be a function of variable time step integration tolerances.
-A record is made of all grid points where $\lambda_2$ is maximal within a circle whose radius equals the defined maximization distance.
-Then the first strainline is integrated from the largest of all the local maxima in both directions, i.e., integrating $\boldsymbol \xi_1$ and $-\boldsymbol \xi_1$.
-Similarly to lambda-lines, this integration must take into account eigenvector field orientation discontinuities.
-Prior to initiating strainline integration, a maximum length is defined.
-Strainlines are then integrated until reaching the domain boundary or the maximum length.
-The maximum length may be seen as safety factor to prevent highly sinuous strainlines leading to inordinately long computations; the maximum length should be set such that most strainlines stop at the domain boundary.
-Once a strainline position has been calculated, local maxima of $\lambda_2$ within the minimization distance are flagged as ineligible initial positions for subsequent strainlines.
-After the first strainline has been obtained and any ineligible local maxima of $\lambda_2$ have been flagged, the next strainline is integrated starting form the highest eligible maximum of $\lambda_2$.
-The process continues until all eligible local maxima of $\lambda_2$ have been used to initialize strainline integration.
-These calculations are performed by the function \lstinline!seed_curves_from_lambda_max!.
-Note that once a strainline integration has begun, it may progress so as to be within the local minimization distance of an existing strainline.
-In other words, the local minimization distance controls where strainlines begin, but not where they stop.
-
-A further point important to note is the existence of hyperbolic LCS inside elliptic LCS.
-This may seem contradictory when accustomed to LCSs not crossing, however, numerical experimentation with LCS calculations overlaid with color tracer advection reveals that hyperbolic LCS can cross elliptic barriers and have an impact on the tracer pattern inside elliptic regions.
-At crossing points a line segment of an elliptic LCS, oriented along $\boldsymbol \eta_\lambda^{\pm}$, does not stretch, while a segment of a strainline, oriented along $\boldsymbol \xi_1$, repels fluid normally at a maximal rate and shrinks at a maximal rate.
-Figure~\ref{fig:ocean_dataset_colortracer} illustrates the effect of strainlines on tracer particles inside elliptic regions.
-Color-coded final positions show large gradients along strainlines, directly showing that particles are strongly repelled by the LCS and therefore end in different positions.
+In classical, infinite-time dynamical systems, elliptic and hyperbolic invariant manifolds cannot intersect, as the trajectories in their intersection would then have to follow two different asymptotic behaviors simultaneously.
+In finite-time dynamical systems, hyperbolic LCSs may well intersect elliptic LCSs, with hyperbolic LCSs continuing to act as cores of mixing patterns observed in the interior of elliptic LCSs.
+Motivated by convention and for clarity in visualization, however, the default setting of LCS Tool removes hyperbolic LCS segments from the interior of elliptic LCSs.
+These removed segments can be turned back on demand, as shown in Figure~\ref{fig:ocean_dataset_colortracer}.
 
@@
 \centering
-\includegraphics[width=0.49\textwidth]{graphics/ocean_dataset/LCS_fwd_coherent_eddy}
-\includegraphics[width=0.49\textwidth]{graphics/ocean_dataset/LCS_fwd_colortracer}
-\caption{(Left) Elliptic LCS (green) with hyperbolic normally repelling LCS (red) inside.
-(Right) Final positions $\boldsymbol x_t = F_{t_0}^t(\boldsymbol x_0)$ of tracer particles are plotted with the x-position encoded in red and y-position encoded in green color.
-The impact of the repulsive property of hyperbolic LCS on the tracer particles can not only be observed outside, but also inside the elliptic region.
-Hyperbolic LCS mark lines of high color gradients, corresponding to tracer particles that start on both sides of the LCS, are repelled, and end up in distinct locations.}
+\includegraphics[width=.49\textwidth]{graphics/ocean_dataset/LCS_fwd_coherent_eddy}
+\includegraphics[width=.49\textwidth]{graphics/ocean_dataset/LCS_fwd_colortracer}
+\caption{(Left) Elliptic LCS (green) with repelling LCSs (red) continuing into its interior.
+(Right) Final positions $x(t) = (x_1(t),x_2(t)) = F_{t_0}^t(x_0)$ of tracer particles, are plotted with their $x_1(t)$ position encoded in red and $x_2(t)$ position encoded in green.
+Note the continued impact of the repelling hyperbolic LCSs (black) on tracer patterns inside the elliptic LCS (white.)}
 \label{fig:ocean_dataset_colortracer}
@@
 \section{Examples}
-In this section we present examples showing how to use LCS Tool to obtain LCSs in different flows, namely a double gyre, a jet, and an oceanic geostrophic flow.
+
+In this section we present examples of the use of LCS Tool to obtain LCSs in three different examples: a double gyre, a jet, and an oceanic geostrophic flow.
 All three examples are available as demonstration files in the demo folder of LCS Tool.
-By executing these scripts, readers can follow LCS Tool computations step-wise.
+By executing these scripts, readers can follow LCS Tool computations step by step.
 
@@
 The double gyre is a model for a time-dependent two gyre system observed in geophysical flows \parencite{shadden05:_defin_lagran_lyapun}.
-It consists of two counter rotating sinusoidal vortices with a harmonically oscillating line in-between.
-The velocity equations are:
+The model consists of two counter rotating sinusoidal vortices with a harmonically oscillating line in-between.
+Lagrangian particle motions satisfy the non-autonomous dynamical system
 \begin{equation}
 \begin{split}
-\dot x = -\pi A \sin[\pi f(x,t)] \cos(\pi y)\\
-\dot y = \pi A \cos[\pi f(x,t)] \sin(\pi y) \frac{\partial f(x,t)}{\partial x}\\
+\od{x}{t} = -\pi A \sin[\pi f(x,t)] \cos(\pi y),\\
+\od{y}{t} = \pi A \cos[\pi f(x,t)] \sin(\pi y) \pd{f(x,t)}{x},\\
 f(x,t) = \epsilon \sin(\omega t) x^2 + [1 - 2 \epsilon \sin(\omega t)] x.
@@
 \end{equation}
-The MATLAB function corresponding to these equations is given in Listing~\ref{l:double gyre derivative} and shows how to specify a derivative function that supports vector form integration.
+The MATLAB function describing this model velocity field is given in Listing~\ref{l:double gyre derivative}, specifying the right hand side of the particle ODE in a way that supports vectorized integration.
 
@@
 
-In what follows, the parameter values are: $A = 0.1$, $\epsilon = 0.1$, $\omega = \pi/5$. The flow timespan is $t \in [0,5]$ and the domain is $x \in [0,2]$, $y \in [0,1]$.
+In what follows, the parameter values are: $A = 0.1$, $\epsilon = 0.1$, $\omega = \pi/5$.
+The flow timespan is $t \in [0,10]$ and the domain is $x \in [0,2]$, $y \in [0,1]$.
+By examining the FTLE field (which
+is optionally provided by LCS Tool), we position Poincare sections to capture elliptic LCSs.
+An LCS Tool demo script to perform this operation is given in Listing~\ref{l:double gyre lambda-line LCS} where two Poincare sections are defined.
+The free stretching parameter $\lambda$ (Equation \eqref{eq:etafields}) is varied over the range $[0.93,1.07]$ with increments of $0.01$.
+In a loop , closed orbits for all predefined $\lambda$ values are computed and the outermost closed orbit is kept as the Lagrangian vortex boundary.
+In this script, we have chosen small values for the error tolerances of the integration of the Cauchy-Green strain tensor and $\lambda$-lines to facilitate convergence.
 
-By examining the FTLE, we can position Poincare sections to detect lambda-line LCSs.
-An LCS Tool script to perform this operation is given in Listing~\ref{l:double gyre lambda-line LCS} where two Poincare sections are defined.
-In the script we have set error tolerances for integration of the Cauchy-Green strain tensor and lambda lines.
-These numerical parameters help achieve convergence.
+\lstinputlisting[caption={Double gyre script for elliptic lambda-line LCSs},label=l:double gyre lambda-line LCS]{listings/double_gyre/lambda_lcs.m}
 
-\lstinputlisting[caption={Double gyre script for lambda-line LCSs},label=l:double gyre lambda-line LCS]{listings/double_gyre/lambda_lcs.m}
-
-In Figure~\ref{fig:double_gyre_lambda_lcs_convergence} the resolution of the  Cauchy-Green strain tensor is varied from $500 \times 250$ to $1000 \times 500$.
+In Figure~\ref{fig:double_gyre_lambda_lcs_convergence}, the resolution of the  Cauchy-Green strain tensor is varied from $500 \times 250$ to $1000 \times 500$.
 The location of the outermost closed lambda-line changes little, demonstrating convergence.
+For all tested resolutions the $\lambda$ values of the outermost closed orbits are constant.
+This is also suggests that the lowest tested resolution of $500 \times 250$ is sufficient to identify the elliptic LCS in this example.
 
@@
 
-We continue the example by adding the calculation of hyperbolic strainline and stretchline LCSs.
-Figures~\ref{fig:double_gyre_lambda_strain_lcs} and~\ref{fig:double_gyre_lambda_stretch_lcs} show the results and Listing~\ref{lst:double_gyre_lambda_strain_stretch_lcs} gives the LCS Tool commands.
-The local maximization distance is set larger for stretchlines than for strainlines usually.
-This is because strainlines follow ridges of $\lambda_2$ maxima whereas stretchlines do not follow ridges of $\lambda_1$ minima.
+Figure \ref{fig:double_gyre_lambda_strain_lcs} shows the complete picture of elliptic and hyperbolic LCSs of the double gyre at this resolution.
+Listing~\ref{lst:double_gyre_lambda_strain_stretch_lcs} lists the corresponding LCS Tool commands.
+The maximal length of shrinklines and stretchlines, \lstinline!strainlineMaxLength!
+and \lstinline!stretchlineMaxLength!, is set to $20$, a multiple of the domain size, since the hyperbolic LCSs may wind around vortices several times.
+
+As usual, the local maximization distance is set larger for stretchlines than for shrinklines (cf.
+l.\,4 and l\,9).
+The purpose of the maximization distance is to obtain spatially separated LCSs and to avoid a dense tangle of lines that basically indicates the same hyperbolic LCS (cf.
+Table \ref{t:Hyperbolic LCS algorithm}, point 5).
+Setting the local maximization distance for stretchlines larger than for shrinklines allows obtaining a comparable number of stretchlines and shrinklines overall in the flow domain (recall that hyperbolic LCS seed points are discarded if they are within the local maximization distance of an existing hyperbolic LCS.) Shrinklines are locally tangent to ridges of $\lambda_2$ maxima, whereas stretchlines are locally normal to these ridges.
+Setting the local maximization distance of stretchlines and shrinklines equal would therefore produce a greater number of stretchlines than shrinklines overall.
+
+\lstinputlisting[caption={LCS Tool commands for double gyre hyperbolic LCSs.
+Condensed version of the LCS Tool script \lstinline!demo/double_gyre/hyperbolic_shear_lcs.m!
+},label=lst:double_gyre_lambda_strain_stretch_lcs]{listings/double_gyre/lambda_strain_stretch_lcs.m}
 
@@
 \includegraphics[width=\textwidth]{graphics/double_gyre/lambda_strain_lcs}
-  \caption{Double gyre lambda-line and strainline LCSs}
+\caption{Double gyre lambda-line and shrinkline LCSs}
 \label{fig:double_gyre_lambda_strain_lcs}
@@
 
-\lstinputlisting[caption={LCS Tool commands for double gyre strainline and stretchline LCSs.
-Condensed version of the LCS Tool script \lstinline!demo/double_gyre/hyperbolic_shear_lcs.m!
-},label=lst:double_gyre_lambda_strain_stretch_lcs]{listings/double_gyre/lambda_strain_stretch_lcs.m}
-
 \clearpage
@@
 The Bickley jet models a meandering zonal jet flanked above and below by
-counter rotating vortices. This is an idealized model of geophysical flows
-such as the polar night jet perturbed by a Rossby
-wave\parencite{haller12:_geodes_theor_trans_barrier_two_dimen_flows,beron-vera10:_invar_lagran}.
+counter rotating vortices.
+This is an idealized model of geophysical flows such as the Gulf Stream and the polar night jet perturbed by a Rossby wave\parencite{castilloNegrete93:_chaot_rossb,beron-vera10:_invar_lagran}.
 
-The velocity is given by $\boldsymbol{v}(x,y,t) = (-\partial_y \psi, \partial_x \psi)$ where
+The velocity is given by $v(x,y,t) = (-\partial_y \psi, \partial_x \psi)$ where
 \begin{gather*}
@@
 \end{gather*}
-As a forcing function we choose a chaotic solution of the Duffing oscillator, specifically
+As a forcing function, we choose a solution running on the chaotic attractor of the damped and forced Duffing oscillator, specifically
 \begin{gather*}
-\frac{d \phi_1}{dt} = \phi_2,\\
-\frac{d \phi_1}{dt} = -0.1 \phi_2 - \phi_1^3 + 11 \cos(t),\\
+\od{\phi_1}{t} = \phi_2,\\
+\od{\phi_1}{t} = -0.1 \phi_2 - \phi_1^3 + 11 \cos(t),\\
 f_{1,2}(t) = 2.625 \times 10^{-2} \phi_1(t/6.238 \times 10^5)
@@
 
+The integration time is $T=4L_x/U$, a multiple of a the eddy turnover
+time in the flow.
 Figure~\ref{fig:Bickley_jet_lambda_LCS_full} shows lambda-line LCS of the Bickley jet with the FTLE in the background.
@@
 
-Strainline LCSs are shown in Figure~\ref{fig:Bickley_jet_strain_lcs} and stretchline LCSs are shown in Figure~\ref{fig:Bickley_jet_stretch_lcs}.
+Shrinkline LCSs are shown in Figure~\ref{fig:Bickley_jet_strain_lcs} and stretchline LCSs are shown in Figure~\ref{fig:Bickley_jet_stretch_lcs}.
 
@@
 \includegraphics[width=.85\textwidth]{graphics/bickley_jet/strain_lcs}
-  \caption{Bickley jet strainline LCSs}
+\caption{Bickley jet shrinkline LCSs}
 \label{fig:Bickley_jet_strain_lcs}
@@
 
-For the last example, corresponding to the ocean dataset demo files in LCS Tool, we choose a measured geophysical flow.
-In contrast to the previous analytic examples, the two dimensional incompressible velocity field is presented as a dataset with discrete temporal and spatial resolution.
-Our region of interest is a small domain in the South Atlantic Ocean, where exceptionally coherent eddies, Agulhas rings, were recently found using methods provided by LCS Tool \parencite{haller13:_coher_lagran}.
-The oceanic velocity field is obtained from altimetric satellite measurements of the sea surface height under the assumption that the flow is geostrophic.
-Under this assumption the sea surface height $\eta$ serves as a stream-function and in a longitude-latitude $(\varphi,\theta)$ coordinate system the evolution of a fluid particle is given by
+This last example demonstrates the use of LCS Tool on velocity data derived from satellite-observed sea-surface heights under the geostrophic approximation.
+In contrast to the previous two analytic examples, the velocity field is available only with discrete temporal and spatial resolution.
+Our region of interest is a small domain in the South Atlantic Ocean, where exceptionally coherent eddies, Agulhas rings, were recently found by \parencite{haller13:_coher_lagran} using the theory we surveyed in Section \ref{sub:Elliptic-LCSs}.
+
+In the geostrophic approximation, the sea-surface height $\eta$ serves as a stream-function for surface velocities.
+In a longitude-latitude $(\varphi,\theta)$ coordinate system, the evolution of a fluid particle is given by
 \begin{eqnarray}
-\dot{\varphi}(\varphi,\theta,t) = -\frac{g}{R^2 f(\theta) \cos\theta}\partial_{\theta}\eta(\varphi,\theta,t)\label{eq:dtlon}\\
-\dot{\theta}(\varphi,\theta,t) = +\frac{g}{R^2 f(\theta) \cos\theta}\partial_{\varphi}\eta(\varphi,\theta,t)
+\pd{\varphi(\varphi,\theta,t)}{t} = -\frac{g}{R^2 f(\theta) \cos\theta} \pd{\eta(\varphi,\theta,t)}{\theta}\label{eq:dtlon}\\
+\pd{\theta(\varphi,\theta,t)}{t} = \frac{g}{R^2 f(\theta) \cos\theta}\pd{\eta(\varphi,\theta,t)}{\varphi}
 \label{eq:dtlat}
 \end{eqnarray}
-where $g$ is the gravitational acceleration, $R$ is the mean radius of Earth, and $f(\theta) \equiv 2\Omega\sin\theta$ is the Coriolis parameter with $\Omega$ being the Earth's mean angular velocity.
+where $g$ is the constant of gravity, $R$ is the mean radius of the Earth, and $f(\theta) \equiv 2\Omega\sin\theta$ is the Coriolis parameter, with $\Omega$ denoting the Earth's mean angular velocity.
 
-The data is given at a spatial resolution of $1/4\degree$ and a temporal resolution of $7\,\mathrm{days}$.
-Due to the discrete data, defining the right hand side of Equation~\eqref{eq:dtlon} and~\eqref{eq:dtlat} involves interpolation in space and time with a spline interpolation scheme.
+The data is given at a spatial resolution of $1/4°$ and a temporal resolution of $7\,\mathrm{days}$.
+Due to the discrete data, defining the right hand side of Equation~\eqref{eq:dtlon} and~\eqref{eq:dtlat} involves spline interpolation in space and time.
 An interpolant is generated first, then the function \lstinline!flowdata_derivative!
 evaluates the interpolants for the zonal and meridional velocity at the needed coordinates.
-Listing \ref{l:griddedInterp} shows large parts of the code in LCS Tool's ocean demo file \lstinline!hyperbolic_shear_lcs.m!.
-In lines 8-11 the commands for the interpolation are given.
+Listing \ref{l:griddedInterp} shows the relevant part of the code in LCS Tool's ocean demo file \lstinline!hyperbolic_shear_lcs.m!.
+In lines 8-11, the commands for the interpolation are given of the velocity dataset are given.
 
-\begin{lstlisting}[caption={LCS ocean demo: Parts of the demo script included in LCS Tool to compute LCS from an ocean data set},label=l:griddedInterp]
+\begin{lstlisting}[caption={LCS ocean demo: Parts of the demo script included in LCS Tool to compute LCS from an ocean dataset},label=l:griddedInterp]
 % Input parameters
@@
 
-For the ocean data, the integration time was chosen to be $T=30$ days (Listing \ref{l:griddedInterp}, line 4), a time larger than the turn over time of the mesoscale eddies in the domain.
-The resolution of the grid of initial conditions is set to $400\times400$ (line 3) which corresponds to a resolution of roughly 0.015\degree\, and gives good results.
-With this choice the resolution of the tracer grid is of the order of 15 times higher than the resolution of the velocity field.
+We choose the integration time as $T=30$ days (Listing \ref{l:griddedInterp}, line 4), which is larger than the eddy turnover time in this region.
+The resolution of the main computational grid for initial conditions is set to $400 \times 400$ (line 3.) This corresponds to a resolution of roughly 0.015° and gives good results.
+With this choice, the resolution of the tracer grid is 15 times higher than the resolution of the velocity field.
 The flow is integrated and the Cauchy-Green strain tensor is computed by the function \lstinline!eig_cgStrain!
@@
 Incompressibility of the flow is enforced (line 12), the auxiliary grid distance is set to 1\% of the main grid distance (line 17), and eigenvalues are computed from the auxiliary grid (line 16).
-Lambda-line LCS are computed in line 54, after the Poincare section have been set (line 41-42) and the $\boldsymbol\eta_\pm$ fields have been defined (line 53).
-Hyperbolic LCS, strainlines and strechlines, are computed in lines 57 and 59, respectively.
+Elliptic LCSs are computed in line 54, after the Poincare sections have been set (line 41-42) and the $\eta_\pm^\lambda$ fields have been defined (line 53).
+Hyperbolic LCSs are computed in lines 57 and 59.
 
-Figures~\ref{f:ocean dataset hyperbolic shear lcs details strainline} and \ref{f:ocean dataset hyperbolic shear lcs details stretchline} show elliptic and hyperbolic LCS on 22 November 2006, i.e., the same time instant shown in~\textcite{haller13:_coher_lagran,beron-vera13:_objec_agulh}.
-Note that here the integration time is $T=30$ days, as opposed to 90 days in their analysis.
-Since for 90 days the hyperbolic LCSs would be a confusing entanglement of lines, we choose a shorter integration time here, to show hyperbolic LCS together with elliptic LCSs.
-For the shorter integration time, the analysis with LCS Tool reveals a large coherent eddy bounded by a closed lambda-line at $(3,-32)$ which corresponds to eddy \#2 in Figure~3 of \textcite{beron-vera13:_objec_agulh}.
-Additionally, a further smaller coherent eddy core at approximately $(1,-31)$ is found that does not stay coherent over a time of 90 days.
-Both closed orbits are stable if the resolution for the Cauchy-Green strain tensor field is increased.
-Strainlines (red) and stretchlines (blue) mark the core lines of strong normal attraction and repulsion.
+Figures~\ref{f:ocean dataset hyperbolic shear lcs details strainline} and \ref{f:ocean dataset hyperbolic shear lcs details stretchline} show the positions of elliptic and hyperbolic LCSs on 22 November 2006, which is the same time instant analyzed in~\textcite{haller13:_coher_lagran,beron-vera13:_objec_agulh}.
+Here the integration time is chosen as $T = 30$ days, as opposed to 90 days in the cited references, to avoid the length tangling of hyperbolic LCSs.
+Our analysis via LCS Tool reveals five coherent eddy boundaries.
+The large coherent eddy at $(3,-32)$ has a non-stretching boundary, i.e., $\lambda = 1$.
+It corresponds to eddy \#2 in Figure~3 of \textcite{beron-vera13:_objec_agulh}.
+Additionally, four further smaller coherent eddy cores are found.
+They do not stay coherent over a time of 90 days.
+All closed orbits are stable under increased spatial resolution for the Cauchy-Green strain tensor field.
+Repelling LCSs (red) and attracting LCSs (blue) are superimposed over the plot of FTLE values.
+The hyperbolic LCSs determine the deformation of the fluid in-between coherent eddy cores.
 
@@
 \begin{center}
-\includegraphics[width=0.8\textwidth]{graphics/ocean_dataset/hyperbolic_shear_lcs_strainlines}
+\includegraphics[width=.8\textwidth]{graphics/ocean_dataset/hyperbolic_shear_lcs_strainlines}
 \end{center}
-\caption{Lambda-line LCSs (green) and strainline LCSs (red) of the ocean data set.
+\caption{Elliptic LCSs (green) and repelling LCSs (red) in the ocean dataset.
 Background gray-scale is the finite time Lyapunov exponent.
 Dashed green lines are Poincare sections.
-White dots indicate $\lambda_2$ local maxima used for strainline integration initial positions.}
+White dots indicate $\lambda_2$ local maxima used as initial positions for shrinkline integration.}
 \label{f:ocean dataset hyperbolic shear lcs details strainline}
@@
 \end{center}
-\caption{Lambda-line LCSs (green) and stretchline LCSs (blue) of the ocean data set.
+\caption{Elliptic LCSs (green) and attracting LCSs (blue) in the ocean dataset.
 Background gray-scale is the finite time Lyapunov exponent.
 Dashed green lines are Poincare sections.
-White dots indicate $\lambda_1$ local minima used for stretchline integration initial positions.}
+White dots indicate $\lambda_1$ local minima used as initial positions for stretchline integration.}
 \label{f:ocean dataset hyperbolic shear lcs details stretchline}
@@
 
-Algorithms to identify LCSs using variational principles have been presented, and a software library, LCS Tool, that implements these algorithms has been introduced and demonstrated on analytic flow models and a geophysical dataset.
-The software library provides functions to enable exploring variational LCS methods without requiring detailed knowledge of variational LCS theory.
-These functions make it possible to write compact programs to identify and visualize LCSs.
-Since LCS Tool is programmed in MATLAB, testing new methods and visualizing their results can be done quite quickly.
+We have described a computational toolbox, LCS Tool, that implements recent variational results for Lagrangian Coherent Structures (LCSs) in two-dimensional unsteady flows.
+We have also demonstrated the performance of LCS Tool on two analytic flow models and a geophysical velocity dataset.
+The publicly available software library producing these results enables the exploration of variational LCS methods without assuming a detailed knowledge of geodesic LCS theory.
 
-LCS Tool is built atop MATLAB and leverages that software's capabilities extensively.
-With FTLE-based LCSs, computational performance has received considerable attention\parencite{conti12:_gpu_apu_finit_time_lyapun_expon,miron12:_anisot_lagran_coher_struc}.
-LCS Tool is free software and we hope it will evolve to incorporate similar advances.
-Optimizing computational performance will likely help apply variational LCS methods to large-scale real-time forecasting applications, for example to allow tracking environmental contaminants.
+LCS Tool leverages the capabilities of MATLAB extensively.
+For FTLE based extraction of LCSs, computational performance has received considerable attention \parencite{conti12:_gpu_apu_finit_time_lyapun_expon,miron12:_anisot_lagran_coher_struc}.
+We hope to see similar computational advances for the publicly available LCS Tool platform.
+Optimizing computational performance will likely aid the application of variational LCS methods to large-scale real-time forecasting applications, such as the tracking of environmental contaminants\parencite{olascoaga12:_forec}.
 
-In addition to improving computational performance, we hope LCS Tool will be the foundation to integrate new theoretical advances, for example the variational LCS jet core identification method\parencite{farazmand13:_shearless} and the analysis of LCSs in higher dimensions\parencite{blazevski:_hyper_ellip_trans_barrier_three}.
+We also hope that LCS Tool will serve as a foundation for the numerical
+implementation of recent theoretical advances.
+These include the geodesic theory of parabolic LCSs (jet cores)\parencite{farazmand14:_shearless} and the variational theory of hyperbolic and elliptic LCSs for three-dimensional
+flows\parencite{blazevski14:_hyper_ellip_trans_barrier_three}.
 
@@
 % End:
-% LocalWords:  stretchline stretchlines incompressible gyre eigenvector eigenvectors incompressibility dataset datasets LCS LCSs strainline Lyapunov Bickley Rossby timespan variational adaptively vortices parametrise advected strainlines minima extremal FTLE MATLAB maxima geostrophic
+% LocalWords:  stretchline stretchlines incompressible gyre eigenvector eigenvectors incompressibility dataset datasets LCS LCSs strainline Lyapunov Bickley Rossby timespan variational adaptively vortices parametrise advected strainlines minima extremal FTLE MATLAB maxima geostrophic shrinkline shrinklines

### 2014-08-23 main.tex
Correct spelling and grammar in main.tex
Commit da33cca incorporated changes made in the arXiv e-print
including mistakes. The present commit corrects those mistakes.

@@
 
-Lagrangian Coherent Structures (LCSs) are evolving organizing centers of trajectory patterns in non-autonomous dynamical systems\parencite{haller00:_lagran,peacock13:_lagran,Haller2014}.
+Lagrangian Coherent Structures (LCSs) are evolving organizing centers of trajectory patterns in non-autonomous dynamical systems\parencite{haller00:_lagran,peacock13:_lagran,haller15:_langr_coher_struc}.
 Applications of LCSs include oceanic and atmospheric flows\parencite{beron-vera13:_objec_agulh,koh02:_hyper}, biological transport problems\parencite{wilson09:_lagran_reynol,tallapragada11:_lagran,huhn12:_south_indian_ocean_count_madag}, aeronautics\parencite{tang10:_accur_lagran_hong_kong_inter_airpor}, celestial mechanics\parencite{gawlik09:_lagran}, crowd dynamics\parencite{ali07:_lagran_partic_dynam_approac_crowd}, and aperiodically forced mechanical oscillators\parencite{hadjighasem13:_detec_kam}.
 
-\textcite{haller01:_distin} proposed that ridges of the Finite-Time Lyapunov exponent (FTLE) are heuristic indicators of hyperbolic (i.e., repelling- and attracting-type) LCSs.
+\textcite{haller01:_distin} proposed that ridges of the finite-time Lyapunov exponent (FTLE) are heuristic indicators of hyperbolic (i.e., repelling- and attracting-type) LCSs.
 A number of examples indeed support this principle\parencite{peacock13:_lagran}.
@@
 \end{equation}
-measuring Lagrangian strain in the velocity field.
-This tensor is symmetric and positive definite\parencite{truesdell04}.
-We label the eigenvalues and eigenvectors of $C_{t_0}^t(\boldsymbol x_0)$ as follows:
+measuring Lagrangian strain in the velocity field. This tensor is symmetric and positive definite\parencite{truesdell04}. We label the eigenvalues and eigenvectors of $C_{t_0}^t(x_0)$ as follows:
 \[
@@
 \frac{x_1(t;t_0,x_0 + \delta x_1) - x_1(t;t_0,x_0 - \delta x_1)}{2\vert\delta x_1\vert} & \frac{x_1(t;t_0,x_0+\delta x_2) - x_1(t;t_0,x_0-\delta x_2)}{2\vert\delta x_2\vert}\\
-\frac{x_2(t;t_0,x_0+\delta x_1) - x^2(t;t_0,x_0 - \delta x_1)}{2\vert\delta x_1\vert} & \frac{x_2(t;t_0,x_0 + \delta x_2) - x_2(t;t_0,x_0 - \delta x_2)}{2\vert\delta x_2\vert}
+\frac{x_2(t;t_0,x_0+\delta x_1) - x_2(t;t_0,x_0 - \delta x_1)}{2\vert\delta x_1\vert} & \frac{x_2(t;t_0,x_0 + \delta x_2) - x_2(t;t_0,x_0 - \delta x_2)}{2\vert\delta x_2\vert}
 \end{array}\right)
@@
 \end{center}
-\caption{Illustration of the main grid (filled circles) and the auxiliary grid (empty circles) used in the computation of the derivative of the flow map in the definition \eqref{eq:CG} of the Cauchy--Green stain tensor.
+\caption{Illustration of the main grid (filled circles) and the auxiliary grid (empty circles) used in the computation of the derivative of the flow map in definition \eqref{eq:CG} of the Cauchy-Green strain tensor.
 The variable \lstinline!auxGridRelDelta!
@@
 The function \lstinline!eig_cgStrain!
-of LCS Tool provides the option to calculate Cauchy--Green eigenvectors from the auxiliary grid using eigenvalues calculated from the main grid.
+of LCS Tool provides the option to calculate Cauchy-Green eigenvectors from the auxiliary grid using eigenvalues calculated from the main grid.
 We have found that for flows defined analytically, the eigenvalues can reliably be calculated from the main grid.
-For the flows defined by datasets, using the auxiliary grid for eigenvalue calculations gives better results.
+For flows defined by datasets, using the auxiliary grid for eigenvalue calculations gives better results.
 
-As mentioned above, a typical main grid for the Cauchy-Green strain tensor comprises $500 \times 500$.
-This means that after the addition of 4 auxiliary grid points around each main grid points, eq.
+As mentioned above, a typical main grid for the Cauchy-Green strain tensor comprises $500 \times 500$ points.
+This means that after the addition of 4 auxiliary grid points around each main grid point, eq.
 \eqref{eq:v(x,t)} must be integrated over 1.25 million initial conditions.
 
-To avoid excessive computational times in MATLAB, we vectorize \eqref{eq:v(x,t)}, i.e., combine its right-hand-side evaluated over each initial points into a single system of equations.
+To avoid excessive computational times in MATLAB, we vectorize \eqref{eq:v(x,t)}, i.e., combine its right-hand-side evaluated over each initial point into a single system of equations.
 The resulting system is composed of independent blocks of two-dimensional first-order ODEs.
@@
 
-The Cauchy-Green strain tensor generically admits points singularities, i.e., points where $C_{t_0}^t(x_0)$ has repeated eigenvalues.
+The Cauchy-Green strain tensor generically admits point singularities, i.e., points where $C_{t_0}^t(x_0)$ has repeated eigenvalues.
 At these points the eigenvectors $\xi_1(x_0)$ and $\xi_2(x_0)$ are no longer well-defined.
@@
 \hline
-\lstinline!nBisection! & optional number of bisection steps to refine zero crossings of Poincare map. Default values: 5.\tabularnewline
+\lstinline!nBisection! & optional number of bisection steps to refine zero crossings of Poincare map. Default value: 5.\tabularnewline
 \hline
@@
 Similarly, we identify the strongest attracting stretchlines as those crossing a local minimum of the $\lambda_1(x_0)$ field.
-These local maxima and minima of the appropriate $\lambda_i(x_0)$ eigenvalue field can be thought of as the extensions of the concept of a saddle points to the present finite-time, temporally aperiodic flow setting.
+These local maxima and minima of the appropriate $\lambda_i(x_0)$ eigenvalue field can be thought of as the extensions of the concept of saddle points to the present finite-time, temporally aperiodic flow setting.
 
@@
 \end{equation}
-The MATLAB function describing this model velocity field is given in Listing~\ref{l:double gyre derivative}, specifying the right hand side of the particle ODE in a way that supports vectorized integration.
+The MATLAB function describing this velocity field is given in Listing~\ref{l:double gyre derivative}, specifying the right hand side of the particle ODE in a way that supports vectorized integration.
 
@@
 For all tested resolutions the $\lambda$ values of the outermost closed orbits are constant.
-This is also suggests that the lowest tested resolution of $500 \times 250$ is sufficient to identify the elliptic LCS in this example.
+This is also suggests that the lowest tested resolution of $500 \times 250$ is sufficient to identify the elliptic LCSs in this example.
 
@@
 
-This last example demonstrates the use of LCS Tool on velocity data derived from satellite-observed sea-surface heights under the geostrophic approximation.
+The final example demonstrates the use of LCS Tool on velocity data derived from satellite-observed sea-surface heights under the geostrophic approximation.
 In contrast to the previous two analytic examples, the velocity field is available only with discrete temporal and spatial resolution.
@@
 Figures~\ref{f:ocean dataset hyperbolic shear lcs details strainline} and \ref{f:ocean dataset hyperbolic shear lcs details stretchline} show the positions of elliptic and hyperbolic LCSs on 22 November 2006, which is the same time instant analyzed in~\textcite{haller13:_coher_lagran,beron-vera13:_objec_agulh}.
-Here the integration time is chosen as $T = 30$ days, as opposed to 90 days in the cited references, to avoid the length tangling of hyperbolic LCSs.
+Here the integration time is chosen as $T = 30$ days, as opposed to 90 days in the cited references, to avoid tangling of hyperbolic LCSs.
 Our analysis via LCS Tool reveals five coherent eddy boundaries.
@@
 It corresponds to eddy \#2 in Figure~3 of \textcite{beron-vera13:_objec_agulh}.
-Additionally, four further smaller coherent eddy cores are found.
+Additionally, four smaller coherent eddy cores are found.
 They do not stay coherent over a time of 90 days.

### 2014-08-23 main.tex
Change grammar and style in main.tex

@@
 LCSs are most repelling, attracting and shearing material lines that form the centerpieces of observed tracer patterns in two-dimensional unsteady dynamical systems.
-LCS Tool implements the latest geodesic theory of LCSs for two-dimensional flows, uncovering key transport barriers in unsteady flow velocity data as explicit solutions of differential equations.
-After a review of the underlying theory, we explain the steps and numerical methods used by LCS Tool, and illustrate its capabilities on three unsteady fluid flow examples.
+LCS Tool implements the latest geodesic theory for LCSs in two-dimensional flows, uncovering key transport barriers in unsteady flow velocity data as explicit solutions of differential equations.
+After a review of the underlying theory, we explain the numerical algorithms used by LCS Tool, and illustrate its capabilities on three unsteady fluid flow examples: a double gyre, a Bickley jet and an ocean dataset for the Agulhas current.
 }
@@
 Equating FTLE ridges with LCSs, however, would create theoretical inconsistencies, as well as false positives and negatives in hyperbolic LCS detection\parencite{haller11:_lagran_coher_struc,norgard12:_secon_lagran_coher_struc}.
-In addition, the role of the FTLE field in the accurate detection of elliptic (vortex-type) and parabolic (jet-core type) LCSs has remained an open question (but see \parencite{beron-vera10:_invar_lagran}).
+In addition, the role of the FTLE field in the accurate detection of elliptic (vortex-type) and parabolic (jet-core type) LCSs has remained an open question (but see \textcite{beron-vera10:_invar_lagran}.)
 
@@
 \begin{equation}
-C_{t_0}^t(x_0) = \left[\nabla F_{t_0}^t(x_0)\right]^T \nabla F_{t_0}^t(x_0),
+C_{t_0}^t(x_0) \equiv \left[\nabla F_{t_0}^t(x_0)\right]^T \nabla F_{t_0}^t(x_0),
 \label{eq:CG}
@@
 For this reason, we refer to trajectories of \eqref{eq:etafields} as $\lambda$-lines.
-Following \parencite{haller13:_coher_lagran}, we call the outermost member of a closed family of $\lambda$-lines a coherent Lagrangian vortex boundary.
+Following \textcite{haller13:_coher_lagran}, we call the outermost member of a closed family of $\lambda$-lines a coherent Lagrangian vortex boundary.
 
@@
 To compute the normal repulsion of shrinklines, we note that an infinitesimal
-normal perturbation to a shrinkline $\gamma$ at its point $r$ grows under the flow map $F_{t_{0}}^{t}$ by a factor of $\lambda_{2}(r)$ in the direction normal to $F_{t_{0}}^{t}(\gamma).$ Similarly, small normal perturbations to a stretchline decay by a factor $\lambda_{1}(r)$ in the direction normal to the evolving stretchline.
+normal perturbation to a shrinkline $\gamma$ at its point $r$ grows under the flow map $F_{t_{0}}^{t}$ by a factor of $\lambda_{2}(r)$ in the direction normal to $F_{t_{0}}^{t}(\gamma)$.
+Similarly, small normal perturbations to a stretchline decay by a factor $\lambda_{1}(r)$ in the direction normal to the evolving stretchline.
 
@@
 of $\nabla F_{t_0}^t$ at a point $x_0$ of the Cartesian grid.
-Here $\delta x_i$ is a vector of length $\vert\delta x_i\vert > 0$ that points from the Cartesian grid-point $x_0$ in the $i^{th}$ coordinate direction (Figure~\ref{f:main and auxiliary grids}).
-Computational improvements arising from the use of the auxiliary grid over simply using the nearest points of the main grid were reported in Ref.\,\parencite{farazmand12:_comput_lagran}.
-Experience suggests an auxiliary grid spacing of 1-10\% of the main grid spacing.
+Here $\delta x_i$ is a vector of length $\vert\delta x_i\vert > 0$ that points from the Cartesian grid-point $x_0$ in the $i^\text{th}$ coordinate direction (Figure~\ref{f:main and auxiliary grids}.) Computational improvements arising from the use of the auxiliary grid over simply using the nearest points of the main grid were reported in \textcite{farazmand12:_comput_lagran}.
+Experience suggests setting the auxiliary grid spacing to 1-10\% of the main grid spacing.
 
@@
 
-To avoid excessive computational times in MATLAB, we vectorize \eqref{eq:v(x,t)}, i.e., combine its right-hand-side evaluated over each initial point into a single system of equations.
+To avoid excessive computational times in MATLAB, we vectorize \eqref{eq:v(x,t)}, i.e., combine its right hand side evaluated over each initial point into a single system of equations.
 The resulting system is composed of independent blocks of two-dimensional first-order ODEs.
 We then use MATLAB's \lstinline!ode45!
-to perform trajectory integration from all grid points simultaneously.
-This process typically takes five to ten minutes with default error tolerances.
+function to perform trajectory integration from all grid points simultaneously.
+This calculation typically takes five to ten minutes with default error tolerances.
 A potential drawback of vector form integration is that memory requirements may become excessive at high resolutions.
@@
 
-The Cauchy-Green strain tensor generically admits point singularities, i.e., points where $C_{t_0}^t(x_0)$ has repeated eigenvalues.
-At these points the eigenvectors $\xi_1(x_0)$ and $\xi_2(x_0)$ are no longer well-defined.
-This generically arises at a finite set of isolated points within the computational domain\parencite{delmarcelle94}, and hence lie off the computational grid with probability one.
+The Cauchy-Green strain tensor generally admits point singularities, i.e., points where $C_{t_0}^t(x_0)$ has repeated eigenvalues.
+At these points the eigenvectors $\xi_1(x_0)$ and $\xi_2(x_0)$ are no longer well defined.
+This generally occurs at a finite set of isolated points within the computational domain\parencite{delmarcelle94}, and hence the point singularities lie off the computational grid with probability one.
 
@@
 Incompressible flows (i.e.
-those satisfying $\nabla \cdot v=0$) admit the relationship $\lambda_1(x_0) \lambda_2(x_0) = 1$ at all points of the computational domain \parencite{arnold78:_mathem}.
+those satisfying $\nabla \cdot v=0$) satisfy the relation $\lambda_1(x_0) \lambda_2(x_0) = 1$ at all points of the computational domain \parencite{arnold78:_mathem}.
 Incompressibility can be computationally imposed by first calculating $\lambda_2(x_0)$, then setting $\lambda_1(x_0) = 1/\lambda_2(x_0)$ and calculating the strain eigenvectors $\xi_2$ from $\lambda_2$, then $\xi_1$ from the relationship in \eqref{eq:CG_invariants}.
 Experience shows that computing $\lambda_i$ in this order gives higher accuracy than in the reverse order \parencite{farazmand12:_comput_lagran}.
+This procedure should be applied carefully since by changing $\lambda_1$ but not $\xi_1$, the error in satisfying the eigenvalue equation for the Cauchy-Green strain tensor may increase.
 
@@
 
-As discussed in section \eqref{sub:Elliptic-LCSs}, positions of elliptic LCSs at time $t_0$ are found as closed orbits of the $\eta_\pm^\lambda$ vector fields defined in \eqref{eq:eta}.
-We find such orbits by integrating \eqref{eq:etafields} from points of an appropriately chosen section (Poincare section), and evaluating the first return map (Poincaré map) onto this section.
+As discussed in section \ref{sub:Elliptic-LCSs}, positions of elliptic LCSs at time $t_0$ are found as closed orbits of the $\eta_\pm^\lambda$ vector fields defined in \eqref{eq:eta}.
+We find such orbits by integrating \eqref{eq:etafields} from points of an appropriately chosen section (Poincare section), and evaluating the first return map (Poincare map) onto this section.
 A $\lambda$-line returning to its starting point is then an elliptic LCS.
@@
 The Poincare section is to be oriented such that the first endpoint is close to the center of the expected Lagrangian vortex, and the second endpoint is outside this vortex.
-There is no automated procedure procedure implemented yet in LCS Tool for the positioning of Poincare sections (see, however, \parencite{karrasch14:_autom_lagran} for a recently developed algorithm).
-Additionally, the number of lambda-lines launched from the Poincaré section, \lstinline!poincareSection.numPoints!, must be defined.
+There is no automated procedure procedure implemented in LCS Tool for the positioning of Poincare sections (see however \textcite{karrasch14:_autom_lagran} for a recently developed algorithm.) Additionally, the number of lambda-lines launched from the Poincare section, \lstinline!poincareSection.numPoints!, must be defined.
 A reasonable default value is 100.
@@
 Integration of $\lambda$-lines is performed using the $\eta_\pm^\lambda$ vector fields defined in \eqref{eq:eta} over the main grid.
-The underlying eigenvector fields $\xi_i(x_0)$ have generic but removable orientation discontinuities, which require monitoring and reorientation, as needed.
+The underlying eigenvector fields, $\xi_i(x_0)$, have generic but removable orientation discontinuities, which require monitoring and local reorientation.
 This process is sketched in Table~\ref{t:variable step integration} and illustrated in Figure~\ref{f:variable step integration}.
@@
 Rotations exceeding 90°, between two such neighboring vectors are classified as orientation discontinuities and are corrected before linear interpolation.
-When setting the Cauchy-Green strain tensor main grid resolution, one may find it helpful to calculate a histogram of eigenvector-field rotations to ensure that all rotations are well below 90°.
+When setting the Cauchy-Green strain tensor main grid resolution, one may find it helpful to calculate a histogram of eigenvector-field rotations to ensure that all rotations are well below 90° or almost 180°.
 
@@
 
-In Figure~\ref{f:Poincare return map}, circle markers indicate fixed points of the Poincare map, i.e., points where the distance between the final and the initial point of the orbit $P(s)-s$, is zero.
+In Figure~\ref{f:Poincare return map}, circle markers indicate fixed points of the Poincare map, i.e., points where the distance between the final and the initial point of the orbit $P(s) - s$, is zero.
 The function \lstinline!poincare_closed_orbit_multi!
 performs the computations.
-As seen in Figure~\ref{f:Poincare return map}, LCS Tool uses a filtering parameter \lstinline!dThresh!
-to discard sign changes of $P(s)-s$ that are likely due to numerical sensitivity or a jump discontinuity of the Poincare map.
+As seen in Figure~\ref{f:Poincare return map}, not all zero crossings have circles.
+This is because LCS Tool uses a filtering parameter, \lstinline!dThresh!, to discard sign changes of $P(s) - s$ that are likely due to numerical sensitivity or a jump discontinuity of the Poincare map.
 Specifically, the location of each detected zero crossing is first refined by the bisection method.
-If after a predetermined number of iterations, \lstinline!nBisection!
-(default value 5) the two points around the zero crossing still have absolute values above \lstinline!dThresh!, the zero crossing is discarded.
+If after a predetermined number of iterations, \lstinline!nBisection!, the two points around the zero crossing still have absolute values above \lstinline!dThresh!, the zero crossing is discarded.
 Once all valid closed $\lambda$-lines have been located, the outermost closed $\lambda$-orbit associated with every Poincare section is identified as a coherent Lagrangian vortex boundary.
@@
 \end{enumerate}
-\caption{Algorithm to calculate initial positions of repelling LCSs at time $t_0$. The algorithm for attracting LCSs is similar.}
+\caption{Algorithm to calculate positions of repelling LCSs at time $t_0$. The algorithm for attracting LCSs is similar.}
 \label{t:Hyperbolic LCS algorithm}
@@
 These removed segments can be turned back on demand, as shown in Figure~\ref{fig:ocean_dataset_colortracer}.
+
+The local maximization distance is the number of grid points over which the Cauchy-Green strain tensor is susceptible to numerical noise from integration.
+This distance is thus expected to be a function of variable time step integration tolerances.
+
+Prior to initiating shrinkline and stretchline integration, a maximum length is defined.
+Shrinklines and stretchlines are then integrated until reaching the domain boundary or the maximum length.
+The maximum length may be seen as safety factor to prevent highly sinuous shrinklines and stretchlines leading to inordinately long computations; the maximum length should be set such that most shrinklines and stretchlines stop at the domain boundary.
+
+Once shrinkline and stretchline integration has begun, it may evolve so as to be within the local minimization distance of an existing shrinkline or stretchline.
+In other words, the local minimization distance affects where shrinklines and stretchlines begin, but not where they finish.
 
@@
 As usual, the local maximization distance is set larger for stretchlines than for shrinklines (cf.
-l.\,4 and l\,9).
-The purpose of the maximization distance is to obtain spatially separated LCSs and to avoid a dense tangle of lines that basically indicates the same hyperbolic LCS (cf.
-Table \ref{t:Hyperbolic LCS algorithm}, point 5).
-Setting the local maximization distance for stretchlines larger than for shrinklines allows obtaining a comparable number of stretchlines and shrinklines overall in the flow domain (recall that hyperbolic LCS seed points are discarded if they are within the local maximization distance of an existing hyperbolic LCS.) Shrinklines are locally tangent to ridges of $\lambda_2$ maxima, whereas stretchlines are locally normal to these ridges.
+l.\,4 and l.\,9.) The purpose of the maximization distance is to obtain spatially separated LCSs and to avoid a dense tangle of lines that basically indicates the same hyperbolic LCS (cf.
+Table \ref{t:Hyperbolic LCS algorithm}, point 5.) Setting the local maximization distance for stretchlines larger than for shrinklines allows obtaining a comparable number of stretchlines and shrinklines overall in the flow domain (recall that hyperbolic LCS seed points are discarded if they are within the local maximization distance of an existing hyperbolic LCS.) Shrinklines are locally tangent to ridges of $\lambda_2$ maxima, whereas stretchlines are locally normal to these ridges.
 Setting the local maximization distance of stretchlines and shrinklines equal would therefore produce a greater number of stretchlines than shrinklines overall.
@@
 In contrast to the previous two analytic examples, the velocity field is available only with discrete temporal and spatial resolution.
-Our region of interest is a small domain in the South Atlantic Ocean, where exceptionally coherent eddies, Agulhas rings, were recently found by \parencite{haller13:_coher_lagran} using the theory we surveyed in Section \ref{sub:Elliptic-LCSs}.
+Our region of interest is a small domain in the South Atlantic Ocean, where exceptionally coherent eddies, Agulhas rings, were recently found by \textcite{haller13:_coher_lagran} using the theory we surveyed in Section \ref{sub:Elliptic-LCSs}.
 
@@
 The flow is integrated and the Cauchy-Green strain tensor is computed by the function \lstinline!eig_cgStrain!
-(line 35).
-Incompressibility of the flow is enforced (line 12), the auxiliary grid distance is set to 1\% of the main grid distance (line 17), and eigenvalues are computed from the auxiliary grid (line 16).
-Elliptic LCSs are computed in line 54, after the Poincare sections have been set (line 41-42) and the $\eta_\pm^\lambda$ fields have been defined (line 53).
-Hyperbolic LCSs are computed in lines 57 and 59.
+(line 35.) Incompressibility of the flow is enforced (line 12), the auxiliary grid distance is set to 1\% of the main grid distance (line 17), and eigenvalues are computed from the auxiliary grid (line 16.) Elliptic LCSs are computed in line 54, after the Poincare sections have been set (line 41-42) and the $\eta_\pm^\lambda$ fields have been defined (line 53.) Hyperbolic LCSs are computed in lines 57 and 59.
 

### 2014-08-23 main.tex
Use the cleveref package
Conflicts:
	main.tex

@@
 \usepackage[hidelinks]{hyperref}
+\usepackage{cleveref}
 
@@
 LCSs are most repelling, attracting and shearing material lines that form the centerpieces of observed tracer patterns in two-dimensional unsteady dynamical systems.
-LCS Tool implements the latest geodesic theory for LCSs in two-dimensional flows, uncovering key transport barriers in unsteady flow velocity data as explicit solutions of differential equations.
-After a review of the underlying theory, we explain the numerical algorithms used by LCS Tool, and illustrate its capabilities on three unsteady fluid flow examples: a double gyre, a Bickley jet and an ocean dataset for the Agulhas current.
+LCS Tool implements the latest geodesic theory of LCSs for two-dimensional flows, uncovering key transport barriers in unsteady flow velocity data as explicit solutions of differential equations.
+After a review of the underlying theory, we explain the steps and numerical methods used by LCS Tool, and illustrate its capabilities on three unsteady fluid flow examples.
 }
@@
 Equating FTLE ridges with LCSs, however, would create theoretical inconsistencies, as well as false positives and negatives in hyperbolic LCS detection\parencite{haller11:_lagran_coher_struc,norgard12:_secon_lagran_coher_struc}.
-In addition, the role of the FTLE field in the accurate detection of elliptic (vortex-type) and parabolic (jet-core type) LCSs has remained an open question (but see \textcite{beron-vera10:_invar_lagran}.)
+In addition, the role of the FTLE field in the accurate detection of elliptic (vortex-type) and parabolic (jet-core type) LCSs has remained an open question (but see \parencite{beron-vera10:_invar_lagran}).
 
@@
 \od{x}{t} = v(x,t), && x \in U \subset \mathbf R^2, && t \in [t_-,t_+].
-\label{eq:v(x,t)}
+\label{eq:velocity}
 \end{align}
-Trajectories of \eqref{eq:v(x,t)} are denoted $x(t;t_0,x_0)$, with $x_0 \in U$ denoting their initial position in the open set $U$ at an initial time $t_0 \in [t_-,t_+]$.
+Trajectories of \cref{eq:velocity} are denoted $x(t;t_0,x_0)$, with $x_0 \in U$ denoting their initial position in the open set $U$ at an initial time $t_0 \in [t_-,t_+]$.
 The flow map is then defined as
@@
 mapping initial positions to current positions at time $t$.
-The time interval $[t_-,t_+]$ is part of the definition of the finite-time dynamical system \eqref{eq:v(x,t)}.
+The time interval $[t_-,t_+]$ is part of the definition of the finite-time dynamical system in \cref{eq:velocity}.
 This interval may be a time scale of interest or the maximum interval over which velocity data is available from simulations or observations.
@@
 \begin{equation}
-C_{t_0}^t(x_0) \equiv \left[\nabla F_{t_0}^t(x_0)\right]^T \nabla F_{t_0}^t(x_0),
+C_{t_0}^t(x_0) = \left[\nabla F_{t_0}^t(x_0)\right]^T \nabla F_{t_0}^t(x_0),
 \label{eq:CG}
@@
 \subsection{Elliptic LCSs}
-\label{sub:Elliptic-LCSs}
+\label{sec:Elliptic LCSs}
 
@@
 \end{equation}
-which coincide with null-geodesics of the Lorentzian metric family $e_\lambda(u,v) = \left\langle u,\left[D_{t_0}^t(r) - \lambda^2 I\right] v \right\rangle$.
-For this reason, we refer to the computation of elliptic LCSs as limit cycles of \eqref{eq:etafields} as \emph{geodesic detection of elliptic LCSs}.
+which coincide with null-geodesics of the Lorentzian metric family 
+\[
+e_\lambda(u,v) = \left\langle u,\left[D_{t_0}^t(r) - \lambda^2 I\right] v \right\rangle.
+\]
+For this reason, we refer to the computation of elliptic LCSs as limit cycles of \cref{eq:etafields} as \emph{geodesic detection of elliptic LCSs}.
 
-Any orbit of \eqref{eq:etafields} turns out to stretch uniformly under the flow map $F_{t_0}^t$.
-Specifically, any subset of an orbit of \eqref{eq:etafields} increases its arclength precisely by a factor of $\lambda$.
-For this reason, we refer to trajectories of \eqref{eq:etafields} as $\lambda$-lines.
+Any orbit of \cref{eq:etafields} turns out to stretch uniformly under the flow map $F_{t_0}^t$.
+Specifically, any subset of an orbit of \cref{eq:etafields} increases its arclength precisely by a factor of $\lambda$.
+For this reason, we refer to trajectories of \cref{eq:etafields} as $\lambda$-lines.
 Following \textcite{haller13:_coher_lagran}, we call the outermost member of a closed family of $\lambda$-lines a coherent Lagrangian vortex boundary.
@@
 \subsection{Hyperbolic LCSs}
-\label{sub:Hyperbolic-LCSs}
+\label{sec:Hyperbolic LCSs}
 
@@
 Shrinklines and stretchlines coincide with the null-geodesics of the Lorentzian metric $h(u,v) = \left\langle u,D_{t_0}^t(r) v \right\rangle$.
-For this reason, we refer to the computation of hyperbolic LCSs as strongest normally-repelling or normally-attracting orbits of \eqref{eq:etafields} as \emph{geodesic detection of hyperbolic LCSs}.
+For this reason, we refer to the computation of hyperbolic LCSs as strongest normally-repelling or normally-attracting orbits of \cref{eq:etafields} as \emph{geodesic detection of hyperbolic LCSs}.
 
 To compute the normal repulsion of shrinklines, we note that an infinitesimal
-normal perturbation to a shrinkline $\gamma$ at its point $r$ grows under the flow map $F_{t_{0}}^{t}$ by a factor of $\lambda_{2}(r)$ in the direction normal to $F_{t_{0}}^{t}(\gamma)$.
-Similarly, small normal perturbations to a stretchline decay by a factor $\lambda_{1}(r)$ in the direction normal to the evolving stretchline.
+normal perturbation to a shrinkline $\gamma$ at its point $r$ grows under the flow map $F_{t_{0}}^{t}$ by a factor of $\lambda_{2}(r)$ in the direction normal to $F_{t_{0}}^{t}(\gamma).$ Similarly, small normal perturbations to a stretchline decay by a factor $\lambda_{1}(r)$ in the direction normal to the evolving stretchline.
 
@@
 
-Here we describe a step-by-step numerical implementation of the geodesic detection of elliptic and hyperbolic LCSs through the differential equations \eqref{eq:etafields}, \eqref{eq:strainline}, and \eqref{eq:stretchline}.
+Here we describe a step-by-step numerical implementation of the geodesic detection of elliptic and hyperbolic LCSs through the differential \cref{eq:etafields,eq:strainline,eq:stretchline}.
 
@@
 
-The common first step in calculating elliptic and hyperbolic LCSs is the computation of the Cauchy-Green strain tensor field $C_{t_0}^t(x_0)$, as defined in \eqref{eq:CG}.
+The common first step in calculating elliptic and hyperbolic LCSs is the computation of the Cauchy-Green strain tensor field $C_{t_0}^t(x_0)$, as defined in \cref{eq:CG}.
 The function performing this calculation in LCS Tool is \lstinline!eig_cgStrain!.
-The main steps executed by this function are enumerated in Table~\ref{t:Cauchy-Green algorithm}, while Table~\ref{t:eig_cgStrain syntax} summarizes the syntax of \lstinline!eig_cgStrain!.
+The main steps executed by this function are enumerated in \cref{t:Cauchy-Green algorithm}, while \cref{t:eig_cgStrain syntax} summarizes the syntax of \lstinline!eig_cgStrain!.
 
@@
 an auxiliary grid for differentiating with respect to initial conditions.
-\item Solve \eqref{eq:v(x,t)} starting from each grid point and auxiliary grid point over the time interval $[t_0,t]$. This gives a discrete approximation to the flow map $F_{t_0}^t(x_0)$.
+\item Solve \cref{eq:velocity} starting from each grid point and auxiliary grid point over the time interval $[t_0,t]$. This gives a discrete approximation to the flow map $F_{t_0}^t(x_0)$.
 \item Use finite differencing over the auxiliary grid to compute numerically the derivative of the flow map $DF_{t_0}^t(x_0)$.
-\item Compute the Cauchy-Green strain tensor field $C_{t_0}^t(x_0) = \left(DF_{t_0}^t(x_0)\right)^T DF_{t_0}^t(x_{0})$, its eigenvalue field $\lambda_{1,2}(x_0)$, and eigenvector fields $\xi_{1,2}(x_0)$ over the initial condition grid.
+\item Compute the Cauchy-Green strain tensor field $C_{t_0}^t(x_0) = \left(DF_{t_0}^t(x_0)\right)^T DF_{t_0}^t(x_0)$, its eigenvalue field $\lambda_{1,2}(x_0)$, and eigenvector fields $\xi_{1,2}(x_0)$ over the initial condition grid.
 \end{enumerate}
@@
 
-The Cartesian grid of initial conditions mentioned in Table~\ref{t:Cauchy-Green algorithm} is rectangular, with user-defined vertical and horizontal ranges and resolutions.
+The Cartesian grid of initial conditions mentioned in \cref{t:Cauchy-Green algorithm} is rectangular, with user-defined vertical and horizontal ranges and resolutions.
 The optimal resolution may be determined by a successive doubling of the initial resolution until convergence of the extracted LCSs is observed visually.
@@
 The auxiliary grid (cf.
-Table~\ref{t:Cauchy-Green algorithm}) comprises four points placed symmetrically around each point of the Cartesian grid (Figure~\ref{f:main and auxiliary grids}.) These points are used to achieve increased accuracy in the finite-difference approximation
+\cref{t:Cauchy-Green algorithm}) comprises four points placed symmetrically around each point of the Cartesian grid (\cref{f:main and auxiliary grids}.) These points are used to achieve increased accuracy in the finite-difference approximation
 \[
@@
 of $\nabla F_{t_0}^t$ at a point $x_0$ of the Cartesian grid.
-Here $\delta x_i$ is a vector of length $\vert\delta x_i\vert > 0$ that points from the Cartesian grid-point $x_0$ in the $i^\text{th}$ coordinate direction (Figure~\ref{f:main and auxiliary grids}.) Computational improvements arising from the use of the auxiliary grid over simply using the nearest points of the main grid were reported in \textcite{farazmand12:_comput_lagran}.
+Here $\delta x_i$ is a vector of length $\vert\delta x_i\vert > 0$ that points from the Cartesian grid-point $x_0$ in the $i^\text{th}$ coordinate direction (\cref{f:main and auxiliary grids}.) Computational improvements arising from the use of the auxiliary grid over simply using the nearest points of the main grid were reported in \textcite{farazmand12:_comput_lagran}.
 Experience suggests setting the auxiliary grid spacing to 1-10\% of the main grid spacing.
@@
 \end{center}
-\caption{Illustration of the main grid (filled circles) and the auxiliary grid (empty circles) used in the computation of the derivative of the flow map in definition \eqref{eq:CG} of the Cauchy-Green strain tensor.
+\caption{Illustration of the main grid (filled circles) and the auxiliary grid (empty circles) used in the computation of the derivative of the flow map in \cref{eq:CG} of the Cauchy-Green strain tensor.
 The variable \lstinline!auxGridRelDelta!
@@
 As mentioned above, a typical main grid for the Cauchy-Green strain tensor comprises $500 \times 500$ points.
-This means that after the addition of 4 auxiliary grid points around each main grid point, eq.
-\eqref{eq:v(x,t)} must be integrated over 1.25 million initial conditions.
+This means that after the addition of 4 auxiliary grid points around each main grid point, \cref{eq:velocity} must be integrated over 1.25 million initial conditions.
 
-To avoid excessive computational times in MATLAB, we vectorize \eqref{eq:v(x,t)}, i.e., combine its right hand side evaluated over each initial point into a single system of equations.
+To avoid excessive computational times in MATLAB, we vectorize \cref{eq:velocity}, i.e., combine its right hand side evaluated over each initial point into a single system of equations.
 The resulting system is composed of independent blocks of two-dimensional first-order ODEs.
@@
 
-The Cauchy-Green strain tensor generally admits point singularities, i.e., points where $C_{t_0}^t(x_0)$ has repeated eigenvalues.
-At these points the eigenvectors $\xi_1(x_0)$ and $\xi_2(x_0)$ are no longer well defined.
-This generally occurs at a finite set of isolated points within the computational domain\parencite{delmarcelle94}, and hence the point singularities lie off the computational grid with probability one.
+The Cauchy-Green strain tensor generically admits point singularities, i.e., points where $C_{t_0}^t(x_0)$ has repeated eigenvalues.
+At these points the eigenvectors $\xi_1(x_0)$ and $\xi_2(x_0)$ are no longer well-defined.
+This generically arises at a finite set of isolated points within the computational domain\parencite{delmarcelle94}, and hence lie off the computational grid with probability one.
 
@@
 those satisfying $\nabla \cdot v=0$) satisfy the relation $\lambda_1(x_0) \lambda_2(x_0) = 1$ at all points of the computational domain \parencite{arnold78:_mathem}.
-Incompressibility can be computationally imposed by first calculating $\lambda_2(x_0)$, then setting $\lambda_1(x_0) = 1/\lambda_2(x_0)$ and calculating the strain eigenvectors $\xi_2$ from $\lambda_2$, then $\xi_1$ from the relationship in \eqref{eq:CG_invariants}.
+Incompressibility can be computationally imposed by first calculating $\lambda_2(x_0)$, then setting $\lambda_1(x_0) = 1/\lambda_2(x_0)$ and calculating the strain eigenvectors $\xi_2$ from $\lambda_2$, then $\xi_1$ from the relationship in \cref{eq:CG_invariants}.
 Experience shows that computing $\lambda_i$ in this order gives higher accuracy than in the reverse order \parencite{farazmand12:_comput_lagran}.
-This procedure should be applied carefully since by changing $\lambda_1$ but not $\xi_1$, the error in satisfying the eigenvalue equation for the Cauchy-Green strain tensor may increase.
 
@@
 Velocity fields defined by datasets require pre-processing before
-they are used in the numerical integration of \eqref{eq:v(x,t)}.
+they are used in the numerical integration of \cref{eq:velocity}.
 This requires spatial and temporal interpolation that enables the evaluation of the velocity function at arbitrary points in $U$ and at arbitrary times between $t_-$ and $t_+$.
-In section \ref{sec:oceandataset}, we present an ocean dataset example with details of possible interpolation functions.
+In \cref{sec:oceandataset}, we present an ocean dataset example with details of possible interpolation functions.
 
@@
 
-As discussed in section \ref{sub:Elliptic-LCSs}, positions of elliptic LCSs at time $t_0$ are found as closed orbits of the $\eta_\pm^\lambda$ vector fields defined in \eqref{eq:eta}.
-We find such orbits by integrating \eqref{eq:etafields} from points of an appropriately chosen section (Poincare section), and evaluating the first return map (Poincare map) onto this section.
+As discussed in \cref{sec:Elliptic LCSs}, positions of elliptic LCSs at time $t_0$ are found as closed orbits of the $\eta_\pm^\lambda$ vector fields defined in \cref{eq:eta}.
+We find such orbits by integrating \cref{eq:etafields} from points of an appropriately chosen section (Poincare section), and evaluating the first return map (Poincare map) onto this section.
 A $\lambda$-line returning to its starting point is then an elliptic LCS.
@@
 
-The main steps in calculating elliptic LCSs are enumerated in Table~\ref{t:Elliptic LCS algorithm} and described in further detail below.
-The syntax of elliptic LCS functions in LCS Tool is shown in Table~\ref{t:Elliptic LCS functions}.
+The main steps in calculating elliptic LCSs are enumerated in \cref{t:Elliptic LCS algorithm} and described in further detail below.
+The syntax of elliptic LCS functions in LCS Tool is shown in \cref{t:Elliptic LCS functions}.
 
@@
 \item Position Poincare sections in flow domain to specify initial positions of lambda-lines
-\item Integrate $\lambda$-lines tangent to $\eta_\pm^\lambda$ (see ~\eqref{eq:etafields}.)
+\item Integrate $\lambda$-lines tangent to $\eta_\pm^\lambda$ (see \cref{eq:etafields}.)
 \item Calculate Poincare map
@@
 \hline
-\lstinline!lambda! & scalar lambda value in Equation~\eqref{eq:eta}\tabularnewline
+\lstinline!lambda! & scalar lambda value in \cref{eq:eta}\tabularnewline
 \hline \hline
@@
 The second step is to integrate the $\lambda$-lines  starting from the Poincare section to obtain the corresponding Poincare map.
-Integration of $\lambda$-lines is performed using the $\eta_\pm^\lambda$ vector fields defined in \eqref{eq:eta} over the main grid.
+Integration of $\lambda$-lines is performed using the $\eta_\pm^\lambda$ vector fields defined in \cref{eq:eta} over the main grid.
 The underlying eigenvector fields, $\xi_i(x_0)$, have generic but removable orientation discontinuities, which require monitoring and local reorientation.
-This process is sketched in Table~\ref{t:variable step integration} and illustrated in Figure~\ref{f:variable step integration}.
+This process is sketched in \cref{t:variable step integration} and illustrated in \cref{f:variable step integration}.
 Linear interpolation is used in the interpolation of $\eta_\pm^\lambda$ within a grid element, since using higher-order interpolation would necessitate verifying that there are no orientation discontinuities beyond the four nearest grid points.
@@
 
-An example of a Poincare map produced from integration of the $\eta_\pm^\lambda$ field is shown in Figure~\ref{f:Poincare return map}.
+An example of a Poincare map produced from integration of the $\eta_\pm^\lambda$ field is shown in \cref{f:Poincare return map}.
 Most orbits will return to the Poincare section and their integration will then be stopped using the ODE event detection function of MATLAB.
@@
 
-In Figure~\ref{f:Poincare return map}, circle markers indicate fixed points of the Poincare map, i.e., points where the distance between the final and the initial point of the orbit $P(s) - s$, is zero.
+In \cref{f:Poincare return map}, circle markers indicate fixed points of the Poincare map, i.e., points where the distance between the final and the initial point of the orbit $P(s) - s$, is zero.
 The function \lstinline!poincare_closed_orbit_multi!
 performs the computations.
-As seen in Figure~\ref{f:Poincare return map}, not all zero crossings have circles.
+As seen in \cref{f:Poincare return map}, not all zero crossings have circles.
 This is because LCS Tool uses a filtering parameter, \lstinline!dThresh!, to discard sign changes of $P(s) - s$ that are likely due to numerical sensitivity or a jump discontinuity of the Poincare map.
@@
 
-As discussed in section \ref{sub:Hyperbolic-LCSs}, positions of hyperbolic LCSs at time $t_{0}$ are found as the strongest repelling orbits of the vector field \eqref{eq:strainline} (repelling LCSs), and strongest attracting orbits of the vector field \eqref{eq:stretchline} (attracting LCSs.) By repulsion and attraction we mean a property of the LCS (as an evolving material line) under the flow map $F_{t_0}^t$.
+As discussed in \cref{sec:Hyperbolic LCSs}, positions of hyperbolic LCSs at time $t_0$ are found as the strongest repelling orbits of the vector field \cref{eq:strainline} (repelling LCSs), and strongest attracting orbits of the vector field \cref{eq:stretchline} (attracting LCSs.) By repulsion and attraction we mean a property of the LCS (as an evolving material line) under the flow map $F_{t_0}^t$.
 We identify the strongest repelling shrinklines as those crossing a local maximum of the $\lambda_2(x_0)$ field.
@@
 
-The main steps of our hyperbolic LCS detection algorithm are enumerated in Table~\ref{t:Hyperbolic LCS algorithm}.
+The main steps of our hyperbolic LCS detection algorithm are enumerated in \cref{t:Hyperbolic LCS algorithm}.
 The basic function to compute hyperbolic LCSs in LCS Tool is \lstinline!seed_curves_from_lambda_max!
-and its syntax is given in Table~\ref{t:seed_curves_from_lambda_max syntax}.
+and its syntax is given in \cref{t:seed_curves_from_lambda_max syntax}.
 
@@
 \item Define a maximum shrinkline length
-\item Integrate a shrinkline forward and backward according to Equation~\eqref{eq:strainline} and using the largest $\lambda_2$ local maximum as the initial position.
+\item Integrate a shrinkline forward and backward according to \cref{eq:strainline} and using the largest $\lambda_2$ local maximum as the initial position.
 Integrate until the shrinkline has attained the maximum shrinkline length, or until it has reached the domain boundary.
@@
 \end{enumerate}
-\caption{Algorithm to calculate positions of repelling LCSs at time $t_0$. The algorithm for attracting LCSs is similar.}
+\caption{Algorithm to calculate initial positions of repelling LCSs at time $t_0$. The algorithm for attracting LCSs is similar.}
 \label{t:Hyperbolic LCS algorithm}
@@
 Motivated by convention and for clarity in visualization, however, the default setting of LCS Tool removes hyperbolic LCS segments from the interior of elliptic LCSs.
-These removed segments can be turned back on demand, as shown in Figure~\ref{fig:ocean_dataset_colortracer}.
-
-The local maximization distance is the number of grid points over which the Cauchy-Green strain tensor is susceptible to numerical noise from integration.
-This distance is thus expected to be a function of variable time step integration tolerances.
-
-Prior to initiating shrinkline and stretchline integration, a maximum length is defined.
-Shrinklines and stretchlines are then integrated until reaching the domain boundary or the maximum length.
-The maximum length may be seen as safety factor to prevent highly sinuous shrinklines and stretchlines leading to inordinately long computations; the maximum length should be set such that most shrinklines and stretchlines stop at the domain boundary.
-
-Once shrinkline and stretchline integration has begun, it may evolve so as to be within the local minimization distance of an existing shrinkline or stretchline.
-In other words, the local minimization distance affects where shrinklines and stretchlines begin, but not where they finish.
+These removed segments can be turned back on demand, as shown in \cref{fig:ocean_dataset_colortracer}.
 
@@
 \end{equation}
-The MATLAB function describing this velocity field is given in Listing~\ref{l:double gyre derivative}, specifying the right hand side of the particle ODE in a way that supports vectorized integration.
+The MATLAB function describing this velocity field is given in \cref{l:double gyre derivative}, specifying the right hand side of the particle ODE in a way that supports vectorized integration.
 
-\begin{lstlisting}[caption={Double gyre derivative function corresponding to Equations~\eqref{eq:double gyre derivative equations}.},label=l:double gyre derivative]
+\begin{lstlisting}[caption={Double gyre derivative function corresponding to \cref{eq:double gyre derivative equations}.},label=l:double gyre derivative]
 function derivative_ = derivative(t,x,epsilon,amplitude,omega)
@@
 is optionally provided by LCS Tool), we position Poincare sections to capture elliptic LCSs.
-An LCS Tool demo script to perform this operation is given in Listing~\ref{l:double gyre lambda-line LCS} where two Poincare sections are defined.
-The free stretching parameter $\lambda$ (Equation \eqref{eq:etafields}) is varied over the range $[0.93,1.07]$ with increments of $0.01$.
+An LCS Tool demo script to perform this operation is given in \cref{l:double gyre lambda-line LCS} where two Poincare sections are defined.
+The free stretching parameter $\lambda$ (\cref{eq:etafields}) is varied over the range $[0.93,1.07]$ with increments of $0.01$.
 In a loop , closed orbits for all predefined $\lambda$ values are computed and the outermost closed orbit is kept as the Lagrangian vortex boundary.
@@
 
-In Figure~\ref{fig:double_gyre_lambda_lcs_convergence}, the resolution of the  Cauchy-Green strain tensor is varied from $500 \times 250$ to $1000 \times 500$.
+In \cref{fig:double_gyre_lambda_lcs_convergence}, the resolution of the  Cauchy-Green strain tensor is varied from $500 \times 250$ to $1000 \times 500$.
 The location of the outermost closed lambda-line changes little, demonstrating convergence.
@@
 
-Figure \ref{fig:double_gyre_lambda_strain_lcs} shows the complete picture of elliptic and hyperbolic LCSs of the double gyre at this resolution.
-Listing~\ref{lst:double_gyre_lambda_strain_stretch_lcs} lists the corresponding LCS Tool commands.
+\Cref{fig:double_gyre_lambda_strain_lcs} shows the complete picture of elliptic and hyperbolic LCSs of the double gyre at this resolution.
+\Cref{lst:double_gyre_lambda_strain_stretch_lcs} lists the corresponding LCS Tool commands.
 The maximal length of shrinklines and stretchlines, \lstinline!strainlineMaxLength!
@@
 l.\,4 and l.\,9.) The purpose of the maximization distance is to obtain spatially separated LCSs and to avoid a dense tangle of lines that basically indicates the same hyperbolic LCS (cf.
-Table \ref{t:Hyperbolic LCS algorithm}, point 5.) Setting the local maximization distance for stretchlines larger than for shrinklines allows obtaining a comparable number of stretchlines and shrinklines overall in the flow domain (recall that hyperbolic LCS seed points are discarded if they are within the local maximization distance of an existing hyperbolic LCS.) Shrinklines are locally tangent to ridges of $\lambda_2$ maxima, whereas stretchlines are locally normal to these ridges.
+\cref{t:Hyperbolic LCS algorithm}, point 5.) Setting the local maximization distance for stretchlines larger than for shrinklines allows obtaining a comparable number of stretchlines and shrinklines overall in the flow domain (recall that hyperbolic LCS seed points are discarded if they are within the local maximization distance of an existing hyperbolic LCS.) Shrinklines are locally tangent to ridges of $\lambda_2$ maxima, whereas stretchlines are locally normal to these ridges.
 Setting the local maximization distance of stretchlines and shrinklines equal would therefore produce a greater number of stretchlines than shrinklines overall.
@@
 time in the flow.
-Figure~\ref{fig:Bickley_jet_lambda_LCS_full} shows lambda-line LCS of the Bickley jet with the FTLE in the background.
+\Cref{fig:Bickley_jet_lambda_LCS_full} shows lambda-line LCS of the Bickley jet with the FTLE in the background.
 The figure demonstrates a poor result since the LCSs change shape significantly between eddies.
@@
 
-Figure ~\ref{fig:Bickley jet lambda lcs convergence} shows convergence of the lambda-line elliptic LCSs.
+\Cref{fig:Bickley jet lambda lcs convergence} shows convergence of the lambda-line elliptic LCSs.
 The LCS Tool script used to produce that figure are based on the script \lstinline!demo/bickley_jet/hyperbolic_shear_lcs.m!
-which is given in  Listing~\ref{lst:Bickley jet hyperbolic_shear_lcs subset}.
+which is given in \cref{lst:Bickley jet hyperbolic_shear_lcs subset}.
 It is worth pointing out that in order to obtain closed orbits, $\lambda = 0.995$; with $\lambda = 1$ no closed orbits exist.
@@
 
-Shrinkline LCSs are shown in Figure~\ref{fig:Bickley_jet_strain_lcs} and stretchline LCSs are shown in Figure~\ref{fig:Bickley_jet_stretch_lcs}.
+Shrinkline LCSs are shown in \cref{fig:Bickley_jet_strain_lcs} and stretchline LCSs are shown in \cref{fig:Bickley_jet_stretch_lcs}.
 
@@
 In contrast to the previous two analytic examples, the velocity field is available only with discrete temporal and spatial resolution.
-Our region of interest is a small domain in the South Atlantic Ocean, where exceptionally coherent eddies, Agulhas rings, were recently found by \textcite{haller13:_coher_lagran} using the theory we surveyed in Section \ref{sub:Elliptic-LCSs}.
+Our region of interest is a small domain in the South Atlantic Ocean, where exceptionally coherent eddies, Agulhas rings, were recently found by \textcite{haller13:_coher_lagran} using the theory we surveyed in \cref{sec:Elliptic LCSs}.
 
@@
 The data is given at a spatial resolution of $1/4°$ and a temporal resolution of $7\,\mathrm{days}$.
-Due to the discrete data, defining the right hand side of Equation~\eqref{eq:dtlon} and~\eqref{eq:dtlat} involves spline interpolation in space and time.
+Due to the discrete data, defining the right hand side of \cref{eq:dtlon} and \cref{eq:dtlat} involves spline interpolation in space and time.
 An interpolant is generated first, then the function \lstinline!flowdata_derivative!
 evaluates the interpolants for the zonal and meridional velocity at the needed coordinates.
-Listing \ref{l:griddedInterp} shows the relevant part of the code in LCS Tool's ocean demo file \lstinline!hyperbolic_shear_lcs.m!.
+\Cref{l:griddedInterp} shows the relevant part of the code in LCS Tool's ocean demo file \lstinline!hyperbolic_shear_lcs.m!.
 In lines 8-11, the commands for the interpolation are given of the velocity dataset are given.
@@
 
-We choose the integration time as $T=30$ days (Listing \ref{l:griddedInterp}, line 4), which is larger than the eddy turnover time in this region.
+We choose the integration time as $T=30$ days (\cref{l:griddedInterp}, line 4), which is larger than the eddy turnover time in this region.
 The resolution of the main computational grid for initial conditions is set to $400 \times 400$ (line 3.) This corresponds to a resolution of roughly 0.015° and gives good results.
@@
 
-Figures~\ref{f:ocean dataset hyperbolic shear lcs details strainline} and \ref{f:ocean dataset hyperbolic shear lcs details stretchline} show the positions of elliptic and hyperbolic LCSs on 22 November 2006, which is the same time instant analyzed in~\textcite{haller13:_coher_lagran,beron-vera13:_objec_agulh}.
+\Cref{f:ocean dataset hyperbolic shear lcs details strainline,f:ocean dataset hyperbolic shear lcs details stretchline} show the positions of elliptic and hyperbolic LCSs on 22 November 2006, which is the same time instant analyzed in \textcite{haller13:_coher_lagran,beron-vera13:_objec_agulh}.
 Here the integration time is chosen as $T = 30$ days, as opposed to 90 days in the cited references, to avoid tangling of hyperbolic LCSs.
@@
 % End:
-% LocalWords:  stretchline stretchlines incompressible gyre eigenvector eigenvectors incompressibility dataset datasets LCS LCSs strainline Lyapunov Bickley Rossby timespan variational adaptively vortices parametrise advected strainlines minima extremal FTLE MATLAB maxima geostrophic shrinkline shrinklines
+% LocalWords:  stretchline stretchlines incompressible gyre eigenvector eigenvectors incompressibility dataset datasets LCS LCSs Lyapunov Bickley Rossby timespan variational adaptively vortices parametrise advected strainlines minima extremal FTLE MATLAB maxima geostrophic shrinkline shrinklines

### 2014-08-23 main.tex
Use the cleveref package

@@
 Experience shows that computing $\lambda_i$ in this order gives higher accuracy than in the reverse order \parencite{farazmand12:_comput_lagran}.
+This procedure should be applied carefully since by changing $\lambda_1$ but not $\xi_1$, the error in satisfying the eigenvalue equation for the Cauchy-Green strain tensor may increase.
 
@@
 These removed segments can be turned back on demand, as shown in \cref{fig:ocean_dataset_colortracer}.
+
+The local maximization distance is the number of grid points over which the Cauchy-Green strain tensor is susceptible to numerical noise from integration.
+This distance is thus expected to be a function of variable time step integration tolerances.
+
+Prior to initiating shrinkline and stretchline integration, a maximum length is defined.
+Shrinklines and stretchlines are then integrated until reaching the domain boundary or the maximum length.
+The maximum length may be seen as safety factor to prevent highly sinuous shrinklines and stretchlines leading to inordinately long computations; the maximum length should be set such that most shrinklines and stretchlines stop at the domain boundary.
+
+Once shrinkline and stretchline integration has begun, it may evolve so as to be within the local minimization distance of an existing shrinkline or stretchline.
+In other words, the local minimization distance affects where shrinklines and stretchlines begin, but not where they finish.
 

### 2014-08-23 main.tex
Make style consistent in tables
Change lambda-line LCSs to elliptic LCSs

@@
 \hline
-\lstinline!auxGridRelDelta! & optional scalar between 0 and 0.5 to specify auxiliary grid spacing. Default is $10^{-2}$.\tabularnewline
+\lstinline!auxGridRelDelta! & optional scalar between 0 and 0.5 to specify auxiliary grid spacing. Default: $10^{-2}$.\tabularnewline
 \hline
-\lstinline!eigenvalueFromMainGrid! & optional logical to control whether eigenvalues of Cauchy-Green strain are calculated from main or auxiliary grid. Default is \lstinline!true!.\tabularnewline
+\lstinline!eigenvalueFromMainGrid! & optional logical to control whether eigenvalues of Cauchy-Green strain are calculated from main or auxiliary grid. Default: \lstinline!true!.\tabularnewline
 \hline
-\lstinline!incompressible! & optional logical to specify if incompressibility is imposed. Default is \lstinline!false!.\tabularnewline
+\lstinline!incompressible! & optional logical to specify if incompressibility is imposed. Default: \lstinline!false!.\tabularnewline
 \hline
@@
 \hline
-\lstinline!nBisection! & optional number of bisection steps to refine zero crossings of Poincare map. Default value: 5.\tabularnewline
+\lstinline!nBisection! & optional number of bisection steps to refine zero crossings of Poincare map. Default: 5.\tabularnewline
 \hline
-\lstinline!dThresh! & optional threshold to discard discontinuous zero crossings of Poincare map. Default value: $10^{-2}$.\tabularnewline
+\lstinline!dThresh! & optional threshold to discard discontinuous zero crossings of Poincare map. Default: $10^{-2}$.\tabularnewline
 \hline
@@
 \hline
-\lstinline!periodicBc! & optional $1 \times 2$ logical array to specify periodic boundary conditions. Default is \lstinline![false,false]!.\tabularnewline
+\lstinline!periodicBc! & optional $1 \times 2$ logical array to specify periodic boundary conditions. Default: \lstinline![false,false]!.\tabularnewline
 \hline
@@
 \hline
-\lstinline!periodicBc! & optional $1 \times 2$ logical array to specify periodic boundary conditions. Default is \lstinline![false,false]!.\tabularnewline
+\lstinline!periodicBc! & optional $1 \times 2$ logical array to specify periodic boundary conditions. Default: \lstinline![false,false]!.\tabularnewline
 \hline
-\lstinline!nMaxCurves! & optional maximum number of curves (i.e. shrinklines of stretchlines) to generate. Default is \lstinline!numel(cgEigenvalue)!.\tabularnewline
+\lstinline!nMaxCurves! & optional maximum number of curves (i.e. shrinklines of stretchlines) to generate. Default: \lstinline!numel(cgEigenvalue)!.\tabularnewline
 \hline
@@
 
-\lstinputlisting[caption={Double gyre script for elliptic lambda-line LCSs},label=l:double gyre lambda-line LCS]{listings/double_gyre/lambda_lcs.m}
+\lstinputlisting[caption={Double gyre script for elliptic LCSs},label=l:double gyre lambda-line LCS]{listings/double_gyre/lambda_lcs.m}
 
@@
 \includegraphics[width=\textwidth]{graphics/double_gyre/lambda_strain_lcs}
-\caption{Double gyre lambda-line and shrinkline LCSs}
+\caption{Double gyre elliptic and shrinkline LCSs}
 \label{fig:double_gyre_lambda_strain_lcs}
@@
 \includegraphics[width=\textwidth]{graphics/double_gyre/lambda_stretch_lcs}
-\caption{Double gyre lambda-line and stretchline LCSs}
+\caption{Double gyre elliptic and stretchline LCSs}
 \label{fig:double_gyre_lambda_stretch_lcs}
@@
 time in the flow.
-\Cref{fig:Bickley_jet_lambda_LCS_full} shows lambda-line LCS of the Bickley jet with the FTLE in the background.
+\Cref{fig:Bickley_jet_lambda_LCS_full} shows elliptic LCSs of the Bickley jet with the FTLE in the background.
 The figure demonstrates a poor result since the LCSs change shape significantly between eddies.
@@
 \includegraphics[width=.85\textwidth]{graphics/bickley_jet/lambda_lcs_full}
-\caption{Bickley jet lambda-line LCSs}
+\caption{Bickley jet elliptic LCSs}
 \label{fig:Bickley_jet_lambda_LCS_full}
@@
 
-\Cref{fig:Bickley jet lambda lcs convergence} shows convergence of the lambda-line elliptic LCSs.
+\Cref{fig:Bickley jet lambda lcs convergence} shows convergence of the elliptic LCSs.
 The LCS Tool script used to produce that figure are based on the script \lstinline!demo/bickley_jet/hyperbolic_shear_lcs.m!
@@
 \includegraphics[width=.7\textwidth]{graphics/bickley_jet/lambda_lcs_convergence_600}
-\caption{Bickley jet closed lambda-line at varying resolutions}
+\caption{Bickley jet closed lambda-lines at varying resolutions}
 \label{fig:Bickley jet lambda lcs convergence}
@@
 
-% Lambda-line LCSs
+% Elliptic LCSs
 % Define Poincare sections; ...

### 2014-09-05 main.tex
Remove unused entries from BibTeX file
By switching the Biblatex backend to Biber, the following command
removes unused entries from a BibTeX file:

	biber main.bcf --output_format=bibtex
Move periods outside parentheses

@@
 \usepackage{amsmath}
-\usepackage[uniquename=init,firstinits,backend=bibtex,sorting=none]{biblatex}
+\usepackage[uniquename=init,firstinits,backend=biber,sorting=none]{biblatex}
 \usepackage{calc}
@@
 
-\addbibresource{main}
+\addbibresource{main.bib}
 
@@
 Equating FTLE ridges with LCSs, however, would create theoretical inconsistencies, as well as false positives and negatives in hyperbolic LCS detection\parencite{haller11:_lagran_coher_struc,norgard12:_secon_lagran_coher_struc}.
-In addition, the role of the FTLE field in the accurate detection of elliptic (vortex-type) and parabolic (jet-core type) LCSs has remained an open question (but see \textcite{beron-vera10:_invar_lagran}.)
+In addition, the role of the FTLE field in the accurate detection of elliptic (vortex-type) and parabolic (jet-core type) LCSs has remained an open question (but see \textcite{beron-vera10:_invar_lagran}).
 
@@
 We note that parabolic LCSs (Lagrangian jet cores) are also stationary curves of $Q(\gamma)$, but under variable endpoint perturbations (cf.
-\textcite{farazmand14:_shearless}.)
+\textcite{farazmand14:_shearless}).
 
@@
 The auxiliary grid (cf.
-\cref{t:Cauchy-Green algorithm}) comprises four points placed symmetrically around each point of the Cartesian grid (\cref{f:main and auxiliary grids}.) These points are used to achieve increased accuracy in the finite-difference approximation
+\cref{t:Cauchy-Green algorithm}) comprises four points placed symmetrically around each point of the Cartesian grid (\cref{f:main and auxiliary grids}).
+These points are used to achieve increased accuracy in the finite-difference approximation
 \[
@@
 of $\nabla F_{t_0}^t$ at a point $x_0$ of the Cartesian grid.
-Here $\delta x_i$ is a vector of length $\vert\delta x_i\vert > 0$ that points from the Cartesian grid-point $x_0$ in the $i^\text{th}$ coordinate direction (\cref{f:main and auxiliary grids}.) Computational improvements arising from the use of the auxiliary grid over simply using the nearest points of the main grid were reported in \textcite{farazmand12:_comput_lagran}.
+Here $\delta x_i$ is a vector of length $\vert\delta x_i\vert > 0$ that points from the Cartesian grid-point $x_0$ in the $i^\text{th}$ coordinate direction (\cref{f:main and auxiliary grids}).
+Computational improvements arising from the use of the auxiliary grid over simply using the nearest points of the main grid were reported in \textcite{farazmand12:_comput_lagran}.
 Experience suggests setting the auxiliary grid spacing to 1-10\% of the main grid spacing.
@@
 \item Position Poincare sections in flow domain to specify initial positions of lambda-lines
-\item Integrate $\lambda$-lines tangent to $\eta_\pm^\lambda$ (see \cref{eq:etafields}.)
+\item Integrate $\lambda$-lines tangent to $\eta_\pm^\lambda$ (see \cref{eq:etafields})
 \item Calculate Poincare map
@@
 The Poincare section is to be oriented such that the first endpoint is close to the center of the expected Lagrangian vortex, and the second endpoint is outside this vortex.
-There is no automated procedure procedure implemented in LCS Tool for the positioning of Poincare sections (see however \textcite{karrasch14:_autom_lagran} for a recently developed algorithm.) Additionally, the number of lambda-lines launched from the Poincare section, \lstinline!poincareSection.numPoints!, must be defined.
+There is no automated procedure procedure implemented in LCS Tool for the positioning of Poincare sections (see however \textcite{karrasch14:_autom_lagran} for a recently developed algorithm).
+Additionally, the number of lambda-lines launched from the Poincare section, \lstinline!poincareSection.numPoints!, must be defined.
 A reasonable default value is 100.
@@
 
-As discussed in \cref{sec:Hyperbolic LCSs}, positions of hyperbolic LCSs at time $t_0$ are found as the strongest repelling orbits of the vector field \cref{eq:strainline} (repelling LCSs), and strongest attracting orbits of the vector field \cref{eq:stretchline} (attracting LCSs.) By repulsion and attraction we mean a property of the LCS (as an evolving material line) under the flow map $F_{t_0}^t$.
+As discussed in \cref{sec:Hyperbolic LCSs}, positions of hyperbolic LCSs at time $t_0$ are found as the strongest repelling orbits of the vector field \cref{eq:strainline} (repelling LCSs), and strongest attracting orbits of the vector field \cref{eq:stretchline} (attracting LCSs).
+By repulsion and attraction we mean a property of the LCS (as an evolving material line) under the flow map $F_{t_0}^t$.
 We identify the strongest repelling shrinklines as those crossing a local maximum of the $\lambda_2(x_0)$ field.
@@
 (Right) Final positions $x(t) = (x_1(t),x_2(t)) = F_{t_0}^t(x_0)$ of tracer particles, are plotted with their $x_1(t)$ position encoded in red and $x_2(t)$ position encoded in green.
-Note the continued impact of the repelling hyperbolic LCSs (black) on tracer patterns inside the elliptic LCS (white.)}
+Note the continued impact of the repelling hyperbolic LCSs (black) on tracer patterns inside the elliptic LCS (white).}
 \label{fig:ocean_dataset_colortracer}
@@
 As usual, the local maximization distance is set larger for stretchlines than for shrinklines (cf.
-l.\,4 and l.\,9.) The purpose of the maximization distance is to obtain spatially separated LCSs and to avoid a dense tangle of lines that basically indicates the same hyperbolic LCS (cf.
-\cref{t:Hyperbolic LCS algorithm}, point 5.) Setting the local maximization distance for stretchlines larger than for shrinklines allows obtaining a comparable number of stretchlines and shrinklines overall in the flow domain (recall that hyperbolic LCS seed points are discarded if they are within the local maximization distance of an existing hyperbolic LCS.) Shrinklines are locally tangent to ridges of $\lambda_2$ maxima, whereas stretchlines are locally normal to these ridges.
+l.\,4 and l.\,9).
+The purpose of the maximization distance is to obtain spatially separated LCSs and to avoid a dense tangle of lines that basically indicates the same hyperbolic LCS (cf.
+\cref{t:Hyperbolic LCS algorithm}, point 5).
+Setting the local maximization distance for stretchlines larger than for shrinklines allows obtaining a comparable number of stretchlines and shrinklines overall in the flow domain (recall that hyperbolic LCS seed points are discarded if they are within the local maximization distance of an existing hyperbolic LCS).
+Shrinklines are locally tangent to ridges of $\lambda_2$ maxima, whereas stretchlines are locally normal to these ridges.
 Setting the local maximization distance of stretchlines and shrinklines equal would therefore produce a greater number of stretchlines than shrinklines overall.
@@
 We choose the integration time as $T=30$ days (\cref{l:griddedInterp}, line 4), which is larger than the eddy turnover time in this region.
-The resolution of the main computational grid for initial conditions is set to $400 \times 400$ (line 3.) This corresponds to a resolution of roughly 0.015° and gives good results.
+The resolution of the main computational grid for initial conditions is set to $400 \times 400$ (line 3).
+This corresponds to a resolution of roughly 0.015° and gives good results.
 With this choice, the resolution of the tracer grid is 15 times higher than the resolution of the velocity field.
 The flow is integrated and the Cauchy-Green strain tensor is computed by the function \lstinline!eig_cgStrain!
-(line 35.) Incompressibility of the flow is enforced (line 12), the auxiliary grid distance is set to 1\% of the main grid distance (line 17), and eigenvalues are computed from the auxiliary grid (line 16.) Elliptic LCSs are computed in line 54, after the Poincare sections have been set (line 41-42) and the $\eta_\pm^\lambda$ fields have been defined (line 53.) Hyperbolic LCSs are computed in lines 57 and 59.
+(line 35).
+Incompressibility of the flow is enforced (line 12), the auxiliary grid distance is set to 1\% of the main grid distance (line 17), and eigenvalues are computed from the auxiliary grid (line 16).
+Elliptic LCSs are computed in line 54, after the Poincare sections have been set (line 41-42) and the $\eta_\pm^\lambda$ fields have been defined (line 53).
+Hyperbolic LCSs are computed in lines 57 and 59.
 

### 2014-09-05 main.tex
Move periods outside parentheses
Conflicts:
	main.tex

@@
 To compute the normal repulsion of shrinklines, we note that an infinitesimal
-normal perturbation to a shrinkline $\gamma$ at its point $r$ grows under the flow map $F_{t_{0}}^{t}$ by a factor of $\lambda_{2}(r)$ in the direction normal to $F_{t_{0}}^{t}(\gamma)$.
-Similarly, small normal perturbations to a stretchline decay by a factor $\lambda_{1}(r)$ in the direction normal to the evolving stretchline.
+normal perturbation to a shrinkline $\gamma$ at its point $r$ grows under the flow map $F_{t_{0}}^{t}$ by a factor of $\lambda_{2}(r)$ in the direction normal to $F_{t_{0}}^{t}(\gamma).$ Similarly, small normal perturbations to a stretchline decay by a factor $\lambda_{1}(r)$ in the direction normal to the evolving stretchline.
 
@@
 Experience shows that computing $\lambda_i$ in this order gives higher accuracy than in the reverse order \parencite{farazmand12:_comput_lagran}.
-This procedure should be applied carefully since by changing $\lambda_1$ but not $\xi_1$, the error in satisfying the eigenvalue equation for the Cauchy-Green strain tensor may increase.
 
@@
 These removed segments can be turned back on demand, as shown in \cref{fig:ocean_dataset_colortracer}.
-
-The local maximization distance is the number of grid points over which the Cauchy-Green strain tensor is susceptible to numerical noise from integration.
-This distance is thus expected to be a function of variable time step integration tolerances.
-
-Prior to initiating shrinkline and stretchline integration, a maximum length is defined.
-Shrinklines and stretchlines are then integrated until reaching the domain boundary or the maximum length.
-The maximum length may be seen as safety factor to prevent highly sinuous shrinklines and stretchlines leading to inordinately long computations; the maximum length should be set such that most shrinklines and stretchlines stop at the domain boundary.
-
-Once shrinkline and stretchline integration has begun, it may evolve so as to be within the local minimization distance of an existing shrinkline or stretchline.
-In other words, the local minimization distance affects where shrinklines and stretchlines begin, but not where they finish.
 
@@
 
-\lstinputlisting[caption={Double gyre script for elliptic LCSs},label=l:double gyre lambda-line LCS]{listings/double_gyre/lambda_lcs.m}
+\lstinputlisting[caption={Double gyre script for elliptic lambda-line LCSs},label=l:double gyre lambda-line LCS]{listings/double_gyre/lambda_lcs.m}
 
@@
 \includegraphics[width=\textwidth]{graphics/double_gyre/lambda_strain_lcs}
-\caption{Double gyre elliptic and shrinkline LCSs}
+\caption{Double gyre lambda-line and shrinkline LCSs}
 \label{fig:double_gyre_lambda_strain_lcs}
@@
 \includegraphics[width=\textwidth]{graphics/double_gyre/lambda_stretch_lcs}
-\caption{Double gyre elliptic and stretchline LCSs}
+\caption{Double gyre lambda-line and stretchline LCSs}
 \label{fig:double_gyre_lambda_stretch_lcs}
@@
 time in the flow.
-\Cref{fig:Bickley_jet_lambda_LCS_full} shows elliptic LCSs of the Bickley jet with the FTLE in the background.
+\Cref{fig:Bickley_jet_lambda_LCS_full} shows lambda-line LCS of the Bickley jet with the FTLE in the background.
 The figure demonstrates a poor result since the LCSs change shape significantly between eddies.
@@
 \includegraphics[width=.85\textwidth]{graphics/bickley_jet/lambda_lcs_full}
-\caption{Bickley jet elliptic LCSs}
+\caption{Bickley jet lambda-line LCSs}
 \label{fig:Bickley_jet_lambda_LCS_full}
@@
 
-\Cref{fig:Bickley jet lambda lcs convergence} shows convergence of the elliptic LCSs.
+\Cref{fig:Bickley jet lambda lcs convergence} shows convergence of the lambda-line elliptic LCSs.
 The LCS Tool script used to produce that figure are based on the script \lstinline!demo/bickley_jet/hyperbolic_shear_lcs.m!
@@
 \includegraphics[width=.7\textwidth]{graphics/bickley_jet/lambda_lcs_convergence_600}
-\caption{Bickley jet closed lambda-lines at varying resolutions}
+\caption{Bickley jet closed lambda-line at varying resolutions}
 \label{fig:Bickley jet lambda lcs convergence}
@@
 
-% Elliptic LCSs
+% Lambda-line LCSs
 % Define Poincare sections; ...

### 2014-09-05 main.tex
Replace explicit reference numbers by LaTeX label and reference tags

@@
 \lstloadlanguages{Matlab}
-\lstset{basicstyle=\small,breaklines=true,frame=single,float,columns=fullflexible,numbers=left,stepnumber=2,backgroundcolor=\color{gray}}
+\lstset{
+  basicstyle=\small,
+  breaklines=true,
+  frame=single,
+  float,
+  columns=fullflexible,
+  numbers=left,
+  stepnumber=2,
+  backgroundcolor=\color{gray},
+  escapeinside={\%@}{\^^M}
+}
 
@@
 \item Flag any remaining local maxima of $\lambda_2$ within the maximization distance of the shrinkline as ineligible initial positions for subsequent shrinklines
+\label{i:flag maxima}
 \item Continue integrating shrinklines using local maxima of $\lambda_2$ as initial positions until no eligible local maxima of $\lambda_2$ remain.
@@
 As usual, the local maximization distance is set larger for stretchlines than for shrinklines (cf.
-l.\,4 and l.\,9).
+\cref{ll:strainlineLocalMaxDistance} and \cref{ll:stretchlineLocalMaxDistance}).
 The purpose of the maximization distance is to obtain spatially separated LCSs and to avoid a dense tangle of lines that basically indicates the same hyperbolic LCS (cf.
-\cref{t:Hyperbolic LCS algorithm}, point 5).
+\cref{t:Hyperbolic LCS algorithm}, \cref{i:flag maxima}).
 Setting the local maximization distance for stretchlines larger than for shrinklines allows obtaining a comparable number of stretchlines and shrinklines overall in the flow domain (recall that hyperbolic LCS seed points are discarded if they are within the local maximization distance of an existing hyperbolic LCS).
@@
 
-\lstinputlisting[caption={LCS Tool commands for double gyre hyperbolic LCSs.
-Condensed version of the LCS Tool script \lstinline!demo/double_gyre/hyperbolic_shear_lcs.m!
-},label=lst:double_gyre_lambda_strain_stretch_lcs]{listings/double_gyre/lambda_strain_stretch_lcs.m}
+\lstinputlisting[
+caption={LCS Tool commands for double gyre hyperbolic LCSs. Condensed version of the LCS Tool script \lstinline!demo/double_gyre/hyperbolic_shear_lcs.m!},
+label=lst:double_gyre_lambda_strain_stretch_lcs
+]
+{listings/double_gyre/lambda_strain_stretch_lcs.m}
 
@@
 \Cref{l:griddedInterp} shows the relevant part of the code in LCS Tool's ocean demo file \lstinline!hyperbolic_shear_lcs.m!.
-In lines 8-11, the commands for the interpolation are given of the velocity dataset are given.
+In \crefrange{ll:interpolation start}{ll:interpolation end}, the commands for the interpolation are given of the velocity dataset are given.
 
@@
 domain = [0,6;-34,-28];
-resolution = [400,400];
-timespan = [100,130];
+resolution = [400,400]; %@\label{ll:resolution}
+timespan = [100,130]; %@\label{ll:timespan}
 
@@
 ...
-interpMethod = 'spline';
+interpMethod = 'spline'; %@\label{ll:interpolation start}
 vlon_interpolant = griddedInterpolant({time,lat,lon},vlon,interpMethod);
 vlat_interpolant = griddedInterpolant({time,lat,lon},vlat,interpMethod);
-lDerivative = @(t,x,~)flowdata_derivative(t,x,vlon_interpolant,vlat_interpolant);
-incompressible = true;
+lDerivative = @(t,x,~)flowdata_derivative(t,x,vlon_interpolant,vlat_interpolant); %@\label{ll:interpolation end}
+incompressible = true; %@\label{ll:incompressible}
 
@@
 % Cauchy-Green strain
-cgEigenvalueFromMainGrid = false;
-cgAuxGridRelDelta = 0.01;
+cgEigenvalueFromMainGrid = false; %@\label{ll:cgEigenvalueFromMainGrid}
+cgAuxGridRelDelta = 0.01; %@\label{ll:cgAuxGridRelDelta}
 
@@
 % Cauchy-Green strain eigenvalues and eigenvectors
-[cgEigenvector,cgEigenvalue] = eig_cgStrain(lDerivative,domain,resolution,timespan,'incompressible',incompressible,'eigenvalueFromMainGrid',cgEigenvalueFromMainGrid,'auxGridRelDelta',cgAuxGridRelDelta);
+[cgEigenvector,cgEigenvalue] = eig_cgStrain(lDerivative,domain,resolution,timespan,'incompressible',incompressible,'eigenvalueFromMainGrid',cgEigenvalueFromMainGrid,'auxGridRelDelta',cgAuxGridRelDelta); %@\label{ll:eig_cgStrain}
 
@@
 ...
-poincareSection(1).endPosition = [3.3,-32.1;3.7,-31.6];
-poincareSection(2).endPosition = [1.3,-30.9;1.9,-31.1];
+poincareSection(1).endPosition = [3.3,-32.1;3.7,-31.6]; %@\label{ll:poincareSection(1)}
+poincareSection(2).endPosition = [1.3,-30.9;1.9,-31.1]; %@\label{ll:poincareSection(2)}
 
@@
 end
-[shearline.etaPos,shearline.etaNeg] = lambda_line(cgEigenvector,cgEigenvalue,lambda);
-closedLambdaLine = poincare_closed_orbit_multi(domain,resolution,shearline,poincareSection,'odeSolverOptions',lambdaLineOdeSolverOptions);
+[shearline.etaPos,shearline.etaNeg] = lambda_line(cgEigenvector,cgEigenvalue,lambda); %@\label{ll:lambda_line}
+closedLambdaLine = poincare_closed_orbit_multi(domain,resolution,shearline,poincareSection,'odeSolverOptions',lambdaLineOdeSolverOptions); %@\label{ll:poincare_closed_orbit_multi}
 ...
 % Hyperbolic strainline LCSs
-strainlineLcs = seed_curves_from_lambda_max(strainlineLocalMaxDistance,strainlineMaxLength,cgEigenvalue(:,2),cgEigenvector(:,1:2),domain,resolution,'odeSolverOptions',strainlineOdeSolverOptions);
+strainlineLcs = seed_curves_from_lambda_max(strainlineLocalMaxDistance,strainlineMaxLength,cgEigenvalue(:,2),cgEigenvector(:,1:2),domain,resolution,'odeSolverOptions',strainlineOdeSolverOptions); %@\label{ll:strainlineLcs}
 ...
-stretchlineLcs = seed_curves_from_lambda_max(stretchlineLocalMaxDistance,stretchlineMaxLength,-cgEigenvalue(:,1),cgEigenvector(:,3:4),domain,resolution,'odeSolverOptions',stretchlineOdeSolverOptions);
+stretchlineLcs = seed_curves_from_lambda_max(stretchlineLocalMaxDistance,stretchlineMaxLength,-cgEigenvalue(:,1),cgEigenvector(:,3:4),domain,resolution,'odeSolverOptions',stretchlineOdeSolverOptions); %@\label{ll:stretchlineLcs}
 \end{lstlisting}
 
-We choose the integration time as $T=30$ days (\cref{l:griddedInterp}, line 4), which is larger than the eddy turnover time in this region.
-The resolution of the main computational grid for initial conditions is set to $400 \times 400$ (line 3).
+We choose the integration time as $T=30$ days (\cref{l:griddedInterp}, \cref{ll:timespan}), which is larger than the eddy turnover time in this region.
+The resolution of the main computational grid for initial conditions is set to $400 \times 400$ (\cref{ll:resolution}).
 This corresponds to a resolution of roughly 0.015° and gives good results.
@@
 The flow is integrated and the Cauchy-Green strain tensor is computed by the function \lstinline!eig_cgStrain!
-(line 35).
-Incompressibility of the flow is enforced (line 12), the auxiliary grid distance is set to 1\% of the main grid distance (line 17), and eigenvalues are computed from the auxiliary grid (line 16).
-Elliptic LCSs are computed in line 54, after the Poincare sections have been set (line 41-42) and the $\eta_\pm^\lambda$ fields have been defined (line 53).
-Hyperbolic LCSs are computed in lines 57 and 59.
+(\cref{ll:eig_cgStrain}).
+Incompressibility of the flow is enforced (\cref{ll:incompressible}), the auxiliary grid distance is set to 1\% of the main grid distance (\cref{ll:cgAuxGridRelDelta}), and eigenvalues are computed from the auxiliary grid (\cref{ll:cgEigenvalueFromMainGrid}).
+Elliptic LCSs are computed in \cref{ll:poincare_closed_orbit_multi}, after the Poincare sections have been set (\cref{ll:poincareSection(1),ll:poincareSection(2)}) and the $\eta_\pm^\lambda$ fields have been defined (\cref{ll:lambda_line}).
+Hyperbolic LCSs are computed in \cref{ll:strainlineLcs,ll:stretchlineLcs}.
 

### 2014-09-05 listings/double_gyre/lambda_strain_stretch_lcs.m

@@ -1,12 +1,12 @@
 % Strainlines LCSs
 strainlineMaxLength = 20;
 gridSpace = diff(domain(1,:))/(double(resolution(1))-1);
-strainlineLocalMaxDistance = 2*gridSpace;
+strainlineLocalMaxDistance = 2*gridSpace; %@\label{ll:strainlineLocalMaxDistance}
 
 strainlineLcs = seed_curves_from_lambda_max(strainlineLocalMaxDistance,strainlineMaxLength,cgEigenvalue(:,2),cgEigenvector(:,1:2),domain,resolution);
 
 % Stretchlines LCSs
 stretchlineMaxLength = 20;
-stretchlineLocalMaxDistance = 10*gridSpace;
+stretchlineLocalMaxDistance = 10*gridSpace;  %@\label{ll:stretchlineLocalMaxDistance}
 
 stretchlineLcs = seed_curves_from_lambda_max(stretchlineLocalMaxDistance,stretchlineMaxLength,-cgEigenvalue(:,1),cgEigenvector(:,3:4),domain,resolution);

### 2014-10-02 main.tex
Change LaTeX document class to Elsevier article

--- 2014-09-05 main.tex
@@
-\documentclass{article}
+\documentclass{elsarticle}
 
 \usepackage{amsmath}
-\usepackage[uniquename=init,firstinits,backend=biber,sorting=none]{biblatex}
-\usepackage{calc}
-\usepackage{color}
 \usepackage{commath}
-\usepackage{fontspec}
-\usepackage{gensymb}
-\usepackage{graphicx}
+\usepackage[utf8]{inputenc}
 \usepackage{listings}
 \usepackage[numbered,framed]{mcode}
-\usepackage[space]{grffile}
-\usepackage[hidelinks]{hyperref}
 \usepackage{cleveref}
-
-\hypersetup{breaklinks=true}
 
@@
 
-\title{LCS Tool : A Computational Platform for Lagrangian Coherent Structures}
-
-\author{Kristjan Onu \and Florian Huhn \and George Haller}
-
-\addbibresource{main.bib}
+\journal{Journal of Computational Science}
 
@@
 
-\numberwithin{equation}{section}
-\numberwithin{figure}{section}
-\numberwithin{lstlisting}{section}
-\numberwithin{table}{section}
+\begin{frontmatter}
 
-\maketitle
+\title{LCS Tool: A Computational Platform for Lagrangian Coherent Structures}
 
-\abstract{
+\author{K.~Onu}
+
+\author{F.~Huhn}
+
+\author{G.~Haller\corref{cor}}
+\ead{georgehaller@ethz.ch}
+
+\cortext[cor]{Corresponding author}
+
+\address{Institute of Mechanical Systems, ETH Zurich, Switzerland}
+
+\begin{abstract}
 We give an algorithmic introduction to Lagrangian coherent structures (LCSs) using a newly developed computational engine, LCS Tool.
@@
 After a review of the underlying theory, we explain the steps and numerical methods used by LCS Tool, and illustrate its capabilities on three unsteady fluid flow examples.
-}
+\end{abstract}
 
-\tableofcontents
+\begin{keyword}
+Dynamical systems \sep Lagrangian coherent structures \sep Invariant manifolds \sep Mixing \ Mathematical software
+
+% FIXME Verify these classifications, http://www.aip.org/pacs
+\PACS 47.10.Fg \sep 47.27.De \sep 47.11.-j
+
+% FIXME Verify these classifications http://www.ams.org/msc/msc2010.html
+\MSC[2010] 37N10 \sep 37B25 \sep 97N80
+\end{keyword}
+
+\end{frontmatter}
 
@@
 
-Lagrangian Coherent Structures (LCSs) are evolving organizing centers of trajectory patterns in non-autonomous dynamical systems\parencite{haller00:_lagran,peacock13:_lagran,haller15:_langr_coher_struc}.
-Applications of LCSs include oceanic and atmospheric flows\parencite{beron-vera13:_objec_agulh,koh02:_hyper}, biological transport problems\parencite{wilson09:_lagran_reynol,tallapragada11:_lagran,huhn12:_south_indian_ocean_count_madag}, aeronautics\parencite{tang10:_accur_lagran_hong_kong_inter_airpor}, celestial mechanics\parencite{gawlik09:_lagran}, crowd dynamics\parencite{ali07:_lagran_partic_dynam_approac_crowd}, and aperiodically forced mechanical oscillators\parencite{hadjighasem13:_detec_kam}.
+Lagrangian Coherent Structures (LCSs) are evolving organizing centers of trajectory patterns in non-autonomous dynamical systems\citep{haller00:_lagran,peacock13:_lagran,haller15:_langr_coher_struc}.
+Applications of LCSs include oceanic and atmospheric flows\citep{beron-vera13:_objec_agulh,koh02:_hyper}, biological transport problems\citep{wilson09:_lagran_reynol,tallapragada11:_lagran,huhn12:_south_indian_ocean_count_madag}, aeronautics\citep{tang10:_accur_lagran_hong_kong_inter_airpor}, celestial mechanics\citep{gawlik09:_lagran}, crowd dynamics\citep{ali07:_lagran_partic_dynam_approac_crowd}, and aperiodically forced mechanical oscillators\citep{hadjighasem13:_detec_kam}.
 
-\textcite{haller01:_distin} proposed that ridges of the finite-time Lyapunov exponent (FTLE) are heuristic indicators of hyperbolic (i.e., repelling- and attracting-type) LCSs.
-A number of examples indeed support this principle\parencite{peacock13:_lagran}.
-Equating FTLE ridges with LCSs, however, would create theoretical inconsistencies, as well as false positives and negatives in hyperbolic LCS detection\parencite{haller11:_lagran_coher_struc,norgard12:_secon_lagran_coher_struc}.
-In addition, the role of the FTLE field in the accurate detection of elliptic (vortex-type) and parabolic (jet-core type) LCSs has remained an open question (but see \textcite{beron-vera10:_invar_lagran}).
+\citet{haller01:_distin} proposed that ridges of the finite-time Lyapunov exponent (FTLE) are heuristic indicators of hyperbolic (i.e., repelling- and attracting-type) LCSs.
+A number of examples indeed support this principle\citep{peacock13:_lagran}.
+Equating FTLE ridges with LCSs, however, would create theoretical inconsistencies, as well as false positives and negatives in hyperbolic LCS detection\citep{haller11:_lagran_coher_struc,norgard12:_secon_lagran_coher_struc}.
+In addition, the role of the FTLE field in the accurate detection of elliptic (vortex-type) and parabolic (jet-core type) LCSs has remained an open question (but see \citet{beron-vera10:_invar_lagran}).
 
-More recent work has focused on an exact mathematical formulation of the properties defining LCSs\parencite{haller11:_lagran_coher_struc,farazmand12:_comput_lagran,haller12:_geodes_theor_trans_barrier_two_dimen_flows,haller13:_coher_lagran,blazevski14:_hyper_ellip_trans_barrier_three}.
-In two-dimensional flows, hyperbolic and parabolic LCSs turn out to be stationary curves of the averaged material shear\parencite{farazmand14:_shearless}, whereas elliptic LCSs are stationary curves of the averaged
-strain\parencite{haller13:_coher_lagran}. These variational formulations lead to explicit solutions for LCSs as null-geodesics of appropriate Lorentzian metrics.
+More recent work has focused on an exact mathematical formulation of the properties defining LCSs\citep{haller11:_lagran_coher_struc,farazmand12:_comput_lagran,haller12:_geodes_theor_trans_barrier_two_dimen_flows,haller13:_coher_lagran,blazevski14:_hyper_ellip_trans_barrier_three}.
+In two-dimensional flows, hyperbolic and parabolic LCSs turn out to be stationary curves of the averaged material shear\citep{farazmand14:_shearless}, whereas elliptic LCSs are stationary curves of the averaged
+strain\citep{haller13:_coher_lagran}. These variational formulations lead to explicit solutions for LCSs as null-geodesics of appropriate Lorentzian metrics.
 
 Here we present a simple algorithmic introduction to geodesic LCS detection without details on the the underlying mathematics.
-We then review the implementation of this approach in a computational engine called LCS Tool\footnote{LCS Tool is publicly available at: \emph{\url{https://github.com/jeixav/LCS-Tool}}.}.
-This engine is a library of MATLAB functions that extract LCSs from two-dimensional unsteady flows.
-The examples we present here have been programmed into demonstration scripts.
+% FIXME Replace GitHub link with http://runmycode.org/
+We then review the implementation of this approach in a computational engine called LCS Tool\footnote{LCS Tool is publicly available at: https://github.com/jeixav/LCS-Tool}.
+This engine is a library of MATLAB functions that extract LCSs from two-dimensional unsteady flows. The examples we present here have been programmed into demonstration scripts.
 
@@
 \begin{itemize}
-\item \emph{MANGEN}\parencite{lekien03:_time} calculates FTLE and advects material curves in two dimensional velocity fields.
+\item \emph{MANGEN}\citep{lekien03:_time} calculates FTLE and advects material curves in two dimensional velocity fields.
 It also includes a graphical user interface and uses MPI for parallel calculations.
-\item \emph{LCS MATLAB Kit}\parencite{dabiri09:_lmk} calculates the FTLE from velocity datasets. Includes a graphical user interface.
-\item \emph{Newman}\parencite{toit10:_trans} calculates the FTLE in $N$ dimensions. Assists ridge extraction of FTLE fields. Supports analytic  and dataset based velocity definitions.
-\item \emph{FlowVC}\parencite{shadden10:_flowvc} is a general purpose LCS platform for two and three dimensional datasets. Parallel calculations are supported with Open\-MP, CUDA and OpenCL.
+% FIXME Cannot cite LCS MATLAB Kit without a journal reference
+% \item \emph{LCS MATLAB Kit}\citep{dabiri09:_lmk} calculates the FTLE from velocity datasets. Includes a graphical user interface.
+\item \emph{Newman}\citep{toit10:_trans} calculates the FTLE in $N$ dimensions. Assists ridge extraction of FTLE fields. Supports analytic  and dataset based velocity definitions.
+% FIXME Cannot cite FlowVC without a journal reference
+% \item \emph{FlowVC}\citep{shadden10:_flowvc} is a general purpose LCS platform for two and three dimensional datasets. Parallel calculations are supported with Open\-MP, CUDA and OpenCL.
+% FIXME FlowTK
+% https://github.com/FlowPhysics/FlowTK
+% doi:10.1007%2F978-3-319-04099-8_13
 \end{itemize}
@@
 \end{equation}
-measuring Lagrangian strain in the velocity field. This tensor is symmetric and positive definite\parencite{truesdell04}. We label the eigenvalues and eigenvectors of $C_{t_0}^t(x_0)$ as follows:
+measuring Lagrangian strain in the velocity field. This tensor is symmetric and positive definite\citep{truesdell04}. We label the eigenvalues and eigenvectors of $C_{t_0}^t(x_0)$ as follows:
 \[
@@
 We seek positions of closed material lines at time $t_0$ that prevail as coherent Lagrangian vortex boundaries (or \emph{elliptic LCSs}) over a time interval $[t_0,t]\subset[t_-,t_+]$.
-\textcite{haller13:_coher_lagran} argue that such initial material line positions are closed stationary curves of the averaged strain functional
+\citet{haller13:_coher_lagran} argue that such initial material line positions are closed stationary curves of the averaged strain functional
 \[
@@
 For this reason, we refer to trajectories of \cref{eq:etafields} as $\lambda$-lines.
-Following \textcite{haller13:_coher_lagran}, we call the outermost member of a closed family of $\lambda$-lines a coherent Lagrangian vortex boundary.
+Following \citet{haller13:_coher_lagran}, we call the outermost member of a closed family of $\lambda$-lines a coherent Lagrangian vortex boundary.
 
@@
 Next we consider positions of material lines at time $t_0$ that prevail as most repelling or attracting material lines (or \emph{hyperbolic LCSs}) over a time interval $[t_0,t] \subset [t_-,t_+]$.
-\textcite{farazmand14:_shearless} argue that hyperbolic LCSs are stationary curves of the averaged shear functional
+\citet{farazmand14:_shearless} argue that hyperbolic LCSs are stationary curves of the averaged shear functional
 \begin{gather*}
@@
 We note that parabolic LCSs (Lagrangian jet cores) are also stationary curves of $Q(\gamma)$, but under variable endpoint perturbations (cf.
-\textcite{farazmand14:_shearless}).
+\citet{farazmand14:_shearless}).
 
@@
 normal perturbation to a shrinkline $\gamma$ at its point $r$ grows under the flow map $F_{t_{0}}^{t}$ by a factor of $\lambda_{2}(r)$ in the direction normal to $F_{t_{0}}^{t}(\gamma).$ Similarly, small normal perturbations to a stretchline decay by a factor $\lambda_{1}(r)$ in the direction normal to the evolving stretchline.
-
-\clearpage
 
@@
 Here $\delta x_i$ is a vector of length $\vert\delta x_i\vert > 0$ that points from the Cartesian grid-point $x_0$ in the $i^\text{th}$ coordinate direction (\cref{f:main and auxiliary grids}).
-Computational improvements arising from the use of the auxiliary grid over simply using the nearest points of the main grid were reported in \textcite{farazmand12:_comput_lagran}.
+Computational improvements arising from the use of the auxiliary grid over simply using the nearest points of the main grid were reported in \citet{farazmand12:_comput_lagran}.
 Experience suggests setting the auxiliary grid spacing to 1-10\% of the main grid spacing.
@@
 At these points the eigenvectors $\xi_1(x_0)$ and $\xi_2(x_0)$ are no longer well-defined.
-This generically arises at a finite set of isolated points within the computational domain\parencite{delmarcelle94}, and hence lie off the computational grid with probability one.
+This generically arises at a finite set of isolated points within the computational domain\citep{delmarcelle94}, and hence lie off the computational grid with probability one.
 
@@
 Incompressible flows (i.e.
-those satisfying $\nabla \cdot v=0$) satisfy the relation $\lambda_1(x_0) \lambda_2(x_0) = 1$ at all points of the computational domain \parencite{arnold78:_mathem}.
+those satisfying $\nabla \cdot v=0$) satisfy the relation $\lambda_1(x_0) \lambda_2(x_0) = 1$ at all points of the computational domain \citep{arnold78:_mathem}.
 Incompressibility can be computationally imposed by first calculating $\lambda_2(x_0)$, then setting $\lambda_1(x_0) = 1/\lambda_2(x_0)$ and calculating the strain eigenvectors $\xi_2$ from $\lambda_2$, then $\xi_1$ from the relationship in \cref{eq:CG_invariants}.
-Experience shows that computing $\lambda_i$ in this order gives higher accuracy than in the reverse order \parencite{farazmand12:_comput_lagran}.
+Experience shows that computing $\lambda_i$ in this order gives higher accuracy than in the reverse order \citep{farazmand12:_comput_lagran}.
 
@@
 A $\lambda$-line returning to its starting point is then an elliptic LCS.
-The outermost member of a family of closed $\lambda$-lines (obtained by varying $\lambda$) is a coherent Lagrangian vortex boundary\parencite{haller13:_coher_lagran}.
+The outermost member of a family of closed $\lambda$-lines (obtained by varying $\lambda$) is a coherent Lagrangian vortex boundary\citep{haller13:_coher_lagran}.
 
@@
 The Poincare section is to be oriented such that the first endpoint is close to the center of the expected Lagrangian vortex, and the second endpoint is outside this vortex.
-There is no automated procedure procedure implemented in LCS Tool for the positioning of Poincare sections (see however \textcite{karrasch14:_autom_lagran} for a recently developed algorithm).
+There is no automated procedure procedure implemented in LCS Tool for the positioning of Poincare sections (see however \citet{karrasch14:_autom_lagran} for a recently developed algorithm).
 Additionally, the number of lambda-lines launched from the Poincare section, \lstinline!poincareSection.numPoints!, must be defined.
@@
 
-\clearpage
-
 \section{Examples}
@@
 \subsection{Double gyre}
-The double gyre is a model for a time-dependent two gyre system observed in geophysical flows \parencite{shadden05:_defin_lagran_lyapun}.
+The double gyre is a model for a time-dependent two gyre system observed in geophysical flows \citep{shadden05:_defin_lagran_lyapun}.
 The model consists of two counter rotating sinusoidal vortices with a harmonically oscillating line in-between.
@@
 
-\clearpage
-
 \subsection{Bickley jet}
@@
 counter rotating vortices.
-This is an idealized model of geophysical flows such as the Gulf Stream and the polar night jet perturbed by a Rossby wave\parencite{castilloNegrete93:_chaot_rossb,beron-vera10:_invar_lagran}.
+This is an idealized model of geophysical flows such as the Gulf Stream and the polar night jet perturbed by a Rossby wave\citep{castilloNegrete93:_chaot_rossb,beron-vera10:_invar_lagran}.
 
@@
 
-\clearpage
-
 \subsection{Ocean dataset}
@@
 In contrast to the previous two analytic examples, the velocity field is available only with discrete temporal and spatial resolution.
-Our region of interest is a small domain in the South Atlantic Ocean, where exceptionally coherent eddies, Agulhas rings, were recently found by \textcite{haller13:_coher_lagran} using the theory we surveyed in \cref{sec:Elliptic LCSs}.
+Our region of interest is a small domain in the South Atlantic Ocean, where exceptionally coherent eddies, Agulhas rings, were recently found by \citet{haller13:_coher_lagran} using the theory we surveyed in \cref{sec:Elliptic LCSs}.
 
@@
 
-\Cref{f:ocean dataset hyperbolic shear lcs details strainline,f:ocean dataset hyperbolic shear lcs details stretchline} show the positions of elliptic and hyperbolic LCSs on 22 November 2006, which is the same time instant analyzed in \textcite{haller13:_coher_lagran,beron-vera13:_objec_agulh}.
+\Cref{f:ocean dataset hyperbolic shear lcs details strainline,f:ocean dataset hyperbolic shear lcs details stretchline} show the positions of elliptic and hyperbolic LCSs on 22 November 2006, which is the same time instant analyzed in \citet{haller13:_coher_lagran,beron-vera13:_objec_agulh}.
 Here the integration time is chosen as $T = 30$ days, as opposed to 90 days in the cited references, to avoid tangling of hyperbolic LCSs.
@@
 The large coherent eddy at $(3,-32)$ has a non-stretching boundary, i.e., $\lambda = 1$.
-It corresponds to eddy \#2 in Figure~3 of \textcite{beron-vera13:_objec_agulh}.
+It corresponds to eddy \#2 in Figure~3 of \citet{beron-vera13:_objec_agulh}.
 Additionally, four smaller coherent eddy cores are found.
@@
 
-\clearpage
-
 \section{Conclusions}
@@
 LCS Tool leverages the capabilities of MATLAB extensively.
-For FTLE based extraction of LCSs, computational performance has received considerable attention \parencite{conti12:_gpu_apu_finit_time_lyapun_expon,miron12:_anisot_lagran_coher_struc}.
+For FTLE based extraction of LCSs, computational performance has received considerable attention \citep{conti12:_gpu_apu_finit_time_lyapun_expon,miron12:_anisot_lagran_coher_struc}.
 We hope to see similar computational advances for the publicly available LCS Tool platform.
-Optimizing computational performance will likely aid the application of variational LCS methods to large-scale real-time forecasting applications, such as the tracking of environmental contaminants\parencite{olascoaga12:_forec}.
+Optimizing computational performance will likely aid the application of variational LCS methods to large-scale real-time forecasting applications, such as the tracking of environmental contaminants\citep{olascoaga12:_forec}.
 
@@
 implementation of recent theoretical advances.
-These include the geodesic theory of parabolic LCSs (jet cores)\parencite{farazmand14:_shearless} and the variational theory of hyperbolic and elliptic LCSs for three-dimensional
-flows\parencite{blazevski14:_hyper_ellip_trans_barrier_three}.
+These include the geodesic theory of parabolic LCSs (jet cores)\citep{farazmand14:_shearless} and the variational theory of hyperbolic and elliptic LCSs for three-dimensional
+flows\citep{blazevski14:_hyper_ellip_trans_barrier_three}.
 
-\subsection*{Acknowledgments}
+\section*{Acknowledgments}
 
-The altimeter products used in this work are produced by SSALTO/DUACS and distributed by AVISO, with support from CNES \url{http://www.aviso.oceanobs.com/duacs}.
+The altimeter products used in this work are produced by SSALTO/DUACS and distributed by AVISO, with support from CNES, http://www.aviso.oceanobs.com/duacs.
 
-\printbibliography
+\section*{References}
+
+\bibliographystyle{elsarticle-num-names} 
+\bibliography{main}
 

### 2014-10-03 main.tex
Add FIXME comments to indicate potential problems & needed revisions

@@
+% FIXME Add 5p option to get JCompSci two column layout
 \documentclass{elsarticle}
@@
 \usepackage{commath}
+\usepackage{gensymb}
 \usepackage[utf8]{inputenc}
@@
 
-\author{G.~Haller\corref{cor}}
+% FIXME Confirm George can be corresponding author
+\author{G.~Haller}
 \ead{georgehaller@ethz.ch}
@@
 \begin{keyword}
-Dynamical systems \sep Lagrangian coherent structures \sep Invariant manifolds \sep Mixing \ Mathematical software
+% FIXME Match keywords and PACS and MSC classifications
+% FIXME Consult previous publications to find keywords
+Lagrangian coherent structures \sep Mathematical software \sep Dynamical systems \sep Invariant manifolds \sep Mixing
 
@@
 
-The data is given at a spatial resolution of $1/4°$ and a temporal resolution of $7\,\mathrm{days}$.
+The data is given at a spatial resolution of $1/4\degree$ and a temporal resolution of $7\,\mathrm{days}$.
 Due to the discrete data, defining the right hand side of \cref{eq:dtlon} and \cref{eq:dtlat} involves spline interpolation in space and time.
@@
 
+% FIXME Keywords are required to be in British English (http://www.elsevier.com/journals/journal-of-computational-science/1877-7503/guide-for-authors), so should switch entire document to British English
 % Local Variables:

### 2014-10-03 main.tex
Update figures to match submission
MATLAB code to generate figures removed

@@
 \usepackage{cleveref}
+
+% FIXME Some figures use PDF version 1.6, PDFLaTeX default is 1.5
+\pdfoptionpdfminorversion 6
 
@@
 Motivated by convention and for clarity in visualization, however, the default setting of LCS Tool removes hyperbolic LCS segments from the interior of elliptic LCSs.
-These removed segments can be turned back on demand, as shown in \cref{fig:ocean_dataset_colortracer}.
+These removed segments can be turned back on demand, as shown in \cref{f:ocean_dataset_colortracer}.
 
@@
 Note the continued impact of the repelling hyperbolic LCSs (black) on tracer patterns inside the elliptic LCS (white).}
-\label{fig:ocean_dataset_colortracer}
+\label{f:ocean_dataset_colortracer}
 \end{figure}
@@
 
-In \cref{fig:double_gyre_lambda_lcs_convergence}, the resolution of the  Cauchy-Green strain tensor is varied from $500 \times 250$ to $1000 \times 500$.
+In \cref{f:double gyre lambda LCS convergence}, the resolution of the  Cauchy-Green strain tensor is varied from $500 \times 250$ to $1000 \times 500$.
 The location of the outermost closed lambda-line changes little, demonstrating convergence.
@@
 \centering
-\includegraphics[width=.8\textwidth]{graphics/double_gyre/lambda_lcs_convergence_500}
-\includegraphics[width=.8\textwidth]{graphics/double_gyre/lambda_lcs_convergence_750}
-\includegraphics[width=.8\textwidth]{graphics/double_gyre/lambda_lcs_convergence_1000}
-\caption{Double gyre closed lambda-lines at varying resolutions with FTLE background}
-\label{fig:double_gyre_lambda_lcs_convergence}
+\includegraphics[width=.8\textwidth]{graphics/double_gyre/lambda_lcs_convergence}\caption{Convergence of closed $\lambda$-lines for increasing main-grid resolution in the double-gyre model.
+The outermost closed $\lambda$-line (bold green line) is the vortex boundary.
+(Top) {[}500x250{]}, $\lambda$-values
+of closed orbits from interior to exterior, left gyre: $\lambda=(0.97, 0.98, 0.99, 1.00, 1.01)$, right gyre: $\lambda=(1.00, 1.01, 1.02, 1.03, 1.04)$; (Middle) {[}750x375{]}, left: $\lambda=(0.97, 0.98, 0.99, 1.00, 1.01)$, right: $\lambda=(1.00, 1.01, 1.02, 1.03, 1.04, 1.05)$; (Bottom) {[}1000x500{]}, left: $\lambda=(0.97, 0.98, 0.99, 1.00, 1.01)$, right: $\lambda=(1.00, 1.01, 1.02, 1.03, 1.04, 1.05)$.}
+\label{f:double gyre lambda LCS convergence}
 \end{figure}
 
-\Cref{fig:double_gyre_lambda_strain_lcs} shows the complete picture of elliptic and hyperbolic LCSs of the double gyre at this resolution.
+\Cref{f:double gyre lambda hyperbolic LCS} shows the complete picture of elliptic and hyperbolic LCSs of the double gyre at this resolution.
 \Cref{lst:double_gyre_lambda_strain_stretch_lcs} lists the corresponding LCS Tool commands.
@@
 \centering
-\includegraphics[width=\textwidth]{graphics/double_gyre/lambda_strain_lcs}
-\caption{Double gyre lambda-line and shrinkline LCSs}
-\label{fig:double_gyre_lambda_strain_lcs}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{graphics/double_gyre/lambda_stretch_lcs}
-\caption{Double gyre lambda-line and stretchline LCSs}
-\label{fig:double_gyre_lambda_stretch_lcs}
+\includegraphics[width=\textwidth]{graphics/double_gyre/lambda_hyperbolic_lcs}
+\caption{LCSs in the double gyre flow.
+Resolution $[500\times250]$.
+$\lambda = 1.00$ and $1.04$ for the left and the right gyre.
+$\lambda \in [0.93,1.07],\,\Delta\lambda=0.01$.
+(Top) Elliptic (green) and repelling (red) LCSs with the FTLE shown in the background.
+(Bottom) Elliptic and attracting (blue) LCSs.}
+\label{f:double gyre lambda hyperbolic LCS}
 \end{figure}
@@
 
-The integration time is $T=4L_x/U$, a multiple of a the eddy turnover
-time in the flow.
-\Cref{fig:Bickley_jet_lambda_LCS_full} shows lambda-line LCS of the Bickley jet with the FTLE in the background.
+The integration time is $T=4L_x/U$, a multiple of a the eddy turnover time in the flow.
+\Cref{f:Bickley jet LCS} shows lambda-line LCS of the Bickley jet with the FTLE in the background.
 The figure demonstrates a poor result since the LCSs change shape significantly between eddies.
@@
 
+\Cref{f:Bickley jet LCS} shows elliptic and hyperbolic LCSs of the Bickley jet with the FTLE field in the background.
+
 \begin{figure}
 \centering
-\includegraphics[width=.85\textwidth]{graphics/bickley_jet/lambda_lcs_full}
-\caption{Bickley jet lambda-line LCSs}
-\label{fig:Bickley_jet_lambda_LCS_full}
+\includegraphics[width=.85\textwidth]{graphics/bickley_jet/lambda_hyperbolic_lcs}
+\caption{Elliptic and hyperbolic (top: red - repelling, bottom: blue - attracting) LCSs in the Bickley jet.
+$\lambda$-valuestrains of coherent eddies from left to right [0.95,\,0.80,\,0.94,\,0.80,\,0.94], $\lambda \in [0.80,1.20],\,\Delta\lambda=0.01$.}
+\label{f:Bickley jet LCS}
 \end{figure}
-
-\Cref{fig:Bickley jet lambda lcs convergence} shows convergence of the lambda-line elliptic LCSs.
-The LCS Tool script used to produce that figure are based on the script \lstinline!demo/bickley_jet/hyperbolic_shear_lcs.m!
-which is given in \cref{lst:Bickley jet hyperbolic_shear_lcs subset}.
-It is worth pointing out that in order to obtain closed orbits, $\lambda = 0.995$; with $\lambda = 1$ no closed orbits exist.
 
@@
 Subset of \lstinline!demo/bickley_jet/hyperbolic_shear_lcs.m!},label=lst:Bickley jet hyperbolic_shear_lcs subset]{listings/bickley_jet/hyperbolic_shear_lcs_subset.m}
-
-\begin{figure}[hbt]
-\centering
-\includegraphics[width=.7\textwidth]{graphics/bickley_jet/lambda_lcs_convergence_400}
-\includegraphics[width=.7\textwidth]{graphics/bickley_jet/lambda_lcs_convergence_500}
-\includegraphics[width=.7\textwidth]{graphics/bickley_jet/lambda_lcs_convergence_600}
-\caption{Bickley jet closed lambda-line at varying resolutions}
-\label{fig:Bickley jet lambda lcs convergence}
-\end{figure}
-
-Shrinkline LCSs are shown in \cref{fig:Bickley_jet_strain_lcs} and stretchline LCSs are shown in \cref{fig:Bickley_jet_stretch_lcs}.
-
-\begin{figure}
-\centering
-\includegraphics[width=.85\textwidth]{graphics/bickley_jet/strain_lcs}
-\caption{Bickley jet shrinkline LCSs}
-\label{fig:Bickley_jet_strain_lcs}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=.85\textwidth]{graphics/bickley_jet/stretch_lcs}
-  \caption{Bickley jet stretchline LCSs}
-  \label{fig:Bickley_jet_stretch_lcs}
-\end{figure}
 
@@
 
-\Cref{f:ocean dataset hyperbolic shear lcs details strainline,f:ocean dataset hyperbolic shear lcs details stretchline} show the positions of elliptic and hyperbolic LCSs on 22 November 2006, which is the same time instant analyzed in \citet{haller13:_coher_lagran,beron-vera13:_objec_agulh}.
+\Cref{f:ocean dataset LCS} shows the positions of elliptic and hyperbolic LCSs on 22 November 2006, which is the same time instant analyzed in \citet{haller13:_coher_lagran,beron-vera13:_objec_agulh}.
 Here the integration time is chosen as $T = 30$ days, as opposed to 90 days in the cited references, to avoid tangling of hyperbolic LCSs.
@@
 \begin{center}
-\includegraphics[width=.8\textwidth]{graphics/ocean_dataset/hyperbolic_shear_lcs_strainlines}
+\includegraphics[width=.8\textwidth]{graphics/ocean_dataset/lambda_hyperbolic_lcs}
 \end{center}
-\caption{Elliptic LCSs (green) and repelling LCSs (red) in the ocean dataset.
+\caption{(Top) Elliptic LCSs (green, $\lambda=$ 1.00, 1.08, 0.94, 0.90, 1.06) and repelling LCSs (red) in the ocean data set.
 Background gray-scale is the finite time Lyapunov exponent.
 Dashed green lines are Poincare sections.
-White dots indicate $\lambda_2$ local maxima used as initial positions for shrinkline integration.}
-\label{f:ocean dataset hyperbolic shear lcs details strainline}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=0.8\textwidth]{graphics/ocean_dataset/hyperbolic_shear_lcs_stretchlines}
-\end{center}
-\caption{Elliptic LCSs (green) and attracting LCSs (blue) in the ocean dataset.
-Background gray-scale is the finite time Lyapunov exponent.
-Dashed green lines are Poincare sections.
-White dots indicate $\lambda_1$ local minima used as initial positions for stretchline integration.}
-\label{f:ocean dataset hyperbolic shear lcs details stretchline}
+White dots indicate $\lambda_{2}$ local maxima used as initial positions for shrinkline integration.
+(Bottom) Elliptic LCSs (green) and attracting LCSs (blue).}
+\label{f:ocean dataset LCS}
 \end{figure}

### 2014-10-03 graphics/bickley_jet/lambda_lcs_full.m

@@ -1,90 +1,0 @@
-function lambda_lcs_full
-
-%% Input parameters
-u = 62.66;
-lengthX = pi*earthRadius;
-lengthY = 1.77e6;
-epsilon = [.075,.4,.3];
-domain = [0,lengthX;[-1,1]*2.25*lengthY];
-resolutionX = 750;
-timespan = [0,4*lengthX/u];
-
-%% Velocity definition
-perturbationCase = 3;
-phiTimespan = [0,25];
-phiInitial = [0,0];
-phiSol = ode45(@d_phi,phiTimespan,phiInitial);
-timeResolution = 1e5;
-phi1 = deval(phiSol,linspace(phiTimespan(1),phiTimespan(2),timeResolution),1);
-phi1Max = max(phi1);
-lDerivative = @(t,x,~)derivative(t,x,false,u,lengthX,lengthY,epsilon,perturbationCase,phiSol,phi1Max);
-incompressible = true;
-periodicBc = [true,false];
-
-%% LCS parameters
-% Cauchy-Green strain
-cgStrainOdeSolverOptions = odeset('relTol',1e-3);
-
-% Lambda-lines
-lambda = 1;
-lambdaLineOdeSolverOptions = odeset('relTol',1e-4);
-poincareSection = struct('endPosition',{},'numPoints',{},'orbitMaxLength',{});
-poincareSection(1).endPosition = [6.5,-1.4;5,-3]*1e6;
-poincareSection(2).endPosition = [1.35e7,-1.4e6;1.5e7,-.5e6];
-poincareSection(3).endPosition = [3.25,1.5;1.4,2.6]*1e6;
-poincareSection(4).endPosition = [1e7,1.5e6;8e6,2.6e6];
-poincareSection(5).endPosition = [1.65e7,1.5e6;1.5e7,2.6e6];
-[poincareSection.numPoints] = deal(100);
-nPoincareSection = numel(poincareSection);
-for i = 1:nPoincareSection
-    rOrbit = hypot(diff(poincareSection(i).endPosition(:,1)),diff(poincareSection(i).endPosition(:,2)));
-    poincareSection(i).orbitMaxLength = 2*(2*pi*rOrbit);
-end
-lambdaLineColor = [0,.6,0];
-
-% Make x and y grid spacing as equal as possible
-gridSpace = diff(domain(1,:))/(double(resolutionX)-1);
-resolutionY = round(diff(domain(2,:))/gridSpace) + 1;
[…]

### 2014-10-03 graphics/bickley_jet/startup.m

@@ -1,5 +1,0 @@
-graphics_root = fullfile('..');
-lcs_tool_root = fullfile('..','..','lcs_tool');
-demo = fullfile(lcs_tool_root,'demo','bickley_jet');
-
-addpath(graphics_root,lcs_tool_root,demo)

### 2014-10-03 graphics/bickley_jet/strain_lcs.m

@@ -1,67 +1,0 @@
-function strain_lcs
-
-%% Input parameters
-u = 62.66;
-lengthX = pi*earthRadius;
-lengthY = 1.77e6;
-epsilon = [.075,.4,.3];
-domain = [2e6,.55*lengthX;[-1,.25]*2.25*lengthY];
-resolutionX = 500;
-timespan = [0,2*lengthX/u];
-
-%% Velocity definition
-perturbationCase = 3;
-phiTimespan = [0,25];
-phiInitial = [0,0];
-phiSol = ode45(@d_phi,phiTimespan,phiInitial);
-timeResolution = 1e5;
-phi1 = deval(phiSol,linspace(phiTimespan(1),phiTimespan(2),timeResolution),1);
-phi1Max = max(phi1);
-lDerivative = @(t,x,~)derivative(t,x,false,u,lengthX,lengthY,epsilon,perturbationCase,phiSol,phi1Max);
-incompressible = true;
-
-%% LCS parameters
-% Cauchy-Green strain
-cgStrainOdeSolverOptions = odeset('relTol',1e-4);
-
-% Strainlines
-strainlineMaxLength = 1e8;
-strainlineOdeSolverOptions = odeset('relTol',1e-4);
-strainlineColor = 'r';
-lcsInitialPositionMarkerSize = 2;
-
-warning('off','eig_cgStrain:unequalDelta')
-warning('off','eig_cgStrain:unequalAuxGridDelta')
-warning('off','integrate_line:isDiscontinuousLargeAngle')
-
-% Make x and y grid spacing as equal as possible
-gridSpace = diff(domain(1,:))/(double(resolutionX)-1);
-resolutionY = round(diff(domain(2,:))/gridSpace) + 1;
-resolution = [resolutionX,resolutionY];
-
-strainlineLocalMaxDistance = 4*gridSpace;
-
-%% Cauchy-Green strain eigenvalues and eigenvectors
-[cgEigenvector,cgEigenvalue] = eig_cgStrain(lDerivative,domain,resolution,timespan,'incompressible',incompressible,'odeSolverOptions',cgStrainOdeSolverOptions);
-
-% Plot finite-time Lyapunov exponent
-cgEigenvalue2 = reshape(cgEigenvalue(:,2),fliplr(resolution));
-ftle_ = ftle(cgEigenvalue2,diff(timespan));
-hAxes = setup_figure(domain);
[…]

### 2014-10-03 graphics/bickley_jet/stretch_lcs.m

@@ -1,69 +1,0 @@
-function stretch_lcs
-
-%% Input parameters
-u = 62.66;
-lengthX = pi*earthRadius;
-lengthY = 1.77e6;
-epsilon = [.075,.4,.3];
-domain = [2e6,.55*lengthX;[-1,.25]*2.25*lengthY];
-resolutionX = 500;
-timespan = [0,2*lengthX/u];
-
-%% Velocity definition
-perturbationCase = 3;
-phiTimespan = [0,25];
-phiInitial = [0,0];
-phiSol = ode45(@d_phi,phiTimespan,phiInitial);
-timeResolution = 1e5;
-phi1 = deval(phiSol,linspace(phiTimespan(1),phiTimespan(2),timeResolution),1);
-phi1Max = max(phi1);
-lDerivative = @(t,x,~)derivative(t,x,false,u,lengthX,lengthY,epsilon,perturbationCase,phiSol,phi1Max);
-incompressible = true;
-
-%% LCS parameters
-% Cauchy-Green strain
-cgStrainOdeSolverOptions = odeset('relTol',1e-4);
-
-% Stretchlines
-stretchlineMaxLength = 1e8;
-stretchlineOdeSolverOptions = odeset('relTol',1e-4);
-stretchlineColor = 'b';
-lcsInitialPositionMarkerSize = 2;
-
-% Make x and y grid spacing as equal as possible
-gridSpace = diff(domain(1,:))/(double(resolutionX)-1);
-resolutionY = round(diff(domain(2,:))/gridSpace) + 1;
-resolution = [resolutionX,resolutionY];
-
-stretchlineLocalMaxDistance = 8*gridSpace;
-
-%% Cauchy-Green strain eigenvalues and eigenvectors
-s = warning('off','eig_cgStrain:unequalDelta');
-warning('off','eig_cgStrain:unequalAuxGridDelta')
-[cgEigenvector,cgEigenvalue] = eig_cgStrain(lDerivative,domain,resolution,timespan,'incompressible',incompressible,'odeSolverOptions',cgStrainOdeSolverOptions);
-warning(s)
-
-% Plot finite-time Lyapunov exponent
-cgEigenvalue2 = reshape(cgEigenvalue(:,2),fliplr(resolution));
-ftle_ = ftle(cgEigenvalue2,diff(timespan));
-hAxes = setup_figure(domain);
-plot_ftle(hAxes,domain,resolution,ftle_);
-colormap(hAxes,flipud(gray))
-
[…]

### 2014-10-03 graphics/double_gyre/hyperbolic_shear_lcs_details_highres.m

@@ -1,215 +1,0 @@
-%% Input parameters
-epsilon = .1;
-amplitude = .1;
-omega = pi/5;
-domain = [0,2;0,1];
-resolution = [1000,500];
-timespan = [0,10];
-
-%% Velocity definition
-lDerivative = @(t,x,~)derivative(t,x,false,epsilon,amplitude,omega);
-incompressible = true;
-
-%% LCS parameters
-cgStrainOdeSolverOptions = odeset('relTol',1e-5);
-
-% Lambda-lines
-poincareSection = struct('endPosition',{},'numPoints',{},'orbitMaxLength',{});
-poincareSection(1).endPosition = [0.55,0.55;0.2,0.2];
-poincareSection(2).endPosition = [1.53,.45;1.95,0.1];
-[poincareSection.numPoints] = deal(100);
-nPoincareSection = numel(poincareSection);
-for i = 1:nPoincareSection
-    rOrbit = hypot(diff(poincareSection(i).endPosition(:,1)),diff(poincareSection(i).endPosition(:,2)));
-    poincareSection(i).orbitMaxLength = 2*(2*pi*rOrbit);
-end
-lambdaLineOdeSolverOptions = odeset('relTol',1e-6);
-lambdaStep = 0.01;
-lambdaRange = 0.95:lambdaStep:1.05;
-% set flag to true to show Poincare return maps in closed orbit detection
-showGraph = false;
-
-% Strainlines
-strainlineMaxLength = 20;
-gridSpace = diff(domain(1,:))/(double(resolution(1))-1);
-strainlineLocalMaxDistance = 2*gridSpace;
-strainlineOdeSolverOptions = odeset('relTol',1e-6);
-
-% Stretchlines
-stretchlineMaxLength = 20;
-stretchlineLocalMaxDistance = 10*gridSpace;
-stretchlineOdeSolverOptions = odeset('relTol',1e-6);
-
-% Graphics properties
-strainlineColor = 'r';
-stretchlineColor = 'b';
-lambdaLineColor = [0,.6,0];
-lcsInitialPositionMarkerSize = 2;
-
-%% Cauchy-Green strain eigenvalues and eigenvectors
-[cgEigenvector,cgEigenvalue] = eig_cgStrain(lDerivative,domain,resolution,timespan,'incompressible',incompressible,'odeSolverOptions',cgStrainOdeSolverOptions);
-
[…]

### 2014-10-03 graphics/double_gyre/poincare_return_map.m

@@ -1,45 +1,0 @@
-function poincare_return_map
-
-%% Input parameters
-epsilon = .1;
-amplitude = .1;
-omega = pi/5;
-domain = [0,2;0,1];
-resolution = [750,375];
-timespan = [0,5];
-
-%% Velocity definition
-lDerivative = @(t,x,~)derivative(t,x,false,epsilon,amplitude,omega);
-incompressible = true;
-
-%% LCS parameters
-cgStrainOdeSolverOptions = odeset('relTol',1e-5);
-
-% Lambda-lines
-poincareSection.endPosition = [.55,.55;.2,.5];
-poincareSection.numPoints = 100;
-rOrbit = hypot(diff(poincareSection.endPosition(:,1)),diff(poincareSection.endPosition(:,2)));
-poincareSection.integrationLength = [0,2*(2*pi*rOrbit)];
-
-lambda = 1;
-lambdaLineOdeSolverOptions = odeset('relTol',1e-6);
-
-%% Cauchy-Green strain eigenvalues and eigenvectors
-[cgEigenvector,cgEigenvalue] = eig_cgStrain(lDerivative,domain,resolution,timespan,'incompressible',incompressible,'odeSolverOptions',cgStrainOdeSolverOptions);
-
-etaPos = lambda_line(cgEigenvector,cgEigenvalue,lambda);
-s = warning('off','integrate_line:isDiscontinuousLargeAngle');
-[~,~,hPoincareMap] = poincare_closed_orbit(domain,resolution,etaPos,poincareSection,'odeSolverOptions',lambdaLineOdeSolverOptions,'showGraph',true);
-warning(s)
-
-delete(findobj(hPoincareMap,'type','axes','Tag','legend'))
-hAxes = findobj(hPoincareMap,'type','axes');
-delete(get(hAxes,'title'))
-set(hAxes,'ylim',[-1,1]*1e-4)
-% Change colors to black
-set(findobj(hAxes,'Marker','none'),'color',[0,0,0])
-set(findobj(hAxes,'Marker','o'),'MarkerEdgeColor',[0,0,0])
-set(findobj(hAxes,'Marker','o','MarkerFaceColor',[1,0,0]),'MarkerFaceColor',[0,0,0])
-set(findobj(hAxes,'Marker','o','MarkerFaceColor',[0,0,1]),'MarkerFaceColor','none')
-
-print_pdf(hPoincareMap,'poincare_return_map')

### 2014-10-03 graphics/double_gyre/startup.m

@@ -1,5 +1,0 @@
-graphics_root = fullfile('..');
-lcs_tool_root = fullfile('..','..','lcs_tool');
-demo = fullfile(lcs_tool_root,'demo','double_gyre');
-
-addpath(graphics_root,lcs_tool_root,demo)

### 2014-10-03 graphics/ocean_dataset/hyperbolic_shear_lcs_details.m

@@ -1,244 +1,0 @@
-%% Input parameters
-domain = [0,6;-34,-28];
-resolution = [400,400];
-timespan = [100,130];
-
-%% Velocity definition
-load('ocean_geostrophic_velocity.mat')
-% Set velocity to zero at boundaries
-vlon(:,[1,end],:) = 0;
-vlon(:,:,[1,end]) = 0;
-vlat(:,[1,end],:) = 0;
-vlat(:,:,[1,end]) = 0;
-interpMethod = 'spline';
-vlon_interpolant = griddedInterpolant({time,lat,lon},vlon,interpMethod);
-vlat_interpolant = griddedInterpolant({time,lat,lon},vlat,interpMethod);
-lDerivative = @(t,x,~)flowdata_derivative(t,x,vlon_interpolant,vlat_interpolant);
-incompressible = true;
-
-%% LCS parameters
-% Cauchy-Green strain
-cgEigenvalueFromMainGrid = false;
-cgAuxGridRelDelta = 0.01;
-
-% Lambda-lines
-lambdaLineOdeSolverOptions = odeset('relTol',1e-6,'initialStep',1e-2);
-lambdaStep = 0.02;
-lambdaRange = 0.90:lambdaStep:1.10;
-% set flag to true to show Poincare return maps in closed orbit detection
-showGraph = false;
-
-% Strainlines
-strainlineMaxLength = 20;
-gridSpace = diff(domain(1,:))/(double(resolution(1))-1);
-strainlineLocalMaxDistance = 2*gridSpace;
-strainlineOdeSolverOptions = odeset('relTol',1e-6);
-
-% Stretchlines
-stretchlineMaxLength = 20;
-stretchlineLocalMaxDistance = 4*gridSpace;
-stretchlineOdeSolverOptions = odeset('relTol',1e-6);
-
-% Graphics properties
-strainlineColor = 'r';
-stretchlineColor = 'b';
-lambdaLineColor = [0,.6,0];
-lcsInitialPositionMarkerSize = 2;
-
-hAxes = setup_figure(domain);
-title(hAxes,'Strainline and \lambda-line LCSs')
-xlabel(hAxes,'Longitude (\circ)')
-ylabel(hAxes,'Latitude (\circ)')
-
-%% Cauchy-Green strain eigenvalues and eigenvectors
[…]

### 2014-10-06 main.tex
Update code listings to match J. Comp. Sci. submssion
MATLAB code listings updated to match Journal of Computational Science
submission by Florian Huhn on 2014-06-13.

@@
 More recent work has focused on an exact mathematical formulation of the properties defining LCSs\citep{haller11:_lagran_coher_struc,farazmand12:_comput_lagran,haller12:_geodes_theor_trans_barrier_two_dimen_flows,haller13:_coher_lagran,blazevski14:_hyper_ellip_trans_barrier_three}.
-In two-dimensional flows, hyperbolic and parabolic LCSs turn out to be stationary curves of the averaged material shear\citep{farazmand14:_shearless}, whereas elliptic LCSs are stationary curves of the averaged
-strain\citep{haller13:_coher_lagran}. These variational formulations lead to explicit solutions for LCSs as null-geodesics of appropriate Lorentzian metrics.
+In two-dimensional flows, hyperbolic and parabolic LCSs turn out to be stationary curves of the averaged material shear\citep{farazmand14:_shearless}, whereas elliptic LCSs are stationary curves of the averaged strain\citep{haller13:_coher_lagran}.
+These variational formulations lead to explicit solutions for LCSs as null-geodesics of appropriate Lorentzian metrics.
 
@@
 \begin{enumerate}
-\item Define a Cartesian grid for initial conditions of trajectories. Define
-an auxiliary grid for differentiating with respect to initial conditions.
+\item Define a Cartesian grid for initial conditions of trajectories. Define an auxiliary grid for differentiating with respect to initial conditions.
 \item Solve \cref{eq:velocity} starting from each grid point and auxiliary grid point over the time interval $[t_0,t]$. This gives a discrete approximation to the flow map $F_{t_0}^t(x_0)$.
@@
 
-Velocity fields defined by datasets require pre-processing before
-they are used in the numerical integration of \cref{eq:velocity}.
+Velocity fields defined by datasets require pre-processing before they are used in the numerical integration of \cref{eq:velocity}.
 This requires spatial and temporal interpolation that enables the evaluation of the velocity function at arbitrary points in $U$ and at arbitrary times between $t_-$ and $t_+$.
@@
 
+% FIXME Need to update table to account for lambda range
 \begin{table}
@@
 \end{tabular}
-\caption{Syntax of LCS Tool's elliptic LCS functions.}
+\caption{Syntax for LCS Tool elliptic LCS functions.}
 \label{t:Elliptic LCS functions}
@@
 The Poincare section is to be oriented such that the first endpoint is close to the center of the expected Lagrangian vortex, and the second endpoint is outside this vortex.
-There is no automated procedure procedure implemented in LCS Tool for the positioning of Poincare sections (see however \citet{karrasch14:_autom_lagran} for a recently developed algorithm).
+There is no automated procedure implemented in LCS Tool for the positioning of Poincare sections (see however \citet{karrasch14:_autom_lagran} for a recently developed algorithm).
 Additionally, the number of lambda-lines launched from the Poincare section, \lstinline!poincareSection.numPoints!, must be defined.
@@
 Rotations exceeding 90°, between two such neighboring vectors are classified as orientation discontinuities and are corrected before linear interpolation.
-When setting the Cauchy-Green strain tensor main grid resolution, one may find it helpful to calculate a histogram of eigenvector-field rotations to ensure that all rotations are well below 90° or almost 180°.
+When setting the Cauchy-Green strain tensor main grid resolution, one may find it helpful to calculate a histogram of eigenvector field rotations to ensure that all rotations are well below 90° or almost 180°.
 
@@
 \caption{(Left) Elliptic LCS (green) with repelling LCSs (red) continuing into its interior.
-(Right) Final positions $x(t) = (x_1(t),x_2(t)) = F_{t_0}^t(x_0)$ of tracer particles, are plotted with their $x_1(t)$ position encoded in red and $x_2(t)$ position encoded in green.
+(Right) Final positions $x(t) = (x_1(t),x_2(t)) = F_{t_0}^t(x_0)$ of tracer particles.
+$x_1(t)$ position encoded in red and $x_2(t)$ position encoded in green.
 Note the continued impact of the repelling hyperbolic LCSs (black) on tracer patterns inside the elliptic LCS (white).}
@@
 The flow timespan is $t \in [0,10]$ and the domain is $x \in [0,2]$, $y \in [0,1]$.
-By examining the FTLE field (which
-is optionally provided by LCS Tool), we position Poincare sections to capture elliptic LCSs.
-An LCS Tool demo script to perform this operation is given in \cref{l:double gyre lambda-line LCS} where two Poincare sections are defined.
-The free stretching parameter $\lambda$ (\cref{eq:etafields}) is varied over the range $[0.93,1.07]$ with increments of $0.01$.
-In a loop , closed orbits for all predefined $\lambda$ values are computed and the outermost closed orbit is kept as the Lagrangian vortex boundary.
-In this script, we have chosen small values for the error tolerances of the integration of the Cauchy-Green strain tensor and $\lambda$-lines to facilitate convergence.
+By examining the FTLE field (which LCS Tool can provide), we position Poincare sections to capture elliptic LCSs.
+An LCS Tool demo script to perform this operation is given in \cref{l:double gyre lambda-line LCS} where two Poincare sections are defined (\cref{ll:double gyre poincareSection(1),ll:double gyre poincareSection(2)}).
+The free stretching parameter $\lambda$ (\cref{eq:etafields}) is varied over the range $[0.93,1.07]$ with increments of $0.01$ (\cref{ll:double gyre lambdaRange}).
+% FIXME The cited loop contains ellipses (...), that hide important code and should become an LCS Tool function
+In a loop (\crefrange{ll:double gyre lambda loop start}{ll:double gyre lambda loop end}), closed orbits for all predefined $\lambda$ values are computed and the outermost closed orbit is kept as the Lagrangian vortex boundary.
+% FIXME Small tolerances (i.e. strict tolerances) should make convergence difficult not easy, so the word "facilitate" in the sentence below seems illogical.
+In this script, we have chosen small values for the error tolerances of the integration of the Cauchy-Green strain tensor (\cref{ll:double gyre cgStrainOdeSolverOptions}) and $\lambda$-lines (\cref{ll:double gyre lambdaLineOdeSolverOptions})to facilitate convergence.
 
-\lstinputlisting[caption={Double gyre script for elliptic lambda-line LCSs},label=l:double gyre lambda-line LCS]{listings/double_gyre/lambda_lcs.m}
+\lstinputlisting[caption={LCS Tool commands for double gyre elliptic LCSs. Subset of the LCS Tool script \lstinline!demo/double_gyre/hyperbolic_shear_lcs.m!},
+                 label=l:double gyre lambda-line LCS]{listings/double_gyre/elliptic_lcs_subset.m}
 
 In \cref{f:double gyre lambda LCS convergence}, the resolution of the  Cauchy-Green strain tensor is varied from $500 \times 250$ to $1000 \times 500$.
-The location of the outermost closed lambda-line changes little, demonstrating convergence.
-For all tested resolutions the $\lambda$ values of the outermost closed orbits are constant.
-This is also suggests that the lowest tested resolution of $500 \times 250$ is sufficient to identify the elliptic LCSs in this example.
+The location of the outermost closed $\lambda$-line changes insignificantly, demonstrating convergence.
+For all tested resolutions, the $\lambda$ values of the outermost closed orbits match.
+% FIXME Should reduce lowest resolution used to demonstrate problems if selected resolution is too low
+This also suggests that the lowest tested resolution, $500 \times 250$, is sufficient to identify elliptic LCSs.
 
@@
 \Cref{f:double gyre lambda hyperbolic LCS} shows the complete picture of elliptic and hyperbolic LCSs of the double gyre at this resolution.
-\Cref{lst:double_gyre_lambda_strain_stretch_lcs} lists the corresponding LCS Tool commands.
+\Cref{l:double gyre lambda strain stretch LCS} lists the corresponding LCS Tool commands.
 The maximal length of shrinklines and stretchlines, \lstinline!strainlineMaxLength!
@@
 As usual, the local maximization distance is set larger for stretchlines than for shrinklines (cf.
-\cref{ll:strainlineLocalMaxDistance} and \cref{ll:stretchlineLocalMaxDistance}).
+\cref{ll:shrinklineLocalMaxDistance} and \cref{ll:stretchlineLocalMaxDistance}).
 The purpose of the maximization distance is to obtain spatially separated LCSs and to avoid a dense tangle of lines that basically indicates the same hyperbolic LCS (cf.
@@
 \lstinputlisting[
-caption={LCS Tool commands for double gyre hyperbolic LCSs. Condensed version of the LCS Tool script \lstinline!demo/double_gyre/hyperbolic_shear_lcs.m!},
-label=lst:double_gyre_lambda_strain_stretch_lcs
-]
-{listings/double_gyre/lambda_strain_stretch_lcs.m}
+        caption={LCS Tool commands for double gyre hyperbolic LCSs. Subset of the LCS Tool script \lstinline!demo/double_gyre/hyperbolic_shear_lcs.m!},
+        label=l:double gyre lambda strain stretch LCS]
+        {listings/double_gyre/hyperbolic_lcs_subset.m}
 
@@
 \includegraphics[width=\textwidth]{graphics/double_gyre/lambda_hyperbolic_lcs}
-\caption{LCSs in the double gyre flow.
-Resolution $[500\times250]$.
-$\lambda = 1.00$ and $1.04$ for the left and the right gyre.
-$\lambda \in [0.93,1.07],\,\Delta\lambda=0.01$.
-(Top) Elliptic (green) and repelling (red) LCSs with the FTLE shown in the background.
-(Bottom) Elliptic and attracting (blue) LCSs.}
+\caption{LCSs in the double gyre flow. Resolution $[500\times250]$. $\lambda = 1.00$ and $1.04$ for the left and the right gyre. $\lambda \in [0.93,1.07]$, $\Delta\lambda = 0.01$.
+(Top) Elliptic (green) and repelling (red) LCSs with the FTLE shown in the background. (Bottom) Elliptic and attracting (blue) LCSs.}
 \label{f:double gyre lambda hyperbolic LCS}
@@
 The integration time is $T=4L_x/U$, a multiple of a the eddy turnover time in the flow.
-\Cref{f:Bickley jet LCS} shows lambda-line LCS of the Bickley jet with the FTLE in the background.
-The figure demonstrates a poor result since the LCSs change shape significantly between eddies.
-To address this problem we could adjust numerical parameters such as the resolution, but instead we will analyze lambda-line closed orbits of one vortex for a short timespan.
+\Cref{l:Bickley jet elliptic hyperbolic LCS subset} shows the LCS Tool commands in which the chaotically perturbed velocity is defined (\crefrange{ll:Bickley jet velocity definition start}{ll:Bickley jet velocity definition end}), periodic boundary conditions are imposed in the x-direction (\cref{ll:Bickley jet periodicBc}), and five Poincare sections are defined where we expect coherent vortices (\crefrange{ll:Bickley jet Poincare start}{ll:Bickley jet Poincare end}).
+$\lambda$-values for closed orbit detection are varied over the range $[0.80,1.20]$ with a step of $0.01$.
 
-\Cref{f:Bickley jet LCS} shows elliptic and hyperbolic LCSs of the Bickley jet with the FTLE field in the background.
+\Cref{f:Bickley jet LCS} shows elliptic and hyperbolic LCSs of the Bickley jet with the FTLE in the background.
+
+\lstinputlisting[caption={LCS Tool commands for Bickley jet elliptic and hyperbolic LCSs. Subset of the LCS Tool script \lstinline!demo/bickley_jet/hyperbolic_shear_lcs.m!},
+                 label=l:Bickley jet elliptic hyperbolic LCS subset]{listings/bickley_jet/elliptic_hyperbolic_lcs_subset.m}
 
@@
 \caption{Elliptic and hyperbolic (top: red - repelling, bottom: blue - attracting) LCSs in the Bickley jet.
-$\lambda$-valuestrains of coherent eddies from left to right [0.95,\,0.80,\,0.94,\,0.80,\,0.94], $\lambda \in [0.80,1.20],\,\Delta\lambda=0.01$.}
+$\lambda$ values of elliptic LCSs from left to right are [0.95,\,0.80,\,0.94,\,0.80,\,0.94], $\lambda \in [0.80,1.20]$, $\Delta\lambda = 0.01$.}
 \label{f:Bickley jet LCS}
@@
 
-\lstinputlisting[caption={Bickley jet script for hyperbolic and elliptic LCSs.
-Subset of \lstinline!demo/bickley_jet/hyperbolic_shear_lcs.m!},label=lst:Bickley jet hyperbolic_shear_lcs subset]{listings/bickley_jet/hyperbolic_shear_lcs_subset.m}
-
-\subsection{Ocean dataset}
+\subsection{Ocean velocity data from satellite altimetry}
 \label{sec:oceandataset}
@@
 
-The data is given at a spatial resolution of $1/4\degree$ and a temporal resolution of $7\,\mathrm{days}$.
+The data is given at a spatial resolution of $1/4\degree$ and a temporal resolution of 7 days.
 Due to the discrete data, defining the right hand side of \cref{eq:dtlon} and \cref{eq:dtlat} involves spline interpolation in space and time.
-An interpolant is generated first, then the function \lstinline!flowdata_derivative!
-evaluates the interpolants for the zonal and meridional velocity at the needed coordinates.
-\Cref{l:griddedInterp} shows the relevant part of the code in LCS Tool's ocean demo file \lstinline!hyperbolic_shear_lcs.m!.
-In \crefrange{ll:interpolation start}{ll:interpolation end}, the commands for the interpolation are given of the velocity dataset are given.
+An interpolant is generated first, then the function \lstinline!flowdata_derivative! evaluates the interpolants for the zonal and meridional velocity at the needed coordinates.
+\Cref{l:ocean data elliptic hyperbolic lcs} shows the relevant part of the code in LCS Tool's ocean demo file \lstinline!demo/ocean_dataset/hyperbolic_shear_lcs.m!.
+The commands for the interpolation of the velocity data are given in \crefrange{ll:interpolation start}{ll:interpolation end}.
 
-\begin{lstlisting}[caption={LCS ocean demo: Parts of the demo script included in LCS Tool to compute LCS from an ocean dataset},label=l:griddedInterp]
-% Input parameters
-domain = [0,6;-34,-28];
-resolution = [400,400]; %@\label{ll:resolution}
-timespan = [100,130]; %@\label{ll:timespan}
+\lstinputlisting[
+        caption={LCS Tool commands for ocean data elliptic and hyperbolic LCSs. Subset of the LCS Tool script \lstinline!demo/ocean_dataset/hyperbolic_shear_lcs.m!},
+        label={l:ocean data elliptic hyperbolic lcs}]
+        {listings/ocean_dataset/elliptic_hyperbolic_lcs_subset.m}
 
-% Velocity definition
-...
-interpMethod = 'spline'; %@\label{ll:interpolation start}
-vlon_interpolant = griddedInterpolant({time,lat,lon},vlon,interpMethod);
-vlat_interpolant = griddedInterpolant({time,lat,lon},vlat,interpMethod);
-lDerivative = @(t,x,~)flowdata_derivative(t,x,vlon_interpolant,vlat_interpolant); %@\label{ll:interpolation end}
-incompressible = true; %@\label{ll:incompressible}
-
-% LCS parameters
-% Cauchy-Green strain
-cgEigenvalueFromMainGrid = false; %@\label{ll:cgEigenvalueFromMainGrid}
-cgAuxGridRelDelta = 0.01; %@\label{ll:cgAuxGridRelDelta}
-
-% Lambda-lines
-lambda = 1;
-lambdaLineOdeSolverOptions = odeset('relTol',1e-6);
-
-% Strainlines
-strainlineMaxLength = 20;
-gridSpace = diff(domain(1,:))/(double(resolution(1))-1);
-strainlineLocalMaxDistance = 2*gridSpace;
-strainlineOdeSolverOptions = odeset('relTol',1e-4);
-
-% Stretchlines
-stretchlineMaxLength = 20;
-stretchlineLocalMaxDistance = 4*gridSpace;
-stretchlineOdeSolverOptions = odeset('relTol',1e-4);
-...
-% Cauchy-Green strain eigenvalues and eigenvectors
-[cgEigenvector,cgEigenvalue] = eig_cgStrain(lDerivative,domain,resolution,timespan,'incompressible',incompressible,'eigenvalueFromMainGrid',cgEigenvalueFromMainGrid,'auxGridRelDelta',cgAuxGridRelDelta); %@\label{ll:eig_cgStrain}
-
-% Lambda-line LCSs
-% Define Poincare sections; ...
-poincareSection = struct('endPosition',{},'numPoints',{},'orbitMaxLength',{});
-...
-poincareSection(1).endPosition = [3.3,-32.1;3.7,-31.6]; %@\label{ll:poincareSection(1)}
-poincareSection(2).endPosition = [1.3,-30.9;1.9,-31.1]; %@\label{ll:poincareSection(2)}
-
-% Number of orbit seed points along each Poincare section
-[poincareSection.numPoints] = deal(100);
-
-% Set maximum orbit length to twice the expected circumference
-nPoincareSection = numel(poincareSection);
-for i = 1:nPoincareSection
-    rOrbit = hypot(diff(poincareSection(i).endPosition(:,1)),diff(poincareSection(i).endPosition(:,2)));
-    poincareSection(i).orbitMaxLength = 2*(2*pi*rOrbit);
-end
-[shearline.etaPos,shearline.etaNeg] = lambda_line(cgEigenvector,cgEigenvalue,lambda); %@\label{ll:lambda_line}
-closedLambdaLine = poincare_closed_orbit_multi(domain,resolution,shearline,poincareSection,'odeSolverOptions',lambdaLineOdeSolverOptions); %@\label{ll:poincare_closed_orbit_multi}
-...
-% Hyperbolic strainline LCSs
-strainlineLcs = seed_curves_from_lambda_max(strainlineLocalMaxDistance,strainlineMaxLength,cgEigenvalue(:,2),cgEigenvector(:,1:2),domain,resolution,'odeSolverOptions',strainlineOdeSolverOptions); %@\label{ll:strainlineLcs}
-...
-stretchlineLcs = seed_curves_from_lambda_max(stretchlineLocalMaxDistance,stretchlineMaxLength,-cgEigenvalue(:,1),cgEigenvector(:,3:4),domain,resolution,'odeSolverOptions',stretchlineOdeSolverOptions); %@\label{ll:stretchlineLcs}
-\end{lstlisting}
-
-We choose the integration time as $T=30$ days (\cref{l:griddedInterp}, \cref{ll:timespan}), which is larger than the eddy turnover time in this region.
-The resolution of the main computational grid for initial conditions is set to $400 \times 400$ (\cref{ll:resolution}).
+We choose the integration time as $T=30$ days (\cref{l:ocean data elliptic hyperbolic lcs}, \cref{ll:ocean data timespan}), which is larger than the eddy turnover time in this region.
+The resolution of the main computational grid for initial conditions is set to $400 \times 400$ (\cref{ll:ocean data resolution}).
 This corresponds to a resolution of roughly 0.015° and gives good results.
@@
 Incompressibility of the flow is enforced (\cref{ll:incompressible}), the auxiliary grid distance is set to 1\% of the main grid distance (\cref{ll:cgAuxGridRelDelta}), and eigenvalues are computed from the auxiliary grid (\cref{ll:cgEigenvalueFromMainGrid}).
-Elliptic LCSs are computed in \cref{ll:poincare_closed_orbit_multi}, after the Poincare sections have been set (\cref{ll:poincareSection(1),ll:poincareSection(2)}) and the $\eta_\pm^\lambda$ fields have been defined (\cref{ll:lambda_line}).
-Hyperbolic LCSs are computed in \cref{ll:strainlineLcs,ll:stretchlineLcs}.
+Elliptic LCSs are computed in \cref{ll:ocean data poincare_closed_orbit_multi}, after the Poincare sections have been set (\cref{ll:ocean data poincareSection(1),ll:ocean data poincareSection(2)}) and the $\eta_\pm^\lambda$ fields have been defined (\cref{ll:ocean data lambda_line}).
+$\lambda$-values are varied over a range of $[0.90,1.10]$ with a step of $0.02$ (\cref{ll:ocean data lambdaRange}).
+Hyperbolic LCSs are computed in \cref{ll:ocean data shrinklineLcs,ll:ocean data stretchlineLcs}.
 

### 2014-10-06 listings/bickley_jet/elliptic_hyperbolic_lcs_subset.m

@@ -1,0 +1,73 @@
+%% Input parameters
+u = 62.66;
+lengthX = pi*earthRadius;
+lengthY = 1.77e6;
+epsilon = [.075,.4,.3];
+domain = [0,lengthX;[-1,1]*2.25*lengthY];
+resolutionX = 500;
+timespan = [0,4*lengthX/u];
+
+%% Velocity definition
+perturbationCase = 3; %@\label{ll:Bickley jet velocity definition start}
+phiTimespan = [0,25];
+phiInitial = [0,0];
+phiSol = ode45(@d_phi,phiTimespan,phiInitial);
+timeResolution = 1e5;
+phi1 = deval(phiSol,linspace(phiTimespan(1),phiTimespan(2),timeResolution),1);
+phi1Max = max(phi1);
+lDerivative = @(t,x,~)derivative(t,x,false,u,lengthX,lengthY,epsilon,perturbationCase,phiSol,phi1Max); %@\label{ll:Bickley jet velocity definition end}
+incompressible = true;
+periodicBc = [true,false]; %@\label{ll:Bickley jet periodicBc}
+
+%% Cauchy-Green strain eigenvalues and eigenvectors
+[cgEigenvector,cgEigenvalue] = eig_cgStrain(lDerivative,domain,resolution,timespan,'incompressible',incompressible);
+
+%% LCS parameters
+% Lambda-lines
+lambdaRange = .8:.01:1.1;
+lambdaLineOdeSolverOptions = odeset('relTol',1e-6);
+
+% Shrinklines
+shrinklineMaxLength = 1e8;
+shrinklineLocalMaxDistance = 8*gridSpace;
+shrinklineOdeSolverOptions = odeset('relTol',1e-4);
+
+% Stretchlines
+stretchlineMaxLength = 1e8;
+stretchlineLocalMaxDistance = 4*gridSpace;
+stretchlineOdeSolverOptions = odeset('relTol',1e-4);
+
+%% Lambda-line LCSs
+% Define Poincare sections; first point in center of elliptic region and
+% second point outside elliptic region
+poincareSection = struct('endPosition',{},'numPoints',{},'orbitMaxLength',{});
+
+poincareSection(1).endPosition = [3.25e6,1.5e6;1.4e6,2.6e6]; %@\label{ll:Bickley jet Poincare start}
+poincareSection(2).endPosition = [6.5e6,-1.4e6; 5.e6,-3.e6];
+poincareSection(3).endPosition = [1e7,1.5e6;8e6,2.6e6];
[…]

### 2014-10-06 listings/double_gyre/elliptic_lcs_subset.m

@@ -1,0 +1,38 @@
+%% Input parameters
+epsilon = .1;
+amplitude = .1;
+omega = pi/5;
+domain = [0,2;0,1];
+resolution = [500,250];
+timespan = [0,10];
+
+%% Velocity definition
+lDerivative = @(t,x,~)derivative(t,x,false,epsilon,amplitude,omega);
+incompressible = true;
+
+%% LCS parameters
+cgStrainOdeSolverOptions = odeset('relTol',1e-5); %@\label{ll:double gyre cgStrainOdeSolverOptions}
+
+% Lambda-lines
+poincareSection = struct('endPosition',{},'numPoints',{},'orbitMaxLength',{});
+poincareSection(1).endPosition = [.55,.55;.1,.1]; %@\label{ll:double gyre poincareSection(1)}
+poincareSection(2).endPosition = [1.53,.45;1.95,.05]; %@\label{ll:double gyre poincareSection(2)}
+[poincareSection.numPoints] = deal(100);
+nPoincareSection = numel(poincareSection);
+for i = 1:nPoincareSection
+    rOrbit = hypot(diff(poincareSection(i).endPosition(:,1)),diff(poincareSection(i).endPosition(:,2)));
+    poincareSection(i).orbitMaxLength = 2*(2*pi*rOrbit);
+end
+lambdaLineOdeSolverOptions = odeset('relTol',1e-6); %@\label{ll:double gyre lambdaLineOdeSolverOptions}
+lambdaRange = .93:.01:1.07; %@\label{ll:double gyre lambdaRange}
+
+%% Cauchy-Green strain eigenvalues and eigenvectors
+[cgEigenvector,cgEigenvalue] = eig_cgStrain(lDerivative,domain,resolution,timespan,'incompressible',incompressible,'odeSolverOptions',cgStrainOdeSolverOptions);
+
+%% Lambda-line LCSs
+
+for lambda = lambdaRange %@\label{ll:double gyre lambda loop start}
+...        
+    closedLambdaLineCandidate = poincare_closed_orbit_multi(domain,resolution,shearline,poincareSection,'odeSolverOptions',lambdaLineOdeSolverOptions,'showGraph',showGraph);
+...
+end %@\label{ll:double gyre lambda loop end}

### 2014-10-06 listings/double_gyre/hyperbolic_lcs_subset.m

@@ -1,0 +1,17 @@
+% Shrinklines
+shrinklineMaxLength = 20;
+gridSpace = diff(domain(1,:))/(double(resolution(1))-1);
+shrinklineLocalMaxDistance = 2*gridSpace; %@\label{ll:shrinklineLocalMaxDistance}
+strainlineOdeSolverOptions = odeset('relTol',1e-6);
+
+% Stretchlines
+stretchlineMaxLength = 20;
+stretchlineLocalMaxDistance = 10*gridSpace; %@\label{ll:stretchlineLocalMaxDistance}
+stretchlineOdeSolverOptions = odeset('relTol',1e-6);
+
+%% Hyperbolic shrinkline LCSs
+[shrinklineLcs,shrinklineLcsInitialPosition] = seed_curves_from_lambda_max(shrinklineLocalMaxDistance,shrinklineMaxLength,cgEigenvalue(:,2),cgEigenvector(:,1:2),domain,resolution,'odeSolverOptions',shrinklineOdeSolverOptions);
+...
+%% Hyperbolic stretchline LCSs
+[stretchlineLcs,stretchlineLcsInitialPosition] = seed_curves_from_lambda_max(stretchlineLocalMaxDistance,stretchlineMaxLength,-cgEigenvalue(:,1),cgEigenvector(:,3:4),domain,resolution,'odeSolverOptions',stretchlineOdeSolverOptions);
+...

### 2014-10-06 listings/double_gyre/lambda_strain_stretch_lcs.m

@@ -1,12 +1,0 @@
-% Strainlines LCSs
-strainlineMaxLength = 20;
-gridSpace = diff(domain(1,:))/(double(resolution(1))-1);
-strainlineLocalMaxDistance = 2*gridSpace; %@\label{ll:strainlineLocalMaxDistance}
-
-strainlineLcs = seed_curves_from_lambda_max(strainlineLocalMaxDistance,strainlineMaxLength,cgEigenvalue(:,2),cgEigenvector(:,1:2),domain,resolution);
-
-% Stretchlines LCSs
-stretchlineMaxLength = 20;
-stretchlineLocalMaxDistance = 10*gridSpace;  %@\label{ll:stretchlineLocalMaxDistance}
-
-stretchlineLcs = seed_curves_from_lambda_max(stretchlineLocalMaxDistance,stretchlineMaxLength,-cgEigenvalue(:,1),cgEigenvector(:,3:4),domain,resolution);

### 2014-10-06 listings/ocean_dataset/elliptic_hyperbolic_lcs_subset.m

@@ -1,0 +1,66 @@
+% Input parameters
+domain = [0,6;-34,-28];
+resolution = [400,400]; %@\label{ll:ocean data resolution}
+timespan = [100,130]; %@\label{ll:ocean data timespan}
+
+% Velocity definition
+load('ocean_geostrophic_velocity.mat');
+...
+interpMethod = 'spline'; %@\label{ll:interpolation start}
+vlon_interpolant = griddedInterpolant({time,lat,lon},vlon,interpMethod);
+vlat_interpolant = griddedInterpolant({time,lat,lon},vlat,interpMethod);
+lDerivative = @(t,x,~)flowdata_derivative(t,x,vlon_interpolant,vlat_interpolant); %@\label{ll:interpolation end}
+incompressible = true; %@\label{ll:incompressible}
+
+% LCS parameters
+% Cauchy-Green strain
+cgEigenvalueFromMainGrid = false; %@\label{ll:cgEigenvalueFromMainGrid}
+cgAuxGridRelDelta = .01; %@\label{ll:cgAuxGridRelDelta}
+
+% Lambda-lines
+lambdaRange = .9:.02:1.1; %@\label{ll:ocean data lambdaRange}
+lambdaLineOdeSolverOptions = odeset('relTol',1e-6);
+
+% Shrinklines
+shrinklineMaxLength = 20;
+gridSpace = diff(domain(1,:))/(double(resolution(1))-1);
+shrinklineLocalMaxDistance = 2*gridSpace;
+shrinklineOdeSolverOptions = odeset('relTol',1e-4);
+
+% Stretchlines
+stretchlineMaxLength = 20;
+stretchlineLocalMaxDistance = 4*gridSpace;
+stretchlineOdeSolverOptions = odeset('relTol',1e-4);
+...
+% Cauchy-Green strain eigenvalues and eigenvectors
+[cgEigenvector,cgEigenvalue] = eig_cgStrain(lDerivative,domain,resolution,timespan,'incompressible',incompressible,'eigenvalueFromMainGrid',cgEigenvalueFromMainGrid,'auxGridRelDelta',cgAuxGridRelDelta); %@\label{ll:eig_cgStrain}
+
+% Lambda-line LCSs
+% Define Poincare sections; ...
+poincareSection = struct('endPosition',{},'numPoints',{},'orbitMaxLength',{});
+...
+poincareSection(1).endPosition = [3.3,-32.1;3.7,-31.6]; %@\label{ll:ocean data poincareSection(1)}
[…]

### 2014-10-06 main.tex
Add FIXME comments from J. Comp. Sci. peer review

@@
 \begin{abstract}
+% FIXME Reviewer #5, comment #1: Abstract and introduction the objective and contribution of the paper are not clear; several points need to be clarified further.
+% FIXME Reviewer #5, comment #2: For readers to quickly catch the contribution in this work, it would be better to highlight major difficulties and challenges, and original achievements to overcome them, in a clearer way in abstract and introduction.
 We give an algorithmic introduction to Lagrangian coherent structures (LCSs) using a newly developed computational engine, LCS Tool.
@@
 
+% FIXME Reviewer #5, comment #5: Please provide information regarding practical applications of this algorithmic software tool.
 Lagrangian Coherent Structures (LCSs) are evolving organizing centers of trajectory patterns in non-autonomous dynamical systems\citep{haller00:_lagran,peacock13:_lagran,haller15:_langr_coher_struc}.
@@
 Here we present a simple algorithmic introduction to geodesic LCS detection without details on the the underlying mathematics.
-% FIXME Replace GitHub link with http://runmycode.org/
+% FIXME JCompSci editor: Replace GitHub link with http://runmycode.org/
+% FIXME Reviewer #3, comment #4: Add documentation within the github repository to show some reference to the APIs and how users may start to run a his/her own application.
+% FIXME Reviewer #4: Please make sure the code is downloadable once the paper is published
 We then review the implementation of this approach in a computational engine called LCS Tool\footnote{LCS Tool is publicly available at: https://github.com/jeixav/LCS-Tool}.
@@
 
+% FIXME Reviewer #3, comment #3: The comparison with other publicly available LCS detection software should include more details.
+It contains a very high-level summary of LCS tool being an "all-purpose numerical engine".
+That part should be revisited and become more concrete.
 The list of publicly available software for LCS detection further
@@
 
+% FIXME Reviewer #3, comment #1: Between the three applications, there should be a description about linking them in terms of generic APIs (application programming interface) or common code style.
+It is good to have code in the paper but it doesn't seem straightforward for a reader to find out how this simplifies the effort from coding from scratch (instead of using the platform).
+That part should be highlighted as a key point for judging the merit to the community.
+So, it would be very useful to pinpoint the generics with a flowchart (e.g.
+it always starts from input parameters, velocity definition and then LCS parameters, and how you finally call only 1-2 functions for the core solving routine), and your Matlab code there can further demonstrate the paradigm.
+% FIXME Reviewer #5, comment #6: The functionality and taxonomy of the proposed LCS tool should be introduced for an overview understanding of the tool.
+
 \subsection{Computing the invariants of the Cauchy-Green strain tensor}
@@
 The optimal resolution may be determined by a successive doubling of the initial resolution until convergence of the extracted LCSs is observed visually.
+% FIXME Reviewer #2, comment #1: "500 grid point" → "500 grid points"
 If the domain of interest comprises only a few expected LCSs, e.g., one vortex, then a resolution of about 500 grid point along the longest axis usually gives good results.
@@
 This calculation typically takes five to ten minutes with default error tolerances.
+% FIXME Reviewer #2, comment #3: Add memory use data based on, for example, computations for \cref{f:double gyre lambda LCS convergence}.
 A potential drawback of vector form integration is that memory requirements may become excessive at high resolutions.
@@
 \end{center}
+% FIXME Reviewer #2, comment #5: Indicate what are "s" and "P(s)"
 \caption{Example of a Poincare map obtained for an elliptic LCS.
@@
 
+% FIXME Reviewer #3, comment #2: "As for a platform, efficiency is also a matter in consideration.
+Performance degradation is not expected as a cost of generalized computational paradigm.
+That part should be at least mentioned along with the applications to make the article technically more sound."
+
 \section{Examples}
@@
 
+% FIXME Reviewer #2, comment #8: Check all code listings for awkward splits across pages and line-splits (e.g. semicolon on own line and array indices)
+% FIXME Reviewer #5, comment #3: Format manuscript carefully with regards to code listings
 \begin{lstlisting}[caption={Double gyre derivative function corresponding to \cref{eq:double gyre derivative equations}.},label=l:double gyre derivative]
@@
 
+% FIXME Reviewer #5, comment #4: There is too much source code in the main content such as cref{l:double gyre lambda strain stretch LCS,l:Bickley jet elliptic hyperbolic LCS subset,l:ocean data elliptic hyperbolic lcs}.
+They could be arranged as an appendix at the end.
 \lstinputlisting[
@@
 \includegraphics[width=\textwidth]{graphics/double_gyre/lambda_hyperbolic_lcs}
+% FIXME Reviewer #2, comment #6: Explain meaning of white dots ($\lambda$ maxima) and, if possible improve their appearance
 \caption{LCSs in the double gyre flow. Resolution $[500\times250]$. $\lambda = 1.00$ and $1.04$ for the left and the right gyre. $\lambda \in [0.93,1.07]$, $\Delta\lambda = 0.01$.
@@
 
+% FIXME Reviewer #2, comment #7: Remove "a" in "multiple of a the eddy turnover", and sentence reads awkwardly
 The integration time is $T=4L_x/U$, a multiple of a the eddy turnover time in the flow.
@@
 
-\lstinputlisting[caption={LCS Tool commands for Bickley jet elliptic and hyperbolic LCSs. Subset of the LCS Tool script \lstinline!demo/bickley_jet/hyperbolic_shear_lcs.m!},
-                 label=l:Bickley jet elliptic hyperbolic LCS subset]{listings/bickley_jet/elliptic_hyperbolic_lcs_subset.m}
+\lstinputlisting[
+        caption={LCS Tool commands for Bickley jet elliptic and hyperbolic LCSs. Subset of the LCS Tool script \lstinline!demo/bickley_jet/hyperbolic_shear_lcs.m!},
+        label=l:Bickley jet elliptic hyperbolic LCS subset]
+        {listings/bickley_jet/elliptic_hyperbolic_lcs_subset.m}
 

### 2014-10-06 main.tex
Address some reviewer comments

@@
 So, it would be very useful to pinpoint the generics with a flowchart (e.g.
-it always starts from input parameters, velocity definition and then LCS parameters, and how you finally call only 1-2 functions for the core solving routine), and your Matlab code there can further demonstrate the paradigm.
+it always starts from input parameters, velocity definition and then LCS parameters, and how you finally call only 1-2 functions for the core solving routine), and your MATLAB code there can further demonstrate the paradigm.
 % FIXME Reviewer #5, comment #6: The functionality and taxonomy of the proposed LCS tool should be introduced for an overview understanding of the tool.
@@
 The optimal resolution may be determined by a successive doubling of the initial resolution until convergence of the extracted LCSs is observed visually.
-% FIXME Reviewer #2, comment #1: "500 grid point" → "500 grid points"
-If the domain of interest comprises only a few expected LCSs, e.g., one vortex, then a resolution of about 500 grid point along the longest axis usually gives good results.
+If the domain of interest comprises only a few expected LCSs, e.g., one vortex, then a resolution of about 500 grid points along the longest axis usually gives good results.
 Otherwise, a higher resolution must be chosen.
@@
 \end{center}
-% FIXME Reviewer #2, comment #5: Indicate what are "s" and "P(s)"
 \caption{Example of a Poincare map obtained for an elliptic LCS.
+The abscissa, $s$, represents distance along Poincare section.
+The ordinate, $p(s) - s$, represents distance of Poincare orbit return point from initial position.
 Circle markers indicate closed orbit positions.
@@
 \centering
+% FIXME Reviewer #2, comment #6: Improve appearance of $lambda$ extrema (white dots)
 \includegraphics[width=\textwidth]{graphics/double_gyre/lambda_hyperbolic_lcs}
-% FIXME Reviewer #2, comment #6: Explain meaning of white dots ($\lambda$ maxima) and, if possible improve their appearance
-\caption{LCSs in the double gyre flow. Resolution $[500\times250]$. $\lambda = 1.00$ and $1.04$ for the left and the right gyre. $\lambda \in [0.93,1.07]$, $\Delta\lambda = 0.01$.
-(Top) Elliptic (green) and repelling (red) LCSs with the FTLE shown in the background. (Bottom) Elliptic and attracting (blue) LCSs.}
+\caption{
+LCSs in the double gyre.
+Resolution is $[500\times250]$.
+Elliptic LCSs are green, shrinkline LCSs are red and stretchline LCSs are blue.
+White dots indicate $\lambda_2$ maxima for shrinklines and $\lambda_1$ minima stretchlines.
+FTLE shown in the background.
+$\lambda = 1.00$ and $1.04$ for the left and the right gyre.
+$\lambda \in [0.93,1.07]$, $\Delta\lambda = 0.01$.
+}
 \label{f:double gyre lambda hyperbolic LCS}
@@
 
-% FIXME Reviewer #2, comment #7: Remove "a" in "multiple of a the eddy turnover", and sentence reads awkwardly
-The integration time is $T=4L_x/U$, a multiple of a the eddy turnover time in the flow.
+% FIXME Reviewer #2, comment #7: sentence reads awkwardly
+The integration time is $T = 4L_x/U$, a multiple of the eddy turnover time.
 \Cref{l:Bickley jet elliptic hyperbolic LCS subset} shows the LCS Tool commands in which the chaotically perturbed velocity is defined (\crefrange{ll:Bickley jet velocity definition start}{ll:Bickley jet velocity definition end}), periodic boundary conditions are imposed in the x-direction (\cref{ll:Bickley jet periodicBc}), and five Poincare sections are defined where we expect coherent vortices (\crefrange{ll:Bickley jet Poincare start}{ll:Bickley jet Poincare end}).
@@
 \centering
+% FIXME Shrinkline LCS figure does not treat periodic boundary condition correctly
 \includegraphics[width=.85\textwidth]{graphics/bickley_jet/lambda_hyperbolic_lcs}
-\caption{Elliptic and hyperbolic (top: red - repelling, bottom: blue - attracting) LCSs in the Bickley jet.
-$\lambda$ values of elliptic LCSs from left to right are [0.95,\,0.80,\,0.94,\,0.80,\,0.94], $\lambda \in [0.80,1.20]$, $\Delta\lambda = 0.01$.}
+\caption{
+LCSs in the Bickley jet.
+Resolution is $[500 \times 276]$.
+Elliptic LCSs are green, shrinkline LCSs are red and stretchline LCSs are blue. 
+White dots indicate $\lambda_2$ maxima for shrinklines and $\lambda_1$ minima stretchlines.
+FTLE shown in the background.
+$\lambda$ values of elliptic LCSs from left to right are [0.95,\,0.80,\,0.94,\,0.80,\,0.94], $\lambda \in [0.80,1.20]$, $\Delta\lambda = 0.01$.
+}
 \label{f:Bickley jet LCS}
@@
 \end{center}
-\caption{(Top) Elliptic LCSs (green, $\lambda=$ 1.00, 1.08, 0.94, 0.90, 1.06) and repelling LCSs (red) in the ocean data set.
-Background gray-scale is the finite time Lyapunov exponent.
-Dashed green lines are Poincare sections.
-White dots indicate $\lambda_{2}$ local maxima used as initial positions for shrinkline integration.
-(Bottom) Elliptic LCSs (green) and attracting LCSs (blue).}
+\caption{
+LCSs in the ocean velocity data from satellite altimetry.
+Resolution is $[500\times250]$.
+Elliptic LCSs are green, shrinkline LCSs are red and stretchline LCSs are blue. 
+White dots indicate $\lambda_2$ maxima for shrinklines and $\lambda_1$ minima stretchlines.
+FTLE shown in the background.
+$\lambda=$ 1.00, 1.08, 0.94, 0.90, 1.06
+}
 \label{f:ocean dataset LCS}
@@
 % End:
-% LocalWords:  stretchline stretchlines incompressible gyre eigenvector eigenvectors incompressibility dataset datasets LCS LCSs Lyapunov Bickley Rossby timespan variational adaptively vortices parametrise advected strainlines minima extremal FTLE MATLAB maxima geostrophic shrinkline shrinklines
+% LocalWords:  stretchline stretchlines incompressible gyre eigenvector eigenvectors incompressibility dataset datasets LCS LCSs Lyapunov Bickley Rossby timespan variational adaptively vortices parametrise advected strainlines minima extremal FTLE MATLAB maxima geostrophic shrinkline shrinklines Poincare altimetry

### 2014-10-06 listings/bickley_jet/elliptic_hyperbolic_lcs_subset.m

@@ -4,7 +4,7 @@ lengthX = pi*earthRadius;
 lengthY = 1.77e6;
 epsilon = [.075,.4,.3];
 domain = [0,lengthX;[-1,1]*2.25*lengthY];
-resolutionX = 500;
+resolution = [500,276];
 timespan = [0,4*lengthX/u];
 
 %% Velocity definition

### 2014-10-07 main.tex
Switch spelling to British English
J. Comp. Sci. Guide for Authors indicates article keywords are to be
in British English. For simplicity, change entire document to British
English.

@@
 We give an algorithmic introduction to Lagrangian coherent structures (LCSs) using a newly developed computational engine, LCS Tool.
-LCSs are most repelling, attracting and shearing material lines that form the centerpieces of observed tracer patterns in two-dimensional unsteady dynamical systems.
+LCSs are most repelling, attracting and shearing material lines that form the centrepieces of observed tracer patterns in two-dimensional unsteady dynamical systems.
 LCS Tool implements the latest geodesic theory of LCSs for two-dimensional flows, uncovering key transport barriers in unsteady flow velocity data as explicit solutions of differential equations.
@@
 % FIXME Reviewer #5, comment #5: Please provide information regarding practical applications of this algorithmic software tool.
-Lagrangian Coherent Structures (LCSs) are evolving organizing centers of trajectory patterns in non-autonomous dynamical systems\citep{haller00:_lagran,peacock13:_lagran,haller15:_langr_coher_struc}.
+Lagrangian Coherent Structures (LCSs) are evolving organizing centres of trajectory patterns in non-autonomous dynamical systems\citep{haller00:_lagran,peacock13:_lagran,haller15:_langr_coher_struc}.
 Applications of LCSs include oceanic and atmospheric flows\citep{beron-vera13:_objec_agulh,koh02:_hyper}, biological transport problems\citep{wilson09:_lagran_reynol,tallapragada11:_lagran,huhn12:_south_indian_ocean_count_madag}, aeronautics\citep{tang10:_accur_lagran_hong_kong_inter_airpor}, celestial mechanics\citep{gawlik09:_lagran}, crowd dynamics\citep{ali07:_lagran_partic_dynam_approac_crowd}, and aperiodically forced mechanical oscillators\citep{hadjighasem13:_detec_kam}.
 
-\citet{haller01:_distin} proposed that ridges of the finite-time Lyapunov exponent (FTLE) are heuristic indicators of hyperbolic (i.e., repelling- and attracting-type) LCSs.
+\citet{haller01:_distin} proposed that ridges of the finite-time Lyapunov exponent (FTLE) are heuristic indicators of hyperbolic (i.e., repelling and attracting type) LCSs.
 A number of examples indeed support this principle\citep{peacock13:_lagran}.
@@
 Any orbit of \cref{eq:etafields} turns out to stretch uniformly under the flow map $F_{t_0}^t$.
-Specifically, any subset of an orbit of \cref{eq:etafields} increases its arclength precisely by a factor of $\lambda$.
+Specifically, any subset of an orbit of \cref{eq:etafields} increases its arc length precisely by a factor of $\lambda$.
 For this reason, we refer to trajectories of \cref{eq:etafields} as $\lambda$-lines.
@@
 To avoid excessive computational times in MATLAB, we vectorize \cref{eq:velocity}, i.e., combine its right hand side evaluated over each initial point into a single system of equations.
-The resulting system is composed of independent blocks of two-dimensional first-order ODEs.
+The resulting system is composed of independent blocks of two-dimensional first-order ordinary differential equations.
 We then use MATLAB's \lstinline!ode45!
@@
 The first step is to define the position of Poincare sections in regions where closed $\lambda$-lines are expected based on a visual analysis of the orbit structure of the $\eta_\pm^\lambda$ vector field.
-The Poincare section is to be oriented such that the first endpoint is close to the center of the expected Lagrangian vortex, and the second endpoint is outside this vortex.
+The Poincare section is to be oriented such that the first endpoint is close to the centre of the expected Lagrangian vortex, and the second endpoint is outside this vortex.
 There is no automated procedure implemented in LCS Tool for the positioning of Poincare sections (see however \citet{karrasch14:_autom_lagran} for a recently developed algorithm).
@@
 We identify orientation discontinuities by checking the inner product of the $\eta_\pm^\lambda$ vectors at adjacent grid points.
-Rotations exceeding 90°, between two such neighboring vectors are classified as orientation discontinuities and are corrected before linear interpolation.
+Rotations exceeding 90°, between two such neighbouring vectors are classified as orientation discontinuities and are corrected before linear interpolation.
 When setting the Cauchy-Green strain tensor main grid resolution, one may find it helpful to calculate a histogram of eigenvector field rotations to ensure that all rotations are well below 90° or almost 180°.
@@
 An example of a Poincare map produced from integration of the $\eta_\pm^\lambda$ field is shown in \cref{f:Poincare return map}.
-Most orbits will return to the Poincare section and their integration will then be stopped using the ODE event detection function of MATLAB.
+Most orbits will return to the Poincare section and their integration will then be stopped using the ordinary differential equation event detection function of MATLAB.
 Some orbits may, however, deviate far from the Poincare section and do not return for any reasonable integration time.
-To control this behavior, we specify a maximum orbit length, \lstinline!poincareSection.orbitMaxLength!.
+To control this behaviour, we specify a maximum orbit length, \lstinline!poincareSection.orbitMaxLength!.
 In practice, viewing the Poincare section as the radius of a circle and setting the maximum $\lambda$-line integration length to twice the circumference gives good results.
@@
 
-In classical, infinite-time dynamical systems, elliptic and hyperbolic invariant manifolds cannot intersect, as the trajectories in their intersection would then have to follow two different asymptotic behaviors simultaneously.
+In classical, infinite-time dynamical systems, elliptic and hyperbolic invariant manifolds cannot intersect, as the trajectories in their intersection would then have to follow two different asymptotic behaviours simultaneously.
 In finite-time dynamical systems, hyperbolic LCSs may well intersect elliptic LCSs, with hyperbolic LCSs continuing to act as cores of mixing patterns observed in the interior of elliptic LCSs.
@@
 \end{equation}
-The MATLAB function describing this velocity field is given in \cref{l:double gyre derivative}, specifying the right hand side of the particle ODE in a way that supports vectorized integration.
+The MATLAB function describing this velocity field is given in \cref{l:double gyre derivative}, specifying the right hand side of the particle ordinary differential equation in a way that supports vectorized integration.
 
@@
 
-\Cref{f:ocean dataset LCS} shows the positions of elliptic and hyperbolic LCSs on 22 November 2006, which is the same time instant analyzed in \citet{haller13:_coher_lagran,beron-vera13:_objec_agulh}.
+\Cref{f:ocean dataset LCS} shows the positions of elliptic and hyperbolic LCSs on 22 November 2006, which is the same time instant analysed in \citet{haller13:_coher_lagran,beron-vera13:_objec_agulh}.
 Here the integration time is chosen as $T = 30$ days, as opposed to 90 days in the cited references, to avoid tangling of hyperbolic LCSs.
@@
 
-\section*{Acknowledgments}
+\section*{Acknowledgements}
 
@@
 
-% FIXME Keywords are required to be in British English (http://www.elsevier.com/journals/journal-of-computational-science/1877-7503/guide-for-authors), so should switch entire document to British English
 % Local Variables:
-% ispell-dictionary: "en_US"
+% ispell-dictionary: "en_GB"
 % End:
-% LocalWords:  stretchline stretchlines incompressible gyre eigenvector eigenvectors incompressibility dataset datasets LCS LCSs Lyapunov Bickley Rossby timespan variational adaptively vortices parametrise advected strainlines minima extremal FTLE MATLAB maxima geostrophic shrinkline shrinklines Poincare altimetry
+% LocalWords:  stretchline stretchlines incompressible gyre eigenvector eigenvectors incompressibility dataset datasets LCS LCSs Lyapunov Bickley Rossby timespan variational adaptively vortices parametrise advected strainlines minima extremal FTLE MATLAB maxima geostrophic shrinkline shrinklines Poincare altimetry zonal

### 2014-10-07 main.tex
Remove \mathrm from formula
Change writing style in a few sentences

@@
 \usepackage{cleveref}
+
+\DeclareMathOperator{\sech}{sech}
 
@@
 
-Here we present a simple algorithmic introduction to geodesic LCS detection without details on the the underlying mathematics.
+Here we present a simple algorithmic introduction to geodesic LCS detection.
 % FIXME JCompSci editor: Replace GitHub link with http://runmycode.org/
@@
 % FIXME Reviewer #4: Please make sure the code is downloadable once the paper is published
-We then review the implementation of this approach in a computational engine called LCS Tool\footnote{LCS Tool is publicly available at: https://github.com/jeixav/LCS-Tool}.
-This engine is a library of MATLAB functions that extract LCSs from two-dimensional unsteady flows. The examples we present here have been programmed into demonstration scripts.
+We then review the implementation of this approach in a computational engine called LCS Tool\footnote{LCS Tool is available at: http://github.com/jeixav/LCS-Tool}.
+This engine is a library of MATLAB functions that extract LCSs from two-dimensional unsteady flows.
+The examples we present have been programmed into demonstration scripts.
 
@@
 That part should be revisited and become more concrete.
-The list of publicly available software for LCS detection further
-includes the following packages:
+Software for LCS detection includes the following packages:
 \begin{itemize}
@@
 
-The common first step in calculating elliptic and hyperbolic LCSs is the computation of the Cauchy-Green strain tensor field $C_{t_0}^t(x_0)$, as defined in \cref{eq:CG}.
+The first step in calculating elliptic and hyperbolic LCSs is the computation of the Cauchy-Green strain tensor field $C_{t_0}^t(x_0)$, as defined in \cref{eq:CG}.
 The function performing this calculation in LCS Tool is \lstinline!eig_cgStrain!.
@@
 \psi(x,y,t) = \psi_0(x,y) + \psi_1(x,y,t),\\
-\psi_0(x,y) = c_3 y - U L_y \tanh\frac{y}{L_y} + \epsilon_3 U L_y \mathrm{sech}^2\frac{y}{L_y} \cos k_3 x,\\
-\psi_1(x,y,t) = U L_y \mathrm{sech}^2\frac{y}{L_y} \mathrm{Re}\left[ \sum_{n=1}^2 \epsilon_n f_n(t) e^{i k_n x}\right].
+\psi_0(x,y) = c_3 y - U L_y \tanh\frac{y}{L_y} + \epsilon_3 U L_y \sech^2\frac{y}{L_y} \cos k_3 x,\\
+\psi_1(x,y,t) = U L_y \sech^2\frac{y}{L_y} \Re\left[\sum_{n=1}^2 \epsilon_n f_n(t) e^{i k_n x}\right].
 \end{gather*}

### 2014-10-07 main.tex
Address some reviewer comments
Add LCS Tool overview table at beginning of numerical methods section

@@
 
-Here we describe a step-by-step numerical implementation of the geodesic detection of elliptic and hyperbolic LCSs through the differential \cref{eq:etafields,eq:strainline,eq:stretchline}.
+This section describes stepwise the numerical implementation of geodesic detection of elliptic and hyperbolic LCSs based on \cref{eq:etafields,eq:strainline,eq:stretchline}.
+\Cref{t:LCS Tool overview} gives an overview of the steps, functions and variable names used in LCS Tool to analyse a flow.
 
-% FIXME Reviewer #3, comment #1: Between the three applications, there should be a description about linking them in terms of generic APIs (application programming interface) or common code style.
-It is good to have code in the paper but it doesn't seem straightforward for a reader to find out how this simplifies the effort from coding from scratch (instead of using the platform).
-That part should be highlighted as a key point for judging the merit to the community.
-So, it would be very useful to pinpoint the generics with a flowchart (e.g.
-it always starts from input parameters, velocity definition and then LCS parameters, and how you finally call only 1-2 functions for the core solving routine), and your MATLAB code there can further demonstrate the paradigm.
-% FIXME Reviewer #5, comment #6: The functionality and taxonomy of the proposed LCS tool should be introduced for an overview understanding of the tool.
+\begin{table}
+\begin{enumerate}
+\item Define a velocity field, \lstinline!dx = derivative(t,x,p)!
+\item Compute Cauchy-Green strain tensor invariants, \lstinline![v,d] = eig_cgStrain(derivative)!
+\item Compute $\lambda$-lines, \lstinline![etaPos,etaNeg] = lambda_line(v,d,lambda)!
+\item Define Poincare sections, \lstinline!ps!
+% FIXME function call below does not reflect LCS Tool usage (e.g. poincare_close_orbit_multi)
+\item Detect elliptic LCSs, \lstinline!ellipticLcs = poincare_closed_orbit(etaPos,etaNeg,ps)!
+\item Define a local maximization/minimization distance for hyperbolic LCSs \lstinline!lmd!
+% FIXME function call below does not reflect LCS Tool usage (e.g. remove_strain_in_shear)
+% FIXME seed_curves_from_lambda_max is a poor name for a function that detects hyperbolic LCSs
+\item Detect hyperbolic LCSs, \lstinline!hyperbolicLcs = seed_curves_from_lambda_max(v,d,ellipticLcs,lmd)!
+\end{enumerate}
+\caption{Overview of sequence of computations to detect LCSs with LCS Tool functions.}
+\label{t:LCS Tool overview}
+\end{table}
 

### 2014-10-21 main.tex
Reformat FIXME
Address reviewer comment

@@
 % \item \emph{FlowVC}\citep{shadden10:_flowvc} is a general purpose LCS platform for two and three dimensional datasets. Parallel calculations are supported with Open\-MP, CUDA and OpenCL.
-% FIXME FlowTK
-% https://github.com/FlowPhysics/FlowTK
-% doi:10.1007%2F978-3-319-04099-8_13
+% FIXME Add FlowTK, https://github.com/FlowPhysics/FlowTK, doi:10.1007%2F978-3-319-04099-8_13
 \end{itemize}
@@
 
-% FIXME Reviewer #3, comment #2: "As for a platform, efficiency is also a matter in consideration.
-Performance degradation is not expected as a cost of generalized computational paradigm.
-That part should be at least mentioned along with the applications to make the article technically more sound."
-
 \section{Examples}
 
-In this section we present examples of the use of LCS Tool to obtain LCSs in three different examples: a double gyre, a jet, and an oceanic geostrophic flow.
-All three examples are available as demonstration files in the demo folder of LCS Tool.
-By executing these scripts, readers can follow LCS Tool computations step by step.
+This section presents the use of LCS Tool in three examples: a double gyre, a jet, and an oceanic geostrophic flow.
+The examples are available as scripts in the demo folder of LCS Tool.
+Executing these scripts can help readers follow LCS Tool computations.
+
+% FIXME Performance also described in Conclusions section
+The scripts call LCS Tool functions.
+These functions are written to achieve balance between ease of use and performance.
+% FIXME List demo script run-times in minutes in a table
+Running the demo scripts takes on the order of minutes.
 

### 2014-10-21 main.tex
Reword conclusions section

@@
 
-LCS Tool leverages the capabilities of MATLAB extensively.
+LCS Tool leverages the capabilities of MATLAB.
 For FTLE based extraction of LCSs, computational performance has received considerable attention \citep{conti12:_gpu_apu_finit_time_lyapun_expon,miron12:_anisot_lagran_coher_struc}.
-We hope to see similar computational advances for the publicly available LCS Tool platform.
-Optimizing computational performance will likely aid the application of variational LCS methods to large-scale real-time forecasting applications, such as the tracking of environmental contaminants\citep{olascoaga12:_forec}.
+We think LCS Tool can facilitate similar computational advances for variational LCS methods.
+Optimizing computational performance will aid applications to large-scale forecasting applications, such as the tracking of environmental contaminants\citep{olascoaga12:_forec}.
 
-We also hope that LCS Tool will serve as a foundation for the numerical
-implementation of recent theoretical advances.
-These include the geodesic theory of parabolic LCSs (jet cores)\citep{farazmand14:_shearless} and the variational theory of hyperbolic and elliptic LCSs for three-dimensional
-flows\citep{blazevski14:_hyper_ellip_trans_barrier_three}.
+We hope LCS Tool will serve as a foundation for the numerical implementation of recent theoretical advances.
+These include the geodesic theory of parabolic LCSs (jet cores)\citep{farazmand14:_shearless} and the variational theory of hyperbolic and elliptic LCSs for three-dimensional flows\citep{blazevski14:_hyper_ellip_trans_barrier_three}.
 

### 2014-10-31 main.tex
Correct corresponding author indication
Address reviewer comment

@@
 
-% FIXME Confirm George can be corresponding author
-\author{G.~Haller}
+\author{G.~Haller\corref{cor}}
 \ead{georgehaller@ethz.ch}
@@
 
-Here we present a simple algorithmic introduction to geodesic LCS detection.
+Here we present an algorithmic introduction to geodesic LCS detection.
 % FIXME JCompSci editor: Replace GitHub link with http://runmycode.org/
@@
 This engine is a library of MATLAB functions that extract LCSs from two-dimensional unsteady flows.
-The examples we present have been programmed into demonstration scripts.
+The examples we present form demonstration scripts distributed with LCS Tool.
 
-% FIXME Reviewer #3, comment #3: The comparison with other publicly available LCS detection software should include more details.
-It contains a very high-level summary of LCS tool being an "all-purpose numerical engine".
-That part should be revisited and become more concrete.
-Software for LCS detection includes the following packages:
-\begin{itemize}
-\item \emph{MANGEN}\citep{lekien03:_time} calculates FTLE and advects material curves in two dimensional velocity fields.
-It also includes a graphical user interface and uses MPI for parallel calculations.
-% FIXME Cannot cite LCS MATLAB Kit without a journal reference
-% \item \emph{LCS MATLAB Kit}\citep{dabiri09:_lmk} calculates the FTLE from velocity datasets. Includes a graphical user interface.
-\item \emph{Newman}\citep{toit10:_trans} calculates the FTLE in $N$ dimensions. Assists ridge extraction of FTLE fields. Supports analytic  and dataset based velocity definitions.
-% FIXME Cannot cite FlowVC without a journal reference
-% \item \emph{FlowVC}\citep{shadden10:_flowvc} is a general purpose LCS platform for two and three dimensional datasets. Parallel calculations are supported with Open\-MP, CUDA and OpenCL.
-% FIXME Add FlowTK, https://github.com/FlowPhysics/FlowTK, doi:10.1007%2F978-3-319-04099-8_13
-\end{itemize}
-These packages primarily target the automated generation of FTLE plots to aid the visual assessment of hyperbolic LCSs in the flow.
-The purpose of LCS Tool and this article is to provide an all-purpose numerical engine for the geodesic extraction of LCSs as parametrized material curves in unsteady two-dimensional flows, and extend the scope of such extraction to elliptic LCSs.
+Other software for LCS detection exists.
+ManGen\citep{lekien03:_time} calculates the FTLE and advects material curves in two-dimensional velocity fields.
+It includes a graphical user interface and uses the Message Passing Interface standard for parallel calculations.
+Newman\citep{toit10:_trans} calculates the FTLE using dimension independent code.
+It assists ridge extraction of FTLE fields and supports analytic and dataset based velocity definitions.
+FlowTK\citep{ameli14:_devel_effic_flexib_pipel_lagran} calculates the FTLE in two and three-dimensions on cartesian and unstructured grids.
+The NVIDIA CUDA parrallel computing platform is used for fast computations and Kitware's ParaView data analysis and visualization application is used for a user interface.
+
+These packages target the automated generation of FTLE plots to aid the visual assessment of hyperbolic LCSs.
+LCS Tool also has functions to generate FTLE plots, but its emphasis is to provide geodesic extraction of LCSs as parametrized material curves, and extend the scope of such extraction to elliptic LCSs.
 
@@
 \end{gather*}
+% FIXME Clarify what is meant by _closed_ material lines
 obtained by averaging the Lagrangian shear arising over $[t_0,t]$ along closed material lines parametrized as $r(s)$ with $s \in [0,\sigma]$.

### 2014-10-31 main.tex
Correct typo

@@
 % FIXME Small tolerances (i.e. strict tolerances) should make convergence difficult not easy, so the word "facilitate" in the sentence below seems illogical.
-In this script, we have chosen small values for the error tolerances of the integration of the Cauchy-Green strain tensor (\cref{ll:double gyre cgStrainOdeSolverOptions}) and $\lambda$-lines (\cref{ll:double gyre lambdaLineOdeSolverOptions})to facilitate convergence.
+In this script, we have chosen small values for the error tolerances of the integration of the Cauchy-Green strain tensor (\cref{ll:double gyre cgStrainOdeSolverOptions}) and $\lambda$-lines (\cref{ll:double gyre lambdaLineOdeSolverOptions}) to facilitate convergence.
 

### 2014-11-06 main.tex
Rename strainline to shrinkline

@@
 r' = \xi_1(r),
-\label{eq:strainline}
+\label{eq:shrinkline}
 \end{equation}
@@
 
-This section describes stepwise the numerical implementation of geodesic detection of elliptic and hyperbolic LCSs based on \cref{eq:etafields,eq:strainline,eq:stretchline}.
+This section describes stepwise the numerical implementation of geodesic detection of elliptic and hyperbolic LCSs based on \cref{eq:etafields,eq:shrinkline,eq:stretchline}.
 \Cref{t:LCS Tool overview} gives an overview of the steps, functions and variable names used in LCS Tool to analyse a flow.
@@
 
-As discussed in \cref{sec:Hyperbolic LCSs}, positions of hyperbolic LCSs at time $t_0$ are found as the strongest repelling orbits of the vector field \cref{eq:strainline} (repelling LCSs), and strongest attracting orbits of the vector field \cref{eq:stretchline} (attracting LCSs).
+As discussed in \cref{sec:Hyperbolic LCSs}, positions of hyperbolic LCSs at time $t_0$ are found as the strongest repelling orbits of the vector field \cref{eq:shrinkline} (repelling LCSs), and strongest attracting orbits of the vector field \cref{eq:stretchline} (attracting LCSs).
 By repulsion and attraction we mean a property of the LCS (as an evolving material line) under the flow map $F_{t_0}^t$.
@@
 \item Define a maximum shrinkline length
-\item Integrate a shrinkline forward and backward according to \cref{eq:strainline} and using the largest $\lambda_2$ local maximum as the initial position.
+\item Integrate a shrinkline forward and backward according to \cref{eq:shrinkline} and using the largest $\lambda_2$ local maximum as the initial position.
 Integrate until the shrinkline has attained the maximum shrinkline length, or until it has reached the domain boundary.
@@
 \Cref{l:double gyre lambda strain stretch LCS} lists the corresponding LCS Tool commands.
-The maximal length of shrinklines and stretchlines, \lstinline!strainlineMaxLength!
+The maximal length of shrinklines and stretchlines, \lstinline!shrinklineMaxLength!
 and \lstinline!stretchlineMaxLength!, is set to $20$, a multiple of the domain size, since the hyperbolic LCSs may wind around vortices several times.
@@
 % End:
-% LocalWords:  stretchline stretchlines incompressible gyre eigenvector eigenvectors incompressibility dataset datasets LCS LCSs Lyapunov Bickley Rossby timespan variational adaptively vortices parametrise advected strainlines minima extremal FTLE MATLAB maxima geostrophic shrinkline shrinklines Poincare altimetry zonal
+% LocalWords:  stretchline stretchlines incompressible gyre eigenvector eigenvectors incompressibility dataset datasets LCS LCSs Lyapunov Bickley Rossby timespan variational adaptively vortices parametrise advected minima extremal FTLE MATLAB maxima geostrophic shrinkline shrinklines Poincare altimetry zonal

### 2014-11-06 listings/double_gyre/hyperbolic_lcs_subset.m

@@ -2,7 +2,7 @@
 shrinklineMaxLength = 20;
 gridSpace = diff(domain(1,:))/(double(resolution(1))-1);
 shrinklineLocalMaxDistance = 2*gridSpace; %@\label{ll:shrinklineLocalMaxDistance}
-strainlineOdeSolverOptions = odeset('relTol',1e-6);
+shrinklineOdeSolverOptions = odeset('relTol',1e-6);
 
 % Stretchlines
 stretchlineMaxLength = 20;

### 2014-11-06 main.tex
Rename lambda-line LCS to elliptic LCS

@@
 By examining the FTLE field (which LCS Tool can provide), we position Poincare sections to capture elliptic LCSs.
-An LCS Tool demo script to perform this operation is given in \cref{l:double gyre lambda-line LCS} where two Poincare sections are defined (\cref{ll:double gyre poincareSection(1),ll:double gyre poincareSection(2)}).
+An LCS Tool demo script to perform this operation is given in \cref{l:double gyre elliptic LCS} where two Poincare sections are defined (\cref{ll:double gyre poincareSection(1),ll:double gyre poincareSection(2)}).
 The free stretching parameter $\lambda$ (\cref{eq:etafields}) is varied over the range $[0.93,1.07]$ with increments of $0.01$ (\cref{ll:double gyre lambdaRange}).
@@
 \lstinputlisting[caption={LCS Tool commands for double gyre elliptic LCSs. Subset of the LCS Tool script \lstinline!demo/double_gyre/hyperbolic_shear_lcs.m!},
-                 label=l:double gyre lambda-line LCS]{listings/double_gyre/elliptic_lcs_subset.m}
+                 label=l:double gyre elliptic LCS]{listings/double_gyre/elliptic_lcs_subset.m}
 

### 2014-11-06 main.tex
Correct spelling

@@
 It assists ridge extraction of FTLE fields and supports analytic and dataset based velocity definitions.
-FlowTK\citep{ameli14:_devel_effic_flexib_pipel_lagran} calculates the FTLE in two and three-dimensions on cartesian and unstructured grids.
-The NVIDIA CUDA parrallel computing platform is used for fast computations and Kitware's ParaView data analysis and visualization application is used for a user interface.
+FlowTK\citep{ameli14:_devel_effic_flexib_pipel_lagran} calculates the FTLE in two and three-dimensions on Cartesian and unstructured grids.
+The NVIDIA CUDA parallel computing platform is used for fast computations and Kitware's ParaView data analysis and visualization application is used for a user interface.
 

### 2014-11-07 main.tex
Switch to two column layout
To detect overfull boxes use:
\documentclass[5p,draft]{elsarticle}
\usepackage[final]{listings}

Additional hints: http://tex.stackexchange.com/a/2780/29072

@@
-% FIXME Add 5p option to get JCompSci two column layout
-\documentclass{elsarticle}
+\documentclass[5p]{elsarticle}
 
@@
 
-\definecolor{gray}{rgb}{0.9,0.9,0.9}
+\definecolor{grey}{rgb}{0.9,0.9,0.9}
 
@@
   stepnumber=2,
-  backgroundcolor=\color{gray},
+  backgroundcolor=\color{grey},
   escapeinside={\%@}{\^^M}
@@
 
+\begin{sloppypar}
 % FIXME Reviewer #5, comment #5: Please provide information regarding practical applications of this algorithmic software tool.
@@
 Applications of LCSs include oceanic and atmospheric flows\citep{beron-vera13:_objec_agulh,koh02:_hyper}, biological transport problems\citep{wilson09:_lagran_reynol,tallapragada11:_lagran,huhn12:_south_indian_ocean_count_madag}, aeronautics\citep{tang10:_accur_lagran_hong_kong_inter_airpor}, celestial mechanics\citep{gawlik09:_lagran}, crowd dynamics\citep{ali07:_lagran_partic_dynam_approac_crowd}, and aperiodically forced mechanical oscillators\citep{hadjighasem13:_detec_kam}.
+\end{sloppypar}
 
@@
 
+\begin{sloppypar}
 Here we present an algorithmic introduction to geodesic LCS detection.
@@
 The examples we present form demonstration scripts distributed with LCS Tool.
+\end{sloppypar}
 
+\begin{sloppypar}
 Other software for LCS detection exists.
@@
 The NVIDIA CUDA parallel computing platform is used for fast computations and Kitware's ParaView data analysis and visualization application is used for a user interface.
+\end{sloppypar}
 
+\begin{sloppypar}
 These packages target the automated generation of FTLE plots to aid the visual assessment of hyperbolic LCSs.
 LCS Tool also has functions to generate FTLE plots, but its emphasis is to provide geodesic extraction of LCSs as parametrized material curves, and extend the scope of such extraction to elliptic LCSs.
+\end{sloppypar}
 
@@
 
+\begin{sloppypar}
 We seek positions of closed material lines at time $t_0$ that prevail as coherent Lagrangian vortex boundaries (or \emph{elliptic LCSs}) over a time interval $[t_0,t]\subset[t_-,t_+]$.
@@
 For this reason, we refer to the computation of elliptic LCSs as limit cycles of \cref{eq:etafields} as \emph{geodesic detection of elliptic LCSs}.
+\end{sloppypar}
 
@@
 
+\begin{sloppypar}
 Shrinklines and stretchlines coincide with the null-geodesics of the Lorentzian metric $h(u,v) = \left\langle u,D_{t_0}^t(r) v \right\rangle$.
 For this reason, we refer to the computation of hyperbolic LCSs as strongest normally-repelling or normally-attracting orbits of \cref{eq:etafields} as \emph{geodesic detection of hyperbolic LCSs}.
+\end{sloppypar}
 
@@
 
-\begin{table}
+\begin{table*}
 \begin{center}
@@
 \hline
-\multicolumn{2}{|p{\textwidth}|}{\lstinline![cgEigenvector,cgEigenvalue] = eig_cgStrain(derivative,domain,timespan,resolution)!}\tabularnewline
+\multicolumn{2}{|p{.95\textwidth}|}{\lstinline![cgEigenvector,cgEigenvalue] = eig_cgStrain(derivative,domain,timespan,resolution)!}\tabularnewline
 \hline
@@
 \end{center}
-\end{table}
+\end{table*}
 
@@
 These points are used to achieve increased accuracy in the finite-difference approximation
-\[
+\begin{gather*}
 \nabla F_{t_0}^t(x_0) \approx
 \left(\begin{array}{cc}
-\frac{x_1(t;t_0,x_0 + \delta x_1) - x_1(t;t_0,x_0 - \delta x_1)}{2\vert\delta x_1\vert} & \frac{x_1(t;t_0,x_0+\delta x_2) - x_1(t;t_0,x_0-\delta x_2)}{2\vert\delta x_2\vert}\\
-\frac{x_2(t;t_0,x_0+\delta x_1) - x_2(t;t_0,x_0 - \delta x_1)}{2\vert\delta x_1\vert} & \frac{x_2(t;t_0,x_0 + \delta x_2) - x_2(t;t_0,x_0 - \delta x_2)}{2\vert\delta x_2\vert}
+\alpha_{11} & \alpha_{12}
+\\
+\alpha_{21} & \alpha_{22}
 \end{array}\right)
-\]
+\\
+\alpha_{ij} \equiv \frac{x_i(t;t_0,x_0+\delta x_j) - x_i(t;t_0,x_0-\delta x_j)}{2\vert\delta x_j\vert}
+\end{gather*}
 of $\nabla F_{t_0}^t$ at a point $x_0$ of the Cartesian grid.
-Here $\delta x_i$ is a vector of length $\vert\delta x_i\vert > 0$ that points from the Cartesian grid-point $x_0$ in the $i^\text{th}$ coordinate direction (\cref{f:main and auxiliary grids}).
+Here $\delta x_j$ is a vector of length $\vert\delta x_j\vert > 0$ that points from the Cartesian grid-point $x_0$ in the $j^\text{th}$ coordinate direction (\cref{f:main and auxiliary grids}).
 Computational improvements arising from the use of the auxiliary grid over simply using the nearest points of the main grid were reported in \citet{farazmand12:_comput_lagran}.
@@
 \begin{center}
-\includegraphics[width=.5\textwidth]{graphics/main_aux_grids}
+\includegraphics[width=.475\textwidth]{graphics/main_aux_grids}
 \end{center}
@@
 
+\begin{sloppypar}
 The function \lstinline!eig_cgStrain!
@@
 For flows defined by datasets, using the auxiliary grid for eigenvalue calculations gives better results.
+\end{sloppypar}
 
+\begin{sloppypar}
 As mentioned above, a typical main grid for the Cauchy-Green strain tensor comprises $500 \times 500$ points.
 This means that after the addition of 4 auxiliary grid points around each main grid point, \cref{eq:velocity} must be integrated over 1.25 million initial conditions.
+\end{sloppypar}
 
@@
 
+\begin{sloppypar}
 The Cauchy-Green strain tensor generically admits point singularities, i.e., points where $C_{t_0}^t(x_0)$ has repeated eigenvalues.
@@
 This generically arises at a finite set of isolated points within the computational domain\citep{delmarcelle94}, and hence lie off the computational grid with probability one.
+\end{sloppypar}
 
@@
 
-Velocity fields defined by datasets require pre-processing before they are used in the numerical integration of \cref{eq:velocity}.
+Velocity fields defined by datasets require preprocessing before they are used in the numerical integration of \cref{eq:velocity}.
 This requires spatial and temporal interpolation that enables the evaluation of the velocity function at arbitrary points in $U$ and at arbitrary times between $t_-$ and $t_+$.
@@
 % FIXME Need to update table to account for lambda range
-\begin{table}
+\begin{table*}
 \begin{tabular}{|c|p{.7\textwidth}|}
 \hline
-\multicolumn{2}{|p{\textwidth}|}
+\multicolumn{2}{|p{.95\textwidth}|}
 {\lstinline![shearline.etaPos,shearline.etaNeg] = lambda_line(cgEigenvector,cgEigenvalue,lambda)!}\tabularnewline
@@
 \hline \hline
-\multicolumn{2}{|p{\textwidth}|}{\lstinline![closedOrbits,orbits] = poincare_closed_orbit_multi(domain,resolution,shearline,PSList)!}\tabularnewline
+\multicolumn{2}{|p{.95\textwidth}|}{\lstinline![closedOrbits,orbits] = poincare_closed_orbit_multi(domain,resolution,shearline,PSList)!}\tabularnewline
 \hline
@@
 \label{t:Elliptic LCS functions}
-\end{table}
+\end{table*}
 
@@
 \begin{center}
-\includegraphics[width=.8\textwidth]{graphics/variable_step_integration}
+\includegraphics[width=.475\textwidth]{graphics/variable_step_integration}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=.8\textwidth]{graphics/double_gyre/poincare_return_map}
+\includegraphics[width=.475\textwidth]{graphics/double_gyre/poincare_return_map}
 \end{center}
@@
 
+\begin{sloppypar}
 As discussed in \cref{sec:Hyperbolic LCSs}, positions of hyperbolic LCSs at time $t_0$ are found as the strongest repelling orbits of the vector field \cref{eq:shrinkline} (repelling LCSs), and strongest attracting orbits of the vector field \cref{eq:stretchline} (attracting LCSs).
@@
 These local maxima and minima of the appropriate $\lambda_i(x_0)$ eigenvalue field can be thought of as the extensions of the concept of saddle points to the present finite-time, temporally aperiodic flow setting.
+\end{sloppypar}
 
+\begin{sloppypar}
 The main steps of our hyperbolic LCS detection algorithm are enumerated in \cref{t:Hyperbolic LCS algorithm}.
@@
 and its syntax is given in \cref{t:seed_curves_from_lambda_max syntax}.
+\end{sloppypar}
 
@@
 \hline
-\multicolumn{2}{|p{\textwidth}|}
+\multicolumn{2}{|p{.95\textwidth}|}
 {\lstinline![curvePosition,curveInitialPosition] = seed_curves_from_lambda_max(distance,cgEigenvalue,cgEigenvector,flowDomain,flowResolution)!}\tabularnewline
@@
 \label{t:seed_curves_from_lambda_max syntax}
-\end{table}
+\end{table*}
 
@@
 
-\begin{figure}[hbt]
+\begin{figure}
 \centering
-\includegraphics[width=.49\textwidth]{graphics/ocean_dataset/LCS_fwd_coherent_eddy}
-\includegraphics[width=.49\textwidth]{graphics/ocean_dataset/LCS_fwd_colortracer}
-\caption{(Left) Elliptic LCS (green) with repelling LCSs (red) continuing into its interior.
-(Right) Final positions $x(t) = (x_1(t),x_2(t)) = F_{t_0}^t(x_0)$ of tracer particles.
+\includegraphics[width=.475\textwidth]{graphics/ocean_dataset/LCS_fwd_coherent_eddy}
+\includegraphics[width=.475\textwidth]{graphics/ocean_dataset/LCS_fwd_colortracer}
+\caption{(Top) Elliptic LCS (green) with repelling LCSs (red) continuing into its interior.
+(Bottom) Final positions $x(t) = (x_1(t),x_2(t)) = F_{t_0}^t(x_0)$ of tracer particles.
 $x_1(t)$ position encoded in red and $x_2(t)$ position encoded in green.
@@
 
-\begin{figure}[hbt]
+\begin{figure}
 \centering
-\includegraphics[width=.8\textwidth]{graphics/double_gyre/lambda_lcs_convergence}\caption{Convergence of closed $\lambda$-lines for increasing main-grid resolution in the double-gyre model.
+\includegraphics[width=.475\textwidth]{graphics/double_gyre/lambda_lcs_convergence}\caption{Convergence of closed $\lambda$-lines for increasing main-grid resolution in the double-gyre model.
 The outermost closed $\lambda$-line (bold green line) is the vortex boundary.
@@
 
+\begin{sloppypar}
 \Cref{f:double gyre lambda hyperbolic LCS} shows the complete picture of elliptic and hyperbolic LCSs of the double gyre at this resolution.
@@
 and \lstinline!stretchlineMaxLength!, is set to $20$, a multiple of the domain size, since the hyperbolic LCSs may wind around vortices several times.
+\end{sloppypar}
 
@@
 % FIXME Reviewer #2, comment #6: Improve appearance of $lambda$ extrema (white dots)
-\includegraphics[width=\textwidth]{graphics/double_gyre/lambda_hyperbolic_lcs}
+\includegraphics[width=.475\textwidth]{graphics/double_gyre/lambda_hyperbolic_lcs}
 \caption{
@@
 % FIXME Shrinkline LCS figure does not treat periodic boundary condition correctly
-\includegraphics[width=.85\textwidth]{graphics/bickley_jet/lambda_hyperbolic_lcs}
+\includegraphics[width=.475\textwidth]{graphics/bickley_jet/lambda_hyperbolic_lcs}
 \caption{
@@
 
+\begin{sloppypar}
 In the geostrophic approximation, the sea-surface height $\eta$ serves as a stream-function for surface velocities.
@@
 where $g$ is the constant of gravity, $R$ is the mean radius of the Earth, and $f(\theta) \equiv 2\Omega\sin\theta$ is the Coriolis parameter, with $\Omega$ denoting the Earth's mean angular velocity.
+\end{sloppypar}
 
+\begin{sloppypar}
 The data is given at a spatial resolution of $1/4\degree$ and a temporal resolution of 7 days.
@@
 The commands for the interpolation of the velocity data are given in \crefrange{ll:interpolation start}{ll:interpolation end}.
+\end{sloppypar}
 
@@
 \begin{center}
-\includegraphics[width=.8\textwidth]{graphics/ocean_dataset/lambda_hyperbolic_lcs}
+\includegraphics[width=.475\textwidth]{graphics/ocean_dataset/lambda_hyperbolic_lcs}
 \end{center}

### 2014-11-07 main.tex
Convert PDF figures to version 1.5
Update keywords

@@
 \DeclareMathOperator{\sech}{sech}
-
-% FIXME Some figures use PDF version 1.6, PDFLaTeX default is 1.5
-\pdfoptionpdfminorversion 6
 
@@
 \begin{keyword}
-% FIXME Match keywords and PACS and MSC classifications
-% FIXME Consult previous publications to find keywords
-Lagrangian coherent structures \sep Mathematical software \sep Dynamical systems \sep Invariant manifolds \sep Mixing
+Lagrangian coherent structures \sep Mathematical software \sep Non-autonomous dynamical systems \sep Invariant manifolds \sep Mixing \sep Transport barriers \sep Fluid dynamics \sep Ocean surface flows
 
-% FIXME Verify these classifications, http://www.aip.org/pacs
-\PACS 47.10.Fg \sep 47.27.De \sep 47.11.-j
+\PACS 47.10.Fg \sep 47.27.De \sep 47.11.-j \sep 05.45.-a \sep 47.51.+a \sep 47.54.-r \sep 92.10.hf \sep 93.30.Qn \sep 92.10.ab \sep 92.10.Ty \sep 92.10.ak \sep 92.10.A-
 
-% FIXME Verify these classifications http://www.ams.org/msc/msc2010.html
-\MSC[2010] 37N10 \sep 37B25 \sep 97N80
+\MSC[2010] 37N10 \sep 37B25 \sep 97N80 \sep 37C60 \sep 37B55 \sep 70H33 \sep 76F25
 \end{keyword}

### 2014-11-07 main.tex
Add bibliography references

@@
 
-More recent work has focused on an exact mathematical formulation of the properties defining LCSs\citep{haller11:_lagran_coher_struc,farazmand12:_comput_lagran,haller12:_geodes_theor_trans_barrier_two_dimen_flows,haller13:_coher_lagran,blazevski14:_hyper_ellip_trans_barrier_three}.
-In two-dimensional flows, hyperbolic and parabolic LCSs turn out to be stationary curves of the averaged material shear\citep{farazmand14:_shearless}, whereas elliptic LCSs are stationary curves of the averaged strain\citep{haller13:_coher_lagran}.
+More recent work has focused on an exact mathematical formulation of the properties defining LCSs\citep{haller11:_lagran_coher_struc,farazmand12:_comput_lagran,haller12:_geodes_theor_trans_barrier_two_dimen_flows,haller13:_coher_lagran,haller14:_adden_coher_lagran,farazmand13:_attrac_lagran,blazevski14:_hyper_ellip_trans_barrier_three}.
+In two-dimensional flows, hyperbolic and parabolic LCSs turn out to be stationary curves of the averaged material shear\citep{farazmand14:_shearless}, whereas elliptic LCSs are stationary curves of the averaged strain\citep{haller13:_coher_lagran,haller14:_adden_coher_lagran}.
 These variational formulations lead to explicit solutions for LCSs as null-geodesics of appropriate Lorentzian metrics.
@@
 We seek positions of closed material lines at time $t_0$ that prevail as coherent Lagrangian vortex boundaries (or \emph{elliptic LCSs}) over a time interval $[t_0,t]\subset[t_-,t_+]$.
-\citet{haller13:_coher_lagran} argue that such initial material line positions are closed stationary curves of the averaged strain functional
+\citet{haller13:_coher_lagran,haller14:_adden_coher_lagran} argue that such initial material line positions are closed stationary curves of the averaged strain functional
 \[
@@
 For this reason, we refer to trajectories of \cref{eq:etafields} as $\lambda$-lines.
-Following \citet{haller13:_coher_lagran}, we call the outermost member of a closed family of $\lambda$-lines a coherent Lagrangian vortex boundary.
+Following \citet{haller13:_coher_lagran,haller14:_adden_coher_lagran}, we call the outermost member of a closed family of $\lambda$-lines a coherent Lagrangian vortex boundary.
 
@@
 A $\lambda$-line returning to its starting point is then an elliptic LCS.
-The outermost member of a family of closed $\lambda$-lines (obtained by varying $\lambda$) is a coherent Lagrangian vortex boundary\citep{haller13:_coher_lagran}.
+The outermost member of a family of closed $\lambda$-lines (obtained by varying $\lambda$) is a coherent Lagrangian vortex boundary\citep{haller13:_coher_lagran,haller14:_adden_coher_lagran}.
 
@@
 In contrast to the previous two analytic examples, the velocity field is available only with discrete temporal and spatial resolution.
-Our region of interest is a small domain in the South Atlantic Ocean, where exceptionally coherent eddies, Agulhas rings, were recently found by \citet{haller13:_coher_lagran} using the theory we surveyed in \cref{sec:Elliptic LCSs}.
+Our region of interest is a small domain in the South Atlantic Ocean, where exceptionally coherent eddies, Agulhas rings, were recently found by \citet{haller13:_coher_lagran,haller14:_adden_coher_lagran} using the theory we surveyed in \cref{sec:Elliptic LCSs}.
 
@@
 
-\Cref{f:ocean dataset LCS} shows the positions of elliptic and hyperbolic LCSs on 22 November 2006, which is the same time instant analysed in \citet{haller13:_coher_lagran,beron-vera13:_objec_agulh}.
+\Cref{f:ocean dataset LCS} shows the positions of elliptic and hyperbolic LCSs on 22 November 2006, which is the same time instant analysed in \citet{haller13:_coher_lagran,haller14:_adden_coher_lagran,beron-vera13:_objec_agulh}.
 Here the integration time is chosen as $T = 30$ days, as opposed to 90 days in the cited references, to avoid tangling of hyperbolic LCSs.

### 2014-11-07 main.tex
Use hyperref package

@@
 \usepackage[numbered,framed]{mcode}
+\usepackage[hidelinks]{hyperref}
 \usepackage{cleveref}
@@
 }
+
+% Workaround for bug in elsarticle-num-names.bst
+% Reference: http://www.tug.org/pipermail/tex-live/2012-August/032152.html
+\makeatletter
+\providecommand{\doi}[1]{%
+  \begingroup
+    \let\bibinfo\@secondoftwo
+    \urlstyle{rm}%
+    \href{http://dx.doi.org/#1}{%
+      doi:\discretionary{}{}{}%
+      \nolinkurl{#1}%
+    }%
+  \endgroup
+}
+\makeatother
 
@@
 
-The altimeter products used in this work are produced by SSALTO/DUACS and distributed by AVISO, with support from CNES, http://www.aviso.oceanobs.com/duacs.
+\begin{sloppypar}
+The altimeter products used in this work are produced by SSALTO/DUACS and distributed by AVISO, with support from CNES, \url{http://www.aviso.oceanobs.com/duacs}.
+\end{sloppypar}
 

### 2014-11-07 main.tex
Change LCS Tool download URL from GitHub to RunMyCode

@@
 Here we present an algorithmic introduction to geodesic LCS detection.
-% FIXME JCompSci editor: Replace GitHub link with http://runmycode.org/
 % FIXME Reviewer #3, comment #4: Add documentation within the github repository to show some reference to the APIs and how users may start to run a his/her own application.
 % FIXME Reviewer #4: Please make sure the code is downloadable once the paper is published
-We then review the implementation of this approach in a computational engine called LCS Tool\footnote{LCS Tool is available at: http://github.com/jeixav/LCS-Tool}.
-This engine is a library of MATLAB functions that extract LCSs from two-dimensional unsteady flows.
-The examples we present form demonstration scripts distributed with LCS Tool.
+We then review the implementation of this approach in a computational engine called LCS Tool\footnote{LCS Tool is available at: \url{http://www.runmycode.org/companion/view/908}}.
+This engine is a library of MATLAB functions that extract LCSs from two-dimensional unsteady flows. The examples we present form demonstration scripts distributed with LCS Tool.
 \end{sloppypar}

### 2014-11-08 main.tex
Update bibliography

@@
 Incompressible flows (i.e.
-those satisfying $\nabla \cdot v=0$) satisfy the relation $\lambda_1(x_0) \lambda_2(x_0) = 1$ at all points of the computational domain \citep{arnold78:_mathem}.
+those satisfying $\nabla \cdot v=0$) satisfy the relation $\lambda_1(x_0) \lambda_2(x_0) = 1$ at all points of the computational domain \citep{arnold89:_mathem}.
 Incompressibility can be computationally imposed by first calculating $\lambda_2(x_0)$, then setting $\lambda_1(x_0) = 1/\lambda_2(x_0)$ and calculating the strain eigenvectors $\xi_2$ from $\lambda_2$, then $\xi_1$ from the relationship in \cref{eq:CG_invariants}.
@@
 
+% FIXME Citations without pages (haller15:_langr_coher_struc and karrasch14:_autom_lagran) do not list publication year
 \bibliographystyle{elsarticle-num-names} 

### 2014-11-09 main.tex
Correct equation references in ocean demo section

@@
 In a longitude-latitude $(\varphi,\theta)$ coordinate system, the evolution of a fluid particle is given by
-\begin{eqnarray}
+\begin{gather}
 \pd{\varphi(\varphi,\theta,t)}{t} = -\frac{g}{R^2 f(\theta) \cos\theta} \pd{\eta(\varphi,\theta,t)}{\theta}\label{eq:dtlon}\\
@@
 \label{eq:dtlat}
-\end{eqnarray}
+\end{gather}
 where $g$ is the constant of gravity, $R$ is the mean radius of the Earth, and $f(\theta) \equiv 2\Omega\sin\theta$ is the Coriolis parameter, with $\Omega$ denoting the Earth's mean angular velocity.

### 2014-11-09 main.tex
Use the microtype package
The microtype package reduces the need for sloppypar

@@
 \usepackage[numbered,framed]{mcode}
+\usepackage{microtype}
 \usepackage[hidelinks]{hyperref}
@@
 
-\begin{sloppypar}
 We seek positions of closed material lines at time $t_0$ that prevail as coherent Lagrangian vortex boundaries (or \emph{elliptic LCSs}) over a time interval $[t_0,t]\subset[t_-,t_+]$.
@@
 For this reason, we refer to the computation of elliptic LCSs as limit cycles of \cref{eq:etafields} as \emph{geodesic detection of elliptic LCSs}.
-\end{sloppypar}
 
@@
 
-\begin{sloppypar}
 The data is given at a spatial resolution of $1/4\degree$ and a temporal resolution of 7 days.
@@
 The commands for the interpolation of the velocity data are given in \crefrange{ll:interpolation start}{ll:interpolation end}.
-\end{sloppypar}
 

### 2014-11-09 main.tex
Correct typo
Modify spelling
- null-geodesic → null geodesic
- shrinkline → shrink line
- stretchline → stretch line

Reference: http://arxiv.org/abs/1405.4727

@@
 In two-dimensional flows, hyperbolic and parabolic LCSs turn out to be stationary curves of the averaged material shear\citep{farazmand14:_shearless}, whereas elliptic LCSs are stationary curves of the averaged strain\citep{haller13:_coher_lagran,haller14:_adden_coher_lagran}.
-These variational formulations lead to explicit solutions for LCSs as null-geodesics of appropriate Lorentzian metrics.
+These variational formulations lead to explicit solutions for LCSs as null geodesics of appropriate Lorentzian metrics.
 
@@
 \end{equation}
-which coincide with null-geodesics of the Lorentzian metric family 
+which coincide with null geodesics of the Lorentzian metric family 
 \[
@@
 
-Solutions to this variational problem turn out to be orbits of the $\xi_1$ or $\xi_2$ eigenvector field. Repelling LCSs (\emph{shrinklines}) are obtained as trajectories of the differential equation
+Solutions to this variational problem turn out to be orbits of the $\xi_1$ or $\xi_2$ eigenvector field. Repelling LCSs (\emph{shrink lines}) are obtained as trajectories of the differential equation
 \begin{equation}
 r' = \xi_1(r),
-\label{eq:shrinkline}
+\label{eq:shrink line}
 \end{equation}
-and attracting LCSs (\emph{stretchlines}) are obtained as trajectories of the differential equation
+and attracting LCSs (\emph{stretch lines}) are obtained as trajectories of the differential equation
 \begin{equation}
 r' = \xi_2(r).
-\label{eq:stretchline}
+\label{eq:stretch line}
 \end{equation}
 
-\begin{sloppypar}
-Shrinklines and stretchlines coincide with the null-geodesics of the Lorentzian metric $h(u,v) = \left\langle u,D_{t_0}^t(r) v \right\rangle$.
+Shrink lines and stretch lines coincide with the null geodesics of the Lorentzian metric $h(u,v) = \left\langle u,D_{t_0}^t(r) v \right\rangle$.
 For this reason, we refer to the computation of hyperbolic LCSs as strongest normally-repelling or normally-attracting orbits of \cref{eq:etafields} as \emph{geodesic detection of hyperbolic LCSs}.
-\end{sloppypar}
 
-To compute the normal repulsion of shrinklines, we note that an infinitesimal
-normal perturbation to a shrinkline $\gamma$ at its point $r$ grows under the flow map $F_{t_{0}}^{t}$ by a factor of $\lambda_{2}(r)$ in the direction normal to $F_{t_{0}}^{t}(\gamma).$ Similarly, small normal perturbations to a stretchline decay by a factor $\lambda_{1}(r)$ in the direction normal to the evolving stretchline.
+To compute the normal repulsion of shrink lines, we note that an infinitesimal
+normal perturbation to a shrink line $\gamma$ at its point $r$ grows under the flow map $F_{t_{0}}^{t}$ by a factor of $\lambda_{2}(r)$ in the direction normal to $F_{t_{0}}^{t}(\gamma).$ Similarly, small normal perturbations to a stretch line decay by a factor $\lambda_{1}(r)$ in the direction normal to the evolving stretch line.
 
@@
 
-This section describes stepwise the numerical implementation of geodesic detection of elliptic and hyperbolic LCSs based on \cref{eq:etafields,eq:shrinkline,eq:stretchline}.
+This section describes stepwise the numerical implementation of geodesic detection of elliptic and hyperbolic LCSs based on \cref{eq:etafields,eq:shrink line,eq:stretch line}.
 \Cref{t:LCS Tool overview} gives an overview of the steps, functions and variable names used in LCS Tool to analyse a flow.
@@
 \begin{sloppypar}
-As discussed in \cref{sec:Hyperbolic LCSs}, positions of hyperbolic LCSs at time $t_0$ are found as the strongest repelling orbits of the vector field \cref{eq:shrinkline} (repelling LCSs), and strongest attracting orbits of the vector field \cref{eq:stretchline} (attracting LCSs).
+As discussed in \cref{sec:Hyperbolic LCSs}, positions of hyperbolic LCSs at time $t_0$ are found as the strongest repelling orbits of the vector field \cref{eq:shrink line} (repelling LCSs), and strongest attracting orbits of the vector field \cref{eq:stretch line} (attracting LCSs).
 By repulsion and attraction we mean a property of the LCS (as an evolving material line) under the flow map $F_{t_0}^t$.
-We identify the strongest repelling shrinklines as those crossing a local maximum of the $\lambda_2(x_0)$ field.
-Similarly, we identify the strongest attracting stretchlines as those crossing a local minimum of the $\lambda_1(x_0)$ field.
+We identify the strongest repelling shrink lines as those crossing a local maximum of the $\lambda_2(x_0)$ field.
+Similarly, we identify the strongest attracting stretch lines as those crossing a local minimum of the $\lambda_1(x_0)$ field.
 These local maxima and minima of the appropriate $\lambda_i(x_0)$ eigenvalue field can be thought of as the extensions of the concept of saddle points to the present finite-time, temporally aperiodic flow setting.
@@
 \item Find all points of the main grid that are local maxima of $\lambda_2$ within a circle whose radius is the local  maximization distance.
-\item Define a maximum shrinkline length
-\item Integrate a shrinkline forward and backward according to \cref{eq:shrinkline} and using the largest $\lambda_2$ local maximum as the initial position.
-Integrate until the shrinkline has attained the maximum shrinkline length, or until it has reached the domain boundary.
-\item Flag any remaining local maxima of $\lambda_2$ within the maximization distance of the shrinkline as ineligible initial positions for subsequent shrinklines
+\item Define a maximum shrink line length
+\item Integrate a shrink line forward and backward according to \cref{eq:shrink line} and using the largest $\lambda_2$ local maximum as the initial position.
+Integrate until the shrink line has attained the maximum shrink line length, or until it has reached the domain boundary.
+\item Flag any remaining local maxima of $\lambda_2$ within the maximization distance of the shrink line as ineligible initial positions for subsequent shrink lines
 \label{i:flag maxima}
-\item Continue integrating shrinklines using local maxima of $\lambda_2$ as initial positions until no eligible local maxima of $\lambda_2$ remain.
-\item Remove all shrinkline segments within elliptic LCSs.
+\item Continue integrating shrink lines using local maxima of $\lambda_2$ as initial positions until no eligible local maxima of $\lambda_2$ remain.
+\item Remove all shrink line segments within elliptic LCSs.
 \end{enumerate}
@@
 \hline
-\lstinline!nMaxCurves! & optional maximum number of curves (i.e. shrinklines of stretchlines) to generate. Default: \lstinline!numel(cgEigenvalue)!.\tabularnewline
+\lstinline!nMaxCurves! & optional maximum number of curves (i.e. shrink lines or stretch lines) to generate. Default: \lstinline!numel(cgEigenvalue)!.\tabularnewline
 \hline
@@
 \Cref{l:double gyre lambda strain stretch LCS} lists the corresponding LCS Tool commands.
-The maximal length of shrinklines and stretchlines, \lstinline!shrinklineMaxLength!
-and \lstinline!stretchlineMaxLength!, is set to $20$, a multiple of the domain size, since the hyperbolic LCSs may wind around vortices several times.
+The maximal length of shrink lines and stretch lines, \lstinline!shrinkLineMaxLength!
+and \lstinline!stretchLineMaxLength!, is set to $20$, a multiple of the domain size, since the hyperbolic LCSs may wind around vortices several times.
 \end{sloppypar}
 
-As usual, the local maximization distance is set larger for stretchlines than for shrinklines (cf.
-\cref{ll:shrinklineLocalMaxDistance} and \cref{ll:stretchlineLocalMaxDistance}).
+As usual, the local maximization distance is set larger for stretch lines than for shrink lines (cf.
+\cref{ll:shrinkLineLocalMaxDistance} and \cref{ll:stretchLineLocalMaxDistance}).
 The purpose of the maximization distance is to obtain spatially separated LCSs and to avoid a dense tangle of lines that basically indicates the same hyperbolic LCS (cf.
 \cref{t:Hyperbolic LCS algorithm}, \cref{i:flag maxima}).
-Setting the local maximization distance for stretchlines larger than for shrinklines allows obtaining a comparable number of stretchlines and shrinklines overall in the flow domain (recall that hyperbolic LCS seed points are discarded if they are within the local maximization distance of an existing hyperbolic LCS).
-Shrinklines are locally tangent to ridges of $\lambda_2$ maxima, whereas stretchlines are locally normal to these ridges.
-Setting the local maximization distance of stretchlines and shrinklines equal would therefore produce a greater number of stretchlines than shrinklines overall.
+Setting the local maximization distance for stretch lines larger than for shrink lines allows obtaining a comparable number of stretch lines and shrink lines overall in the flow domain (recall that hyperbolic LCS seed points are discarded if they are within the local maximization distance of an existing hyperbolic LCS).
+Shrink lines are locally tangent to ridges of $\lambda_2$ maxima, whereas stretch lines are locally normal to these ridges.
+Setting the local maximization distance of stretch lines and shrink lines equal would therefore produce a greater number of stretch lines than shrink lines overall.
 
@@
 Resolution is $[500\times250]$.
-Elliptic LCSs are green, shrinkline LCSs are red and stretchline LCSs are blue.
-White dots indicate $\lambda_2$ maxima for shrinklines and $\lambda_1$ minima stretchlines.
+Elliptic LCSs are green, shrink line LCSs are red and stretch line LCSs are blue.
+White dots indicate $\lambda_2$ maxima for shrink lines and $\lambda_1$ minima stretch lines.
 FTLE shown in the background.
@@
 \centering
-% FIXME Shrinkline LCS figure does not treat periodic boundary condition correctly
+% FIXME Shrink line LCS figure does not treat periodic boundary condition correctly
 \includegraphics[width=.475\textwidth]{graphics/bickley_jet/lambda_hyperbolic_lcs}
@@
 Resolution is $[500 \times 276]$.
-Elliptic LCSs are green, shrinkline LCSs are red and stretchline LCSs are blue. 
-White dots indicate $\lambda_2$ maxima for shrinklines and $\lambda_1$ minima stretchlines.
+Elliptic LCSs are green, shrink line LCSs are red and stretch line LCSs are blue. 
+White dots indicate $\lambda_2$ maxima for shrink lines and $\lambda_1$ minima stretch lines.
 FTLE shown in the background.
@@
 \begin{sloppypar}
-In the geostrophic approximation, the sea-surface height $\eta$ serves as a stream-function for surface velocities.
+In the geostrophic approximation, the sea surface height ($\eta$) serves as a stream-function for surface velocities.
 In a longitude-latitude $(\varphi,\theta)$ coordinate system, the evolution of a fluid particle is given by
@@
 $\lambda$-values are varied over a range of $[0.90,1.10]$ with a step of $0.02$ (\cref{ll:ocean data lambdaRange}).
-Hyperbolic LCSs are computed in \cref{ll:ocean data shrinklineLcs,ll:ocean data stretchlineLcs}.
+Hyperbolic LCSs are computed in \cref{ll:ocean data shrinkLineLcs,ll:ocean data stretchLineLcs}.
 
@@
 Resolution is $[500\times250]$.
-Elliptic LCSs are green, shrinkline LCSs are red and stretchline LCSs are blue. 
-White dots indicate $\lambda_2$ maxima for shrinklines and $\lambda_1$ minima stretchlines.
+Elliptic LCSs are green, shrink line LCSs are red and stretch line LCSs are blue. 
+White dots indicate $\lambda_2$ maxima for shrink lines and $\lambda_1$ minima stretch lines.
 FTLE shown in the background.
@@
 % End:
-% LocalWords:  stretchline stretchlines incompressible gyre eigenvector eigenvectors incompressibility dataset datasets LCS LCSs Lyapunov Bickley Rossby timespan variational adaptively vortices parametrise advected minima extremal FTLE MATLAB maxima geostrophic shrinkline shrinklines Poincare altimetry zonal
+% LocalWords:  incompressible gyre eigenvector eigenvectors incompressibility dataset datasets LCS LCSs Lyapunov Bickley Rossby timespan variational adaptively vortices parametrise advected minima extremal FTLE MATLAB maxima geostrophic Poincare altimetry zonal

### 2014-11-09 listings/bickley_jet/elliptic_hyperbolic_lcs_subset.m

@@ -27,15 +27,15 @@ periodicBc = [true,false]; %@\label{ll:Bickley jet periodicBc}
 lambdaRange = .8:.01:1.1;
 lambdaLineOdeSolverOptions = odeset('relTol',1e-6);
 
-% Shrinklines
-shrinklineMaxLength = 1e8;
-shrinklineLocalMaxDistance = 8*gridSpace;
-shrinklineOdeSolverOptions = odeset('relTol',1e-4);
+% Shrink lines
+shrinkLineMaxLength = 1e8;
+shrinkLineLocalMaxDistance = 8*gridSpace;
+shrinkLineOdeSolverOptions = odeset('relTol',1e-4);
 
-% Stretchlines
-stretchlineMaxLength = 1e8;
-stretchlineLocalMaxDistance = 4*gridSpace;
-stretchlineOdeSolverOptions = odeset('relTol',1e-4);
+% Stretch lines
+stretchLineMaxLength = 1e8;
+stretchLineLocalMaxDistance = 4*gridSpace;
+stretchLineOdeSolverOptions = odeset('relTol',1e-4);
 
 %% Lambda-line LCSs
 % Define Poincare sections; first point in center of elliptic region and
@@ -66,8 +66,8 @@ for lambda = lambdaRange
 ...    
 end
 ...   
-%% Hyperbolic shrinkline LCSs
-[shrinklineLcs,shrinklineLcsInitialPosition] = seed_curves_from_lambda_max(shrinklineLocalMaxDistance,shrinklineMaxLength,cgEigenvalue(:,2),cgEigenvector(:,1:2),domain,resolution,'periodicBc',periodicBc,'odeSolverOptions',shrinklineOdeSolverOptions);
+%% Hyperbolic shrink line LCSs
+[shrinkLineLcs,shrinkLineLcsInitialPosition] = seed_curves_from_lambda_max(shrinkLineLocalMaxDistance,shrinkLineMaxLength,cgEigenvalue(:,2),cgEigenvector(:,1:2),domain,resolution,'periodicBc',periodicBc,'odeSolverOptions',shrinkLineOdeSolverOptions);
 
-%% Hyperbolic stretchline LCSs
[…]

### 2014-11-09 listings/double_gyre/hyperbolic_lcs_subset.m

@@ -1,17 +1,17 @@
-% Shrinklines
-shrinklineMaxLength = 20;
+% Shrink lines
+shrinkLineMaxLength = 20;
 gridSpace = diff(domain(1,:))/(double(resolution(1))-1);
-shrinklineLocalMaxDistance = 2*gridSpace; %@\label{ll:shrinklineLocalMaxDistance}
-shrinklineOdeSolverOptions = odeset('relTol',1e-6);
+shrinkLineLocalMaxDistance = 2*gridSpace; %@\label{ll:shrinkLineLocalMaxDistance}
+shrinkLineOdeSolverOptions = odeset('relTol',1e-6);
 
-% Stretchlines
-stretchlineMaxLength = 20;
-stretchlineLocalMaxDistance = 10*gridSpace; %@\label{ll:stretchlineLocalMaxDistance}
-stretchlineOdeSolverOptions = odeset('relTol',1e-6);
+% Stretch lines
+stretchLineMaxLength = 20;
+stretchLineLocalMaxDistance = 10*gridSpace; %@\label{ll:stretchLineLocalMaxDistance}
+stretchLineOdeSolverOptions = odeset('relTol',1e-6);
 
-%% Hyperbolic shrinkline LCSs
-[shrinklineLcs,shrinklineLcsInitialPosition] = seed_curves_from_lambda_max(shrinklineLocalMaxDistance,shrinklineMaxLength,cgEigenvalue(:,2),cgEigenvector(:,1:2),domain,resolution,'odeSolverOptions',shrinklineOdeSolverOptions);
+%% Hyperbolic shrink line LCSs
+[shrinkLineLcs,shrinkLineLcsInitialPosition] = seed_curves_from_lambda_max(shrinkLineLocalMaxDistance,shrinkLineMaxLength,cgEigenvalue(:,2),cgEigenvector(:,1:2),domain,resolution,'odeSolverOptions',shrinkLineOdeSolverOptions);
 ...
-%% Hyperbolic stretchline LCSs
-[stretchlineLcs,stretchlineLcsInitialPosition] = seed_curves_from_lambda_max(stretchlineLocalMaxDistance,stretchlineMaxLength,-cgEigenvalue(:,1),cgEigenvector(:,3:4),domain,resolution,'odeSolverOptions',stretchlineOdeSolverOptions);
+%% Hyperbolic stretch line LCSs
[…]

### 2014-11-09 listings/ocean_dataset/elliptic_hyperbolic_lcs_subset.m

@@ -21,16 +21,16 @@ cgAuxGridRelDelta = .01; %@\label{ll:cgAuxGridRelDelta}
 lambdaRange = .9:.02:1.1; %@\label{ll:ocean data lambdaRange}
 lambdaLineOdeSolverOptions = odeset('relTol',1e-6);
 
-% Shrinklines
-shrinklineMaxLength = 20;
+% Shrink lines
+shrinkLineMaxLength = 20;
 gridSpace = diff(domain(1,:))/(double(resolution(1))-1);
-shrinklineLocalMaxDistance = 2*gridSpace;
-shrinklineOdeSolverOptions = odeset('relTol',1e-4);
+shrinkLineLocalMaxDistance = 2*gridSpace;
+shrinkLineOdeSolverOptions = odeset('relTol',1e-4);
 
-% Stretchlines
-stretchlineMaxLength = 20;
-stretchlineLocalMaxDistance = 4*gridSpace;
-stretchlineOdeSolverOptions = odeset('relTol',1e-4);
+% Stretch lines
+stretchLineMaxLength = 20;
+stretchLineLocalMaxDistance = 4*gridSpace;
+stretchLineOdeSolverOptions = odeset('relTol',1e-4);
 ...
 % Cauchy-Green strain eigenvalues and eigenvectors
 [cgEigenvector,cgEigenvalue] = eig_cgStrain(lDerivative,domain,resolution,timespan,'incompressible',incompressible,'eigenvalueFromMainGrid',cgEigenvalueFromMainGrid,'auxGridRelDelta',cgAuxGridRelDelta); %@\label{ll:eig_cgStrain}
@@ -59,8 +59,8 @@ for lambda = lambdaRange
 ...
 end
 ...
-% Hyperbolic shrinkline LCSs
-shrinklineLcs = seed_curves_from_lambda_max(shrinklineLocalMaxDistance,shrinklineMaxLength,cgEigenvalue(:,2),cgEigenvector(:,1:2),domain,resolution,'odeSolverOptions',shrinklineOdeSolverOptions); %@\label{ll:ocean data shrinklineLcs}
+% Hyperbolic shrink line LCSs
[…]

### 2014-11-09 main.tex
Revise grammar

@@
 \citet{haller01:_distin} proposed that ridges of the finite-time Lyapunov exponent (FTLE) are heuristic indicators of hyperbolic (i.e., repelling and attracting type) LCSs.
-A number of examples indeed support this principle\citep{peacock13:_lagran}.
+A number of examples support this principle\citep{peacock13:_lagran}.
 Equating FTLE ridges with LCSs, however, would create theoretical inconsistencies, as well as false positives and negatives in hyperbolic LCS detection\citep{haller11:_lagran_coher_struc,norgard12:_secon_lagran_coher_struc}.
@@
 \begin{sloppypar}
-Here we present an algorithmic introduction to geodesic LCS detection.
+We present an algorithmic introduction to geodesic LCS detection.
 % FIXME Reviewer #3, comment #4: Add documentation within the github repository to show some reference to the APIs and how users may start to run a his/her own application.
@@
 
-\begin{sloppypar}
 Other software for LCS detection exists.
@@
 Newman\citep{toit10:_trans} calculates the FTLE using dimension independent code.
-It assists ridge extraction of FTLE fields and supports analytic and dataset based velocity definitions.
+It assists ridge extraction of FTLE fields and supports analytic and dataset velocity definitions.
 FlowTK\citep{ameli14:_devel_effic_flexib_pipel_lagran} calculates the FTLE in two and three-dimensions on Cartesian and unstructured grids.
 The NVIDIA CUDA parallel computing platform is used for fast computations and Kitware's ParaView data analysis and visualization application is used for a user interface.
-\end{sloppypar}
 
-\begin{sloppypar}
-These packages target the automated generation of FTLE plots to aid the visual assessment of hyperbolic LCSs.
+These packages generate FTLE plots to aid the visual assessment of hyperbolic LCSs.
 LCS Tool also has functions to generate FTLE plots, but its emphasis is to provide geodesic extraction of LCSs as parametrized material curves, and extend the scope of such extraction to elliptic LCSs.
-\end{sloppypar}
 
@@
 
-\begin{sloppypar}
-The function \lstinline!eig_cgStrain!
-of LCS Tool provides the option to calculate Cauchy-Green eigenvectors from the auxiliary grid using eigenvalues calculated from the main grid.
-We have found that for flows defined analytically, the eigenvalues can reliably be calculated from the main grid.
-For flows defined by datasets, using the auxiliary grid for eigenvalue calculations gives better results.
-\end{sloppypar}
+The function \lstinline!eig_cgStrain! of LCS Tool provides the option to calculate Cauchy-Green eigenvectors from the auxiliary grid using eigenvalues calculated from the main grid.
+We have found that for analytic flows, the eigenvalues can be calculated from the main grid.
+For dataset flows, using the auxiliary grid for eigenvalue calculations gives better results.
 
-\begin{sloppypar}
-As mentioned above, a typical main grid for the Cauchy-Green strain tensor comprises $500 \times 500$ points.
+As stated above, a typical main grid for the Cauchy-Green strain tensor has $500 \times 500$ points.
 This means that after the addition of 4 auxiliary grid points around each main grid point, \cref{eq:velocity} must be integrated over 1.25 million initial conditions.
-\end{sloppypar}
 
@@
 
-\begin{sloppypar}
-The Cauchy-Green strain tensor generically admits point singularities, i.e., points where $C_{t_0}^t(x_0)$ has repeated eigenvalues.
+The Cauchy-Green strain tensor can have point singularities, i.e., points where $C_{t_0}^t(x_0)$ has repeated eigenvalues.
 At these points the eigenvectors $\xi_1(x_0)$ and $\xi_2(x_0)$ are no longer well-defined.
 This generically arises at a finite set of isolated points within the computational domain\citep{delmarcelle94}, and hence lie off the computational grid with probability one.
-\end{sloppypar}
 
@@
 
-\begin{sloppypar}
-As discussed in \cref{sec:Hyperbolic LCSs}, positions of hyperbolic LCSs at time $t_0$ are found as the strongest repelling orbits of the vector field \cref{eq:shrink line} (repelling LCSs), and strongest attracting orbits of the vector field \cref{eq:stretch line} (attracting LCSs).
+% FIXME Clarify use of the term "orbit"
+As discussed in \cref{sec:Hyperbolic LCSs}, hyperbolic LCS positions at time $t_0$ are found as the strongest repelling orbits of the vector field \cref{eq:shrink line} (repelling LCSs), and strongest attracting orbits of the vector field \cref{eq:stretch line} (attracting LCSs).
 By repulsion and attraction we mean a property of the LCS (as an evolving material line) under the flow map $F_{t_0}^t$.
@@
 These local maxima and minima of the appropriate $\lambda_i(x_0)$ eigenvalue field can be thought of as the extensions of the concept of saddle points to the present finite-time, temporally aperiodic flow setting.
-\end{sloppypar}
 
-\begin{sloppypar}
-The main steps of our hyperbolic LCS detection algorithm are enumerated in \cref{t:Hyperbolic LCS algorithm}.
-The basic function to compute hyperbolic LCSs in LCS Tool is \lstinline!seed_curves_from_lambda_max!
-and its syntax is given in \cref{t:seed_curves_from_lambda_max syntax}.
-\end{sloppypar}
+The main steps of hyperbolic LCS detection are enumerated in \cref{t:Hyperbolic LCS algorithm}.
+The core function to compute hyperbolic LCSs is \lstinline!seed_curves_from_lambda_max! and its syntax is given in \cref{t:seed_curves_from_lambda_max syntax}.
 
@@
 
-\begin{sloppypar}
-\Cref{f:double gyre lambda hyperbolic LCS} shows the complete picture of elliptic and hyperbolic LCSs of the double gyre at this resolution.
-\Cref{l:double gyre lambda strain stretch LCS} lists the corresponding LCS Tool commands.
-The maximal length of shrink lines and stretch lines, \lstinline!shrinkLineMaxLength!
+\Cref{f:double gyre lambda hyperbolic LCS} shows elliptic and hyperbolic LCSs for the double gyre at this resolution.
+% FIXME Sentence below is inaccurate since \cref{l:double gyre lambda strain stretch LCS} gives commands for hyperbolic LCSs only
+\Cref{l:double gyre lambda strain stretch LCS} gives the corresponding LCS Tool commands.
+The maximum length of shrink lines and stretch lines, \lstinline!shrinkLineMaxLength!
 and \lstinline!stretchLineMaxLength!, is set to $20$, a multiple of the domain size, since the hyperbolic LCSs may wind around vortices several times.
-\end{sloppypar}
 
@@
 
-\begin{sloppypar}
-In the geostrophic approximation, the sea surface height ($\eta$) serves as a stream-function for surface velocities.
+In the geostrophic approximation, sea surface height ($\eta$) serves as a stream-function for surface velocities.
 In a longitude-latitude $(\varphi,\theta)$ coordinate system, the evolution of a fluid particle is given by
@@
 where $g$ is the constant of gravity, $R$ is the mean radius of the Earth, and $f(\theta) \equiv 2\Omega\sin\theta$ is the Coriolis parameter, with $\Omega$ denoting the Earth's mean angular velocity.
-\end{sloppypar}
 

### 2014-11-12 main.tex
Reduce size of figure illustrating main & auxiliary grids
Add code to generate convergence figure

@@
 \usepackage{microtype}
+\usepackage{pgfplots}
 \usepackage[hidelinks]{hyperref}
@@
 \makeatother
+
+\pgfplotsset{compat=newest}
+\pgfplotsset{plot coordinates/math parser=false}
+\newlength\figureheight
+\newlength\figurewidth
 
@@
 \begin{center}
-\includegraphics[width=.475\textwidth]{graphics/main_aux_grids}
+\includegraphics[width=.325\textwidth]{graphics/main_aux_grids}
 \end{center}
@@
 \centering
-\includegraphics[width=.475\textwidth]{graphics/double_gyre/lambda_lcs_convergence}\caption{Convergence of closed $\lambda$-lines for increasing main-grid resolution in the double-gyre model.
+\setlength\figurewidth{.4\textwidth}
+\input{graphics/double_gyre/elliptic_lcs_convergence_500.tikz}
+\input{graphics/double_gyre/elliptic_lcs_convergence_750.tikz}
+\input{graphics/double_gyre/elliptic_lcs_convergence_1000.tikz}
+\caption{Convergence of closed $\lambda$-lines for increasing main-grid resolution in the double-gyre.
 The outermost closed $\lambda$-line (bold green line) is the vortex boundary.
-(Top) {[}500x250{]}, $\lambda$-values
-of closed orbits from interior to exterior, left gyre: $\lambda=(0.97, 0.98, 0.99, 1.00, 1.01)$, right gyre: $\lambda=(1.00, 1.01, 1.02, 1.03, 1.04)$; (Middle) {[}750x375{]}, left: $\lambda=(0.97, 0.98, 0.99, 1.00, 1.01)$, right: $\lambda=(1.00, 1.01, 1.02, 1.03, 1.04, 1.05)$; (Bottom) {[}1000x500{]}, left: $\lambda=(0.97, 0.98, 0.99, 1.00, 1.01)$, right: $\lambda=(1.00, 1.01, 1.02, 1.03, 1.04, 1.05)$.}
+Top: $500 \times 250$, middle: $750 \times 375$, bottom: $1000 \times 500$.}
 \label{f:double gyre lambda LCS convergence}

### 2014-11-12 graphics/bickley_jet/startup.m

@@ -1,0 +1,2 @@
+addpath(fullfile('..','..','LCS-Tool','demo','bickley_jet'))
+addpath(fullfile('..','..','LCS-Tool'))

### 2014-11-12 graphics/double_gyre/elliptic_lcs_convergence.m

@@ -1,0 +1,77 @@
+function elliptic_lcs_convergence
+
+% Parameters
+epsilon = .1;
+amplitude = .1;
+omega = pi/5;
+domain = [0,2;0,1];
+resolutionX = 500:250:1000;
+timespan = [0,10];
+
+%% Velocity definition
+lDerivative = @(t,x,~)derivative(t,x,false,epsilon,amplitude,omega);
+incompressible = true;
+
+%% LCS parameters
+% Cauchy-Green strain
+cgStrainOdeSolverOptions = odeset('relTol',1e-5);
+
+% Lambda-lines
+poincareSection = struct('endPosition',{},'numPoints',{},'orbitMaxLength',{});
+poincareSection(1).endPosition = [.55,.55;.1,.1];
+poincareSection(2).endPosition = [1.53,.45;1.95,.05];
+[poincareSection.numPoints] = deal(100);
+nPoincareSection = numel(poincareSection);
+for i = 1:nPoincareSection
+    rOrbit = hypot(diff(poincareSection(i).endPosition(:,1)),diff(poincareSection(i).endPosition(:,2)));
+    poincareSection(i).orbitMaxLength = 2*(2*pi*rOrbit);
+end
+lambda = .93:.01:1.07;
+lambdaLineOdeSolverOptions = odeset('relTol',1e-6);
+forceEtaComplexNaN = true;
+
+% Graphics properties
+ellipticColor = [0,.6,0];
+
+for m = 1:numel(resolutionX)
+    % Make x and y grid spacing as equal as possible
+    lResolutionX = resolutionX(m);
+    resolutionY = equal_resolution(domain,resolutionX);
+    resolution = [lResolutionX,resolutionY];
+    
+    hAxes = setup_figure(domain);
+    hFigure = get(hAxes,'parent');
+    delete(get(hAxes,'xlabel'))
+    delete(get(hAxes,'ylabel'))
+    
+    % Cauchy-Green strain eigenvalues and eigenvectors
+    warnState = warning('off','eig_cgStrain:unequalDelta');
+    [cgEigenvector,cgEigenvalue] = eig_cgStrain(lDerivative,domain,resolution,timespan,'incompressible',incompressible,'odeSolverOptions',cgStrainOdeSolverOptions);
+    warning(warnState)
+    
+    %% Elliptic LCSs
+    warnState = warning('off','integrate_line:isDiscontinuousLargeAngle');
[…]

### 2014-11-12 graphics/double_gyre/startup.m

@@ -1,0 +1,3 @@
+addpath(fullfile('..','..','LCS-Tool','demo','double_gyre'))
+addpath(fullfile('..','..','LCS-Tool'))
+addpath(fullfile('..','matlab2tikz','src'))

### 2014-11-12 main.tex
Correct typographical error

@@
 Circle markers indicate closed orbit positions.
-The filled circle indicates the outermost fixed point of the Poincare map, marking the intersection of a coherent Lagrangian cortex boundary with the Poincare section.}
+The filled circle indicates the outermost fixed point of the Poincare map, marking the intersection of a coherent Lagrangian vortex boundary with the Poincare section.}
 \label{f:Poincare return map}

### 2014-11-12 graphics/double_gyre/elliptic_lcs_convergence.m
Add MATLAB program to generate Poincare return map figure

@@ -73,5 +73,5 @@ for m = 1:numel(resolutionX)
     drawnow
 
     filename = strcat('elliptic_lcs_convergence_',num2str(lResolutionX),'.tikz');
-    matlab2tikz(filename,'showInfo',false,'relativeDataPath',fullfile('graphics','double_gyre'),'width','\figurewidth','figurehandle',hFigure)
+    matlab2tikz(filename,'showInfo',false,'width','\figurewidth','figurehandle',hFigure)
 end

### 2014-11-12 graphics/double_gyre/poincare_return_map.m

@@ -1,0 +1,43 @@
+function poincare_return_map
+
+%% Input parameters
+epsilon = .1;
+amplitude = .1;
+omega = pi/5;
+domain = [0,2;0,1];
+resolution = [750,375];
+timespan = [0,5];
+
+%% Velocity definition
+lDerivative = @(t,x,~)derivative(t,x,false,epsilon,amplitude,omega);
+incompressible = true;
+
+%% LCS parameters
+cgStrainOdeSolverOptions = odeset('relTol',1e-5);
+
+% Lambda-lines
+poincareSection.endPosition = [.55,.55;.2,.5];
+poincareSection.numPoints = 100;
+rOrbit = hypot(diff(poincareSection.endPosition(:,1)),diff(poincareSection.endPosition(:,2)));
+poincareSection.integrationLength = [0,2*(2*pi*rOrbit)];
+
+lambda = 1;
+lambdaLineOdeSolverOptions = odeset('relTol',1e-6);
+
+%% Cauchy-Green strain eigenvalues and eigenvectors
+[cgEigenvector,cgEigenvalue] = eig_cgStrain(lDerivative,domain,resolution,timespan,'incompressible',incompressible,'odeSolverOptions',cgStrainOdeSolverOptions);
+
+etaPos = lambda_line(cgEigenvector,cgEigenvalue,lambda);
+s = warning('off','integrate_line:isDiscontinuousLargeAngle');
+[~,~,hPoincareMap] = poincare_closed_orbit(domain,resolution,etaPos,poincareSection,'odeSolverOptions',lambdaLineOdeSolverOptions,'showGraph',true);
+warning(s)
+
+delete(findobj(hPoincareMap,'type','axes','Tag','legend'))
+hAxes = findobj(hPoincareMap,'type','axes');
+delete(get(hAxes,'title'))
+set(hAxes,'ylim',[-1,1]*1e-4)
+% Change colors to black
+set(findobj(hAxes,'Marker','none'),'color',[0,0,0])
+set(findobj(hAxes,'Marker','o'),'MarkerEdgeColor',[0,0,0])
+set(findobj(hAxes,'Marker','o','MarkerFaceColor',[1,0,0]),'MarkerFaceColor',[0,0,0])
+set(findobj(hAxes,'Marker','o','MarkerFaceColor',[0,0,1]),'MarkerFaceColor','none')

### 2014-11-12 main.tex
Add MATLAB program to generate Poincare return map figure
Add MATLAB program to make double gyre LCSs figures

@@
 \begin{center}
-\includegraphics[width=.475\textwidth]{graphics/double_gyre/poincare_return_map}
+\setlength\figurewidth{.4\textwidth}
+\input{graphics/double_gyre/poincare_return_map.tikz}
 \end{center}
@@
 % FIXME Reviewer #2, comment #6: Improve appearance of $lambda$ extrema (white dots)
-\includegraphics[width=.475\textwidth]{graphics/double_gyre/lambda_hyperbolic_lcs}
+\setlength\figurewidth{.35\textwidth}
+\input{graphics/double_gyre/elliptic_repelling_lcs_details.tikz}
+\input{graphics/double_gyre/elliptic_attracting_lcs_details.tikz}
 \caption{

### 2014-11-12 graphics/double_gyre/elliptic_hyperbolic_lcs_details.m

@@ -1,0 +1,162 @@
+function elliptic_hyperbolic_lcs_details
+
+% Parameters
+epsilon = .1;
+amplitude = .1;
+omega = pi/5;
+timespan = [0,10];
+domain = [0,2;0,1];
+resolutionX = 500;
+resolutionY = equal_resolution(domain,resolutionX);
+resolution = [resolutionX,resolutionY];
+
+%% Velocity definition
+lDerivative = @(t,x,~)derivative(t,x,false,epsilon,amplitude,omega);
+incompressible = true;
+
+%% LCS parameters
+% Cauchy-Green strain
+cgStrainOdeSolverOptions = odeset('relTol',1e-5);
+
+% Lambda-lines
+poincareSection = struct('endPosition',{},'numPoints',{},'orbitMaxLength',{});
+poincareSection(1).endPosition = [.55,.55;.1,.1];
+poincareSection(2).endPosition = [1.53,.45;1.95,.05];
+[poincareSection.numPoints] = deal(100);
+nPoincareSection = numel(poincareSection);
+for i = 1:nPoincareSection
+    rOrbit = hypot(diff(poincareSection(i).endPosition(:,1)),diff(poincareSection(i).endPosition(:,2)));
+    poincareSection(i).orbitMaxLength = 2*(2*pi*rOrbit);
+end
+lambda = .93:.01:1.07;
+lambdaLineOdeSolverOptions = odeset('relTol',1e-6);
+forceEtaComplexNaN = true;
+
+% Strainlines
+strainlineMaxLength = 20;
+gridSpace = diff(domain(1,:))/(double(resolution(1))-1);
+strainlineLocalMaxDistance = 2*gridSpace;
+strainlineOdeSolverOptions = odeset('relTol',1e-6);
+
+% Stretchlines
+stretchlineMaxLength = 20;
+stretchlineLocalMaxDistance = 10*gridSpace;
+stretchlineOdeSolverOptions = odeset('relTol',1e-6);
+
+% Graphics properties
+repellingColor = 'r';
+attractingColor = 'b';
+ellipticColor = [0,.6,0];
+lcsInitialPositionMarkerSize = 2;
+
+hAxes = setup_figure(domain);
+delete(get(hAxes,'xlabel'))
+delete(get(hAxes,'ylabel'))
+hFigure = get(hAxes,'parent');
+hColorbar = findobj(hFigure,'Type','ColorBar');
[…]

### 2014-11-13 graphics/bickley_jet/elliptic_hyperbolic_lcs_details.m
Improve the apperance of figures by using native LaTeX fonts

@@ -1,0 +1,167 @@
+function elliptic_hyperbolic_lcs_details
+
+%% Input parameters
+u = 62.66;
+lengthX = pi*earthRadius;
+lengthY = 1.77e6;
+epsilon = [.075,.4,.3];
+domain = [0,lengthX;[-1,1]*2.25*lengthY];
+timespan = [0,2*lengthX/u];
+resolutionX = 500;
+[resolutionY,deltaX] = equal_resolution(domain,resolutionX);
+resolution = [resolutionX,resolutionY];
+
+%% Velocity definition
+perturbationCase = 3;
+phiTimespan = [0,25];
+phiInitial = [0,0];
+phiSol = ode45(@d_phi,phiTimespan,phiInitial);
+timeResolution = 1e5;
+phi1 = deval(phiSol,linspace(phiTimespan(1),phiTimespan(2),timeResolution),1);
+phi1Max = max(phi1);
+lDerivative = @(t,x,~)derivative(t,x,false,u,lengthX,lengthY,epsilon,perturbationCase,phiSol,phi1Max);
+incompressible = true;
+periodicBc = [true,false];
+
+%% LCS parameters
+
+% Lambda-lines
+poincareSection = struct('endPosition',{},'numPoints',{},'orbitMaxLength',{});
+poincareSection(1).endPosition = [3.25e6,1.5e6;1.4e6,2.6e6];
+poincareSection(2).endPosition = [6.5e6,-1.4e6;5e6,-3e6];
+poincareSection(3).endPosition = [1e7,1.5e6;8e6,2.6e6];
+poincareSection(4).endPosition = [1.35e7,-1.4e6;1.5e7,-.5e6];
+poincareSection(5).endPosition = [1.65e7, 1.5e6;1.5e7, 2.6e6];
+[poincareSection.numPoints] = deal(80);
+nPoincareSection = numel(poincareSection);
+for i = 1:nPoincareSection
+    rOrbit = hypot(diff(poincareSection(i).endPosition(:,1)),diff(poincareSection(i).endPosition(:,2)));
+    poincareSection(i).orbitMaxLength = 2*(2*pi*rOrbit);
+end
+lambda = .9:.01:1.1;
+lambdaLineOdeSolverOptions = odeset('relTol',1e-6);
+forceEtaComplexNaN = true;
+
+% Strainlines
+strainlineMaxLength = 1e8;
+strainlineLocalMaxDistance = 8*deltaX;
+strainlineOdeSolverOptions = odeset('relTol',1e-4);
+
[…]

### 2014-11-14 main.tex
Improve the apperance of figures by using native LaTeX fonts
Complete previous commit (8747902d6cb49191bf2e26d5647c22d896b4581c)
Correct ocean data figure caption
Improve grammar

@@
 \usepackage{gensymb}
+\usepackage{import}
 \usepackage[utf8]{inputenc}
@@
 \begin{center}
-\includegraphics[width=.325\textwidth]{graphics/main_aux_grids}
+\def\svgwidth{.325\textwidth}
+\import{graphics/}{main_aux_grids.pdf_tex}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=.475\textwidth]{graphics/variable_step_integration}
+\def\svgwidth{.4\textwidth}
+\import{graphics/}{variable_step_integration.pdf_tex}
 \end{center}
@@
 At the initial point, there is an orientation discontinuity at the lower-right grid point that must be corrected prior to linear interpolation.
-At point \#1, no orientation discontinuities are present.
-At point \#2 all interpolated $\eta_\pm^\lambda$ vectors must be rotated by 180° to match the orientation of the trajectory.}
+At point \#1, no orientation discontinuity is present.
+At point \#2, $\eta_\pm^\lambda$ vectors must be rotated by 180° to match the orientation of the trajectory.}
 \label{f:variable step integration}
@@
 
-In classical, infinite-time dynamical systems, elliptic and hyperbolic invariant manifolds cannot intersect, as the trajectories in their intersection would then have to follow two different asymptotic behaviours simultaneously.
+In classical infinite-time dynamical systems, elliptic and hyperbolic invariant manifolds cannot intersect since trajectories at intersections would follow two different asymptotic behaviours simultaneously.
 In finite-time dynamical systems, hyperbolic LCSs may well intersect elliptic LCSs, with hyperbolic LCSs continuing to act as cores of mixing patterns observed in the interior of elliptic LCSs.
 Motivated by convention and for clarity in visualization, however, the default setting of LCS Tool removes hyperbolic LCS segments from the interior of elliptic LCSs.
-These removed segments can be turned back on demand, as shown in \cref{f:ocean_dataset_colortracer}.
+These removed segments can be turned on, as shown in \cref{f:ocean_dataset_colortracer}.
 
@@
 The maximum length of shrink lines and stretch lines, \lstinline!shrinkLineMaxLength!
-and \lstinline!stretchLineMaxLength!, is set to $20$, a multiple of the domain size, since the hyperbolic LCSs may wind around vortices several times.
+and \lstinline!stretchLineMaxLength!, is set to $20$, a multiple of the domain size since the hyperbolic LCSs may wind around vortices several times.
 
@@
 LCSs in the ocean velocity data from satellite altimetry.
-Resolution is $[500\times250]$.
+Resolution is $400 \times 400$.
 Elliptic LCSs are green, shrink line LCSs are red and stretch line LCSs are blue.
@@
 
-% FIXME Citations without pages (haller15:_langr_coher_struc and karrasch14:_autom_lagran) do not list publication year
+% FIXME Citations without pages (karrasch14:_autom_lagran) do not list publication year
 \bibliographystyle{elsarticle-num-names} 

### 2014-11-28 main.tex
Correct & update figure captions
Add benchmarks table

@@
 \multicolumn{2}{|p{.95\textwidth}|}{\lstinline![cgEigenvector,cgEigenvalue] = eig_cgStrain(derivative,domain,timespan,resolution)!}\tabularnewline
-\hline
+\hline \hline
 \lstinline!derivative! & function handle for flow velocity equations\tabularnewline
@@
 function to perform trajectory integration from all grid points simultaneously.
-This calculation typically takes five to ten minutes with default error tolerances.
-% FIXME Reviewer #2, comment #3: Add memory use data based on, for example, computations for \cref{f:double gyre lambda LCS convergence}.
+This calculation typically takes five to ten minutes.
 A potential drawback of vector form integration is that memory requirements may become excessive at high resolutions.
+\Cref{t:vector integration memory} lists runtime and memory use to integrate a flow map at different resolutions.
 Furthermore, writing the velocity function in vector form is more error-prone than the simpler two-dimensional form.
+
+\begin{table}
+\begin{center}
+\begin{tabular}{|c|c|c|}
+\hline
+Resolution & Runtime (s) & Memory (MB)\tabularnewline
+\hline \hline
+$500 \times 251$ & 29 & 55\tabularnewline
+\hline
+$750 \times 376$ & 63 & 123\tabularnewline
+\hline
+$1000 \times 501$ & 112 & 223\tabularnewline
+\hline
+\end{tabular}
+\end{center}
+\caption{Runtime and memory use associated with vector integration at different resolutions.
+Results are from the function \lstinline!eig_cgStrain!
+applied to the double gyre presented in \cref{sec:double gyre}.
+Results obtained with an Intel Core i3-4130 processor.}
+\label{t:vector integration memory}
+\end{table}
 
@@
 {\lstinline![shearline.etaPos,shearline.etaNeg] = lambda_line(cgEigenvector,cgEigenvalue,lambda)!}\tabularnewline
-\hline
+\hline \hline
 \lstinline!cgEigenvector! & array of Cauchy-Green strain eigenvectors\tabularnewline
@@
 \multicolumn{2}{|p{.95\textwidth}|}{\lstinline![closedOrbits,orbits] = poincare_closed_orbit_multi(domain,resolution,shearline,PSList)!}\tabularnewline
-\hline
+\hline \hline
 \lstinline!domain! & array to define flow domain\tabularnewline
@@
 {\lstinline![curvePosition,curveInitialPosition] = seed_curves_from_lambda_max(distance,cgEigenvalue,cgEigenvector,flowDomain,flowResolution)!}\tabularnewline
-\hline
+\hline \hline
 \lstinline!distance! & threshold distance for placement of $\lambda_2(x_0)$ maxima\tabularnewline
@@
 \subsection{Double gyre}
+\label{sec:double gyre}
+
 The double gyre is a model for a time-dependent two gyre system observed in geophysical flows \citep{shadden05:_defin_lagran_lyapun}.
@@
 LCSs in the double gyre.
-Resolution is $[500\times250]$.
+Resolution is $500 \times 250$.
 Elliptic LCSs are green, shrink line LCSs are red and stretch line LCSs are blue.
@@
 LCSs in the Bickley jet.
-Resolution is $[500 \times 276]$.
+Resolution is $500 \times 200$.
 Elliptic LCSs are green, shrink line LCSs are red and stretch line LCSs are blue.

### 2014-11-28 benchmarks/startup.m

@@ -1,0 +1,2 @@
+addpath(fullfile('..','LCS-Tool','demo','double_gyre'))
+addpath(fullfile('..','LCS-Tool'))

### 2014-11-28 benchmarks/vector_integration_memory.m

@@ -1,0 +1,31 @@
+% Use: "profile -memory on" to measure memory use
+% Reference: http://undocumentedmatlab.com/blog/undocumented-profiler-options/
+%
+% profile('-memory','on')
+% vector_integration_memory
+% stats = profile('info');
+% strcmp('vector_integration_memory',stats.FunctionTable(1).FunctionName)
+% stats.FunctionTable(1).TotalTime
+
+function vector_integration_memory
+
+% Parameters
+epsilon = .1;
+amplitude = .1;
+omega = pi/5;
+domain = [0,2;0,1];
+resolutionX = 1000;
+timespan = [0,10];
+
+%% Velocity definition
+lDerivative = @(t,x,~)derivative(t,x,false,epsilon,amplitude,omega);
+incompressible = true;
+
+%% LCS parameters
+% Cauchy-Green strain
+cgStrainOdeSolverOptions = odeset('relTol',1e-5);
+
+resolutionY = equal_resolution(domain,resolutionX);
+resolution = [resolutionX,resolutionY];
+    
+[~,~] = eig_cgStrain(lDerivative,domain,resolution,timespan,'incompressible',incompressible,'odeSolverOptions',cgStrainOdeSolverOptions);

### 2014-11-28 main.tex
Address reviewer comment
Simplify LaTeX maths mark-up

@@
 % FIXME Reviewer #3, comment #4: Add documentation within the github repository to show some reference to the APIs and how users may start to run a his/her own application.
-% FIXME Reviewer #4: Please make sure the code is downloadable once the paper is published
 We then review the implementation of this approach in a computational engine called LCS Tool\footnote{LCS Tool is available at: \url{http://www.runmycode.org/companion/view/908}}.
@@
 \[
-Q(\gamma) = \frac{1}{\sigma} \int_0^{\sigma} \frac{\sqrt{\langle r'(s),C_{t_0}^t(r(s))r'(s)\rangle}}{\sqrt{\langle r'(s),r'(s)\rangle}} \dif s,
+Q(\gamma) = \frac{1}{\sigma} \int_0^\sigma \frac{\sqrt{\langle r'(s),C_{t_0}^t(r(s))r'(s)\rangle}}{\sqrt{\langle r'(s),r'(s)\rangle}} \dif s,
 \]
@@
 To compute the normal repulsion of shrink lines, we note that an infinitesimal
-normal perturbation to a shrink line $\gamma$ at its point $r$ grows under the flow map $F_{t_{0}}^{t}$ by a factor of $\lambda_{2}(r)$ in the direction normal to $F_{t_{0}}^{t}(\gamma).$ Similarly, small normal perturbations to a stretch line decay by a factor $\lambda_{1}(r)$ in the direction normal to the evolving stretch line.
+normal perturbation to a shrink line $\gamma$ at its point $r$ grows under the flow map $F_{t_0}^t$ by a factor of $\lambda_2(r)$ in the direction normal to $F_{t_0}^t(\gamma)$.
+Similarly, small normal perturbations to a stretch line decay by a factor $\lambda_1(r)$ in the direction normal to the evolving stretch line.
 
@@
 
-% FIXME Reviewer #2, comment #7: sentence reads awkwardly
-The integration time is $T = 4L_x/U$, a multiple of the eddy turnover time.
+The integration time is $T = 4L_x/U$, an integer multiple of the eddy turnover time.
 \Cref{l:Bickley jet elliptic hyperbolic LCS subset} shows the LCS Tool commands in which the chaotically perturbed velocity is defined (\crefrange{ll:Bickley jet velocity definition start}{ll:Bickley jet velocity definition end}), periodic boundary conditions are imposed in the x-direction (\cref{ll:Bickley jet periodicBc}), and five Poincare sections are defined where we expect coherent vortices (\crefrange{ll:Bickley jet Poincare start}{ll:Bickley jet Poincare end}).
@@
 FTLE shown in the background.
-$\lambda=$ 1.00, 1.08, 0.94, 0.90, 1.06
+$\lambda = 1.00, 1.08, 0.94, 0.90, 1.06$.
 }

### 2014-11-28 main.tex
Revise grammar in ocean data section

@@
 
-\Cref{f:ocean dataset LCS} shows the positions of elliptic and hyperbolic LCSs on 22 November 2006, which is the same time instant analysed in \citet{haller13:_coher_lagran,haller14:_adden_coher_lagran,beron-vera13:_objec_agulh}.
-Here the integration time is chosen as $T = 30$ days, as opposed to 90 days in the cited references, to avoid tangling of hyperbolic LCSs.
-Our analysis via LCS Tool reveals five coherent eddy boundaries.
-The large coherent eddy at $(3,-32)$ has a non-stretching boundary, i.e., $\lambda = 1$.
-It corresponds to eddy \#2 in Figure~3 of \citet{beron-vera13:_objec_agulh}.
-Additionally, four smaller coherent eddy cores are found.
-They do not stay coherent over a time of 90 days.
-All closed orbits are stable under increased spatial resolution for the Cauchy-Green strain tensor field.
-Repelling LCSs (red) and attracting LCSs (blue) are superimposed over the plot of FTLE values.
-The hyperbolic LCSs determine the deformation of the fluid in-between coherent eddy cores.
+\Cref{f:ocean dataset LCS} shows elliptic and hyperbolic LCSs on 22 November 2006, the same time as analysed in \citet{haller13:_coher_lagran,haller14:_adden_coher_lagran,beron-vera13:_objec_agulh}.
+In those references, the integration time is 90 days.
+We use a shorter time to avoid tangling hyperbolic LCSs.
+Our analysis via LCS Tool reveals five coherent eddies.
+The largest, at $(3,-32)$, has a non-stretching boundary, i.e., $\lambda = 1$.
+It corresponds to eddy \#2 in figure~3 of \citet{beron-vera13:_objec_agulh}.
+Four additional smaller coherent eddies are found.
+They do not stay coherent over 90 days.
+The hyperbolic LCSs determine the deformation of the fluid between coherent eddies.
 

### 2014-11-28 main.tex
Update FIXME comments
Update Bickley jet and ocean data figures

@@
 These functions are written to achieve balance between ease of use and performance.
-% FIXME List demo script run-times in minutes in a table
+% FIXME With lambda range running demos can take over an hour.
+% FIXME List runtime of demo scripts in a table
 Running the demo scripts takes on the order of minutes.
@@
 \centering
-% FIXME Shrink line LCS figure does not treat periodic boundary condition correctly
-\includegraphics[width=.475\textwidth]{graphics/bickley_jet/lambda_hyperbolic_lcs}
+\setlength\figurewidth{.35\textwidth}
+\import{graphics/bickley_jet/}{elliptic_repelling_lcs_details.tikz}
+\import{graphics/bickley_jet/}{elliptic_attracting_lcs_details.tikz}
 \caption{
@@
 \begin{center}
-\includegraphics[width=.475\textwidth]{graphics/ocean_dataset/lambda_hyperbolic_lcs}
+\setlength\figurewidth{.4\textwidth}
+\import{graphics/ocean_dataset/}{elliptic_repelling_lcs.tikz}
+\import{graphics/ocean_dataset/}{elliptic_attracting_lcs.tikz}
 \end{center}

### 2014-11-28 graphics/bickley_jet/startup.m

@@ -1,2 +1,3 @@
 addpath(fullfile('..','..','LCS-Tool','demo','bickley_jet'))
 addpath(fullfile('..','..','LCS-Tool'))
+addpath(fullfile('..','matlab2tikz','src'))

### 2014-11-28 graphics/ocean_dataset/startup.m

@@ -1,0 +1,3 @@
+addpath(fullfile('..','..','LCS-Tool','demo','ocean_dataset'))
+addpath(fullfile('..','..','LCS-Tool'))
+addpath(fullfile('..','matlab2tikz','src'))

### 2014-11-28 main.tex
Remove section on intersection of elliptic and hyperbolic LCSs
Code to reproduce figures was missing. Florian Huhn might provide it.

@@
 
-In classical infinite-time dynamical systems, elliptic and hyperbolic invariant manifolds cannot intersect since trajectories at intersections would follow two different asymptotic behaviours simultaneously.
-In finite-time dynamical systems, hyperbolic LCSs may well intersect elliptic LCSs, with hyperbolic LCSs continuing to act as cores of mixing patterns observed in the interior of elliptic LCSs.
-Motivated by convention and for clarity in visualization, however, the default setting of LCS Tool removes hyperbolic LCS segments from the interior of elliptic LCSs.
-These removed segments can be turned on, as shown in \cref{f:ocean_dataset_colortracer}.
-
-\begin{figure}
-\centering
-\includegraphics[width=.475\textwidth]{graphics/ocean_dataset/LCS_fwd_coherent_eddy}
-\includegraphics[width=.475\textwidth]{graphics/ocean_dataset/LCS_fwd_colortracer}
-\caption{(Top) Elliptic LCS (green) with repelling LCSs (red) continuing into its interior.
-(Bottom) Final positions $x(t) = (x_1(t),x_2(t)) = F_{t_0}^t(x_0)$ of tracer particles.
-$x_1(t)$ position encoded in red and $x_2(t)$ position encoded in green.
-Note the continued impact of the repelling hyperbolic LCSs (black) on tracer patterns inside the elliptic LCS (white).}
-\label{f:ocean_dataset_colortracer}
-\end{figure}
-
 \section{Examples}

### 2014-12-02 main.tex
Correct overfull warnings
Update demo script names

@@
 \begin{center}
-\setlength\figurewidth{.4\textwidth}
+\setlength\figurewidth{.35\textwidth}
 \input{graphics/double_gyre/poincare_return_map.tikz}
@@
 
-\lstinputlisting[caption={LCS Tool commands for double gyre elliptic LCSs. Subset of the LCS Tool script \lstinline!demo/double_gyre/hyperbolic_shear_lcs.m!},
+\lstinputlisting[caption={LCS Tool commands for double gyre elliptic LCSs. Subset of the LCS Tool script \lstinline!demo/double_gyre/elliptic_hyperbolic_lcs.m!},
                  label=l:double gyre elliptic LCS]{listings/double_gyre/elliptic_lcs_subset.m}
@@
 \lstinputlisting[
-        caption={LCS Tool commands for double gyre hyperbolic LCSs. Subset of the LCS Tool script \lstinline!demo/double_gyre/hyperbolic_shear_lcs.m!},
+        caption={LCS Tool commands for double gyre hyperbolic LCSs. Subset of the LCS Tool script \lstinline!demo/double_gyre/elliptic_hyperbolic_lcs.m!},
         label=l:double gyre lambda strain stretch LCS]
@@
 \lstinputlisting[
-        caption={LCS Tool commands for Bickley jet elliptic and hyperbolic LCSs. Subset of the LCS Tool script \lstinline!demo/bickley_jet/hyperbolic_shear_lcs.m!},
+        caption={LCS Tool commands for Bickley jet elliptic and hyperbolic LCSs. Subset of the LCS Tool script \lstinline!demo/bickley_jet/elliptic_hyperbolic_lcs.m!},
         label=l:Bickley jet elliptic hyperbolic LCS subset]
@@
 \centering
-\setlength\figurewidth{.35\textwidth}
+\setlength\figurewidth{.325\textwidth}
 \import{graphics/bickley_jet/}{elliptic_repelling_lcs_details.tikz}
@@
 An interpolant is generated first, then the function \lstinline!flowdata_derivative! evaluates the interpolants for the zonal and meridional velocity at the needed coordinates.
-\Cref{l:ocean data elliptic hyperbolic lcs} shows the relevant part of the code in LCS Tool's ocean demo file \lstinline!demo/ocean_dataset/hyperbolic_shear_lcs.m!.
+\Cref{l:ocean data elliptic hyperbolic lcs} shows the relevant part of the code in LCS Tool's ocean demo file \lstinline!demo/ocean_dataset/elliptic_hyperbolic_lcs.m!.
 The commands for the interpolation of the velocity data are given in \crefrange{ll:interpolation start}{ll:interpolation end}.
@@
 \lstinputlisting[
-        caption={LCS Tool commands for ocean data elliptic and hyperbolic LCSs. Subset of the LCS Tool script \lstinline!demo/ocean_dataset/hyperbolic_shear_lcs.m!},
+        caption={LCS Tool commands for ocean data elliptic and hyperbolic LCSs. Subset of the LCS Tool script \lstinline!demo/ocean_dataset/elliptic_hyperbolic_lcs.m!},
         label={l:ocean data elliptic hyperbolic lcs}]
@@
 
+\begin{sloppypar}
 \Cref{f:ocean dataset LCS} shows elliptic and hyperbolic LCSs on 22 November 2006, the same time as analysed in \citet{haller13:_coher_lagran,haller14:_adden_coher_lagran,beron-vera13:_objec_agulh}.
@@
 The hyperbolic LCSs determine the deformation of the fluid between coherent eddies.
+\end{sloppypar}
 
@@
 \begin{center}
-\setlength\figurewidth{.4\textwidth}
+\setlength\figurewidth{.375\textwidth}
 \import{graphics/ocean_dataset/}{elliptic_repelling_lcs.tikz}

### 2014-12-02 main.tex
Address reviewer #5, comment #1 and #2

@@
 \begin{abstract}
-% FIXME Reviewer #5, comment #1: Abstract and introduction the objective and contribution of the paper are not clear; several points need to be clarified further.
-% FIXME Reviewer #5, comment #2: For readers to quickly catch the contribution in this work, it would be better to highlight major difficulties and challenges, and original achievements to overcome them, in a clearer way in abstract and introduction.
 We give an algorithmic introduction to Lagrangian coherent structures (LCSs) using a newly developed computational engine, LCS Tool.
@@
 LCS Tool implements the latest geodesic theory of LCSs for two-dimensional flows, uncovering key transport barriers in unsteady flow velocity data as explicit solutions of differential equations.
+LCS Tool makes theoretical results accessible to the fluid mechanics community since implementing these results directly could be time consuming.
 After a review of the underlying theory, we explain the steps and numerical methods used by LCS Tool, and illustrate its capabilities on three unsteady fluid flow examples.

### 2014-12-03 main.tex
Address reviewer #3, comment #4
Update FIXME comments

@@
-\documentclass[5p]{elsarticle}
+\documentclass[draft]{article}
 
 \usepackage{amsmath}
+\usepackage[uniquename=init,firstinits,backend=biber,sorting=none]{biblatex}
 \usepackage{commath}
+\usepackage{fontspec}
 \usepackage{gensymb}
 \usepackage{import}
-\usepackage[utf8]{inputenc}
-\usepackage{listings}
+\usepackage[final]{listings}
 \usepackage[numbered,framed]{mcode}
@@
 
-% Workaround for bug in elsarticle-num-names.bst
-% Reference: http://www.tug.org/pipermail/tex-live/2012-August/032152.html
-\makeatletter
-\providecommand{\doi}[1]{%
-  \begingroup
-    \let\bibinfo\@secondoftwo
-    \urlstyle{rm}%
-    \href{http://dx.doi.org/#1}{%
-      doi:\discretionary{}{}{}%
-      \nolinkurl{#1}%
-    }%
-  \endgroup
-}
-\makeatother
-
 \pgfplotsset{compat=newest}
@@
 
-\journal{Journal of Computational Science}
+\addbibresource{main.bib}
+
+\title{LCS Tool: A Computational Platform for Lagrangian Coherent Structures}
+
+\author{Kristjan Onu \and Florian Huhn \and George Haller}
 
@@
 
-\begin{frontmatter}
+\numberwithin{equation}{section}
+\numberwithin{figure}{section}
+\numberwithin{lstlisting}{section}
+\numberwithin{table}{section}
 
-\title{LCS Tool: A Computational Platform for Lagrangian Coherent Structures}
-
-\author{K.~Onu}
-
-\author{F.~Huhn}
-
-\author{G.~Haller\corref{cor}}
-\ead{georgehaller@ethz.ch}
-
-\cortext[cor]{Corresponding author}
-
-\address{Institute of Mechanical Systems, ETH Zurich, Switzerland}
+\maketitle
 
@@
 LCSs are most repelling, attracting and shearing material lines that form the centrepieces of observed tracer patterns in two-dimensional unsteady dynamical systems.
-LCS Tool implements the latest geodesic theory of LCSs for two-dimensional flows, uncovering key transport barriers in unsteady flow velocity data as explicit solutions of differential equations.
+LCS Tool implements the latest geodesic theory for LCSs in two-dimensional flows, uncovering key transport barriers in unsteady flow velocity data as explicit solutions of differential equations.
 LCS Tool makes theoretical results accessible to the fluid mechanics community since implementing these results directly could be time consuming.
-After a review of the underlying theory, we explain the steps and numerical methods used by LCS Tool, and illustrate its capabilities on three unsteady fluid flow examples.
+After a review of the underlying theory, we explain the numerical algorithms used by LCS Tool, and illustrate its capabilities on three unsteady fluid flow examples: a double gyre, a Bickley jet and an ocean dataset for the Agulhas current.
 \end{abstract}
 
-\begin{keyword}
-Lagrangian coherent structures \sep Mathematical software \sep Non-autonomous dynamical systems \sep Invariant manifolds \sep Mixing \sep Transport barriers \sep Fluid dynamics \sep Ocean surface flows
-
-\PACS 47.10.Fg \sep 47.27.De \sep 47.11.-j \sep 05.45.-a \sep 47.51.+a \sep 47.54.-r \sep 92.10.hf \sep 93.30.Qn \sep 92.10.ab \sep 92.10.Ty \sep 92.10.ak \sep 92.10.A-
-
-\MSC[2010] 37N10 \sep 37B25 \sep 97N80 \sep 37C60 \sep 37B55 \sep 70H33 \sep 76F25
-\end{keyword}
-
-\end{frontmatter}
+\tableofcontents
 
@@
 
+Lagrangian Coherent Structures (LCSs) are evolving organizing centres of trajectory patterns in non-autonomous dynamical systems\parencite{haller00:_lagran,peacock13:_lagran,haller15:_langr_coher_struc}.
+Applications of LCSs include oceanic and atmospheric flows\parencite{beron-vera13:_objec_agulh,koh02:_hyper}, biological transport problems\parencite{wilson09:_lagran_reynol,tallapragada11:_lagran,huhn12:_south_indian_ocean_count_madag}, aeronautics\parencite{tang10:_accur_lagran_hong_kong_inter_airpor}, celestial mechanics\parencite{gawlik09:_lagran}, crowd dynamics\parencite{ali07:_lagran_partic_dynam_approac_crowd}, and aperiodically forced mechanical oscillators\parencite{hadjighasem13:_detec_kam}.
+
 \begin{sloppypar}
-% FIXME Reviewer #5, comment #5: Please provide information regarding practical applications of this algorithmic software tool.
-Lagrangian Coherent Structures (LCSs) are evolving organizing centres of trajectory patterns in non-autonomous dynamical systems\citep{haller00:_lagran,peacock13:_lagran,haller15:_langr_coher_struc}.
-Applications of LCSs include oceanic and atmospheric flows\citep{beron-vera13:_objec_agulh,koh02:_hyper}, biological transport problems\citep{wilson09:_lagran_reynol,tallapragada11:_lagran,huhn12:_south_indian_ocean_count_madag}, aeronautics\citep{tang10:_accur_lagran_hong_kong_inter_airpor}, celestial mechanics\citep{gawlik09:_lagran}, crowd dynamics\citep{ali07:_lagran_partic_dynam_approac_crowd}, and aperiodically forced mechanical oscillators\citep{hadjighasem13:_detec_kam}.
+\textcite{haller01:_distin} proposed that ridges of the finite-time Lyapunov exponent (FTLE) are heuristic indicators of hyperbolic (i.e., repelling and attracting type) LCSs.
+A number of examples support this principle\parencite{peacock13:_lagran}.
+Equating FTLE ridges with LCSs, however, would create theoretical inconsistencies, as well as false positives and negatives in hyperbolic LCS detection\parencite{haller11:_lagran_coher_struc,norgard12:_secon_lagran_coher_struc}.
+In addition, the role of the FTLE field in the accurate detection of elliptic (vortex-type) and parabolic (jet-core type) LCSs has remained an open question (but see \textcite{beron-vera10:_invar_lagran}).
 \end{sloppypar}
 
-\citet{haller01:_distin} proposed that ridges of the finite-time Lyapunov exponent (FTLE) are heuristic indicators of hyperbolic (i.e., repelling and attracting type) LCSs.
-A number of examples support this principle\citep{peacock13:_lagran}.
-Equating FTLE ridges with LCSs, however, would create theoretical inconsistencies, as well as false positives and negatives in hyperbolic LCS detection\citep{haller11:_lagran_coher_struc,norgard12:_secon_lagran_coher_struc}.
-In addition, the role of the FTLE field in the accurate detection of elliptic (vortex-type) and parabolic (jet-core type) LCSs has remained an open question (but see \citet{beron-vera10:_invar_lagran}).
-
-More recent work has focused on an exact mathematical formulation of the properties defining LCSs\citep{haller11:_lagran_coher_struc,farazmand12:_comput_lagran,haller12:_geodes_theor_trans_barrier_two_dimen_flows,haller13:_coher_lagran,haller14:_adden_coher_lagran,farazmand13:_attrac_lagran,blazevski14:_hyper_ellip_trans_barrier_three}.
-In two-dimensional flows, hyperbolic and parabolic LCSs turn out to be stationary curves of the averaged material shear\citep{farazmand14:_shearless}, whereas elliptic LCSs are stationary curves of the averaged strain\citep{haller13:_coher_lagran,haller14:_adden_coher_lagran}.
+More recent work has focused on an exact mathematical formulation of the properties defining LCSs\parencite{haller11:_lagran_coher_struc,farazmand12:_comput_lagran,haller12:_geodes_theor_trans_barrier_two_dimen_flows,haller13:_coher_lagran,haller14:_adden_coher_lagran,farazmand13:_attrac_lagran,blazevski14:_hyper_ellip_trans_barrier_three}.
+In two-dimensional flows, hyperbolic and parabolic LCSs turn out to be stationary curves of the averaged material shear\parencite{farazmand14:_shearless}, whereas elliptic LCSs are stationary curves of the averaged strain\parencite{haller13:_coher_lagran,haller14:_adden_coher_lagran}.
 These variational formulations lead to explicit solutions for LCSs as null geodesics of appropriate Lorentzian metrics.
 
-\begin{sloppypar}
 We present an algorithmic introduction to geodesic LCS detection.
-% FIXME Reviewer #3, comment #4: Add documentation within the github repository to show some reference to the APIs and how users may start to run a his/her own application.
-We then review the implementation of this approach in a computational engine called LCS Tool\footnote{LCS Tool is available at: \url{http://www.runmycode.org/companion/view/908}}.
+We then review the implementation of this approach in a computational engine called LCS Tool\footnote{LCS Tool is available at: \url{http://github.com/jeixav/LCS-Tool}}.
 This engine is a library of MATLAB functions that extract LCSs from two-dimensional unsteady flows. The examples we present form demonstration scripts distributed with LCS Tool.
-\end{sloppypar}
 
 Other software for LCS detection exists.
-ManGen\citep{lekien03:_time} calculates the FTLE and advects material curves in two-dimensional velocity fields.
+ManGen\parencite{lekien03:_time} calculates the FTLE and advects material curves in two-dimensional velocity fields.
 It includes a graphical user interface and uses the Message Passing Interface standard for parallel calculations.
-Newman\citep{toit10:_trans} calculates the FTLE using dimension independent code.
+Newman\parencite{toit10:_trans} calculates the FTLE using dimension independent code.
 It assists ridge extraction of FTLE fields and supports analytic and dataset velocity definitions.
-FlowTK\citep{ameli14:_devel_effic_flexib_pipel_lagran} calculates the FTLE in two and three-dimensions on Cartesian and unstructured grids.
+FlowTK\parencite{ameli14:_devel_effic_flexib_pipel_lagran} calculates the FTLE in two and three-dimensions on Cartesian and unstructured grids.
 The NVIDIA CUDA parallel computing platform is used for fast computations and Kitware's ParaView data analysis and visualization application is used for a user interface.
@@
 \begin{equation}
-C_{t_0}^t(x_0) = \left[\nabla F_{t_0}^t(x_0)\right]^T \nabla F_{t_0}^t(x_0),
+C_{t_0}^t(x_0) \equiv \left[\nabla F_{t_0}^t(x_0)\right]^T \nabla F_{t_0}^t(x_0),
 \label{eq:CG}
 \end{equation}
-measuring Lagrangian strain in the velocity field. This tensor is symmetric and positive definite\citep{truesdell04}. We label the eigenvalues and eigenvectors of $C_{t_0}^t(x_0)$ as follows:
+measuring Lagrangian strain in the velocity field. This tensor is symmetric and positive definite\parencite{truesdell04}. We label the eigenvalues and eigenvectors of $C_{t_0}^t(x_0)$ as follows:
 \[
@@
 We seek positions of closed material lines at time $t_0$ that prevail as coherent Lagrangian vortex boundaries (or \emph{elliptic LCSs}) over a time interval $[t_0,t]\subset[t_-,t_+]$.
-\citet{haller13:_coher_lagran,haller14:_adden_coher_lagran} argue that such initial material line positions are closed stationary curves of the averaged strain functional
+\textcite{haller13:_coher_lagran,haller14:_adden_coher_lagran} argue that such initial material line positions are closed stationary curves of the averaged strain functional
 \[
@@
 For this reason, we refer to trajectories of \cref{eq:etafields} as $\lambda$-lines.
-Following \citet{haller13:_coher_lagran,haller14:_adden_coher_lagran}, we call the outermost member of a closed family of $\lambda$-lines a coherent Lagrangian vortex boundary.
+Following \textcite{haller13:_coher_lagran,haller14:_adden_coher_lagran}, we call the outermost member of a closed family of $\lambda$-lines a coherent Lagrangian vortex boundary.
 
@@
 Next we consider positions of material lines at time $t_0$ that prevail as most repelling or attracting material lines (or \emph{hyperbolic LCSs}) over a time interval $[t_0,t] \subset [t_-,t_+]$.
-\citet{farazmand14:_shearless} argue that hyperbolic LCSs are stationary curves of the averaged shear functional
+\textcite{farazmand14:_shearless} argue that hyperbolic LCSs are stationary curves of the averaged shear functional
 \begin{gather*}
@@
 More precisely, hyperbolic LCSs are stationary curves of $Q(\gamma)$ with respect to fixed-endpoint perturbations.
-We note that parabolic LCSs (Lagrangian jet cores) are also stationary curves of $Q(\gamma)$, but under variable endpoint perturbations (cf.
-\citet{farazmand14:_shearless}).
+We note that parabolic LCSs (Lagrangian jet cores) are also stationary curves of $Q(\gamma)$, but under variable endpoint perturbations (cf. \textcite{farazmand14:_shearless}).
 
-Solutions to this variational problem turn out to be orbits of the $\xi_1$ or $\xi_2$ eigenvector field. Repelling LCSs (\emph{shrink lines}) are obtained as trajectories of the differential equation
+Solutions to this variational problem turn out to be orbits of the $\xi_1$ or $\xi_2$ eigenvector field.
+Repelling LCSs (\emph{shrink lines}) are obtained as trajectories of the differential equation
 \begin{equation}
@@
 
+\begin{sloppypar}
 Shrink lines and stretch lines coincide with the null geodesics of the Lorentzian metric $h(u,v) = \left\langle u,D_{t_0}^t(r) v \right\rangle$.
 For this reason, we refer to the computation of hyperbolic LCSs as strongest normally-repelling or normally-attracting orbits of \cref{eq:etafields} as \emph{geodesic detection of hyperbolic LCSs}.
+\end{sloppypar}
 
@@
 \begin{center}
-\begin{tabular}{|c|p{.7\textwidth}|}
+\begin{tabular}{|c|p{.6\textwidth}|}
 \hline
-\multicolumn{2}{|p{.95\textwidth}|}{\lstinline![cgEigenvector,cgEigenvalue] = eig_cgStrain(derivative,domain,timespan,resolution)!}\tabularnewline
+\multicolumn{2}{|p{.8\textwidth}|}{\lstinline![cgEigenvector,cgEigenvalue] = eig_cgStrain(derivative,domain,timespan,resolution)!}\tabularnewline
 \hline \hline
@@
 Here $\delta x_j$ is a vector of length $\vert\delta x_j\vert > 0$ that points from the Cartesian grid-point $x_0$ in the $j^\text{th}$ coordinate direction (\cref{f:main and auxiliary grids}).
-Computational improvements arising from the use of the auxiliary grid over simply using the nearest points of the main grid were reported in \citet{farazmand12:_comput_lagran}.
+Computational improvements arising from the use of the auxiliary grid over simply using the nearest points of the main grid were reported in \textcite{farazmand12:_comput_lagran}.
 Experience suggests setting the auxiliary grid spacing to 1-10\% of the main grid spacing.
@@
 \begin{center}
-\def\svgwidth{.325\textwidth}
+\def\svgwidth{.75\textwidth}
 \import{graphics/}{main_aux_grids.pdf_tex}
@@
 At these points the eigenvectors $\xi_1(x_0)$ and $\xi_2(x_0)$ are no longer well-defined.
-This generically arises at a finite set of isolated points within the computational domain\citep{delmarcelle94}, and hence lie off the computational grid with probability one.
+This generically arises at a finite set of isolated points within the computational domain\parencite{delmarcelle94}, and hence lie off the computational grid with probability one.
 
@@
 
-Incompressible flows (i.e.
-those satisfying $\nabla \cdot v=0$) satisfy the relation $\lambda_1(x_0) \lambda_2(x_0) = 1$ at all points of the computational domain \citep{arnold89:_mathem}.
+\begin{sloppypar}
+Incompressible flows (i.e. those satisfying $\nabla \cdot v=0$) satisfy the relation $\lambda_1(x_0) \lambda_2(x_0) = 1$ at all points of the computational domain \parencite{arnold89:_mathem}.
 Incompressibility can be computationally imposed by first calculating $\lambda_2(x_0)$, then setting $\lambda_1(x_0) = 1/\lambda_2(x_0)$ and calculating the strain eigenvectors $\xi_2$ from $\lambda_2$, then $\xi_1$ from the relationship in \cref{eq:CG_invariants}.
-Experience shows that computing $\lambda_i$ in this order gives higher accuracy than in the reverse order \citep{farazmand12:_comput_lagran}.
+Experience shows that computing $\lambda_i$ in this order gives higher accuracy than in the reverse order \parencite{farazmand12:_comput_lagran}.
+% FIXME Add table or graph with calculations demonstrating statement below
+This procedure should be applied carefully since by changing $\lambda_1$ but not $\xi_1$, the error in satisfying the eigenvalue equation for the Cauchy-Green strain tensor may increase.
+\end{sloppypar}
 
@@
 A $\lambda$-line returning to its starting point is then an elliptic LCS.
-The outermost member of a family of closed $\lambda$-lines (obtained by varying $\lambda$) is a coherent Lagrangian vortex boundary\citep{haller13:_coher_lagran,haller14:_adden_coher_lagran}.
+The outermost member of a family of closed $\lambda$-lines (obtained by varying $\lambda$) is a coherent Lagrangian vortex boundary\parencite{haller13:_coher_lagran,haller14:_adden_coher_lagran}.
 
@@
 The Poincare section is to be oriented such that the first endpoint is close to the centre of the expected Lagrangian vortex, and the second endpoint is outside this vortex.
-There is no automated procedure implemented in LCS Tool for the positioning of Poincare sections (see however \citet{karrasch14:_autom_lagran} for a recently developed algorithm).
+There is no automated procedure implemented in LCS Tool for the positioning of Poincare sections (see however \textcite{karrasch14:_autom_lagran} for a recently developed algorithm).
 Additionally, the number of lambda-lines launched from the Poincare section, \lstinline!poincareSection.numPoints!, must be defined.
@@
 \begin{center}
-\def\svgwidth{.4\textwidth}
+\def\svgwidth{.75\textwidth}
 \import{graphics/}{variable_step_integration.pdf_tex}
@@
 
+\begin{sloppypar}
 An example of a Poincare map produced from integration of the $\eta_\pm^\lambda$ field is shown in \cref{f:Poincare return map}.
@@
 In practice, viewing the Poincare section as the radius of a circle and setting the maximum $\lambda$-line integration length to twice the circumference gives good results.
+\end{sloppypar}
 
@@
 \begin{center}
-\setlength\figurewidth{.35\textwidth}
+\setlength\figurewidth{.75\textwidth}
 \input{graphics/double_gyre/poincare_return_map.tikz}
@@
 
+The local maximization distance is the number of grid points over which the Cauchy-Green strain tensor is susceptible to numerical noise from integration.
+This distance is thus expected to be a function of variable time step integration tolerances.
+
+Prior to initiating shrinkline and stretchline integration, a maximum length is defined.
+Shrinklines and stretchlines are then integrated until reaching the domain boundary or the maximum length.
+The maximum length may be seen as safety factor to prevent highly sinuous shrinklines and stretchlines leading to inordinately long computations; the maximum length should be set such that most shrinklines and stretchlines stop at the domain boundary.
+
+Once shrinkline and stretchline integration has begun, it may evolve so as to be within the local minimization distance of an existing shrinkline or stretchline.
+In other words, the local minimization distance affects where shrinklines and stretchlines begin, but not where they finish.
+
 \section{Examples}
@@
 These functions are written to achieve balance between ease of use and performance.
-% FIXME With lambda range running demos can take over an hour.
+% FIXME With lambda range demo runtimes can be over an hour.
 % FIXME List runtime of demo scripts in a table
@@
 
-The double gyre is a model for a time-dependent two gyre system observed in geophysical flows \citep{shadden05:_defin_lagran_lyapun}.
+The double gyre is a model for a time-dependent two gyre system observed in geophysical flows \parencite{shadden05:_defin_lagran_lyapun}.
 The model consists of two counter rotating sinusoidal vortices with a harmonically oscillating line in-between.
@@
 
-% FIXME Reviewer #2, comment #8: Check all code listings for awkward splits across pages and line-splits (e.g. semicolon on own line and array indices)
-% FIXME Reviewer #5, comment #3: Format manuscript carefully with regards to code listings
 \begin{lstlisting}[caption={Double gyre derivative function corresponding to \cref{eq:double gyre derivative equations}.},label=l:double gyre derivative]
@@
 The free stretching parameter $\lambda$ (\cref{eq:etafields}) is varied over the range $[0.93,1.07]$ with increments of $0.01$ (\cref{ll:double gyre lambdaRange}).
-% FIXME The cited loop contains ellipses (...), that hide important code and should become an LCS Tool function
+% FIXME Loop replaced by the function poincare_closed_orbit_range
 In a loop (\crefrange{ll:double gyre lambda loop start}{ll:double gyre lambda loop end}), closed orbits for all predefined $\lambda$ values are computed and the outermost closed orbit is kept as the Lagrangian vortex boundary.
@@
 \centering
-\setlength\figurewidth{.4\textwidth}
+\setlength\figurewidth{.875\textwidth}
 \input{graphics/double_gyre/elliptic_lcs_convergence_500.tikz}
@@
 
-% FIXME Reviewer #5, comment #4: There is too much source code in the main content such as cref{l:double gyre lambda strain stretch LCS,l:Bickley jet elliptic hyperbolic LCS subset,l:ocean data elliptic hyperbolic lcs}.
-They could be arranged as an appendix at the end.
 \lstinputlisting[
@@
 \centering
-% FIXME Reviewer #2, comment #6: Improve appearance of $lambda$ extrema (white dots)
-\setlength\figurewidth{.35\textwidth}
+\setlength\figurewidth{.8\textwidth}
 \input{graphics/double_gyre/elliptic_repelling_lcs_details.tikz}
@@
 counter rotating vortices.
-This is an idealized model of geophysical flows such as the Gulf Stream and the polar night jet perturbed by a Rossby wave\citep{castilloNegrete93:_chaot_rossb,beron-vera10:_invar_lagran}.
+This is an idealized model of geophysical flows such as the Gulf Stream and the polar night jet perturbed by a Rossby wave\parencite{castilloNegrete93:_chaot_rossb,beron-vera10:_invar_lagran}.
 
@@
 \centering
-\setlength\figurewidth{.325\textwidth}
+\setlength\figurewidth{.775\textwidth}
 \import{graphics/bickley_jet/}{elliptic_repelling_lcs_details.tikz}
@@
 In contrast to the previous two analytic examples, the velocity field is available only with discrete temporal and spatial resolution.
-Our region of interest is a small domain in the South Atlantic Ocean, where exceptionally coherent eddies, Agulhas rings, were recently found by \citet{haller13:_coher_lagran,haller14:_adden_coher_lagran} using the theory we surveyed in \cref{sec:Elliptic LCSs}.
+Our region of interest is a small domain in the South Atlantic Ocean, where exceptionally coherent eddies, Agulhas rings, were recently found by \textcite{haller13:_coher_lagran,haller14:_adden_coher_lagran} using the theory we surveyed in \cref{sec:Elliptic LCSs}.
 
@@
 
-\begin{sloppypar}
-\Cref{f:ocean dataset LCS} shows elliptic and hyperbolic LCSs on 22 November 2006, the same time as analysed in \citet{haller13:_coher_lagran,haller14:_adden_coher_lagran,beron-vera13:_objec_agulh}.
+\Cref{f:ocean dataset LCS} shows elliptic and hyperbolic LCSs on 22 November 2006, the same time as analysed in \textcite{haller13:_coher_lagran,haller14:_adden_coher_lagran,beron-vera13:_objec_agulh}.
 In those references, the integration time is 90 days.
@@
 The largest, at $(3,-32)$, has a non-stretching boundary, i.e., $\lambda = 1$.
-It corresponds to eddy \#2 in figure~3 of \citet{beron-vera13:_objec_agulh}.
+It corresponds to eddy \#2 in figure~3 of \textcite{beron-vera13:_objec_agulh}.
 Four additional smaller coherent eddies are found.
@@
 The hyperbolic LCSs determine the deformation of the fluid between coherent eddies.
-\end{sloppypar}
 
@@
 \begin{center}
-\setlength\figurewidth{.375\textwidth}
+\setlength\figurewidth{.85\textwidth}
 \import{graphics/ocean_dataset/}{elliptic_repelling_lcs.tikz}
@@
 LCS Tool leverages the capabilities of MATLAB.
-For FTLE based extraction of LCSs, computational performance has received considerable attention \citep{conti12:_gpu_apu_finit_time_lyapun_expon,miron12:_anisot_lagran_coher_struc}.
+For FTLE based extraction of LCSs, computational performance has received considerable attention \parencite{conti12:_gpu_apu_finit_time_lyapun_expon,miron12:_anisot_lagran_coher_struc}.
 We think LCS Tool can facilitate similar computational advances for variational LCS methods.
-Optimizing computational performance will aid applications to large-scale forecasting applications, such as the tracking of environmental contaminants\citep{olascoaga12:_forec}.
+Optimizing computational performance will aid applications to large-scale forecasting applications, such as the tracking of environmental contaminants\parencite{olascoaga12:_forec}.
 
 We hope LCS Tool will serve as a foundation for the numerical implementation of recent theoretical advances.
-These include the geodesic theory of parabolic LCSs (jet cores)\citep{farazmand14:_shearless} and the variational theory of hyperbolic and elliptic LCSs for three-dimensional flows\citep{blazevski14:_hyper_ellip_trans_barrier_three}.
+These include the geodesic theory of parabolic LCSs (jet cores)\parencite{farazmand14:_shearless} and the variational theory of hyperbolic and elliptic LCSs for three-dimensional flows\parencite{blazevski14:_hyper_ellip_trans_barrier_three}.
 
@@
 
-\begin{sloppypar}
 The altimeter products used in this work are produced by SSALTO/DUACS and distributed by AVISO, with support from CNES, \url{http://www.aviso.oceanobs.com/duacs}.
-\end{sloppypar}
-
-\section*{References}
 
 % FIXME Citations without pages (karrasch14:_autom_lagran) do not list publication year
-\bibliographystyle{elsarticle-num-names} 
-\bibliography{main}
+\printbibliography
 
@@
 % Local Variables:
-% ispell-dictionary: "en_GB"
+% ispell-dictionary: "en_CA"
 % End:

### 2014-12-08 main.tex
Remove stray draft options
List demo runtimes

@@
-\documentclass[draft]{article}
+\documentclass{article}
 
@@
 \usepackage{import}
-\usepackage[final]{listings}
+\usepackage{listings}
 \usepackage[numbered,framed]{mcode}
@@
 The examples are available as scripts in the demo folder of LCS Tool.
-Executing these scripts can help readers follow LCS Tool computations.
+Executing these scripts demonstrates how to call LCS Tool functions.
+These functions are written to achieve balance between ease of use and performance.
+Examples of demo runtimes are given in \cref{t:demo runtimes}.
 
-% FIXME Performance also described in Conclusions section
-The scripts call LCS Tool functions.
-These functions are written to achieve balance between ease of use and performance.
-% FIXME With lambda range demo runtimes can be over an hour.
-% FIXME List runtime of demo scripts in a table
-Running the demo scripts takes on the order of minutes.
+\begin{table}
+\begin{center}
+\begin{tabular}{|c|c|}
+\hline
+Demo & Runtime (min)\tabularnewline
+\hline \hline
+double\_gyre/elliptic\_hyperbolic\_lcs & 50\tabularnewline
+\hline
+bickley\_jet/elliptic\_hyperbolic\_lcs\_details & 975\tabularnewline
+\hline
+ocean\_dataset/elliptic\_hyperbolic\_lcs\_details & 127\tabularnewline
+\hline
+\end{tabular}
+\end{center}
+\caption{Runtime of some demo scripts Results obtained with an Intel Core i3-4130 processor with eight gigabytes of memory.}
+\label{t:demo runtimes}
+\end{table}
 

### 2014-12-08 main.tex
Improve writing style for paragraph in conclusions
Improve writing style for paragraph in conclusions
Conflicts:
	main.tex

--- 2014-12-02 main.tex
@@
 We present an algorithmic introduction to geodesic LCS detection.
-% FIXME Reviewer #3, comment #4: Add documentation within the github repository to show some reference to the APIs and how users may start to run a his/her own application.
 We then review the implementation of this approach in a computational engine called LCS Tool\footnote{LCS Tool is available at: \url{http://www.runmycode.org/companion/view/908}}.
@@
 The examples are available as scripts in the demo folder of LCS Tool.
-Executing these scripts can help readers follow LCS Tool computations.
+Executing these scripts demonstrates how to call LCS Tool functions.
+These functions are written to achieve balance between ease of use and performance.
+Examples of demo runtimes are given in \cref{t:demo runtimes}.
 
-% FIXME Performance also described in Conclusions section
-The scripts call LCS Tool functions.
-These functions are written to achieve balance between ease of use and performance.
-% FIXME With lambda range running demos can take over an hour.
-% FIXME List runtime of demo scripts in a table
-Running the demo scripts takes on the order of minutes.
+\begin{table}
+\begin{center}
+\begin{tabular}{|p{.3\textwidth}|c|}
+\hline
+Demo & Runtime (min)\tabularnewline
+\hline \hline
+double gyre & 50\tabularnewline
+\hline
+Bickley jet & 975\tabularnewline
+\hline
+ocean dataset & 127\tabularnewline
+\hline
+\end{tabular}
+\end{center}
+\caption{Runtime of some demo scripts Results obtained with an Intel Core i3-4130 processor with eight gigabytes of memory.}
+\label{t:demo runtimes}
+\end{table}
 
@@
 
-We hope LCS Tool will serve as a foundation for the numerical implementation of recent theoretical advances.
+We think LCS Tool can serve as a foundation for the numerical implementation of recent theoretical advances.
 These include the geodesic theory of parabolic LCSs (jet cores)\citep{farazmand14:_shearless} and the variational theory of hyperbolic and elliptic LCSs for three-dimensional flows\citep{blazevski14:_hyper_ellip_trans_barrier_three}.

### 2014-12-10 main.tex
Reformat Conclusions LaTeX source code
List demo runtimes
Shorten & revise code listings
- use math characters
- abbreviate variable names
- remove some input parameters
- use the Listings breakwhitespace option to control line breaks in
  code listings

@@
   basicstyle=\small,
-  breaklines=true,
+  breaklines,
+  breakatwhitespace,
   frame=single,
@@
   backgroundcolor=\color{grey},
-  escapeinside={\%@}{\^^M}
+  escapeinside={\%@}{\^^M},
+  mathescape
 }
@@
 
-% FIXME Reviewer #2, comment #8: Check all code listings for awkward splits across pages and line-splits (e.g. semicolon on own line and array indices)
-% FIXME Reviewer #5, comment #3: Format manuscript carefully with regards to code listings
-\begin{lstlisting}[caption={Double gyre derivative function corresponding to \cref{eq:double gyre derivative equations}.},label=l:double gyre derivative]
-function derivative_ = derivative(t,x,epsilon,amplitude,omega)
-
-idx1 = 1:2:numel(x)-1;
-idx2 = 2:2:numel(x);
-
-a = epsilon*sin(omega*t);
-b = 1 - 2*epsilon*sin(omega*t);
-forcing = a*x(idx1).^2 + b*x(idx1);
-
-derivative_ = nan(size(x));
-
-derivative_(idx1) = -pi*amplitude*sin(pi*forcing).*cos(pi*x(idx2));
-derivative_(idx2) = pi*amplitude*cos(pi*forcing).*sin(pi*x(idx2)).*(2*a*x(idx1) + b);
-\end{lstlisting}
+\lstinputlisting[caption={Double gyre derivative function corresponding to \cref{eq:double gyre derivative equations}.},
+                 label=l:double gyre derivative]{listings/double_gyre/derivative.m}
 
@@
 The free stretching parameter $\lambda$ (\cref{eq:etafields}) is varied over the range $[0.93,1.07]$ with increments of $0.01$ (\cref{ll:double gyre lambdaRange}).
-% FIXME Loop replaced by the function poincare_closed_orbit_range
-In a loop (\crefrange{ll:double gyre lambda loop start}{ll:double gyre lambda loop end}), closed orbits for all predefined $\lambda$ values are computed and the outermost closed orbit is kept as the Lagrangian vortex boundary.
+Closed orbits for all predefined $\lambda$ values are computed and the outermost closed orbit is kept as the Lagrangian vortex boundary  (\crefrange{ll:double gyre lambda range}{ll:double gyre elliptic_lcs}).
 % FIXME Small tolerances (i.e. strict tolerances) should make convergence difficult not easy, so the word "facilitate" in the sentence below seems illogical.
@@
 
-\lstinputlisting[caption={LCS Tool commands for double gyre elliptic LCSs. Subset of the LCS Tool script \lstinline!demo/double_gyre/elliptic_hyperbolic_lcs.m!},
+\lstinputlisting[caption={LCS Tool commands for double gyre elliptic LCSs. Abridged subset of the LCS Tool script \lstinline!demo/double_gyre/elliptic_hyperbolic_lcs.m!},
                  label=l:double gyre elliptic LCS]{listings/double_gyre/elliptic_lcs_subset.m}
@@
 \Cref{f:double gyre lambda hyperbolic LCS} shows elliptic and hyperbolic LCSs for the double gyre at this resolution.
-% FIXME Sentence below is inaccurate since \cref{l:double gyre lambda strain stretch LCS} gives commands for hyperbolic LCSs only
-\Cref{l:double gyre lambda strain stretch LCS} gives the corresponding LCS Tool commands.
+\Cref{l:double gyre lambda strain stretch LCS} gives the LCS Tool commands to compute hyperbolic LCSs.
 The maximum length of shrink lines and stretch lines, \lstinline!shrinkLineMaxLength!
@@
 
-% FIXME Reviewer #5, comment #4: There is too much source code in the main content such as cref{l:double gyre lambda strain stretch LCS,l:Bickley jet elliptic hyperbolic LCS subset,l:ocean data elliptic hyperbolic lcs}.
-They could be arranged as an appendix at the end.
 \lstinputlisting[
-        caption={LCS Tool commands for double gyre hyperbolic LCSs. Subset of the LCS Tool script \lstinline!demo/double_gyre/elliptic_hyperbolic_lcs.m!},
+        caption={LCS Tool commands for double gyre hyperbolic LCSs. Abridged subset of the LCS Tool script \lstinline!demo/double_gyre/elliptic_hyperbolic_lcs.m!},
         label=l:double gyre lambda strain stretch LCS]
@@
 The integration time is $T = 4L_x/U$, an integer multiple of the eddy turnover time.
-\Cref{l:Bickley jet elliptic hyperbolic LCS subset} shows the LCS Tool commands in which the chaotically perturbed velocity is defined (\crefrange{ll:Bickley jet velocity definition start}{ll:Bickley jet velocity definition end}), periodic boundary conditions are imposed in the x-direction (\cref{ll:Bickley jet periodicBc}), and five Poincare sections are defined where we expect coherent vortices (\crefrange{ll:Bickley jet Poincare start}{ll:Bickley jet Poincare end}).
+\Cref{l:Bickley jet elliptic hyperbolic LCS subset} shows the LCS Tool commands in which the chaotically perturbed velocity is defined (\cref{ll:Bickley jet velocity definition}), periodic boundary conditions are imposed in the x-direction (\cref{ll:Bickley jet periodicBc}), and five Poincare sections are defined where we expect coherent vortices (\crefrange{ll:Bickley jet Poincare start}{ll:Bickley jet Poincare end}).
 $\lambda$-values for closed orbit detection are varied over the range $[0.80,1.20]$ with a step of $0.01$.
@@
 \lstinputlisting[
-        caption={LCS Tool commands for Bickley jet elliptic and hyperbolic LCSs. Subset of the LCS Tool script \lstinline!demo/bickley_jet/elliptic_hyperbolic_lcs.m!},
+        caption={LCS Tool commands for Bickley jet elliptic and hyperbolic LCSs. Abridged version of the LCS Tool script \lstinline!demo/bickley_jet/elliptic_hyperbolic_lcs.m!},
         label=l:Bickley jet elliptic hyperbolic LCS subset]
@@
 \lstinputlisting[
-        caption={LCS Tool commands for ocean data elliptic and hyperbolic LCSs. Subset of the LCS Tool script \lstinline!demo/ocean_dataset/elliptic_hyperbolic_lcs.m!},
+        caption={LCS Tool commands for ocean data elliptic and hyperbolic LCSs. Abridged version of the LCS Tool script \lstinline!demo/ocean_dataset/elliptic_hyperbolic_lcs.m!},
         label={l:ocean data elliptic hyperbolic lcs}]
@@
 With this choice, the resolution of the tracer grid is 15 times higher than the resolution of the velocity field.
-The flow is integrated and the Cauchy-Green strain tensor is computed by the function \lstinline!eig_cgStrain!
-(\cref{ll:eig_cgStrain}).
-Incompressibility of the flow is enforced (\cref{ll:incompressible}), the auxiliary grid distance is set to 1\% of the main grid distance (\cref{ll:cgAuxGridRelDelta}), and eigenvalues are computed from the auxiliary grid (\cref{ll:cgEigenvalueFromMainGrid}).
-Elliptic LCSs are computed in \cref{ll:ocean data poincare_closed_orbit_multi}, after the Poincare sections have been set (\cref{ll:ocean data poincareSection(1),ll:ocean data poincareSection(2)}) and the $\eta_\pm^\lambda$ fields have been defined (\cref{ll:ocean data lambda_line}).
+The flow is integrated and the Cauchy-Green strain tensor is computed by the function \lstinline!eig_cgStrain! (\cref{ll:eig_cgStrain}).
+The auxiliary grid distance is set to 1\% of the main grid distance (\cref{ll:cgAuxGridRelDelta}), and eigenvalues are computed from the auxiliary grid (\cref{ll:cgEigenvalueFromMainGrid}).
+Elliptic LCSs are computed in \cref{ll:ocean data poincare_closed_orbit_multi}, after the Poincare sections have been set (\cref{ll:ocean data poincareSection(1)}).
 $\lambda$-values are varied over a range of $[0.90,1.10]$ with a step of $0.02$ (\cref{ll:ocean data lambdaRange}).

### 2014-12-10 listings/bickley_jet/elliptic_hyperbolic_lcs_subset.m

@@ -1,73 +1,42 @@
 %% Input parameters
-u = 62.66;
-lengthX = pi*earthRadius;
-lengthY = 1.77e6;
-epsilon = [.075,.4,.3];
-domain = [0,lengthX;[-1,1]*2.25*lengthY];
-resolution = [500,276];
-timespan = [0,4*lengthX/u];
+...
+domain = [0,$L_x$;[-1,1]*2.25*$L_y$];
+res = [500,276];
+time = [0,4*$L_x$/u];
 
 %% Velocity definition
-perturbationCase = 3; %@\label{ll:Bickley jet velocity definition start}
-phiTimespan = [0,25];
-phiInitial = [0,0];
-phiSol = ode45(@d_phi,phiTimespan,phiInitial);
-timeResolution = 1e5;
-phi1 = deval(phiSol,linspace(phiTimespan(1),phiTimespan(2),timeResolution),1);
-phi1Max = max(phi1);
-lDerivative = @(t,x,~)derivative(t,x,false,u,lengthX,lengthY,epsilon,perturbationCase,phiSol,phi1Max); %@\label{ll:Bickley jet velocity definition end}
-incompressible = true;
-periodicBc = [true,false]; %@\label{ll:Bickley jet periodicBc}
-
-%% Cauchy-Green strain eigenvalues and eigenvectors
-[cgEigenvector,cgEigenvalue] = eig_cgStrain(lDerivative,domain,resolution,timespan,'incompressible',incompressible);
+...
+lDerivative = @(t,x)derivative(t,x,u,$L_x$,$L_y$,$\epsilon$, perturbationCase,phiSol,phi1Max); %@\label{ll:Bickley jet velocity definition}
+pBc = [true,false]; %@\label{ll:Bickley jet periodicBc}
 
 %% LCS parameters
-% Lambda-lines
-lambdaRange = .8:.01:1.1;
-lambdaLineOdeSolverOptions = odeset('relTol',1e-6);
-
-% Shrink lines
-shrinkLineMaxLength = 1e8;
-shrinkLineLocalMaxDistance = 8*gridSpace;
-shrinkLineOdeSolverOptions = odeset('relTol',1e-4);
-
-% Stretch lines
-stretchLineMaxLength = 1e8;
-stretchLineLocalMaxDistance = 4*gridSpace;
-stretchLineOdeSolverOptions = odeset('relTol',1e-4);
-
-%% Lambda-line LCSs
[…]

### 2014-12-10 listings/double_gyre/elliptic_lcs_subset.m

@@ -1,38 +1,27 @@
 %% Input parameters
-epsilon = .1;
-amplitude = .1;
-omega = pi/5;
+...
 domain = [0,2;0,1];
-resolution = [500,250];
-timespan = [0,10];
+res = [500,250];
+time = [0,10];
 
 %% Velocity definition
-lDerivative = @(t,x,~)derivative(t,x,false,epsilon,amplitude,omega);
-incompressible = true;
+lDerivative = @(t,x)derivative(t,x,$\epsilon$,A,$\omega$);
 
 %% LCS parameters
-cgStrainOdeSolverOptions = odeset('relTol',1e-5); %@\label{ll:double gyre cgStrainOdeSolverOptions}
+cgOptions = odeset('relTol',1e-5); %@\label{ll:double gyre cgStrainOdeSolverOptions}
 
-% Lambda-lines
-poincareSection = struct('endPosition',{},'numPoints',{},'orbitMaxLength',{});
-poincareSection(1).endPosition = [.55,.55;.1,.1]; %@\label{ll:double gyre poincareSection(1)}
-poincareSection(2).endPosition = [1.53,.45;1.95,.05]; %@\label{ll:double gyre poincareSection(2)}
-[poincareSection.numPoints] = deal(100);
-nPoincareSection = numel(poincareSection);
-for i = 1:nPoincareSection
-    rOrbit = hypot(diff(poincareSection(i).endPosition(:,1)),diff(poincareSection(i).endPosition(:,2)));
-    poincareSection(i).orbitMaxLength = 2*(2*pi*rOrbit);
-end
-lambdaLineOdeSolverOptions = odeset('relTol',1e-6); %@\label{ll:double gyre lambdaLineOdeSolverOptions}
-lambdaRange = .93:.01:1.07; %@\label{ll:double gyre lambdaRange}
+% Lambda lines
+...
+ps(1).endPosition = [.55,.55;.1,.1]; %@\label{ll:double gyre poincareSection(1)}
+ps(2).endPosition = [1.53,.45;1.95,.05]; %@\label{ll:double gyre poincareSection(2)}
+...
+lambda = .93:.01:1.07; %@\label{ll:double gyre lambdaRange}
+llOptions = odeset('relTol',1e-6); %@\label{ll:double gyre lambdaLineOdeSolverOptions}
 
 %% Cauchy-Green strain eigenvalues and eigenvectors
[…]

### 2014-12-10 listings/double_gyre/hyperbolic_lcs_subset.m

@@ -1,17 +1,14 @@
 % Shrink lines
-shrinkLineMaxLength = 20;
-gridSpace = diff(domain(1,:))/(double(resolution(1))-1);
-shrinkLineLocalMaxDistance = 2*gridSpace; %@\label{ll:shrinkLineLocalMaxDistance}
-shrinkLineOdeSolverOptions = odeset('relTol',1e-6);
+shrinkLMaxLength = 20;
+shrinkLMaxDistance = 2*gridSpace; %@\label{ll:shrinkLineLocalMaxDistance}
 
 % Stretch lines
-stretchLineMaxLength = 20;
-stretchLineLocalMaxDistance = 10*gridSpace; %@\label{ll:stretchLineLocalMaxDistance}
-stretchLineOdeSolverOptions = odeset('relTol',1e-6);
+stretchLMaxLength = 20;
+stretchLMaxDistance = 10*gridSpace; %@\label{ll:stretchLineLocalMaxDistance}
 
 %% Hyperbolic shrink line LCSs
-[shrinkLineLcs,shrinkLineLcsInitialPosition] = seed_curves_from_lambda_max(shrinkLineLocalMaxDistance,shrinkLineMaxLength,cgEigenvalue(:,2),cgEigenvector(:,1:2),domain,resolution,'odeSolverOptions',shrinkLineOdeSolverOptions);
+shrinkL = seed_curves_from_lambda_max( shrinkLMaxDistance,shrinkLMaxLength,cgV(:,2), cgD(:,1:2),domain,res);
 ...
 %% Hyperbolic stretch line LCSs
-[stretchLineLcs,stretchLineLcsInitialPosition] = seed_curves_from_lambda_max(stretchLineLocalMaxDistance,stretchLineMaxLength,-cgEigenvalue(:,1),cgEigenvector(:,3:4),domain,resolution,'odeSolverOptions',stretchLineOdeSolverOptions);
+stretchL = seed_curves_from_lambda_max( stretchLMaxDistance,stretchLMaxLength,-cgV(:,1), cgD(:,3:4),domain,res);
 ...

### 2014-12-10 listings/ocean_dataset/elliptic_hyperbolic_lcs_subset.m

@@ -1,66 +1,43 @@
 % Input parameters
 domain = [0,6;-34,-28];
-resolution = [400,400]; %@\label{ll:ocean data resolution}
-timespan = [100,130]; %@\label{ll:ocean data timespan}
+res = [400,400]; %@\label{ll:ocean data resolution}
+time = [100,130]; %@\label{ll:ocean data timespan}
 
 % Velocity definition
 load('ocean_geostrophic_velocity.mat');
 ...
-interpMethod = 'spline'; %@\label{ll:interpolation start}
-vlon_interpolant = griddedInterpolant({time,lat,lon},vlon,interpMethod);
-vlat_interpolant = griddedInterpolant({time,lat,lon},vlat,interpMethod);
-lDerivative = @(t,x,~)flowdata_derivative(t,x,vlon_interpolant,vlat_interpolant); %@\label{ll:interpolation end}
-incompressible = true; %@\label{ll:incompressible}
+intMet = 'spline'; %@\label{ll:interpolation start}
+vLonI = griddedInterpolant({time,lat,lon},vLon,intMet);
+vLatI = griddedInterpolant({time,lat,lon},vLat,intMet);
+lDerivative = @(t,x,~)flowdata_derivative(t,x,vLonI,vLatI); %@\label{ll:interpolation end}
 
 % LCS parameters
 % Cauchy-Green strain
-cgEigenvalueFromMainGrid = false; %@\label{ll:cgEigenvalueFromMainGrid}
-cgAuxGridRelDelta = .01; %@\label{ll:cgAuxGridRelDelta}
+cgVmg = false; %@\label{ll:cgEigenvalueFromMainGrid}
+cgAgd = .01; %@\label{ll:cgAuxGridRelDelta}
 
-% Lambda-lines
-lambdaRange = .9:.02:1.1; %@\label{ll:ocean data lambdaRange}
-lambdaLineOdeSolverOptions = odeset('relTol',1e-6);
-
+% Lambda lines
+...
+ps(1).endPosition = [3.3,-32.1;3.7,-31.6]; %@\label{ll:ocean data poincareSection(1)}
+...
+lambda = .9:.02:1.1; %@\label{ll:ocean data lambdaRange}
+llOptions = odeset('relTol',1e-6);
+...
 % Shrink lines
-shrinkLineMaxLength = 20;
-gridSpace = diff(domain(1,:))/(double(resolution(1))-1);
-shrinkLineLocalMaxDistance = 2*gridSpace;
[…]

### 2014-12-10 main.tex
Revise grammar

--- 2014-12-08 main.tex
@@
 By examining the FTLE field (which LCS Tool can provide), we position Poincare sections to capture elliptic LCSs.
-An LCS Tool demo script to perform this operation is given in \cref{l:double gyre elliptic LCS} where two Poincare sections are defined (\cref{ll:double gyre poincareSection(1),ll:double gyre poincareSection(2)}).
+A script to perform this operation is given in \cref{l:double gyre elliptic LCS} where two Poincare sections are defined (\cref{ll:double gyre poincareSection(1),ll:double gyre poincareSection(2)}).
 The free stretching parameter $\lambda$ (\cref{eq:etafields}) is varied over the range $[0.93,1.07]$ with increments of $0.01$ (\cref{ll:double gyre lambdaRange}).
@@
 % FIXME Small tolerances (i.e. strict tolerances) should make convergence difficult not easy, so the word "facilitate" in the sentence below seems illogical.
-In this script, we have chosen small values for the error tolerances of the integration of the Cauchy-Green strain tensor (\cref{ll:double gyre cgStrainOdeSolverOptions}) and $\lambda$-lines (\cref{ll:double gyre lambdaLineOdeSolverOptions}) to facilitate convergence.
+We have chosen small values for the error tolerances of the integration of the Cauchy-Green strain tensor (\cref{ll:double gyre cgStrainOdeSolverOptions}) and $\lambda$-lines (\cref{ll:double gyre lambdaLineOdeSolverOptions}) to facilitate convergence.
 
@@
 Shrink lines are locally tangent to ridges of $\lambda_2$ maxima, whereas stretch lines are locally normal to these ridges.
-Setting the local maximization distance of stretch lines and shrink lines equal would therefore produce a greater number of stretch lines than shrink lines overall.
+Setting the local maximization distance of stretch lines and shrink lines equal would therefore produce a greater number of stretch lines than shrink lines.
 
@@
 
-We hope LCS Tool will serve as a foundation for the numerical implementation of recent theoretical advances.
+We think LCS Tool can serve as a foundation for the numerical implementation of recent theoretical advances.
 These include the geodesic theory of parabolic LCSs (jet cores)\parencite{farazmand14:_shearless} and the variational theory of hyperbolic and elliptic LCSs for three-dimensional flows\parencite{blazevski14:_hyper_ellip_trans_barrier_three}.

### 2014-12-10 main.tex

--- 2014-12-10 main.tex
@@
   basicstyle=\small,
-  breaklines=true,
+  breaklines,
   frame=single,
@@
 
-\begin{lstlisting}[caption={Double gyre derivative function corresponding to \cref{eq:double gyre derivative equations}.},label=l:double gyre derivative]
-function derivative_ = derivative(t,x,epsilon,amplitude,omega)
-
-idx1 = 1:2:numel(x)-1;
-idx2 = 2:2:numel(x);
-
-a = epsilon*sin(omega*t);
-b = 1 - 2*epsilon*sin(omega*t);
-forcing = a*x(idx1).^2 + b*x(idx1);
-
-derivative_ = nan(size(x));
-
-derivative_(idx1) = -pi*amplitude*sin(pi*forcing).*cos(pi*x(idx2));
-derivative_(idx2) = pi*amplitude*cos(pi*forcing).*sin(pi*x(idx2)).*(2*a*x(idx1) + b);
-\end{lstlisting}
+\lstinputlisting[caption={Double gyre derivative function corresponding to \cref{eq:double gyre derivative equations}.},
+                 label=l:double gyre derivative]{listings/double_gyre/derivative.m}
 
@@
 The free stretching parameter $\lambda$ (\cref{eq:etafields}) is varied over the range $[0.93,1.07]$ with increments of $0.01$ (\cref{ll:double gyre lambdaRange}).
-% FIXME Loop replaced by the function poincare_closed_orbit_range
-In a loop (\crefrange{ll:double gyre lambda loop start}{ll:double gyre lambda loop end}), closed orbits for all predefined $\lambda$ values are computed and the outermost closed orbit is kept as the Lagrangian vortex boundary.
+Closed orbits for all $\lambda$ values are computed and the outermost closed orbit is kept as the Lagrangian vortex boundary  (\crefrange{ll:double gyre lambda range}{ll:double gyre elliptic_lcs}).
 % FIXME Small tolerances (i.e. strict tolerances) should make convergence difficult not easy, so the word "facilitate" in the sentence below seems illogical.
@@
 \lstinputlisting[
-        caption={LCS Tool commands for double gyre hyperbolic LCSs. Subset of the LCS Tool script \lstinline!demo/double_gyre/elliptic_hyperbolic_lcs.m!},
+        caption={LCS Tool commands for double gyre hyperbolic LCSs. Abridged subset of the LCS Tool script \lstinline!demo/double_gyre/elliptic_hyperbolic_lcs.m!},
         label=l:double gyre lambda strain stretch LCS]
@@
 
-We choose the integration time as $T=30$ days (\cref{l:ocean data elliptic hyperbolic lcs}, \cref{ll:ocean data timespan}), which is larger than the eddy turnover time in this region.
-The resolution of the main computational grid for initial conditions is set to $400 \times 400$ (\cref{ll:ocean data resolution}).
-This corresponds to a resolution of roughly 0.015° and gives good results.
+We choose the integration time as $T=30$ days (\cref{l:ocean data elliptic hyperbolic lcs}, \cref{ll:ocean timespan}), which is larger than the eddy turnover time in this region.
+The resolution of the main computational grid for initial conditions is set to $400 \times 400$ (\cref{ll:ocean resolution}).
+This corresponds to a resolution of roughly 0.015°.
 With this choice, the resolution of the tracer grid is 15 times higher than the resolution of the velocity field.
-The flow is integrated and the Cauchy-Green strain tensor is computed by the function \lstinline!eig_cgStrain!
-(\cref{ll:eig_cgStrain}).
+The flow is integrated and the Cauchy-Green strain tensor is computed by the function \lstinline!eig_cgStrain! (\cref{ll:eig_cgStrain}).
 Incompressibility of the flow is enforced (\cref{ll:incompressible}), the auxiliary grid distance is set to 1\% of the main grid distance (\cref{ll:cgAuxGridRelDelta}), and eigenvalues are computed from the auxiliary grid (\cref{ll:cgEigenvalueFromMainGrid}).
-Elliptic LCSs are computed in \cref{ll:ocean data poincare_closed_orbit_multi}, after the Poincare sections have been set (\cref{ll:ocean data poincareSection(1),ll:ocean data poincareSection(2)}) and the $\eta_\pm^\lambda$ fields have been defined (\cref{ll:ocean data lambda_line}).
-$\lambda$-values are varied over a range of $[0.90,1.10]$ with a step of $0.02$ (\cref{ll:ocean data lambdaRange}).
-Hyperbolic LCSs are computed in \cref{ll:ocean data shrinkLineLcs,ll:ocean data stretchLineLcs}.
+Elliptic LCSs are computed in \crefrange{ll:ocean elliptic LCS start}{ll:ocean elliptic LCS end}, after the Poincare sections have been set (\crefrange{ll:ocean data poincareSection(1)}{ll:ocean data poincareSection(5)}).
+$\lambda$-values are varied over a range of $[0.90,1.10]$ with a step of $0.02$ (\cref{ll:ocean lambda}).
+Hyperbolic LCSs are computed in \crefrange{ll:ocean hyperbolic start}{ll:ocean hyperbolic end}.
 

### 2014-12-10 main.tex
Change graphics \input commands to \import
Conflicts:
	main.tex

--- 2014-12-10 main.tex
@@
 \setlength\figurewidth{.4\textwidth}
-\input{graphics/double_gyre/elliptic_lcs_convergence_500.tikz}
-\input{graphics/double_gyre/elliptic_lcs_convergence_750.tikz}
-\input{graphics/double_gyre/elliptic_lcs_convergence_1000.tikz}
+\import{graphics/double_gyre/}{elliptic_lcs_convergence_500.tikz}
+\import{graphics/double_gyre/}{elliptic_lcs_convergence_750.tikz}
+\import{graphics/double_gyre/}{elliptic_lcs_convergence_1000.tikz}
 \caption{Convergence of closed $\lambda$-lines for increasing main-grid resolution in the double-gyre.
@@
 \setlength\figurewidth{.35\textwidth}
-\input{graphics/double_gyre/elliptic_repelling_lcs_details.tikz}
-\input{graphics/double_gyre/elliptic_attracting_lcs_details.tikz}
+\import{graphics/double_gyre/}{elliptic_repelling_lcs_details.tikz}
+\import{graphics/double_gyre/}{elliptic_attracting_lcs_details.tikz}
 \caption{
@@
 The resolution of the main computational grid for initial conditions is set to $400 \times 400$ (\cref{ll:ocean data resolution}).
-This corresponds to a resolution of roughly 0.015° and gives good results.
+This corresponds to a resolution of roughly 0.015°.
 With this choice, the resolution of the tracer grid is 15 times higher than the resolution of the velocity field.

### 2014-12-10 main.tex
Change graphics \input commands to \import

--- 2014-12-10 main.tex
@@
 \setlength\figurewidth{.875\textwidth}
-\input{graphics/double_gyre/elliptic_lcs_convergence_500.tikz}
-\input{graphics/double_gyre/elliptic_lcs_convergence_750.tikz}
-\input{graphics/double_gyre/elliptic_lcs_convergence_1000.tikz}
+\import{graphics/double_gyre/}{elliptic_lcs_convergence_500.tikz}
+\import{graphics/double_gyre/}{elliptic_lcs_convergence_750.tikz}
+\import{graphics/double_gyre/}{elliptic_lcs_convergence_1000.tikz}
 \caption{Convergence of closed $\lambda$-lines for increasing main-grid resolution in the double-gyre.
@@
 \setlength\figurewidth{.8\textwidth}
-\input{graphics/double_gyre/elliptic_repelling_lcs_details.tikz}
-\input{graphics/double_gyre/elliptic_attracting_lcs_details.tikz}
+\import{graphics/double_gyre/}{elliptic_repelling_lcs_details.tikz}
+\import{graphics/double_gyre/}{elliptic_attracting_lcs_details.tikz}
 \caption{

### 2014-12-10 main.tex
Revise grammar

--- 2014-12-10 main.tex
@@
 \setlength\figurewidth{.4\textwidth}
-\import{graphics/double_gyre/}{elliptic_lcs_convergence_500.tikz}
-\import{graphics/double_gyre/}{elliptic_lcs_convergence_750.tikz}
-\import{graphics/double_gyre/}{elliptic_lcs_convergence_1000.tikz}
+\input{graphics/double_gyre/elliptic_lcs_convergence_500.tikz}
+\input{graphics/double_gyre/elliptic_lcs_convergence_750.tikz}
+\input{graphics/double_gyre/elliptic_lcs_convergence_1000.tikz}
 \caption{Convergence of closed $\lambda$-lines for increasing main-grid resolution in the double-gyre.
@@
 \setlength\figurewidth{.35\textwidth}
-\import{graphics/double_gyre/}{elliptic_repelling_lcs_details.tikz}
-\import{graphics/double_gyre/}{elliptic_attracting_lcs_details.tikz}
+\input{graphics/double_gyre/elliptic_repelling_lcs_details.tikz}
+\input{graphics/double_gyre/elliptic_attracting_lcs_details.tikz}
 \caption{

### 2014-12-10 main.tex
Fix formatting

--- 2014-12-10 main.tex
@@
 \multicolumn{2}{|p{.95\textwidth}|}
-{\lstinline![curvePosition,curveInitialPosition] = seed_curves_from_lambda_max(distance,cgEigenvalue,cgEigenvector,flowDomain,flowResolution)!}\tabularnewline
+{\lstinline![curvePosition,curveInitialPosition] = seed_curves_from_lambda_max(distance,cgEigenvalue,cgEigenvector,flowDomain, flowResolution)!}\tabularnewline
 \hline \hline

### 2014-12-11 main.tex
Remove unpublished reference
Remove & revise FIXMEs

--- 2014-12-10 main.tex
@@
 \end{gather*}
-% FIXME Clarify what is meant by _closed_ material lines
-obtained by averaging the Lagrangian shear arising over $[t_0,t]$ along closed material lines parametrized as $r(s)$ with $s \in [0,\sigma]$.
+obtained by averaging the Lagrangian shear arising over $[t_0,t]$ along material lines parametrized as $r(s)$ with $s \in [0,\sigma]$.
 More precisely, hyperbolic LCSs are stationary curves of $Q(\gamma)$ with respect to fixed-endpoint perturbations.
@@
 
-Solutions to this variational problem turn out to be orbits of the $\xi_1$ or $\xi_2$ eigenvector field.
+Solutions to this variational problem turn out to be curves of the $\xi_1$ or $\xi_2$ eigenvector field.
 Repelling LCSs (\emph{shrink lines}) are obtained as trajectories of the differential equation
@@
 Shrink lines and stretch lines coincide with the null geodesics of the Lorentzian metric $h(u,v) = \left\langle u,D_{t_0}^t(r) v \right\rangle$.
-For this reason, we refer to the computation of hyperbolic LCSs as strongest normally-repelling or normally-attracting orbits of \cref{eq:etafields} as \emph{geodesic detection of hyperbolic LCSs}.
+For this reason, we refer to the computation of hyperbolic LCSs as strongest normally-repelling or normally-attracting curves of \cref{eq:etafields} as \emph{geodesic detection of hyperbolic LCSs}.
 \end{sloppypar}
@@
 \item Compute Cauchy-Green strain tensor invariants, \lstinline![v,d] = eig_cgStrain(derivative)!
+% FIXME lambda_line and poincare_closed_orbit have been removed from demo scripts. Now use poincare_closed_orbit_range and elliptic_lcs.
 \item Compute $\lambda$-lines, \lstinline![etaPos,etaNeg] = lambda_line(v,d,lambda)!
 \item Define Poincare sections, \lstinline!ps!
-% FIXME function call below does not reflect LCS Tool usage (e.g. poincare_close_orbit_multi)
 \item Detect elliptic LCSs, \lstinline!ellipticLcs = poincare_closed_orbit(etaPos,etaNeg,ps)!
@@
 
-% FIXME Need to update table to account for lambda range
+% FIXME lambda_line and poincare_closed_orbit have been removed from demo scripts. Now use poincare_closed_orbit_range and elliptic_lcs.
 \begin{table*}
@@
 
-% FIXME Clarify use of the term "orbit"
-As discussed in \cref{sec:Hyperbolic LCSs}, hyperbolic LCS positions at time $t_0$ are found as the strongest repelling orbits of the vector field \cref{eq:shrink line} (repelling LCSs), and strongest attracting orbits of the vector field \cref{eq:stretch line} (attracting LCSs).
+As discussed in \cref{sec:Hyperbolic LCSs}, hyperbolic LCS positions at time $t_0$ are found as the strongest repelling curves of the vector field \cref{eq:shrink line} (repelling LCSs), and strongest attracting curves of the vector field \cref{eq:stretch line} (attracting LCSs).
 By repulsion and attraction we mean a property of the LCS (as an evolving material line) under the flow map $F_{t_0}^t$.

### 2014-12-11 main.tex
Remove & revise FIXMEs
Conflicts:
	main.tex

--- 2014-12-10 main.tex
@@
 \end{gather*}
-% FIXME Clarify what is meant by _closed_ material lines
-obtained by averaging the Lagrangian shear arising over $[t_0,t]$ along closed material lines parametrized as $r(s)$ with $s \in [0,\sigma]$.
+obtained by averaging the Lagrangian shear arising over $[t_0,t]$ along material lines parametrized as $r(s)$ with $s \in [0,\sigma]$.
 More precisely, hyperbolic LCSs are stationary curves of $Q(\gamma)$ with respect to fixed-endpoint perturbations.
-We note that parabolic LCSs (Lagrangian jet cores) are also stationary curves of $Q(\gamma)$, but under variable endpoint perturbations (cf.
-\citet{farazmand14:_shearless}).
+We note that parabolic LCSs (Lagrangian jet cores) are also stationary curves of $Q(\gamma)$, but under variable endpoint perturbations (cf. \citet{farazmand14:_shearless}).
 
-Solutions to this variational problem turn out to be orbits of the $\xi_1$ or $\xi_2$ eigenvector field. Repelling LCSs (\emph{shrink lines}) are obtained as trajectories of the differential equation
+Solutions to this variational problem turn out to be curves of the $\xi_1$ or $\xi_2$ eigenvector field.
+Repelling LCSs (\emph{shrink lines}) are obtained as trajectories of the differential equation
 \begin{equation}
@@
 Shrink lines and stretch lines coincide with the null geodesics of the Lorentzian metric $h(u,v) = \left\langle u,D_{t_0}^t(r) v \right\rangle$.
-For this reason, we refer to the computation of hyperbolic LCSs as strongest normally-repelling or normally-attracting orbits of \cref{eq:etafields} as \emph{geodesic detection of hyperbolic LCSs}.
+For this reason, we refer to the computation of hyperbolic LCSs as strongest normally-repelling or normally-attracting curves of \cref{eq:etafields} as \emph{geodesic detection of hyperbolic LCSs}.
 
@@
 The Poincare section is to be oriented such that the first endpoint is close to the centre of the expected Lagrangian vortex, and the second endpoint is outside this vortex.
-There is no automated procedure implemented in LCS Tool for the positioning of Poincare sections (see however \citet{karrasch14:_autom_lagran} for a recently developed algorithm).
 Additionally, the number of lambda-lines launched from the Poincare section, \lstinline!poincareSection.numPoints!, must be defined.
@@
 
-% FIXME Citations without pages (karrasch14:_autom_lagran) do not list publication year
 \bibliographystyle{elsarticle-num-names} 

### 2014-12-12 Review Response.md
Convert Review Response to Markdown

Response to reviewer comments

Reviewer #2
===========

Comment #1

Typo corrected as suggested

Comment #2

Typo corrected as suggested

Comment #3

A table has been added with benchmarks on memory use.

Comment #4

Typo corrected as suggested

Comment #5

Figure caption modified to explain notation of axes

Comment #6

Figure captions modified to explain labelling of lambda extrema and figures have been regenerated to improve the appearance of initial conditions

Comment #7

Typo corrected and sentence modified to improve grammar

Comment #8

Code listings have been revised and checked for poor breaks across lines and pages

Reviewer #3
===========

Comment #1

LCS Tool functions overview table added at beginning of numerical methods section.

Comment #2

Paragraph added at the beginning of Examples section to explain LCS Tool functions balance between performance and ease of use. Give approximate run-times for demo scripts.

Comment #3

The number of LCS detection software packages in the list of other LCS packages has been reduced and more details have been given. The “all-purpose” description of LCS Tool has been removed and the paragraph about LCS Tool's capabilities reworded.

Comment #4

A link to our article has been added to the LCS Tool README file.

Comment #5

The Highlights have been revised. The core contribution of the article is to present LCS Tool, a software tool to enable easy geodesic LCS computation. We believe the highlights explain this.

Reviewer #4
===========

The LCS Tool source code has been uploaded to RunMyCode, as suggested by the editor.

Reviewer #5
===========

Comment #1 and #2

The abstract and introduction have been revised based on the review comments.

Comment #3 and #4

Code listings have been revised and shortened.

Comment #6

LCS Tool functions overview table added at beginning of numerical methods section.

### 2014-12-12 Review Response.md
Address FIXMEs and reviewer comment

@@
 The number of LCS detection software packages in the list of other LCS packages has been reduced and more details have been given.
-The “all-purpose” description of LCS Tool has been removed and the paragraph about LCS Tool's capabilities reworded.
+The "all-purpose" description of LCS Tool has been removed and the paragraph about LCS Tool's capabilities reworded.
 
@@
 
+Comment # 5
+
+Applications of LCSs are cited in the introduction.
+Applications of the software 'LCS Tool' specifically are shown in the example part of the article.
+A manuscript about the detection of elliptic LCSs with LCS Tool in large scale ocean data is in preparation (Huhn et al., Coherent Lagrangian eddies and coherent transport in the Agulhas leakage).
+
 Comment #6

### 2014-12-12 main.tex
Address reviewer #2, comment #6
Address FIXMEs and reviewer comment

@@
 \begin{sloppypar}
-% FIXME Reviewer #5, comment #5: Please provide information regarding practical applications of this algorithmic software tool.
 Lagrangian Coherent Structures (LCSs) are evolving organizing centres of trajectory patterns in non-autonomous dynamical systems\citep{haller00:_lagran,peacock13:_lagran,haller15:_langr_coher_struc}.
@@
 \item Compute Cauchy-Green strain tensor invariants, \lstinline![v,d] = eig_cgStrain(derivative)!
-% FIXME lambda_line and poincare_closed_orbit have been removed from demo scripts. Now use poincare_closed_orbit_range and elliptic_lcs.
-\item Compute $\lambda$-lines, \lstinline![etaPos,etaNeg] = lambda_line(v,d,lambda)!
+\item Define a range of $\lambda$ values, \lstinline!lambda!
 \item Define Poincare sections, \lstinline!ps!
-\item Detect elliptic LCSs, \lstinline!ellipticLcs = poincare_closed_orbit(etaPos,etaNeg,ps)!
-\item Define a local maximization/minimization distance for hyperbolic LCSs \lstinline!lmd!
-% FIXME function call below does not reflect LCS Tool usage (e.g. remove_strain_in_shear)
-% FIXME seed_curves_from_lambda_max is a poor name for a function that detects hyperbolic LCSs
-\item Detect hyperbolic LCSs, \lstinline!hyperbolicLcs = seed_curves_from_lambda_max(v,d,ellipticLcs,lmd)!
+\item Detect elliptic LCSs, \lstinline!closedLambdaLine = poincare_closed_orbit_range(v,d,lambda,ps)! and \lstinline!ellipticLcs = elliptic_lcs(closedLambdaLine)!
+\item Define a local maximization/minimization distance for hyperbolic LCSs, \lstinline!lmd!
+\item Detect hyperbolic LCSs, \lstinline!hyperbolicLine = seed_curves_from_lambda_max(v,d,ellipticLcs,lmd)! and \lstinline!hyperbolicLcs = remove_strain_in_elliptic( hyperbolicLine,ellipticLcs)!
 \end{enumerate}
@@
 \end{center}
-\caption{Runtime and memory use associated with vector integration at different resolutions.
-Results are from the function \lstinline!eig_cgStrain!
-applied to the double gyre presented in \cref{sec:double gyre}.
-Results obtained with an Intel Core i3-4130 processor.}
+\caption{
+Runtime and memory use associated with vector integration at different resolutions.
+Results are from the function \lstinline!eig_cgStrain! applied to the double gyre presented in \cref{sec:double gyre}.
+Results obtained with an Intel Core i3-4130 3.4 GHz processor.
+}
 \label{t:vector integration memory}
@@
 
-% FIXME lambda_line and poincare_closed_orbit have been removed from demo scripts. Now use poincare_closed_orbit_range and elliptic_lcs.
 \begin{table*}
@@
 \end{tabular}
-\caption{Syntax for LCS Tool elliptic LCS functions.}
+\caption{
+Syntax for LCS Tool elliptic LCS functions.
+Both functions are called by the function \lstinline!poincare_closed_orbit_range!.
+}
 \label{t:Elliptic LCS functions}
@@
 \end{center}
-\caption{Runtime of some demo scripts Results obtained with an Intel Core i3-4130 processor with eight gigabytes of memory.}
+\caption{
+Runtime of some demo scripts.
+Results obtained with an Intel Core i3-4130 3.4 GHz processor with 8 GB of memory.
+}
 \label{t:demo runtimes}
@@
 Closed orbits for all predefined $\lambda$ values are computed and the outermost closed orbit is kept as the Lagrangian vortex boundary  (\crefrange{ll:double gyre lambda range}{ll:double gyre elliptic_lcs}).
-% FIXME Small tolerances (i.e. strict tolerances) should make convergence difficult not easy, so the word "facilitate" in the sentence below seems illogical.
-We have chosen small values for the error tolerances of the integration of the Cauchy-Green strain tensor (\cref{ll:double gyre cgStrainOdeSolverOptions}) and $\lambda$-lines (\cref{ll:double gyre lambdaLineOdeSolverOptions}) to facilitate convergence.
+We have chosen small values for the error tolerances of the integration of the Cauchy-Green strain tensor (\cref{ll:double gyre cgStrainOdeSolverOptions}) and $\lambda$-lines (\cref{ll:double gyre lambdaLineOdeSolverOptions}) to ensure convergence.
 
@@
 For all tested resolutions, the $\lambda$ values of the outermost closed orbits match.
-% FIXME Should reduce lowest resolution used to demonstrate problems if selected resolution is too low
 This also suggests that the lowest tested resolution, $500 \times 250$, is sufficient to identify elliptic LCSs.
@@
 \centering
-% FIXME Reviewer #2, comment #6: Improve appearance of $lambda$ extrema (white dots)
 \setlength\figurewidth{.35\textwidth}
@@
 
+\begin{sloppypar}
 The data is given at a spatial resolution of $1/4\degree$ and a temporal resolution of 7 days.

### 2014-12-12 main.tex
Correct typo in Bickley jet section
Correct overfull warnings

@@
 
+\begin{sloppypar}
 Next we consider positions of material lines at time $t_0$ that prevail as most repelling or attracting material lines (or \emph{hyperbolic LCSs}) over a time interval $[t_0,t] \subset [t_-,t_+]$.
@@
 We note that parabolic LCSs (Lagrangian jet cores) are also stationary curves of $Q(\gamma)$, but under variable endpoint perturbations (cf. \citet{farazmand14:_shearless}).
+\end{sloppypar}
 
@@
 \od{\phi_1}{t} = \phi_2,\\
-\od{\phi_1}{t} = -0.1 \phi_2 - \phi_1^3 + 11 \cos(t),\\
+\od{\phi_2}{t} = -0.1 \phi_2 - \phi_1^3 + 11 \cos(t),\\
 f_{1,2}(t) = 2.625 \times 10^{-2} \phi_1(t/6.238 \times 10^5)

### 2014-12-12 listings/bickley_jet/elliptic_hyperbolic_lcs_subset.m

@@ -38,5 +38,5 @@ ps(5).endPosition = [1.65e7,1.5e6;1.5e7,2.6e6]; %@\label{ll:Bickley jet Poincare
 shrinkL = seed_curves_from_lambda_max(shrinkLMaxDistance, shrinkLMaxLength,cgV(:,2),cgD(:,1:2),domain,res, 'periodicBc',pBc);
 ...
 %% Hyperbolic stretch line LCSs
-stretchL = seed_curves_from_lambda_max(stretchLMaxDistance, stretchLMaxLength,-cgV(:,1),cgD(:,3:4),domain,res, 'periodicBc',pBc);
+stretchL = seed_curves_from_lambda_max( stretchLMaxDistance,stretchLMaxLength,-cgV(:,1), cgD(:,3:4),domain,res,'periodicBc',pBc);
 ...

### 2014-12-12 listings/ocean_dataset/elliptic_hyperbolic_lcs_subset.m

@@ -32,9 +32,9 @@ stretchLMaxLength = 20;
 stretchLMaxDistance = 4*gridSpace;
 ...  
 %% Cauchy-Green strain eigenvalues and eigenvectors
-[cgV,cgD] = eig_cgStrain(lDerivative,domain,res,time, 'eigenvalueFromMainGrid',cgVmg,'auxGridRelDelta', cgAgd); %@\label{ll:eig_cgStrain}
+[cgV,cgD] = eig_cgStrain(lDerivative,domain,res,time, 'eigenvalueFromMainGrid',cgVmg, 'auxGridRelDelta',cgAgd); %@\label{ll:eig_cgStrain}
 %% Elliptic LCSs
-[closedLlp,closedLln] = poincare_closed_orbit_range(domain,res,cgV,cgD, lambda,ps,'lambdaLineOdeSolverOptions',llOptions);%@\label{ll:ocean data poincare_closed_orbit_multi}
+[closedLlp,closedLln] = poincare_closed_orbit_range(domain,res,cgV,cgD, lambda,ps,'odeSolverOptions',llOptions);%@\label{ll:ocean data poincare_closed_orbit_multi}
 ...
 % Hyperbolic shrink line LCSs
 shrinkL = seed_curves_from_lambda_max( shrinkLMaxDistance,shrinkLMaxLength,cgV(:,2), cgD(:,1:2),domain,res); %@\label{ll:ocean data shrinkLineLcs}

### 2014-12-12 main.tex

--- 2014-12-12 main.tex
@@
-\documentclass[5p]{elsarticle}
+\documentclass{article}
 
 \usepackage{amsmath}
+\usepackage[uniquename=init,firstinits,backend=biber,sorting=none]{biblatex}
 \usepackage{commath}
+\usepackage{fontspec}
 \usepackage{gensymb}
 \usepackage{import}
-\usepackage[utf8]{inputenc}
 \usepackage{listings}
@@
   breaklines,
-  breakatwhitespace,
   frame=single,
@@
   backgroundcolor=\color{grey},
-  escapeinside={\%@}{\^^M},
-  mathescape
+  escapeinside={\%@}{\^^M}
 }
-
-% Workaround for bug in elsarticle-num-names.bst
-% Reference: http://www.tug.org/pipermail/tex-live/2012-August/032152.html
-\makeatletter
-\providecommand{\doi}[1]{%
-  \begingroup
-    \let\bibinfo\@secondoftwo
-    \urlstyle{rm}%
-    \href{http://dx.doi.org/#1}{%
-      doi:\discretionary{}{}{}%
-      \nolinkurl{#1}%
-    }%
-  \endgroup
-}
-\makeatother
 
@@
 
+Lagrangian Coherent Structures (LCSs) are evolving organizing centres of trajectory patterns in non-autonomous dynamical systems\parencite{haller00:_lagran,peacock13:_lagran,haller15:_langr_coher_struc}.
+Applications of LCSs include oceanic and atmospheric flows\parencite{beron-vera13:_objec_agulh,koh02:_hyper}, biological transport problems\parencite{wilson09:_lagran_reynol,tallapragada11:_lagran,huhn12:_south_indian_ocean_count_madag}, aeronautics\parencite{tang10:_accur_lagran_hong_kong_inter_airpor}, celestial mechanics\parencite{gawlik09:_lagran}, crowd dynamics\parencite{ali07:_lagran_partic_dynam_approac_crowd}, and aperiodically forced mechanical oscillators\parencite{hadjighasem13:_detec_kam}.
+
 \begin{sloppypar}
-Lagrangian Coherent Structures (LCSs) are evolving organizing centres of trajectory patterns in non-autonomous dynamical systems\citep{haller00:_lagran,peacock13:_lagran,haller15:_langr_coher_struc}.
-Applications of LCSs include oceanic and atmospheric flows\citep{beron-vera13:_objec_agulh,koh02:_hyper}, biological transport problems\citep{wilson09:_lagran_reynol,tallapragada11:_lagran,huhn12:_south_indian_ocean_count_madag}, aeronautics\citep{tang10:_accur_lagran_hong_kong_inter_airpor}, celestial mechanics\citep{gawlik09:_lagran}, crowd dynamics\citep{ali07:_lagran_partic_dynam_approac_crowd}, and aperiodically forced mechanical oscillators\citep{hadjighasem13:_detec_kam}.
+\textcite{haller01:_distin} proposed that ridges of the finite-time Lyapunov exponent (FTLE) are heuristic indicators of hyperbolic (i.e., repelling and attracting type) LCSs.
+A number of examples support this principle\parencite{peacock13:_lagran}.
+Equating FTLE ridges with LCSs, however, would create theoretical inconsistencies, as well as false positives and negatives in hyperbolic LCS detection\parencite{haller11:_lagran_coher_struc,norgard12:_secon_lagran_coher_struc}.
+In addition, the role of the FTLE field in the accurate detection of elliptic (vortex-type) and parabolic (jet-core type) LCSs has remained an open question (but see \textcite{beron-vera10:_invar_lagran}).
 \end{sloppypar}
 
-\citet{haller01:_distin} proposed that ridges of the finite-time Lyapunov exponent (FTLE) are heuristic indicators of hyperbolic (i.e., repelling and attracting type) LCSs.
-A number of examples support this principle\citep{peacock13:_lagran}.
-Equating FTLE ridges with LCSs, however, would create theoretical inconsistencies, as well as false positives and negatives in hyperbolic LCS detection\citep{haller11:_lagran_coher_struc,norgard12:_secon_lagran_coher_struc}.
-In addition, the role of the FTLE field in the accurate detection of elliptic (vortex-type) and parabolic (jet-core type) LCSs has remained an open question (but see \citet{beron-vera10:_invar_lagran}).
-
-More recent work has focused on an exact mathematical formulation of the properties defining LCSs\citep{haller11:_lagran_coher_struc,farazmand12:_comput_lagran,haller12:_geodes_theor_trans_barrier_two_dimen_flows,haller13:_coher_lagran,haller14:_adden_coher_lagran,farazmand13:_attrac_lagran,blazevski14:_hyper_ellip_trans_barrier_three}.
-In two-dimensional flows, hyperbolic and parabolic LCSs turn out to be stationary curves of the averaged material shear\citep{farazmand14:_shearless}, whereas elliptic LCSs are stationary curves of the averaged strain\citep{haller13:_coher_lagran,haller14:_adden_coher_lagran}.
+More recent work has focused on an exact mathematical formulation of the properties defining LCSs\parencite{haller11:_lagran_coher_struc,farazmand12:_comput_lagran,haller12:_geodes_theor_trans_barrier_two_dimen_flows,haller13:_coher_lagran,haller14:_adden_coher_lagran,farazmand13:_attrac_lagran,blazevski14:_hyper_ellip_trans_barrier_three}.
+In two-dimensional flows, hyperbolic and parabolic LCSs turn out to be stationary curves of the averaged material shear\parencite{farazmand14:_shearless}, whereas elliptic LCSs are stationary curves of the averaged strain\parencite{haller13:_coher_lagran,haller14:_adden_coher_lagran}.
 These variational formulations lead to explicit solutions for LCSs as null geodesics of appropriate Lorentzian metrics.
 
-\begin{sloppypar}
 We present an algorithmic introduction to geodesic LCS detection.
-We then review the implementation of this approach in a computational engine called LCS Tool\footnote{LCS Tool is available at: \url{http://www.runmycode.org/companion/view/908}}.
+We then review the implementation of this approach in a computational engine called LCS Tool\footnote{LCS Tool is available at: \url{http://github.com/jeixav/LCS-Tool}}.
 This engine is a library of MATLAB functions that extract LCSs from two-dimensional unsteady flows. The examples we present form demonstration scripts distributed with LCS Tool.
-\end{sloppypar}
 
 Other software for LCS detection exists.
-ManGen\citep{lekien03:_time} calculates the FTLE and advects material curves in two-dimensional velocity fields.
+ManGen\parencite{lekien03:_time} calculates the FTLE and advects material curves in two-dimensional velocity fields.
 It includes a graphical user interface and uses the Message Passing Interface standard for parallel calculations.
-Newman\citep{toit10:_trans} calculates the FTLE using dimension independent code.
+Newman\parencite{toit10:_trans} calculates the FTLE using dimension independent code.
 It assists ridge extraction of FTLE fields and supports analytic and dataset velocity definitions.
-FlowTK\citep{ameli14:_devel_effic_flexib_pipel_lagran} calculates the FTLE in two and three-dimensions on Cartesian and unstructured grids.
+FlowTK\parencite{ameli14:_devel_effic_flexib_pipel_lagran} calculates the FTLE in two and three-dimensions on Cartesian and unstructured grids.
 The NVIDIA CUDA parallel computing platform is used for fast computations and Kitware's ParaView data analysis and visualization application is used for a user interface.
@@
 More precisely, hyperbolic LCSs are stationary curves of $Q(\gamma)$ with respect to fixed-endpoint perturbations.
-We note that parabolic LCSs (Lagrangian jet cores) are also stationary curves of $Q(\gamma)$, but under variable endpoint perturbations (cf. \citet{farazmand14:_shearless}).
+We note that parabolic LCSs (Lagrangian jet cores) are also stationary curves of $Q(\gamma)$, but under variable endpoint perturbations (cf. \textcite{farazmand14:_shearless}).
 
@@
 
+\begin{sloppypar}
 Shrink lines and stretch lines coincide with the null geodesics of the Lorentzian metric $h(u,v) = \left\langle u,D_{t_0}^t(r) v \right\rangle$.
 For this reason, we refer to the computation of hyperbolic LCSs as strongest normally-repelling or normally-attracting curves of \cref{eq:etafields} as \emph{geodesic detection of hyperbolic LCSs}.
+\end{sloppypar}
 
@@
 \item Define a local maximization/minimization distance for hyperbolic LCSs, \lstinline!lmd!
-\item Detect hyperbolic LCSs, \lstinline!hyperbolicLine = seed_curves_from_lambda_max(v,d,ellipticLcs,lmd)! and \lstinline!hyperbolicLcs = remove_strain_in_elliptic( hyperbolicLine,ellipticLcs)!
+% FIXME seed_curves_from_lambda_max is a poor name for a function that detects hyperbolic LCSs
+\item Detect hyperbolic LCSs, \lstinline!hyperbolicLine = seed_curves_from_lambda_max(v,d,ellipticLcs,lmd)! and \lstinline!hyperbolicLcs = remove_strain_in_elliptic(hyperbolicLine,ellipticLcs)!
 \end{enumerate}
@@
 Results are from the function \lstinline!eig_cgStrain! applied to the double gyre presented in \cref{sec:double gyre}.
-Results obtained with an Intel Core i3-4130 3.4 GHz processor.
+Results obtained with an Intel Core i3-4130 processor.
 }
@@
 
+% FIXME lambda_line and poincare_closed_orbit have been removed from demo scripts. Now use poincare_closed_orbit_range and elliptic_lcs.
 \begin{table*}
@@
 The Poincare section is to be oriented such that the first endpoint is close to the centre of the expected Lagrangian vortex, and the second endpoint is outside this vortex.
+% FIXME Move citation to conclusions, as a candidate for future work.
+There is no automated procedure implemented in LCS Tool for the positioning of Poincare sections (see however \textcite{karrasch14:_autom_lagran} for a recently developed algorithm).
 Additionally, the number of lambda-lines launched from the Poincare section, \lstinline!poincareSection.numPoints!, must be defined.
@@
 \begin{center}
-\begin{tabular}{|p{.3\textwidth}|c|}
+\begin{tabular}{|c|c|}
 \hline
@@
 \hline \hline
-double gyre & 50\tabularnewline
+double\_gyre/elliptic\_hyperbolic\_lcs & 50\tabularnewline
 \hline
-Bickley jet & 975\tabularnewline
+bickley\_jet/elliptic\_hyperbolic\_lcs\_details & 975\tabularnewline
 \hline
-ocean dataset & 127\tabularnewline
+ocean\_dataset/elliptic\_hyperbolic\_lcs\_details & 127\tabularnewline
 \hline
@@
 Runtime of some demo scripts.
-Results obtained with an Intel Core i3-4130 3.4 GHz processor with 8 GB of memory.
+Results obtained with an Intel Core i3-4130 processor with eight gigabytes of memory.
 }
@@
 The free stretching parameter $\lambda$ (\cref{eq:etafields}) is varied over the range $[0.93,1.07]$ with increments of $0.01$ (\cref{ll:double gyre lambdaRange}).
-Closed orbits for all predefined $\lambda$ values are computed and the outermost closed orbit is kept as the Lagrangian vortex boundary  (\crefrange{ll:double gyre lambda range}{ll:double gyre elliptic_lcs}).
+Closed orbits for all $\lambda$ values are computed and the outermost closed orbit is kept as the Lagrangian vortex boundary  (\crefrange{ll:double gyre lambda range}{ll:double gyre elliptic_lcs}).
 We have chosen small values for the error tolerances of the integration of the Cauchy-Green strain tensor (\cref{ll:double gyre cgStrainOdeSolverOptions}) and $\lambda$-lines (\cref{ll:double gyre lambdaLineOdeSolverOptions}) to ensure convergence.
@@
 For all tested resolutions, the $\lambda$ values of the outermost closed orbits match.
+% FIXME Should reduce lowest resolution used to demonstrate problems if selected resolution is too low
 This also suggests that the lowest tested resolution, $500 \times 250$, is sufficient to identify elliptic LCSs.
@@
 \centering
-\setlength\figurewidth{.4\textwidth}
+\setlength\figurewidth{.875\textwidth}
 \import{graphics/double_gyre/}{elliptic_lcs_convergence_500.tikz}
@@
 \centering
-\setlength\figurewidth{.35\textwidth}
+\setlength\figurewidth{.8\textwidth}
 \import{graphics/double_gyre/}{elliptic_repelling_lcs_details.tikz}
@@
 The integration time is $T = 4L_x/U$, an integer multiple of the eddy turnover time.
-\Cref{l:Bickley jet elliptic hyperbolic LCS subset} shows the LCS Tool commands in which the chaotically perturbed velocity is defined (\cref{ll:Bickley jet velocity definition}), periodic boundary conditions are imposed in the x-direction (\cref{ll:Bickley jet periodicBc}), and five Poincare sections are defined where we expect coherent vortices (\crefrange{ll:Bickley jet Poincare start}{ll:Bickley jet Poincare end}).
+\Cref{l:Bickley jet elliptic hyperbolic LCS subset} shows the LCS Tool commands in which the chaotically perturbed velocity is defined (\crefrange{ll:Bickley jet velocity definition start}{ll:Bickley jet velocity definition end}), periodic boundary conditions are imposed in the x-direction (\cref{ll:Bickley jet periodicBc}), and five Poincare sections are defined where we expect coherent vortices (\crefrange{ll:Bickley jet Poincare start}{ll:Bickley jet Poincare end}).
 $\lambda$-values for closed orbit detection are varied over the range $[0.80,1.20]$ with a step of $0.01$.
@@
 
-We choose the integration time as $T=30$ days (\cref{l:ocean data elliptic hyperbolic lcs}, \cref{ll:ocean data timespan}), which is larger than the eddy turnover time in this region.
-The resolution of the main computational grid for initial conditions is set to $400 \times 400$ (\cref{ll:ocean data resolution}).
+We choose the integration time as $T=30$ days (\cref{l:ocean data elliptic hyperbolic lcs}, \cref{ll:ocean timespan}), which is larger than the eddy turnover time in this region.
+The resolution of the main computational grid for initial conditions is set to $400 \times 400$ (\cref{ll:ocean resolution}).
 This corresponds to a resolution of roughly 0.015°.
@@
 The flow is integrated and the Cauchy-Green strain tensor is computed by the function \lstinline!eig_cgStrain! (\cref{ll:eig_cgStrain}).
-The auxiliary grid distance is set to 1\% of the main grid distance (\cref{ll:cgAuxGridRelDelta}), and eigenvalues are computed from the auxiliary grid (\cref{ll:cgEigenvalueFromMainGrid}).
-Elliptic LCSs are computed in \cref{ll:ocean data poincare_closed_orbit_multi}, after the Poincare sections have been set (\cref{ll:ocean data poincareSection(1)}).
-$\lambda$-values are varied over a range of $[0.90,1.10]$ with a step of $0.02$ (\cref{ll:ocean data lambdaRange}).
-Hyperbolic LCSs are computed in \cref{ll:ocean data shrinkLineLcs,ll:ocean data stretchLineLcs}.
+Incompressibility of the flow is enforced (\cref{ll:incompressible}), the auxiliary grid distance is set to 1\% of the main grid distance (\cref{ll:cgAuxGridRelDelta}), and eigenvalues are computed from the auxiliary grid (\cref{ll:cgEigenvalueFromMainGrid}).
+Elliptic LCSs are computed in \crefrange{ll:ocean elliptic LCS start}{ll:ocean elliptic LCS end}, after the Poincare sections have been set (\crefrange{ll:ocean data poincareSection(1)}{ll:ocean data poincareSection(5)}).
+$\lambda$-values are varied over a range of $[0.90,1.10]$ with a step of $0.02$ (\cref{ll:ocean lambda}).
+Hyperbolic LCSs are computed in \crefrange{ll:ocean hyperbolic start}{ll:ocean hyperbolic end}.
 
-\begin{sloppypar}
-\Cref{f:ocean dataset LCS} shows elliptic and hyperbolic LCSs on 22 November 2006, the same time as analysed in \citet{haller13:_coher_lagran,haller14:_adden_coher_lagran,beron-vera13:_objec_agulh}.
+\Cref{f:ocean dataset LCS} shows elliptic and hyperbolic LCSs on 22 November 2006, the same time as analysed in \textcite{haller13:_coher_lagran,haller14:_adden_coher_lagran,beron-vera13:_objec_agulh}.
 In those references, the integration time is 90 days.
@@
 LCS Tool leverages the capabilities of MATLAB.
-For FTLE based extraction of LCSs, computational performance has received considerable attention \citep{conti12:_gpu_apu_finit_time_lyapun_expon,miron12:_anisot_lagran_coher_struc}.
+For FTLE based extraction of LCSs, computational performance has received considerable attention \parencite{conti12:_gpu_apu_finit_time_lyapun_expon,miron12:_anisot_lagran_coher_struc}.
 We think LCS Tool can facilitate similar computational advances for variational LCS methods.
-Optimizing computational performance will aid applications to large-scale forecasting applications, such as the tracking of environmental contaminants\citep{olascoaga12:_forec}.
+Optimizing computational performance will aid applications to large-scale forecasting applications, such as the tracking of environmental contaminants\parencite{olascoaga12:_forec}.
 
 We think LCS Tool can serve as a foundation for the numerical implementation of recent theoretical advances.
-These include the geodesic theory of parabolic LCSs (jet cores)\citep{farazmand14:_shearless} and the variational theory of hyperbolic and elliptic LCSs for three-dimensional flows\citep{blazevski14:_hyper_ellip_trans_barrier_three}.
+These include the geodesic theory of parabolic LCSs (jet cores)\parencite{farazmand14:_shearless} and the variational theory of hyperbolic and elliptic LCSs for three-dimensional flows\parencite{blazevski14:_hyper_ellip_trans_barrier_three}.
 
@@
 
-\begin{sloppypar}
 The altimeter products used in this work are produced by SSALTO/DUACS and distributed by AVISO, with support from CNES, \url{http://www.aviso.oceanobs.com/duacs}.
-\end{sloppypar}
 
-\section*{References}
-
-\bibliographystyle{elsarticle-num-names} 
-\bibliography{main}
+% FIXME Citations without pages (karrasch14:_autom_lagran) do not list publication year
+\printbibliography
 

### 2014-12-12 main.tex
Shorten unambiguous variable names

@@
 \setlength\figurewidth{.75\textwidth}
-\input{graphics/double_gyre/poincare_return_map.tikz}
+\import{graphics/double_gyre/}{poincare_return_map}
 \end{center}
@@
 \setlength\figurewidth{.875\textwidth}
-\import{graphics/double_gyre/}{elliptic_lcs_convergence_500.tikz}
-\import{graphics/double_gyre/}{elliptic_lcs_convergence_750.tikz}
-\import{graphics/double_gyre/}{elliptic_lcs_convergence_1000.tikz}
+\import{graphics/double_gyre/}{elliptic_lcs_500}
+\import{graphics/double_gyre/}{elliptic_lcs_750}
+\import{graphics/double_gyre/}{elliptic_lcs_1000}
 \caption{Convergence of closed $\lambda$-lines for increasing main-grid resolution in the double-gyre.
@@
 \setlength\figurewidth{.8\textwidth}
-\import{graphics/double_gyre/}{elliptic_repelling_lcs_details.tikz}
-\import{graphics/double_gyre/}{elliptic_attracting_lcs_details.tikz}
+\import{graphics/double_gyre/}{elliptic_repelling_lcs}
+\import{graphics/double_gyre/}{elliptic_attracting_lcs}
 \caption{
@@
 \setlength\figurewidth{.775\textwidth}
-\import{graphics/bickley_jet/}{elliptic_repelling_lcs_details.tikz}
-\import{graphics/bickley_jet/}{elliptic_attracting_lcs_details.tikz}
+\import{graphics/bickley_jet/}{elliptic_repelling_lcs}
+\import{graphics/bickley_jet/}{elliptic_attracting_lcs}
 \caption{
@@
 \setlength\figurewidth{.85\textwidth}
-\import{graphics/ocean_dataset/}{elliptic_repelling_lcs.tikz}
-\import{graphics/ocean_dataset/}{elliptic_attracting_lcs.tikz}
+\import{graphics/ocean_dataset/}{elliptic_repelling_lcs}
+\import{graphics/ocean_dataset/}{elliptic_attracting_lcs}
 \end{center}

### 2014-12-12 graphics/bickley_jet/elliptic_hyperbolic_lcs.m

@@ -1,4 +1,4 @@
-function elliptic_hyperbolic_lcs_details
+function elliptic_hyperbolic_lcs
 
 %% Input parameters
 u = 62.66;
@@ -79,7 +79,7 @@ set(hPoincareSection,'marker','o')
 set(hPoincareSection,'MarkerFaceColor',ellipticColor)
 set(hPoincareSection,'MarkerEdgeColor','w')
 
-[closedLambdaLinePos,closedLambdaLineNeg] = poincare_closed_orbit_range(domain,resolution,cgEigenvector,cgEigenvalue,lambda,poincareSection,'forceEtaComplexNaN',forceEtaComplexNaN,'lambdaLineOdeSolverOptions',lambdaLineOdeSolverOptions,'periodicBc',periodicBc);
+[closedLambdaLinePos,closedLambdaLineNeg] = poincare_closed_orbit_range(domain,resolution,cgEigenvector,cgEigenvalue,lambda,poincareSection,'forceEtaComplexNaN',forceEtaComplexNaN,'odeSolverOptions',lambdaLineOdeSolverOptions,'periodicBc',periodicBc);
 
 ellipticLcs = elliptic_lcs(closedLambdaLinePos);
 ellipticLcs = [ellipticLcs,elliptic_lcs(closedLambdaLineNeg)];
@@ -120,7 +120,7 @@ uistack(hEllipticLcs,'top')
 uistack(hClosedLambdaLine,'top')
 uistack(hPoincareSection,'top')
 
-filename = 'elliptic_repelling_lcs_details.tikz';
+filename = 'elliptic_repelling_lcs.tex';
 hFigure = get(hAxes,'parent');
 matlab2tikz(filename,'showInfo',false,'width','\figurewidth','figurehandle',hFigure)
 
@@ -168,6 +168,6 @@ uistack(hEllipticLcs,'top')
 uistack(hClosedLambdaLine,'top')
 uistack(hPoincareSection,'top')
 
-filename = 'elliptic_attracting_lcs_details.tikz';
+filename = 'elliptic_attracting_lcs.tex';
 hFigure = get(hAxes,'parent');
 matlab2tikz(filename,'showInfo',false,'width','\figurewidth','figurehandle',hFigure)

### 2014-12-12 graphics/bickley_jet/startup.m

@@ -1,3 +1,2 @@
 addpath(fullfile('..','..','LCS-Tool','demo','bickley_jet'))
 addpath(fullfile('..','..','LCS-Tool'))
-addpath(fullfile('..','matlab2tikz','src'))

### 2014-12-12 graphics/double_gyre/elliptic_hyperbolic_lcs.m

@@ -1,4 +1,4 @@
-function elliptic_hyperbolic_lcs_details
+function elliptic_hyperbolic_lcs
 
 % Parameters
 epsilon = .1;
@@ -112,7 +112,7 @@ uistack(hEllipticLcs,'top')
 uistack(hClosedLambdaLine,'top')
 uistack(hPoincareSection,'top')
 
-filename = 'elliptic_repelling_lcs_details.tikz';
+filename = 'elliptic_repelling_lcs.tex';
 if ~verLessThan('matlab','8.4')
     warning([mfilename,':matlabVersion'],'matlab2tikz colormap problem with MATLAB newer than R2014a')
 end
@@ -159,5 +159,5 @@ uistack(hEllipticLcs,'top')
 uistack(hClosedLambdaLine,'top')
 uistack(hPoincareSection,'top')
 
-filename = 'elliptic_attracting_lcs_details.tikz';
+filename = 'elliptic_attracting_lcs.tex';
 matlab2tikz(filename,'showInfo',false,'relativeDataPath',fullfile('graphics','double_gyre'),'width','\figurewidth','figurehandle',hFigure,'colormap',colormap(hAxes))

### 2014-12-12 graphics/double_gyre/elliptic_lcs.m

@@ -1,4 +1,4 @@
-function elliptic_lcs_convergence
+function elliptic_lcs
 
 % Parameters
 epsilon = .1;
@@ -79,6 +79,6 @@ for m = 1:numel(resolutionX)
         end
     end
     
-    filename = strcat('elliptic_lcs_convergence_',num2str(lResolutionX),'.tikz');
+    filename = strcat('elliptic_lcs_convergence_',num2str(lResolutionX),'.tex');
     matlab2tikz(filename,'showInfo',false,'width','\figurewidth','figurehandle',hFigure)
 end

### 2014-12-12 graphics/double_gyre/poincare_return_map.m

@@ -42,5 +42,5 @@ set(findobj(hAxes,'Marker','o','MarkerFaceColor','r'),'MarkerFaceColor','k')
 set(findobj(hAxes,'Marker','o','MarkerFaceColor','b'),'MarkerFaceColor','none')
 
 hFigure = get(hAxes,'parent');
-filename = strcat('poincare_return_map.tikz');
+filename = strcat('poincare_return_map.tex');
 matlab2tikz(filename,'showInfo',false,'width','\figurewidth','figurehandle',hFigure)

### 2014-12-12 graphics/double_gyre/startup.m

@@ -1,3 +1,2 @@
 addpath(fullfile('..','..','LCS-Tool','demo','double_gyre'))
 addpath(fullfile('..','..','LCS-Tool'))
-addpath(fullfile('..','matlab2tikz','src'))

### 2014-12-12 graphics/ocean_dataset/elliptic_hyperbolic_lcs.m

@@ -106,7 +106,7 @@ set(hStrainlineLcs,'color',repellingColor)
 uistack(hEllipticLcs,'top')
 uistack(hPoincareSection,'top')
 
-filename = 'elliptic_repelling_lcs.tikz';
+filename = 'elliptic_repelling_lcs.tex';
 hFigure = get(hAxes,'parent');
 matlab2tikz(filename,'showInfo',false,'width','\figurewidth','figurehandle',hFigure)
 
@@ -145,6 +145,6 @@ set(hStretchlineLcs,'color',attractingColor)
 uistack(hEllipticLcs,'top')
 uistack(hPoincareSection,'top')
 
-filename = 'elliptic_attracting_lcs.tikz';
+filename = 'elliptic_attracting_lcs.tex';
 hFigure = get(hAxes,'parent');
 matlab2tikz(filename,'showInfo',false,'width','\figurewidth','figurehandle',hFigure)

### 2014-12-12 main.tex
Add FIXME comments

@@
 We give an algorithmic introduction to Lagrangian coherent structures (LCSs) using a newly developed computational engine, LCS Tool.
+% FIXME "centrepiece" is vague and may confuse with centre fixed point in classical dynamical systems
 LCSs are most repelling, attracting and shearing material lines that form the centrepieces of observed tracer patterns in two-dimensional unsteady dynamical systems.
@@
 
+% FIXME "Centres" may be associated with center fixed points in the context of classical dynamical systems
 Lagrangian Coherent Structures (LCSs) are evolving organizing centres of trajectory patterns in non-autonomous dynamical systems\parencite{haller00:_lagran,peacock13:_lagran,haller15:_langr_coher_struc}.
@@
 \subsection{Elliptic LCSs}
-\label{sec:Elliptic LCSs}
+\label{s:Elliptic LCSs}
 
@@
 \subsection{Hyperbolic LCSs}
-\label{sec:Hyperbolic LCSs}
+\label{s:Hyperbolic LCSs}
 
@@
 Runtime and memory use associated with vector integration at different resolutions.
-Results are from the function \lstinline!eig_cgStrain! applied to the double gyre presented in \cref{sec:double gyre}.
+Results are from the function \lstinline!eig_cgStrain! applied to the double gyre presented in \cref{s:double gyre}.
 Results obtained with an Intel Core i3-4130 processor.
@@
 This requires spatial and temporal interpolation that enables the evaluation of the velocity function at arbitrary points in $U$ and at arbitrary times between $t_-$ and $t_+$.
-In \cref{sec:oceandataset}, we present an ocean dataset example with details of possible interpolation functions.
+In \cref{s:oceandataset}, we present an ocean dataset example with details of possible interpolation functions.
 
@@
 
-As discussed in \cref{sec:Elliptic LCSs}, positions of elliptic LCSs at time $t_0$ are found as closed orbits of the $\eta_\pm^\lambda$ vector fields defined in \cref{eq:eta}.
+As discussed in \cref{s:Elliptic LCSs}, positions of elliptic LCSs at time $t_0$ are found as closed orbits of the $\eta_\pm^\lambda$ vector fields defined in \cref{eq:eta}.
 We find such orbits by integrating \cref{eq:etafields} from points of an appropriately chosen section (Poincare section), and evaluating the first return map (Poincare map) onto this section.
@@
 
-As discussed in \cref{sec:Hyperbolic LCSs}, hyperbolic LCS positions at time $t_0$ are found as the strongest repelling curves of the vector field \cref{eq:shrink line} (repelling LCSs), and strongest attracting curves of the vector field \cref{eq:stretch line} (attracting LCSs).
+As discussed in \cref{s:Hyperbolic LCSs}, hyperbolic LCS positions at time $t_0$ are found as the strongest repelling curves of the vector field \cref{eq:shrink line} (repelling LCSs), and strongest attracting curves of the vector field \cref{eq:stretch line} (attracting LCSs).
 By repulsion and attraction we mean a property of the LCS (as an evolving material line) under the flow map $F_{t_0}^t$.
@@
 \end{enumerate}
-\caption{Algorithm to calculate initial positions of repelling LCSs at time $t_0$. The algorithm for attracting LCSs is similar.}
+\caption{Algorithm to calculate initial positions of repelling LCSs at time $t_0$.
+The algorithm for attracting LCSs is similar.}
 \label{t:Hyperbolic LCS algorithm}
@@
 \subsection{Double gyre}
-\label{sec:double gyre}
+\label{s:double gyre}
 
@@
 
-\lstinputlisting[caption={LCS Tool commands for double gyre elliptic LCSs. Abridged subset of the LCS Tool script \lstinline!demo/double_gyre/elliptic_hyperbolic_lcs.m!},
-                 label=l:double gyre elliptic LCS]{listings/double_gyre/elliptic_lcs_subset.m}
+\lstinputlisting[
+caption={LCS Tool commands for double gyre elliptic LCSs.
+Abridged subset of the LCS Tool script \lstinline!demo/double_gyre/elliptic_hyperbolic_lcs.m!},
+label=l:double gyre elliptic LCS]
+{listings/double_gyre/elliptic_lcs_subset.m}
 
@@
 
-As usual, the local maximization distance is set larger for stretch lines than for shrink lines (cf.
-\cref{ll:shrinkLineLocalMaxDistance} and \cref{ll:stretchLineLocalMaxDistance}).
+As usual, the local maximization distance is set larger for stretch lines than for shrink lines (cf. \cref{ll:shrinkLineLocalMaxDistance} and \cref{ll:stretchLineLocalMaxDistance}).
 The purpose of the maximization distance is to obtain spatially separated LCSs and to avoid a dense tangle of lines that basically indicates the same hyperbolic LCS (cf.
@@
 \lstinputlisting[
-        caption={LCS Tool commands for double gyre hyperbolic LCSs. Abridged subset of the LCS Tool script \lstinline!demo/double_gyre/elliptic_hyperbolic_lcs.m!},
+caption={LCS Tool commands for double gyre hyperbolic LCSs.
+Abridged subset of the LCS Tool script \lstinline!demo/double_gyre/elliptic_hyperbolic_lcs.m!},
 label=l:double gyre lambda strain stretch LCS]
@@
 \lstinputlisting[
-        caption={LCS Tool commands for Bickley jet elliptic and hyperbolic LCSs. Abridged version of the LCS Tool script \lstinline!demo/bickley_jet/elliptic_hyperbolic_lcs.m!},
+caption={LCS Tool commands for Bickley jet elliptic and hyperbolic LCSs.
+Abridged version of the LCS Tool script \lstinline!demo/bickley_jet/elliptic_hyperbolic_lcs.m!},
 label=l:Bickley jet elliptic hyperbolic LCS subset]
@@
 In contrast to the previous two analytic examples, the velocity field is available only with discrete temporal and spatial resolution.
-Our region of interest is a small domain in the South Atlantic Ocean, where exceptionally coherent eddies, Agulhas rings, were recently found by \textcite{haller13:_coher_lagran,haller14:_adden_coher_lagran} using the theory we surveyed in \cref{sec:Elliptic LCSs}.
+Our region of interest is a small domain in the South Atlantic Ocean, where exceptionally coherent eddies, Agulhas rings, were recently found by \textcite{haller13:_coher_lagran,haller14:_adden_coher_lagran} using the theory we surveyed in \cref{s:Elliptic LCSs}.
 

### 2014-12-16 main.tex
Correct reference label
Remove overfull box warning
Update Bickley jet demo parameters
Conflicts:
	graphics/bickley_jet/elliptic_hyperbolic_lcs_details.m
	main.tex

--- 2014-12-12 main.tex
@@
-\documentclass{article}
+\documentclass[5p]{elsarticle}
 
 \usepackage{amsmath}
-\usepackage[uniquename=init,firstinits,backend=biber,sorting=none]{biblatex}
 \usepackage{commath}
-\usepackage{fontspec}
 \usepackage{gensymb}
 \usepackage{import}
+\usepackage[utf8]{inputenc}
 \usepackage{listings}
@@
   breaklines,
+  breakatwhitespace,
   frame=single,
@@
   backgroundcolor=\color{grey},
-  escapeinside={\%@}{\^^M}
+  escapeinside={\%@}{\^^M},
+  mathescape
 }
+
+% Workaround for bug in elsarticle-num-names.bst
+% Reference: http://www.tug.org/pipermail/tex-live/2012-August/032152.html
+\makeatletter
+\providecommand{\doi}[1]{%
+  \begingroup
+    \let\bibinfo\@secondoftwo
+    \urlstyle{rm}%
+    \href{http://dx.doi.org/#1}{%
+      doi:\discretionary{}{}{}%
+      \nolinkurl{#1}%
+    }%
+  \endgroup
+}
+\makeatother
 
@@
 
-\addbibresource{main.bib}
+\journal{Journal of Computational Science}
+
+\begin{document}
+
+\begin{frontmatter}
 
@@
 
-\author{Kristjan Onu \and Florian Huhn \and George Haller}
+\author{K.~Onu}
 
-\begin{document}
+\author{F.~Huhn}
 
-\numberwithin{equation}{section}
-\numberwithin{figure}{section}
-\numberwithin{lstlisting}{section}
-\numberwithin{table}{section}
+\author{G.~Haller\corref{cor}}
+\ead{georgehaller@ethz.ch}
 
-\maketitle
+\cortext[cor]{Corresponding author}
+
+\address{Institute of Mechanical Systems, ETH Zurich, Switzerland}
 
@@
 LCSs are most repelling, attracting and shearing material lines that form the centrepieces of observed tracer patterns in two-dimensional unsteady dynamical systems.
-LCS Tool implements the latest geodesic theory for LCSs in two-dimensional flows, uncovering key transport barriers in unsteady flow velocity data as explicit solutions of differential equations.
+LCS Tool implements the latest geodesic theory of LCSs for two-dimensional flows, uncovering key transport barriers in unsteady flow velocity data as explicit solutions of differential equations.
 LCS Tool makes theoretical results accessible to the fluid mechanics community since implementing these results directly could be time consuming.
-After a review of the underlying theory, we explain the numerical algorithms used by LCS Tool, and illustrate its capabilities on three unsteady fluid flow examples: a double gyre, a Bickley jet and an ocean dataset for the Agulhas current.
+After a review of the underlying theory, we explain the steps and numerical methods used by LCS Tool, and illustrate its capabilities on three unsteady fluid flow examples.
 \end{abstract}
 
-\tableofcontents
+\begin{keyword}
+Lagrangian coherent structures \sep Mathematical software \sep Non-autonomous dynamical systems \sep Invariant manifolds \sep Mixing \sep Transport barriers \sep Fluid dynamics \sep Ocean surface flows
+
+\PACS 47.10.Fg \sep 47.27.De \sep 47.11.-j \sep 05.45.-a \sep 47.51.+a \sep 47.54.-r \sep 92.10.hf \sep 93.30.Qn \sep 92.10.ab \sep 92.10.Ty \sep 92.10.ak \sep 92.10.A-
+
+\MSC[2010] 37N10 \sep 37B25 \sep 97N80 \sep 37C60 \sep 37B55 \sep 70H33 \sep 76F25
+\end{keyword}
+
+\end{frontmatter}
 
@@
 
-Lagrangian Coherent Structures (LCSs) are evolving organizing centres of trajectory patterns in non-autonomous dynamical systems\parencite{haller00:_lagran,peacock13:_lagran,haller15:_langr_coher_struc}.
-Applications of LCSs include oceanic and atmospheric flows\parencite{beron-vera13:_objec_agulh,koh02:_hyper}, biological transport problems\parencite{wilson09:_lagran_reynol,tallapragada11:_lagran,huhn12:_south_indian_ocean_count_madag}, aeronautics\parencite{tang10:_accur_lagran_hong_kong_inter_airpor}, celestial mechanics\parencite{gawlik09:_lagran}, crowd dynamics\parencite{ali07:_lagran_partic_dynam_approac_crowd}, and aperiodically forced mechanical oscillators\parencite{hadjighasem13:_detec_kam}.
+\begin{sloppypar}
+Lagrangian Coherent Structures (LCSs) are evolving organizing centres of trajectory patterns in non-autonomous dynamical systems\citep{haller00:_lagran,peacock13:_lagran,haller15:_langr_coher_struc}.
+Applications of LCSs include oceanic and atmospheric flows\citep{beron-vera13:_objec_agulh,koh02:_hyper}, biological transport problems\citep{wilson09:_lagran_reynol,tallapragada11:_lagran,huhn12:_south_indian_ocean_count_madag}, aeronautics\citep{tang10:_accur_lagran_hong_kong_inter_airpor}, celestial mechanics\citep{gawlik09:_lagran}, crowd dynamics\citep{ali07:_lagran_partic_dynam_approac_crowd}, and aperiodically forced mechanical oscillators\citep{hadjighasem13:_detec_kam}.
+\end{sloppypar}
+
+\citet{haller01:_distin} proposed that ridges of the finite-time Lyapunov exponent (FTLE) are heuristic indicators of hyperbolic (i.e., repelling and attracting type) LCSs.
+A number of examples support this principle\citep{peacock13:_lagran}.
+Equating FTLE ridges with LCSs, however, would create theoretical inconsistencies, as well as false positives and negatives in hyperbolic LCS detection\citep{haller11:_lagran_coher_struc,norgard12:_secon_lagran_coher_struc}.
+In addition, the role of the FTLE field in the accurate detection of elliptic (vortex-type) and parabolic (jet-core type) LCSs has remained an open question (but see \citet{beron-vera10:_invar_lagran}).
+
+More recent work has focused on an exact mathematical formulation of the properties defining LCSs\citep{haller11:_lagran_coher_struc,farazmand12:_comput_lagran,haller12:_geodes_theor_trans_barrier_two_dimen_flows,haller13:_coher_lagran,haller14:_adden_coher_lagran,farazmand13:_attrac_lagran,blazevski14:_hyper_ellip_trans_barrier_three}.
+In two-dimensional flows, hyperbolic and parabolic LCSs turn out to be stationary curves of the averaged material shear\citep{farazmand14:_shearless}, whereas elliptic LCSs are stationary curves of the averaged strain\citep{haller13:_coher_lagran,haller14:_adden_coher_lagran}.
+These variational formulations lead to explicit solutions for LCSs as null geodesics of appropriate Lorentzian metrics.
 
 \begin{sloppypar}
-\textcite{haller01:_distin} proposed that ridges of the finite-time Lyapunov exponent (FTLE) are heuristic indicators of hyperbolic (i.e., repelling and attracting type) LCSs.
-A number of examples support this principle\parencite{peacock13:_lagran}.
-Equating FTLE ridges with LCSs, however, would create theoretical inconsistencies, as well as false positives and negatives in hyperbolic LCS detection\parencite{haller11:_lagran_coher_struc,norgard12:_secon_lagran_coher_struc}.
-In addition, the role of the FTLE field in the accurate detection of elliptic (vortex-type) and parabolic (jet-core type) LCSs has remained an open question (but see \textcite{beron-vera10:_invar_lagran}).
+We present an algorithmic introduction to geodesic LCS detection.
+We then review the implementation of this approach in a computational engine called LCS Tool\footnote{LCS Tool is available at: \url{http://www.runmycode.org/companion/view/908}}.
+This engine is a library of MATLAB functions that extract LCSs from two-dimensional unsteady flows. The examples we present form demonstration scripts distributed with LCS Tool.
 \end{sloppypar}
 
-More recent work has focused on an exact mathematical formulation of the properties defining LCSs\parencite{haller11:_lagran_coher_struc,farazmand12:_comput_lagran,haller12:_geodes_theor_trans_barrier_two_dimen_flows,haller13:_coher_lagran,haller14:_adden_coher_lagran,farazmand13:_attrac_lagran,blazevski14:_hyper_ellip_trans_barrier_three}.
-In two-dimensional flows, hyperbolic and parabolic LCSs turn out to be stationary curves of the averaged material shear\parencite{farazmand14:_shearless}, whereas elliptic LCSs are stationary curves of the averaged strain\parencite{haller13:_coher_lagran,haller14:_adden_coher_lagran}.
-These variational formulations lead to explicit solutions for LCSs as null geodesics of appropriate Lorentzian metrics.
-
-We present an algorithmic introduction to geodesic LCS detection.
-We then review the implementation of this approach in a computational engine called LCS Tool\footnote{LCS Tool is available at: \url{http://github.com/jeixav/LCS-Tool}}.
-This engine is a library of MATLAB functions that extract LCSs from two-dimensional unsteady flows. The examples we present form demonstration scripts distributed with LCS Tool.
-
 Other software for LCS detection exists.
-ManGen\parencite{lekien03:_time} calculates the FTLE and advects material curves in two-dimensional velocity fields.
+ManGen\citep{lekien03:_time} calculates the FTLE and advects material curves in two-dimensional velocity fields.
 It includes a graphical user interface and uses the Message Passing Interface standard for parallel calculations.
-Newman\parencite{toit10:_trans} calculates the FTLE using dimension independent code.
+Newman\citep{toit10:_trans} calculates the FTLE using dimension independent code.
 It assists ridge extraction of FTLE fields and supports analytic and dataset velocity definitions.
-FlowTK\parencite{ameli14:_devel_effic_flexib_pipel_lagran} calculates the FTLE in two and three-dimensions on Cartesian and unstructured grids.
+FlowTK\citep{ameli14:_devel_effic_flexib_pipel_lagran} calculates the FTLE in two and three-dimensions on Cartesian and unstructured grids.
 The NVIDIA CUDA parallel computing platform is used for fast computations and Kitware's ParaView data analysis and visualization application is used for a user interface.
@@
 \begin{equation}
-C_{t_0}^t(x_0) \equiv \left[\nabla F_{t_0}^t(x_0)\right]^T \nabla F_{t_0}^t(x_0),
+C_{t_0}^t(x_0) = \left[\nabla F_{t_0}^t(x_0)\right]^T \nabla F_{t_0}^t(x_0),
 \label{eq:CG}
 \end{equation}
-measuring Lagrangian strain in the velocity field. This tensor is symmetric and positive definite\parencite{truesdell04}. We label the eigenvalues and eigenvectors of $C_{t_0}^t(x_0)$ as follows:
+measuring Lagrangian strain in the velocity field. This tensor is symmetric and positive definite\citep{truesdell04}. We label the eigenvalues and eigenvectors of $C_{t_0}^t(x_0)$ as follows:
 \[
@@
 We seek positions of closed material lines at time $t_0$ that prevail as coherent Lagrangian vortex boundaries (or \emph{elliptic LCSs}) over a time interval $[t_0,t]\subset[t_-,t_+]$.
-\textcite{haller13:_coher_lagran,haller14:_adden_coher_lagran} argue that such initial material line positions are closed stationary curves of the averaged strain functional
+\citet{haller13:_coher_lagran,haller14:_adden_coher_lagran} argue that such initial material line positions are closed stationary curves of the averaged strain functional
 \[
@@
 For this reason, we refer to trajectories of \cref{eq:etafields} as $\lambda$-lines.
-Following \textcite{haller13:_coher_lagran,haller14:_adden_coher_lagran}, we call the outermost member of a closed family of $\lambda$-lines a coherent Lagrangian vortex boundary.
+Following \citet{haller13:_coher_lagran,haller14:_adden_coher_lagran}, we call the outermost member of a closed family of $\lambda$-lines a coherent Lagrangian vortex boundary.
 
@@
 
+\begin{sloppypar}
 Next we consider positions of material lines at time $t_0$ that prevail as most repelling or attracting material lines (or \emph{hyperbolic LCSs}) over a time interval $[t_0,t] \subset [t_-,t_+]$.
-\textcite{farazmand14:_shearless} argue that hyperbolic LCSs are stationary curves of the averaged shear functional
+\citet{farazmand14:_shearless} argue that hyperbolic LCSs are stationary curves of the averaged shear functional
 \begin{gather*}
@@
 More precisely, hyperbolic LCSs are stationary curves of $Q(\gamma)$ with respect to fixed-endpoint perturbations.
-We note that parabolic LCSs (Lagrangian jet cores) are also stationary curves of $Q(\gamma)$, but under variable endpoint perturbations (cf. \textcite{farazmand14:_shearless}).
+We note that parabolic LCSs (Lagrangian jet cores) are also stationary curves of $Q(\gamma)$, but under variable endpoint perturbations (cf. \citet{farazmand14:_shearless}).
+\end{sloppypar}
 
@@
 
-\begin{sloppypar}
 Shrink lines and stretch lines coincide with the null geodesics of the Lorentzian metric $h(u,v) = \left\langle u,D_{t_0}^t(r) v \right\rangle$.
 For this reason, we refer to the computation of hyperbolic LCSs as strongest normally-repelling or normally-attracting curves of \cref{eq:etafields} as \emph{geodesic detection of hyperbolic LCSs}.
-\end{sloppypar}
 
@@
 \item Define a local maximization/minimization distance for hyperbolic LCSs, \lstinline!lmd!
-% FIXME seed_curves_from_lambda_max is a poor name for a function that detects hyperbolic LCSs
-\item Detect hyperbolic LCSs, \lstinline!hyperbolicLine = seed_curves_from_lambda_max(v,d,ellipticLcs,lmd)! and \lstinline!hyperbolicLcs = remove_strain_in_elliptic(hyperbolicLine,ellipticLcs)!
+\item Detect hyperbolic LCSs, \lstinline!hyperbolicLine = seed_curves_from_lambda_max(v,d,ellipticLcs,lmd)! and \lstinline!hyperbolicLcs = remove_strain_in_elliptic( hyperbolicLine,ellipticLcs)!
 \end{enumerate}
@@
 \begin{center}
-\begin{tabular}{|c|p{.6\textwidth}|}
+\begin{tabular}{|c|p{.7\textwidth}|}
 \hline
-\multicolumn{2}{|p{.8\textwidth}|}{\lstinline![cgEigenvector,cgEigenvalue] = eig_cgStrain(derivative,domain,timespan,resolution)!}\tabularnewline
+\multicolumn{2}{|p{.95\textwidth}|}{\lstinline![cgEigenvector,cgEigenvalue] = eig_cgStrain(derivative,domain,timespan,resolution)!}\tabularnewline
 \hline \hline
@@
 Here $\delta x_j$ is a vector of length $\vert\delta x_j\vert > 0$ that points from the Cartesian grid-point $x_0$ in the $j^\text{th}$ coordinate direction (\cref{f:main and auxiliary grids}).
-Computational improvements arising from the use of the auxiliary grid over simply using the nearest points of the main grid were reported in \textcite{farazmand12:_comput_lagran}.
+Computational improvements arising from the use of the auxiliary grid over simply using the nearest points of the main grid were reported in \citet{farazmand12:_comput_lagran}.
 Experience suggests setting the auxiliary grid spacing to 1-10\% of the main grid spacing.
@@
 \begin{center}
-\def\svgwidth{.75\textwidth}
+\def\svgwidth{.325\textwidth}
 \import{graphics/}{main_aux_grids.pdf_tex}
@@
 Results are from the function \lstinline!eig_cgStrain! applied to the double gyre presented in \cref{sec:double gyre}.
-Results obtained with an Intel Core i3-4130 processor.
+Results obtained with an Intel Core i3-4130 3.4 GHz processor.
 }
@@
 
-\begin{sloppypar}
-Incompressible flows (i.e. those satisfying $\nabla \cdot v=0$) satisfy the relation $\lambda_1(x_0) \lambda_2(x_0) = 1$ at all points of the computational domain \parencite{arnold89:_mathem}.
+Incompressible flows (i.e.
+those satisfying $\nabla \cdot v=0$) satisfy the relation $\lambda_1(x_0) \lambda_2(x_0) = 1$ at all points of the computational domain \citep{arnold89:_mathem}.
 Incompressibility can be computationally imposed by first calculating $\lambda_2(x_0)$, then setting $\lambda_1(x_0) = 1/\lambda_2(x_0)$ and calculating the strain eigenvectors $\xi_2$ from $\lambda_2$, then $\xi_1$ from the relationship in \cref{eq:CG_invariants}.
-Experience shows that computing $\lambda_i$ in this order gives higher accuracy than in the reverse order \parencite{farazmand12:_comput_lagran}.
-% FIXME Add table or graph with calculations demonstrating statement below
-This procedure should be applied carefully since by changing $\lambda_1$ but not $\xi_1$, the error in satisfying the eigenvalue equation for the Cauchy-Green strain tensor may increase.
-\end{sloppypar}
+Experience shows that computing $\lambda_i$ in this order gives higher accuracy than in the reverse order \citep{farazmand12:_comput_lagran}.
 
@@
 A $\lambda$-line returning to its starting point is then an elliptic LCS.
-The outermost member of a family of closed $\lambda$-lines (obtained by varying $\lambda$) is a coherent Lagrangian vortex boundary\parencite{haller13:_coher_lagran,haller14:_adden_coher_lagran}.
+The outermost member of a family of closed $\lambda$-lines (obtained by varying $\lambda$) is a coherent Lagrangian vortex boundary\citep{haller13:_coher_lagran,haller14:_adden_coher_lagran}.
 
@@
 The Poincare section is to be oriented such that the first endpoint is close to the centre of the expected Lagrangian vortex, and the second endpoint is outside this vortex.
-% FIXME Move citation to conclusions, as a candidate for future work.
-There is no automated procedure implemented in LCS Tool for the positioning of Poincare sections (see however \textcite{karrasch14:_autom_lagran} for a recently developed algorithm).
 Additionally, the number of lambda-lines launched from the Poincare section, \lstinline!poincareSection.numPoints!, must be defined.
@@
 \begin{center}
-\def\svgwidth{.75\textwidth}
+\def\svgwidth{.4\textwidth}
 \import{graphics/}{variable_step_integration.pdf_tex}
@@
 
-\begin{sloppypar}
 An example of a Poincare map produced from integration of the $\eta_\pm^\lambda$ field is shown in \cref{f:Poincare return map}.
@@
 In practice, viewing the Poincare section as the radius of a circle and setting the maximum $\lambda$-line integration length to twice the circumference gives good results.
-\end{sloppypar}
 
@@
 \begin{center}
-\setlength\figurewidth{.75\textwidth}
+\setlength\figurewidth{.35\textwidth}
 \input{graphics/double_gyre/poincare_return_map.tikz}
@@
 
-The local maximization distance is the number of grid points over which the Cauchy-Green strain tensor is susceptible to numerical noise from integration.
-This distance is thus expected to be a function of variable time step integration tolerances.
-
-Prior to initiating shrinkline and stretchline integration, a maximum length is defined.
-Shrinklines and stretchlines are then integrated until reaching the domain boundary or the maximum length.
-The maximum length may be seen as safety factor to prevent highly sinuous shrinklines and stretchlines leading to inordinately long computations; the maximum length should be set such that most shrinklines and stretchlines stop at the domain boundary.
-
-Once shrinkline and stretchline integration has begun, it may evolve so as to be within the local minimization distance of an existing shrinkline or stretchline.
-In other words, the local minimization distance affects where shrinklines and stretchlines begin, but not where they finish.
-
 \section{Examples}
@@
 \begin{center}
-\begin{tabular}{|c|c|}
+\begin{tabular}{|p{.3\textwidth}|c|}
 \hline
@@
 \hline \hline
-double\_gyre/elliptic\_hyperbolic\_lcs & 50\tabularnewline
+double gyre & 50\tabularnewline
 \hline
-bickley\_jet/elliptic\_hyperbolic\_lcs\_details & 975\tabularnewline
+Bickley jet & 25\tabularnewline
 \hline
-ocean\_dataset/elliptic\_hyperbolic\_lcs\_details & 127\tabularnewline
+ocean dataset & 127\tabularnewline
 \hline
@@
 Runtime of some demo scripts.
-Results obtained with an Intel Core i3-4130 processor with eight gigabytes of memory.
+Results obtained with an Intel Core i3-4130 3.4 GHz processor with 8 GB of memory.
 }
@@
 
-The double gyre is a model for a time-dependent two gyre system observed in geophysical flows \parencite{shadden05:_defin_lagran_lyapun}.
+The double gyre is a model for a time-dependent two gyre system observed in geophysical flows \citep{shadden05:_defin_lagran_lyapun}.
 The model consists of two counter rotating sinusoidal vortices with a harmonically oscillating line in-between.
@@
 The free stretching parameter $\lambda$ (\cref{eq:etafields}) is varied over the range $[0.93,1.07]$ with increments of $0.01$ (\cref{ll:double gyre lambdaRange}).
-Closed orbits for all $\lambda$ values are computed and the outermost closed orbit is kept as the Lagrangian vortex boundary  (\crefrange{ll:double gyre lambda range}{ll:double gyre elliptic_lcs}).
+Closed orbits for all predefined $\lambda$ values are computed and the outermost closed orbit is kept as the Lagrangian vortex boundary  (\crefrange{ll:double gyre lambda range}{ll:double gyre elliptic_lcs}).
 We have chosen small values for the error tolerances of the integration of the Cauchy-Green strain tensor (\cref{ll:double gyre cgStrainOdeSolverOptions}) and $\lambda$-lines (\cref{ll:double gyre lambdaLineOdeSolverOptions}) to ensure convergence.
@@
 \centering
-\setlength\figurewidth{.875\textwidth}
+\setlength\figurewidth{.4\textwidth}
 \import{graphics/double_gyre/}{elliptic_lcs_convergence_500.tikz}
@@
 \centering
-\setlength\figurewidth{.8\textwidth}
+\setlength\figurewidth{.35\textwidth}
 \import{graphics/double_gyre/}{elliptic_repelling_lcs_details.tikz}
@@
 \centering
-\setlength\figurewidth{.775\textwidth}
+\setlength\figurewidth{.325\textwidth}
 \import{graphics/bickley_jet/}{elliptic_repelling_lcs_details.tikz}
@@
 In contrast to the previous two analytic examples, the velocity field is available only with discrete temporal and spatial resolution.
-Our region of interest is a small domain in the South Atlantic Ocean, where exceptionally coherent eddies, Agulhas rings, were recently found by \textcite{haller13:_coher_lagran,haller14:_adden_coher_lagran} using the theory we surveyed in \cref{sec:Elliptic LCSs}.
+Our region of interest is a small domain in the South Atlantic Ocean, where exceptionally coherent eddies, Agulhas rings, were recently found by \citet{haller13:_coher_lagran,haller14:_adden_coher_lagran} using the theory we surveyed in \cref{sec:Elliptic LCSs}.
 
@@
 
-We choose the integration time as $T=30$ days (\cref{l:ocean data elliptic hyperbolic lcs}, \cref{ll:ocean timespan}), which is larger than the eddy turnover time in this region.
-The resolution of the main computational grid for initial conditions is set to $400 \times 400$ (\cref{ll:ocean resolution}).
+We choose the integration time as $T=30$ days (\cref{l:ocean data elliptic hyperbolic lcs}, \cref{ll:ocean data timespan}), which is larger than the eddy turnover time in this region.
+The resolution of the main computational grid for initial conditions is set to $400 \times 400$ (\cref{ll:ocean data resolution}).
 This corresponds to a resolution of roughly 0.015°.
@@
 The flow is integrated and the Cauchy-Green strain tensor is computed by the function \lstinline!eig_cgStrain! (\cref{ll:eig_cgStrain}).
-Incompressibility of the flow is enforced (\cref{ll:incompressible}), the auxiliary grid distance is set to 1\% of the main grid distance (\cref{ll:cgAuxGridRelDelta}), and eigenvalues are computed from the auxiliary grid (\cref{ll:cgEigenvalueFromMainGrid}).
-Elliptic LCSs are computed in \crefrange{ll:ocean elliptic LCS start}{ll:ocean elliptic LCS end}, after the Poincare sections have been set (\crefrange{ll:ocean data poincareSection(1)}{ll:ocean data poincareSection(5)}).
-$\lambda$-values are varied over a range of $[0.90,1.10]$ with a step of $0.02$ (\cref{ll:ocean lambda}).
-Hyperbolic LCSs are computed in \crefrange{ll:ocean hyperbolic start}{ll:ocean hyperbolic end}.
+The auxiliary grid distance is set to 1\% of the main grid distance (\cref{ll:cgAuxGridRelDelta}), and eigenvalues are computed from the auxiliary grid (\cref{ll:cgEigenvalueFromMainGrid}).
+Elliptic LCSs are computed in \cref{ll:ocean data poincare_closed_orbit_multi}, after the Poincare sections have been set (\cref{ll:ocean data poincareSection(1)}).
+$\lambda$-values are varied over a range of $[0.90,1.10]$ with a step of $0.02$ (\cref{ll:ocean data lambdaRange}).
+Hyperbolic LCSs are computed in \cref{ll:ocean data shrinkLineLcs,ll:ocean data stretchLineLcs}.
 
-\Cref{f:ocean dataset LCS} shows elliptic and hyperbolic LCSs on 22 November 2006, the same time as analysed in \textcite{haller13:_coher_lagran,haller14:_adden_coher_lagran,beron-vera13:_objec_agulh}.
+\begin{sloppypar}
+\Cref{f:ocean dataset LCS} shows elliptic and hyperbolic LCSs on 22 November 2006, the same time as analysed in \citet{haller13:_coher_lagran,haller14:_adden_coher_lagran,beron-vera13:_objec_agulh}.
 In those references, the integration time is 90 days.
@@
 The largest, at $(3,-32)$, has a non-stretching boundary, i.e., $\lambda = 1$.
-It corresponds to eddy \#2 in figure~3 of \textcite{beron-vera13:_objec_agulh}.
+It corresponds to eddy \#2 in figure~3 of \citet{beron-vera13:_objec_agulh}.
 Four additional smaller coherent eddies are found.
@@
 \begin{center}
-\setlength\figurewidth{.85\textwidth}
+\setlength\figurewidth{.375\textwidth}
 \import{graphics/ocean_dataset/}{elliptic_repelling_lcs.tikz}
@@
 LCS Tool leverages the capabilities of MATLAB.
-For FTLE based extraction of LCSs, computational performance has received considerable attention \parencite{conti12:_gpu_apu_finit_time_lyapun_expon,miron12:_anisot_lagran_coher_struc}.
+For FTLE based extraction of LCSs, computational performance has received considerable attention \citep{conti12:_gpu_apu_finit_time_lyapun_expon,miron12:_anisot_lagran_coher_struc}.
 We think LCS Tool can facilitate similar computational advances for variational LCS methods.
-Optimizing computational performance will aid applications to large-scale forecasting applications, such as the tracking of environmental contaminants\parencite{olascoaga12:_forec}.
+Optimizing computational performance will aid applications to large-scale forecasting applications, such as the tracking of environmental contaminants\citep{olascoaga12:_forec}.
 
 We think LCS Tool can serve as a foundation for the numerical implementation of recent theoretical advances.
-These include the geodesic theory of parabolic LCSs (jet cores)\parencite{farazmand14:_shearless} and the variational theory of hyperbolic and elliptic LCSs for three-dimensional flows\parencite{blazevski14:_hyper_ellip_trans_barrier_three}.
+These include the geodesic theory of parabolic LCSs (jet cores)\citep{farazmand14:_shearless} and the variational theory of hyperbolic and elliptic LCSs for three-dimensional flows\citep{blazevski14:_hyper_ellip_trans_barrier_three}.
 
@@
 
+\begin{sloppypar}
 The altimeter products used in this work are produced by SSALTO/DUACS and distributed by AVISO, with support from CNES, \url{http://www.aviso.oceanobs.com/duacs}.
+\end{sloppypar}
 
-% FIXME Citations without pages (karrasch14:_autom_lagran) do not list publication year
-\printbibliography
+\section*{References}
+
+\bibliographystyle{elsarticle-num-names} 
+\bibliography{main}
 
@@
 % Local Variables:
-% ispell-dictionary: "en_CA"
+% ispell-dictionary: "en_GB"
 % End:

### 2014-12-16 graphics/bickley_jet/elliptic_hyperbolic_lcs_details.m

@@ -1,173 +1,243 @@
-function elliptic_hyperbolic_lcs_details
+commit d1700275fbb35a7c16481b37cb6b61e53810e421
+Author: Kristjan Onu <jeixav@gmail.com>
+Date:   Tue Dec 16 17:00:00 2014 -0500
 
-%% Input parameters
-u = 62.66;
-lengthX = pi*earthRadius;
-lengthY = 1.77e6;
-epsilon = [.075,.4,.3];
-domain = [0,lengthX;[-1,1]*2.25*lengthY];
-timespan = [0,4*lengthX/u];
-resolutionX = 500;
-[resolutionY,deltaX] = equal_resolution(domain,resolutionX);
-resolution = [resolutionX,resolutionY];
+    Update Bickley jet demo parameters
 
-%% Velocity definition
-perturbationCase = 3;
-phiTimespan = [0,25];
-phiInitial = [0,0];
-phiSol = ode45(@d_phi,phiTimespan,phiInitial);
-timeResolution = 1e5;
-phi1 = deval(phiSol,linspace(phiTimespan(1),phiTimespan(2),timeResolution),1);
-phi1Max = max(phi1);
-lDerivative = @(t,x,~)derivative(t,x,false,u,lengthX,lengthY,epsilon,perturbationCase,phiSol,phi1Max);
-incompressible = true;
-periodicBc = [true,false];
-
-%% LCS parameters
-
-% Lambda-lines
-poincareSection = struct('endPosition',{},'numPoints',{},'orbitMaxLength',{});
-poincareSection(1).endPosition = [3.25e6,1.5e6;1.4e6,2.6e6];
-poincareSection(2).endPosition = [6.5e6,-1.4e6;5e6,-3e6];
-poincareSection(3).endPosition = [1e7,1.5e6;8e6,2.6e6];
-poincareSection(4).endPosition = [1.35e7,-1.4e6;1.5e7,-.5e6];
-poincareSection(5).endPosition = [1.65e7, 1.5e6;1.5e7, 2.6e6];
-[poincareSection.numPoints] = deal(80);
-nPoincareSection = numel(poincareSection);
-for i = 1:nPoincareSection
-    rOrbit = hypot(diff(poincareSection(i).endPosition(:,1)),diff(poincareSection(i).endPosition(:,2)));
-    poincareSection(i).orbitMaxLength = 2*(2*pi*rOrbit);
-end
-lambda = .9:.01:1.1;
[…]

### 2014-12-16 graphics/bickley_jet/elliptic_hyperbolic_lcs.m
Update Bickley jet demo parameters

@@ -5,8 +5,8 @@ u = 62.66;
 lengthX = pi*earthRadius;
 lengthY = 1.77e6;
 epsilon = [.075,.4,.3];
-domain = [0,lengthX;[-1,1]*2.25*lengthY];
 timespan = [0,4*lengthX/u];
+domain = [0,lengthX;[-1,1]*2.25*lengthY];
 resolutionX = 500;
 [resolutionY,deltaX] = equal_resolution(domain,resolutionX);
 resolution = [resolutionX,resolutionY];
@@ -17,51 +17,56 @@ phiTimespan = [0,25];
 phiInitial = [0,0];
 phiSol = ode45(@d_phi,phiTimespan,phiInitial);
 timeResolution = 1e5;
-phi1 = deval(phiSol,linspace(phiTimespan(1),phiTimespan(2),timeResolution),1);
+phi1 = deval(phiSol,linspace(phiTimespan(1),phiTimespan(2),...
+    timeResolution),1);
 phi1Max = max(phi1);
-lDerivative = @(t,x,~)derivative(t,x,false,u,lengthX,lengthY,epsilon,perturbationCase,phiSol,phi1Max);
+lDerivative = @(t,x,~)derivative(t,x,false,u,lengthX,lengthY,epsilon,...
+    perturbationCase,phiSol,phi1Max);
 incompressible = true;
 periodicBc = [true,false];
 
 %% LCS parameters
-
-% Lambda-lines
-poincareSection = struct('endPosition',{},'numPoints',{},'orbitMaxLength',{});
+% Lambda lines
+poincareSection = struct('endPosition',{},'numPoints',{},...
+    'orbitMaxLength',{});
 poincareSection(1).endPosition = [3.25e6,1.5e6;1.4e6,2.6e6];
 poincareSection(2).endPosition = [6.5e6,-1.4e6;5e6,-3e6];
 poincareSection(3).endPosition = [1e7,1.5e6;8e6,2.6e6];
 poincareSection(4).endPosition = [1.35e7,-1.4e6;1.5e7,-.5e6];
 poincareSection(5).endPosition = [1.65e7, 1.5e6;1.5e7, 2.6e6];
-[poincareSection.numPoints] = deal(80);
+[poincareSection.numPoints] = deal(20);
 nPoincareSection = numel(poincareSection);
 for i = 1:nPoincareSection
-    rOrbit = hypot(diff(poincareSection(i).endPosition(:,1)),diff(poincareSection(i).endPosition(:,2)));
[…]
